### 2019-08-19 maciek.tex
init

\documentclass[conference]{IEEEtran}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts,mathtools}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{url}
\usepackage{tikz}
\usepackage{comment}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{claim}{Claim}
\newtheorem{lemma}{Lemma}
\newtheorem{fact}{Fact}
\newtheorem{corollary}{Corollary}
\newtheorem{rem}{Remark}
\newtheorem{observation}{Observation}
\newtheorem{definition}{Definition}
\newtheorem{property}{Property}

\DeclarePairedDelimiter\pair{(}{)}
\DeclarePairedDelimiter\set{\{}{\}}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

\newcommand\mahmoud[1]{\color{green}\textbf{\\ Mahmoud: #1}\\\color{black}}
\newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}

\begin{document}

Maciek

\end{document}

### 2019-08-19 maciek.tex
A simple algorithm for k=3

@@
-\documentclass[conference]{IEEEtran}
+\section{Lower bound $\Omega(\ell)$}
 
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts,mathtools}
-\usepackage[noend]{algpseudocode}
-\usepackage{algorithm}
-\usepackage{url}
-\usepackage{tikz}
-\usepackage{comment}
+Model assumes the existence of a perfect partition of the input graph.
+We present a lower-bound for the competitive ratio of $\Omega(\ell)$ for any deterministic algorithm.
+To this end, we use $k = 3$, i.e., the capacity of each server is $3$.
 
-\newtheorem{theorem}{Theorem}
-\newtheorem{proposition}{Proposition}
-\newtheorem{claim}{Claim}
-\newtheorem{lemma}{Lemma}
-\newtheorem{fact}{Fact}
-\newtheorem{corollary}{Corollary}
-\newtheorem{rem}{Remark}
-\newtheorem{observation}{Observation}
-\newtheorem{definition}{Definition}
-\newtheorem{property}{Property}
+\section{An $O(\ell)$-competitive algorithm for $k=3$}
 
-\DeclarePairedDelimiter\pair{(}{)}
-\DeclarePairedDelimiter\set{\{}{\}}
+\begin{theorem}
+  The cost of rebalancing cost for algorithm is $O(k)$.
+  \label{rebalancing-cost}
+\end{theorem}
 
-\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
-\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
+\begin{proof}
+  \maciek{Note: we consider a cost of cheapest feasible rebalancing after insertion of one edge. This in contrast to bounding the distance between any two reconfigurations (we do not do that).}
 
-\newcommand\mahmoud[1]{\color{green}\textbf{\\ Mahmoud: #1}\\\color{black}}
-\newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}
+  We distinguish among three configuration types of algorithm's clusters: $C_1, C_2, C_3$.
+In $C_1$ we have $3$ singleton components (this is also the initial configuration of any cluster).
+In $C_2$ we have one component of size $2$ and one component of size $1$.
+Finally, in $C_3$ we have one component of size $3$.
 
-\begin{document}
+  Consider a request $(u, v)$ and let $U, V$ be their clusters.
+  If $U=V$, then the request does not require a reconfiguration.
+  Now, we consider cases upon the type of clusters $U$ and $V$.
+  Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise the resulting component would have the size $4$, and this contradicts the existence of the perfect partition.
+  If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
 
-Maciek
+  Now, we focus on the case where both $U$ and $V$ are $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as the merged component would have size $4$.
+  If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
+  Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+  Note that in such case, a $C_1$-type cluster exists, as otherwise the input graph would not have a perfect partition.
+The cost of rebalancing is then at most $12$ \maciek{can show smaller constant than 12, but it does not matter that much}.
+\end{proof}
 
-\end{document}
+\begin{theorem}
+  There exists a $O(\ell)$-competitive algorithm for $k=3$.
+\end{theorem}
+
+\begin{proof}
+  The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
+Each request issues at most one rebalancing, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
+\end{proof}

### 2019-08-19 maciek.tex
A sketch of the lower-bound

@@
 To this end, we use $k = 3$, i.e., the capacity of each server is $3$.
+
+\begin{theorem}
+  The competitive ratio of any deterministic algorithm for Online Balanced Partition (with existance of a perfect partition of the input graph) is $\Omega(\ell)$.
+\end{theorem}
+%Note that in the scenario with a perfect partition, both OPT and ALG always pays at most $O(\ell)$.
+
+\begin{proof}
+We say that nodes that belong to the same cluster in the initial configuration \emph{have the same color}.
+
+Consider clusters $A$ and $B$ that contain nodes $a_1, a_2, a_3$ and $b_1, b_2, b_3$, respectively.
+The adversary begins with issuing three requests: $(a_1, a_2)$, $(b_1, b_2)$, and $(a_3, b_3)$.
+Note that these require moving one of these pairs to a third cluster.
+Our idea is that the OPT pays for the request $(a_3, b_3)$ only (i.e., it moves this pair to some cluster $C$), and all other requests are free for the OPT.
+
+The algorithm must place these three components in different clusters.
+Let's call the cluster of $(a_1, a_2)$ the cluster $A_{ALG}$, the cluster of $(b_1, b_2)$ the cluster $B_{ALG}$ and the cluster of $(a_3, b_3)$ the cluster $C$.
+The configuration of ALG's clusters is then: $A$ contains $a_1, a_2,x$, $B$ contains $b_1, b_2,y$ and $C$ contains $a_3, b_3, z$, for some $x, y, z$ (of choice of the algorithm).
+For this, ALG pays at least $2$.
+
+The OPT places these three components in clusters $A, B and D$, paying $2$ for migrating $a_3$ and $b_3$.
+The OPT chooses the cluster $D$ so that it does not contain neither $x$, $y$ nor $z$.
+The configuration of OPT's clusters is then: $A$ contains $a_1, a_2,d_1$, $B$ contains $b_1, b_2,d_2$ and $D$ contains $a_3, b_3, d_3$.
+
+The next action of the adversary depends on the choice of $x, y, z$.
+If $x$ and $y$ have different color (i.e., initially they belong to different clusters), then we issue a request between $x$ and any other node with the same color as $x$.
+As $x$ do not intersect with $D$, the OPT has them co-located in one server, and it pays no cost for such requests.
+We issue such requests until $x$ and $y$ have the same color.
+
+Then, we issue a request $(x, y)$.
+
+\end{proof}
 

### 2019-08-19 maciek.tex
A sketch of the lower-bound

@@
 Model assumes the existence of a perfect partition of the input graph.
-We present a lower-bound for the competitive ratio of $\Omega(\ell)$ for any deterministic algorithm.
+We present a lower bound for the competitive ratio of $\Omega(\ell)$ for any deterministic algorithm.
 To this end, we use $k = 3$, i.e., the capacity of each server is $3$.
@@
 
-Then, we issue a request $(x, y)$.
+Note that the adversary can foresee the actions of the algorithm, and may choose $D$ in such way, that $x$ cycles through all other clusters before reaching $D$, paying the cost of $\Omega(k)$, which would already suffice to show the lower bound.
+
+Then, $x$ and $y$ have the same color, and we issue a request $(x, y)$.
+As these have the same color (and there are different than $d_1, d_2$ and $d_3$), OPT pays no cost for this request, and ALG pays $2$ for migrating $x$ and $y$ to other cluster.
+We repeat the entire process, i.e., we force ALG to choose $x$ and $y$ of the same color, and issue such request that is free for OPT.
+This process can continue until ALG has no more clusters with singleton components, or it pays the cost $\Omega(k)$ already.
 

### 2019-08-19 maciek.tex
Lower bound, simplifications

--- 2019-08-19 maciek.tex
@@
 The OPT places these three components in clusters $A, B and D$, paying $2$ for migrating $a_3$ and $b_3$.
-The OPT chooses the cluster $D$ so that it does not contain neither $x$, $y$ nor $z$.
+The OPT has the freedom of choice of the cluster $D$ depending on the choices of the algorithm.
+We choose $D$ in such way that it does not contain neither $x$ nor $y$, and furthermore that the algorithm places any of nodes from $D$ at $x$ or $y$'s place as the last possible move.
 The configuration of OPT's clusters is then: $A$ contains $a_1, a_2,d_1$, $B$ contains $b_1, b_2,d_2$ and $D$ contains $a_3, b_3, d_3$.
 
-The next action of the adversary depends on the choice of $x, y, z$.
-If $x$ and $y$ have different color (i.e., initially they belong to different clusters), then we issue a request between $x$ and any other node with the same color as $x$.
+From this point, OPT does not change its configuration, and the adversary continues to issue a request between $x$ and any other node with the same color as $x$.
+Upon receiving this request, ALG pays for it the cost $1$ and it increases its number of clusters that contain non-singleton components.
+This procedure ends when ALG has no more clusters with singleton-only components left or ALG manages to put $d_1, d_2$ or $d_3$ in place of $x$ or $y$.
+
+At the begining, ALG has $\ell$ singleton-only components and each request increases their number by $1$.
+Furthermore, before reaching either $d_1, d_2$ or $d_3$ in place of $x$ or $y$, the algorithm must cycle through $\Omega(\ell)$ nodes from other clusters.
+This is guaranteed as the choice of $D$ can depend on the actions of the deterministic algorithm.
+In total, OPT pays $2$, and ALG pays $\Omega(\ell)$.
+
+%The next action of the adversary depends on the choice of $x, y, z$.
+%If $x$ and $y$ have different color (i.e., initially they belong to different clusters), then we issue a request between $x$ and any other node with the same color as $x$.
 As $x$ do not intersect with $D$, the OPT has them co-located in one server, and it pays no cost for such requests.
@@
 
-Then, we issue a request $(x, y)$.
+%Note that the adversary can foresee the actions of the algorithm, and may choose $D$ in such way, that $x$ cycles through all other clusters before reaching $D$, paying the cost of $\Omega(k)$, which would already suffice to show the lower bound.
+
+%Then, $x$ and $y$ have the same color, and we issue a request $(x, y)$.
+As these have the same color (and there are different than $d_1, d_2$ and $d_3$), OPT pays no cost for this request, and ALG pays $2$ for migrating $x$ and $y$ to other cluster.
+%We repeat the entire process, i.e., we force ALG to choose $x$ and $y$ of the same color, and issue such request that is free for OPT.
+This process can continue until ALG has no more clusters with singleton components, or it pays the cost $\Omega(k)$ already.
 

### 2019-08-19 maciek.tex
Lower bound, simplifications

@@
 
-The OPT places these three components in clusters $A, B and D$, paying $2$ for migrating $a_3$ and $b_3$.
+The OPT places these three components in clusters $A, B$ and $D$, paying $2$ for migrating $a_3$ and $b_3$.
 The OPT has the freedom of choice of the cluster $D$ depending on the choices of the algorithm.
@@
 The configuration of OPT's clusters is then: $A$ contains $a_1, a_2,d_1$, $B$ contains $b_1, b_2,d_2$ and $D$ contains $a_3, b_3, d_3$.
+Note that the requests between nodes of the same color are free for OPT, with the exception of nodes from $A$, $B$, and most notably $D$.
 

### 2019-08-19 maciek.tex
Lower bound, simplifications
Whiteboard pics

@@
 Note that these require moving one of these pairs to a third cluster.
-Our idea is that the OPT pays for the request $(a_3, b_3)$ only (i.e., it moves this pair to some cluster $C$), and all other requests are free for the OPT.
+Our idea is that the OPT pays for the request $(a_3, b_3)$ only (i.e., it moves this pair to a cluster $D$), and for OPT all other requests are free.
 
@@
 This is guaranteed as the choice of $D$ can depend on the actions of the deterministic algorithm.
-In total, OPT pays $2$, and ALG pays $\Omega(\ell)$.
+In total, OPT pays $2$, and ALG pays $\Omega(\ell)$ (for each request besides the first two).
 

### 2019-08-19 maciek.tex
Lower bound, simplifications

@@
 Note that these require moving one of these pairs to a third cluster.
-Our idea is that the OPT pays for the request $(a_3, b_3)$ only (i.e., it moves this pair to a cluster $D$), and for OPT all other requests are free.
+Our idea is that the OPT pays for the request $(a_3, b_3)$ only (i.e., it moves this pair to a carefuly chosen cluster, that we denote by $D$), and for OPT all other requests are free.
 
@@
 Let's call the cluster of $(a_1, a_2)$ the cluster $A_{ALG}$, the cluster of $(b_1, b_2)$ the cluster $B_{ALG}$ and the cluster of $(a_3, b_3)$ the cluster $C$.
-The configuration of ALG's clusters is then: $A$ contains $a_1, a_2,x$, $B$ contains $b_1, b_2,y$ and $C$ contains $a_3, b_3, z$, for some $x, y, z$ (of choice of the algorithm).
+The configuration of ALG's clusters is then: $A_{ALG}$ contains $a_1, a_2,x$, the cluster $B_{ALG}$ contains $b_1, b_2,y$ and the cluster $C$ contains $a_3, b_3, z$, for some $x, y, z$ (of the algorithm's choice).
 For this, ALG pays at least $2$.
@@
 The OPT has the freedom of choice of the cluster $D$ depending on the choices of the algorithm.
-We choose $D$ in such way that it does not contain neither $x$ nor $y$, and furthermore that the algorithm places any of nodes from $D$ at $x$ or $y$'s place as the last possible move.
+We choose $D$ in such way that it does not contain neither $x$ nor $y$, and furthermore that the algorithm places any of nodes from $D$ at $x$'s or $y$'s place as the last possible move.
 The configuration of OPT's clusters is then: $A$ contains $a_1, a_2,d_1$, $B$ contains $b_1, b_2,d_2$ and $D$ contains $a_3, b_3, d_3$.
@@
 
-From this point, OPT does not change its configuration, and the adversary continues to issue a request between $x$ and any other node with the same color as $x$.
-Upon receiving this request, ALG pays for it the cost $1$ and it increases its number of clusters that contain non-singleton components.
-This procedure ends when ALG has no more clusters with singleton-only components left or ALG manages to put $d_1, d_2$ or $d_3$ in place of $x$ or $y$.
+From this point, OPT does not change its configuration, and the adversary continues to issue a request between a node colocated with the pair $(a_1, a_2)$, which we call $x$ (we update $x$ in every iteration, as it gets moved from this place) and any other node with the same color as $x$.
+Upon receiving this request, ALG pays for it the cost at least $1$ and it increases its number of clusters that contain non-singleton components by $1$.
+This procedure ends when ALG has no more clusters with singleton-only components left or ALG manages to put $d_1, d_2$ or $d_3$ in place of either $x$ or $y$.
 
 At the begining, ALG has $\ell$ singleton-only components and each request increases their number by $1$.
-Furthermore, before reaching either $d_1, d_2$ or $d_3$ in place of $x$ or $y$, the algorithm must cycle through $\Omega(\ell)$ nodes from other clusters.
+Furthermore, before reaching either $d_1, d_2$ or $d_3$ in place of either $x$ or $y$, the algorithm must cycle through $\Omega(\ell)$ nodes from other clusters.
 This is guaranteed as the choice of $D$ can depend on the actions of the deterministic algorithm.

### 2019-08-19 maciek.tex
Some fixes

@@
 \begin{theorem}
-  The cost of rebalancing cost for algorithm is $O(k)$.
+  The rebalancing cost for the greedy algorithm is $O(k)$.
+  \maciek{Greedy = find the cheapest feasible partition}
   \label{rebalancing-cost}
@@
 \begin{proof} 
-  \maciek{Note: we consider a cost of cheapest feasible rebalancing after insertion of one edge. This in contrast to bounding the distance between any two reconfigurations (we do not do that).}
+ \maciek{Note: we consider a cost of cheapest feasible rebalancing after insertion of one edge.
+This in contrast to bounding the distance between any two reconfigurations (this theorem does not bound the latter).}
 

### 2019-08-19 maciek.tex
Some fixes

@@
 \begin{theorem}
-  The rebalancing cost for the greedy algorithm is $O(k)$.
+  The rebalancing cost for the greedy algorithm for $k=3$ is $O(1)$.
   \maciek{Greedy = find the cheapest feasible partition}
@@
   The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
-Each request issues at most one rebalancing, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
+Each request issues at most one rebalancingfor the greedy algorithm, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
 \end{proof}

### 2019-08-19 maciek.tex
Some fixes

@@
 The OPT has the freedom of choice of the cluster $D$ depending on the choices of the algorithm.
-We choose $D$ in such way that it does not contain neither $x$ nor $y$, and furthermore that the algorithm places any of nodes from $D$ at $x$'s or $y$'s place as the last possible move.
+We choose $D$ in such way that it does not contain neither $x$ nor $y$, and furthermore that the algorithm places any of nodes from $D$ at $x$'s or $y$'s place as the last possible move \maciek{elaborate}.
 The configuration of OPT's clusters is then: $A$ contains $a_1, a_2,d_1$, $B$ contains $b_1, b_2,d_2$ and $D$ contains $a_3, b_3, d_3$.
@@
 Furthermore, before reaching either $d_1, d_2$ or $d_3$ in place of either $x$ or $y$, the algorithm must cycle through $\Omega(\ell)$ nodes from other clusters.
-This is guaranteed as the choice of $D$ can depend on the actions of the deterministic algorithm.
+This can be guaranteed as the choice of $D$ can depend on the actions of the deterministic algorithm.
 In total, OPT pays $2$, and ALG pays $\Omega(\ell)$ (for each request besides the first two).

### 2019-08-20 maciek.tex
Lower bound for augmentation

@@
 \begin{theorem}
-  The competitive ratio of any deterministic algorithm for Online Balanced Partition (with existance of a perfect partition of the input graph) is $\Omega(\ell)$.
+  The competitive ratio of any deterministic algorithm for Online Balanced Partition (assuming the existence of a perfect partition of the input graph) is $\Omega(\ell)$.
+  \label{th:lb_omega_l}
 \end{theorem}
@@
 
-Consider clusters $A$ and $B$ that contain nodes $a_1, a_2, a_3$ and $b_1, b_2, b_3$, respectively.
+Consider clusters $A$ and $B$ containing nodes $a_1, a_2, a_3$ and $b_1, b_2, b_3$, respectively.
 The adversary begins with issuing three requests: $(a_1, a_2)$, $(b_1, b_2)$, and $(a_3, b_3)$.
-Note that these require moving one of these pairs to a third cluster.
-Our idea is that the OPT pays for the request $(a_3, b_3)$ only (i.e., it moves this pair to a carefuly chosen cluster, that we denote by $D$), and for OPT all other requests are free.
+Note that the algorithm is forced to move at least one of these components to a third cluster:
+in general, if the algorithm does not keep each component within one cluster, the adversary might continue issuing requests among nodes of this component.
 
-The algorithm must place these three components in different clusters.
+The algorithm places these three components in different clusters.
 Let's call the cluster of $(a_1, a_2)$ the cluster $A_{ALG}$, the cluster of $(b_1, b_2)$ the cluster $B_{ALG}$ and the cluster of $(a_3, b_3)$ the cluster $C$.
@@
 
+Our idea is that the OPT pays for the request $(a_3, b_3)$ only (i.e., it moves this pair to a carefuly chosen cluster, that we denote by $D$), and for OPT all other requests are free. 
 The OPT places these three components in clusters $A, B$ and $D$, paying $2$ for migrating $a_3$ and $b_3$.
@@
 \end{proof}
+
+\begin{observation}
+  The lower bound of $\Omega(\ell)$ can be generalized for arbitrary $k = c\cdot 3$.
+  The adversary starts with a serie of ``glueing requests'', so that each cluster contains 3 components of size $k/3$.
+  We repeat the argument of Theorem~\ref{th:lb_omega_l} using these glued components instead of singletons.
+  OPT pays $O(k)$ for migration of one such component, and ALG pays $\Omega(k \cdot \ell)$.
+  \label{obs:lb_general_k}
+\end{observation}
+
+\begin{observation}
+  The lower bound of $\Omega(\ell)$ holds even with $1/3-1/k$-augmentation, i.e., in scenarios where the capacity of each cluster of the algorithm is $(1+1/3)\cdot k - 1$, while the capacity of OPT's clusters is $k$.
+  To this end we use the construction from Observation~\ref{obs:lb_general_k}.
+  We observe that each cluster of the algorithm can fit exactly $3$ components of size $k/3$.
+  \maciek{Note that this does not contradict the existence of $O((1+1/\epsilon)\cdot k \cdot \log(k))$ algorithm of Avin, Bienkowski et al, as it requires $2+\epsilon$-augmentation.}
+\end{observation}
 

### 2019-08-20 maciek.tex
Refinement of lower bound writeup

@@
 \section{Lower bound $\Omega(\ell)$}
+
+\maciek{I assume $\alpha = 1$ to simplify calculations, but everything should hold for general $\alpha$}
 
@@
 
-The algorithm places these three components in different clusters.
-Let's call the cluster of $(a_1, a_2)$ the cluster $A_{ALG}$, the cluster of $(b_1, b_2)$ the cluster $B_{ALG}$ and the cluster of $(a_3, b_3)$ the cluster $C$.
+The algorithm places these three components in different clusters, and for this it pays at least $2$.
+By $x$ (resp. $y$) we denote the third node from the cluster containing the component $(a_1, a_2)$ (resp. $(b_1, b_2)$).
+%These nodes $x$ and $y$ are of the algorithm's choice, but.
+During the execution of the algorithm, these nodes change, and by $x$ and $y$ we always denote the current nodes of the algorithm colocated with components $(a_1, a_2)$ and $(b_1, b_2)$.
+
+%Let's call the cluster of $(a_1, a_2)$ the cluster $A_{ALG}$, the cluster of $(b_1, b_2)$ the cluster $B_{ALG}$ and the cluster of $(a_3, b_3)$ the cluster $C$.
 The configuration of ALG's clusters is then: $A_{ALG}$ contains $a_1, a_2,x$, the cluster $B_{ALG}$ contains $b_1, b_2,y$ and the cluster $C$ contains $a_3, b_3, z$, for some $x, y, z$ (of the algorithm's choice).
-For this, ALG pays at least $2$.
 
-Our idea is that the OPT pays for the request $(a_3, b_3)$ only (i.e., it moves this pair to a carefuly chosen cluster, that we denote by $D$), and for OPT all other requests are free. 
-The OPT places these three components in clusters $A, B$ and $D$, paying $2$ for migrating $a_3$ and $b_3$.
-The OPT has the freedom of choice of the cluster $D$ depending on the choices of the algorithm.
-We choose $D$ in such way that it does not contain neither $x$ nor $y$, and furthermore that the algorithm places any of nodes from $D$ at $x$'s or $y$'s place as the last possible move \maciek{elaborate}.
+Our idea is that OPT pays for the request $(a_3, b_3)$ only (i.e., it moves this pair to a carefuly chosen cluster that we denote by $D$), and for OPT all other requests are free. 
+OPT places these three components in clusters $A, B$ and $D$, paying $2$ for migrating $a_3$ and $b_3$.
 The configuration of OPT's clusters is then: $A$ contains $a_1, a_2,d_1$, $B$ contains $b_1, b_2,d_2$ and $D$ contains $a_3, b_3, d_3$.
 Note that the requests between nodes of the same color are free for OPT, with the exception of nodes from $A$, $B$, and most notably $D$.
+
+OPT has the freedom of choice of the cluster $D$ depending on the choices of the algorithm.
+We choose $D$ in such way that initially it does not contain neither $x$ nor $y$, and furthermore that the algorithm places any of nodes from $D$ at $x$'s or $y$'s place as the last possible move.
+In other words, we inspect the sequence of $x$ and $y$ of the algorithm, and we choose $D$ to be the last cluster whose initial contents appears as the $x$ and $y$.
+As there are $\ell$ clusters to choose from, ALG has to serve at least $\Omega(\ell)$ requests before any of $D$'s contents appear as $x$ or $y$.
 
@@
 This can be guaranteed as the choice of $D$ can depend on the actions of the deterministic algorithm.
-In total, OPT pays $2$, and ALG pays $\Omega(\ell)$ (for each request besides the first two).
+In total, OPT pays $2$, and moreover ALG pays $\Omega(\ell)$, as it peforms a migration of two nodes for each request besides the first two.
 

### 2019-08-20 maciek.tex
Refinement of lower bound writeup

@@
 
-From this point, OPT does not change its configuration, and the adversary continues to issue a request between a node colocated with the pair $(a_1, a_2)$, which we call $x$ (we update $x$ in every iteration, as it gets moved from this place) and any other node with the same color as $x$.
+From this point, OPT does not change its configuration, and the adversary continues to issue a request between $x$ and any other node with the same color as $x$
+(note that by $x$ we might denote different node in every iteration, as $x$ is defined as a node colocated with a component $(a_1, a_2)$).
 Upon receiving this request, ALG pays for it the cost at least $1$ and it increases its number of clusters that contain non-singleton components by $1$.

### 2019-08-21 maciek.tex
Costly rebalance
Omega(kl)

@@
 
+\section{Lower bound $\Omega(k\cdot \ell)$}
+
+\maciek{Warning: sketch only, I did not finish writing it yet!}
+
+In this section we assume the Online Clustering problem with the existence of a perfect partition of the input graph.
+
+\begin{theorem}
+  The competitive ratio of any deterministic algorithm for Online Clustering (with perfect partition) is $\Omega(k\cdot \ell)$.
+\end{theorem}
+
+\begin{proof}
+
+\end{proof}
+
 \section{An $O(\ell)$-competitive algorithm for $k=3$}
@@
 \end{proof}
+
+\section{Costly rebalance example $\Omega(k^2)$}

### 2019-08-21 maciek.tex
Omega(kl)

@@
 
+  \begin{enumerate}
+    \item Clusters A and B with components of size k-1. Define x and y.
+    \item The input sequence (after these glueing) is to keep asking for x and someone of x's color
+    \item Define a sequence of xs and ys for any deterministic algorithm.
+    \item Define what OPT should do: the cluster D and nodes d1 d2 that should be hidden from the sequence of xs (and ys?)
+    \item (somewhere) define c(x): it is easy when x is not from D: then any would do; if x is from D, then issue request between non-d1/d2
+    \item Argue that OPT does not pay for input sequence (after glueing)
+    \item bound the number of times we can issue (x, c(x)) request
+  \end{enumerate}
 \end{proof}

### 2019-08-21 maciek.tex

@@
 
-In this section we assume the Online Clustering problem with the existence of a perfect partition of the input graph.
+In this section we assume the Online Balanced Partition problem with the existence of a perfect partition of the input graph.
 
 \begin{theorem}
-  The competitive ratio of any deterministic algorithm for Online Clustering (with perfect partition) is $\Omega(k\cdot \ell)$.
+  The competitive ratio of any deterministic algorithm for Online Balanced Partition (with perfect partition) is $\Omega(k\cdot \ell)$.
 \end{theorem}
@@
     \item Argue that OPT does not pay for input sequence (after glueing)
-    \item bound the number of times we can issue (x, c(x)) request
+    \item lower-bound the number of times we can issue (x, c(x)) request
   \end{enumerate}
@@
   The rebalancing cost for the greedy algorithm for $k=3$ is $O(1)$.
-  \maciek{Greedy = find the cheapest feasible partition}
+  \maciek{Greedy = find the cheapest (closest to the current) feasible partition}
   \label{rebalancing-cost}

### 2019-08-21 maciek.tex

@@
   \begin{enumerate}
-    \item Clusters A and B with components of size k-1. Define x and y.
+    \item Cluster A component of size k-1. Define x.
     \item The input sequence (after these glueing) is to keep asking for x and someone of x's color
-    \item Define a sequence of xs and ys for any deterministic algorithm.
-    \item Define what OPT should do: the cluster D and nodes d1 d2 that should be hidden from the sequence of xs (and ys?)
-    \item (somewhere) define c(x): it is easy when x is not from D: then any would do; if x is from D, then issue request between non-d1/d2
+    \item Define a sequence of xs for any deterministic algorithm.
+    \item Define what OPT should do: the cluster D and nodes d1 d2 that should be hidden from the sequence of xs?)
+    \item (somewhere) define c(x): it is easy when x is not from D: then any would do; if x is from D, then issue request between non-d1/d2; also not defined for a1,ak-1, b1,bk-1 and ak, bk
     \item Argue that OPT does not pay for input sequence (after glueing)
-    \item lower-bound the number of times we can issue (x, c(x)) request
+    \item lower-bound the number of times we can issue (x, c(x)) request; until it encounters either d1 or d2 in place of x; this is lower-bounded by the total number of nodes-2 (d1 and d2 are last?
+they cannot be, though, but the furthest away should be at min distance k*l/2)
+    \item also, this k*l/2 sequence is feasible, as the perfect partition exists: we can always issue this request (unless d1d2)
+    \item the total cost of ALG, comp ratio
   \end{enumerate}
+
+  Fix a deterministic algorithm DET.
+  Fix a cluster $A$ containing nodes $a_1, \ldots, a_k$ and a cluster $B$ containing (among others) a node $b$.
+  The adversary begins with issuing $k-2$ internal requests $(a_i, a_{i+1})$, so that nodes $a_1, \ldots a_{k-1}$ form a single component.
+  The internal requests are free for OPT, and it doesn't perform additional actions at this point.
+
+  \maciek{Possibly we need a better name than $x$. Something that indicates that this changes over time, possibly tied to the name DET.}
+  We define $x$ to be the node in the configuration of DET that is co-located in one cluster with the component $a_1, \ldots, a_{k-1}$.
+  Initially, $x = a_k$.
+
+  Then, the adversary issues an external request $(x, b)$, which at this point is the request $(a_k, b)$.
+  Note that DET must relocate the component $\{a_k, b\}$ to some other cluster.
+  OPT also relocates the component $\{a_k, b\}$ to some carefuly chosen cluster $D$ (to be fixed later), and it migrates nodes $d_1$ and $d_2$ to the previous locations $a_k$ and $b$ (i.e., to the clusters $A$ and $B$).
+
+  Now, for a node $p$ we define $c(p)$ in the following way: for each node outside of $a_i$ and $b$, we define $c(p)$ to be any node that in the initial configuration was placed at the same cluster as $p$, and if $p$ belonged to $D$, then $c(p)$ is any such node that is different from $d_1$ and $d_2$.
+  \maciek{we have a cyclic reference at this point. D is not yet defined (depends on c), and c depends on d1, d2...}
+  Then, the adversary continues to issue requests $(x, c(x))$.
 \end{proof}

### 2019-08-21 mahmoud.tex
init
Alg 1 and the upperbound

--- 2019-08-19 maciek.tex
@@
-\documentclass[conference]{IEEEtran}
+\section{Mahmoud}
 
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts,mathtools}
-\usepackage[noend]{algpseudocode}
-\usepackage{algorithm}
-\usepackage{url}
-\usepackage{tikz}
-\usepackage{comment}
+Let $P_I = I_1, \dots, I_{\ell}$ denote the initial partition with which both ON and OPT begin and
+$P_F = F_1, \dots, F_{\ell}$ the final partition.
+\begin{definition}	\label{def:dist}
+	The \emph{distance} of a partition $P = C_1, \dots, C_{\ell}$ is the number of nodes in $P$ that do not reside in their initial cluster.
+	That is,
+	$dist(P, P_I) = \sum_{j=1}^{\ell} | C_j \setminus I_j |$. 
+\end{definition}
 
-\newtheorem{theorem}{Theorem}
-\newtheorem{proposition}{Proposition}
-\newtheorem{claim}{Claim}
-\newtheorem{lemma}{Lemma}
-\newtheorem{fact}{Fact}
-\newtheorem{corollary}{Corollary}
-\newtheorem{rem}{Remark}
-\newtheorem{observation}{Observation}
-\newtheorem{definition}{Definition}
-\newtheorem{property}{Property}
+In other words,
+at least $dist(P, P_I)/2$ node swaps are required in order to reach the a partition $P$ from $P_I$.
+Therefore,
+$OPT \geq dist(P_F, P_I) $.
+We devise an online algorithm that mimics OPT by minimizing the distance to the initial partition $P_I$ on every re-partitioning.
+As a result,
+ON never ends up a partition that is father away from $P_I$ than $OPT$.
+This invariant ensures that ON does not pay too much while reaching $P_F$.
 
-\DeclarePairedDelimiter\pair{(}{)}
-\DeclarePairedDelimiter\set{\{}{\}}
+\begin{algorithm}
+	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+	\renewcommand{\algorithmicensure}{\textbf{Output:}}
+	\begin{algorithmic}[1]
+		\Require 
+		$k, \ell$,
+		initial partition $P_I$,
+		sequence of communication edges $<\sigma_t>$ covering all ground truth components
+		\Ensure Final partition $P_F = F_1, \dots, F_{\ell}$ 
+		\State for each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$ \label{line:initcomponents}
+		\State on communication request $\sigma_t=\{u,v\}$:
+		\State Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
+		\If{$C_1 \neq C_2$}
+		\State unite the two components into a single component $C'$ and
+		$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$ \label{line:mergecomponents}
+		\If{cluster$(C_1) \neq$ cluster$(C_2)$}		
+		\State \textit{re-partition()} \label{line:rebalance} 
+		\EndIf
+		\EndIf
+	\end{algorithmic}
+	\caption{Perfect Partition Learner}
+	\label{alg:ppl}
+\end{algorithm}
 
-\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
-\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
+\begin{property} \label{prop:dist<OPT}
+	Let $P$ be any partition chosen Algorithm \ref{alg:ppl} at Line $\ref{line:rebalance}$.
+	Then, $dist(P) \leq OPT$.
+\end{property}
 
-\newcommand\mahmoud[1]{\color{green}\textbf{\\ Mahmoud: #1}\\\color{black}}
-\newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}
+\begin{lemma}	\label{lemma:rebalancecost}
+	The cost of re-partitioning at Line \ref{line:rebalance} is at most $2.OPT$.
+\end{lemma}
+\begin{IEEEproof}
+	Consider the re-partitioning that transforms $P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
+	Let $M \subset V$ denote the set of nodes that are migrated during this process.
+	We have $M = M^+ \cup M^- \cup M^\circ$,
+	where $M^-$ ($M^+$) denotes the subset of nodes that
+	enter (leave) their original cluster during the re-partitioning,
+	and $M^\circ$ denotes the set of remaining nodes in $M$.
+	Since $|M^- \cup M^\circ|$ nodes are not in their original cluster before the re-partitioning,
+	by Definition \ref{def:dist},
+	the distance before the re-partitioning is $dist(P_{t-1}) \geq | M^- \cup M^\circ |$.
+	Analogously,
+	 the distance afterwards is $dist(P_{t}) \geq | M^+ \cup M^\circ |$.
+	Thus,
+	$|M| \leq dist(P_{t-1}) + dist(P_{t})$.
+	By Property \ref{prop:dist<OPT},
+	$dist(P_{t-1}) , dist(P_{t}) \leq OPT$
+	and therefore we have	
+	$|M| \leq 2.OPT$.
+\end{IEEEproof}
 
-\begin{document}
-
-Maciek
-
-\end{document}
+\begin{theorem}	\label{thm:upperbound}
+	Algorithm \ref{alg:ppl} reaches the final partition while being $(2.k.\ell)$-competitive.
+\end{theorem}
+\begin{IEEEproof}
+	The algorithm eventually reaches the final partition since it
+	 evaluates all possible $\ell$-way partitioning of components on every external request,
+	including the request that completes revealing of all ground truth components.
+	There are at most $k.\ell$ calls to re-partition() and by Lemma \ref{lemma:rebalancecost} each costs at most $2.OPT$.
+	The total cost is therefore at most $2.OPT.k.\ell$ which implies the competitive ratio.
+\end{IEEEproof}

### 2019-08-22 maciek.tex
Omega(kl)

@@
-\section{Lower bound $\Omega(\ell)$}
+\section{Lower bound $\Omega(k\cdot \ell)$}
+
+\maciek{Warning: sketch only, I did not finish writing it yet!}
+
+In this section we consider the Online Balanced Partition problem.
+We show a lower bound of $\Omega(k \cdot \ell)$ that uses an input that has a perfect partition.
+Therefore, this result is a lower bound for the general Online Balanced Partition problem, as well as a lower bound for Online Balanced Partition with perfect partition.
+The latter complements an upper bound $O(k \cdot \ell)$ for Online Balanced Partition with perfect partition that we present in Section TODO.
+
+\begin{theorem}
+  The competitive ratio of any deterministic algorithm for Online Balanced Partition is $\Omega(k\cdot \ell)$.
+\end{theorem}
+
+\begin{proof}
+
+  \begin{enumerate}
+    \item Cluster A component of size k-1. Define x.
+    \item The input sequence (after these glueing) is to keep asking for x and someone of x's color
+    \item Define a sequence of xs for any deterministic algorithm.
+    \item Define what OPT should do: the cluster D and nodes d1 d2 that should be hidden from the sequence of xs?)
+    \item (somewhere) define c(x): it is easy when x is not from D: then any would do; if x is from D, then issue request between non-d1/d2; also not defined for a1,ak-1, b1,bk-1 and ak, bk
+    \item Argue that OPT does not pay for input sequence (after glueing)
+    \item lower-bound the number of times we can issue (x, c(x)) request; until it encounters either d1 or d2 in place of x; this is lower-bounded by the total number of nodes-2 (d1 and d2 are last?
+they cannot be, though, but the furthest away should be at min distance k*l/2)
+    \item also, this k*l/2 sequence is feasible, as the perfect partition exists: we can always issue this request (unless d1d2)
+    \item the total cost of ALG, comp ratio
+  \end{enumerate}
+
+  Fix a deterministic algorithm DET.
+  Fix a cluster $A$ containing nodes $a_1, \ldots, a_k$ and a cluster $B$ containing (among others) a node $b$.
+  The adversary begins with issuing $k-2$ internal requests $(a_i, a_{i+1})$, so that nodes $a_1, \ldots a_{k-1}$ form a single component.
+  The internal requests are free for OPT, and it doesn't perform additional actions at this point.
+
+  \maciek{Possibly we need a better name than $x$. Something that indicates that this changes over time, possibly tied to the name DET.}
+  We define $x$ to be the node in the configuration of DET that is co-located in one cluster with the component $a_1, \ldots, a_{k-1}$.
+  Initially, $x = a_k$.
+
+  Then, the adversary issues an external request $(x, b)$, which at this point is the request $(a_k, b)$.
+  Note that DET must relocate the component $\{a_k, b\}$ to some other cluster.
+  OPT also relocates the component $\{a_k, b\}$ to some carefuly chosen cluster $D$ (to be fixed later), and it migrates nodes $d_1$ and $d_2$ to the previous locations $a_k$ and $b$ (i.e., to the clusters $A$ and $B$).
+
+  Now, for a node $p$ we define $c(p)$ in the following way: for each node outside of $a_i$ and $b$, we define $c(p)$ to be any node that in the initial configuration was placed at the same cluster as $p$, and if $p$ belonged to $D$, then $c(p)$ is any such node that is different from $d_1$ and $d_2$.
+  \maciek{we have a cyclic reference at this point. D is not yet defined (depends on c), and c depends on d1, d2...}
+  Then, the adversary continues to issue requests $(x, c(x))$.
+\end{proof}
+
+\section{An $O(\ell)$-competitive algorithm for $k=3$}
+
+\begin{theorem}
+  The rebalancing cost for the greedy algorithm for $k=3$ is $O(1)$.
+  \maciek{Greedy = find the cheapest (closest to the current) feasible partition}
+  \label{rebalancing-cost}
+\end{theorem}
+
+\begin{proof} 
+ \maciek{Note: we consider a cost of cheapest feasible rebalancing after insertion of one edge.
+This in contrast to bounding the distance between any two reconfigurations (this theorem does not bound the latter).}
+
+  We distinguish among three configuration types of algorithm's clusters: $C_1, C_2, C_3$.
+In $C_1$ we have $3$ singleton components (this is also the initial configuration of any cluster).
+In $C_2$ we have one component of size $2$ and one component of size $1$.
+Finally, in $C_3$ we have one component of size $3$.
+
+  Consider a request $(u, v)$ and let $U, V$ be their clusters.
+  If $U=V$, then the request does not require a reconfiguration.
+  Now, we consider cases upon the type of clusters $U$ and $V$.
+  Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise the resulting component would have the size $4$, and this contradicts the existence of the perfect partition.
+  If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
+
+  Now, we focus on the case where both $U$ and $V$ are $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as the merged component would have size $4$.
+  If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
+  Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+  Note that in such case, a $C_1$-type cluster exists, as otherwise the input graph would not have a perfect partition.
+The cost of rebalancing is then at most $12$ \maciek{can show smaller constant than 12, but it does not matter that much}.
+\end{proof}
+
+\begin{theorem}
+  There exists a $O(\ell)$-competitive algorithm for $k=3$.
+\end{theorem}
+
+\begin{proof}
+  The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
+Each request issues at most one rebalancingfor the greedy algorithm, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
+\end{proof}
+
+\section{Costly rebalance example $\Omega(k^2)$}
+
+\maciek{ToDo}
+
+\section{Trash: old versions etc.}
+
+\subsection{Lower bound $\Omega(\ell)$}
+
+\maciek{The most recent version of the lower bound is in $\Omega(l\cdot k)$ section. This is an early version!}
 
@@
 \end{observation}
-
-\section{Lower bound $\Omega(k\cdot \ell)$}
-
-\maciek{Warning: sketch only, I did not finish writing it yet!}
-
-In this section we assume the Online Balanced Partition problem with the existence of a perfect partition of the input graph.
-
-\begin{theorem}
-  The competitive ratio of any deterministic algorithm for Online Balanced Partition (with perfect partition) is $\Omega(k\cdot \ell)$.
-\end{theorem}
-
-\begin{proof}
-
-  \begin{enumerate}
-    \item Cluster A component of size k-1. Define x.
-    \item The input sequence (after these glueing) is to keep asking for x and someone of x's color
-    \item Define a sequence of xs for any deterministic algorithm.
-    \item Define what OPT should do: the cluster D and nodes d1 d2 that should be hidden from the sequence of xs?)
-    \item (somewhere) define c(x): it is easy when x is not from D: then any would do; if x is from D, then issue request between non-d1/d2; also not defined for a1,ak-1, b1,bk-1 and ak, bk
-    \item Argue that OPT does not pay for input sequence (after glueing)
-    \item lower-bound the number of times we can issue (x, c(x)) request; until it encounters either d1 or d2 in place of x; this is lower-bounded by the total number of nodes-2 (d1 and d2 are last?
-they cannot be, though, but the furthest away should be at min distance k*l/2)
-    \item also, this k*l/2 sequence is feasible, as the perfect partition exists: we can always issue this request (unless d1d2)
-    \item the total cost of ALG, comp ratio
-  \end{enumerate}
-
-  Fix a deterministic algorithm DET.
-  Fix a cluster $A$ containing nodes $a_1, \ldots, a_k$ and a cluster $B$ containing (among others) a node $b$.
-  The adversary begins with issuing $k-2$ internal requests $(a_i, a_{i+1})$, so that nodes $a_1, \ldots a_{k-1}$ form a single component.
-  The internal requests are free for OPT, and it doesn't perform additional actions at this point.
-
-  \maciek{Possibly we need a better name than $x$. Something that indicates that this changes over time, possibly tied to the name DET.}
-  We define $x$ to be the node in the configuration of DET that is co-located in one cluster with the component $a_1, \ldots, a_{k-1}$.
-  Initially, $x = a_k$.
-
-  Then, the adversary issues an external request $(x, b)$, which at this point is the request $(a_k, b)$.
-  Note that DET must relocate the component $\{a_k, b\}$ to some other cluster.
-  OPT also relocates the component $\{a_k, b\}$ to some carefuly chosen cluster $D$ (to be fixed later), and it migrates nodes $d_1$ and $d_2$ to the previous locations $a_k$ and $b$ (i.e., to the clusters $A$ and $B$).
-
-  Now, for a node $p$ we define $c(p)$ in the following way: for each node outside of $a_i$ and $b$, we define $c(p)$ to be any node that in the initial configuration was placed at the same cluster as $p$, and if $p$ belonged to $D$, then $c(p)$ is any such node that is different from $d_1$ and $d_2$.
-  \maciek{we have a cyclic reference at this point. D is not yet defined (depends on c), and c depends on d1, d2...}
-  Then, the adversary continues to issue requests $(x, c(x))$.
-\end{proof}
-
-\section{An $O(\ell)$-competitive algorithm for $k=3$}
-
-\begin{theorem}
-  The rebalancing cost for the greedy algorithm for $k=3$ is $O(1)$.
-  \maciek{Greedy = find the cheapest (closest to the current) feasible partition}
-  \label{rebalancing-cost}
-\end{theorem}
-
-\begin{proof} 
- \maciek{Note: we consider a cost of cheapest feasible rebalancing after insertion of one edge.
-This in contrast to bounding the distance between any two reconfigurations (this theorem does not bound the latter).}
-
-  We distinguish among three configuration types of algorithm's clusters: $C_1, C_2, C_3$.
-In $C_1$ we have $3$ singleton components (this is also the initial configuration of any cluster).
-In $C_2$ we have one component of size $2$ and one component of size $1$.
-Finally, in $C_3$ we have one component of size $3$.
-
-  Consider a request $(u, v)$ and let $U, V$ be their clusters.
-  If $U=V$, then the request does not require a reconfiguration.
-  Now, we consider cases upon the type of clusters $U$ and $V$.
-  Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise the resulting component would have the size $4$, and this contradicts the existence of the perfect partition.
-  If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
-
-  Now, we focus on the case where both $U$ and $V$ are $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as the merged component would have size $4$.
-  If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
-  Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-  Note that in such case, a $C_1$-type cluster exists, as otherwise the input graph would not have a perfect partition.
-The cost of rebalancing is then at most $12$ \maciek{can show smaller constant than 12, but it does not matter that much}.
-\end{proof}
-
-\begin{theorem}
-  There exists a $O(\ell)$-competitive algorithm for $k=3$.
-\end{theorem}
-
-\begin{proof}
-  The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
-Each request issues at most one rebalancingfor the greedy algorithm, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
-\end{proof}
-
-\section{Costly rebalance example $\Omega(k^2)$}

### 2019-08-22 maciek.tex
Omega(kl)

@@
 \maciek{Warning: sketch only, I did not finish writing it yet!}
+
+\maciek{I assume $\alpha = 1$ to simplify calculations, but everything should hold for general $\alpha$}
 
@@
 
-  \begin{enumerate}
-    \item Cluster A component of size k-1. Define x.
-    \item The input sequence (after these glueing) is to keep asking for x and someone of x's color
-    \item Define a sequence of xs for any deterministic algorithm.
-    \item Define what OPT should do: the cluster D and nodes d1 d2 that should be hidden from the sequence of xs?)
-    \item (somewhere) define c(x): it is easy when x is not from D: then any would do; if x is from D, then issue request between non-d1/d2; also not defined for a1,ak-1, b1,bk-1 and ak, bk
-    \item Argue that OPT does not pay for input sequence (after glueing)
-    \item lower-bound the number of times we can issue (x, c(x)) request; until it encounters either d1 or d2 in place of x; this is lower-bounded by the total number of nodes-2 (d1 and d2 are last?
-they cannot be, though, but the furthest away should be at min distance k*l/2)
-    \item also, this k*l/2 sequence is feasible, as the perfect partition exists: we can always issue this request (unless d1d2)
-    \item the total cost of ALG, comp ratio
-  \end{enumerate}
+  \maciek{cluster B is special?}
+  \maciek{If DET splits a component, we issue intra-component requests until DET joins it}
 
   Fix a deterministic algorithm DET.
-  Fix a cluster $A$ containing nodes $a_1, \ldots, a_k$ and a cluster $B$ containing (among others) a node $b$.
+  Fix a cluster $A$ containing nodes $a_1, \ldots, a_k$ and a cluster $B$ containing (among others) a node $b_1$.
   The adversary begins with issuing $k-2$ internal requests $(a_i, a_{i+1})$, so that nodes $a_1, \ldots a_{k-1}$ form a single component.
   The internal requests are free for OPT, and it doesn't perform additional actions at this point.
+  The next request issued by the adversary is an external request $(a_k, b_1)$.%, which at this point is the request $(a_k, b)$.
+  Note that DET must relocate the component $\{a_k, b_1\}$ to some other cluster.
+  OPT also relocates the component $\{a_k, b_1\}$ to some carefuly chosen cluster $D$, and it evicts nodes $d^{(1)}$ and $d^{(2)}$ (to be fixed later) from $D$ to the previous locations of $a_k$ and $b_1$ (i.e., to the clusters $A$ and $B$).
+  For these actions, OPT pays $2$, and DET pays at least $2$.
+  We name the requests issued so far as $I_1$.
 
-  \maciek{Possibly we need a better name than $x$. Something that indicates that this changes over time, possibly tied to the name DET.}
-  We define $x$ to be the node in the configuration of DET that is co-located in one cluster with the component $a_1, \ldots, a_{k-1}$.
-  Initially, $x = a_k$.
+  Now, we construct a sequence of requests $I_2$, but we do not feed them to DET yet.
+  Instead, we simulate DET's actions on $I_1I_2$ to terminate the input sequence earlier, i.e., we choose a prefix $I_2'$ of $I_2$ and we feed DET the input $I_1I_2'$.
+  \maciek{We can produce $I_2'$ on-the-fly, but this was convinient to write it down this way}
 
-  Then, the adversary issues an external request $(x, b)$, which at this point is the request $(a_k, b)$.
-  Note that DET must relocate the component $\{a_k, b\}$ to some other cluster.
-  OPT also relocates the component $\{a_k, b\}$ to some carefuly chosen cluster $D$ (to be fixed later), and it migrates nodes $d_1$ and $d_2$ to the previous locations $a_k$ and $b$ (i.e., to the clusters $A$ and $B$).
+  The input $I_2$ consists of requests $(x^{(1)}, y^{(1)}), (x^{(2)},y^{(2)}),\ldots$.
+  The node $x^{(i)}$ is a node in the current configuration of DET that is co-located in one cluster with the component $a_1, \ldots, a_{k-1}$.
+  After a request $(x^{(i)}, y^{(i)})$, the node $x^{(i)}$ is a part of a non-singleton component and must be relocated to some other cluster (as its current cluster hosts a component $a_1, \ldots, a_{k-1}$).
+  Then, DET places the node $x^{(i+1)}$ in place of $x^{(i)}$, and the next request concerns that node.
+  % Initially, $x^{(1)} = a_k$, and the next node $x^{(2)}$ is the that DET locates in place of the
 
-  Now, for a node $p$ we define $c(p)$ in the following way: for each node outside of $a_i$ and $b$, we define $c(p)$ to be any node that in the initial configuration was placed at the same cluster as $p$, and if $p$ belonged to $D$, then $c(p)$ is any such node that is different from $d_1$ and $d_2$.
-  \maciek{we have a cyclic reference at this point. D is not yet defined (depends on c), and c depends on d1, d2...}
-  Then, the adversary continues to issue requests $(x, c(x))$.
+  For each $x^{(i)}$, we set $y^{(i)}$ to some node that in the intial configuration was co-located with $x^{(i)}$.
+
+  Note that in the initial configuration, nodes of one cluster are indistinguishible for DET \maciek{and hence we can order them on-the-fly}.
+
+  Now, we formally fix nodes $y^{(i)}$.
+  Consider a set of nodes from an initial configuration of a cluster $C$ (where $C \neq A, C \neq B$).
+  Formally, by $v^C_1$ we label the first node of a cluster $C$ that appears in the sequence of $x$'es at the position $i$.
+  At this point, we choose an arbitrary other node of $C$ and we label it $v^C_0$.
+  We set $y^{(i)} = v^C_0$.
+  For the following $k-2$ nodes of the cluster $C$, we label them $v^C_2, v^C_3, \ldots$ in the order of appearance in the sequence of $x$'es.
+  For a node $v^C_j = x^{(i)}$, we set $y^{(i)} = v^C_0$.
+  Some nodes of $C$ might never appear in the sequence of $x$'es, and we number them arbitrarily (with the restriction that the nodes that appear in the sequence of $x$'es have smaller number than the ones that did not).
+
+ % During the runtime of DET, by $x(A)$ we denote the node in the configuration of DET that is co-located in one cluster with the component $a_1, \ldots, a_{k-1}$.
+
+ % \maciek{Possibly we need a better name than $x(A)$.
+This should be tied to (1) DET, and (2) the component $a_1, \ldots, a_{k-1}$ (not the server A, from which it might be relocated by DET).
+Maybe we should number them, i.e., $x_1, x_2,\ldots$}
+ % During the runtime of DET, by $x(A)$ we denote the node in the configuration of DET that is co-located in one cluster with the component $a_1, \ldots, a_{k-1}$.
+ % Initially, $x(A) = a_k$.
+
+ % We consider a sequence of requests between a node $x(A)$ and some other nodes, i.e., each request concerns the node that is currently co-located with the component $a_1,\ldots,a_{k-1}$.
+ % As a result
+
+  Fix a node $m^C$ for each cluster $C$.
+
+  % Note that nodes of one cluster are indistinguishible for DET.
+
+ % For the purpose of constructing $I_2$, we order nodes of any cluster $C$: $v^C_1, \ldots, v^C_k$.
+ % For any node $v^C_i\neq v^C_1$ we define $c(v^C_i) = v^C_1$, and $c(v^C_1) = v^C_2$.
+
+  %Now, for a node $p$ we define $c(p)$ in the following way: for each node outside of $a_i$ and $b$, we define $c(p)$ to be any node that in the initial configuration was placed at the same cluster as $p$, and if $p$ belonged to $D$, then $c(p)$ is any such node that is different from $d_1$ and $d_2$.
+  %\maciek{we have a cyclic reference at this point. D is not yet defined (depends on c), and c depends on d1, d2...}
+  %Then, the adversary continues to issue requests $(x, c(x))$.
 \end{proof}
@@
 \maciek{The most recent version of the lower bound is in $\Omega(l\cdot k)$ section. This is an early version!}
+  \begin{enumerate}
+    \item Cluster A component of size k-1. Define x.
+    \item The input sequence (after these glueing) is to keep asking for x and someone of x's color
+    \item Define a sequence of xs for any deterministic algorithm.
+    \item Define what OPT should do: the cluster D and nodes d1 d2 that should be hidden from the sequence of xs?)
+    \item (somewhere) define c(x): it is easy when x is not from D: then any would do; if x is from D, then issue request between non-d1/d2; also not defined for a1,ak-1, b1,bk-1 and ak, bk
+    \item Argue that OPT does not pay for input sequence (after glueing)
+    \item lower-bound the number of times we can issue (x, c(x)) request; until it encounters either d1 or d2 in place of x; this is lower-bounded by the total number of nodes-2 (d1 and d2 are last?
+they cannot be, though, but the furthest away should be at min distance k*l/2)
+    \item also, this k*l/2 sequence is feasible, as the perfect partition exists: we can always issue this request (unless d1d2)
+    \item the total cost of ALG, comp ratio
+  \end{enumerate}
 

### 2019-08-22 maciek.tex
Omega(kl)

@@
 
-  The input $I_2$ consists of requests $(x^{(1)}, y^{(1)}), (x^{(2)},y^{(2)}),\ldots$.
+  The input $I_2$ consists of requests $(x^{(1)}, c(x^{(1)})), (x^{(2)}, c(x^{(2)})),\ldots$.
   The node $x^{(i)}$ is a node in the current configuration of DET that is co-located in one cluster with the component $a_1, \ldots, a_{k-1}$.
-  After a request $(x^{(i)}, y^{(i)})$, the node $x^{(i)}$ is a part of a non-singleton component and must be relocated to some other cluster (as its current cluster hosts a component $a_1, \ldots, a_{k-1}$).
+  After a request $(x^{(i)}, c(x^{(i)}))$, the node $x^{(i)}$ is a part of a non-singleton component and must be relocated to some other cluster (as its current cluster hosts a component $a_1, \ldots, a_{k-1}$).
   Then, DET places the node $x^{(i+1)}$ in place of $x^{(i)}$, and the next request concerns that node.
@@
 
-  For each $x^{(i)}$, we set $y^{(i)}$ to some node that in the intial configuration was co-located with $x^{(i)}$.
+  For each $x^{(i)}$, we set $c(x^{(i)})$ to some node that in the initial configuration was co-located with $x^{(i)}$.
+  While producing an input sequence, we would like to additionally order the initial nodes $V^C$ of each cluster (i.e., order the nodes that were present in each cluster in its initial configuration).
+  The ordering needs to be produced on-the-fly, as this ordering is used to produce subsequent requests.
 
-  Note that in the initial configuration, nodes of one cluster are indistinguishible for DET \maciek{and hence we can order them on-the-fly}.
+  %Note that in the initial configuration, nodes of one cluster are indistinguishible for DET \maciek{and hence we can order them on-the-fly}.
 
-  Now, we formally fix nodes $y^{(i)}$.
+  Now, we formally define $c(x^{(i)})$, and the ordering of nodes in the initial configuration.
   Consider a set of nodes from an initial configuration of a cluster $C$ (where $C \neq A, C \neq B$).
-  Formally, by $v^C_1$ we label the first node of a cluster $C$ that appears in the sequence of $x$'es at the position $i$.
-  At this point, we choose an arbitrary other node of $C$ and we label it $v^C_0$.
-  We set $y^{(i)} = v^C_0$.
-  For the following $k-2$ nodes of the cluster $C$, we label them $v^C_2, v^C_3, \ldots$ in the order of appearance in the sequence of $x$'es.
-  For a node $v^C_j = x^{(i)}$, we set $y^{(i)} = v^C_0$.
-  Some nodes of $C$ might never appear in the sequence of $x$'es, and we number them arbitrarily (with the restriction that the nodes that appear in the sequence of $x$'es have smaller number than the ones that did not).
+  By $v^C_1$ we label the first node of a cluster $C$ that appears in the sequence $\{x^{(i)}\}$.
+  At this point, we fix an arbitrary other node of $C$ and we label it $v^C_0$, and we set $c(v^C_i) = v^C_0$ for all $i$.
+  For the following $k-2$ nodes of the cluster $C$, we label them $v^C_2, v^C_3, \ldots$ in the order of appearance in the sequence $\{x^{(i)}\}$.
+  %For a node $v^C_p = x^{(j)}$, we set $y^{(j)} = v^C_0$.
+  Some nodes of $C$ might never appear in the sequence $\{x^{(i)}\}$, and we number them arbitrarily, with the restriction that the nodes that do appear in the sequence have smaller number than the ones that do not.
 

### 2019-08-22 mahmoud.tex
some corrections

@@
 Therefore,
-$OPT \geq dist(P_F, P_I) $.
-We devise an online algorithm that mimics OPT by minimizing the distance to the initial partition $P_I$ on every re-partitioning.
+$OPT \geq \Delta:= dist(P_F, P_I) $.
+We devise an online algorithm that mimics OPT by minimizing the distance to the initial partition $P_I$ with each re-partitioning.
 As a result,
-ON never ends up a partition that is father away from $P_I$ than $OPT$.
+ON never ends up in a partition that is father away from $P_I$ than $\Delta$.
 This invariant ensures that ON does not pay too much while reaching $P_F$.
@@
 		initial partition $P_I$,
-		sequence of communication edges $<\sigma_t>$ covering all ground truth components
-		\Ensure Final partition $P_F = F_1, \dots, F_{\ell}$ 
+		sequence of  edges $\sigma_1, \dots, \sigma_N$ covering all ground truth components
+		\Ensure Final partition $P_F$ 
 		\State for each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$ \label{line:initcomponents}
@@
 		\If{cluster$(C_1) \neq$ cluster$(C_2)$}		
-		\State \textit{re-partition()} \label{line:rebalance} 
+		\State \textit{re-partition($k, \ell, P_I, \mathcal{C}$)} \label{line:rebalance} 
 		\EndIf
@@
 \begin{property} \label{prop:dist<OPT}
-	Let $P$ be any partition chosen Algorithm \ref{alg:ppl} at Line $\ref{line:rebalance}$.
-	Then, $dist(P) \leq OPT$.
+	Let $P$ be any partition chosen by Algorithm \ref{alg:ppl} at Line $\ref{line:rebalance}$.
+	Then, $dist(P,P_I) \leq \Delta$.
 \end{property}
@@
 	By Property \ref{prop:dist<OPT},
-	$dist(P_{t-1}) , dist(P_{t}) \leq OPT$
-	and therefore we have	
+	$dist(P_{t-1}) , dist(P_{t}) \leq \Delta \leq OPT$
+	and thereby we have	
 	$|M| \leq 2.OPT$.
@@
 	The algorithm eventually reaches the final partition since it
-	 evaluates all possible $\ell$-way partitioning of components on every external request,
+	 evaluates all possible $\ell$-way partitioning of components on each external request,
 	including the request that completes revealing of all ground truth components.
-	There are at most $k.\ell$ calls to re-partition() and by Lemma \ref{lemma:rebalancecost} each costs at most $2.OPT$.
+	There are at most $(k-1).\ell $ calls to re-partition() and by Lemma \ref{lemma:rebalancecost} each costs at most $2.OPT$.
 	The total cost is therefore at most $2.OPT.k.\ell$ which implies the competitive ratio.

### 2019-08-22 maciek.tex
Omega(kl)

@@
 
-  \maciek{cluster B is special?}
+ We say that an online algorithm DET is component-respecting if it never reaches a configuration with a component split between clusters.
+  %Recall that if DET resides in a configuration with a component split between clusters, we issue intra-component requests until
    \maciek{If DET splits a component, we issue intra-component requests until DET joins it}
@@
   % Initially, $x^{(1)} = a_k$, and the next node $x^{(2)}$ is the that DET locates in place of the
+  The sequence continues as long as after serving the request, a perfect partition of the input graph exists.
 
@@
   Now, we formally define $c(x^{(i)})$, and the ordering of nodes in the initial configuration.
-  Consider a set of nodes from an initial configuration of a cluster $C$ (where $C \neq A, C \neq B$).
+  Consider a set of nodes from an initial configuration of a cluster $C$ (where $C \neq A$).
   By $v^C_1$ we label the first node of a cluster $C$ that appears in the sequence $\{x^{(i)}\}$.
@@
   %For a node $v^C_p = x^{(j)}$, we set $y^{(j)} = v^C_0$.
-  Some nodes of $C$ might never appear in the sequence $\{x^{(i)}\}$, and we number them arbitrarily, with the restriction that the nodes that do appear in the sequence have smaller number than the ones that do not.
+  Some nodes of $C$ might never appear in the sequence $\{x^{(i)}\}$ (e.g.
+DET might terminate prematurely or might reside in a configuration with components split), and we number them arbitrarily, with the restriction that the nodes that do appear in the sequence have smaller number than the ones that do not.
+
+  Now, we claim that given the input $I_1I_2$, every component-respecting DET 
+  % (that does not terminate prematurely and never resides in a configuration with components split)
+  must produce a sequence $\{x^{(i)}\}$ of length $L = k\cdot \ell - (k-1) - 2 - (\ell-1) - 1$.
+  Consider a request $(x^{(i)}, c(x^{(i)})$.
+  If in this configuration at least one singleton component is present (other than the components of $(x^{(i)}, c(x^{(i)})$), then serving this request is possible (i.e., the perfect partition exists), and a singleton node appears as $x^{(i+1)}$, and the sequence continues.
+  Otherwise, as we have a component $a_1, \ldots, a_{k-1}$, no perfect partition exists.
+
+  Hence, the length of a sequence $\{x^{(i)}\}$ depends on the number of singleton components.
+  Before the arrival of the first request from $I_2$, the number of singleton components is $L + (\ell-1) + 1$, and after serving the last request, the number of singleton components is $1$.
+  %Every node $x^{(i)}$ must be a singleton component, otherwise DET resides in a configuration with components split (and such DETs either leave such configuration eventually or issue an infinite cost).
+  After each request, we join this singleton component with another component.
+  This join decreases the number of singleton components by one or by two: if this is the join of two singletons $v^C_0, v^C_1$ (there are $\ell-1$ such joins: for all clusters but $A$), then the number of singletons decrease by $2$, and it decreases by $1$ otherwise.
+  We conclude that the number of requests $\{x^{(i)}\}$ to decrease the number of singletons to $1$ is $L$.
+
+  \medskip
+
+  Now, we define the input $I_2'$.
 

### 2019-08-22 maciek.tex
Omega(kl)

@@
 \begin{proof}
+
+  \maciek{Sequence $\lbrace \rbrace$ not the set notation}
 
@@
 
-  Now, we define the input $I_2'$.
+  Now, we define the input $I_2'$ as a prefix of $I_2$.
+  Recall that OPT serves the first external request $(a_k, b_1)$ by placing the component $\{a_k, b_1\}$ in some cluster $D$, and it evicts some nodes $d^{(1)}$ and $d^{(2)}$ from $D$ to the previous locations of $a_k$ and $b_1$.
+  It remains to determine the cluster $D$ and nodes $d^{(1)}$ and $d^{(2)}$,
+  and our aim is that OPT would not pay any cost for $I_2'$.
+  As nodes $d^{(1)}$ and $d^{(2)}$ are not in their initial cluster, a request between a node from (intitial configuration of) $D$ and either $d^{(1)}$ or $d^{(2)}$ would not be free for OPT.
+
+  At this point we already know the sequence $\{x^{(i)}\}$ of DET.
+  For the requests of $I_2'$ to be free for OPT, it suffices to choose $v^D_0$ that is neither $d^{(1)}$ nor $d^{(2)}$,
+  and to guarantee that $d^{(1)}$, $d^{(2)}$ does not appear in the sequence $\{x^{(i)}\}$, i.e., to finish the input sequence before that.
+  Furthermore, we would like to guarantee the length of $\{x^{(i)}\}$ to remain $\Omega(k\cdot \ell)$.
+
+  To this end, we inspect the sequences $v^{C}_i$ for each cluster, and select the cluster $C'$ whose node $v^{C'}_{k-2}$, i.e., the penultimate node appear last in the sequence $\{x^{(i)}\}$, and end the sequence $I_2'$ just before its appearance.
+  Then, we set $d^{(1)} = v^{C'}_{k-2}$, $d^{(2)} = v^{C'}_{k-1}$.
+  Such choice guarantees that requests from $I_2'$ are free for OPT.
+  Furthermore, the number of requests of $I_2'$ is at least $|I_2| - \ell - 1$: as $D$ is the last cluster to remain with two singleton clusters, all other $\ell-1$ clusters have at most $1$ singleton cluster.
+
+  Finally, OPT pays $2$, and DET pays $L - \ell - 1 = \Omega(\ell \cdot k)$.
 
@@
 
-  Fix a node $m^C$ for each cluster $C$.
+  % Fix a node $m^C$ for each cluster $C$.
 

### 2019-08-22 maciek.tex
Omega(kl)

--- 2019-08-22 maciek.tex
@@
   The adversary begins with issuing $k-2$ internal requests $(a_i, a_{i+1})$, so that nodes $a_1, \ldots a_{k-1}$ form a single component.
-  The internal requests are free for OPT, and it doesn't perform additional actions at this point.
+  The internal requests are free for OPT, and it does not perform additional actions at this point.
   The next request issued by the adversary is an external request $(a_k, b_1)$.%, which at this point is the request $(a_k, b)$.
-  Note that DET must relocate the component $\{a_k, b_1\}$ to some other cluster.
-  OPT also relocates the component $\{a_k, b_1\}$ to some carefuly chosen cluster $D$, and it evicts nodes $d^{(1)}$ and $d^{(2)}$ (to be fixed later) from $D$ to the previous locations of $a_k$ and $b_1$ (i.e., to the clusters $A$ and $B$).
+  Note that DET must relocate the component $\langle a_k, b_1\rangle$ to some other cluster.
+  OPT also relocates the component $\langle a_k, b_1\rangle$ to some carefuly chosen cluster $D$, and it evicts nodes $d^{(1)}$ and $d^{(2)}$ (to be fixed later) from $D$ to the previous locations of $a_k$ and $b_1$ (i.e., to the clusters $A$ and $B$).
   For these actions, OPT pays $2$, and DET pays at least $2$.
@@
   Consider a set of nodes from an initial configuration of a cluster $C$ (where $C \neq A$).
-  By $v^C_1$ we label the first node of a cluster $C$ that appears in the sequence $\{x^{(i)}\}$.
+  By $v^C_1$ we label the first node of a cluster $C$ that appears in the sequence $\langle x^{(i)}\rangle$.
   At this point, we fix an arbitrary other node of $C$ and we label it $v^C_0$, and we set $c(v^C_i) = v^C_0$ for all $i$.
-  For the following $k-2$ nodes of the cluster $C$, we label them $v^C_2, v^C_3, \ldots$ in the order of appearance in the sequence $\{x^{(i)}\}$.
+  For the following $k-2$ nodes of the cluster $C$, we label them $v^C_2, v^C_3, \ldots$ in the order of appearance in the sequence $\langle x^{(i)}\rangle$.
   %For a node $v^C_p = x^{(j)}$, we set $y^{(j)} = v^C_0$.
-  Some nodes of $C$ might never appear in the sequence $\{x^{(i)}\}$ (e.g.
+  Some nodes of $C$ might never appear in the sequence $\langle x^{(i)}\rangle$ (e.g.
 DET might terminate prematurely or might reside in a configuration with components split), and we number them arbitrarily, with the restriction that the nodes that do appear in the sequence have smaller number than the ones that do not.
@@
   % (that does not terminate prematurely and never resides in a configuration with components split)
-  must produce a sequence $\{x^{(i)}\}$ of length $L = k\cdot \ell - (k-1) - 2 - (\ell-1) - 1$.
-  Consider a request $(x^{(i)}, c(x^{(i)})$.
-  If in this configuration at least one singleton component is present (other than the components of $(x^{(i)}, c(x^{(i)})$), then serving this request is possible (i.e., the perfect partition exists), and a singleton node appears as $x^{(i+1)}$, and the sequence continues.
+  must produce a sequence $\langle x^{(i)}\rangle$ of length $L = k\cdot \ell - (k-1) - 2 - (\ell-1) - 1$.
+  Consider a request $(x^{(i)}, c(x^{(i)}))$.
+  If in this configuration at least one singleton component is present (other than the components of $(x^{(i)}, c(x^{(i)}))$), then serving this request is possible (i.e., the perfect partition exists), and a singleton node appears as $x^{(i+1)}$, and the sequence continues.
   Otherwise, as we have a component $a_1, \ldots, a_{k-1}$, no perfect partition exists.
 
-  Hence, the length of a sequence $\{x^{(i)}\}$ depends on the number of singleton components.
+  Hence, the length of a sequence $\langle x^{(i)}\rangle$ depends on the number of singleton components.
   Before the arrival of the first request from $I_2$, the number of singleton components is $L + (\ell-1) + 1$, and after serving the last request, the number of singleton components is $1$.
@@
   This join decreases the number of singleton components by one or by two: if this is the join of two singletons $v^C_0, v^C_1$ (there are $\ell-1$ such joins: for all clusters but $A$), then the number of singletons decrease by $2$, and it decreases by $1$ otherwise.
-  We conclude that the number of requests $\{x^{(i)}\}$ to decrease the number of singletons to $1$ is $L$.
+  We conclude that the number of requests $\langle x^{(i)}\rangle$ to decrease the number of singletons to $1$ is $L$.
 
@@
 
-  Now, we define the input $I_2'$.
+  Now, we define the input $I_2'$ as a prefix of $I_2$.
+  Recall that OPT serves the first external request $(a_k, b_1)$ by placing the component $\langle a_k, b_1\rangle$ in some cluster $D$, and it evicts some nodes $d^{(1)}$ and $d^{(2)}$ from $D$ to the previous locations of $a_k$ and $b_1$.
+  It remains to determine the cluster $D$ and nodes $d^{(1)}$ and $d^{(2)}$,
+  and our aim is that OPT would not pay any cost for $I_2'$.
+  As nodes $d^{(1)}$ and $d^{(2)}$ are not in their initial cluster, a request between a node from (intitial configuration of) $D$ and either $d^{(1)}$ or $d^{(2)}$ would not be free for OPT.
+
+  At this point we already know the sequence $\langle x^{(i)}\rangle$ of DET.
+  For the requests of $I_2'$ to be free for OPT, it suffices to choose $v^D_0$ that is neither $d^{(1)}$ nor $d^{(2)}$,
+  and to guarantee that $d^{(1)}$, $d^{(2)}$ does not appear in the sequence $\langle x^{(i)}\rangle$, i.e., to finish the input sequence before that.
+  Furthermore, we would like to guarantee the length of $\langle x^{(i)}\rangle$ to remain $\Omega(k\cdot \ell)$.
+
+  To this end, we inspect the sequences $v^{C}_i$ for each cluster, and select the cluster $C'$ whose node $v^{C'}_{k-2}$, i.e., the penultimate node appear last in the sequence $\langle x^{(i)}\rangle$, and end the sequence $I_2'$ just before its appearance.
+  Then, we set $d^{(1)} = v^{C'}_{k-2}$, $d^{(2)} = v^{C'}_{k-1}$.
+  Such choice guarantees that requests from $I_2'$ are free for OPT.
+  Furthermore, the number of requests of $I_2'$ is at least $|I_2| - \ell - 1$: as $D$ is the last cluster to remain with two singleton clusters, all other $\ell-1$ clusters have at most $1$ singleton cluster.
+
+  Finally, OPT pays $2$, and DET pays $L - \ell - 1 = \Omega(\ell \cdot k)$.
 

### 2019-08-22 maciek.tex
Omega(kl)

@@
   %Recall that if DET resides in a configuration with a component split between clusters, we issue intra-component requests until
-   \maciek{If DET splits a component, we issue intra-component requests until DET joins it}
+ If DET splits a component, we issue intra-component requests until DET joins it.
+ If it never joins a component, it is not competitive.
 
-  Fix a deterministic algorithm DET.
+  Fix a component-respecting deterministic algorithm DET.
   Fix a cluster $A$ containing nodes $a_1, \ldots, a_k$ and a cluster $B$ containing (among others) a node $b_1$.
   The adversary begins with issuing $k-2$ internal requests $(a_i, a_{i+1})$, so that nodes $a_1, \ldots a_{k-1}$ form a single component.
-  The internal requests are free for OPT, and it does not perform additional actions at this point.
-  The next request issued by the adversary is an external request $(a_k, b_1)$.%, which at this point is the request $(a_k, b)$.
-  Note that DET must relocate the component $\langle a_k, b_1\rangle$ to some other cluster.
-  OPT also relocates the component $\langle a_k, b_1\rangle$ to some carefuly chosen cluster $D$, and it evicts nodes $d^{(1)}$ and $d^{(2)}$ (to be fixed later) from $D$ to the previous locations of $a_k$ and $b_1$ (i.e., to the clusters $A$ and $B$).
+  These internal requests are free for OPT, and it does not perform additional actions at this point.
+  The next request issued by the adversary is an external request $(a_k, b_1)$. %, which at this point is the request $(a_k, b)$.
+  Note that DET must relocate the component $\langle a_k, b_1\rangle$ to some other cluster, as otherwise it is not component-respecting.
+  OPT also relocates the component $\{ a_k, b_1\}$ to some carefuly chosen cluster $D$, and it evicts nodes $d^{(1)}$ and $d^{(2)}$ (to be fixed later) from $D$ to the previous locations of $a_k$ and $b_1$ (i.e., to the clusters $A$ and $B$).
   For these actions, OPT pays $2$, and DET pays at least $2$.
-  We name the requests issued so far as $I_1$.
+  We call the requests issued so far the requests $I_1$.
 
   Now, we construct a sequence of requests $I_2$, but we do not feed them to DET yet.
-  Instead, we simulate DET's actions on $I_1I_2$ to terminate the input sequence earlier, i.e., we choose a prefix $I_2'$ of $I_2$ and we feed DET the input $I_1I_2'$.
-  \maciek{We can produce $I_2'$ on-the-fly, but this was convinient to write it down this way}
+  Instead, we simulate DET's actions on $I_1I_2$ to decide at which point to terminate the input sequence earlier, i.e., we choose a prefix $I_2'$ of $I_2$ and we feed DET the input $I_1I_2'$.
+  %\maciek{We can produce $I_2'$ on-the-fly, but this was convinient to write it down this way}
 
@@
 
-  For each $x^{(i)}$, we set $c(x^{(i)})$ to some node that in the initial configuration was co-located with $x^{(i)}$.
+  For each $x^{(i)}$, we set $c(x^{(i)})$ to some (to-be-determined) node that in the initial configuration was co-located with $x^{(i)}$.
   While producing an input sequence, we would like to additionally order the initial nodes $V^C$ of each cluster (i.e., order the nodes that were present in each cluster in its initial configuration).
@@
   %For a node $v^C_p = x^{(j)}$, we set $y^{(j)} = v^C_0$.
-  Some nodes of $C$ might never appear in the sequence $\langle x^{(i)}\rangle$ (e.g.
-DET might terminate prematurely or might reside in a configuration with components split), and we number them arbitrarily, with the restriction that the nodes that do appear in the sequence have smaller number than the ones that do not.
+  If some nodes of $C$ might never appear in the sequence $\langle x^{(i)}\rangle$ %(e.g. DET might terminate prematurely or might reside in a configuration with components split)
+  , we number them arbitrarily, with the restriction that the nodes that do appear in the sequence have smaller number than the ones that do not \maciek{This should not happen at all for component-respecting DET, though}
 

### 2019-08-22 maciek.tex
Omega(kl)

@@
   %For a node $v^C_p = x^{(j)}$, we set $y^{(j)} = v^C_0$.
-  If some nodes of $C$ might never appear in the sequence $\langle x^{(i)}\rangle$ %(e.g. DET might terminate prematurely or might reside in a configuration with components split)
+  If some nodes of $C$ never appear in the sequence $\langle x^{(i)}\rangle$ %(e.g. DET might terminate prematurely or might reside in a configuration with components split)
   , we number them arbitrarily, with the restriction that the nodes that do appear in the sequence have smaller number than the ones that do not \maciek{This should not happen at all for component-respecting DET, though}
@@
   Consider a request $(x^{(i)}, c(x^{(i)}))$.
-  If in this configuration at least one singleton component is present (other than the components of $(x^{(i)}, c(x^{(i)}))$), then serving this request is possible (i.e., the perfect partition exists), and a singleton node appears as $x^{(i+1)}$, and the sequence continues.
-  Otherwise, as we have a component $a_1, \ldots, a_{k-1}$, no perfect partition exists.
+  If in this configuration at least one singleton component is present (other than the components of $(x^{(i)}, c(x^{(i)}))$), then serving this request is possible (i.e., the perfect partition exists), and hence a singleton node appears as $x^{(i+1)}$ and the sequence continues.
+  Otherwise, no perfect partition exists, as we have a component $\{a_1, \ldots, a_{k-1}\}$ of size $k-1$.
 
-  Hence, the length of a sequence $\langle x^{(i)}\rangle$ depends on the number of singleton components.
+  Hence, the length of a sequence $\langle x^{(i)}\rangle$ is closely related to the number of singleton components.
   Before the arrival of the first request from $I_2$, the number of singleton components is $L + (\ell-1) + 1$, and after serving the last request, the number of singleton components is $1$.
@@
   After each request, we join this singleton component with another component.
-  This join decreases the number of singleton components by one or by two: if this is the join of two singletons $v^C_0, v^C_1$ (there are $\ell-1$ such joins: for all clusters but $A$), then the number of singletons decrease by $2$, and it decreases by $1$ otherwise.
+  This join decreases the number of singleton components by either $1$ or $2$: if we join two singletons $\{v^C_0\}, \{v^C_1\}$ (there are $\ell-1$ such joins: for all clusters but $A$), then the number of singletons decrease by $2$, and it decreases by $1$ otherwise.
   We conclude that the number of requests $\langle x^{(i)}\rangle$ to decrease the number of singletons to $1$ is $L$.
@@
 
-  To this end, we inspect the sequences $v^{C}_i$ for each cluster, and select the cluster $C'$ whose node $v^{C'}_{k-2}$, i.e., the penultimate node appear last in the sequence $\langle x^{(i)}\rangle$, and end the sequence $I_2'$ just before its appearance.
+  To this end, we inspect the sequences $v^{C}_i$ for each cluster, and select the cluster $C'$ whose node $v^{C'}_{k-2}$, i.e., the penultimate node appears last in the sequence $\langle x^{(i)}\rangle$, and we end the sequence $I_2'$ just before its appearance.
   Then, we set $d^{(1)} = v^{C'}_{k-2}$, $d^{(2)} = v^{C'}_{k-1}$.
@@
   Finally, OPT pays $2$, and DET pays $L - \ell - 1 = \Omega(\ell \cdot k)$.
+
+  \maciek{ToDo: check all off-by-one errors}
+  \maciek{ToDo: check if the existence of $b_1$ was considered in all calculations}
 

### 2019-08-22 maciek.tex
Omega(kl)

@@
 
-  To this end, we inspect the sequences $v^{C}_i$ for each cluster, and select the cluster $C'$ whose node $v^{C'}_{k-2}$, i.e., the penultimate node appears last in the sequence $\langle x^{(i)}\rangle$, and we end the sequence $I_2'$ just before its appearance.
-  Then, we set $d^{(1)} = v^{C'}_{k-2}$, $d^{(2)} = v^{C'}_{k-1}$.
+  To this end, we inspect the sequences $v^{C}_i$ for each cluster, and select such the cluster $D$ whose node $v^{D}_{k-2}$, i.e., the penultimate node appears last in the sequence $\langle x^{(i)}\rangle$, and we end the sequence $I_2'$ just before its appearance.
+  Then, we set $d^{(1)} = v^{D}_{k-2}$ and $d^{(2)} = v^{D}_{k-1}$.
   Such choice guarantees that requests from $I_2'$ are free for OPT.
@@
 
-  Finally, OPT pays $2$, and DET pays $L - \ell - 1 = \Omega(\ell \cdot k)$.
+  We conclude that for the input $I_1I_2'$ OPT pays $2$, and DET pays $L - \ell - 1 = \Omega(\ell \cdot k)$.
 

### 2019-08-22 maciek.tex
Omega(kl)

@@
   At this point we already know the sequence $\langle x^{(i)}\rangle$ of DET.
-  For the requests of $I_2'$ to be free for OPT, it suffices to choose $v^D_0$ that is neither $d^{(1)}$ nor $d^{(2)}$,
+  For the requests of $I_2'$ to be free for OPT, it suffices to choose $v^D_0$ that is neither $d^{(1)}$ nor $d^{(2)}$ (which is always possible for $k>3$),
   and to guarantee that $d^{(1)}$, $d^{(2)}$ does not appear in the sequence $\langle x^{(i)}\rangle$, i.e., to finish the input sequence before that.

### 2019-08-22 mahmoud.tex
Minor fixes

@@
 In other words,
-at least $dist(P, P_I)/2$ node swaps are required in order to reach the a partition $P$ from $P_I$.
+at least $dist(P, P_I)/2$ node swaps are required in order to reach the partition $P$ from $P_I$.
 Therefore,
@@
 As a result,
-ON never ends up in a partition that is father away from $P_I$ than $\Delta$.
+ON never ends up in a partition that is further away from $P_I$ than $\Delta$.
 This invariant ensures that ON does not pay too much while reaching $P_F$.
@@
 	including the request that completes revealing of all ground truth components.
-	There are at most $(k-1).\ell $ calls to re-partition() and by Lemma \ref{lemma:rebalancecost} each costs at most $2.OPT$.
+	There are at most $(k-1).\ell < k.\ell $ calls to \emph{re-partition()} and by Lemma \ref{lemma:rebalancecost} each costs at most $2.OPT$.
 	The total cost is therefore at most $2.OPT.k.\ell$ which implies the competitive ratio.

### 2019-08-22 appendix.tex
Proposition of a layout of the paper

--- 2019-08-22 maciek.tex
@@
-\section{Lower bound $\Omega(k\cdot \ell)$}
+\section{Trash: old versions etc.}
 
-\maciek{Warning: sketch only, I did not finish writing it yet!}
-
-\maciek{I assume $\alpha = 1$ to simplify calculations, but everything should hold for general $\alpha$}
-
-In this section we consider the Online Balanced Partition problem.
-We show a lower bound of $\Omega(k \cdot \ell)$ that uses an input that has a perfect partition.
-Therefore, this result is a lower bound for the general Online Balanced Partition problem, as well as a lower bound for Online Balanced Partition with perfect partition.
-The latter complements an upper bound $O(k \cdot \ell)$ for Online Balanced Partition with perfect partition that we present in Section TODO.
-
-\begin{theorem}
-  The competitive ratio of any deterministic algorithm for Online Balanced Partition is $\Omega(k\cdot \ell)$.
-\end{theorem}
-
-\begin{proof}
-
- We say that an online algorithm DET is component-respecting if it never reaches a configuration with a component split between clusters.
-  %Recall that if DET resides in a configuration with a component split between clusters, we issue intra-component requests until
- If DET splits a component, we issue intra-component requests until DET joins it.
- If it never joins a component, it is not competitive.
-
-  Fix a component-respecting deterministic algorithm DET.
-  Fix a cluster $A$ containing nodes $a_1, \ldots, a_k$ and a cluster $B$ containing (among others) a node $b_1$.
-  The adversary begins with issuing $k-2$ internal requests $(a_i, a_{i+1})$, so that nodes $a_1, \ldots a_{k-1}$ form a single component.
-  These internal requests are free for OPT, and it does not perform additional actions at this point.
-  The next request issued by the adversary is an external request $(a_k, b_1)$. %, which at this point is the request $(a_k, b)$.
-  Note that DET must relocate the component $\langle a_k, b_1\rangle$ to some other cluster, as otherwise it is not component-respecting.
-  OPT also relocates the component $\{ a_k, b_1\}$ to some carefuly chosen cluster $D$, and it evicts nodes $d^{(1)}$ and $d^{(2)}$ (to be fixed later) from $D$ to the previous locations of $a_k$ and $b_1$ (i.e., to the clusters $A$ and $B$).
-  For these actions, OPT pays $2$, and DET pays at least $2$.
-  We call the requests issued so far the requests $I_1$.
-
-  Now, we construct a sequence of requests $I_2$, but we do not feed them to DET yet.
-  Instead, we simulate DET's actions on $I_1I_2$ to decide at which point to terminate the input sequence earlier, i.e., we choose a prefix $I_2'$ of $I_2$ and we feed DET the input $I_1I_2'$.
-  %\maciek{We can produce $I_2'$ on-the-fly, but this was convinient to write it down this way}
-
-  The input $I_2$ consists of requests $(x^{(1)}, c(x^{(1)})), (x^{(2)}, c(x^{(2)})),\ldots$.
-  The node $x^{(i)}$ is a node in the current configuration of DET that is co-located in one cluster with the component $a_1, \ldots, a_{k-1}$.
-  After a request $(x^{(i)}, c(x^{(i)}))$, the node $x^{(i)}$ is a part of a non-singleton component and must be relocated to some other cluster (as its current cluster hosts a component $a_1, \ldots, a_{k-1}$).
-  Then, DET places the node $x^{(i+1)}$ in place of $x^{(i)}$, and the next request concerns that node.
-  % Initially, $x^{(1)} = a_k$, and the next node $x^{(2)}$ is the that DET locates in place of the
-  The sequence continues as long as after serving the request, a perfect partition of the input graph exists.
-
-  For each $x^{(i)}$, we set $c(x^{(i)})$ to some (to-be-determined) node that in the initial configuration was co-located with $x^{(i)}$.
-  While producing an input sequence, we would like to additionally order the initial nodes $V^C$ of each cluster (i.e., order the nodes that were present in each cluster in its initial configuration).
-  The ordering needs to be produced on-the-fly, as this ordering is used to produce subsequent requests.
-
-  %Note that in the initial configuration, nodes of one cluster are indistinguishible for DET \maciek{and hence we can order them on-the-fly}.
-
-  Now, we formally define $c(x^{(i)})$, and the ordering of nodes in the initial configuration.
-  Consider a set of nodes from an initial configuration of a cluster $C$ (where $C \neq A$).
-  By $v^C_1$ we label the first node of a cluster $C$ that appears in the sequence $\langle x^{(i)}\rangle$.
-  At this point, we fix an arbitrary other node of $C$ and we label it $v^C_0$, and we set $c(v^C_i) = v^C_0$ for all $i$.
-  For the following $k-2$ nodes of the cluster $C$, we label them $v^C_2, v^C_3, \ldots$ in the order of appearance in the sequence $\langle x^{(i)}\rangle$.
-  %For a node $v^C_p = x^{(j)}$, we set $y^{(j)} = v^C_0$.
-  If some nodes of $C$ never appear in the sequence $\langle x^{(i)}\rangle$ %(e.g. DET might terminate prematurely or might reside in a configuration with components split)
-  , we number them arbitrarily, with the restriction that the nodes that do appear in the sequence have smaller number than the ones that do not \maciek{This should not happen at all for component-respecting DET, though}
-
-  Now, we claim that given the input $I_1I_2$, every component-respecting DET 
-  % (that does not terminate prematurely and never resides in a configuration with components split)
-  must produce a sequence $\langle x^{(i)}\rangle$ of length $L = k\cdot \ell - (k-1) - 2 - (\ell-1) - 1$.
-  Consider a request $(x^{(i)}, c(x^{(i)}))$.
-  If in this configuration at least one singleton component is present (other than the components of $(x^{(i)}, c(x^{(i)}))$), then serving this request is possible (i.e., the perfect partition exists), and hence a singleton node appears as $x^{(i+1)}$ and the sequence continues.
-  Otherwise, no perfect partition exists, as we have a component $\{a_1, \ldots, a_{k-1}\}$ of size $k-1$.
-
-  Hence, the length of a sequence $\langle x^{(i)}\rangle$ is closely related to the number of singleton components.
-  Before the arrival of the first request from $I_2$, the number of singleton components is $L + (\ell-1) + 1$, and after serving the last request, the number of singleton components is $1$.
-  %Every node $x^{(i)}$ must be a singleton component, otherwise DET resides in a configuration with components split (and such DETs either leave such configuration eventually or issue an infinite cost).
-  After each request, we join this singleton component with another component.
-  This join decreases the number of singleton components by either $1$ or $2$: if we join two singletons $\{v^C_0\}, \{v^C_1\}$ (there are $\ell-1$ such joins: for all clusters but $A$), then the number of singletons decrease by $2$, and it decreases by $1$ otherwise.
-  We conclude that the number of requests $\langle x^{(i)}\rangle$ to decrease the number of singletons to $1$ is $L$.
-
-  \medskip
-
-  Now, we define the input $I_2'$ as a prefix of $I_2$.
-  Recall that OPT serves the first external request $(a_k, b_1)$ by placing the component $\langle a_k, b_1\rangle$ in some cluster $D$, and it evicts some nodes $d^{(1)}$ and $d^{(2)}$ from $D$ to the previous locations of $a_k$ and $b_1$.
-  It remains to determine the cluster $D$ and nodes $d^{(1)}$ and $d^{(2)}$,
-  and our aim is that OPT would not pay any cost for $I_2'$.
-  As nodes $d^{(1)}$ and $d^{(2)}$ are not in their initial cluster, a request between a node from (intitial configuration of) $D$ and either $d^{(1)}$ or $d^{(2)}$ would not be free for OPT.
-
-  At this point we already know the sequence $\langle x^{(i)}\rangle$ of DET.
-  For the requests of $I_2'$ to be free for OPT, it suffices to choose $v^D_0$ that is neither $d^{(1)}$ nor $d^{(2)}$,
-  and to guarantee that $d^{(1)}$, $d^{(2)}$ does not appear in the sequence $\langle x^{(i)}\rangle$, i.e., to finish the input sequence before that.
-  Furthermore, we would like to guarantee the length of $\langle x^{(i)}\rangle$ to remain $\Omega(k\cdot \ell)$.
-
-  To this end, we inspect the sequences $v^{C}_i$ for each cluster, and select the cluster $C'$ whose node $v^{C'}_{k-2}$, i.e., the penultimate node appears last in the sequence $\langle x^{(i)}\rangle$, and we end the sequence $I_2'$ just before its appearance.
-  Then, we set $d^{(1)} = v^{C'}_{k-2}$, $d^{(2)} = v^{C'}_{k-1}$.
-  Such choice guarantees that requests from $I_2'$ are free for OPT.
-  Furthermore, the number of requests of $I_2'$ is at least $|I_2| - \ell - 1$: as $D$ is the last cluster to remain with two singleton clusters, all other $\ell-1$ clusters have at most $1$ singleton cluster.
-
-  Finally, OPT pays $2$, and DET pays $L - \ell - 1 = \Omega(\ell \cdot k)$.
-
-  \maciek{ToDo: check all off-by-one errors}
-  \maciek{ToDo: check if the existence of $b_1$ was considered in all calculations}
-
- % During the runtime of DET, by $x(A)$ we denote the node in the configuration of DET that is co-located in one cluster with the component $a_1, \ldots, a_{k-1}$.
-
- % \maciek{Possibly we need a better name than $x(A)$.
-This should be tied to (1) DET, and (2) the component $a_1, \ldots, a_{k-1}$ (not the server A, from which it might be relocated by DET).
-Maybe we should number them, i.e., $x_1, x_2,\ldots$}
- % During the runtime of DET, by $x(A)$ we denote the node in the configuration of DET that is co-located in one cluster with the component $a_1, \ldots, a_{k-1}$.
- % Initially, $x(A) = a_k$.
-
- % We consider a sequence of requests between a node $x(A)$ and some other nodes, i.e., each request concerns the node that is currently co-located with the component $a_1,\ldots,a_{k-1}$.
- % As a result
-
-  % Fix a node $m^C$ for each cluster $C$.
-
-  % Note that nodes of one cluster are indistinguishible for DET.
-
- % For the purpose of constructing $I_2$, we order nodes of any cluster $C$: $v^C_1, \ldots, v^C_k$.
- % For any node $v^C_i\neq v^C_1$ we define $c(v^C_i) = v^C_1$, and $c(v^C_1) = v^C_2$.
-
-  %Now, for a node $p$ we define $c(p)$ in the following way: for each node outside of $a_i$ and $b$, we define $c(p)$ to be any node that in the initial configuration was placed at the same cluster as $p$, and if $p$ belonged to $D$, then $c(p)$ is any such node that is different from $d_1$ and $d_2$.
-  %\maciek{we have a cyclic reference at this point. D is not yet defined (depends on c), and c depends on d1, d2...}
-  %Then, the adversary continues to issue requests $(x, c(x))$.
-\end{proof}
-
-\section{An $O(\ell)$-competitive algorithm for $k=3$}
-
-\begin{theorem}
-  The rebalancing cost for the greedy algorithm for $k=3$ is $O(1)$.
-  \maciek{Greedy = find the cheapest (closest to the current) feasible partition}
-  \label{rebalancing-cost}
-\end{theorem}
-
-\begin{proof} 
- \maciek{Note: we consider a cost of cheapest feasible rebalancing after insertion of one edge.
-This in contrast to bounding the distance between any two reconfigurations (this theorem does not bound the latter).}
-
-  We distinguish among three configuration types of algorithm's clusters: $C_1, C_2, C_3$.
-In $C_1$ we have $3$ singleton components (this is also the initial configuration of any cluster).
-In $C_2$ we have one component of size $2$ and one component of size $1$.
-Finally, in $C_3$ we have one component of size $3$.
-
-  Consider a request $(u, v)$ and let $U, V$ be their clusters.
-  If $U=V$, then the request does not require a reconfiguration.
-  Now, we consider cases upon the type of clusters $U$ and $V$.
-  Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise the resulting component would have the size $4$, and this contradicts the existence of the perfect partition.
-  If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
-
-  Now, we focus on the case where both $U$ and $V$ are $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as the merged component would have size $4$.
-  If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
-  Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-  Note that in such case, a $C_1$-type cluster exists, as otherwise the input graph would not have a perfect partition.
-The cost of rebalancing is then at most $12$ \maciek{can show smaller constant than 12, but it does not matter that much}.
-\end{proof}
-
-\begin{theorem}
-  There exists a $O(\ell)$-competitive algorithm for $k=3$.
-\end{theorem}
-
-\begin{proof}
-  The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
-Each request issues at most one rebalancingfor the greedy algorithm, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
-\end{proof}
-
-\section{Costly rebalance example $\Omega(k^2)$}
-
-\maciek{ToDo}
-
-\section{Trash: old versions etc.}
+\begin{comment}
 
@@
 \end{proof}
-
-\begin{observation}
-  The lower bound of $\Omega(\ell)$ can be generalized for arbitrary $k = c\cdot 3$.
-  The adversary starts with a serie of ``glueing requests'', so that each cluster contains 3 components of size $k/3$.
-  We repeat the argument of Theorem~\ref{th:lb_omega_l} using these glued components instead of singletons.
-  OPT pays $O(k)$ for migration of one such component, and ALG pays $\Omega(k \cdot \ell)$.
-  \label{obs:lb_general_k}
-\end{observation}
-
-\begin{observation}
-  The lower bound of $\Omega(\ell)$ holds even with $1/3-1/k$-augmentation, i.e., in scenarios where the capacity of each cluster of the algorithm is $(1+1/3)\cdot k - 1$, while the capacity of OPT's clusters is $k$.
-  To this end we use the construction from Observation~\ref{obs:lb_general_k}.
-  We observe that each cluster of the algorithm can fit exactly $3$ components of size $k/3$.
-  \maciek{Note that this does not contradict the existence of $O((1+1/\epsilon)\cdot k \cdot \log(k))$ algorithm of Avin, Bienkowski et al, as it requires $2+\epsilon$-augmentation.}
-\end{observation}
+\end{comment}

### 2019-08-22 lower.tex
Proposition of a layout of the paper

--- 2019-08-22 maciek.tex
@@
+\section{Warm-up: lower-bound $\Omega(\ell)$ for $k=3$}
+
+\maciek{The intention of this section is to present a simpler version of a lower bound (that should be skipped if we are short on space) AND to present a lower bound for the case with augmentation and discuss the gap}
+
+\begin{theorem}
+  The competitive ratio of any deterministic algorithm for Online Balanced Partition with $k=3$ is $\Omega(\ell)$.
+  \label{th:lb_omega_l}
+\end{theorem}
+
+\begin{proof}
+  \maciek{ToDo}
+\end{proof}
+
+\begin{observation}
+  The lower bound of $\Omega(\ell)$ can be generalized for arbitrary $k = c\cdot 3$.
+  The adversary starts with a serie of ``glueing requests'', so that each cluster contains 3 components of size $k/3$.
+  We repeat the argument of Theorem~\ref{th:lb_omega_l} using these glued components instead of singletons.
+  OPT pays $O(k)$ for migration of one such component, and ALG pays $\Omega(k \cdot \ell)$.
+  \label{obs:lb_general_k}
+\end{observation}
+
+\begin{observation}
+  The lower bound of $\Omega(\ell)$ holds even with $1/3-1/k$-augmentation, i.e., in scenarios where the capacity of each cluster of the algorithm is $(1+1/3)\cdot k - 1$, while the capacity of OPT's clusters is $k$.
+  To this end we use the construction from Observation~\ref{obs:lb_general_k}.
+  We observe that each cluster of the algorithm can fit exactly $3$ components of size $k/3$.
+\end{observation}
+
+  \maciek{Note that this does not contradict the existence of $O((1+1/\epsilon)\cdot k \cdot \log(k))$ algorithm of Avin, Bienkowski et al, as the algorithm requires $2+\epsilon$-augmentation.}
+
+  \maciek{Note that we have a gap here. What is the maximum augmentation (between 1+1/3 and 2) that still has the $\Omega(\ell)$ lower bound?}
+
 \section{Lower bound $\Omega(k\cdot \ell)$}
-
-\maciek{Warning: sketch only, I did not finish writing it yet!}
 
@@
 \begin{proof}
-
- We say that an online algorithm DET is component-respecting if it never reaches a configuration with a component split between clusters.
-  %Recall that if DET resides in a configuration with a component split between clusters, we issue intra-component requests until
- If DET splits a component, we issue intra-component requests until DET joins it.
- If it never joins a component, it is not competitive.
 
@@
 
-  We conclude that for the input $I_1I_2'$ OPT pays $2$, and DET pays $L - \ell - 1 = \Omega(\ell \cdot k)$.
+  We conclude that for the input $I_1I_2'$ OPT pays $2$, and DET pays at least $2$ for each of $L - \ell - 1 = \Omega(\ell \cdot k)$ requests of $I_2'$.
 
@@
 \end{proof}
-
-\section{An $O(\ell)$-competitive algorithm for $k=3$}
-
-\begin{theorem}
-  The rebalancing cost for the greedy algorithm for $k=3$ is $O(1)$.
-  \maciek{Greedy = find the cheapest (closest to the current) feasible partition}
-  \label{rebalancing-cost}
-\end{theorem}
-
-\begin{proof} 
- \maciek{Note: we consider a cost of cheapest feasible rebalancing after insertion of one edge.
-This in contrast to bounding the distance between any two reconfigurations (this theorem does not bound the latter).}
-
-  We distinguish among three configuration types of algorithm's clusters: $C_1, C_2, C_3$.
-In $C_1$ we have $3$ singleton components (this is also the initial configuration of any cluster).
-In $C_2$ we have one component of size $2$ and one component of size $1$.
-Finally, in $C_3$ we have one component of size $3$.
-
-  Consider a request $(u, v)$ and let $U, V$ be their clusters.
-  If $U=V$, then the request does not require a reconfiguration.
-  Now, we consider cases upon the type of clusters $U$ and $V$.
-  Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise the resulting component would have the size $4$, and this contradicts the existence of the perfect partition.
-  If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
-
-  Now, we focus on the case where both $U$ and $V$ are $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as the merged component would have size $4$.
-  If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
-  Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-  Note that in such case, a $C_1$-type cluster exists, as otherwise the input graph would not have a perfect partition.
-The cost of rebalancing is then at most $12$ \maciek{can show smaller constant than 12, but it does not matter that much}.
-\end{proof}
-
-\begin{theorem}
-  There exists a $O(\ell)$-competitive algorithm for $k=3$.
-\end{theorem}
-
-\begin{proof}
-  The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
-Each request issues at most one rebalancingfor the greedy algorithm, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
-\end{proof}
-
-\section{Costly rebalance example $\Omega(k^2)$}
-
-\maciek{ToDo}
-
-\section{Trash: old versions etc.}
-
-\subsection{Lower bound $\Omega(\ell)$}
-
-\maciek{The most recent version of the lower bound is in $\Omega(l\cdot k)$ section. This is an early version!}
-  \begin{enumerate}
-    \item Cluster A component of size k-1. Define x.
-    \item The input sequence (after these glueing) is to keep asking for x and someone of x's color
-    \item Define a sequence of xs for any deterministic algorithm.
-    \item Define what OPT should do: the cluster D and nodes d1 d2 that should be hidden from the sequence of xs?)
-    \item (somewhere) define c(x): it is easy when x is not from D: then any would do; if x is from D, then issue request between non-d1/d2; also not defined for a1,ak-1, b1,bk-1 and ak, bk
-    \item Argue that OPT does not pay for input sequence (after glueing)
-    \item lower-bound the number of times we can issue (x, c(x)) request; until it encounters either d1 or d2 in place of x; this is lower-bounded by the total number of nodes-2 (d1 and d2 are last?
-they cannot be, though, but the furthest away should be at min distance k*l/2)
-    \item also, this k*l/2 sequence is feasible, as the perfect partition exists: we can always issue this request (unless d1d2)
-    \item the total cost of ALG, comp ratio
-  \end{enumerate}
-
-\maciek{I assume $\alpha = 1$ to simplify calculations, but everything should hold for general $\alpha$}
-
-Model assumes the existence of a perfect partition of the input graph.
-We present a lower bound for the competitive ratio of $\Omega(\ell)$ for any deterministic algorithm.
-To this end, we use $k = 3$, i.e., the capacity of each server is $3$.
-
-\begin{theorem}
-  The competitive ratio of any deterministic algorithm for Online Balanced Partition (assuming the existence of a perfect partition of the input graph) is $\Omega(\ell)$.
-  \label{th:lb_omega_l}
-\end{theorem}
-%Note that in the scenario with a perfect partition, both OPT and ALG always pays at most $O(\ell)$.
-
-\begin{proof}
-We say that nodes that belong to the same cluster in the initial configuration \emph{have the same color}.
-
-Consider clusters $A$ and $B$ containing nodes $a_1, a_2, a_3$ and $b_1, b_2, b_3$, respectively.
-The adversary begins with issuing three requests: $(a_1, a_2)$, $(b_1, b_2)$, and $(a_3, b_3)$.
-Note that the algorithm is forced to move at least one of these components to a third cluster:
-in general, if the algorithm does not keep each component within one cluster, the adversary might continue issuing requests among nodes of this component.
-
-The algorithm places these three components in different clusters, and for this it pays at least $2$.
-By $x$ (resp. $y$) we denote the third node from the cluster containing the component $(a_1, a_2)$ (resp. $(b_1, b_2)$).
-%These nodes $x$ and $y$ are of the algorithm's choice, but.
-During the execution of the algorithm, these nodes change, and by $x$ and $y$ we always denote the current nodes of the algorithm colocated with components $(a_1, a_2)$ and $(b_1, b_2)$.
-
-%Let's call the cluster of $(a_1, a_2)$ the cluster $A_{ALG}$, the cluster of $(b_1, b_2)$ the cluster $B_{ALG}$ and the cluster of $(a_3, b_3)$ the cluster $C$.
-The configuration of ALG's clusters is then: $A_{ALG}$ contains $a_1, a_2,x$, the cluster $B_{ALG}$ contains $b_1, b_2,y$ and the cluster $C$ contains $a_3, b_3, z$, for some $x, y, z$ (of the algorithm's choice).
-
-Our idea is that OPT pays for the request $(a_3, b_3)$ only (i.e., it moves this pair to a carefuly chosen cluster that we denote by $D$), and for OPT all other requests are free. 
-OPT places these three components in clusters $A, B$ and $D$, paying $2$ for migrating $a_3$ and $b_3$.
-The configuration of OPT's clusters is then: $A$ contains $a_1, a_2,d_1$, $B$ contains $b_1, b_2,d_2$ and $D$ contains $a_3, b_3, d_3$.
-Note that the requests between nodes of the same color are free for OPT, with the exception of nodes from $A$, $B$, and most notably $D$.
-
-OPT has the freedom of choice of the cluster $D$ depending on the choices of the algorithm.
-We choose $D$ in such way that initially it does not contain neither $x$ nor $y$, and furthermore that the algorithm places any of nodes from $D$ at $x$'s or $y$'s place as the last possible move.
-In other words, we inspect the sequence of $x$ and $y$ of the algorithm, and we choose $D$ to be the last cluster whose initial contents appears as the $x$ and $y$.
-As there are $\ell$ clusters to choose from, ALG has to serve at least $\Omega(\ell)$ requests before any of $D$'s contents appear as $x$ or $y$.
-
-From this point, OPT does not change its configuration, and the adversary continues to issue a request between $x$ and any other node with the same color as $x$
-(note that by $x$ we might denote different node in every iteration, as $x$ is defined as a node colocated with a component $(a_1, a_2)$).
-Upon receiving this request, ALG pays for it the cost at least $1$ and it increases its number of clusters that contain non-singleton components by $1$.
-This procedure ends when ALG has no more clusters with singleton-only components left or ALG manages to put $d_1, d_2$ or $d_3$ in place of either $x$ or $y$.
-
-At the begining, ALG has $\ell$ singleton-only components and each request increases their number by $1$.
-Furthermore, before reaching either $d_1, d_2$ or $d_3$ in place of either $x$ or $y$, the algorithm must cycle through $\Omega(\ell)$ nodes from other clusters.
-This can be guaranteed as the choice of $D$ can depend on the actions of the deterministic algorithm.
-In total, OPT pays $2$, and moreover ALG pays $\Omega(\ell)$, as it peforms a migration of two nodes for each request besides the first two.
-
-%The next action of the adversary depends on the choice of $x, y, z$.
-%If $x$ and $y$ have different color (i.e., initially they belong to different clusters), then we issue a request between $x$ and any other node with the same color as $x$.
-As $x$ do not intersect with $D$, the OPT has them co-located in one server, and it pays no cost for such requests.
-We issue such requests until $x$ and $y$ have the same color.
-
-%Note that the adversary can foresee the actions of the algorithm, and may choose $D$ in such way, that $x$ cycles through all other clusters before reaching $D$, paying the cost of $\Omega(k)$, which would already suffice to show the lower bound.
-
-%Then, $x$ and $y$ have the same color, and we issue a request $(x, y)$.
-As these have the same color (and there are different than $d_1, d_2$ and $d_3$), OPT pays no cost for this request, and ALG pays $2$ for migrating $x$ and $y$ to other cluster.
-%We repeat the entire process, i.e., we force ALG to choose $x$ and $y$ of the same color, and issue such request that is free for OPT.
-This process can continue until ALG has no more clusters with singleton components, or it pays the cost $\Omega(k)$ already.
-
-\end{proof}
-
-\begin{observation}
-  The lower bound of $\Omega(\ell)$ can be generalized for arbitrary $k = c\cdot 3$.
-  The adversary starts with a serie of ``glueing requests'', so that each cluster contains 3 components of size $k/3$.
-  We repeat the argument of Theorem~\ref{th:lb_omega_l} using these glued components instead of singletons.
-  OPT pays $O(k)$ for migration of one such component, and ALG pays $\Omega(k \cdot \ell)$.
-  \label{obs:lb_general_k}
-\end{observation}
-
-\begin{observation}
-  The lower bound of $\Omega(\ell)$ holds even with $1/3-1/k$-augmentation, i.e., in scenarios where the capacity of each cluster of the algorithm is $(1+1/3)\cdot k - 1$, while the capacity of OPT's clusters is $k$.
-  To this end we use the construction from Observation~\ref{obs:lb_general_k}.
-  We observe that each cluster of the algorithm can fit exactly $3$ components of size $k/3$.
-  \maciek{Note that this does not contradict the existence of $O((1+1/\epsilon)\cdot k \cdot \log(k))$ algorithm of Avin, Bienkowski et al, as it requires $2+\epsilon$-augmentation.}
-\end{observation}

### 2019-08-22 paper.tex
Merge into one file
Proposition of a layout of the paper

--- 2019-08-19 maciek.tex
@@
 \usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts,mathtools}
+\usepackage{amsmath,amssymb,amsfonts,mathtools,amsthm}
 \usepackage[noend]{algpseudocode}
@@
 \newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}
+\newcommand\maciek[1]{\color{brown}\textbf{\\ Maciek: #1}\color{black}}
 
@@
 
-Maciek
+\section{Introduction}
+
+Main technical contribution of this paper is $\Omega(k\ell)$ lower bound for any deterministic algorithm for Online Balanced Partition problem.
+Previous known lower bound were $\Omega(k)$ (cite Marcin-Chen).
+The gap between upper and lower bound is still substantial, as the best known algorithm is $O(k^2\ell^2)$ (cite Marcin-Chen).
+We contribute to the search for the optimal algorithm by presenting $O(k\ell)$-competitive algorithm for the special case where the perfect partition of the input graph is assumed.
+
+\maciek{Future directions of research: How well does the Perfect Partition Learner performs when applied to a single epoch in general online clustering problem (without the perfect partition?)?
+Epoch = C * l * (k-1) external requests OR per-component epochs.}
+
+\section{The Model}
+
+Formal introduction of the model.
+
+ We say that an online algorithm DET is component-respecting if it never reaches a configuration with a component split between clusters.
+  %Recall that if DET resides in a configuration with a component split between clusters, we issue intra-component requests until
+ If DET splits a component, we issue intra-component requests until DET joins it.
+ If it never joins a component, it is not competitive.
+
+\input{lower}
+
+\input{upper}
+
+\appendix
+
+\input{appendix}
 

### 2019-08-22 upper.tex
Proposition of a layout of the paper

--- 2019-08-22 mahmoud.tex
@@
-\section{Mahmoud}
+\section{An $O(kl)$-competitive Algorithm For the Scenario With the Perfect Partition}
+
+\subsection{On a Cost of a Single Repartition}
+
+\subsubsection{A Costly Repartition Example}
+
+A repartition cost is bounded by $O(kl)$ (cite marcin-chen).
+Now we show that the rebalance may be expensive (cubic) in terms of $k$, even for $l$ linear in $k$.
+
+This justifies taking a different approach to repartitioning (see Section PPLA).
+
+\maciek{Possibly simplify $\Omega(k^3)$? Or just present $\Omega(k^2)$?}
+
+\subsubsection{For Constant k, Repartition Cost is Constant?}
+
+\maciek{Interesting research problem: is a single repartition cost bounded by $O(poly(k))$? We have such examples for $k=3,4,5$. It gives us an $O(\ell)$-competitive algorithm for $k=3$}
+
+\begin{theorem}
+  The rebalancing cost for the greedy algorithm for $k=3$ is $O(1)$.
+  \maciek{Greedy = find the cheapest (closest to the current) feasible partition}
+  \label{rebalancing-cost}
+\end{theorem}
+
+\begin{proof} 
+ \maciek{Note: we consider a cost of cheapest feasible rebalancing after insertion of one edge.
+This in contrast to bounding the distance between any two reconfigurations (this theorem does not bound the latter).}
+
+  We distinguish among three configuration types of algorithm's clusters: $C_1, C_2, C_3$.
+In $C_1$ we have $3$ singleton components (this is also the initial configuration of any cluster).
+In $C_2$ we have one component of size $2$ and one component of size $1$.
+Finally, in $C_3$ we have one component of size $3$.
+
+  Consider a request $(u, v)$ and let $U, V$ be their clusters.
+  If $U=V$, then the request does not require a reconfiguration.
+  Now, we consider cases upon the type of clusters $U$ and $V$.
+  Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise the resulting component would have the size $4$, and this contradicts the existence of the perfect partition.
+  If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
+
+  Now, we focus on the case where both $U$ and $V$ are $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as the merged component would have size $4$.
+  If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
+  Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+  Note that in such case, a $C_1$-type cluster exists, as otherwise the input graph would not have a perfect partition.
+The cost of rebalancing is then at most $12$ \maciek{can show smaller constant than 12, but it does not matter that much}.
+\end{proof}
+
+\begin{theorem}
+  There exists a $O(\ell)$-competitive algorithm for $k=3$.
+\end{theorem}
+
+\begin{proof}
+  The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
+Each request issues at most one rebalancingfor the greedy algorithm, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
+\end{proof}
+
+\subsection{Perfect Partition Learner Algorithm}
 
@@
 
+      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial configuration}
+
 \begin{property} \label{prop:dist<OPT}

### 2019-08-23 lower.tex
Future directions

@@
 \begin{proof}
+
+  \maciek{Warning! This proof requires still major rewrite before submission!}
+  \maciek{The argument of connecting the length of sequence and the number of merged singletons should be stated as an auxiliary lemma.}
+  \maciek{Maybe we do not need to construct $I_2$ and then $I_2'$: we can construct $I_2'$ on-the-fly and bound its length. However, the subtracting argument (as it is written now) is quite simple.}
+  \maciek{ToDo: (one) figure containing: input $I_1$, cluster $D$, growing components $v^C_i$, and the ``searched area (by ALG)'' (in searching for $d_1, d_2$)}
+
+  \maciek{ToDo: check all off-by-one errors, especially this one: check if the existence of $b_1$ in one of the clusters was considered in all calculations}
 
@@
 
-  \maciek{ToDo: check all off-by-one errors}
-  \maciek{ToDo: check if the existence of $b_1$ was considered in all calculations}
-
  % During the runtime of DET, by $x(A)$ we denote the node in the configuration of DET that is co-located in one cluster with the component $a_1, \ldots, a_{k-1}$.

### 2019-08-23 lower.tex
Minor fixes

@@
   For the following $k-2$ nodes of the cluster $C$, we label them $v^C_2, v^C_3, \ldots$ in the order of appearance in the sequence $\langle x^{(i)}\rangle$.
-  %For a node $v^C_p = x^{(j)}$, we set $y^{(j)} = v^C_0$.
-  If some nodes of $C$ never appear in the sequence $\langle x^{(i)}\rangle$ %(e.g. DET might terminate prematurely or might reside in a configuration with components split)
-  , we number them arbitrarily, with the restriction that the nodes that do appear in the sequence have smaller number than the ones that do not \maciek{This should not happen at all for component-respecting DET, though}
+  % For a node $v^C_p = x^{(j)}$, we set $y^{(j)} = v^C_0$.
+
+  If some nodes of $C$ never appear in the sequence $\langle x^{(i)}\rangle$ (e.g. DET might terminate prematurely or might reside in a configuration with components split)
+  , we number them arbitrarily, with the restriction that the nodes that do appear in the sequence have smaller number than the ones that do not \maciek{This should not happen at all for component-respecting DET, though.
+Would be better to provide an additional lemma that shows that every node of each cluster appears in the sequence (outside of $v_0$'s, $a_1,\ldots,a_{k-1}, a_k, b_1$)}
 

### 2019-08-23 lower.tex
Minor fixes

@@
   If in this configuration at least one singleton component is present (other than the components of $(x^{(i)}, c(x^{(i)}))$), then serving this request is possible (i.e., the perfect partition exists), and hence a singleton node appears as $x^{(i+1)}$ and the sequence continues.
-  Otherwise, no perfect partition exists, as we have a component $\{a_1, \ldots, a_{k-1}\}$ of size $k-1$.
+  Otherwise, no perfect partition exists, as we have a component $\{a_1, \ldots, a_{k-1}\}$ of size $k-1$ and it can only be complemented by a singleton component.
 
+  \maciek{A candidate for auxiliary lemma: intuitions before the claim, then the formal formulation, and a rewrite of the argument from this paragraph}
   Hence, the length of a sequence $\langle x^{(i)}\rangle$ is closely related to the number of singleton components.

### 2019-08-23 upper.tex
Minor review
Future directions
Minor fixes

@@
 
-\maciek{Interesting research problem: is a single repartition cost bounded by $O(poly(k))$? We have such examples for $k=3,4,5$. It gives us an $O(\ell)$-competitive algorithm for $k=3$}
+\maciek{Interesting research problem: is a single repartition cost bounded by $O(poly(k))$?
+This would improve currently best-known $O(k^2\ell^2)$ to $O(k\cdot poly(k) \cdot \ell)$, i.e., would tighten the lower bound in terms of $\ell$.}
+  \maciek{We have such examples for $k=3,4,5$. It gives us an $O(\ell)$-competitive algorithm for $k=3$}
 
@@
 \begin{proof} 
- \maciek{Note: we consider a cost of cheapest feasible rebalancing after insertion of one edge.
+% \maciek{Note: we consider a cost of cheapest feasible rebalancing after insertion of one edge.
 This in contrast to bounding the distance between any two reconfigurations (this theorem does not bound the latter).}
@@
 \subsection{Perfect Partition Learner Algorithm}
+
+\maciek{$. \rightarrow \cdot$}
+\maciek{Font for OPT, ALG, DET, dist etc}
 

### 2019-08-23 lower.tex
Minor fixes
simpler LB proof, first draft

@@
 Therefore, this result is a lower bound for the general Online Balanced Partition problem, as well as a lower bound for Online Balanced Partition with perfect partition.
-The latter complements an upper bound $O(k \cdot \ell)$ for Online Balanced Partition with perfect partition that we present in Section TODO.
+The latter complements an upper bound $O(k \cdot \ell)$ for Online Balanced Partition with perfect partition that we present in Section \ref{sec:upperbound}.
 
@@
 \end{theorem}
+
+\begin{proof}
+A simpler proof: \\
+We construct an instance of the problem with $\ell$ clusters 
+$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
+Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
+Consider any online algorithm ALG.
+We construct the input sequence of requests for ALG as follows.
+First,
+we issue $k-2$ (internal) requests so that ALG form a component of $k-1$
+nodes on clusters $S_1$.
+Note that this does not cost ALG.
+Let $x_0$  be the only single nodes left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
+Next,
+we issue the request between $x_0$ and $y_0$.
+Since this is an external request,
+ALG joins the into one component and collocates them in some cluster other than $S_1$ and $S_2$.
+For this,
+ALG moves to a new configuration that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
+Next,
+we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(v)$.
+ALG must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
+consequently replaces $x_1$.
+We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
+ until there are only two single nodes left that  originate from the same cluster.
+I.e.,
+two single nodes $x^*, y^*,I(x^*) = I(y^*)$
+that constitute the last pair of such nodes.
+At this point there are at most $\ell+1$ single nodes left
+since there must be at least two nodes with the same initial cluster in any subset of $\ell+1$
+nodes.
+Given this sequence of requests,
+the optimal strategy is to swap $x_0$ and $y_0$ with $x^*$ and $y^*$ respectively.
+Hence,
+OPT pays for $4$ migrations and
+ALG pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
+We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k.\ell$ nodes,
+therefore $|X| \geq k.\ell - k - \ell \in \Omega(k.\ell)$.
+\end{proof}
 
@@
   must produce a sequence $\langle x^{(i)}\rangle$ of length $L = k\cdot \ell - (k-1) - 2 - (\ell-1) - 1$.
+  \maciek{ToDo: a sentence that explains each summand of $L$}
+  \maciek{Name of $L$ should indicate a connection to $I_2$, e.g., maybe just $|I_2|$, and then $L-\ell-1$ from last paragraph should be related to $I_2'$}
   Consider a request $(x^{(i)}, c(x^{(i)}))$.

### 2019-08-23 upper.tex
simpler LB proof, first draft

@@
 		initial partition $P_I$,
-		sequence of  edges $\sigma_1, \dots, \sigma_N$ covering all ground truth components
+		sequence of  edges $\sigma_1, \dots, \sigma_N$ 
 		\Ensure Final partition $P_F$ 
@@
       \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial configuration}
+
+\mahmoud{address the \textit{re-partition()}}
 
@@
 	The algorithm eventually reaches the final partition since it
-	 evaluates all possible $\ell$-way partitioning of components on each external request,
+	 evaluates all possible $\ell$-way partitions of components on each external request,
 	including the request that completes revealing of all ground truth components.

### 2019-08-23 lower.tex
minor

@@
 Given this sequence of requests,
-the optimal strategy is to swap $x_0$ and $y_0$ with $x^*$ and $y^*$ respectively.
+the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
+ swapping $x_0$ and $y_0$ with $x^*$ and $y^*$ respectively.
 Hence,
-OPT pays for $4$ migrations and
+OPT pays $4$ for node migrations and
 ALG pays at least one for each node in the sequence $X := x_0, x_1,\dots$.

### 2019-08-23 lower.tex
comment

@@
 \begin{proof}
-A simpler proof: \\
+\mahmoud{A simpler proof but still not fully rigorous. E.g., what if ALG collocates on $S_1$? Easy technicalities most likely.}
 We construct an instance of the problem with $\ell$ clusters 

### 2019-08-23 lower.tex
minor

@@
 Since this is an external request,
-ALG joins the into one component and collocates them in some cluster other than $S_1$ and $S_2$.
+ALG joins them into one component and collocates them in some cluster other than $S_1$.
 For this,
@@
 Next,
-we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(v)$.
+we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
 ALG must collocate $x_1$ and $C$ on some cluster other than $S_1$ and

### 2019-08-29 paper.tex
Proposition of a layout of the paper
Intro sketch
Recent observations

@@
 The gap between upper and lower bound is still substantial, as the best known algorithm is $O(k^2\ell^2)$ (cite Marcin-Chen).
-We contribute to the search for the optimal algorithm by presenting $O(k\ell)$-competitive algorithm for the special case where the perfect partition of the input graph is assumed.
+
+We contribute to the search for the optimal algorithm by presenting $O(k\ell)$-competitive algorithm for the special case where the perfect partition of the input graph is assumed and every algorithm must colocate every communicating pair.
+Note that our $\Omega(k\ell)$ lower bound uses a perfect partition input graph and colocates communicating pairs, and hence we provide a tight result for such scenario.
 
 \maciek{Future directions of research: How well does the Perfect Partition Learner performs when applied to a single epoch in general online clustering problem (without the perfect partition?)?
-Epoch = C * l * (k-1) external requests OR per-component epochs.}
+Epoch = C * l * (k-1) external requests.}
 
@@
 
-Formal introduction of the model.
+Formal introduction of the general model.
 
@@
 
+ \maciek{+ description of perfect partition + ground components model}
 \input{lower}

### 2019-08-29 upper.tex
Recent observations

@@
 
-This justifies taking a different approach to repartitioning (see Section PPLA).
+This justifies taking a different approach to repartitioning then bounding the repartition cost in $O(k^2\cdot \ell^2)$ algorithm (described in Section ``A simple upper bound'' in the DISC paper).
+As the cost of a specific repartition is $\Omega(k^2)$, then the cost of DET during a single phase is: for each of $O(k \cdot l)$ edges revealed, we already have the cost $O(k^2)$ (precisely, $O(min\{k^2, k\ell\})$, and the competitive ratio would be $O(k^3\cdot \ell)$.
+\maciek{This would be an improvement in terms of $\ell$ still, but no hopes for $O(k\ell)$ algorithm just by improving the rebalancing analysis}
 

### 2019-09-13 upper.tex
general online for k=3

@@
 
-\begin{theorem}
+\begin{lemma} \label{lemma:k=3}
   The rebalancing cost for the greedy algorithm for $k=3$ is $O(1)$.
@@
   \label{rebalancing-cost}
-\end{theorem}
+\end{lemma}
 
@@
   There exists a $O(\ell)$-competitive algorithm for $k=3$.
+\mahmoud{is this needed anymore?}
 \end{theorem}
@@
   The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
-Each request issues at most one rebalancingfor the greedy algorithm, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
+Each request issues at most one re-balancing for the greedy algorithm, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
+\end{proof}
+
+We leverage the simple algorithm $ON_{k=3}$ that seeks a perfect partition for $k=3$ towards an algorithm for the general online problem.
+Note that in the general model, an optimal partition is not a necessarily a perfect partition.
+We divide the sequence of requests into phases.
+A phase terminates when there is no perfect partition of components after arrival of some request $r$.
+We apply requests to $ON_{k=3}$ as before. 
+Once the current phase terminates,
+we reset all components by removing (forgetting) all revealed edges,
+ and then we proceed to the next phase.
+
+\begin{corollary} \label{cor:k=3}
+	There exists $O(\ell)$-competitive  algorithm for the general online problem when $k=3$. 
+\end{corollary}
+\begin{proof}
+	Consider any phase and assume it terminates upon arrival of some request $r$.
+	 By Lemma  \ref{lemma:k=3},
+	ON pays $O(\ell)$ during the phase.
+	Regardless of the of OPT's partition at the beginning of the phase,
+	as long as OPT stays in the same configuration,
+	 it incurs remote communication cost for at least one request $r'$ (possibly $r'=r$) by the end of the phase.
+	 Else,
+	 OPT incurs the cost of moving to new partitions.
+	 In any case,
+	 OPT must pay at least one during the phase.
 \end{proof}

### 2019-09-13 upper.tex
general online for k=3 (1)

@@
 
-We leverage the simple algorithm $ON_{k=3}$ that seeks a perfect partition for $k=3$ towards an algorithm for the general online problem.
-Note that in the general model, an optimal partition is not a necessarily a perfect partition.
+We leverage the simple algorithm $ON_{k=3}$ that seeks a perfect partition for $k=3$,
+towards an algorithm for the general online problem.
+Note that in the general model,
+an optimal partition is not a necessarily a perfect partition.
 We divide the sequence of requests into phases.
-A phase terminates when there is no perfect partition of components after arrival of some request $r$.
+A phase terminates when there is no perfect partition of components.
 We apply requests to $ON_{k=3}$ as before. 

### 2019-12-05 upper.tex
Ring communication pattern

@@
       \end{IEEEproof}
+
+      \section{$O(k)$-competitive algorithm for the ring}
+
+      \maciek{todo}
+
+      We know that the lower bound is $k$ for the ring communication pattern (cite Marcin-Chen).
+      The simple matching (assymptotically \maciek{maybe it is easy to have $k$ exactly?}) upper bound is to fix the reconfiguration to be the cyclic shift.
+We reconfigure upon encountering inter-cluster request.
+We use phases: upon discovering the edge on each cut we charge OPT 1 and reset counters.

### 2020-01-21 paper.tex
Quick notes on what to do

@@
 
+\section{ToDos}
+
+\begin{enumerate}
+    \item Make sure that we loose nothing while preparing the short paper (backup: in branch, in directory(svn-style branch)?)
+    \item Apply the ACM style and see how much space do we have
+    \item Decide on the subset of results to publish (I would minimize those: if a figure on LB fits, it would be vital to include it)
+    \item (and hence, draw a figure)
+    \item Decide on the order of results
+    \item Couple iterations on existing results (these are mostly fine already, though)
+    \item Make an ACM account for the submission and sum up all the steps needed for the submission
+\end{enumerate}
+
 \section{Introduction}

### 2020-01-22 paper/paper.tex
ACM style application

 
\documentclass[sigconf,screen=true]{acmart}

\usepackage[utf8]{inputenc}
\usepackage{xspace}
\usepackage{balance}

\begin{document}

some text

\end{document}

### 2020-01-22 paper/paper.tex
Related Work + Reference fix

@@
  
-\documentclass[sigconf,screen=true]{acmart}
+\documentclass[manuscript,screen=true]{acmart}
 
@@
 
-some text
+\title{Brief Announcement: An Improved Lower Bound for Dynamic Graph Partitioning}
+
+\author{Maciej Pacut}
+%\orcid{0000-0002-6379-1490}
+\affiliation{%
+  \department{Faculty of Computer Science}
+  \institution{University of Vienna}
+  \country{Austria}
+}
+
+\author{Mahmoud Parham} 
+\affiliation{%
+  \department{Faculty of Computer Science}
+  \institution{University of Vienna}
+  \country{Austria}
+}
+
+\author{Stefan Schmid} 
+\affiliation{%
+  \department{Faculty of Computer Science}
+  \institution{University of Vienna}
+  \country{Austria}
+}
+
+\copyrightyear{2020} 
+\acmYear{2020} 
+\setcopyright{acmlicensed}
+\acmConference{PODC '20}{July 29-August 2, 2020}{Toronto, Canada}
+
+\newcommand{\todo}[1]{\noindent\color{brown}{todo: #1}\color{black}}
+
+\begin{abstract}
+    Lorem
+\end{abstract}
+
+\maketitle
+
+\renewcommand{\shortauthors}{M.~Pacut, M.~Parham, S.~Schmid}
+
+\section{Related Work}
+
+In this paper we study an online balanced partitioning problem.
+The static offline version of~the~problem, i.e., a problem variant where
+migration is not allowed, where all requests are known in advance, and where
+the goal is to find an assignment of $n$ nodes to $\ell$~physical machines, each of~capacity $n/\ell$, is known as the
+\emph{$\ell$-balanced graph partitioning problem}. The problem is 
+NP-complete, and cannot even be approximated within any finite factor unless P
+= NP~\cite{AndRae06}.  The static
+variant where $\ell = 2$ corresponds to the minimum bisection problem, which
+is already NP-hard~\cite{GaJoSt76}, and 
+the currently best approximation ratio is $O(\log n)$~\cite{SarVaz95,ArKaKa99,FeKrNi00,FeiKra02,KraFei06,Raec08}.
+The inapproximability of the static variant for general values of $\ell$
+motivated research on the bicriteria variant, which can be seen as the offline
+counterpart of our capacity augmentation approach. Here, the~goal
+is~to~compute a~graph partitioning into $\ell$ components of~size at most~$k$ (where $k > n/\ell$) and the cost of the cut is compared to the optimal (non-augmented)
+solution where all components are of a~size at most $n/k$. The variant where
+$n \geq 2 \cdot k \cdot \ell$ was considered in
+\cite{LeMaTr90,SimTen97,EvNaRS00,EvNaRS99,KrNaSc09}. So far, the~best result~is~an~$O(\!\sqrt{\log n \cdot \log \ell})$-approximation algorithm~\cite{KrNaSc09}.
+
+Our dynamic model is related to online
+caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}, sometimes also referred to
+as online caching, where requests for data items (nodes) arrive over time and
+need to be served from a cache of finite capacity, and where the number of
+cache misses must be minimized. Classic problem variants usually boil down to
+finding a smart eviction strategy, such as Least Recently Used (LRU)~\cite{SleTar85}. In our
+setting, requests can be served remotely (i.e.,~without fetching the
+corresponding nodes to a single physical machine). In this light, our model is more
+reminiscent of caching models \emph{with
+bypassing}~\cite{EpImLN11,EpImLN15,Irani02}. As a~side result, we show that our problem is
+capable of emulating online caching.
+A major difference between  these problems is that in the caching problems, each request involves a~single element of the universe, while in our model \emph{both} endpoints of a communication request are subject to~optimization.
+
+\section{Todos}
+
+\todo{Make ORCID visible in ACM style}
+
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{references}
 

### 2020-01-22 paper/paper.tex
A lot of text to write intro, related work, practical motivations, etc + references

@@
 \begin{abstract}
-    Lorem
+    Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 \end{abstract}
@@
 
-Our dynamic model is related to online
+Our model is related to online
 caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}, sometimes also referred to
@@
 
+\section{Practical Motivation}
+
+Distributed cloud applications, including batch processing
+applications such as MapReduce, streaming applications such as Apache Flink or
+Apache Spark, and scale-out databases and key-value stores such as Cassandra,
+generate a~significant amount of network traffic and a~considerable fraction
+of their runtime is due to network activity~\cite{MogPop12}. For example,
+traces of jobs from a Facebook data center reveal that network transfers on
+average account for 33\% of the execution time~\cite{orchestra}. In such
+applications, it is desirable that frequently communicating virtual machines
+are \emph{collocated}, i.e., mapped to the same physical server: 
+communication across the network (i.e., inter-server communication) induces
+network load and latency. However, migrating virtual machines between servers
+also comes at a price: the state transfer is bandwidth intensive, and may even
+lead to short service interruptions. Therefore the goal is to design online
+algorithms that find a good trade-off between the inter-server communication
+cost and the migration cost.
+
+\vspace{1cm}
+
+We study virtual network embeddings in the scenario where virtual machines can be migrated during runtime to another physical machine.
+The possibility of migration allows reacting to unpredictable communication patterns.
+For example, if some distant nodes communicate often, it is vital to reduce their distance to save network bandwidth.
+The objective is to~minimize the total network bandwidth used for communication and for migration.
+
+We assume that the communication patterns are not known in advance to our algorithm.
+We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.
+In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.
+
+In the dynamic scenario, we assume that the physical substrate network is a~tree of height one.
+That is, every physical machine (leaf) is connected directly to the root (that has no hosting capabilities).
+A single physical machine hosts a fixed number of virtual machines.
+The model restricted to such networks becomes a variant of online graph clustering.
+That is, we are given a~set of~$n$ nodes (virtual machines) with time-varying pairwise
+communication patterns, which have to be partitioned into~$\ell$~physical machines, each of
+capacity $k=n/\ell$.
+
+Intuitively, we would like to minimize inter-machine
+interactions by mapping frequently communicating nodes to the same physical machine
+Since communication patterns change over time, the~nodes should be \emph{repartitioned}, in
+an online manner, by \emph{migrating} them between physical machines.
+The~objective is to minimize the weighted sum of inter-machine communication and repartitioning costs.
+The former is defined as the number of communication requests between nodes placed at distinct physical machines, and the latter as the number of migrations.
+
+The possibility to perform a migration uncovers algorithmic challenges:
+\begin{itemize}
+
+\item \emph{Serve remotely or migrate?} For a brief communication
+pattern, it may not be worthwhile to collocate the nodes: the migration cost might
+be too large in comparison to~communication costs.
+
+\item \emph{Where to migrate, and what?}
+If an algorithm decides to collocate nodes $x$ and~$y$, the~question becomes
+how. Should $x$ be migrated to the physical machine holding $y$, $y$ to the one holding
+$x$, or should both nodes be migrated to a new machine?
+
+\item \emph{Which nodes to evict?}
+The space of the desired destination physical machine may not be sufficient. In
+this case, the~algorithm needs to decide which nodes to ``evict'' (migrate to
+other machines), to free up space.
+
+\end{itemize}
+
+\section{Problem Definition}
+
+Formally, the online \emph{Balanced RePartitioning} problem (BRP) is defined as
+follows. There is a set of $n$ nodes, initially distributed arbitrarily
+across $\ell$~clusters, each of size~$k$. We call two nodes~$u,v\in V$
+\emph{collocated} if they are in the same cluster.
+
+An input to the problem is a sequence of communication requests $\sigma =
+(u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$, where pair $(u_t,v_t)$ means that
+the nodes $u_t,v_t$ exchange a fixed amount of data.  At any time~$t$, an online algorithm needs to serve the~communication
+request~$(u_t,v_t)$. Right before serving the request, the online algorithm
+can repartition the nodes into new clusters. We assume that
+a~communication request between two collocated nodes costs 0. The cost of a~communication request between two nodes located in different clusters is
+normalized to~1, and the cost of migrating a node from one cluster to another
+is~$\alpha \geq 1$, where $\alpha$ is a parameter (an~integer). For any
+algorithm $ALG$, we denote its total cost (consisting of communication plus
+migration costs) on sequence $\sigma$ by $ALG(\sigma)$.
+
 \section{Todos}

### 2020-01-22 paper/paper.tex
Start with including all the results

@@
 \usepackage{balance}
+\usepackage{amsmath,amssymb,amsfonts,mathtools,amsthm}
+\usepackage[noend]{algpseudocode}
+\usepackage{algorithm}
 
@@
 \acmConference{PODC '20}{July 29-August 2, 2020}{Toronto, Canada}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%&&
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%&&
+%  our macros start
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%&&
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%&&
+
+\newtheorem{claim}{Claim}
+\newtheorem{fact}{Fact}
+\newtheorem{rem}{Remark}
+\newtheorem{observation}{Observation}
+\newtheorem{property}{Property}
+
+\DeclarePairedDelimiter\pair{(}{)}
+\DeclarePairedDelimiter\set{\{}{\}}
+
+\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
+\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
+
+\newcommand\mahmoud[1]{\color{green}\textbf{\\ Mahmoud: #1}\\\color{black}}
+\newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}
+\newcommand\maciek[1]{\color{brown}\textbf{\\ Maciek: #1}\color{black}}
 
@@
 
+\section{Introduction}
+
+Main technical contribution of this paper is $\Omega(k\ell)$ lower bound for any deterministic algorithm for Online Balanced Partition problem.
+Previously the best known lower bound was $\Omega(k)$ \cite{repartition-disc}.
+The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\ell^2)$.
+
+We contribute to the search for the optimal algorithm by presenting $O(k\ell)$-competitive algorithm for the special case where the perfect partition of the input graph is assumed and every algorithm must colocate every communicating pair.
+Note that our $\Omega(k\ell)$ lower bound uses a perfect partition input graph and colocates communicating pairs, and hence we provide a tight result for such scenario.
+
 \section{Problem Definition}
@@
 
+\section{Lower bound $\Omega(k\cdot \ell)$}
+
+\maciek{I assume $\alpha = 1$ to simplify calculations, but everything should hold for general $\alpha$}
+
+In this section we consider the Online Balanced Partition problem.
+We show a lower bound of $\Omega(k \cdot \ell)$ that uses an input that has a perfect partition.
+Therefore, this result is a lower bound for the general Online Balanced Partition problem, as well as a lower bound for Online Balanced Partition with perfect partition.
+The latter complements an upper bound $O(k \cdot \ell)$ for Online Balanced Partition with perfect partition that we present in Section \ref{sec:upperbound}.
+
+\begin{theorem}
+  The competitive ratio of any deterministic algorithm for Online Balanced Partition is $\Omega(k\cdot \ell)$.
+\end{theorem}
+
+\begin{proof}
+\mahmoud{A simpler proof but still not fully rigorous. E.g., what if ALG collocates on $S_1$? Easy technicalities most likely.}
+We construct an instance of the problem with $\ell$ clusters 
+$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
+Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
+Consider any online algorithm ALG.
+We construct the input sequence of requests for ALG as follows.
+First,
+we issue $k-2$ (internal) requests so that ALG form a component of $k-1$
+nodes on clusters $S_1$.
+Note that this does not cost ALG.
+Let $x_0$  be the only single nodes left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
+Next,
+we issue the request between $x_0$ and $y_0$.
+Since this is an external request,
+ALG joins them into one component and collocates them in some cluster other than $S_1$.
+For this,
+ALG moves to a new configuration that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
+Next,
+we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
+ALG must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
+consequently replaces $x_1$.
+We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
+ until there are only two single nodes left that  originate from the same cluster.
+I.e.,
+two single nodes $x^*, y^*,I(x^*) = I(y^*)$
+that constitute the last pair of such nodes.
+At this point there are at most $\ell+1$ single nodes left
+since there must be at least two nodes with the same initial cluster in any subset of $\ell+1$
+nodes.
+Given this sequence of requests,
+the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
+ swapping $x_0$ and $y_0$ with $x^*$ and $y^*$ respectively.
+Hence,
+OPT pays $4$ for node migrations and
+ALG pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
+We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k.\ell$ nodes,
+therefore $|X| \geq k.\ell - k - \ell \in \Omega(k.\ell)$.
+\end{proof}
+
+\begin{observation}
+    The lower bound of $\Omega(\ell)$ holds even with $1/3-1/k$-augmentation, i.e., in scenarios where the capacity of each cluster of the algorithm is $(1+1/3)\cdot k - 1$, while the capacity of OPT's clusters is $k$.
+    To this end we use the construction from Observation~\ref{obs:lb_general_k}.
+    We observe that each cluster of the algorithm can fit exactly $3$ components of size $k/3$.
+
+    \maciek{This refers to simpler proof with $k=3$}
+  \end{observation}
+
+    \maciek{Note that this does not contradict the existence of $O((1+1/\epsilon)\cdot k \cdot \log(k))$ algorithm of Avin, Bienkowski et al, as the algorithm requires $2+\epsilon$-augmentation.}
+
+    \section{An $O(kl)$-competitive Algorithm For the Scenario With the Perfect Partition}
+
+    \subsection{On a Cost of a Single Repartition}
+
+    \subsubsection{A Costly Repartition Example}
+
+    A repartition cost is bounded by $O(kl)$ (cite marcin-chen).
+    Now we show that the rebalance may be expensive (cubic) in terms of $k$, even for $l$ linear in $k$.
+
+    This justifies taking a different approach to repartitioning then bounding the repartition cost in $O(k^2\cdot \ell^2)$ algorithm (described in Section ``A simple upper bound'' in the DISC paper).
+    As the cost of a specific repartition is $\Omega(k^2)$, then the cost of DET during a single phase is: for each of $O(k \cdot l)$ edges revealed, we already have the cost $O(k^2)$ (precisely, $O(min\{k^2, k\ell\})$, and the competitive ratio would be $O(k^3\cdot \ell)$.
+\maciek{This would be an improvement in terms of $\ell$ still, but no hopes for $O(k\ell)$ algorithm just by improving the rebalancing analysis}
+
+    \maciek{Possibly simplify $\Omega(k^3)$? Or just present $\Omega(k^2)$?}
+
+    \subsubsection{For Constant k, Repartition Cost is Constant?}
+
+    \maciek{Interesting research problem: is a single repartition cost bounded by $O(poly(k))$?
+This would improve currently best-known $O(k^2\ell^2)$ to $O(k\cdot poly(k) \cdot \ell)$, i.e., would tighten the lower bound in terms of $\ell$.}
+      \maciek{We have such examples for $k=3,4,5$. It gives us an $O(\ell)$-competitive algorithm for $k=3$}
+
+    \begin{lemma} \label{lemma:k=3}
+      The rebalancing cost for the greedy algorithm for $k=3$ is $O(1)$.
+      \maciek{Greedy = find the cheapest (closest to the current) feasible partition}
+      \label{rebalancing-cost}
+    \end{lemma}
+
+    \begin{proof} 
+    % \maciek{Note: we consider a cost of cheapest feasible rebalancing after insertion of one edge.
+This in contrast to bounding the distance between any two reconfigurations (this theorem does not bound the latter).}
+
+      We distinguish among three configuration types of algorithm's clusters: $C_1, C_2, C_3$.
+In $C_1$ we have $3$ singleton components (this is also the initial configuration of any cluster).
+In $C_2$ we have one component of size $2$ and one component of size $1$.
+Finally, in $C_3$ we have one component of size $3$.
+
+      Consider a request $(u, v)$ and let $U, V$ be their clusters.
+      If $U=V$, then the request does not require a reconfiguration.
+      Now, we consider cases upon the type of clusters $U$ and $V$.
+      Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise the resulting component would have the size $4$, and this contradicts the existence of the perfect partition.
+      If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
+
+      Now, we focus on the case where both $U$ and $V$ are $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as the merged component would have size $4$.
+      If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
+      Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+      Note that in such case, a $C_1$-type cluster exists, as otherwise the input graph would not have a perfect partition.
+The cost of rebalancing is then at most $12$ \maciek{can show smaller constant than 12, but it does not matter that much}.
+    \end{proof}
+
+    \begin{theorem}
+      There exists a $O(\ell)$-competitive algorithm for $k=3$.
+    \mahmoud{is this needed anymore?}
+    \end{theorem}
+
+    \begin{proof}
+      The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
+Each request issues at most one re-balancing for the greedy algorithm, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
+    \end{proof}
+
+    We leverage the simple algorithm $ON_{k=3}$ that seeks a perfect partition for $k=3$,
+    towards an algorithm for the general online problem.
+    Note that in the general model,
+    an optimal partition is not a necessarily a perfect partition.
+    We divide the sequence of requests into phases.
+    A phase terminates when there is no perfect partition of components.
+    We apply requests to $ON_{k=3}$ as before. 
+    Once the current phase terminates,
+    we reset all components by removing (forgetting) all revealed edges,
+     and then we proceed to the next phase.
+
+    \begin{corollary} \label{cor:k=3}
+        There exists $O(\ell)$-competitive  algorithm for the general online problem when $k=3$. 
+    \end{corollary}
+    \begin{proof}
+        Consider any phase and assume it terminates upon arrival of some request $r$.
+         By Lemma  \ref{lemma:k=3},
+        ON pays $O(\ell)$ during the phase.
+        Regardless of the of OPT's partition at the beginning of the phase,
+        as long as OPT stays in the same configuration,
+         it incurs remote communication cost for at least one request $r'$ (possibly $r'=r$) by the end of the phase.
+         Else,
+         OPT incurs the cost of moving to new partitions.
+         In any case,
+         OPT must pay at least one during the phase.
+    \end{proof}
+
+    \subsection{Perfect Partition Learner Algorithm}
+
+    \maciek{$. \rightarrow \cdot$}
+    \maciek{Font for OPT, ALG, DET, dist etc}
+
+    Let $P_I = I_1, \dots, I_{\ell}$ denote the initial partition with which both ON and OPT begin and
+    $P_F = F_1, \dots, F_{\ell}$ the final partition.
+    \begin{definition}	\label{def:dist}
+        The \emph{distance} of a partition $P = C_1, \dots, C_{\ell}$ is the number of nodes in $P$ that do not reside in their initial cluster.
+        That is,
+        $dist(P, P_I) = \sum_{j=1}^{\ell} | C_j \setminus I_j |$. 
+    \end{definition}
+
+    In other words,
+    at least $dist(P, P_I)/2$ node swaps are required in order to reach the partition $P$ from $P_I$.
+    Therefore,
+    $OPT \geq \Delta:= dist(P_F, P_I) $.
+    We devise an online algorithm that mimics OPT by minimizing the distance to the initial partition $P_I$ with each re-partitioning.
+    As a result,
+    ON never ends up in a partition that is further away from $P_I$ than $\Delta$.
+    This invariant ensures that ON does not pay too much while reaching $P_F$.
+
+    \begin{algorithm}
+        \renewcommand{\algorithmicrequire}{\textbf{Input:}}
+        \renewcommand{\algorithmicensure}{\textbf{Output:}}
+        \begin{algorithmic}[1]
+            \Require 
+            $k, \ell$,
+            initial partition $P_I$,
+            sequence of  edges $\sigma_1, \dots, \sigma_N$ 
+            \Ensure Final partition $P_F$ 
+            \State for each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$ \label{line:initcomponents}
+            \State on communication request $\sigma_t=\{u,v\}$:
+            \State Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
+            \If{$C_1 \neq C_2$}
+            \State unite the two components into a single component $C'$ and
+            $\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$ \label{line:mergecomponents}
+            \If{cluster$(C_1) \neq$ cluster$(C_2)$}		
+            \State \textit{re-partition($k, \ell, P_I, \mathcal{C}$)} \label{line:rebalance} 
+            \EndIf
+            \EndIf
+        \end{algorithmic}
+        \caption{Perfect Partition Learner}
+        \label{alg:ppl}
+          \end{algorithm}
+
+          \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial configuration}
+
+    \mahmoud{address the \textit{re-partition()}}
+
+    \begin{property} \label{prop:dist<OPT}
+        Let $P$ be any partition chosen by Algorithm \ref{alg:ppl} at Line $\ref{line:rebalance}$.
+        Then, $dist(P,P_I) \leq \Delta$.
+    \end{property}
+
+    \begin{lemma}	\label{lemma:rebalancecost}
+        The cost of re-partitioning at Line \ref{line:rebalance} is at most $2.OPT$.
+    \end{lemma}
+    \begin{proof}
+        Consider the re-partitioning that transforms $P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
+        Let $M \subset V$ denote the set of nodes that are migrated during this process.
+        We have $M = M^+ \cup M^- \cup M^\circ$,
+        where $M^-$ ($M^+$) denotes the subset of nodes that
+        enter (leave) their original cluster during the re-partitioning,
+        and $M^\circ$ denotes the set of remaining nodes in $M$.
+        Since $|M^- \cup M^\circ|$ nodes are not in their original cluster before the re-partitioning,
+        by Definition \ref{def:dist},
+        the distance before the re-partitioning is $dist(P_{t-1}) \geq | M^- \cup M^\circ |$.
+        Analogously,
+         the distance afterwards is $dist(P_{t}) \geq | M^+ \cup M^\circ |$.
+        Thus,
+        $|M| \leq dist(P_{t-1}) + dist(P_{t})$.
+        By Property \ref{prop:dist<OPT},
+        $dist(P_{t-1}) , dist(P_{t}) \leq \Delta \leq OPT$
+        and thereby we have	
+        $|M| \leq 2.OPT$.
+    \end{proof}
+
+    \begin{theorem}	\label{thm:upperbound}
+        Algorithm \ref{alg:ppl} reaches the final partition while being $(2.k.\ell)$-competitive.
+    \end{theorem}
+    \begin{proof}
+        The algorithm eventually reaches the final partition since it
+         evaluates all possible $\ell$-way partitions of components on each external request,
+        including the request that completes revealing of all ground truth components.
+        There are at most $(k-1).\ell < k.\ell $ calls to \emph{re-partition()} and by Lemma \ref{lemma:rebalancecost} each costs at most $2.OPT$.
+        The total cost is therefore at most $2.OPT.k.\ell$ which implies the competitive ratio.
+          \end{proof}
+
+          \section{$O(k)$-competitive algorithm for the ring}
+
+          \maciek{todo}
+
+          We know that the lower bound is $k$ for the ring communication pattern (cite Marcin-Chen).
+          The simple matching (assymptotically \maciek{maybe it is easy to have $k$ exactly?}) upper bound is to fix the reconfiguration to be the cyclic shift.
+We reconfigure upon encountering inter-cluster request.
+We use phases: upon discovering the edge on each cut we charge OPT 1 and reset counters.
+
 \section{Todos}
@@
 \todo{Make ORCID visible in ACM style}
+
+\todo{Check ACM package whitelist (especially these algorithm environments)}
+
+\todo{Make an ACM account for the submission}
+
+\todo{Sum up all the steps needed for the submission}
+
+\todo{ACM keywords are mandatory}
+
+\todo{CCS concepts are mandatory}
 

### 2020-01-22 paper/paper.tex
Distinguishing between core results and other
ACM keywords
ACM Whitelist of packages

@@
 \usepackage{algorithm}
-
-\begin{document}
 
@@
 
+\keywords{online algorithms, competitive analysis, etc}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%&&
@@
 
+\begin{document}
+
 \begin{abstract}
@@
 \renewcommand{\shortauthors}{M.~Pacut, M.~Parham, S.~Schmid}
+
+\section{Introduction}
+
+Main technical contribution of this paper is $\Omega(k\ell)$ lower bound for any deterministic algorithm for Online Balanced Partition problem.
+Previously the best known lower bound was $\Omega(k)$ \cite{repartition-disc}.
+The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\ell^2)$.
+
+We contribute to the search for the optimal algorithm by presenting $O(k\ell)$-competitive algorithm for the special case where the perfect partition of the input graph is assumed and every algorithm must colocate every communicating pair.
+Note that our $\Omega(k\ell)$ lower bound uses a perfect partition input graph and colocates communicating pairs, and hence we provide a tight result for such scenario.
 
@@
 
-In the dynamic scenario, we assume that the physical substrate network is a~tree of height one.
+We assume that the physical substrate network is a~tree of height one.
 That is, every physical machine (leaf) is connected directly to the root (that has no hosting capabilities).
@@
 
-\item \emph{Where to migrate, and what?}
+h\item \emph{Where to migrate, and what?}
 If an algorithm decides to collocate nodes $x$ and~$y$, the~question becomes
@@
 \end{itemize}
-
-\section{Introduction}
-
-Main technical contribution of this paper is $\Omega(k\ell)$ lower bound for any deterministic algorithm for Online Balanced Partition problem.
-Previously the best known lower bound was $\Omega(k)$ \cite{repartition-disc}.
-The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\ell^2)$.
-
-We contribute to the search for the optimal algorithm by presenting $O(k\ell)$-competitive algorithm for the special case where the perfect partition of the input graph is assumed and every algorithm must colocate every communicating pair.
-Note that our $\Omega(k\ell)$ lower bound uses a perfect partition input graph and colocates communicating pairs, and hence we provide a tight result for such scenario.
 
@@
 
-\section{Lower bound $\Omega(k\cdot \ell)$}
-
-\maciek{I assume $\alpha = 1$ to simplify calculations, but everything should hold for general $\alpha$}
+\section{Core results}
+\subsection{Lower bound $\Omega(k\cdot \ell)$}
 
@@
 therefore $|X| \geq k.\ell - k - \ell \in \Omega(k.\ell)$.
-\end{proof}
-
-\begin{observation}
-    The lower bound of $\Omega(\ell)$ holds even with $1/3-1/k$-augmentation, i.e., in scenarios where the capacity of each cluster of the algorithm is $(1+1/3)\cdot k - 1$, while the capacity of OPT's clusters is $k$.
-    To this end we use the construction from Observation~\ref{obs:lb_general_k}.
-    We observe that each cluster of the algorithm can fit exactly $3$ components of size $k/3$.
-
-    \maciek{This refers to simpler proof with $k=3$}
-  \end{observation}
-
-    \maciek{Note that this does not contradict the existence of $O((1+1/\epsilon)\cdot k \cdot \log(k))$ algorithm of Avin, Bienkowski et al, as the algorithm requires $2+\epsilon$-augmentation.}
-
-    \section{An $O(kl)$-competitive Algorithm For the Scenario With the Perfect Partition}
-
-    \subsection{On a Cost of a Single Repartition}
-
-    \subsubsection{A Costly Repartition Example}
-
-    A repartition cost is bounded by $O(kl)$ (cite marcin-chen).
-    Now we show that the rebalance may be expensive (cubic) in terms of $k$, even for $l$ linear in $k$.
-
-    This justifies taking a different approach to repartitioning then bounding the repartition cost in $O(k^2\cdot \ell^2)$ algorithm (described in Section ``A simple upper bound'' in the DISC paper).
-    As the cost of a specific repartition is $\Omega(k^2)$, then the cost of DET during a single phase is: for each of $O(k \cdot l)$ edges revealed, we already have the cost $O(k^2)$ (precisely, $O(min\{k^2, k\ell\})$, and the competitive ratio would be $O(k^3\cdot \ell)$.
-\maciek{This would be an improvement in terms of $\ell$ still, but no hopes for $O(k\ell)$ algorithm just by improving the rebalancing analysis}
-
-    \maciek{Possibly simplify $\Omega(k^3)$? Or just present $\Omega(k^2)$?}
-
-    \subsubsection{For Constant k, Repartition Cost is Constant?}
-
-    \maciek{Interesting research problem: is a single repartition cost bounded by $O(poly(k))$?
-This would improve currently best-known $O(k^2\ell^2)$ to $O(k\cdot poly(k) \cdot \ell)$, i.e., would tighten the lower bound in terms of $\ell$.}
-      \maciek{We have such examples for $k=3,4,5$. It gives us an $O(\ell)$-competitive algorithm for $k=3$}
-
-    \begin{lemma} \label{lemma:k=3}
-      The rebalancing cost for the greedy algorithm for $k=3$ is $O(1)$.
-      \maciek{Greedy = find the cheapest (closest to the current) feasible partition}
-      \label{rebalancing-cost}
-    \end{lemma}
-
-    \begin{proof} 
-    % \maciek{Note: we consider a cost of cheapest feasible rebalancing after insertion of one edge.
-This in contrast to bounding the distance between any two reconfigurations (this theorem does not bound the latter).}
-
-      We distinguish among three configuration types of algorithm's clusters: $C_1, C_2, C_3$.
-In $C_1$ we have $3$ singleton components (this is also the initial configuration of any cluster).
-In $C_2$ we have one component of size $2$ and one component of size $1$.
-Finally, in $C_3$ we have one component of size $3$.
-
-      Consider a request $(u, v)$ and let $U, V$ be their clusters.
-      If $U=V$, then the request does not require a reconfiguration.
-      Now, we consider cases upon the type of clusters $U$ and $V$.
-      Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise the resulting component would have the size $4$, and this contradicts the existence of the perfect partition.
-      If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
-
-      Now, we focus on the case where both $U$ and $V$ are $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as the merged component would have size $4$.
-      If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
-      Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-      Note that in such case, a $C_1$-type cluster exists, as otherwise the input graph would not have a perfect partition.
-The cost of rebalancing is then at most $12$ \maciek{can show smaller constant than 12, but it does not matter that much}.
-    \end{proof}
-
-    \begin{theorem}
-      There exists a $O(\ell)$-competitive algorithm for $k=3$.
-    \mahmoud{is this needed anymore?}
-    \end{theorem}
-
-    \begin{proof}
-      The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
-Each request issues at most one re-balancing for the greedy algorithm, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
-    \end{proof}
-
-    We leverage the simple algorithm $ON_{k=3}$ that seeks a perfect partition for $k=3$,
-    towards an algorithm for the general online problem.
-    Note that in the general model,
-    an optimal partition is not a necessarily a perfect partition.
-    We divide the sequence of requests into phases.
-    A phase terminates when there is no perfect partition of components.
-    We apply requests to $ON_{k=3}$ as before. 
-    Once the current phase terminates,
-    we reset all components by removing (forgetting) all revealed edges,
-     and then we proceed to the next phase.
-
-    \begin{corollary} \label{cor:k=3}
-        There exists $O(\ell)$-competitive  algorithm for the general online problem when $k=3$. 
-    \end{corollary}
-    \begin{proof}
-        Consider any phase and assume it terminates upon arrival of some request $r$.
-         By Lemma  \ref{lemma:k=3},
-        ON pays $O(\ell)$ during the phase.
-        Regardless of the of OPT's partition at the beginning of the phase,
-        as long as OPT stays in the same configuration,
-         it incurs remote communication cost for at least one request $r'$ (possibly $r'=r$) by the end of the phase.
-         Else,
-         OPT incurs the cost of moving to new partitions.
-         In any case,
-         OPT must pay at least one during the phase.
 \end{proof}
@@
 
-          \section{$O(k)$-competitive algorithm for the ring}
+\section{Additional results (most likely should be deleted)}
+
+\begin{observation}
+    The lower bound of $\Omega(\ell)$ holds even with $1/3-1/k$-augmentation, i.e., in scenarios where the capacity of each cluster of the algorithm is $(1+1/3)\cdot k - 1$, while the capacity of OPT's clusters is $k$.
+    To this end we use the construction from Observation~\ref{obs:lb_general_k}.
+    We observe that each cluster of the algorithm can fit exactly $3$ components of size $k/3$.
+
+    \maciek{This refers to simpler proof with $k=3$}
+  \end{observation}
+
+    \maciek{Note that this does not contradict the existence of $O((1+1/\epsilon)\cdot k \cdot \log(k))$ algorithm of Avin, Bienkowski et al, as the algorithm requires $2+\epsilon$-augmentation.}
+
+    \subsection{On a Cost of a Single Repartition}
+
+    \subsubsection{A Costly Repartition Example}
+
+    A repartition cost is bounded by $O(kl)$ (cite marcin-chen).
+    Now we show that the rebalance may be expensive (cubic) in terms of $k$, even for $l$ linear in $k$.
+
+    This justifies taking a different approach to repartitioning then bounding the repartition cost in $O(k^2\cdot \ell^2)$ algorithm (described in Section ``A simple upper bound'' in the DISC paper).
+    As the cost of a specific repartition is $\Omega(k^2)$, then the cost of DET during a single phase is: for each of $O(k \cdot l)$ edges revealed, we already have the cost $O(k^2)$ (precisely, $O(min\{k^2, k\ell\})$, and the competitive ratio would be $O(k^3\cdot \ell)$.
+\maciek{This would be an improvement in terms of $\ell$ still, but no hopes for $O(k\ell)$ algorithm just by improving the rebalancing analysis}
+
+    \maciek{Possibly simplify $\Omega(k^3)$? Or just present $\Omega(k^2)$?}
+
+    \subsubsection{For Constant k, Repartition Cost is Constant?}
+
+    \maciek{Interesting research problem: is a single repartition cost bounded by $O(poly(k))$?
+This would improve currently best-known $O(k^2\ell^2)$ to $O(k\cdot poly(k) \cdot \ell)$, i.e., would tighten the lower bound in terms of $\ell$.}
+      \maciek{We have such examples for $k=3,4,5$. It gives us an $O(\ell)$-competitive algorithm for $k=3$}
+
+    \begin{lemma} \label{lemma:k=3}
+      The rebalancing cost for the greedy algorithm for $k=3$ is $O(1)$.
+      \maciek{Greedy = find the cheapest (closest to the current) feasible partition}
+      \label{rebalancing-cost}
+    \end{lemma}
+
+    \begin{proof} 
+    % \maciek{Note: we consider a cost of cheapest feasible rebalancing after insertion of one edge.
+This in contrast to bounding the distance between any two reconfigurations (this theorem does not bound the latter).}
+
+      We distinguish among three configuration types of algorithm's clusters: $C_1, C_2, C_3$.
+In $C_1$ we have $3$ singleton components (this is also the initial configuration of any cluster).
+In $C_2$ we have one component of size $2$ and one component of size $1$.
+Finally, in $C_3$ we have one component of size $3$.
+
+      Consider a request $(u, v)$ and let $U, V$ be their clusters.
+      If $U=V$, then the request does not require a reconfiguration.
+      Now, we consider cases upon the type of clusters $U$ and $V$.
+      Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise the resulting component would have the size $4$, and this contradicts the existence of the perfect partition.
+      If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
+
+      Now, we focus on the case where both $U$ and $V$ are $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as the merged component would have size $4$.
+      If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
+      Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+      Note that in such case, a $C_1$-type cluster exists, as otherwise the input graph would not have a perfect partition.
+The cost of rebalancing is then at most $12$ \maciek{can show smaller constant than 12, but it does not matter that much}.
+    \end{proof}
+
+    \begin{theorem}
+      There exists a $O(\ell)$-competitive algorithm for $k=3$.
+    \mahmoud{is this needed anymore?}
+    \end{theorem}
+
+    \begin{proof}
+      The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
+Each request issues at most one re-balancing for the greedy algorithm, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
+    \end{proof}
+
+    We leverage the simple algorithm $ON_{k=3}$ that seeks a perfect partition for $k=3$,
+    towards an algorithm for the general online problem.
+    Note that in the general model,
+    an optimal partition is not a necessarily a perfect partition.
+    We divide the sequence of requests into phases.
+    A phase terminates when there is no perfect partition of components.
+    We apply requests to $ON_{k=3}$ as before. 
+    Once the current phase terminates,
+    we reset all components by removing (forgetting) all revealed edges,
+     and then we proceed to the next phase.
+
+    \begin{corollary} \label{cor:k=3}
+        There exists $O(\ell)$-competitive  algorithm for the general online problem when $k=3$. 
+    \end{corollary}
+    \begin{proof}
+        Consider any phase and assume it terminates upon arrival of some request $r$.
+         By Lemma  \ref{lemma:k=3},
+        ON pays $O(\ell)$ during the phase.
+        Regardless of the of OPT's partition at the beginning of the phase,
+        as long as OPT stays in the same configuration,
+         it incurs remote communication cost for at least one request $r'$ (possibly $r'=r$) by the end of the phase.
+         Else,
+         OPT incurs the cost of moving to new partitions.
+         In any case,
+         OPT must pay at least one during the phase.
+    \end{proof}
+
+          \subsection{$O(k)$-competitive algorithm for the ring}
 
@@
 
-\todo{Check ACM package whitelist (especially these algorithm environments)}
+\todo{Include e-mails}
+
+\todo{Check ACM package whitelist}
+
+\todo{Replace package algorithmpseudocode, it is not whitelisted https://www.acm.org/publications/taps/whitelist-of-latex-packages}
 

### 2020-01-22 paper/paper.tex
Minor rewrite of intro

@@
 
-Main technical contribution of this paper is $\Omega(k\ell)$ lower bound for any deterministic algorithm for Online Balanced Partition problem.
-Previously the best known lower bound was $\Omega(k)$ \cite{repartition-disc}.
-The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\ell^2)$.
+Main technical contribution of this paper is $\Omega(k\ell)$ lower bound for the competitive ratio of any deterministic algorithm for Online Balanced Partition problem.
+The best known lower bound so far was $\Omega(k)$ \cite{repartition-disc}.
 
-We contribute to the search for the optimal algorithm by presenting $O(k\ell)$-competitive algorithm for the special case where the perfect partition of the input graph is assumed and every algorithm must colocate every communicating pair.
-Note that our $\Omega(k\ell)$ lower bound uses a perfect partition input graph and colocates communicating pairs, and hence we provide a tight result for such scenario.
+Additionally, we contribute to search for the optimal algorithm by presenting $O(k\ell)$-competitive algorithm for the special case, where the perfect partition of the input graph is assumed, and every algorithm must colocate every communicating pair.
+Note that our $\Omega(k\ell)$ lower bound uses a perfect partition input graph and colocates communicating pairs, and hence for this scenario, our results are tight.
+\maciek{Also, it is easy to construct the adversary that forces collocation of requested nodes}
 
@@
 
-In this paper we study an online balanced partitioning problem.
+In this paper we study the Online Balanced Partitioning problem (formally presented in Section~\ref{sec:problem-definition}).
 The static offline version of~the~problem, i.e., a problem variant where
@@
 
-h\item \emph{Where to migrate, and what?}
+\item \emph{Where to migrate, and what?}
 If an algorithm decides to collocate nodes $x$ and~$y$, the~question becomes
@@
 \section{Problem Definition}
+\label{sec:problem-definition}
 
@@
 
-\section{Core results}
-\subsection{Lower bound $\Omega(k\cdot \ell)$}
+\section{Lower bound $\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
 
@@
 
-\subsection{Perfect Partition Learner Algorithm}
+\section{Perfect Partition Learner Algorithm}
 
@@
       \end{proof}
+
+\section{Corollary and Future Directions}
+
+The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\ell^2)$.
+
+\section{Todos}
+
+\todo{Make ORCID visible in ACM style}
+
+\todo{Include e-mails}
+
+\todo{Check ACM package whitelist}
+
+\todo{Replace package algorithmpseudocode, it is not whitelisted https://www.acm.org/publications/taps/whitelist-of-latex-packages}
+
+\todo{Make an ACM account for the submission}
+
+\todo{Sum up all the steps needed for the submission}
+
+\todo{ACM keywords are mandatory}
+
+\todo{CCS concepts are mandatory}
+
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{references}
+
+\begin{appendix}
 
@@
 
-\section{Todos}
-
-\todo{Make ORCID visible in ACM style}
-
-\todo{Include e-mails}
-
-\todo{Check ACM package whitelist}
-
-\todo{Replace package algorithmpseudocode, it is not whitelisted https://www.acm.org/publications/taps/whitelist-of-latex-packages}
-
-\todo{Make an ACM account for the submission}
-
-\todo{Sum up all the steps needed for the submission}
-
-\todo{ACM keywords are mandatory}
-
-\todo{CCS concepts are mandatory}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{references}
+\end{appendix}
 

### 2020-01-22 paper/paper.tex
Minor rewrite of intro
Comments
More comments

@@
 
-\maciek{$. \rightarrow \cdot$}
-\maciek{Font for OPT, ALG, DET, dist etc}
-
 Let $P_I = I_1, \dots, I_{\ell}$ denote the initial partition with which both ON and OPT begin and
@@
 \end{definition}
+\maciek{We should figure out a better name than "distance".
+We usually apply two arguments to distance.
+Examples of names that can be applied to only one configuration as a parameter: radius, span, stretch, spread, width.
+Also: difference, disparity, distinction, alternation}
 
@@
 
+\todo{$. \rightarrow \cdot$}
+
+\todo{Font for OPT, ALG, DET, dist etc}
+
 \bibliographystyle{ACM-Reference-Format}
@@
     \mahmoud{is this needed anymore?}
+    \maciek{It does not overlap with any other result from this paper.}
     \end{theorem}

### 2020-01-22 paper/paper.tex
Deleting minor results and emphasizing 3 major ones

@@
 
-\section{Corollary and Future Directions}
-
-The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\ell^2)$.
-
-\section{Todos}
-
-\todo{Make ORCID visible in ACM style}
-
-\todo{Include e-mails}
-
-\todo{Check ACM package whitelist}
-
-\todo{Replace package algorithmpseudocode, it is not whitelisted https://www.acm.org/publications/taps/whitelist-of-latex-packages}
-
-\todo{Make an ACM account for the submission}
-
-\todo{Sum up all the steps needed for the submission}
-
-\todo{ACM keywords are mandatory}
-
-\todo{CCS concepts are mandatory}
-
-\todo{$. \rightarrow \cdot$}
-
-\todo{Font for OPT, ALG, DET, dist etc}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{references}
-
-\begin{appendix}
-
-\section{Additional results (most likely should be deleted)}
-
-\begin{observation}
-    The lower bound of $\Omega(\ell)$ holds even with $1/3-1/k$-augmentation, i.e., in scenarios where the capacity of each cluster of the algorithm is $(1+1/3)\cdot k - 1$, while the capacity of OPT's clusters is $k$.
-    To this end we use the construction from Observation~\ref{obs:lb_general_k}.
-    We observe that each cluster of the algorithm can fit exactly $3$ components of size $k/3$.
-
-    \maciek{This refers to simpler proof with $k=3$}
-  \end{observation}
-
-    \maciek{Note that this does not contradict the existence of $O((1+1/\epsilon)\cdot k \cdot \log(k))$ algorithm of Avin, Bienkowski et al, as the algorithm requires $2+\epsilon$-augmentation.}
-
-    \subsection{On a Cost of a Single Repartition}
-
-    \subsubsection{A Costly Repartition Example}
-
-    A repartition cost is bounded by $O(kl)$ (cite marcin-chen).
-    Now we show that the rebalance may be expensive (cubic) in terms of $k$, even for $l$ linear in $k$.
-
-    This justifies taking a different approach to repartitioning then bounding the repartition cost in $O(k^2\cdot \ell^2)$ algorithm (described in Section ``A simple upper bound'' in the DISC paper).
-    As the cost of a specific repartition is $\Omega(k^2)$, then the cost of DET during a single phase is: for each of $O(k \cdot l)$ edges revealed, we already have the cost $O(k^2)$ (precisely, $O(min\{k^2, k\ell\})$, and the competitive ratio would be $O(k^3\cdot \ell)$.
-\maciek{This would be an improvement in terms of $\ell$ still, but no hopes for $O(k\ell)$ algorithm just by improving the rebalancing analysis}
-
-    \maciek{Possibly simplify $\Omega(k^3)$? Or just present $\Omega(k^2)$?}
-
-    \subsubsection{For Constant k, Repartition Cost is Constant?}
-
-    \maciek{Interesting research problem: is a single repartition cost bounded by $O(poly(k))$?
-This would improve currently best-known $O(k^2\ell^2)$ to $O(k\cdot poly(k) \cdot \ell)$, i.e., would tighten the lower bound in terms of $\ell$.}
-      \maciek{We have such examples for $k=3,4,5$. It gives us an $O(\ell)$-competitive algorithm for $k=3$}
+\section{$O(\ell)$-competitive algorithm for $k=3$}
 
@@
   There exists a $O(\ell)$-competitive algorithm for $k=3$.
-    \mahmoud{is this needed anymore?}
-    \maciek{It does not overlap with any other result from this paper.}
 \end{theorem}
@@
 
-          \subsection{$O(k)$-competitive algorithm for the ring}
+\section{Corollary and Future Directions}
 
-          \maciek{todo}
+The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\ell^2)$.
 
-          We know that the lower bound is $k$ for the ring communication pattern (cite Marcin-Chen).
-          The simple matching (assymptotically \maciek{maybe it is easy to have $k$ exactly?}) upper bound is to fix the reconfiguration to be the cyclic shift.
-We reconfigure upon encountering inter-cluster request.
-We use phases: upon discovering the edge on each cut we charge OPT 1 and reset counters.
+\section{Todos}
+
+\todo{Make ORCID visible in ACM style}
+
+\todo{Include e-mails}
+
+\todo{Check ACM package whitelist}
+
+\todo{Replace package algorithmpseudocode, it is not whitelisted https://www.acm.org/publications/taps/whitelist-of-latex-packages}
+
+\todo{Make an ACM account for the submission}
+
+\todo{Sum up all the steps needed for the submission}
+
+\todo{ACM keywords are mandatory}
+
+\todo{CCS concepts are mandatory}
+
+\todo{$. \rightarrow \cdot$}
+
+\todo{Font for OPT, ALG, DET, dist etc}
+
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{references}
+
+\begin{appendix}
 

### 2020-01-22 paper/paper.tex
Summing up ommited results

@@
 
+\section{Ommited results}
+
+\begin{enumerate}
+  \item The lower bound of $\Omega(\ell)$ for the scenario with at most $(1/3-1/k)$-augmentation
+  \item Costly cascade example (it is easy to present $k^2$ and $k\cdot \sqrt{k}$ (with different approach), and then we can combine these to obtain $k^3$)
+  \item $k$-competitive algorithm for the ring communication pattern
+  \item $O(\ell)$-competitive algorithm for $k=4$ and $k=5$
+\end{enumerate}
+
 \end{appendix}

### 2020-01-27 paper/paper.tex
small changes for k=3

--- 2020-01-22 paper/paper.tex
@@
 \begin{proof}
-\mahmoud{A simpler proof but still not fully rigorous. E.g., what if ALG collocates on $S_1$? Easy technicalities most likely.}
 We construct an instance of the problem with $\ell$ clusters 
@@
 
+   Let \emph{Arb-partition} be the algorithm that moves to an arbitrary perfect partition
+on each inter-cluster request.
 \begin{theorem}
-  There exists a $O(\ell)$-competitive algorithm for $k=3$.
+	Algorithm \emph{Arb-partition} is $O(\ell)$-competitive when $k=3$.
 \end{theorem}
@@
 	The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
-Each request issues at most one re-balancing for the greedy algorithm, and by Theorem~\ref{rebalancing-cost}, the total cost of any algorithm is $O(k\cdot \ell)$.
+	Each request triggers at most one re-balancing in \emph{Arb-partition}
+	which costs $O(1)$ by Theorem~\ref{rebalancing-cost}.
+	Then,
+	the total cost of any algorithm is $O(\ell)$.
 \end{proof}
 
-We leverage the simple algorithm $ON_{k=3}$ that seeks a perfect partition for $k=3$,
-towards an algorithm for the general online problem.
+We leverage the simple algorithm \emph{Arb-partition} for $k=3$
+towards an algorithm for the general problem
+where the communication graph is not restricted.
 Note that in the general model,
@@
 A phase terminates when there is no perfect partition of components.
-We apply requests to $ON_{k=3}$ as before. 
+We apply requests to \emph{Arb-partition} as before. 
 Once the current phase terminates,
@@
 
-\begin{corollary} \label{cor:k=3}
+\begin{theorem} \label{cor:k=3}
 	There exists $O(\ell)$-competitive algorithm for the general online problem when $k=3$. 
-\end{corollary}
+\end{theorem}
 \begin{proof}
@@
 	By Lemma  \ref{lemma:k=3},
-    ON pays $O(\ell)$ during the phase.
-    Regardless of the of OPT's partition at the beginning of the phase,
-    as long as OPT stays in the same configuration,
+	\emph{Arb-partition} pays $O(\ell)$ during the phase.
+	If OPT stays in the same configuration throughout the phase,
+	regardless of its partition at the beginning of the phase,        
 	it incurs remote communication cost for at least one request $r'$ (possibly $r'=r$) by the end of the phase.
@@
 	In any case,
-     OPT must pay at least one during the phase.
+	OPT pays at least one during the phase and 
+	the competitive ratio is $O(\ell)$.
 \end{proof}
@@
 \begin{appendix}
-
 \end{appendix}

### 2020-01-27 paper/paper.tex
undo deleting appendix

@@
 \begin{appendix}
+	-
+	-\section{Ommited results}
+	-
+	-\begin{enumerate}
+		-  \item The lower bound of $\Omega(\ell)$ for the scenario with at most $(1/3-1/k)$-augmentation
+		-  \item Costly cascade example (it is easy to present $k^2$ and $k\cdot \sqrt{k}$ (with different approach), and then we can combine these to obtain $k^3$)
+		-  \item $k$-competitive algorithm for the ring communication pattern
+		-  \item $O(\ell)$-competitive algorithm for $k=4$ and $k=5$
+		-\end{enumerate}
+	-
 \end{appendix}
-
 \end{document}

### 2020-01-27 paper/paper.tex
undo deleting appendix (1)

--- 2020-01-27 paper/paper.tex
@@
 \begin{appendix}
+	\section{Ommited results}
+	\begin{enumerate}
+		 \item The lower bound of $\Omega(\ell)$ for the scenario with at most $(1/3-1/k)$-augmentation
+		 \item Costly cascade example (it is easy to present $k^2$ and $k\cdot \sqrt{k}$ (with different approach), and then we can combine these to obtain $k^3$)
+		 \item $k$-competitive algorithm for the ring communication pattern
+		 \item $O(\ell)$-competitive algorithm for $k=4$ and $k=5$
+	\end{enumerate}
+
 \end{appendix}

### 2020-01-28 paper/paper.tex
PPL and its proof

@@
 \usepackage{algorithm}
+
+\mathchardef\mhyphen="2D
+\newcommand{\OPT}{\mathit{OPT}}
 
@@
 
-Formally, the online \emph{Balanced RePartitioning} problem (BRP) is defined as
+Formally, the online \emph{Balanced Re-Partitioning} problem (BRP) is defined as
 follows. There is a set of $n$ nodes, initially distributed arbitrarily
@@
 We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k.\ell$ nodes,
-therefore $|X| \geq k.\ell - k - \ell \in \Omega(k.\ell)$.
+therefore $|X| \geq k.\ell - k - \ell \in \Omega(k\cdot\ell)$.
 \end{proof}
@@
 
-Let $P_I = I_1, \dots, I_{\ell}$ denote the initial partition with which both ON and OPT begin and
-$P_F = F_1, \dots, F_{\ell}$ the final partition.
-\begin{definition}	\label{def:dist}
-    The \emph{distance} of a partition $P = C_1, \dots, C_{\ell}$ is the number of nodes in $P$ that do not reside in their initial cluster.
+We assume OPT begins with the initial partition
+$P_I = I_1, \dots, I_{\ell}$ and moves to the final partition
+$P_F = F_1, \dots, F_{\ell}$.
+ The \emph{distance} of a partition $P = C_1, \dots, C_{\ell}$ from the initial partition
+ is the number of nodes in $P$ that do not reside in their initial cluster.
     That is,
-    $dist(P, P_I) = \sum_{j=1}^{\ell} | C_j \setminus I_j |$. 
-\end{definition}
+    $\mathit{dist}(P, P_I) = \sum_{j=1}^{\ell} | C_j \setminus I_j |$. 
+In other words,
+at least $\mathit{dist}(P, P_I)/2$ node swaps are required in order to reach the partition $P$ from $P_I$.
+Therefore,
+$\OPT \geq \Delta:= dist(P_F, P_I) $.
+The online algorithm PPL mimics OPT by minimizing the distance to the initial partition $P_I$ with each re-partitioning.
+As a result,
+ON never ends up with a partition that is more than $\Delta$ away from $P_I$.
+This invariant ensures that ON does not pay too much while reaching $P_F$.
+
 \maciek{We should figure out a better name than "distance".
@@
 Also: difference, disparity, distinction, alternation}
-
-In other words,
-at least $dist(P, P_I)/2$ node swaps are required in order to reach the partition $P$ from $P_I$.
-Therefore,
-$OPT \geq \Delta:= dist(P_F, P_I) $.
-We devise an online algorithm that mimics OPT by minimizing the distance to the initial partition $P_I$ with each re-partitioning.
-As a result,
-ON never ends up in a partition that is further away from $P_I$ than $\Delta$.
-This invariant ensures that ON does not pay too much while reaching $P_F$.
 
@@
         initial partition $P_I$,
-        sequence of  edges $\sigma_1, \dots, \sigma_N$ 
-        \Ensure Final partition $P_F$ 
-        \State for each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$ \label{line:initcomponents}
-        \State on communication request $\sigma_t=\{u,v\}$:
+        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
+        \Ensure A final partition $P_F$ 
+        \State For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$
+        \State{$P_0 := P_I$}
+         \label{line:initcomponents}
+        \For {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
         \State Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
         \If{$C_1 \neq C_2$}
-        \State unite the two components into a single component $C'$ and
+        \State Unite the two components into a single component $C'$ and
         $\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$ \label{line:mergecomponents}
-        \If{cluster$(C_1) \neq$ cluster$(C_2)$}		
-        \State \textit{re-partition($k, \ell, P_I, \mathcal{C}$)} \label{line:rebalance} 
+%        \If{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$}
+        \If{$C_1$ and $C_2$ are not in the same cluster w.r.t.~$P_{t-1}$}
+        \State $P_{t} = \mathit{re\mhyphen partition}(P_{t-1}, P_I, \mathcal{C}$) \label{line:rebalance} 
+        \Comment{move to $P$ closest to $P_I$}
         \EndIf
         \EndIf
+        \EndFor
     \end{algorithmic}
-    \caption{Perfect Partition Learner}
+    \caption{Perfect Partition Learner (PPL)}
     \label{alg:ppl}
@@
 
-\mahmoud{address the \textit{re-partition()}}
-
+      We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current partition $P$ with a (perfect) partition closest to $P_I$.
+Hence it never moves to a partition beyond distance $\Delta$.      
 \begin{property} \label{prop:dist<OPT}
     Let $P$ be any partition chosen by Algorithm \ref{alg:ppl} at Line $\ref{line:rebalance}$.
-    Then, $dist(P,P_I) \leq \Delta$.
+    Then, $\mathit{dist}(P,P_I) \leq \Delta$.
 \end{property}
@@
 \begin{lemma}	\label{lemma:rebalancecost}
-    The cost of re-partitioning at Line \ref{line:rebalance} is at most $2.OPT$.
+    The cost of re-partitioning at Line \ref{line:rebalance} is at most $2\cdot\OPT$.
 \end{lemma}
@@
     Consider the re-partitioning that transforms $P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
-    Let $M \subset V$ denote the set of nodes that are migrated during this process.
-    We have $M = M^+ \cup M^- \cup M^\circ$,
-    where $M^-$ ($M^+$) denotes the subset of nodes that
-    enter (leave) their original cluster during the re-partitioning,
-    and $M^\circ$ denotes the set of remaining nodes in $M$.
-    Since $|M^- \cup M^\circ|$ nodes are not in their original cluster before the re-partitioning,
-    by Definition \ref{def:dist},
-    the distance before the re-partitioning is $dist(P_{t-1}) \geq | M^- \cup M^\circ |$.
+    Let $M \subset V$ denote the set of nodes that migrate during this process.
+	Let $M^-$ and $M^+$ denote the subset of nodes that (respectively)
+    enter or leave their original cluster during the re-partitioning.    
+    Then,
+    $M = M^+ \cup M^-$.
+    Since $|M^-|$ nodes are not in their original cluster before the re-partitioning (i.e., in $P_{t-1}$),
+    the distance before the re-partitioning is $\mathit{dist}(P_{t-1},P_I) \geq | M^-|$.
     Analogously,
-     the distance afterwards is $dist(P_{t}) \geq | M^+ \cup M^\circ |$.
+     the distance afterwards is $\mathit{dist}(P_{t},P_I) \geq | M^+|$.
     Thus,
-    $|M| \leq dist(P_{t-1}) + dist(P_{t})$.
+    $|M| \leq \mathit{dist}(P_{t-1},P_I) + \mathit{dist}(P_{t},P_I)$.
     By Property \ref{prop:dist<OPT},
-    $dist(P_{t-1}) , dist(P_{t}) \leq \Delta \leq OPT$
+    $\mathit{dist}(P_{t-1},P_I) , \mathit{dist}(P_{t},P_I) \leq \Delta \leq \OPT$
     and thereby we have	
-    $|M| \leq 2.OPT$.
+    $|M| \leq 2\cdot\OPT$.
 \end{proof}
@@
 \begin{theorem}	\label{thm:upperbound}
-    Algorithm \ref{alg:ppl} reaches the final partition while being $(2.k.\ell)$-competitive.
+    Algorithm \ref{alg:ppl} reaches the final partition $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
 \end{theorem}
 \begin{proof}
-    The algorithm eventually reaches the final partition since it
-     evaluates all possible $\ell$-way partitions of components on each external request,
-    including the request that completes revealing of all ground truth components.
-    There are at most $(k-1).\ell < k.\ell $ calls to \emph{re-partition()} and by Lemma \ref{lemma:rebalancecost} each costs at most $2.OPT$.
-    The total cost is therefore at most $2.OPT.k.\ell$ which implies the competitive ratio.
+      On each inter-cluster request,
+     the algorithm enumerates all $\ell$-way partitions of components
+     that are in the same (closest) distance of $P_I$.
+     That is, 
+     once it reaches $\Delta = \mathit{dist} (P_{t}, P_I)$,
+     it does not move to a partition
+     $P_{t'}, t'>t,\mathit{dist} (P_{t'}, P_I) > \Delta$
+     before it enumerates all partitions at distance $\Delta$.
+     Therefore,
+     PPL eventually reaches $\Delta=\OPT$ and the partition $P_F$.
+%    including the request that completes revealing of all components that are collocated in $P_F$.
+    There are at most $(k-1)\cdot\ell < k\cdot\ell $ calls   to $\mathit{re\mhyphen partition}$
+     (i.e., the number of internal edges in $P_F$).
+    By Lemma \ref{lemma:rebalancecost},
+    each costs at most $2\cdot\OPT$.
+    The total cost is therefore at most $2\cdot\OPT\cdot k\cdot\ell$ which implies the competitive ratio.
+    \mahmoud{This is the cost of moving and the cost of remote comm. is not counted.
+    	So it is 4-competitive (?)}
  \end{proof}
@@
    Let \emph{Arb-partition} be the algorithm that moves to an arbitrary perfect partition
-on each inter-cluster request.
+on each inter-cluster (i.e., external) request.
 \begin{theorem}

### 2020-01-28 paper/paper.tex
Comments on the algorithm definition

--- 2020-01-27 paper/paper.tex
@@
 
+Let \emph{Arb-partition} be the algorithm that moves to an arbitrary perfect partition
+on each inter-cluster request.
+
+\maciek{The definition of the algorithm should be: it moves to an arbitrary perfect partition that is the closest from the current partition.}
+\maciek{Therefore, I would name it differently. The easiest would be to call it "greedy" with an explaination what we mean. Alternatives: CPP (closest perfect partition)}
+
 \begin{lemma} \label{lemma:k=3}
-  The rebalancing cost for the greedy algorithm for $k=3$ is $O(1)$.
-  \maciek{Greedy = find the cheapest (closest to the current) feasible partition}
+  The rebalancing cost for the Arb-partition algorithm for $k=3$ is $O(1)$.
+  %\maciek{Greedy = find the cheapest (closest to the current) feasible partition}
   \label{rebalancing-cost}
@@
 
-   Let \emph{Arb-partition} be the algorithm that moves to an arbitrary perfect partition
-on each inter-cluster request.
 \begin{theorem}

### 2020-01-28 paper/paper.tex

--- 2020-01-28 paper/paper.tex
@@
 
-   Let \emph{Arb-partition} be the algorithm that moves to an arbitrary perfect partition
-on each inter-cluster (i.e., external) request.
 \begin{theorem}

### 2020-01-28 paper/paper.tex
Greedy it should be

@@
 
-Let \emph{Arb-partition} be the algorithm that moves to an arbitrary perfect partition
-on each inter-cluster request.
+Let \emph{Greedy} be the algorithm that on each inter-cluster request
+ moves to a perfect partition closest to the current one.
+ That is,
+ a partition reachable with the least number of node swaps.
 
@@
 \begin{lemma} \label{lemma:k=3}
-  The rebalancing cost for the Arb-partition algorithm for $k=3$ is $O(1)$.
+  The re-balancing cost for  Greedy algorithm for $k=3$ is $O(1)$.
   %\maciek{Greedy = find the cheapest (closest to the current) feasible partition}
@@
   Note that in such case, a $C_1$-type cluster exists, as otherwise the input graph would not have a perfect partition.
-The cost of rebalancing is then at most $12$ \maciek{can show smaller constant than 12, but it does not matter that much}.
+The cost of re-balancing is then at most $12$ \maciek{can show smaller constant than 12, but it does not matter that much}.
 \end{proof}
@@
 \begin{theorem}
-	Algorithm \emph{Arb-partition} is $O(\ell)$-competitive when $k=3$.
+	Algorithm \emph{Greedy} is $O(\ell)$-competitive when $k=3$.
 \end{theorem}
@@
 	The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
-	Each request triggers at most one re-balancing in \emph{Arb-partition}
+	Each request triggers at most one re-balancing in \emph{Greedy}
 	which costs $O(1)$ by Theorem~\ref{rebalancing-cost}.
@@
 
-We leverage the simple algorithm \emph{Arb-partition} for $k=3$
+We leverage the simple algorithm \emph{Greedy} for $k=3$
 towards an algorithm for the general problem
@@
 A phase terminates when there is no perfect partition of components.
-We apply requests to \emph{Arb-partition} as before. 
+We apply requests to \emph{Greedy} as before. 
 Once the current phase terminates,
@@
 	By Lemma  \ref{lemma:k=3},
-	\emph{Arb-partition} pays $O(\ell)$ during the phase.
+	\emph{Greedy} pays $O(\ell)$ during the phase.
 	If OPT stays in the same configuration throughout the phase,

### 2020-01-29 paper/paper.tex
Greedy it should be (1)
minor

@@
  That is,
- a partition reachable with the least number of node swaps.
+ a partition reachable with the least number of node swaps from the current partition.
 
@@
 \begin{lemma} \label{lemma:k=3}
-  The re-balancing cost for  Greedy algorithm for $k=3$ is $O(1)$.
+  The re-balancing cost of Greedy for $k=3$ is $O(1)$.
   %\maciek{Greedy = find the cheapest (closest to the current) feasible partition}

### 2020-01-29 paper/paper.tex
Sections Model and Lower bound

@@
 
-Formally, the online \emph{Balanced Re-Partitioning} problem (BRP) is defined as
-follows. There is a set of $n$ nodes, initially distributed arbitrarily
-across $\ell$~clusters, each of size~$k$. We call two nodes~$u,v\in V$
-\emph{collocated} if they are in the same cluster.
+Formally, the online \emph{Balanced Re-Partitioning} problem (OBRP) is defined as
+follows. There is a set of $n$ nodes,
+initially arbitrarily partitioned into $\ell$~clusters,
+each of size~$k$.
+Nodes in~$C \subset V$ are \emph{collocated} by the partition $P$
+if $P$ assigns them to the same cluster.
+I.e.,
+$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.
+The input to the problem is a sequence of communication requests
+$\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
+where a pair $(u_t,v_t)$ indicates that nodes $u_t,v_t$ exchange a fixed amount of data.
+An online algorithm needs to serve the communication request~$(u_t,v_t)$
+as soon as it arrives.
+Before serving the request,
+the online algorithm
+can re-partition,
+i.e.,
+it may move some of the nodes into clusters different than their current assigned clusters.
+We assume that
+a~communication request between two collocated nodes costs 0,
+the cost of a~communication request between two nodes located in different clusters is~1,
+and the cost of migrating a node from one cluster to another
+is~$\alpha \in \mathbb{Z}^+$.
+For any algorithm $ALG$,
+its cost,
+denoted by $ALG(\sigma)$,
+is the total cost of communications and
+the cost of migrations on the sequence $\sigma$.
 
-An input to the problem is a sequence of communication requests $\sigma =
-(u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$, where pair $(u_t,v_t)$ means that
-the nodes $u_t,v_t$ exchange a fixed amount of data.  At any time~$t$, an online algorithm needs to serve the~communication
-request~$(u_t,v_t)$. Right before serving the request, the online algorithm
-can repartition the nodes into new clusters. We assume that
-a~communication request between two collocated nodes costs 0. The cost of a~communication request between two nodes located in different clusters is
-normalized to~1, and the cost of migrating a node from one cluster to another
-is~$\alpha \geq 1$, where $\alpha$ is a parameter (an~integer). For any
-algorithm $ALG$, we denote its total cost (consisting of communication plus
-migration costs) on sequence $\sigma$ by $ALG(\sigma)$.
+\section{Lower bound} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
 
-\section{Lower bound $\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
-
-In this section we consider the Online Balanced Partition problem.
-We show a lower bound of $\Omega(k \cdot \ell)$ that uses an input that has a perfect partition.
-Therefore, this result is a lower bound for the general Online Balanced Partition problem, as well as a lower bound for Online Balanced Partition with perfect partition.
-The latter complements an upper bound $O(k \cdot \ell)$ for Online Balanced Partition with perfect partition that we present in Section \ref{sec:upperbound}.
+In this section we show a lower bound of $\Omega(k \cdot \ell)$
+using an input sequence that allows OPT to move to a perfect partition.
+Therefore,
+it is a lower bound for OBRP with perfect partition,
+which directly implies a lower bound for the (general) OBRP.
+The former complements an upper bound in $O(k \cdot \ell)$
+for the OBRP with perfect partition which we present in Section \ref{sec:upperbound}.
 
 \begin{theorem}
-  The competitive ratio of any deterministic algorithm for Online Balanced Partition is $\Omega(k\cdot \ell)$.
+  The competitive ratio of any deterministic algorithm for OBRP is in $\Omega(k\cdot \ell)$.
 \end{theorem}
@@
 Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
-Consider any online algorithm ALG.
+Let ALG be any online algorithm for OBRP.
 We construct the input sequence of requests for ALG as follows.
 First,
-we issue $k-2$ (internal) requests so that ALG form a component of $k-1$
-nodes on clusters $S_1$.
-Note that this does not cost ALG.
-Let $x_0$  be the only single nodes left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
+we issue $k-2$ (internal) requests so that ALG can form a component of $k-1$
+nodes on the cluster $S_1$.
+Recall that internal requests does not incur ALG any cost.
+If ALG at any point splits a component then
+we issue  requests between the parts as many as it is necessary for ALG to collocate the parts,
+as otherwise ALG is not competitive.
+Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
 Next,
@@
 For this,
-ALG moves to a new configuration that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
+ALG moves to a new configuration (i.e., partition)
+that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
 Next,
@@
 ALG must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
-consequently replaces $x_1$.
+consequently replaces $x_1$ with some other (single) node $x_2$.
 We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
@@
 I.e.,
-two single nodes $x^*, y^*,I(x^*) = I(y^*)$
-that constitute the last pair of such nodes.
-At this point there are at most $\ell+1$ single nodes left
-since there must be at least two nodes with the same initial cluster in any subset of $\ell+1$
-nodes.
+two single nodes $x^*, y^*,I(x^*) = I(y^*)$ s.t.~for any other pair of single nodes
+$x'$ and $y'$,
+we have $I(x') \neq I(y')$.
+At this point there are at most $\ell+1$ single nodes left,
+otherwise there would be more pairs of single nodes that were initially in the same cluster.
 Given this sequence of requests,
 the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
- swapping $x_0$ and $y_0$ with $x^*$ and $y^*$ respectively.
+ swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
 Hence,
@@
 
-\section{Perfect Partition Learner Algorithm}
+\section{Perfect Partition Learner Algorithm}	\label{sec:upperbound}
 
@@
 %        \If{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$}
-        \If{$C_1$ and $C_2$ are not in the same cluster w.r.t.~$P_{t-1}$}
+        \If{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$}
+        \Comment{if not in the same cluster}
         \State $P_{t} = \mathit{re\mhyphen partition}(P_{t-1}, P_I, \mathcal{C}$) \label{line:rebalance} 
@@
 
-Let \emph{Greedy} be the algorithm that on each inter-cluster request
+Let \emph{GREEDY} be the algorithm that on each inter-cluster request
  moves to a perfect partition closest to the current one.
@@
 \begin{lemma} \label{lemma:k=3}
-  The re-balancing cost of Greedy for $k=3$ is $O(1)$.
-  %\maciek{Greedy = find the cheapest (closest to the current) feasible partition}
+  The re-balancing cost of GREEDY for $k=3$ is $O(1)$.
+  %\maciek{GREEDY = find the cheapest (closest to the current) feasible partition}
   \label{rebalancing-cost}
@@
 \begin{theorem}
-	Algorithm \emph{Greedy} is $O(\ell)$-competitive when $k=3$.
+	Algorithm \emph{GREEDY} is $O(\ell)$-competitive when $k=3$.
 \end{theorem}
@@
 	The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
-	Each request triggers at most one re-balancing in \emph{Greedy}
+	Each request triggers at most one re-balancing in \emph{GREEDY}
 	which costs $O(1)$ by Theorem~\ref{rebalancing-cost}.
@@
 
-We leverage the simple algorithm \emph{Greedy} for $k=3$
+We leverage the simple algorithm \emph{GREEDY} for $k=3$
 towards an algorithm for the general problem
@@
 A phase terminates when there is no perfect partition of components.
-We apply requests to \emph{Greedy} as before. 
+We apply requests to \emph{GREEDY} as before. 
 Once the current phase terminates,
@@
 	By Lemma  \ref{lemma:k=3},
-	\emph{Greedy} pays $O(\ell)$ during the phase.
+	\emph{GREEDY} pays $O(\ell)$ during the phase.
 	If OPT stays in the same configuration throughout the phase,

### 2020-01-29 paper/paper.tex
introduction

@@
 
-Main technical contribution of this paper is $\Omega(k\ell)$ lower bound for the competitive ratio of any deterministic algorithm for Online Balanced Partition problem.
-The best known lower bound so far was $\Omega(k)$ \cite{repartition-disc}.
+Main technical contribution of this paper is the lower bound $\Omega(k\ell)$ for the competitive ratio of any deterministic algorithm for Online Balanced Partition problem.
+The best known lower bound so far was $\Omega(k)$ \cite{repartition-disc}
+which involves only two cluster.
 
-Additionally, we contribute to search for the optimal algorithm by presenting $O(k\ell)$-competitive algorithm for the special case, where the perfect partition of the input graph is assumed, and every algorithm must colocate every communicating pair.
-Note that our $\Omega(k\ell)$ lower bound uses a perfect partition input graph and colocates communicating pairs, and hence for this scenario, our results are tight.
-\maciek{Also, it is easy to construct the adversary that forces collocation of requested nodes}
+Additionally, we contribute to the research on the optimal online algorithm by presenting a $O(k\ell)$-competitive algorithm for the special case
+where the communication graph (i.e., request sequence) admits a \emph{perfect partition},
+i.e., a (multi-way) cut of cost zero that is also an optimal partition for the offline problem.
+%and every algorithm must colocate every communicating pair.
+Note that our $\Omega(k\ell)$ lower bound generates the input sequence
+so that the optimal offline choice is a perfect partition.
+\maciek{Also, it is easy to construct the adversary that forces collocation of requested nodes
+\mahmoud{this is already the case in the lower bound proof. Do we still need the last sentence here?}
+}
 

### 2020-01-29 paper/paper.tex
Minor rewrite, comments

@@
 \newcommand{\OPT}{\mathit{OPT}}
+\newcommand{\ALG}{ALG}
+\newcommand{\OBRP}{OBRP}
+\newcommand{\PPOBRP}{PPOBRP}
 
@@
 so that the optimal offline choice is a perfect partition.
-\maciek{Also, it is easy to construct the adversary that forces collocation of requested nodes
+\maciek{Also, it is easy to construct the adversary that forces collocation of requested nodes}
 \mahmoud{this is already the case in the lower bound proof. Do we still need the last sentence here?}
-}
+\maciek{In general it is OK to emphasize some parts of the paper in the intro. Not sure if this sentence in particular is that important, though.}
 
@@
 
-Formally, the online \emph{Balanced Re-Partitioning} problem (OBRP) is defined as
-follows. There is a set of $n$ nodes,
+\maciek{I suggest to rename the problem to either: (1) Dynamic Balanced Graph Partition or (2) Online Balanced Graph Partition, as in the SIDMA paper}
+
+Formally, the online \emph{Balanced Re-Partitioning} problem (\OBRP) is defined as
+follows. We are given a set of $n$ nodes,
 initially arbitrarily partitioned into $\ell$~clusters,
 each of size~$k$.
-Nodes in~$C \subset V$ are \emph{collocated} by the partition $P$
+We say that a set of nodes $C \subset V$ is \emph{collocated} in a partition $P$
 if $P$ assigns them to the same cluster.
-I.e.,
-$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.
-The input to the problem is a sequence of communication requests
+\maciek{We can ommit the formal definition: "I.e.,
+$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". However, if we use the notion of cluster later, then TODO: define cluster}
+The input consists of a sequence of pairwise communication requests
 $\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
-where a pair $(u_t,v_t)$ indicates that nodes $u_t,v_t$ exchange a fixed amount of data.
+where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a fixed amount of data.
 An online algorithm needs to serve the communication request~$(u_t,v_t)$
@@
 is the total cost of communications and
-the cost of migrations on the sequence $\sigma$.
+the cost of migrations performed by $ALG$ while serving the sequence $\sigma$.
+\maciek{Not an expert on commas in english, but here it feels like we have too much. Use some tool to determine (Grammarly?)}
+
+\maciek{Here introduce a restricted model of \PPOBRP, the restricted variant where the input sequence can be perfectly partitioned}
 
@@
 
-In this section we show a lower bound of $\Omega(k \cdot \ell)$
-using an input sequence that allows OPT to move to a perfect partition.
-Therefore,
-it is a lower bound for OBRP with perfect partition,
-which directly implies a lower bound for the (general) OBRP.
-The former complements an upper bound in $O(k \cdot \ell)$
-for the OBRP with perfect partition which we present in Section \ref{sec:upperbound}.
+In this section we show a lower bound of $\Omega(k \cdot \ell)$.
+The constructed input sequence allows OPT to move to a perfect partition, and hence
+the construction also constitues a lower bound for \PPOBRP.
+This complements an upper bound of $O(k \cdot \ell)$
+for \PPOBRP that we present in Section \ref{sec:upperbound}.
 
 \begin{theorem}
-  The competitive ratio of any deterministic algorithm for OBRP is in $\Omega(k\cdot \ell)$.
+  The competitive ratio of any deterministic algorithm for \OBRP is in $\Omega(k\cdot \ell)$.
 \end{theorem}
@@
 Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
-Let ALG be any online algorithm for OBRP.
-We construct the input sequence of requests for ALG as follows.
+\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}
+Fix any online algorithm \ALG.
+We construct the input sequence of requests for \ALG as follows.
 First,
-we issue $k-2$ (internal) requests so that ALG can form a component of $k-1$
+we issue $k-2$ (internal) requests so that \ALG can form a component of $k-1$
 nodes on the cluster $S_1$.
-Recall that internal requests does not incur ALG any cost.
-If ALG at any point splits a component then
-we issue  requests between the parts as many as it is necessary for ALG to collocate the parts,
-as otherwise ALG is not competitive.
+Recall that internal requests does not incur any cost for \ALG.
+If \ALG at any point splits a component\maciek{define splitting}, then
+we continue to issue requests between non-collocated nodes of the component until \ALG collocates the nodes of the component.
+If \ALG never collocates them then it is not competitive.
+
 Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
@@
 that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
+
 Next,
@@
 otherwise there would be more pairs of single nodes that were initially in the same cluster.
+
 Given this sequence of requests,
@@
 $P_F = F_1, \dots, F_{\ell}$.
- The \emph{distance} of a partition $P = C_1, \dots, C_{\ell}$ from the initial partition
- is the number of nodes in $P$ that do not reside in their initial cluster.
+ The \emph{distance} of a partition $P = C_1, \dots, C_{\ell}$ from the initial partition is the number of nodes in $P$ that do not reside in their initial cluster.
     That is,
@@
      it does not move to a partition
-     $P_{t'}, t'>t,\mathit{dist} (P_{t'}, P_I) > \Delta$
+     $P_{t'}$, for any $t'>t,\mathit{dist} (P_{t'}, P_I) > \Delta$,
      before it enumerates all partitions at distance $\Delta$.
@@
     By Lemma \ref{lemma:rebalancecost},
-    each costs at most $2\cdot\OPT$.
+    each re-partition costs at most $2\cdot\OPT$.
     The total cost is therefore at most $2\cdot\OPT\cdot k\cdot\ell$ which implies the competitive ratio.
@@
 Let \emph{GREEDY} be the algorithm that on each inter-cluster request
- moves to a perfect partition closest to the current one.
- That is,
- a partition reachable with the least number of node swaps from the current partition.
-
-\maciek{The definition of the algorithm should be: it moves to an arbitrary perfect partition that is the closest from the current partition.}
-\maciek{Therefore, I would name it differently. The easiest would be to call it "greedy" with an explaination what we mean. Alternatives: CPP (closest perfect partition)}
+ moves to arbitrary perfect partition \maciek{of input} closest (i.e. reachable in minimum number of node swaps) to the current one.
 
 \begin{lemma} \label{lemma:k=3}
-  The re-balancing cost of GREEDY for $k=3$ is $O(1)$.
+  Every re-balancing cost of GREEDY for $k=3$ is $O(1)$.
   %\maciek{GREEDY = find the cheapest (closest to the current) feasible partition}
+  \maciek{We need to add to the assumption that the input has a perfect partition}
   \label{rebalancing-cost}
@@
 
-  We distinguish among three configuration types of algorithm's clusters: $C_1, C_2, C_3$.
-In $C_1$ we have $3$ singleton components (this is also the initial configuration of any cluster).
-In $C_2$ we have one component of size $2$ and one component of size $1$.
-Finally, in $C_3$ we have one component of size $3$.
+  \maciek{We probably should start with something along: Note that the assumption of perfect partition means that there is no component of size 4. Then refer to it (twice) by some short ref.}
 
-  Consider a request $(u, v)$ and let $U, V$ be their clusters.
-  If $U=V$, then the request does not require a reconfiguration.
-  Now, we consider cases upon the type of clusters $U$ and $V$.
-  Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise the resulting component would have the size $4$, and this contradicts the existence of the perfect partition.
+  We distinguish among three configuration types of any algorithm's cluster: $C_1, C_2, C_3$, which we define as follows.
+In $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+In $C_2$ it contains one component of size $2$ and one component of size $1$.
+Finally, in $C_3$ it contains one component of size $3$.
+
+  Consider a request $(u, v)$ and let $U, V$ be clusters of these nodes.
+  If $U=V$, then the request does not require a reconfiguration and GREEDY incurs no cost.
+  Otherwise, we consider cases upon the type of clusters $U$ and $V$.
+  Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise we encounter a component of size $4$ and this would contradict the existence of the perfect partition of the input.
   If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
 
-  Now, we focus on the case where both $U$ and $V$ are $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as the merged component would have size $4$.
+  Finally, we consider the case where both $U$ and $V$ are $C_2$.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that an input sequence includes the component of size $4$, a contradiction.
   If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
   Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-  Note that in such case, a $C_1$-type cluster exists, as otherwise the input graph would not have a perfect partition.
-The cost of re-balancing is then at most $12$ \maciek{can show smaller constant than 12, but it does not matter that much}.
+  Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise the input graph would not have a perfect partition.
+In this case GREEDY exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost of $2\cdot \alpha$.
 \end{proof}
@@
   Algorithm \emph{GREEDY} is $O(\ell)$-competitive when $k=3$.
+  \maciek{Assumming the perfect partition of input sequence}
 \end{theorem}
@@
 \begin{proof}
+  \maciek{TODO: this proof is incomplete!}
 	Consider any phase and assume it terminates upon arrival of some request $r$.

### 2020-02-13 paper/paper.tex
Title was incomplete
Format question
Abstract WIP

@@
 
-\title{Brief Announcement: An Improved Lower Bound for Dynamic Graph Partitioning}
+\title{Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning}
 
@@
 
-\keywords{online algorithms, competitive analysis, etc}
+\keywords{online algorithms, competitive analysis}
 
@@
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+  In the Dynamic Graph Partitioning problem, we are given a sequence of pairwise communication requests between n nodes, with patterns that may change over time.
+  The objective is to service these requests by partitioning the nodes into $\ell$ clusters, each of size $k$, such that frequently communicating nodes are located in the same cluster.
+  The partitioning can be updated dynamically by migrating nodes between clusters.
+  The goal is to devise online algorithms which jointly minimize the amount of inter-cluster communication and migration cost.
+
 \end{abstract}
@@
 
+\todo{ACM format for review? With line numbers etc.}
+
 \todo{Make ORCID visible in ACM style}

### 2020-02-13 paper/paper.tex
Abstract v1

@@
 
+  So far, for this natural problem, the best known deterministic algorithm was $O(k^2\cdot \ell^2)$-competitive, and the best known lower bound was $\Omega(k)$ for the competitive ratio of any deterministic algorithm that used the reduction from the caching problem.
+  In this paper, we significantly improve this lower bound to $\Omega(k\cdot \ell)$.
+  Furthermore, we provide a $O(k\cdot \ell)$-competitive algorithm for the scenario, where the perfect partition of input exists.
 \end{abstract}

### 2020-02-13 paper/paper.tex
correcting cost model, fixing macros

@@
 
-Formally, the online \emph{Balanced Re-Partitioning} problem (\OBRP) is defined as
+Formally, the online \emph{Balanced Re-Partitioning} problem (\OBRP{}) is defined as
 follows. We are given a set of $n$ nodes,
@@
 each of size~$k$.
-We say that a set of nodes $C \subset V$ is \emph{collocated} in a partition $P$
+We say that nodes in $C \subset V$ are \emph{collocated} in a partition $P$
 if $P$ assigns them to the same cluster.
@@
 where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a fixed amount of data.
-An online algorithm needs to serve the communication request~$(u_t,v_t)$
-as soon as it arrives.
 Before serving the request,
@@
 i.e.,
-it may move some of the nodes into clusters different than their current assigned clusters.
-We assume that
-a~communication request between two collocated nodes costs 0,
-the cost of a~communication request between two nodes located in different clusters is~1,
-and the cost of migrating a node from one cluster to another
+it may (optionally) move some of the nodes into clusters different than their current assigned clusters.
+Once the re-partition is executed,
+we serve
+a~communication request between two  nodes at cost~0 if they are collocated,
+or at cost~1 if they are located in different clusters.
+The cost of migrating a node from one cluster to another
 is~$\alpha \in \mathbb{Z}^+$.
@@
 
-\maciek{Here introduce a restricted model of \PPOBRP, the restricted variant where the input sequence can be perfectly partitioned}
+\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
 
@@
 The constructed input sequence allows OPT to move to a perfect partition, and hence
-the construction also constitues a lower bound for \PPOBRP.
+the construction also constitutes a lower bound for \PPOBRP{}.
 This complements an upper bound of $O(k \cdot \ell)$
-for \PPOBRP that we present in Section \ref{sec:upperbound}.
+for \PPOBRP{} that we present in Section \ref{sec:upperbound}.
 
 \begin{theorem}
-  The competitive ratio of any deterministic algorithm for \OBRP is in $\Omega(k\cdot \ell)$.
+  The competitive ratio of any deterministic algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$.
 \end{theorem}
@@
 Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
-\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}
+\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
 Fix any online algorithm \ALG.
-We construct the input sequence of requests for \ALG as follows.
+We construct the input sequence of requests for \ALG{} as follows.
 First,
-we issue $k-2$ (internal) requests so that \ALG can form a component of $k-1$
+we issue $k-2$ (internal) requests so that \ALG{} can form a component of $k-1$
 nodes on the cluster $S_1$.
 Recall that internal requests does not incur any cost for \ALG.
-If \ALG at any point splits a component\maciek{define splitting}, then
-we continue to issue requests between non-collocated nodes of the component until \ALG collocates the nodes of the component.
-If \ALG never collocates them then it is not competitive.
+If \ALG{}at any point splits a component\maciek{define splitting}, then
+we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
+If \ALG{} never collocates them then it is not competitive.
 

### 2020-02-13 paper/paper.tex
ORCIDs resolved
minor

@@
 we serve
-a~communication request between two  nodes at cost~0 if they are collocated,
+the communication request between two  nodes at cost~0 if they are collocated,
 or at cost~1 if they are located in different clusters.

### 2020-02-13 paper/paper.tex
dots

--- 2020-02-13 paper/paper.tex
@@
 \newcommand{\ALG}{ALG}
-\newcommand{\OBRP}{OBRP}
-\newcommand{\PPOBRP}{PPOBRP}
+\newcommand{\OBRP}{BRP}
+\newcommand{\PPOBRP}{PP-BRP}
 
@@
 \author{Maciej Pacut}
-%\orcid{0000-0002-6379-1490}
+\orcid{0000-0002-6379-1490}
 \affiliation{%
-  \department{Faculty of Computer Science}
-  \institution{University of Vienna}
+  \institution{Faculty of Computer Science, University of Vienna}
   \country{Austria}
@@
 \affiliation{%
-  \department{Faculty of Computer Science}
-  \institution{University of Vienna}
+  \institution{Faculty of Computer Science, University of Vienna}
   \country{Austria}
@@
 
-Additionally, we contribute to the research on the optimal online algorithm by presenting a $O(k\ell)$-competitive algorithm for the special case
+Additionally, we contribute to the research on the optimal online algorithm by presenting a $O(k\cdot \ell)$-competitive algorithm for the special case
 where the communication graph (i.e., request sequence) admits a \emph{perfect partition},
@@
 %and every algorithm must colocate every communicating pair.
-Note that our $\Omega(k\ell)$ lower bound generates the input sequence
+Note that our $\Omega(k\cdot \ell)$ lower bound generates the input sequence
 so that the optimal offline choice is a perfect partition.
-\maciek{Also, it is easy to construct the adversary that forces collocation of requested nodes}
-\mahmoud{this is already the case in the lower bound proof. Do we still need the last sentence here?}
-\maciek{In general it is OK to emphasize some parts of the paper in the intro. Not sure if this sentence in particular is that important, though.}
+%\maciek{Also, it is easy to construct the adversary that forces collocation of requested nodes}
+%\mahmoud{this is already the case in the lower bound proof. Do we still need the last sentence here?}
+%\maciek{In general it is OK to emphasize some parts of the paper in the intro. %Not sure if this sentence in particular is that important, though.}
 
@@
 In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.
-
+t
 We assume that the physical substrate network is a~tree of height one.
@@
 
-\maciek{I suggest to rename the problem to either: (1) Dynamic Balanced Graph Partition or (2) Online Balanced Graph Partition, as in the SIDMA paper}
-
-Formally, the online \emph{Balanced Re-Partitioning} problem (\OBRP{}) is defined as
+Formally, the online \emph{Balanced Graph RePartitioning} problem (\OBRP{}) is defined as
 follows. We are given a set of $n$ nodes,
@@
 ALG pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
-We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k.\ell$ nodes,
-therefore $|X| \geq k.\ell - k - \ell \in \Omega(k\cdot\ell)$.
+We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
+therefore $|X| \geq k \cdot \ell - k - \ell \in \Omega(k\cdot\ell)$.
 \end{proof}
@@
 
-The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\ell^2)$.
+The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\cdot\ell^2)$.
 
@@
 
-\todo{Make ORCID visible in ACM style}
+\todo{Make ORCID visible in ACM style: "Right now we do not typeset ORCIDs" (Source: ACM Style Docs)}
 

### 2020-02-13 paper/paper.tex
Final name of the problem
Cutting intro

@@
 
-Main technical contribution of this paper is the lower bound $\Omega(k\ell)$ for the competitive ratio of any deterministic algorithm for Online Balanced Partition problem.
+We study (...) in the scenario where virtual machines can be migrated during runtime to another physical machine.
+The possibility of migration allows reacting to unpredictable communication patterns.
+If some distant nodes communicate often, it is vital to reduce their distance to save network bandwidth.
+The objective is to~minimize the total network bandwidth used for communication and for migration.
+
+Main technical contribution of this paper is the lower bound $\Omega(k\cdot\ell)$ for the competitive ratio of any deterministic algorithm for Online Balanced Partition problem.
 The best known lower bound so far was $\Omega(k)$ \cite{repartition-disc}
@@
 
-In this paper we study the Online Balanced Partitioning problem (formally presented in Section~\ref{sec:problem-definition}).
+In this paper we study the online Balanced RePartitioning problem (formally presented in Section~\ref{sec:problem-definition}).
 The static offline version of~the~problem, i.e., a problem variant where
-migration is not allowed, where all requests are known in advance, and where
-the goal is to find an assignment of $n$ nodes to $\ell$~physical machines, each of~capacity $n/\ell$, is known as the
+migration is not allowed, and all requests are known in advance is known as the
 \emph{$\ell$-balanced graph partitioning problem}. The problem is 
@@
 the currently best approximation ratio is $O(\log n)$~\cite{SarVaz95,ArKaKa99,FeKrNi00,FeiKra02,KraFei06,Raec08}.
-The inapproximability of the static variant for general values of $\ell$
-motivated research on the bicriteria variant, which can be seen as the offline
-counterpart of our capacity augmentation approach. Here, the~goal
-is~to~compute a~graph partitioning into $\ell$ components of~size at most~$k$ (where $k > n/\ell$) and the cost of the cut is compared to the optimal (non-augmented)
-solution where all components are of a~size at most $n/k$. The variant where
-$n \geq 2 \cdot k \cdot \ell$ was considered in
-\cite{LeMaTr90,SimTen97,EvNaRS00,EvNaRS99,KrNaSc09}. So far, the~best result~is~an~$O(\!\sqrt{\log n \cdot \log \ell})$-approximation algorithm~\cite{KrNaSc09}.
 
-Our model is related to online
-caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}, sometimes also referred to
-as online caching, where requests for data items (nodes) arrive over time and
-need to be served from a cache of finite capacity, and where the number of
-cache misses must be minimized. Classic problem variants usually boil down to
-finding a smart eviction strategy, such as Least Recently Used (LRU)~\cite{SleTar85}. In our
-setting, requests can be served remotely (i.e.,~without fetching the
-corresponding nodes to a single physical machine). In this light, our model is more
+The model is related to online
+caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}. In our
+setting, requests can be served remotely, which is
 reminiscent of caching models \emph{with
-bypassing}~\cite{EpImLN11,EpImLN15,Irani02}. As a~side result, we show that our problem is
-capable of emulating online caching.
-A major difference between  these problems is that in the caching problems, each request involves a~single element of the universe, while in our model \emph{both} endpoints of a communication request are subject to~optimization.
-
-\section{Practical Motivation}
-
-Distributed cloud applications, including batch processing
-applications such as MapReduce, streaming applications such as Apache Flink or
-Apache Spark, and scale-out databases and key-value stores such as Cassandra,
-generate a~significant amount of network traffic and a~considerable fraction
-of their runtime is due to network activity~\cite{MogPop12}. For example,
-traces of jobs from a Facebook data center reveal that network transfers on
-average account for 33\% of the execution time~\cite{orchestra}. In such
-applications, it is desirable that frequently communicating virtual machines
-are \emph{collocated}, i.e., mapped to the same physical server: 
-communication across the network (i.e., inter-server communication) induces
-network load and latency. However, migrating virtual machines between servers
-also comes at a price: the state transfer is bandwidth intensive, and may even
-lead to short service interruptions. Therefore the goal is to design online
-algorithms that find a good trade-off between the inter-server communication
-cost and the migration cost.
-
-\vspace{1cm}
-
-We study virtual network embeddings in the scenario where virtual machines can be migrated during runtime to another physical machine.
-The possibility of migration allows reacting to unpredictable communication patterns.
-For example, if some distant nodes communicate often, it is vital to reduce their distance to save network bandwidth.
-The objective is to~minimize the total network bandwidth used for communication and for migration.
-
-We assume that the communication patterns are not known in advance to our algorithm.
-We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.
-In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.
-t
-We assume that the physical substrate network is a~tree of height one.
-That is, every physical machine (leaf) is connected directly to the root (that has no hosting capabilities).
-A single physical machine hosts a fixed number of virtual machines.
-The model restricted to such networks becomes a variant of online graph clustering.
-That is, we are given a~set of~$n$ nodes (virtual machines) with time-varying pairwise
-communication patterns, which have to be partitioned into~$\ell$~physical machines, each of
-capacity $k=n/\ell$.
-
-Intuitively, we would like to minimize inter-machine
-interactions by mapping frequently communicating nodes to the same physical machine
-Since communication patterns change over time, the~nodes should be \emph{repartitioned}, in
-an online manner, by \emph{migrating} them between physical machines.
-The~objective is to minimize the weighted sum of inter-machine communication and repartitioning costs.
-The former is defined as the number of communication requests between nodes placed at distinct physical machines, and the latter as the number of migrations.
-
-The possibility to perform a migration uncovers algorithmic challenges:
-\begin{itemize}
-
-\item \emph{Serve remotely or migrate?} For a brief communication
-pattern, it may not be worthwhile to collocate the nodes: the migration cost might
-be too large in comparison to~communication costs.
-
-\item \emph{Where to migrate, and what?}
-If an algorithm decides to collocate nodes $x$ and~$y$, the~question becomes
-how. Should $x$ be migrated to the physical machine holding $y$, $y$ to the one holding
-$x$, or should both nodes be migrated to a new machine?
-
-\item \emph{Which nodes to evict?}
-The space of the desired destination physical machine may not be sufficient. In
-this case, the~algorithm needs to decide which nodes to ``evict'' (migrate to
-other machines), to free up space.
-
-\end{itemize}
+bypassing}~\cite{EpImLN11,EpImLN15,Irani02}.
 
@@
 
-Formally, the online \emph{Balanced Graph RePartitioning} problem (\OBRP{}) is defined as
+Formally, the online \emph{Balanced RePartitioning} problem (\OBRP{}) is defined as
 follows. We are given a set of $n$ nodes,

### 2020-02-13 paper/paper.tex
More cutting

@@
 We say that nodes in $C \subset V$ are \emph{collocated} in a partition $P$
-if $P$ assigns them to the same cluster.
-\maciek{We can ommit the formal definition: "I.e.,
-$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". However, if we use the notion of cluster later, then TODO: define cluster}
+if they are assigned to the same cluster.
+%\maciek{We can ommit the formal definition: "I.e.,
+%$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". %However, if we use the notion of cluster later, then TODO: define cluster}
 The input consists of a sequence of pairwise communication requests
@@
 the cost of migrations performed by $ALG$ while serving the sequence $\sigma$.
-\maciek{Not an expert on commas in english, but here it feels like we have too much. Use some tool to determine (Grammarly?)}
 
@@
 
-\todo{$. \rightarrow \cdot$}
+\todo{Check grammar}
 

### 2020-02-13 paper/paper.tex
perfect partition and new sections

@@
 %$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". %However, if we use the notion of cluster later, then TODO: define cluster}
+%\maciek{We can ommit the formal definition: "I.e.,
+%$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". However, if we use the notion of cluster later, then TODO: define cluster}
 The input consists of a sequence of pairwise communication requests
@@
 
+\subsection{Perfect Partition Model}\label{sec:perfectPartition}
 \maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
+
+In this paper we mostly focus on a restricted variant of   \PPOBRP{}
+where the communication graph can be partitioned into $\ell$ clusters without any inter-cluster edges.
+Moreover, we assume an optimal offline algorithm (OPT) moves to this \emph{perfect partition}
+at the beginning and stays there permanently.
+Then the task of an online algorithm is to "recover" (or learn) the perfect partition while not paying too much more than OPT.
 
@@
 
-\section{Perfect Partition Learner Algorithm}	\label{sec:upperbound}
+\section{Recovering a Perfect Partition}	\label{sec:upperbound}
 
@@
 
-      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial configuration}
-
+%      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial configuration}
       We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current partition $P$ with a (perfect) partition closest to $P_I$.
@@
     The total cost is therefore at most $2\cdot\OPT\cdot k\cdot\ell$ which implies the competitive ratio.
-    \mahmoud{This is the cost of moving and the cost of remote comm. is not counted.
-    	So it is 4-competitive (?)}
+%    \mahmoud{This is the cost of moving and the cost of remote comm. is not counted.
+%    	So it is 4-competitive (?)}
  \end{proof}
 
-\section{$O(\ell)$-competitive algorithm for $k=3$}
+\subsection{$O(\ell)$-competitive algorithm for $k=3$}
 
@@
 \end{proof}
+
+\section{Arbitrary Communication Graph} \label{sec:generalModel}
 

### 2020-02-13 paper/paper.tex
Intro

--- 2020-02-13 paper/paper.tex
@@
 \begin{abstract}
-  In the Dynamic Graph Partitioning problem, we are given a sequence of pairwise communication requests between n nodes, with patterns that may change over time.
+  In the online Graph RePartitioning problem, we are given a sequence of pairwise communication requests between n nodes, revealed online.
   The objective is to service these requests by partitioning the nodes into $\ell$ clusters, each of size $k$, such that frequently communicating nodes are located in the same cluster.
-  The partitioning can be updated dynamically by migrating nodes between clusters.
-  The goal is to devise online algorithms which jointly minimize the amount of inter-cluster communication and migration cost.
+  The partitioning can be updated by migrating nodes between clusters for a fixed cost.
+  The goal is to jointly minimize the amount of inter-cluster communication and migration cost.
 
   So far, for this natural problem, the best known deterministic algorithm was $O(k^2\cdot \ell^2)$-competitive, and the best known lower bound was $\Omega(k)$ for the competitive ratio of any deterministic algorithm that used the reduction from the caching problem.
-  In this paper, we significantly improve this lower bound to $\Omega(k\cdot \ell)$.
-  Furthermore, we provide a $O(k\cdot \ell)$-competitive algorithm for the scenario, where the perfect partition of input exists.
+  In this paper, we provide the lower bound of $\Omega(k\cdot \ell)$.
+  %Furthermore, we provide a $O(k\cdot \ell)$-competitive algorithm for the scenario, where the perfect partition of input exists.
 \end{abstract}
@@
 
-\section{Introduction}
+\section{Introduction and Related Work}
 
-We study (...) in the scenario where virtual machines can be migrated during runtime to another physical machine.
-The possibility of migration allows reacting to unpredictable communication patterns.
-If some distant nodes communicate often, it is vital to reduce their distance to save network bandwidth.
-The objective is to~minimize the total network bandwidth used for communication and for migration.
+%We study (...) in the scenario where virtual machines can be migrated during runtime to another physical machine.
+%The possibility of migration allows reacting to unpredictable communication patterns.
+%If some distant nodes communicate often, it is vital to reduce their distance to save network bandwidth.
+%The objective is to~minimize the total network bandwidth used for communication and for migration.
 
@@
 
-Additionally, we contribute to the research on the optimal online algorithm by presenting a $O(k\cdot \ell)$-competitive algorithm for the special case
-where the communication graph (i.e., request sequence) admits a \emph{perfect partition},
-i.e., a (multi-way) cut of cost zero that is also an optimal partition for the offline problem.
-%and every algorithm must colocate every communicating pair.
-Note that our $\Omega(k\cdot \ell)$ lower bound generates the input sequence
-so that the optimal offline choice is a perfect partition.
-%\maciek{Also, it is easy to construct the adversary that forces collocation of requested nodes}
-%\mahmoud{this is already the case in the lower bound proof. Do we still need the last sentence here?}
-%\maciek{In general it is OK to emphasize some parts of the paper in the intro. %Not sure if this sentence in particular is that important, though.}
+So far, the competitive ratio of the best known algorithm was $O(k^2\cdot \ell^2)$ \cite{repartition-disc}.
+We present a $O(k\cdot \ell)$-competitive algorithm for the restricted scenario
+where the communication graph (i.e., request sequence) admits a \emph{perfect partition} (i.e., a multi-way cut of cost zero) that is also an optimal partition for the offline problem.
+For this setting, the algorithm is optimal, as the lower bound utilized an input sequence that has a perfect partition.
 
-\section{Related Work}
+The problem was studied under resource augmentation \cite{repartition-disc}, and under stochastic assumptions of input\maciek{cite}.\maciek{cite other papers that cite DISC?}
 
-In this paper we study the online Balanced RePartitioning problem (formally presented in Section~\ref{sec:problem-definition}).
-The static offline version of~the~problem, i.e., a problem variant where
-migration is not allowed, and all requests are known in advance is known as the
-\emph{$\ell$-balanced graph partitioning problem}. The problem is 
+The static offline version of~the~problem, called the
+\emph{$\ell$-balanced graph partitioning problem} is 
 NP-complete, and cannot even be approximated within any finite factor unless P
@@
 \maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
+\maciek{ every algorithm must colocate every communicating pair.}
 
@@
 This invariant ensures that ON does not pay too much while reaching $P_F$.
-
-\maciek{We should figure out a better name than "distance".
-We usually apply two arguments to distance.
-Examples of names that can be applied to only one configuration as a parameter: radius, span, stretch, spread, width.
-Also: difference, disparity, distinction, alternation}
 

### 2020-02-13 paper/paper.tex

@@
 
+\subsection{Perfect Partition Model}\label{sec:perfectPartition}
 \maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
 \maciek{ every algorithm must colocate every communicating pair.}
+
+In this paper we mostly focus on a restricted variant of   \PPOBRP{}
+where the communication graph can be partitioned into $\ell$ clusters without any inter-cluster edges.
+Moreover, we assume an optimal offline algorithm (OPT) moves to this \emph{perfect partition}
+at the beginning and stays there permanently.
+Then the task of an online algorithm is to "recover" (or learn) the perfect partition while not paying too much more than OPT.
 

### 2020-02-13 paper/paper.tex
section title for k=3 case

@@
 
-\subsection{$O(\ell)$-competitive algorithm for $k=3$}
+%\subsection{$O(\ell)$-competitive algorithm for $k=3$}
+\subsection{A Simpler Algorithm for $k=3$}
+In this section,
+we give an alternative algorithm for the special case of $k=3$.
+The algorithm does not  improve upon PPL (asymptotically),
+but we will use it in Section \ref{sec:generalModel} for the same special case in the general model.
 
 Let \emph{GREEDY} be the algorithm that on each inter-cluster request
- moves to arbitrary perfect partition \maciek{of input} closest (i.e. reachable in minimum number of node swaps) to the current one.
+ moves to a perfect partition
+% \maciek{of input}
+closest to the current partition,
+i.e.,
+reachable in minimum number of node swaps (ties broken arbitrarily).
 
@@
   %\maciek{GREEDY = find the cheapest (closest to the current) feasible partition}
-  \maciek{We need to add to the assumption that the input has a perfect partition}
+  \maciek{We need to add to the assumption that the input has a perfect partition
+  \mahmoud{the section now is only about PP model}}
   \label{rebalancing-cost}
@@
   Algorithm \emph{GREEDY} is $O(\ell)$-competitive when $k=3$.
-  \maciek{Assumming the perfect partition of input sequence}
+  \maciek{Assumming the perfect partition of input sequence
+  \mahmoud{the section now is only about PP model}}
 \end{theorem}

### 2020-02-13 paper/paper.tex
any ON must collocate

@@
 \maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
-\maciek{ every algorithm must colocate every communicating pair.}
-
+\maciek{ every algorithm must colocate every communicating pair.
+	\mahmoud{updated!}}
 In this paper we mostly focus on a restricted variant of   \PPOBRP{}
@@
 Then the task of an online algorithm is to "recover" (or learn) the perfect partition while not paying too much more than OPT.
+This implies that on a request between any two non-collocated nodes,
+any online algorithm must eventually collocate them,
+as otherwise it not competitive under a sequence that repeats the same request arbitrarily many times.
 

### 2020-02-13 paper/paper.tex
emails
Bring back the related work
k=3 sketch

@@
 \mathchardef\mhyphen="2D
-\newcommand{\OPT}{\mathit{OPT}}
-\newcommand{\ALG}{ALG}
-\newcommand{\OBRP}{BRP}
-\newcommand{\PPOBRP}{PP-BRP}
 
@@
 \author{Maciej Pacut}
+\email{maciek.pacut@gmail.com}
 \orcid{0000-0002-6379-1490}
@@
 \author{Mahmoud Parham} 
+\email{mahmoud.parham@univie.ac.at}
+\orcid{0000-0002-6211-077X}
 \affiliation{%
@@
 \author{Stefan Schmid} 
+\email{stefan_schmid@univie.ac.at}
 \affiliation{%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%&&
+
+\newcommand{\OPT}{\mathit{OPT}}
+\newcommand{\ALG}{ALG}
+\newcommand{\OBRP}{BRP}
+\newcommand{\PPOBRP}{PP-BRP}
+\newcommand{\TAlg}{{\ensuremath{ALG_{k=3}}}\xspace} % we should change this name
 
@@
   In this paper, we provide the lower bound of $\Omega(k\cdot \ell)$.
-  %Furthermore, we provide a $O(k\cdot \ell)$-competitive algorithm for the scenario, where the perfect partition of input exists.
+
+  We also present the optimal algorithm for $k=3$.
+  Finally, we provide a $O(k\cdot \ell)$-competitive algorithm for the scenario, where the perfect partition of input exists.
 \end{abstract}
@@
 
-\section{Introduction and Related Work}
+\section{Introduction}
 
-%We study (...) in the scenario where virtual machines can be migrated during runtime to another physical machine.
-%The possibility of migration allows reacting to unpredictable communication patterns.
-%If some distant nodes communicate often, it is vital to reduce their distance to save network bandwidth.
-%The objective is to~minimize the total network bandwidth used for communication and for migration.
-
-Main technical contribution of this paper is the lower bound $\Omega(k\cdot\ell)$ for the competitive ratio of any deterministic algorithm for Online Balanced Partition problem.
+Main technical contribution of this paper is the lower bound $\Omega(k\cdot\ell)$ for the competitive ratio of any deterministic algorithm for Online Balanced Partition problem (cf.
+Section~\ref{sec:lowerbound}).
 The best known lower bound so far was $\Omega(k)$ \cite{repartition-disc}
@@
 
-So far, the competitive ratio of the best known algorithm was $O(k^2\cdot \ell^2)$ \cite{repartition-disc}.
-We present a $O(k\cdot \ell)$-competitive algorithm for the restricted scenario
+So far, the competitive ratio of the best known algorithm was the component-based algorithm with the ratio $O(k^2\cdot \ell^2)$ \cite{repartition-disc}.
+We present an optimal algorithm for $k=3$ in Section~\ref{sec:k3}.
+In Section~\ref{sec:ppl}, we present a $O(k\cdot \ell)$-competitive algorithm for the restricted scenario
 where the communication graph (i.e., request sequence) admits a \emph{perfect partition} (i.e., a multi-way cut of cost zero) that is also an optimal partition for the offline problem.
@@
 
+\section{Related Work}
+
 The problem was studied under resource augmentation \cite{repartition-disc}, and under stochastic assumptions of input\maciek{cite}.\maciek{cite other papers that cite DISC?}
+
+The model is related to online
+caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}. In our
+setting, requests can be served remotely, which is
+reminiscent of caching models \emph{with
+bypassing}~\cite{EpImLN11,EpImLN15,Irani02}.
 
@@
 the currently best approximation ratio is $O(\log n)$~\cite{SarVaz95,ArKaKa99,FeKrNi00,FeiKra02,KraFei06,Raec08}.
-
-The model is related to online
-caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}. In our
-setting, requests can be served remotely, which is
-reminiscent of caching models \emph{with
-bypassing}~\cite{EpImLN11,EpImLN15,Irani02}.
 
@@
 
-\subsection{Perfect Partition Model}\label{sec:perfectPartition}
-\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
-\maciek{ every algorithm must colocate every communicating pair.
-	\mahmoud{updated!}}
-In this paper we mostly focus on a restricted variant of   \PPOBRP{}
-where the communication graph can be partitioned into $\ell$ clusters without any inter-cluster edges.
-Moreover, we assume an optimal offline algorithm (OPT) moves to this \emph{perfect partition}
-at the beginning and stays there permanently.
-Then the task of an online algorithm is to "recover" (or learn) the perfect partition while not paying too much more than OPT.
-This implies that on a request between any two non-collocated nodes,
-any online algorithm must eventually collocate them,
-as otherwise it not competitive under a sequence that repeats the same request arbitrarily many times.
-
 \section{Lower bound} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
+\label{sec:lowerbound}
 
@@
 This complements an upper bound of $O(k \cdot \ell)$
-for \PPOBRP{} that we present in Section \ref{sec:upperbound}.
+for \PPOBRP{} that we present in Section \ref{sec:ppl}.
 
@@
 
-\section{Recovering a Perfect Partition}	\label{sec:upperbound}
+\section{An Optimal Algorithm for $k=3$}
+
+In this section we provide an optimal algorithm for the case $k=3$.
+
+We define the algorithm \TAlg in the following way.
+\TAlg maintains the components for each node.
+Intially, each node is in its own singleton component.
+Upon receiving a request $(u,v)$, \TAlg merges the components of $u$ and $v$, and moves to the closest partition, where it each component has all of its nodes in one server.
+If no such partition exists, \TAlg resets all its components and finishes the phase.
+
+\begin{lemma}
+  Every repartition of \TAlg has the cost $O(1)$.
+\end{lemma}
+
+\begin{theorem}
+  In a single phase, \TAlg pays $O(\ell)$.
+\end{theorem}
+
+\begin{theorem}
+  \TAlg is $O(\ell)$-competitive.
+\end{theorem}
+\begin{proof}
+  OPT pays at least $1$ in a single phase.
+\end{proof}
+
+\section{Recovering a Perfect Partition}
+\label{sec:ppl}
+\subsection{Perfect Partition Model}\label{sec:perfectPartition}
+\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
+\maciek{ every algorithm must colocate every communicating pair.
+	\mahmoud{updated!}}
+In this paper we mostly focus on a restricted variant of   \PPOBRP{}
+where the communication graph can be partitioned into $\ell$ clusters without any inter-cluster edges.
+Moreover, we assume an optimal offline algorithm (OPT) moves to this \emph{perfect partition}
+at the beginning and stays there permanently.
+Then the task of an online algorithm is to "recover" (or learn) the perfect partition while not paying too much more than OPT.
+This implies that on a request between any two non-collocated nodes,
+any online algorithm must eventually collocate them,
+as otherwise it not competitive under a sequence that repeats the same request arbitrarily many times.
+
+\maciek{Note that the algorithm from Section~\ref{sec:k3} is $O(\ell)$-competitive for the Perfect Partition Model, as during the first phase, a perfect partition of input exists.}
+
+\subsection{An $O(k\cdot \ell)$-competitive Algorithm for Perfect Partition}
 
@@
 \begin{theorem}	\label{thm:upperbound}
-    Algorithm \ref{alg:ppl} reaches the final partition $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
+    PPL reaches the final partition $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
 \end{theorem}
@@
 
-%\subsection{$O(\ell)$-competitive algorithm for $k=3$}
-\subsection{A Simpler Algorithm for $k=3$}
+\section{Corollary and Future Directions}
+
+The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\cdot\ell^2)$.
+
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{references}
+
+\begin{appendix}
+
+\section{Todos}
+
+\todo{ACM format for review? With line numbers etc.}
+
+\todo{Make ORCID visible in ACM style: "Right now we do not typeset ORCIDs" (Source: ACM Style Docs)}
+
+\todo{Include e-mails}
+
+\todo{Check ACM package whitelist}
+
+\todo{Replace package algorithmpseudocode, it is not whitelisted https://www.acm.org/publications/taps/whitelist-of-latex-packages}
+
+\todo{Make an ACM account for the submission}
+
+\todo{Sum up all the steps needed for the submission}
+
+\todo{ACM keywords are mandatory}
+
+\todo{CCS concepts are mandatory}
+
+\todo{Check grammar}
+
+\todo{Font for OPT, ALG, DET, dist etc}
+
+	\section{Ommited results}
+	\begin{enumerate}
+		 \item The lower bound of $\Omega(\ell)$ for the scenario with at most $(1/3-1/k)$-augmentation
+		 \item Costly cascade example (it is easy to present $k^2$ and $k\cdot \sqrt{k}$ (with different approach), and then we can combine these to obtain $k^3$)
+		 \item $k$-competitive algorithm for the ring communication pattern
+		 \item $O(\ell)$-competitive algorithm for $k=4$ and $k=5$
+	\end{enumerate}
+
+  \section{An old section about $k=3$}
+
+  \label{sec:k3}
   In this section,
@@
   but we will use it in Section \ref{sec:generalModel} for the same special case in the general model.
+
+  \maciek{We must introduce phases in the description of this algorithm. If no PP exists, the algorithm resets its components.}
 
@@
 
-\section{Corollary and Future Directions}
-
-The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\cdot\ell^2)$.
-
-\section{Todos}
-
-\todo{ACM format for review? With line numbers etc.}
-
-\todo{Make ORCID visible in ACM style: "Right now we do not typeset ORCIDs" (Source: ACM Style Docs)}
-
-\todo{Include e-mails}
-
-\todo{Check ACM package whitelist}
-
-\todo{Replace package algorithmpseudocode, it is not whitelisted https://www.acm.org/publications/taps/whitelist-of-latex-packages}
-
-\todo{Make an ACM account for the submission}
-
-\todo{Sum up all the steps needed for the submission}
-
-\todo{ACM keywords are mandatory}
-
-\todo{CCS concepts are mandatory}
-
-\todo{Check grammar}
-
-\todo{Font for OPT, ALG, DET, dist etc}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{references}
-
-\begin{appendix}
-	\section{Ommited results}
-	\begin{enumerate}
-		 \item The lower bound of $\Omega(\ell)$ for the scenario with at most $(1/3-1/k)$-augmentation
-		 \item Costly cascade example (it is easy to present $k^2$ and $k\cdot \sqrt{k}$ (with different approach), and then we can combine these to obtain $k^3$)
-		 \item $k$-competitive algorithm for the ring communication pattern
-		 \item $O(\ell)$-competitive algorithm for $k=4$ and $k=5$
-	\end{enumerate}
-
 \end{appendix}

### 2020-02-13 paper/paper.tex
Confusion about k=3

@@
 
+\maciek{Define "component-respecting" algorithm}
+
 \begin{lemma}
-  Every repartition of \TAlg has the cost $O(1)$.
+  \label{lem:1req}
+  Serving each request incurs the cost at most $3\cdot\alpha$ for \TAlg.
 \end{lemma}
 
-\begin{theorem}
-  In a single phase, \TAlg pays $O(\ell)$.
-\end{theorem}
+\begin{proof}
+  If no partition of \TAlg's components exist, then it performs no repartition, and it pays nothing.
+  Hence, in the following we bound the cost of moving to component-respecting partition.
+
+  We distinguish among three configuration types of any algorithm's cluster: $C_1, C_2, C_3$, which we define as follows.
+In $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+In $C_2$ it contains one component of size $2$ and one component of size $1$.
+Finally, in $C_3$ it contains one component of size $3$.
+
+    Consider a request $(u, v)$ and let $U, V$ be clusters of these nodes.
+    If $U=V$, then the request does not require a reconfiguration and GREEDY incurs no cost.
+    Otherwise, we consider cases upon the type of clusters $U$ and $V$.
+    Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise we encounter a component of size $4$ and this would contradict the case assumption that the component-respecting partition exists.
+    If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
+
+    Finally, we consider the case where both $U$ and $V$ are $C_2$.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component-respecting partition exists.
+    If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
+    Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise the input graph would not have a perfect partition.
+In this case GREEDY exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost of $2\cdot \alpha$.
+\end{proof}
 
@@
 \begin{proof}
-  OPT pays at least $1$ in a single phase.
+  Consider a request $(u,v)$.
+  If $u$ and $v$ are collocated in the same server, then \TAlg pays nothing.
+
+  \TAlg collocates nodes belonging to one component, hence during a single phase, we have at most $2\cdot \ell$ requests to non-collocated nodes.
+  We use Lemma~\ref{lem:1req} to bound the cost of each such request, and therefore \TAlg pays $6\cdot\alpha\cdot \ell$ in a single phase.
+  On the other hand, during each phase, OPT pays at least $1$.
 \end{proof}

### 2020-02-13 paper/paper.tex
k=3 to the appendix

@@
 
-\section{An Optimal Algorithm for $k=3$}
-
-In this section we provide an optimal algorithm for the case $k=3$.
-
-We define the algorithm \TAlg in the following way.
-\TAlg maintains the components for each node.
-Intially, each node is in its own singleton component.
-Upon receiving a request $(u,v)$, \TAlg merges the components of $u$ and $v$, and moves to the closest partition, where it each component has all of its nodes in one server.
-If no such partition exists, \TAlg resets all its components and finishes the phase.
-
-\maciek{Define "component-respecting" algorithm}
-
-\begin{lemma}
-  \label{lem:1req}
-  Serving each request incurs the cost at most $3\cdot\alpha$ for \TAlg.
-\end{lemma}
-
-\begin{proof}
-  If no partition of \TAlg's components exist, then it performs no repartition, and it pays nothing.
-  Hence, in the following we bound the cost of moving to component-respecting partition.
-
-  We distinguish among three configuration types of any algorithm's cluster: $C_1, C_2, C_3$, which we define as follows.
-In $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
-In $C_2$ it contains one component of size $2$ and one component of size $1$.
-Finally, in $C_3$ it contains one component of size $3$.
-
-    Consider a request $(u, v)$ and let $U, V$ be clusters of these nodes.
-    If $U=V$, then the request does not require a reconfiguration and GREEDY incurs no cost.
-    Otherwise, we consider cases upon the type of clusters $U$ and $V$.
-    Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise we encounter a component of size $4$ and this would contradict the case assumption that the component-respecting partition exists.
-    If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
-
-    Finally, we consider the case where both $U$ and $V$ are $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component-respecting partition exists.
-    If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
-    Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise the input graph would not have a perfect partition.
-In this case GREEDY exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost of $2\cdot \alpha$.
-\end{proof}
-
-\begin{theorem}
-  \TAlg is $O(\ell)$-competitive.
-\end{theorem}
-\begin{proof}
-  Consider a request $(u,v)$.
-  If $u$ and $v$ are collocated in the same server, then \TAlg pays nothing.
-
-  \TAlg collocates nodes belonging to one component, hence during a single phase, we have at most $2\cdot \ell$ requests to non-collocated nodes.
-  We use Lemma~\ref{lem:1req} to bound the cost of each such request, and therefore \TAlg pays $6\cdot\alpha\cdot \ell$ in a single phase.
-  On the other hand, during each phase, OPT pays at least $1$.
-\end{proof}
-
 \section{Recovering a Perfect Partition}
@@
 
+  \section{An Optimal Algorithm for $k=3$ (most likely wrong)}
+
+  In this section we provide an optimal algorithm for the case $k=3$.
+
+  We define the algorithm \TAlg in the following way.
+  \TAlg maintains components for each node.
+  Intially, each node is in its own singleton component.
+  Upon receiving a request $(u,v)$, \TAlg merges the components of $u$ and $v$, and moves to the closest partition, where it each component has all of its nodes in one server.
+
+  \maciek{Component reseting magic}
+  Now we define what \TAlg does when no such partition exists.
+  If the resulting component (of $u$ and $v$) has size $4$, then \TAlg resets all internal requests that concern this component.
+  Otherwise, \TAlg resets all its components and ends the phase.
+
+  \maciek{Define "component-respecting" algorithm}
+
+  \begin{lemma}
+    \label{lem:1req}
+    Serving each request incurs the cost at most $3\cdot\alpha$ for \TAlg.
+  \end{lemma}
+
+  \begin{proof}
+    If no partition of \TAlg's components exist, then it performs no repartition, and it pays nothing.
+    Hence, in the following we bound the cost of moving to component-respecting partition.
+
+    We distinguish among three configuration types of any algorithm's cluster: $C_1, C_2, C_3$, which we define as follows.
+In $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+In $C_2$ it contains one component of size $2$ and one component of size $1$.
+Finally, in $C_3$ it contains one component of size $3$.
+
+      Consider a request $(u, v)$ and let $U, V$ be clusters of these nodes.
+      If $U=V$, then the request does not require a reconfiguration and GREEDY incurs no cost.
+      Otherwise, we consider cases upon the type of clusters $U$ and $V$.
+      Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise we encounter a component of size $4$ and this would contradict the case assumption that the component-respecting partition exists.
+      If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
+
+      Finally, we consider the case where both $U$ and $V$ are $C_2$.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component-respecting partition exists.
+      If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
+      Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+      Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise the input graph would not have a perfect partition.
+In this case GREEDY exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost of $2\cdot \alpha$.
+  \end{proof}
+
+  \begin{theorem}
+    \TAlg is $O(\ell)$-competitive.
+  \end{theorem}
+  \begin{proof}
+    Consider a request $(u,v)$.
+    If $u$ and $v$ are collocated in the same server, then \TAlg pays nothing.
+
+    \TAlg collocates nodes belonging to one component, hence during a single phase, we have at most $2\cdot \ell$ requests to non-collocated nodes.
+    We use Lemma~\ref{lem:1req} to bound the cost of each such request, and therefore \TAlg pays $6\cdot\alpha\cdot \ell$ in a single phase.
+    On the other hand, during each phase, OPT pays at least $1$.
+  \end{proof}
+
   \section{An old section about $k=3$}

### 2020-02-14 paper/paper.tex
remiving unneeded section header

@@
 
-\section{Recovering a Perfect Partition}
-\label{sec:ppl}
-\subsection{Perfect Partition Model}\label{sec:perfectPartition}
+\section{Recovering a Perfect Partition}	\label{sec:ppl}
+%\subsection{Perfect Partition Model}\label{sec:perfectPartition}
 \maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
@@
 	\mahmoud{updated!}}
-In this paper we mostly focus on a restricted variant of   \PPOBRP{}
+We  focus on a restricted variant of   \PPOBRP{}
 where the communication graph can be partitioned into $\ell$ clusters without any inter-cluster edges.

### 2020-02-14 paper/paper.tex
explin what split means

@@
 Recall that internal requests does not incur any cost for \ALG.
-If \ALG{}at any point splits a component\maciek{define splitting}, then
-we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
+If \ALG{} at any point splits a component
+(i.e., spreads its nodes over two or more clusters),
+then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
 If \ALG{} never collocates them then it is not competitive.
@@
     That is,
-    $\mathit{dist}(P, P_I) = \sum_{j=1}^{\ell} | C_j \setminus I_j |$. 
+    $\mathit{dist}(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$. 
 In other words,

### 2020-02-14 paper/paper.tex
minor

@@
 \maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
-Fix any online algorithm \ALG.
+Fix any online algorithm \ALG{}.
 We construct the input sequence of requests for \ALG{} as follows.
@@
 nodes on the cluster $S_1$.
-Recall that internal requests does not incur any cost for \ALG.
+Recall that internal requests does not incur any cost for \ALG{}.
 If \ALG{} at any point splits a component
@@
 Since this is an external request,
-ALG joins them into one component and collocates them in some cluster other than $S_1$.
+\ALG{} joins them into one component and collocates them in some cluster other than $S_1$.
 For this,
-ALG moves to a new configuration (i.e., partition)
+\ALG{} moves to a new configuration (i.e., partition)
 that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
@@
 we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
-ALG must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
+\ALG{} must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
 consequently replaces $x_1$ with some other (single) node $x_2$.
@@
 OPT pays $4$ for node migrations and
-ALG pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
+\ALG{} pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
 We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,

### 2020-02-14 paper/paper.tex
k=3 algorithm description

--- 2020-02-14 paper/paper.tex
@@
   Intially, each node is in its own singleton component.
-  Upon receiving a request $(u,v)$, \TAlg merges the components of $u$ and $v$, and moves to the closest partition, where it each component has all of its nodes in one server.
+  For each pair of nodes \TAlg maintains a counter, initiated $0$, and upon receiving a request to this pair while its nodes are located on different servers, it increases the counter by $1$.
+  If a counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest partition, where it each component has all of its nodes in one server.
+  If no such partition exists, \TAlg resets all components to singletons and all counters to $0$.
 
-  \maciek{Component reseting magic}
-  Now we define what \TAlg does when no such partition exists.
-  If the resulting component (of $u$ and $v$) has size $4$, then \TAlg resets all internal requests that concern this component.
-  Otherwise, \TAlg resets all its components and ends the phase.
+  %\maciek{Shouldn't we think about whole component cuts instead of single edge? Hmm. NO!}
 
-  \maciek{Define "component-respecting" algorithm}
+  We say that \TAlg is \emph{component respecting}, i.e., it maintaints the invariant that nodes of each component are located at the same server.
 

### 2020-02-14 paper/paper.tex
swapping sections because notions of PP and components must come first
define component and component-respecting

@@
 
-\section{Lower bound} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
-\label{sec:lowerbound}
-
-In this section we show a lower bound of $\Omega(k \cdot \ell)$.
-The constructed input sequence allows OPT to move to a perfect partition, and hence
-the construction also constitutes a lower bound for \PPOBRP{}.
-This complements an upper bound of $O(k \cdot \ell)$
-for \PPOBRP{} that we present in Section \ref{sec:ppl}.
-
-\begin{theorem}
-  The competitive ratio of any deterministic algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$.
-\end{theorem}
-
-\begin{proof}
-We construct an instance of the problem with $\ell$ clusters 
-$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
-Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
-\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
-Fix any online algorithm \ALG{}.
-We construct the input sequence of requests for \ALG{} as follows.
-First,
-we issue $k-2$ (internal) requests so that \ALG{} can form a component of $k-1$
-nodes on the cluster $S_1$.
-Recall that internal requests does not incur any cost for \ALG{}.
-If \ALG{} at any point splits a component
-(i.e., spreads its nodes over two or more clusters),
-then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
-If \ALG{} never collocates them then it is not competitive.
-
-Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
-Next,
-we issue the request between $x_0$ and $y_0$.
-Since this is an external request,
-\ALG{} joins them into one component and collocates them in some cluster other than $S_1$.
-For this,
-\ALG{} moves to a new configuration (i.e., partition)
-that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
-
-Next,
-we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
-\ALG{} must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
-consequently replaces $x_1$ with some other (single) node $x_2$.
-We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
- until there are only two single nodes left that  originate from the same cluster.
-I.e.,
-two single nodes $x^*, y^*,I(x^*) = I(y^*)$ s.t.~for any other pair of single nodes
-$x'$ and $y'$,
-we have $I(x') \neq I(y')$.
-At this point there are at most $\ell+1$ single nodes left,
-otherwise there would be more pairs of single nodes that were initially in the same cluster.
-
-Given this sequence of requests,
-the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
- swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
-Hence,
-OPT pays $4$ for node migrations and
-\ALG{} pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
-We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
-therefore $|X| \geq k \cdot \ell - k - \ell \in \Omega(k\cdot\ell)$.
-\end{proof}
-
 \section{Recovering a Perfect Partition}	\label{sec:ppl}
@@
 as otherwise it not competitive under a sequence that repeats the same request arbitrarily many times.
-
-\maciek{Note that the algorithm from Section~\ref{sec:k3} is $O(\ell)$-competitive for the Perfect Partition Model, as during the first phase, a perfect partition of input exists.}
+Hence,
+any competitive algorithm,
+must at some point collocate all nodes of a  connected sub-graph of the communication graph.
+We refer to this connected sub-graph as a \emph{component}.
+An algorithm that always keeps a component collocated is \emph{component-respecting}.
+%\maciek{Note that the algorithm from Section~\ref{sec:k3} is $O(\ell)$-competitive for the Perfect Partition Model, as during the first phase, a perfect partition of input exists.}
 
@@
 %    	So it is 4-competitive (?)}
+ \end{proof}
+
+\section{Lower bound} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
+\label{sec:lowerbound}
+
+In this section we show a lower bound of $\Omega(k \cdot \ell)$.
+The constructed input sequence allows OPT to move to a perfect partition, and hence
+the construction also constitutes a lower bound for \PPOBRP{}.
+This complements an upper bound of $O(k \cdot \ell)$
+for \PPOBRP{} that we present in Section \ref{sec:ppl}.
+
+\begin{theorem}
+	The competitive ratio of any deterministic algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$.
+\end{theorem}
+
+\begin{proof}
+	We construct an instance of the problem with $\ell$ clusters 
+	$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
+	Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
+	\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
+	Fix any online algorithm \ALG{}.
+	We construct the input sequence of requests for \ALG{} as follows.
+	First,
+	we issue $k-2$ (internal) requests so that \ALG{} can form a component of $k-1$
+	nodes on the cluster $S_1$.
+	Recall that internal requests does not incur any cost for \ALG{}.
+	If \ALG{} at any point splits a component
+	(i.e., spreads its nodes over two or more clusters),
+	then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
+	If \ALG{} never collocates them then it is not competitive.
+
+	Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
+	Next,
+	we issue the request between $x_0$ and $y_0$.
+	Since this is an external request,
+	\ALG{} joins them into one component and collocates them in some cluster other than $S_1$.
+	For this,
+	\ALG{} moves to a new configuration (i.e., partition)
+	that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
+
+	Next,
+	we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
+	\ALG{} must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
+	consequently replaces $x_1$ with some other (single) node $x_2$.
+	We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
+	until there are only two single nodes left that  originate from the same cluster.
+	I.e.,
+	two single nodes $x^*, y^*,I(x^*) = I(y^*)$ s.t.~for any other pair of single nodes
+	$x'$ and $y'$,
+	we have $I(x') \neq I(y')$.
+	At this point there are at most $\ell+1$ single nodes left,
+	otherwise there would be more pairs of single nodes that were initially in the same cluster.
+
+	Given this sequence of requests,
+	the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
+	swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
+	Hence,
+	OPT pays $4$ for node migrations and
+	\ALG{} pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
+	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
+	therefore $|X| \geq k \cdot \ell - k - \ell \in \Omega(k\cdot\ell)$.
 \end{proof}

### 2020-02-14 paper/paper.tex
introduce \PPOBRP{}

@@
 	\mahmoud{updated!}}
-We  focus on a restricted variant of   \PPOBRP{}
+We  focus on a restricted variant of  \OBRP{},
 where the communication graph can be partitioned into $\ell$ clusters without any inter-cluster edges.
@@
 at the beginning and stays there permanently.
-Then the task of an online algorithm is to "recover" (or learn) the perfect partition while not paying too much more than OPT.
+We refer this variant as \PPOBRP{}.
+The task of an online algorithm for \PPOBRP{} is to "recover" (or learn) the perfect partition while not paying too much more than OPT.
 This implies that on a request between any two non-collocated nodes,
@@
 The constructed input sequence allows OPT to move to a perfect partition, and hence
-the construction also constitutes a lower bound for \PPOBRP{}.
+the construction also constitutes a lower bound for \OBRP{}.
 This complements an upper bound of $O(k \cdot \ell)$

### 2020-02-14 paper/paper.tex
hiding comments

@@
 %\subsection{Perfect Partition Model}\label{sec:perfectPartition}
-\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
-\maciek{ every algorithm must colocate every communicating pair.
-	\mahmoud{updated!}}
+%\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
+%\maciek{ every algorithm must colocate every communicating pair.
+%	\mahmoud{updated!}}
 We  focus on a restricted variant of  \OBRP{},

### 2020-02-14 paper/paper.tex
undoing a mistake

@@
 
-In this section we show a lower bound of $\Omega(k \cdot \ell)$.
+In this section we show a lower bound of $\Omega(k \cdot \ell)$ for \OBRP{}.
 The constructed input sequence allows OPT to move to a perfect partition, and hence
-the construction also constitutes a lower bound for \OBRP{}.
+the construction also constitutes a lower bound for \PPOBRP{}.
 This complements an upper bound of $O(k \cdot \ell)$

### 2020-02-14 paper/paper.tex
hiding comment
fixing \TAlg

@@
 \usepackage[utf8]{inputenc}
-\usepackage{xspace}
+%\usepackage{xspace}
 \usepackage{balance}
@@
 \newcommand{\PPOBRP}{PP-BRP}
-\newcommand{\TAlg}{{\ensuremath{ALG_{k=3}}}\xspace} % we should change this name
+\newcommand{\TAlg}{{\ensuremath{ALG_{k=3}}}} % we should change this name
 
@@
 	Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
-	\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
+%	\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
 	Fix any online algorithm \ALG{}.
@@
 
-  We define the algorithm \TAlg in the following way.
-  \TAlg maintains components for each node.
+  We define the algorithm \TAlg{} in the following way.
+  \TAlg{} maintains components for each node.
   Intially, each node is in its own singleton component.
-  For each pair of nodes \TAlg maintains a counter, initiated $0$, and upon receiving a request to this pair while its nodes are located on different servers, it increases the counter by $1$.
-  If a counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest partition, where it each component has all of its nodes in one server.
-  If no such partition exists, \TAlg resets all components to singletons and all counters to $0$.
+  For each pair of nodes \TAlg{} maintains a counter, initiated $0$, and upon receiving a request to this pair while its nodes are located on different servers, it increases the counter by $1$.
+  If a counter for a pair $(u,v)$ reaches $\alpha$, \TAlg{} merges the components of $u$ and $v$, and moves to the closest partition, where it each component has all of its nodes in one server.
+  If no such partition exists, \TAlg{} resets all components to singletons and all counters to $0$.
 
@@
 
-  We say that \TAlg is \emph{component respecting}, i.e., it maintaints the invariant that nodes of each component are located at the same server.
+  We say that \TAlg{} is \emph{component respecting}, i.e., it maintaints the invariant that nodes of each component are located at the same server.
 
@@
       Finally, we consider the case where both $U$ and $V$ are $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component-respecting partition exists.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg{} has the component of size $4$, a contradiction with the case assumption that the component-respecting partition exists.
       If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
@@
   \begin{theorem}
-    \TAlg is $O(\ell)$-competitive.
+    \TAlg{} is $O(\ell)$-competitive.
   \end{theorem}
@@
     Consider a request $(u,v)$.
-    If $u$ and $v$ are collocated in the same server, then \TAlg pays nothing.
+    If $u$ and $v$ are collocated in the same server, then \TAlg{} pays nothing.
 
-    \TAlg collocates nodes belonging to one component, hence during a single phase, we have at most $2\cdot \ell$ requests to non-collocated nodes.
-    We use Lemma~\ref{lem:1req} to bound the cost of each such request, and therefore \TAlg pays $6\cdot\alpha\cdot \ell$ in a single phase.
+    \TAlg{} collocates nodes belonging to one component, hence during a single phase, we have at most $2\cdot \ell$ requests to non-collocated nodes.
+    We use Lemma~\ref{lem:1req} to bound the cost of each such request, and therefore \TAlg{} pays $6\cdot\alpha\cdot \ell$ in a single phase.
     On the other hand, during each phase, OPT pays at least $1$.

### 2020-02-14 paper/paper.tex
Cost of TAlg

--- 2020-02-14 paper/paper.tex
@@
 
+\section{Lower bound} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
+\label{sec:lowerbound}
+
+In this section we show a lower bound of $\Omega(k \cdot \ell)$.
+The constructed input sequence allows OPT to move to a perfect partition, and hence
+the construction also constitutes a lower bound for \PPOBRP{}.
+This complements an upper bound of $O(k \cdot \ell)$
+for \PPOBRP{} that we present in Section \ref{sec:ppl}.
+
+\begin{theorem}
+  The competitive ratio of any deterministic algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$.
+\end{theorem}
+
+\begin{proof}
+We construct an instance of the problem with $\ell$ clusters 
+$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
+Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
+\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
+Fix any online algorithm \ALG{}.
+We construct the input sequence of requests for \ALG{} as follows.
+First,
+we issue $k-2$ (internal) requests so that \ALG{} can form a component of $k-1$
+nodes on the cluster $S_1$.
+Recall that internal requests does not incur any cost for \ALG{}.
+If \ALG{} at any point splits a component
+(i.e., spreads its nodes over two or more clusters),
+then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
+If \ALG{} never collocates them then it is not competitive.
+
+Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
+Next,
+we issue the request between $x_0$ and $y_0$.
+Since this is an external request,
+\ALG{} joins them into one component and collocates them in some cluster other than $S_1$.
+For this,
+\ALG{} moves to a new configuration (i.e., partition)
+that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
+
+Next,
+we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
+\ALG{} must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
+consequently replaces $x_1$ with some other (single) node $x_2$.
+We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
+ until there are only two single nodes left that  originate from the same cluster.
+I.e.,
+two single nodes $x^*, y^*,I(x^*) = I(y^*)$ s.t.~for any other pair of single nodes
+$x'$ and $y'$,
+we have $I(x') \neq I(y')$.
+At this point there are at most $\ell+1$ single nodes left,
+otherwise there would be more pairs of single nodes that were initially in the same cluster.
+
+Given this sequence of requests,
+the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
+ swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
+Hence,
+OPT pays $4$ for node migrations and
+\ALG{} pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
+We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
+therefore $|X| \geq k \cdot \ell - k - \ell \in \Omega(k\cdot\ell)$.
+\end{proof}
+
 \section{Recovering a Perfect Partition}	\label{sec:ppl}
 %\subsection{Perfect Partition Model}\label{sec:perfectPartition}
-%\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
-%\maciek{ every algorithm must colocate every communicating pair.
-%	\mahmoud{updated!}}
-We  focus on a restricted variant of  \OBRP{},
+\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
+\maciek{ every algorithm must colocate every communicating pair.
+	\mahmoud{updated!}}
+We  focus on a restricted variant of   \PPOBRP{}
 where the communication graph can be partitioned into $\ell$ clusters without any inter-cluster edges.
@@
 at the beginning and stays there permanently.
-We refer this variant as \PPOBRP{}.
-The task of an online algorithm for \PPOBRP{} is to "recover" (or learn) the perfect partition while not paying too much more than OPT.
+Then the task of an online algorithm is to "recover" (or learn) the perfect partition while not paying too much more than OPT.
 This implies that on a request between any two non-collocated nodes,
@@
 as otherwise it not competitive under a sequence that repeats the same request arbitrarily many times.
-Hence,
-any competitive algorithm,
-must at some point collocate all nodes of a  connected sub-graph of the communication graph.
-We refer to this connected sub-graph as a \emph{component}.
-An algorithm that always keeps a component collocated is \emph{component-respecting}.
-%\maciek{Note that the algorithm from Section~\ref{sec:k3} is $O(\ell)$-competitive for the Perfect Partition Model, as during the first phase, a perfect partition of input exists.}
+
+\maciek{Note that the algorithm from Section~\ref{sec:k3} is $O(\ell)$-competitive for the Perfect Partition Model, as during the first phase, a perfect partition of input exists.}
 
@@
 
-\section{Lower bound} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
-\label{sec:lowerbound}
-
-In this section we show a lower bound of $\Omega(k \cdot \ell)$ for \OBRP{}.
-The constructed input sequence allows OPT to move to a perfect partition, and hence
-the construction also constitutes a lower bound for \PPOBRP{}.
-This complements an upper bound of $O(k \cdot \ell)$
-for \PPOBRP{} that we present in Section \ref{sec:ppl}.
-
-\begin{theorem}
-	The competitive ratio of any deterministic algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$.
-\end{theorem}
-
-\begin{proof}
-	We construct an instance of the problem with $\ell$ clusters 
-	$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
-	Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
-	\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
-	Fix any online algorithm \ALG{}.
-	We construct the input sequence of requests for \ALG{} as follows.
-	First,
-	we issue $k-2$ (internal) requests so that \ALG{} can form a component of $k-1$
-	nodes on the cluster $S_1$.
-	Recall that internal requests does not incur any cost for \ALG{}.
-	If \ALG{} at any point splits a component
-	(i.e., spreads its nodes over two or more clusters),
-	then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
-	If \ALG{} never collocates them then it is not competitive.
-
-	Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
-	Next,
-	we issue the request between $x_0$ and $y_0$.
-	Since this is an external request,
-	\ALG{} joins them into one component and collocates them in some cluster other than $S_1$.
-	For this,
-	\ALG{} moves to a new configuration (i.e., partition)
-	that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
-
-	Next,
-	we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
-	\ALG{} must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
-	consequently replaces $x_1$ with some other (single) node $x_2$.
-	We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
-	until there are only two single nodes left that  originate from the same cluster.
-	I.e.,
-	two single nodes $x^*, y^*,I(x^*) = I(y^*)$ s.t.~for any other pair of single nodes
-	$x'$ and $y'$,
-	we have $I(x') \neq I(y')$.
-	At this point there are at most $\ell+1$ single nodes left,
-	otherwise there would be more pairs of single nodes that were initially in the same cluster.
-
-	Given this sequence of requests,
-	the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
-	swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
-	Hence,
-	OPT pays $4$ for node migrations and
-	\ALG{} pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
-	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
-	therefore $|X| \geq k \cdot \ell - k - \ell \in \Omega(k\cdot\ell)$.
-\end{proof}
-
 \section{Corollary and Future Directions}
@@
   We say that \TAlg is \emph{component respecting}, i.e., it maintaints the invariant that nodes of each component are located at the same server.
+  Observe that this implies that \TAlg never increases any counter above $\alpha$.
 
@@
     \label{lem:1req}
-    Serving each request incurs the cost at most $3\cdot\alpha$ for \TAlg.
+    A single repartition incurs the cost at most $3\cdot\alpha$ for \TAlg.
   \end{lemma}
@@
   \begin{proof}
-    If no partition of \TAlg's components exist, then it performs no repartition, and it pays nothing.
+    If no partition of \TAlg's components exist, then it performs no repartition, and serves the request remotely, incurring the cost $1$.
     Hence, in the following we bound the cost of moving to component-respecting partition.
@@
   \begin{proof}
-    Consider a request $(u,v)$.
-    If $u$ and $v$ are collocated in the same server, then \TAlg pays nothing.
+    Fix a phase of \TAlg, and consider the state of its counters at the end of it.
 
-    \TAlg collocates nodes belonging to one component, hence during a single phase, we have at most $2\cdot \ell$ requests to non-collocated nodes.
-    We use Lemma~\ref{lem:1req} to bound the cost of each such request, and therefore \TAlg pays $6\cdot\alpha\cdot \ell$ in a single phase.
-    On the other hand, during each phase, OPT pays at least $1$.
+    Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
+    Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components throughout this phase.
+    Let $A^M$ be the cost of migrations performed by \TAlg in this phase.
+
+    We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
+
+    During a single phase, at most $2\cdot \ell$ component merge operations are performed.
+    Exceeding this number would mean that a component of size $4$ would emerge, a contradiction.
+    Together with Lemma~\ref{lem:1req}, this gives us $A^M \leq 6\cdot\alpha\cdot\ell$.
+
+    Now we bound $A^I$.
+    Each component of size $3$ contributes at most $(3\cdot \alpha - 1)$ to $A^I$, as at most $2$ of its pairs are saturated and contribute $\alpha$ each, and the unsaturated pair contributes at most $\alpha-1$.
+    Each of $\ell$ clusters of \TAlg contain at most one component of size $3$ (other cluster configurations contribute less: $3$ singletons contribute $0$ and a component of size $2$ contributes at most $\alpha$ to $A^I$), and in total, $A^I \leq \ell \cdot (3\cdot \alpha-1)$.
+
+    We bound $A^E$ by $k^2 \cdot (\alpha - 1)$, as no more than $k^2$ pairs are unsaturated, and each of them contributes at most $\alpha -1$.
+
+    In total, the cost of \TAlg is at most $A^I + A^E + A^M$ during the phase.
+
+    Now we lower bound the cost of $\OPT$.
   \end{proof}

### 2020-02-14 paper/paper.tex
Merge by hand

--- 2020-02-14 paper/paper.tex
@@
 
-  We say that \TAlg{} is \emph{component respecting}, i.e., it maintaints the invariant that nodes of each component are located at the same server.
+  We say that \TAlg is \emph{component respecting}, i.e., it maintains the invariant that nodes of each component are located at the same server.
+  Observe that this implies that \TAlg never increases any counter above $\alpha$.
 
@@
   \begin{proof}
-    Consider a request $(u,v)$.
-    If $u$ and $v$ are collocated in the same server, then \TAlg{} pays nothing.
+    Fix a phase of \TAlg, and consider the state of its counters at the end of it.
 
-    \TAlg{} collocates nodes belonging to one component, hence during a single phase, we have at most $2\cdot \ell$ requests to non-collocated nodes.
-    We use Lemma~\ref{lem:1req} to bound the cost of each such request, and therefore \TAlg{} pays $6\cdot\alpha\cdot \ell$ in a single phase.
-    On the other hand, during each phase, OPT pays at least $1$.
+    Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
+    Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components throughout this phase.
+    Let $A^M$ be the cost of migrations performed by \TAlg in this phase.
+
+    We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
+
+    During a single phase, at most $2\cdot \ell$ component merge operations are performed.
+    Exceeding this number would mean that a component of size $4$ would emerge, a contradiction.
+    Together with Lemma~\ref{lem:1req}, this gives us $A^M \leq 6\cdot\alpha\cdot\ell$.
+
+    Now we bound $A^I$.
+    Each component of size $3$ contributes at most $(3\cdot \alpha - 1)$ to $A^I$, as at most $2$ of its pairs are saturated and contribute $\alpha$ each, and the unsaturated pair contributes at most $\alpha-1$.
+    Each of $\ell$ clusters of \TAlg contain at most one component of size $3$ (other cluster configurations contribute less: $3$ singletons contribute $0$ and a component of size $2$ contributes at most $\alpha$ to $A^I$), and in total, $A^I \leq \ell \cdot (3\cdot \alpha-1)$.
+
+    We bound $A^E$ by $k^2 \cdot (\alpha - 1)$, as no more than $k^2$ pairs are unsaturated, and each of them contributes at most $\alpha -1$.
+
+    In total, the cost of \TAlg is at most $A^I + A^E + A^M$ during the phase.
+
+    Now we lower bound the cost of $\OPT$.
   \end{proof}

### 2020-02-14 paper/paper.tex
general editing, partition => configuration

@@
 \usepackage{balance}
-\usepackage{amsmath,amssymb,amsfonts,mathtools,amsthm}
+\usepackage{amsmath,amsfonts,mathtools,amsthm}
 \usepackage[noend]{algpseudocode}
@@
 In Section~\ref{sec:ppl}, we present a $O(k\cdot \ell)$-competitive algorithm for the restricted scenario
-where the communication graph (i.e., request sequence) admits a \emph{perfect partition} (i.e., a multi-way cut of cost zero) that is also an optimal partition for the offline problem.
+where the communication graph (i.e., request sequence) admits a \emph{perfect partition} (i.e., a multi-way cut of cost zero) that is also an optimal configuration for the offline problem.
 For this setting, the algorithm is optimal, as the lower bound utilized an input sequence that has a perfect partition.
@@
 each of size~$k$.
-We say that nodes in $C \subset V$ are \emph{collocated} in a partition $P$
+We say that nodes in $C \subset V$ are \emph{collocated}% in a partition $P$
 if they are assigned to the same cluster.
@@
 Before serving the request,
-the online algorithm
-can re-partition,
+the online algorithm can re-partition,
 i.e.,
@@
 
-We assume OPT begins with the initial partition
+We assume OPT begins with the initial configuration
 $P_I = I_1, \dots, I_{\ell}$ and moves to the final partition
 $P_F = F_1, \dots, F_{\ell}$.
- The \emph{distance} of a partition $P = C_1, \dots, C_{\ell}$ from the initial partition is the number of nodes in $P$ that do not reside in their initial cluster.
+ The \emph{distance} of a configuration $P = C_1, \dots, C_{\ell}$ from the initial configuration is the number of nodes in $P$ that do not reside in their initial cluster.
     That is,
@@
 In other words,
-at least $\mathit{dist}(P, P_I)/2$ node swaps are required in order to reach the partition $P$ from $P_I$.
+at least $\mathit{dist}(P, P_I)/2$ node swaps are required in order to reach the configuration $P$ from $P_I$.
 Therefore,
 $\OPT \geq \Delta:= dist(P_F, P_I) $.
-The online algorithm PPL mimics OPT by minimizing the distance to the initial partition $P_I$ with each re-partitioning.
+ With each re-partitioning,
+  PPL moves to a configuration that minimizes distance to the initial configuration $P_I$.
 As a result,
-ON never ends up with a partition that is more than $\Delta$ away from $P_I$.
+ON never ends up in a configuration that is more than $\Delta$ away from $P_I$.
 This invariant ensures that ON does not pay too much while reaching $P_F$.
@@
         $k, \ell$,
-        initial partition $P_I$,
+        initial configuration $P_I$,
         sequence of  requests $\sigma_1, \dots, \sigma_N$ 
-        \Ensure A final partition $P_F$ 
+        \Ensure A final configuration $P_F$ 
         \State For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$
@@
 
-%      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial configuration}
-      We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current partition $P$ with a (perfect) partition closest to $P_I$.
-Hence it never moves to a partition beyond distance $\Delta$.      
+%      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
+      We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a (perfect) partition closest to $P_I$.
+Hence it never moves to a configuration beyond distance $\Delta$.      
 \begin{property} \label{prop:dist<OPT}
-    Let $P$ be any partition chosen by Algorithm \ref{alg:ppl} at Line $\ref{line:rebalance}$.
+    Let $P$ be any configuration chosen by Algorithm \ref{alg:ppl} at Line $\ref{line:rebalance}$.
     Then, $\mathit{dist}(P,P_I) \leq \Delta$.
@@
 \begin{theorem}	\label{thm:upperbound}
-    PPL reaches the final partition $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
+    PPL reaches the final configuration $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
 \end{theorem}
@@
      That is, 
-     once it reaches $\Delta = \mathit{dist} (P_{t}, P_I)$,
+     once it reaches a configuration $P$ at distance $\Delta = \mathit{dist} (P, P_I)$,
      it does not move to a partition
-     $P_{t'}$, for any $t'>t,\mathit{dist} (P_{t'}, P_I) > \Delta$,
+     $P', \mathit{dist} (P', P_I) > \Delta$,
      before it enumerates all partitions at distance $\Delta$.
      Therefore,
-     PPL eventually reaches $\Delta=\OPT$ and the partition $P_F$.
+     PPL eventually reaches $\Delta=\OPT$ and the configuration $P_F$.
 %    including the request that completes revealing of all components that are collocated in $P_F$.
@@
 \begin{proof}
+	\mahmoud{mention $\alpha=1$}
 	We construct an instance of the problem with $\ell$ clusters 
@@
 	For this,
-	\ALG{} moves to a new configuration (i.e., partition)
+	\ALG{} moves to a new configuration
 	that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
@@
 
-    We distinguish among three configuration types of any algorithm's cluster: $C_1, C_2, C_3$, which we define as follows.
-In $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
-In $C_2$ it contains one component of size $2$ and one component of size $1$.
-Finally, in $C_3$ it contains one component of size $3$.
+    We distinguish between three types of clusters: $C_1, C_2, C_3$,
+    which we define as follows.
+    A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+    A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
+    Finally, a cluster of type $C_3$  contains one component of size $3$.
 

### 2020-02-14 paper/paper.tex
lower bound proof independent from \alpha

@@
      once it reaches a configuration $P$ at distance $\Delta = \mathit{dist} (P, P_I)$,
-     it does not move to a partition
+     it does not move to a configuration
      $P', \mathit{dist} (P', P_I) > \Delta$,
-     before it enumerates all partitions at distance $\Delta$.
+     before it enumerates all configurations at distance $\Delta$.
      Therefore,
@@
 \begin{proof}
-	\mahmoud{mention $\alpha=1$}
 	We construct an instance of the problem with $\ell$ clusters 
@@
 	Hence,
-	OPT pays $4$ for node migrations and
-	\ALG{} pays at least one for each node in the sequence $X := x_0, x_1,\dots$.
+	OPT pays for $4$ node migrations and
+	\ALG{} incurs at least one migration for each node in the sequence $X := x_0, x_1,\dots$.
 	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,

### 2020-02-14 paper/paper.tex
move PP def to Section 3

@@
 
-\section{Recovering a Perfect Partition}	\label{sec:ppl}
-%\subsection{Perfect Partition Model}\label{sec:perfectPartition}
-%\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
-%\maciek{ every algorithm must colocate every communicating pair.
-%	\mahmoud{updated!}}
-We  focus on a restricted variant of  \OBRP{},
+\subsection{The Perfect Partition Variant}	\label{sec:PP}
+In Section \ref{sec:ppl},
+we give an algorithm for a restricted variant of  \OBRP{},
 where the communication graph can be partitioned into $\ell$ clusters without any inter-cluster edges.
@@
 An algorithm that always keeps a component collocated is \emph{component-respecting}.
+
+\section{Recovering a Perfect Partition}	\label{sec:ppl}
+%\subsection{Perfect Partition Model}\label{sec:perfectPartition}
+%\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
+%\maciek{ every algorithm must colocate every communicating pair.
+%	\mahmoud{updated!}}
 %\maciek{Note that the algorithm from Section~\ref{sec:k3} is $O(\ell)$-competitive for the Perfect Partition Model, as during the first phase, a perfect partition of input exists.}
 
-\subsection{An $O(k\cdot \ell)$-competitive Algorithm for Perfect Partition}
+%\subsection{An $O(k\cdot \ell)$-competitive Algorithm for Perfect Partition}
 

### 2020-02-14 paper/paper.tex
minor
editing the abstract

@@
 \begin{abstract}
-  In the online Graph RePartitioning problem, we are given a sequence of pairwise communication requests between n nodes, revealed online.
-  The objective is to service these requests by partitioning the nodes into $\ell$ clusters, each of size $k$, such that frequently communicating nodes are located in the same cluster.
-  The partitioning can be updated by migrating nodes between clusters for a fixed cost.
-  The goal is to jointly minimize the amount of inter-cluster communication and migration cost.
+  In the online Graph RePartitioning problem,
+  a sequence of communication requests arrives online,
+  between pairs of $n$ nodes that are partitioned into $\ell$ clusters
+  each of size $k$ (i.e., $n = \ell \cdot k$).
+  The objective is to serve these requests while paying a cost that is reasonably close to the optimal cost of serving them given  the whole sequence in advance. 
+  This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
+  such that frequently communicating nodes (from different clusters)
+  are relocated to the same cluster before they inflict too much communication cost.
+  A \emph{reconfiguration} (i.e., re-partitioning) is performed by
+  moving (i.e., migrating) nodes between clusters at a cost.
+  The goal is to jointly minimize communication and migration costs while serving the sequence.
 
-  So far, for this natural problem, the best known deterministic algorithm was $O(k^2\cdot \ell^2)$-competitive, and the best known lower bound was $\Omega(k)$ for the competitive ratio of any deterministic algorithm that used the reduction from the caching problem.
+	The best known deterministic algorithm is $O(k^2\cdot \ell^2)$-competitive \cite{?????},
+	and the best known lower bound for the competitive ratio of
+	deterministic algorithms is $\Omega(k)$,
+	which uses a reduction from the online paging problem.
 	  In this paper, we provide the lower bound of $\Omega(k\cdot \ell)$.
-
-  We also present the optimal algorithm for $k=3$.
-  Finally, we provide a $O(k\cdot \ell)$-competitive algorithm for the scenario, where the perfect partition of input exists.
+    We also present an optimal algorithm for the case of $k=3$.
+  Finally,
+  we provide a $O(k\cdot \ell)$-competitive algorithm for the variant
+  where the optimal configuration is always a fixed partition
+  that incurs zero communication cost through the whole request sequence.
 \end{abstract}
@@
 each of size~$k$.
-We say that nodes in $C \subset V$ are \emph{collocated}% in a partition $P$
+We say that nodes in $C \subset V$ are \emph{collocated}
+% in a partition $P$
 if they are assigned to the same cluster.
@@
 We refer this variant as \PPOBRP{}.
-The task of an online algorithm for \PPOBRP{} is to "recover" (or learn) the perfect partition while not paying too much more than OPT.
+The task of an online algorithm for \PPOBRP{} is to ``recover" (or learn) the perfect partition while not paying too much more than OPT.
 This implies that on a request between any two non-collocated nodes,

### 2020-02-14 paper/paper.tex
swapping sections
rewording in related works

@@
 
-The model is related to online
-caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}. In our
-setting, requests can be served remotely, which is
-reminiscent of caching models \emph{with
-bypassing}~\cite{EpImLN11,EpImLN15,Irani02}.
+%The model is related to online
+%paging~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}.
+In our
+setting,
+each  requests between two nodes  not in the same cluster
+(i.e., inter-cluster requests) inflicts a cost until the nodes are placed in the same cluster.
+This is a reminiscent of caching models \emph{with
+rejection} penalty~\cite{EpImLN11,EpImLN15,Irani02}.
 
@@
 We refer to this connected sub-graph as a \emph{component}.
-An algorithm that always keeps a component collocated is \emph{component-respecting}.
+An algorithm that  keeps all components collocated is \emph{component-respecting}.
+
+\section{Lower bound} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
+\label{sec:lowerbound}
+
+In this section we show a lower bound of $\Omega(k \cdot \ell)$ for \OBRP{}.
+The constructed input sequence allows OPT to move to a perfect partition, and hence
+the construction also constitutes a lower bound for \PPOBRP{}.
+This complements an upper bound of $O(k \cdot \ell)$
+for \PPOBRP{} that we present in Section \ref{sec:ppl}.
+
+\begin{theorem}
+	The competitive ratio of any deterministic algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$.
+\end{theorem}
+
+\begin{proof}
+	We construct an instance of the problem with $\ell$ clusters 
+	$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
+	Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
+	%	\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
+	Fix any online algorithm \ALG{}.
+	We construct the input sequence of requests for \ALG{} as follows.
+	First,
+	we issue $k-2$ (internal) requests so that \ALG{} can form a component of $k-1$
+	nodes on the cluster $S_1$.
+	Recall that internal requests does not incur any cost for \ALG{}.
+	If \ALG{} at any point splits a component
+	(i.e., spreads its nodes over two or more clusters),
+	then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
+	If \ALG{} never collocates them then it is not competitive.
+
+	Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
+	Next,
+	we issue the request between $x_0$ and $y_0$.
+	Since this is an external request,
+	\ALG{} joins them into one component and collocates them in some cluster other than $S_1$.
+	For this,
+	\ALG{} moves to a new configuration
+	that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
+
+	Next,
+	we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
+	\ALG{} must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
+	consequently replaces $x_1$ with some other (single) node $x_2$.
+	We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
+	until there are only two single nodes left that  originate from the same cluster.
+	I.e.,
+	two single nodes $x^*, y^*,I(x^*) = I(y^*)$ s.t.~for any other pair of single nodes
+	$x'$ and $y'$,
+	we have $I(x') \neq I(y')$.
+	At this point there are at most $\ell+1$ single nodes left,
+	otherwise there would be more pairs of single nodes that were initially in the same cluster.
+
+	Given this sequence of requests,
+	the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
+	swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
+	Hence,
+	OPT pays for $4$ node migrations and
+	\ALG{} incurs at least one migration for each node in the sequence $X := x_0, x_1,\dots$.
+	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
+	therefore $|X| \geq k \cdot \ell - k - \ell \in \Omega(k\cdot\ell)$.
+\end{proof}
 
@@
 
-\section{Lower bound} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
-\label{sec:lowerbound}
-
-In this section we show a lower bound of $\Omega(k \cdot \ell)$ for \OBRP{}.
-The constructed input sequence allows OPT to move to a perfect partition, and hence
-the construction also constitutes a lower bound for \PPOBRP{}.
-This complements an upper bound of $O(k \cdot \ell)$
-for \PPOBRP{} that we present in Section \ref{sec:ppl}.
-
-\begin{theorem}
-	The competitive ratio of any deterministic algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$.
-\end{theorem}
-
-\begin{proof}
-	We construct an instance of the problem with $\ell$ clusters 
-	$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
-	Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
-%	\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
-	Fix any online algorithm \ALG{}.
-	We construct the input sequence of requests for \ALG{} as follows.
-	First,
-	we issue $k-2$ (internal) requests so that \ALG{} can form a component of $k-1$
-	nodes on the cluster $S_1$.
-	Recall that internal requests does not incur any cost for \ALG{}.
-	If \ALG{} at any point splits a component
-	(i.e., spreads its nodes over two or more clusters),
-	then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
-	If \ALG{} never collocates them then it is not competitive.
-
-	Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
-	Next,
-	we issue the request between $x_0$ and $y_0$.
-	Since this is an external request,
-	\ALG{} joins them into one component and collocates them in some cluster other than $S_1$.
-	For this,
-	\ALG{} moves to a new configuration
-	that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
-
-	Next,
-	we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
-	\ALG{} must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
-	consequently replaces $x_1$ with some other (single) node $x_2$.
-	We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
-	until there are only two single nodes left that  originate from the same cluster.
-	I.e.,
-	two single nodes $x^*, y^*,I(x^*) = I(y^*)$ s.t.~for any other pair of single nodes
-	$x'$ and $y'$,
-	we have $I(x') \neq I(y')$.
-	At this point there are at most $\ell+1$ single nodes left,
-	otherwise there would be more pairs of single nodes that were initially in the same cluster.
-
-	Given this sequence of requests,
-	the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
-	swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
-	Hence,
-	OPT pays for $4$ node migrations and
-	\ALG{} incurs at least one migration for each node in the sequence $X := x_0, x_1,\dots$.
-	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
-	therefore $|X| \geq k \cdot \ell - k - \ell \in \Omega(k\cdot\ell)$.
-\end{proof}
-
 \section{Corollary and Future Directions}
 
-The gap between upper and lower bound is still substantial, as the best known algorithm is the component-based algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\cdot\ell^2)$.
+The gap between upper and lower bound is still substantial, as the best known algorithm is the component-respecting algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\cdot\ell^2)$.
 

### 2020-02-14 paper/paper.tex
got rid of algpseudocode

@@
 \usepackage{amsmath,amsfonts,mathtools,amsthm}
-\usepackage[noend]{algpseudocode}
+\usepackage{algorithmic}
 \usepackage{algorithm}
@@
     \renewcommand{\algorithmicensure}{\textbf{Output:}}
-    \begin{algorithmic}[1]
-        \Require 
-        $k, \ell$,
-        initial configuration $P_I$,
-        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
-        \Ensure A final configuration $P_F$ 
-        \State For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$
-        \State{$P_0 := P_I$}
+    \begin{algorithmic}
+%        \Require 
+%        $k, \ell$,
+%        initial configuration $P_I$,
+%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
+%        \Ensure A final configuration $P_F$ 
+        \STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$}
+        \STATE{$P_0 := P_I$}
          \label{line:initcomponents}
-        \For {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
-        \State Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
-        \If{$C_1 \neq C_2$}
-        \State Unite the two components into a single component $C'$ and
-        $\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$ \label{line:mergecomponents}
-%        \If{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$}
-        \If{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$}
-        \Comment{if not in the same cluster}
-        \State $P_{t} = \mathit{re\mhyphen partition}(P_{t-1}, P_I, \mathcal{C}$) \label{line:rebalance} 
-        \Comment{move to $P$ closest to $P_I$}
-        \EndIf
-        \EndIf
-        \EndFor
+        \FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
+        \STATE Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
+        \IF{$C_1 \neq C_2$}
+        \STATE {Unite the two components into a single component $C'$ and
+        $\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$} \label{line:mergecomponents}
+        \IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
+	     \COMMENT{i.e.~if not in the same cluster}    
+    	}       
+        \STATE {$P_{t} = \mathit{re\mhyphen partition}(P_{t-1}, P_I, \mathcal{C})$} 
+         \COMMENT{move to $P$ closest to $P_I$}
+        \label{line:rebalance} 
+        \ENDIF
+        \ENDIF
+        \ENDFOR
     \end{algorithmic}

### 2020-02-14 paper/paper.tex
Better notation

--- 2020-02-14 paper/paper.tex
@@
 \begin{abstract}
-  In the online Graph RePartitioning problem,
-  a sequence of communication requests arrives online,
-  between pairs of $n$ nodes that are partitioned into $\ell$ clusters
-  each of size $k$ (i.e., $n = \ell \cdot k$).
-  The objective is to serve these requests while paying a cost that is reasonably close to the optimal cost of serving them given  the whole sequence in advance. 
-  This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
-  such that frequently communicating nodes (from different clusters)
-  are relocated to the same cluster before they inflict too much communication cost.
-  A \emph{reconfiguration} (i.e., re-partitioning) is performed by
-  moving (i.e., migrating) nodes between clusters at a cost.
-  The goal is to jointly minimize communication and migration costs while serving the sequence.
+  In the online Graph RePartitioning problem, we are given a sequence of pairwise communication requests between n nodes, revealed online.
+  The objective is to service these requests by partitioning the nodes into $\ell$ clusters, each of size $k$, such that frequently communicating nodes are located in the same cluster.
+  The partitioning can be updated by migrating nodes between clusters for a fixed cost.
+  The goal is to jointly minimize the amount of inter-cluster communication and migration cost.
 
-	The best known deterministic algorithm is $O(k^2\cdot \ell^2)$-competitive \cite{?????},
-	and the best known lower bound for the competitive ratio of
-	deterministic algorithms is $\Omega(k)$,
-	which uses a reduction from the online paging problem.
+  So far, for this natural problem, the best known deterministic algorithm was $O(k^2\cdot \ell^2)$-competitive, and the best known lower bound was $\Omega(k)$ for the competitive ratio of any deterministic algorithm that used the reduction from the caching problem.
   In this paper, we provide the lower bound of $\Omega(k\cdot \ell)$.
-    We also present an optimal algorithm for the case of $k=3$.
-  Finally,
-  we provide a $O(k\cdot \ell)$-competitive algorithm for the variant
-  where the optimal configuration is always a fixed partition
-  that incurs zero communication cost through the whole request sequence.
+
+  We also present the optimal algorithm for $k=3$.
+  Finally, we provide a $O(k\cdot \ell)$-competitive algorithm for the scenario, where the perfect partition of input exists.
 \end{abstract}
@@
 each of size~$k$.
-We say that nodes in $C \subset V$ are \emph{collocated}
-% in a partition $P$
+We say that nodes in $C \subset V$ are \emph{collocated}% in a partition $P$
 if they are assigned to the same cluster.
@@
 
-  \section{An Optimal Algorithm for $k=3$ (most likely wrong)}
+  \section{An Optimal Algorithm for $k=3$}
 
@@
   \begin{proof}
-    Fix a phase of \TAlg, and consider the state of its counters at the end of it.
+    Fix a completed phase of \TAlg, and consider the state of its counters at the end of it.
+    We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
 
-    Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
-    Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components throughout this phase.
-    Let $A^M$ be the cost of migrations performed by \TAlg in this phase.
+    We partition the input sequence $\sigma$ that arrived during the phase into subsequences $\sigma_I$ (called \emph{internal}) and $\sigma_E$ (called \emph{external}), defined as follows.
+    Requests $\sigma_E$ concern pairs of nodes that belong to the same component of \TAlg at the end of the phase, and $\sigma_E$ denotes all other requests issued during this phase.
 
-    We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
+    %Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
+    %Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
+    Let $ALG(M)$ be the cost of migrations performed by \TAlg in this phase.
 
@@
     Exceeding this number would mean that a component of size $4$ would emerge, a contradiction.
-    Together with Lemma~\ref{lem:1req}, this gives us $A^M \leq 6\cdot\alpha\cdot\ell$.
+    Together with Lemma~\ref{lem:1req}, this gives us $ALG(M) \leq 6\cdot\alpha\cdot\ell$.
 
-    Now we bound $A^I$.
+    Now we bound $ALG(\sigma_I)$.
     Each component of size $3$ contributes at most $(3\cdot \alpha - 1)$ to $A^I$, as at most $2$ of its pairs are saturated and contribute $\alpha$ each, and the unsaturated pair contributes at most $\alpha-1$.
-    Each of $\ell$ clusters of \TAlg contain at most one component of size $3$ (other cluster configurations contribute less: $3$ singletons contribute $0$ and a component of size $2$ contributes at most $\alpha$ to $A^I$), and in total, $A^I \leq \ell \cdot (3\cdot \alpha-1)$.
+    Each of $\ell$ clusters of \TAlg contain at most one component of size $3$ (other cluster configurations contribute less: $3$ singletons contribute $0$ and a component of size $2$ contributes at most $\alpha$ to $A^I$), and in total, $ALG(\sigma_I) \leq \ell \cdot (3\cdot \alpha-1)$.
 
-    We bound $A^E$ by $k^2 \cdot (\alpha - 1)$, as no more than $k^2$ pairs are unsaturated, and each of them contributes at most $\alpha -1$.
+    %We bound $A^E$ by $k^2 \cdot (\alpha - 1)$, as no more than $k^2$ pairs are unsaturated, and each of them contributes at most $\alpha -1$.
+    %\maciek{not needed most likely}
 
-    In total, the cost of \TAlg is at most $A^I + A^E + A^M$ during the phase.
+    In total, the cost of \TAlg is at most $ALG(\sigma_I) + ALG(\sigma_E) + ALG(M) \leq 9\alpha\cdot \ell + ALG(\sigma_E)$ during this phase.
 
-    Now we lower bound the cost of $\OPT$.
+    \medskip
+
+    Now we lower bound the cost of $\OPT$ by $\OPT(\sigma_I) +\OPT(\sigma_E)+\OPT(M)$, defined with respect to components of \TAlg.
+    We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
+    While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts.
+
+    %Let $O^M$ be the cost of migrations performed by $\OPT$ during the phase.
+    %Let $O^I$ be the cost of serving requests between nodes that were put in one component by \TAlg during this phase.
+    %Let $O^E$ be the cost serving requests between nodes that \TAlg did not put in the same component during that phase.
+
+    We have $OPT(M) + OPT(\sigma_I) \geq \alpha$, as the components of \TAlg could not be partitioned without splitting them, and there exist a pair of nodes with at least $\alpha$ requests between them.
+    $\OPT$ either served them remotely or performed a migration.
+
+    Now we bound $\OPT(M) + \OPT(\sigma_E))$.
+
+    \medskip
+
+    Now we consider the last, unfinished phase.
+    If the unfinished phase is also the first one, then we use the fact that we started with the same initial configuration.
+    Then, we charge $\OPT$ $\alpha$ for first $\alpha$ requests or a migration,
+    and we follow the analysis regarding the unsaturated requests.
+    \maciek{todo}
+
+    If we have at least two phases, then we assume that OPT did not pay during that, and the cost of ALG increases at most twofold.
   \end{proof}

### 2020-02-15 paper/paper.tex
Section intro

@@
 
-	The best known deterministic algorithm is $O(k^2\cdot \ell^2)$-competitive \cite{?????},
+	The best known deterministic algorithm is $O(k^2\cdot \ell^2)$-competitive \cite{?????} \maciek{We should cite \cite{repartition-disc} here.},
 	and the best known lower bound for the competitive ratio of
@@
 \label{sec:lowerbound}
+
+\maciek{emphasize that lower bound works for $k=3$ already. Also in the proof?}
 
@@
 
+\todo{Make sure our manuscript is autor anonymous}
+
 \todo{Font for OPT, ALG, DET, dist etc}
@@
 
-  In this section we provide an optimal algorithm for the case $k=3$.
+  In this section we show that the classic \emph{rent-or-buy} approach yields an optimal algorithm for $k=3$.
+  Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one server until the end of an epoch (that we define precisely later).
+  To this end, we maintain \emph{components of nodes}, in similar way to \cite{repartition-disc}.
 

### 2020-02-15 paper/paper.tex
Clean up todos
Remove omitted results

@@
 
-\todo{ACM format for review? With line numbers etc.}
+\todo{ACM format for review? Is there any specific format like this? With line numbers etc.}
 
-\todo{Make ORCID visible in ACM style: "Right now we do not typeset ORCIDs" (Source: ACM Style Docs)}
-
-\todo{Include e-mails}
+\todo{Include e-mails? Or shall we skip it for now.}
 
@@
 \todo{Make an ACM account for the submission}
-
-\todo{Sum up all the steps needed for the submission}
 
@@
 \todo{Font for OPT, ALG, DET, dist etc}
-
-	\section{Ommited results}
-	\begin{enumerate}
-		 \item The lower bound of $\Omega(\ell)$ for the scenario with at most $(1/3-1/k)$-augmentation
-		 \item Costly cascade example (it is easy to present $k^2$ and $k\cdot \sqrt{k}$ (with different approach), and then we can combine these to obtain $k^3$)
-		 \item $k$-competitive algorithm for the ring communication pattern
-		 \item $O(\ell)$-competitive algorithm for $k=4$ and $k=5$
-  \end{enumerate}
 

### 2020-02-15 paper/paper.tex
k=3 in the paper
Previous (wrong) proof of k=3 incorporated already

@@
 
+\section{An Optimal Algorithm for $k=3$}
+
+In this section we show that the classic \emph{rent-or-buy} approach yields an optimal algorithm for $k=3$.
+Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one server until the end of an epoch (that we define precisely later).
+To this end, we maintain \emph{components of nodes}, in similar way to \cite{repartition-disc}.
+
+We define the algorithm \TAlg{} in the following way.
+\TAlg{} maintains components for each node.
+Intially, each node is in its own singleton component.
+For each pair of nodes \TAlg{} maintains a counter, initiated $0$, and upon receiving a request to this pair while its nodes are located on different servers, it increases the counter by $1$.
+If a counter for a pair $(u,v)$ reaches $\alpha$, \TAlg{} merges the components of $u$ and $v$, and moves to the closest partition, where it each component has all of its nodes in one server.
+If no such partition exists, \TAlg{} resets all components to singletons and all counters to $0$.
+
+%\maciek{Shouldn't we think about whole component cuts instead of single edge? Hmm. NO!}
+
+We say that \TAlg is \emph{component respecting}, i.e., it maintains the invariant that nodes of each component are located at the same server.
+Observe that this implies that \TAlg never increases any counter above $\alpha$.
+
+\begin{lemma}
+  \label{lem:1req}
+  A single repartition incurs the cost at most $3\cdot\alpha$ for \TAlg.
+\end{lemma}
+
+\begin{proof}
+  If no partition of \TAlg's components exist, then it performs no repartition, and serves the request remotely, incurring the cost $1$.
+  Hence, in the following we bound the cost of moving to component-respecting partition.
+
+  We distinguish between three types of clusters: $C_1, C_2, C_3$,
+  which we define as follows.
+  A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+  A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
+  Finally, a cluster of type $C_3$  contains one component of size $3$.
+
+    Consider a request $(u, v)$ and let $U, V$ be clusters of these nodes.
+    If $U=V$, then the request does not require a reconfiguration and GREEDY incurs no cost.
+    Otherwise, we consider cases upon the type of clusters $U$ and $V$.
+    Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise we encounter a component of size $4$ and this would contradict the case assumption that the component-respecting partition exists.
+    If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
+
+    Finally, we consider the case where both $U$ and $V$ are $C_2$.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg{} has the component of size $4$, a contradiction with the case assumption that the component-respecting partition exists.
+    If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
+    Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise the input graph would not have a perfect partition.
+In this case GREEDY exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost of $2\cdot \alpha$.
+\end{proof}
+
+\begin{theorem}
+  \TAlg{} is $O(\ell)$-competitive.
+\end{theorem}
+\begin{proof}
+  Fix a completed phase of \TAlg, and consider the state of its counters at the end of it.
+  We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
+
+  We partition the input sequence $\sigma$ that arrived during the phase into subsequences $\sigma_I$ (called \emph{internal}) and $\sigma_E$ (called \emph{external}), defined as follows.
+  Requests $\sigma_E$ concern pairs of nodes that belong to the same component of \TAlg at the end of the phase, and $\sigma_E$ denotes all other requests issued during this phase.
+
+  %Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
+  %Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
+  Let $ALG(M)$ be the cost of migrations performed by \TAlg in this phase.
+
+  During a single phase, at most $2\cdot \ell$ component merge operations are performed.
+  Exceeding this number would mean that a component of size $4$ would emerge, a contradiction.
+  Together with Lemma~\ref{lem:1req}, this gives us $ALG(M) \leq 6\cdot\alpha\cdot\ell$.
+
+  Now we bound $ALG(\sigma_I)$.
+  Each component of size $3$ contributes at most $(3\cdot \alpha - 1)$ to $A^I$, as at most $2$ of its pairs are saturated and contribute $\alpha$ each, and the unsaturated pair contributes at most $\alpha-1$.
+  Each of $\ell$ clusters of \TAlg contain at most one component of size $3$ (other cluster configurations contribute less: $3$ singletons contribute $0$ and a component of size $2$ contributes at most $\alpha$ to $A^I$), and in total, $ALG(\sigma_I) \leq \ell \cdot (3\cdot \alpha-1)$.
+
+  %We bound $A^E$ by $k^2 \cdot (\alpha - 1)$, as no more than $k^2$ pairs are unsaturated, and each of them contributes at most $\alpha -1$.
+  %\maciek{not needed most likely}
+
+  In total, the cost of \TAlg is at most $ALG(\sigma_I) + ALG(\sigma_E) + ALG(M) \leq 9\alpha\cdot \ell + ALG(\sigma_E)$ during this phase.
+
+  \medskip
+
+  Now we lower bound the cost of $\OPT$ by $\OPT(\sigma_I) +\OPT(\sigma_E)+\OPT(M)$, defined with respect to components of \TAlg.
+  We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
+  While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts.
+
+  %Let $O^M$ be the cost of migrations performed by $\OPT$ during the phase.
+  %Let $O^I$ be the cost of serving requests between nodes that were put in one component by \TAlg during this phase.
+  %Let $O^E$ be the cost serving requests between nodes that \TAlg did not put in the same component during that phase.
+
+  We have $OPT(M) + OPT(\sigma_I) \geq \alpha$, as the components of \TAlg could not be partitioned without splitting them, and there exist a pair of nodes with at least $\alpha$ requests between them.
+  $\OPT$ either served them remotely or performed a migration.
+
+  Now we bound $\OPT(M) + \OPT(\sigma_E))$.
+
+  \medskip
+
+  Now we consider the last, unfinished phase.
+  If the unfinished phase is also the first one, then we use the fact that we started with the same initial configuration.
+  Then, we charge $\OPT$ $\alpha$ for first $\alpha$ requests or a migration,
+  and we follow the analysis regarding the unsaturated requests.
+  \maciek{todo}
+
+  If we have at least two phases, then we assume that OPT did not pay during that, and the cost of ALG increases at most twofold.
+\end{proof}
+
 \section{Recovering a Perfect Partition}	\label{sec:ppl}
@@
 
-  \section{An Optimal Algorithm for $k=3$}
-
-  In this section we show that the classic \emph{rent-or-buy} approach yields an optimal algorithm for $k=3$.
-  Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one server until the end of an epoch (that we define precisely later).
-  To this end, we maintain \emph{components of nodes}, in similar way to \cite{repartition-disc}.
-
-  We define the algorithm \TAlg{} in the following way.
-  \TAlg{} maintains components for each node.
-  Intially, each node is in its own singleton component.
-  For each pair of nodes \TAlg{} maintains a counter, initiated $0$, and upon receiving a request to this pair while its nodes are located on different servers, it increases the counter by $1$.
-  If a counter for a pair $(u,v)$ reaches $\alpha$, \TAlg{} merges the components of $u$ and $v$, and moves to the closest partition, where it each component has all of its nodes in one server.
-  If no such partition exists, \TAlg{} resets all components to singletons and all counters to $0$.
-
-  %\maciek{Shouldn't we think about whole component cuts instead of single edge? Hmm. NO!}
-
-  We say that \TAlg is \emph{component respecting}, i.e., it maintains the invariant that nodes of each component are located at the same server.
-  Observe that this implies that \TAlg never increases any counter above $\alpha$.
-
-  \begin{lemma}
-    \label{lem:1req}
-    A single repartition incurs the cost at most $3\cdot\alpha$ for \TAlg.
-  \end{lemma}
-
-  \begin{proof}
-    If no partition of \TAlg's components exist, then it performs no repartition, and serves the request remotely, incurring the cost $1$.
-    Hence, in the following we bound the cost of moving to component-respecting partition.
-
-    We distinguish between three types of clusters: $C_1, C_2, C_3$,
-    which we define as follows.
-    A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
-    A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
-    Finally, a cluster of type $C_3$  contains one component of size $3$.
-
-      Consider a request $(u, v)$ and let $U, V$ be clusters of these nodes.
-      If $U=V$, then the request does not require a reconfiguration and GREEDY incurs no cost.
-      Otherwise, we consider cases upon the type of clusters $U$ and $V$.
-      Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise we encounter a component of size $4$ and this would contradict the case assumption that the component-respecting partition exists.
-      If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
-
-      Finally, we consider the case where both $U$ and $V$ are $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg{} has the component of size $4$, a contradiction with the case assumption that the component-respecting partition exists.
-      If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
-      Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-      Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise the input graph would not have a perfect partition.
-In this case GREEDY exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost of $2\cdot \alpha$.
-  \end{proof}
-
-  \begin{theorem}
-    \TAlg{} is $O(\ell)$-competitive.
-  \end{theorem}
-  \begin{proof}
-    Fix a completed phase of \TAlg, and consider the state of its counters at the end of it.
-    We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
-
-    We partition the input sequence $\sigma$ that arrived during the phase into subsequences $\sigma_I$ (called \emph{internal}) and $\sigma_E$ (called \emph{external}), defined as follows.
-    Requests $\sigma_E$ concern pairs of nodes that belong to the same component of \TAlg at the end of the phase, and $\sigma_E$ denotes all other requests issued during this phase.
-
-    %Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
-    %Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
-    Let $ALG(M)$ be the cost of migrations performed by \TAlg in this phase.
-
-    During a single phase, at most $2\cdot \ell$ component merge operations are performed.
-    Exceeding this number would mean that a component of size $4$ would emerge, a contradiction.
-    Together with Lemma~\ref{lem:1req}, this gives us $ALG(M) \leq 6\cdot\alpha\cdot\ell$.
-
-    Now we bound $ALG(\sigma_I)$.
-    Each component of size $3$ contributes at most $(3\cdot \alpha - 1)$ to $A^I$, as at most $2$ of its pairs are saturated and contribute $\alpha$ each, and the unsaturated pair contributes at most $\alpha-1$.
-    Each of $\ell$ clusters of \TAlg contain at most one component of size $3$ (other cluster configurations contribute less: $3$ singletons contribute $0$ and a component of size $2$ contributes at most $\alpha$ to $A^I$), and in total, $ALG(\sigma_I) \leq \ell \cdot (3\cdot \alpha-1)$.
-
-    %We bound $A^E$ by $k^2 \cdot (\alpha - 1)$, as no more than $k^2$ pairs are unsaturated, and each of them contributes at most $\alpha -1$.
-    %\maciek{not needed most likely}
-
-    In total, the cost of \TAlg is at most $ALG(\sigma_I) + ALG(\sigma_E) + ALG(M) \leq 9\alpha\cdot \ell + ALG(\sigma_E)$ during this phase.
-
-    \medskip
-
-    Now we lower bound the cost of $\OPT$ by $\OPT(\sigma_I) +\OPT(\sigma_E)+\OPT(M)$, defined with respect to components of \TAlg.
-    We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
-    While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts.
-
-    %Let $O^M$ be the cost of migrations performed by $\OPT$ during the phase.
-    %Let $O^I$ be the cost of serving requests between nodes that were put in one component by \TAlg during this phase.
-    %Let $O^E$ be the cost serving requests between nodes that \TAlg did not put in the same component during that phase.
-
-    We have $OPT(M) + OPT(\sigma_I) \geq \alpha$, as the components of \TAlg could not be partitioned without splitting them, and there exist a pair of nodes with at least $\alpha$ requests between them.
-    $\OPT$ either served them remotely or performed a migration.
-
-    Now we bound $\OPT(M) + \OPT(\sigma_E))$.
-
-    \medskip
-
-    Now we consider the last, unfinished phase.
-    If the unfinished phase is also the first one, then we use the fact that we started with the same initial configuration.
-    Then, we charge $\OPT$ $\alpha$ for first $\alpha$ requests or a migration,
-    and we follow the analysis regarding the unsaturated requests.
-    \maciek{todo}
-
-    If we have at least two phases, then we assume that OPT did not pay during that, and the cost of ALG increases at most twofold.
-  \end{proof}
-
-  \section{An old section about $k=3$}
-
-  \label{sec:k3}
-  In this section,
-  we give an alternative algorithm for the special case of $k=3$.
-  The algorithm does not  improve upon PPL (asymptotically),
-  but we will use it in Section \ref{sec:generalModel} for the same special case in the general model.
-
-  \maciek{We must introduce phases in the description of this algorithm. If no PP exists, the algorithm resets its components.}
-
-  Let \emph{GREEDY} be the algorithm that on each inter-cluster request
-   moves to a perfect partition
-  % \maciek{of input}
-  closest to the current partition,
-  i.e.,
-  reachable in minimum number of node swaps (ties broken arbitrarily).
-
-  \begin{lemma} \label{lemma:k=3}
-    Every re-balancing cost of GREEDY for $k=3$ is $O(1)$.
-    %\maciek{GREEDY = find the cheapest (closest to the current) feasible partition}
-    \maciek{We need to add to the assumption that the input has a perfect partition
-    \mahmoud{the section now is only about PP model}}
-    \label{rebalancing-cost}
-  \end{lemma}
-
-  \begin{proof} 
-  % \maciek{Note: we consider a cost of cheapest feasible rebalancing after insertion of one edge.
-This in contrast to bounding the distance between any two reconfigurations (this theorem does not bound the latter).}
-
-    \maciek{We probably should start with something along: Note that the assumption of perfect partition means that there is no component of size 4. Then refer to it (twice) by some short ref.}
-
-    We distinguish among three configuration types of any algorithm's cluster: $C_1, C_2, C_3$, which we define as follows.
-In $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
-In $C_2$ it contains one component of size $2$ and one component of size $1$.
-Finally, in $C_3$ it contains one component of size $3$.
-
-    Consider a request $(u, v)$ and let $U, V$ be clusters of these nodes.
-    If $U=V$, then the request does not require a reconfiguration and GREEDY incurs no cost.
-    Otherwise, we consider cases upon the type of clusters $U$ and $V$.
-    Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise we encounter a component of size $4$ and this would contradict the existence of the perfect partition of the input.
-    If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
-
-    Finally, we consider the case where both $U$ and $V$ are $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that an input sequence includes the component of size $4$, a contradiction.
-    If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
-    Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise the input graph would not have a perfect partition.
-In this case GREEDY exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost of $2\cdot \alpha$.
-  \end{proof}
-
-  \begin{theorem}
-    Algorithm \emph{GREEDY} is $O(\ell)$-competitive when $k=3$.
-    \maciek{Assumming the perfect partition of input sequence
-    \mahmoud{the section now is only about PP model}}
-  \end{theorem}
-
-  \begin{proof}
-    The adversary issues at most $3\ell$ requests, as otherwise the perfect partition would not exist.
-    Each request triggers at most one re-balancing in \emph{GREEDY}
-    which costs $O(1)$ by Theorem~\ref{rebalancing-cost}.
-    Then,
-    the total cost of any algorithm is $O(\ell)$.
-  \end{proof}
-
-  \section{Arbitrary Communication Graph} \label{sec:generalModel}
-
-  We leverage the simple algorithm \emph{GREEDY} for $k=3$
-  towards an algorithm for the general problem
-  where the communication graph is not restricted.
-  Note that in the general model,
-  an optimal partition is not a necessarily a perfect partition.
-  We divide the sequence of requests into phases.
-  A phase terminates when there is no perfect partition of components.
-  We apply requests to \emph{GREEDY} as before. 
-  Once the current phase terminates,
-  we reset all components by removing (forgetting) all revealed edges,
-  and then we proceed to the next phase.
-
-  \begin{theorem} \label{cor:k=3}
-    There exists $O(\ell)$-competitive algorithm for the general online problem when $k=3$. 
-  \end{theorem}
-  \begin{proof}
-    \maciek{TODO: this proof is incomplete!}
-    Consider any phase and assume it terminates upon arrival of some request $r$.
-    By Lemma  \ref{lemma:k=3},
-    \emph{GREEDY} pays $O(\ell)$ during the phase.
-    If OPT stays in the same configuration throughout the phase,
-    regardless of its partition at the beginning of the phase,        
-    it incurs remote communication cost for at least one request $r'$ (possibly $r'=r$) by the end of the phase.
-    Else,
-    OPT incurs the cost of moving to new partitions.
-    In any case,
-    OPT pays at least one during the phase and 
-    the competitive ratio is $O(\ell)$.
-  \end{proof}
-
 \end{appendix}

### 2020-02-15 paper/paper.tex
Fixing lem:1req

@@
 If a counter for a pair $(u,v)$ reaches $\alpha$, \TAlg{} merges the components of $u$ and $v$, and moves to the closest partition, where it each component has all of its nodes in one server.
-If no such partition exists, \TAlg{} resets all components to singletons and all counters to $0$.
+If no such partition exists, \TAlg{} resets all components to singletons and all counters to $0$ and ends the phase.
 
@@
   \label{lem:1req}
-  A single repartition incurs the cost at most $3\cdot\alpha$ for \TAlg.
+  A single repartition performed by \TAlg{} incurs the cost at most $3\cdot\alpha$.
 \end{lemma}
@@
 \begin{proof}
-  If no partition of \TAlg's components exist, then it performs no repartition, and serves the request remotely, incurring the cost $1$.
+  If no partition of \TAlg's components exist, then it performs no repartition and the cost is $0$.
   Hence, in the following we bound the cost of moving to component-respecting partition.
+  We define three types of clusters, $C_1, C_2, C_3$; in cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
 
-  We distinguish between three types of clusters: $C_1, C_2, C_3$,
-  which we define as follows.
-  A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
-  A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
-  Finally, a cluster of type $C_3$  contains one component of size $3$.
+  %We distinguish between three types of clusters: $C_1, C_2, C_3$,
+  %which we define as follows.
+  %A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+  %A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
+  %Finally, a cluster of type $C_3$  contains one component of size $3$.
 
     Consider a request $(u, v)$ and let $U, V$ be clusters of these nodes.
-    If $U=V$, then the request does not require a reconfiguration and GREEDY incurs no cost.
+    If $U=V$, then the request does not require a reconfiguration and \TAlg{} incurs no cost.
     Otherwise, we consider cases upon the type of clusters $U$ and $V$.
     Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise we encounter a component of size $4$ and this would contradict the case assumption that the component-respecting partition exists.
-    If either $U$ or $V$ is $C_1$, then either cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
+    If either $U$ or $V$ is $C_1$, then this cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
 
@@
 Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg{} has the component of size $4$, a contradiction with the case assumption that the component-respecting partition exists.
-    If either $u$ or $v$ belongs to a component of size $2$ then it suffices to exchange components of size $1$ between $U$ and $V$.
+    If either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
     Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise the input graph would not have a perfect partition.
-In this case GREEDY exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost of $2\cdot \alpha$.
+    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component-respecting partition exists.
+In this case \TAlg{} exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost of $2\cdot \alpha$.
 \end{proof}

### 2020-02-15 paper/paper.tex
Restructuring

@@
 
+In our analysis, we distinguish among three types of clusters, $C_1, C_2, C_3$. In cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
+
 \begin{lemma}
@@
   Hence, in the following we bound the cost of moving to component-respecting partition.
-  We define three types of clusters, $C_1, C_2, C_3$; in cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
 
@@
   Fix a completed phase of \TAlg, and consider the state of its counters at the end of it.
+  We consider the incomplete phase later in this proof.
   We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
 
-  We partition the input sequence $\sigma$ that arrived during the phase into subsequences $\sigma_I$ (called \emph{internal}) and $\sigma_E$ (called \emph{external}), defined as follows.
-  Requests $\sigma_E$ concern pairs of nodes that belong to the same component of \TAlg at the end of the phase, and $\sigma_E$ denotes all other requests issued during this phase.
+  We partition the input sequence $\sigma$ that arrived during the phase into subsequences $\sigma_I$ and $\sigma_E$, called \emph{internal} and \emph{external}, respectively.
+  The subsequence $\sigma_E$ contains requests to pairs that belong to the same component of \TAlg{} at the end of the phase, and $\sigma_E$ denotes the remaining requests issued during this phase.
 
@@
   %Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
-  Let $ALG(M)$ be the cost of migrations performed by \TAlg in this phase.
-
+  Let $\TAlg(M)$ be the cost of migrations performed by \TAlg in this phase.
   During a single phase, at most $2\cdot \ell$ component merge operations are performed.
   Exceeding this number would mean that a component of size $4$ would emerge, a contradiction.
-  Together with Lemma~\ref{lem:1req}, this gives us $ALG(M) \leq 6\cdot\alpha\cdot\ell$.
+  Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
 
-  Now we bound $ALG(\sigma_I)$.
-  Each component of size $3$ contributes at most $(3\cdot \alpha - 1)$ to $A^I$, as at most $2$ of its pairs are saturated and contribute $\alpha$ each, and the unsaturated pair contributes at most $\alpha-1$.
-  Each of $\ell$ clusters of \TAlg contain at most one component of size $3$ (other cluster configurations contribute less: $3$ singletons contribute $0$ and a component of size $2$ contributes at most $\alpha$ to $A^I$), and in total, $ALG(\sigma_I) \leq \ell \cdot (3\cdot \alpha-1)$.
+  Now we bound $\TAlg(\sigma_I)$.
+  Cluster of type $C_3$ contributes at most $(3\cdot \alpha - 1)$ to $\TAlg(\sigma_I)$, as at most $2$ of pairs of the component are saturated and contribute $\alpha$ each, and the remaining unsaturated pair contributes at most $\alpha-1$.
+  Other cluster configurations contribute less: $C_1$ contributes $0$ and $C_2$ contributes $\alpha$ to $\TAlg(\sigma_I)$.
+  Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq \ell \cdot (3\cdot \alpha-1)$.
 
@@
 
-  In total, the cost of \TAlg is at most $ALG(\sigma_I) + ALG(\sigma_E) + ALG(M) \leq 9\alpha\cdot \ell + ALG(\sigma_E)$ during this phase.
+  In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 9\alpha\cdot \ell + \TAlg(\sigma_E)$ during this phase.
 
@@
 
-  Now we lower bound the cost of $\OPT$ by $\OPT(\sigma_I) +\OPT(\sigma_E)+\OPT(M)$, defined with respect to components of \TAlg.
-  We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
+  Now we lower bound the cost of $\OPT$ by $\OPT(\sigma_I) +\OPT(\sigma_E)+\OPT(M)$, with $\sigma_I$ and $\sigma_E$ defined with respect to components of \TAlg, and $\OPT(M)$ denoting the cost of migrations performed by $\OPT$.
+  %We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
   While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts.
@@
 
-  We have $OPT(M) + OPT(\sigma_I) \geq \alpha$, as the components of \TAlg could not be partitioned without splitting them, and there exist a pair of nodes with at least $\alpha$ requests between them.
+  We have $OPT(M) + OPT(\sigma_I) \geq \alpha$, as the components of \TAlg{} could not be partitioned without splitting them, and there exist a pair of nodes with at least $\alpha$ requests between them.
   $\OPT$ either served them remotely or performed a migration.

### 2020-02-15 paper/paper.tex
Better bound on OPT

@@
 
-  Now we lower bound the cost of $\OPT$ by $\OPT(\sigma_I) +\OPT(\sigma_E)+\OPT(M)$, with $\sigma_I$ and $\sigma_E$ defined with respect to components of \TAlg, and $\OPT(M)$ denoting the cost of migrations performed by $\OPT$.
-  %We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
+  Now we lower bound the cost of $\OPT$ in this phase.
+  We define $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$, with $\sigma_I$ and $\sigma_E$ defined with respect to components of \TAlg.
+  By $\OPT(M)$ we denote the cost of migrations performed by $\OPT$.
+
+  We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
   While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts.
+  We have $\OPT \geq (\OPT(\sigma_I) + \OPT(M)) + (\OPT(\sigma_E) + \OPT(M))$.
 
@@
 
+  %First, we estimate the cost related to $\sigma_I$.
   We have $OPT(M) + OPT(\sigma_I) \geq \alpha$, as the components of \TAlg{} could not be partitioned without splitting them, and there exist a pair of nodes with at least $\alpha$ requests between them.
   $\OPT$ either served them remotely or performed a migration.
+  We obtain $\OPT \geq \max\{\alpha, \OPT(\sigma_E) + \OPT(M)}$.
 
-  Now we bound $\OPT(M) + \OPT(\sigma_E))$.
+  \medskip
+
+  Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ related to $\sigma_E$.
+
+  Second, we estimate the cost related to $\sigma_E$.
+  Now we bound the cost that $\OPT$ incurred while serving requests $\sigma_E$.
+  If $\OPT$ would
+
+  \medskip
+
+  Now we bound the competitive ratio.
 

### 2020-02-15 paper/paper.tex
Better bound on OPT
xspace

@@
 \usepackage[utf8]{inputenc}
-%\usepackage{xspace}
+\usepackage{xspace}
 \usepackage{balance}
@@
 \newcommand{\PPOBRP}{PP-BRP}
-\newcommand{\TAlg}{{\ensuremath{ALG_{k=3}}}} % we should change this name
+\newcommand{\TAlg}{{\ensuremath{ALG_{k=3}}}\xspace} % we should change this name
 
@@
   $\OPT$ either served them remotely or performed a migration.
-  We obtain $\OPT \geq \max\{\alpha, \OPT(\sigma_E) + \OPT(M)}$.
+  We obtain $\OPT \geq \max\{\alpha, \OPT(\sigma_E) + \OPT(M)\}$.
 
@@
 
-  Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ related to $\sigma_E$.
+  \maciek{unfinished, don't read :)}
 
-  Second, we estimate the cost related to $\sigma_E$.
-  Now we bound the cost that $\OPT$ incurred while serving requests $\sigma_E$.
-  If $\OPT$ would
+  Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to $\sigma_E$.
+  We begin by lower-bounding the number of requests from $\sigma_E$ that $\OPT$ must account (by either serving remotely or migrating), and we denote their number by $x$.
+  It holds that $x \geq \TAlg(\sigma_E) - 3\cdot\ell\cdot(\alpha - 1)$, as it received all the requests that \TAlg did, but it could ignore the cost of serving at most $3\cdot\ell$ pairs of nodes, as it might have them collocated in its clusters.
+  Each such pair of nodes account for at most $\alpha-1$ requests (the pairs from $\sigma_E$ are unsaturated) in the cost of $\TAlg(\sigma_E)$, hence collocating them would decrease the number by at most $3\cdot\ell\cdot(\alpha - 1)$.
 
-  \medskip
-
-  Now we bound the competitive ratio.
+  Now we bound the cost that $\OPT$ pays while serving these $x$ requests.
+  Instead of serving all of them remotely, $\OPT$ can perform migrations to decrease its cost.
+  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
+  (express that between a pair of nodes we accounted $\alpha-1$ requests in $x$)
+  Hence, by performing a migration for the cost $\alpha$, $\OPT$ does not pay the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $x$ ($x$ is not a sequence but its length!).
+  Finally, for serving $x$, $\OPT$ pays at least $x \cdot \frac{\alpha}{4\cdot (\alpha-1)}$.
+  This gives us
+  \begin{align*}
+    \OPT(\sigma_E) + \OPT(M) & \geq (\TAlg(\sigma_E)-3\cdot\ell\cdot(\alpha - 1)) \cdot \frac{\alpha}{4\cdot (\alpha-1)}\\
+    &\geq ...
+  \end{align*}
 

### 2020-02-15 paper/paper.tex
Obtaining the bound on sigmaE
The competitive ratio
Fixing OPT bound

@@
 \begin{theorem}
-  \TAlg{} is $O(\ell)$-competitive.
+  \TAlg is $O(\ell)$-competitive.
 \end{theorem}
@@
   While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts.
-  We have $\OPT \geq (\OPT(\sigma_I) + \OPT(M)) + (\OPT(\sigma_E) + \OPT(M))$.
+  Combining those bounds gives us $\OPT \geq \max\{\OPT(\sigma_I) + \OPT(M), \OPT(\sigma_E) + \OPT(M)\} \geq (\OPT(\sigma_I) + \OPT(M) + \OPT(\sigma_E) + \OPT(M)) / 2$.
 
@@
   $\OPT$ either served them remotely or performed a migration.
-  We obtain $\OPT \geq \max\{\alpha, \OPT(\sigma_E) + \OPT(M)\}$.
 
   \medskip
-
-  \maciek{unfinished, don't read :)}
 
@@
     \OPT(\sigma_E) + \OPT(M) & \geq (\TAlg(\sigma_E)-3\cdot\ell\cdot(\alpha - 1)) \cdot \frac{\alpha}{4\cdot (\alpha-1)}\\
-    &\geq ...
+    &\geq \TAlg(\sigma_E) / 4 - 4\cdot \ell \cdot \alpha
   \end{align*}
+  From that, we obtain $\TAlg(\sigma_E) \geq 4\cdot(\OPT(\sigma_E)+\OPT(M)) + 16\ell \cdot \alpha$.
+
+  Now we are ready to bound the competitive ratio.
+  Let $E := \OPT(\sigma_E) + \OPT(M)$. Then, using above bounds,
+  \begin{equation*}
+    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{9\alpha \cdot \ell + \TAlg(\sigma_E)}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{50\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 50 \ell = O(\ell).
+  \end{equation*}
 
@@
   Now we consider the last, unfinished phase.
-  If the unfinished phase is also the first one, then we use the fact that we started with the same initial configuration.
-  Then, we charge $\OPT$ $\alpha$ for first $\alpha$ requests or a migration,
+  Consider the case, where the unfinished phase is also the first one.
+  Then, we cannot charge $\OPT$ due to unability to partiton the components.
+  Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
+  We charge $\OPT$ $\alpha$ for the first external $\alpha$ requests or a migration,
   and we follow the analysis regarding the unsaturated requests.
-  \maciek{todo}
+  If the input finished before first $\alpha$ external requests, then \TAlg is optimal.
 

### 2020-02-15 paper/paper.tex
The last phase

@@
 
-  If we have at least two phases, then we assume that OPT did not pay during that, and the cost of ALG increases at most twofold.
+  If we have at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests
+  This way, the competitive ratio increases at most twofold.
 \end{proof}

### 2020-02-15 paper/paper.tex
Rewrite

@@
 
-In this section we show that the classic \emph{rent-or-buy} approach yields an optimal algorithm for $k=3$.
+In this section we show that the classic \emph{rent-or-buy} approach\cite{borodin-book} allows to obtain an optimal algorithm for $k=3$.
 Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one server until the end of an epoch (that we define precisely later).
-To this end, we maintain \emph{components of nodes}, in similar way to \cite{repartition-disc}.
+To this end, we maintain \emph{components of nodes}, in similar fashion to previous approaches\cite{repartition-disc}.
 
-We define the algorithm \TAlg{} in the following way.
-\TAlg{} maintains components for each node.
+We say that the request $(u,v)$ is \emph{external}, if the algorithm that receives this request is in the configuration, where $u$ and $v$ are located at different servers.
+
+We define the algorithm \TAlg in the following way.
+\TAlg maintains components for each node.
 Intially, each node is in its own singleton component.
-For each pair of nodes \TAlg{} maintains a counter, initiated $0$, and upon receiving a request to this pair while its nodes are located on different servers, it increases the counter by $1$.
-If a counter for a pair $(u,v)$ reaches $\alpha$, \TAlg{} merges the components of $u$ and $v$, and moves to the closest partition, where it each component has all of its nodes in one server.
-If no such partition exists, \TAlg{} resets all components to singletons and all counters to $0$ and ends the phase.
+For each pair of nodes \TAlg maintains a counter, initiated $0$. 
+Upon receiving an external request to a pair, it increases the counter by $1$.
+If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest partition, in which each component has all of its nodes collocated in one server.
+If no such partition exists, \TAlg resets all components to singletons and all counters to $0$ and ends the phase.
 
@@
   Consider the case, where the unfinished phase is also the first one.
-  Then, we cannot charge $\OPT$ due to unability to partiton the components.
+  Then, we cannot charge $\OPT$ due to unability to partition the components.
   Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.

### 2020-02-15 paper/paper.tex
wip

@@
 
-We say that the request $(u,v)$ is \emph{external}, if the algorithm that receives this request is in the configuration, where $u$ and $v$ are located at different servers.
+We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different servers.
 
 We define the algorithm \TAlg in the following way.
-\TAlg maintains components for each node.
+\TAlg partition nodes into components.
 Intially, each node is in its own singleton component.

### 2020-02-15 paper/paper.tex
Todo
External was already used

@@
 
-We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different servers.
+%We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different servers.
 
@@
 For each pair of nodes \TAlg maintains a counter, initiated $0$. 
-Upon receiving an external request to a pair, it increases the counter by $1$.
+Upon receiving a request to a pair that is not collocated at one server, it increases the counter by $1$.
 If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest partition, in which each component has all of its nodes collocated in one server.
@@
 
+\todo{Check naming consistency inside sections, and across them. Possibly contain some definitions in preliminaries.}
+
 \todo{ACM format for review? Is there any specific format like this? With line numbers etc.}

### 2020-02-15 paper/paper.tex
More precise lemma 1req formulation

@@
 
-In our analysis, we distinguish among three types of clusters, $C_1, C_2, C_3$. In cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
+In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
 
@@
   \label{lem:1req}
-  A single repartition performed by \TAlg{} incurs the cost at most $3\cdot\alpha$.
+  The cost of migrations during a single repartition of \TAlg is at most $3\cdot\alpha$.
 \end{lemma}
@@
   \medskip
-
   Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to $\sigma_E$.
@@
 
+\todo{Line endings (add tildes where necessary)}
 \end{appendix}

### 2020-02-15 paper/paper.tex
Better case distinction in lem:1req

@@
   If no partition of \TAlg's components exist, then it performs no repartition and the cost is $0$.
-  Hence, in the following we bound the cost of moving to component-respecting partition.
+  Hence, in the following we bound the cost of moving to a component-respecting partition.
 
@@
 
-    Consider a request $(u, v)$ and let $U, V$ be clusters of these nodes.
-    If $U=V$, then the request does not require a reconfiguration and \TAlg{} incurs no cost.
-    Otherwise, we consider cases upon the type of clusters $U$ and $V$.
-    Note that this is impossible that either $U$ or $V$ is $C_3$ as otherwise we encounter a component of size $4$ and this would contradict the case assumption that the component-respecting partition exists.
-    If either $U$ or $V$ is $C_1$, then this cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
+    Consider a request $(u, v)$ that triggered the repartitiong and let $U, V$ be clusters of these nodes.
+    Note that $U\neq V$, as such a request cannot trigger a reconfiguration.
+    We consider cases upon the type of clusters $U$ and $V$.
+    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we encounter a component of size $4$ and this would contradict the case assumption that the component-respecting partition exists.
+    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
 
-    Finally, we consider the case where both $U$ and $V$ are $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg{} has the component of size $4$, a contradiction with the case assumption that the component-respecting partition exists.
+    Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component-respecting partition exists.
     If either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
@@
     Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component-respecting partition exists.
-In this case \TAlg{} exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost of $2\cdot \alpha$.
+In this case \TAlg exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost of $2\cdot \alpha$.
 \end{proof}

### 2020-02-15 paper/paper.tex
Better case distinction in lem:1req
Much better name for k=3 ALG

@@
 \newcommand{\PPOBRP}{PP-BRP}
-\newcommand{\TAlg}{{\ensuremath{ALG_{k=3}}}\xspace} % we should change this name
+\newcommand{\TAlg}{{\ensuremath{\textsc{ALG}_{3}}}\xspace} % we should change this name
 
@@
     Note that $U\neq V$, as such a request cannot trigger a reconfiguration.
-    We consider cases upon the type of clusters $U$ and $V$.
+    We consider cases upon types of clusters $U$ and $V$.
     Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we encounter a component of size $4$ and this would contradict the case assumption that the component-respecting partition exists.
@@
 \begin{proof}
-  Fix a completed phase of \TAlg, and consider the state of its counters at the end of it.
+  Fix a completed phase, and consider the state of \TAlg's counters at the end of it.
   We consider the incomplete phase later in this proof.

### 2020-02-15 paper/paper.tex
wip

@@
 \newcommand{\PPOBRP}{PP-BRP}
-\newcommand{\TAlg}{{\ensuremath{\textsc{ALG}_{3}}}\xspace} % we should change this name
+\newcommand{\TAlg}{{\ensuremath{\textsf{ALG}_{3}}}\xspace} % we should change this name
 
@@
   Let $\TAlg(M)$ be the cost of migrations performed by \TAlg in this phase.
-  During a single phase, at most $2\cdot \ell$ component merge operations are performed.
-  Exceeding this number would mean that a component of size $4$ would emerge, a contradiction.
-  Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
+  During a single phase, it performs at most $2\cdot \ell$ component merge operations, as
+  exceeding this number means that a component of size $4$ exists.
+  Combining this with Lemma~\ref{lem:1req} gives us $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
+  %Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
 
@@
   Cluster of type $C_3$ contributes at most $(3\cdot \alpha - 1)$ to $\TAlg(\sigma_I)$, as at most $2$ of pairs of the component are saturated and contribute $\alpha$ each, and the remaining unsaturated pair contributes at most $\alpha-1$.
-  Other cluster configurations contribute less: $C_1$ contributes $0$ and $C_2$ contributes $\alpha$ to $\TAlg(\sigma_I)$.
+  Other cluster types contribute less: $C_1$ contributes $0$ and $C_2$ contributes $\alpha$.
   Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq \ell \cdot (3\cdot \alpha-1)$.
@@
 
-  Now we lower bound the cost of $\OPT$ in this phase.
-  We define $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$, with $\sigma_I$ and $\sigma_E$ defined with respect to components of \TAlg.
-  By $\OPT(M)$ we denote the cost of migrations performed by $\OPT$.
+  Now we lower-bound the cost of $\OPT$.
+  By $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$ we denote the cost of $\OPT$ on these input sequences ($\sigma_I$ and $\sigma_E$ are defined with respect to components of \TAlg in this phase).
+  By $\OPT(M)$ we denote the cost of migrations performed by $\OPT$ in this phase.
 
   We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
-  While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts.
-  Combining those bounds gives us $\OPT \geq \max\{\OPT(\sigma_I) + \OPT(M), \OPT(\sigma_E) + \OPT(M)\} \geq (\OPT(\sigma_I) + \OPT(M) + \OPT(\sigma_E) + \OPT(M)) / 2$.
+  While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts: we separately bound $\OPT$ by $\OPT(\sigma_I) + \OPT(M)$ and $\OPT(\sigma_E) + \OPT(M)$.
+  Combining those bounds gives us $\OPT \geq \max\{\OPT(\sigma_I) + \OPT(M), \OPT(\sigma_E) + \OPT(M)\} \geq (\OPT(\sigma_I) + \OPT(M)) / 2 + (\OPT(\sigma_E) + \OPT(M)) / 2$.
 
@@
   %First, we estimate the cost related to $\sigma_I$.
-  We have $OPT(M) + OPT(\sigma_I) \geq \alpha$, as the components of \TAlg{} could not be partitioned without splitting them, and there exist a pair of nodes with at least $\alpha$ requests between them.
+  We have $OPT(M) + OPT(\sigma_I) \geq \alpha$, as the components of \TAlg{} could not be partitioned without splitting them, and there exists a pair of nodes with at least $\alpha$ requests between them.
   $\OPT$ either served them remotely or performed a migration.

### 2020-02-15 paper/paper.tex
Quick pass

@@
 
-  We partition the input sequence $\sigma$ that arrived during the phase into subsequences $\sigma_I$ and $\sigma_E$, called \emph{internal} and \emph{external}, respectively.
-  The subsequence $\sigma_E$ contains requests to pairs that belong to the same component of \TAlg{} at the end of the phase, and $\sigma_E$ denotes the remaining requests issued during this phase.
+  \maciek{Rewrite the paragraph to emphasize in the definition that requests that did not increase the counter are not included}
+  We distinguish two disjoint subsequences of input $\sigma$ that arrived during the phase: $\sigma_I$ and $\sigma_E$, called \emph{internal} and \emph{external}, respectively.
+  We only account for the requests to nodes that were not collocated by \TAlg at the moment of their arrival, i.e., the requests that increased a counter.
+  The subsequence $\sigma_E$ contains requests to pairs that belong to the same component of \TAlg{} at the end of the phase, and $\sigma_E$ denotes the requests to pairs that did not belong to the same component of \TAlg.
 
@@
   \medskip
+  \maciek{The weakest paragraph}
+  \maciek{x must be a sequence of requests, not the number; better name for x}
   Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to $\sigma_E$.
@@
   It holds that $x \geq \TAlg(\sigma_E) - 3\cdot\ell\cdot(\alpha - 1)$, as it received all the requests that \TAlg did, but it could ignore the cost of serving at most $3\cdot\ell$ pairs of nodes, as it might have them collocated in its clusters.
-  Each such pair of nodes account for at most $\alpha-1$ requests (the pairs from $\sigma_E$ are unsaturated) in the cost of $\TAlg(\sigma_E)$, hence collocating them would decrease the number by at most $3\cdot\ell\cdot(\alpha - 1)$.
+  Each such pair of nodes account for at most $\alpha-1$ requests (the pairs from $\sigma_E$ are unsaturated) in the cost of $\TAlg(\sigma_E)$, hence collocating them would decrease their number by at most $3\cdot\ell\cdot(\alpha - 1)$.
 
-  Now we bound the cost that $\OPT$ pays while serving these $x$ requests.
+  Now we bound the cost incurred by $\OPT$ while serving requests from $x$.
   Instead of serving all of them remotely, $\OPT$ can perform migrations to decrease its cost.
   By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
+  This allows to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
   (express that between a pair of nodes we accounted $\alpha-1$ requests in $x$)
@@
   This gives us
-  \begin{align*}
-    \OPT(\sigma_E) + \OPT(M) & \geq (\TAlg(\sigma_E)-3\cdot\ell\cdot(\alpha - 1)) \cdot \frac{\alpha}{4\cdot (\alpha-1)}\\
-    &\geq \TAlg(\sigma_E) / 4 - 4\cdot \ell \cdot \alpha
-  \end{align*}
+  \begin{equation*}
+    \OPT(\sigma_E) + \OPT(M)  \geq (\TAlg(\sigma_E)-3\cdot\ell\cdot(\alpha - 1)) \cdot \frac{\alpha}{4\cdot (\alpha-1)}
+    \geq \TAlg(\sigma_E) / 4 - 4\cdot \ell \cdot \alpha.
+  \end{equation*}
   From that, we obtain $\TAlg(\sigma_E) \geq 4\cdot(\OPT(\sigma_E)+\OPT(M)) + 16\ell \cdot \alpha$.
-
-  Now we are ready to bound the competitive ratio.
-  Let $E := \OPT(\sigma_E) + \OPT(M)$. Then, using above bounds,
+  With that, we are ready to bound the competitive ratio.
+  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, using above bounds,
   \begin{equation*}
@@
   If the input finished before first $\alpha$ external requests, then \TAlg is optimal.
-
-  If we have at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests
+  Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests
   This way, the competitive ratio increases at most twofold.

### 2020-02-15 paper/paper.tex
minireview of the lower bound

@@
 
-\maciek{emphasize that lower bound works for $k=3$ already. Also in the proof?}
+\maciek{Here, we would like to introduce insights about the proof, emphasize the conditions for the bound to work (such as $k\geq 3$), and maybe compare to previously known lower bounds.
+Also, introduce the notation and definitions that are needed, such as the concepts of components (that we force alg to keep on the same server).}
 
+\maciek{This paragraph should be included in the corollary from PPL, right after its competitive proof.}
 In this section we show a lower bound of $\Omega(k \cdot \ell)$ for \OBRP{}.
@@
 \begin{theorem}
-	The competitive ratio of any deterministic algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$.
+	The competitive ratio of any deterministic algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$ for any $k\geq 3$.
 \end{theorem}
@@
 	Fix any online algorithm \ALG{}.
-	We construct the input sequence of requests for \ALG{} as follows.
+	We construct the input sequence for \ALG{} as follows.
 	First,
@@
 	then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
-	If \ALG{} never collocates them then it is not competitive.
+	Hence, if \ALG{} never collocates them then it is not competitive.
 
+  \maciek{Components are not introduced. Thus, the reader does not know what a single node means.}
 	Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
 	Next,
-	we issue the request between $x_0$ and $y_0$.
-	Since this is an external request,
+	we issue a request between $x_0$ and $y_0$.
+	Since this is an external \maciek{external request is not defined at this point} request,
 	\ALG{} joins them into one component and collocates them in some cluster other than $S_1$.
@@
   until there are only two single nodes left that  originate from the same cluster.
-	I.e.,
-	two single nodes $x^*, y^*,I(x^*) = I(y^*)$ s.t.~for any other pair of single nodes
+	Formally, we denote these two single nodes by $x^*, y^*$, and we have $I(x^*) = I(y^*)$, and~for any other pair of single nodes
 	$x'$ and $y'$,
@@
   If the input finished before first $\alpha$ external requests, then \TAlg is optimal.
-  Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests
+  Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
   This way, the competitive ratio increases at most twofold.

### 2020-02-15 paper/paper.tex
minireview of the lower bound

@@
 \maciek{Here, we would like to introduce insights about the proof, emphasize the conditions for the bound to work (such as $k\geq 3$), and maybe compare to previously known lower bounds.
-Also, introduce the notation and definitions that are needed, such as the concepts of components (that we force alg to keep on the same server).}
+Also, introduce the notation and definitions that are needed.}
+\maciek{We should describe what we mean by  components (that we force alg to keep on the same server).
+We use components differently in $k=3$, but we denote them by components, too.
+This is understandable, but must be emphasized.}
 
@@
 
-In this section we show that the classic \emph{rent-or-buy} approach\cite{borodin-book} allows to obtain an optimal algorithm for $k=3$.
+In this section we show that the classic \emph{rent-or-buy} approach\cite{borodin-book}\maciek{any better source? ski rental?} allows to obtain an optimal algorithm for $k=3$.
 Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one server until the end of an epoch (that we define precisely later).

### 2020-02-15 paper/paper.tex
minireview

@@
 In other words,
-at least $\mathit{dist}(P, P_I)/2$ node swaps are required in order to reach the configuration $P$ from $P_I$.
-Therefore,
+at least $\mathit{dist}(P, P_I)/2$ node swaps are required in order to reach the configuration $P$ from $P_I$, and thus
 $\OPT \geq \Delta:= dist(P_F, P_I) $.
- With each re-partitioning,
+ With each re-partitioning\maciek{Use consistently: repartition or re-partition. IDK which one is better.},
   PPL moves to a configuration that minimizes distance to the initial configuration $P_I$.
@@
     each re-partition costs at most $2\cdot\OPT$.
-    The total cost is therefore at most $2\cdot\OPT\cdot k\cdot\ell$ which implies the competitive ratio.
+    The total cost is therefore at most $2\cdot\OPT\cdot k\cdot\ell$, which implies the competitive ratio.
 %    \mahmoud{This is the cost of moving and the cost of remote comm. is not counted.
@@
 
-\section{Corollary and Future Directions}
+% \section{Corollary and Future Directions}
 
-The gap between upper and lower bound is still substantial, as the best known algorithm is the component-respecting algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\cdot\ell^2)$.
+% The gap between upper and lower bound is still substantial, as the best known algorithm is the component-respecting algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\cdot\ell^2)$.
 
@@
 \begin{appendix}
+
+\section{Ommited proofs}
 
@@
 
-\todo{Check grammar}
+\todo{Spellcheck, grammar (grammarly?)}
 

### 2020-02-15 paper/paper.tex
typo

@@
   We only account for the requests to nodes that were not collocated by \TAlg at the moment of their arrival, i.e., the requests that increased a counter.
-  The subsequence $\sigma_E$ contains requests to pairs that belong to the same component of \TAlg{} at the end of the phase, and $\sigma_E$ denotes the requests to pairs that did not belong to the same component of \TAlg.
+  The subsequence $\sigma_E$ contains requests to pairs that belong to the same component of \TAlg{} at the end of the phase, and $\sigma_I$ denotes the requests to pairs that did not belong to the same component of \TAlg.
 

### 2020-02-15 paper/paper.tex
Grammar issue

@@
 Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one server until the end of an epoch (that we define precisely later).
-To this end, we maintain \emph{components of nodes}, in similar fashion to previous approaches\cite{repartition-disc}.
+To this end, we maintain \emph{components of nodes}, in the similar fashion to previous approaches\cite{repartition-disc}.
 
@@
 
-We define the algorithm \TAlg in the following way.
-\TAlg partition nodes into components.
-Intially, each node is in its own singleton component.
-For each pair of nodes \TAlg maintains a counter, initiated $0$. 
+%We define the algorithm \TAlg in the following way.
+Now we describe the algorithm \TAlg.
+\TAlg partition nodes into components, and
+intially each node belongs to its own singleton component.
+For each pair of nodes \TAlg maintains a counter, initiated to $0$. 
 Upon receiving a request to a pair that is not collocated at one server, it increases the counter by $1$.

### 2020-02-15 paper/paper.tex
Grammar issue

@@
 For each pair of nodes \TAlg maintains a counter, initiated to $0$. 
-Upon receiving a request to a pair that is not collocated at one server, it increases the counter by $1$.
-If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest partition, in which each component has all of its nodes collocated in one server.
-If no such partition exists, \TAlg resets all components to singletons and all counters to $0$ and ends the phase.
+Upon receiving a request to a pair that is not collocated at one server, it increases their counter by $1$.
+If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest partition where each component has all of its nodes collocated in one server.
+If no such partition exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
 

### 2020-02-15 paper/paper.tex
Grammar issue

@@
 \begin{proof}
-  If no partition of \TAlg's components exist, then it performs no repartition and the cost is $0$.
+  If no partition of \TAlg's components exists, then it simply ends the phase and performs no repartition, thus the cost is $0$.
   Hence, in the following we bound the cost of moving to a component-respecting partition.
@@
     Consider a request $(u, v)$ that triggered the repartitiong and let $U, V$ be clusters of these nodes.
-    Note that $U\neq V$, as such a request cannot trigger a reconfiguration.
+    Note that $U\neq V$, as otherwise such a request would not trigger a reconfiguration.
     We consider cases upon types of clusters $U$ and $V$.

### 2020-02-16 paper/paper.tex
Better case distinction
Fixed references

@@
 
-	The best known deterministic algorithm is $O(k^2\cdot \ell^2)$-competitive \cite{?????} \maciek{We should cite \cite{repartition-disc} here.},
+	The best known deterministic algorithm known so far is $O(k^2\cdot \ell^2)$-competitive \cite{repartition-disc},
 	and the best known lower bound for the competitive ratio of
 	deterministic algorithms is $\Omega(k)$,
-	which uses a reduction from the online paging problem.
+ by the reduction from online paging.
 	  In this paper, we provide the lower bound of $\Omega(k\cdot \ell)$.
@@
 \section{An Optimal Algorithm for $k=3$}
+\label{sec:k3}
 
@@
 Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component-respecting partition exists.
-    If either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
+    Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
     Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.

### 2020-02-16 paper/paper.tex
Moving things to right places

@@
 This is understandable, but must be emphasized.}
-
-\maciek{This paragraph should be included in the corollary from PPL, right after its competitive proof.}
-In this section we show a lower bound of $\Omega(k \cdot \ell)$ for \OBRP{}.
-The constructed input sequence allows OPT to move to a perfect partition, and hence
-the construction also constitutes a lower bound for \PPOBRP{}.
-This complements an upper bound of $O(k \cdot \ell)$
-for \PPOBRP{} that we present in Section \ref{sec:ppl}.
 
@@
 
+ In Section~\ref{sec:lowerbound} we constructed a $\Omega(k \cdot \ell)$ for \OBRP{}.
+ Note that the lower bound holds also in the perfect partition model, as the constructed input sequence allows OPT to move to a perfect partition.
+ The corollary is that PPL is optimal.
+
 % \section{Corollary and Future Directions}

### 2020-02-16 paper/paper.tex
Moving things to right places

@@
 the cost of migrations performed by $ALG$ while serving the sequence $\sigma$.
-
-\subsection{The Perfect Partition Variant}	\label{sec:PP}
-In Section \ref{sec:ppl},
-we give an algorithm for a restricted variant of  \OBRP{},
-where the communication graph can be partitioned into $\ell$ clusters without any inter-cluster edges.
-Moreover, we assume an optimal offline algorithm (OPT) moves to this \emph{perfect partition}
-at the beginning and stays there permanently.
-We refer this variant as \PPOBRP{}.
-The task of an online algorithm for \PPOBRP{} is to ``recover" (or learn) the perfect partition while not paying too much more than OPT.
-This implies that on a request between any two non-collocated nodes,
-any online algorithm must eventually collocate them,
-as otherwise it not competitive under a sequence that repeats the same request arbitrarily many times.
-Hence,
-any competitive algorithm,
-must at some point collocate all nodes of a  connected sub-graph of the communication graph.
-We refer to this connected sub-graph as a \emph{component}.
-An algorithm that  keeps all components collocated is \emph{component-respecting}.
 
@@
 
-\section{Recovering a Perfect Partition}	\label{sec:ppl}
+\section{An Optimal Algorithm for Perfect Partition}	\label{sec:ppl}
 %\subsection{Perfect Partition Model}\label{sec:perfectPartition}
@@
 %\subsection{An $O(k\cdot \ell)$-competitive Algorithm for Perfect Partition}
+
+%\subsection{The Perfect Partition Variant}	\label{sec:PP}
+In this section,
+we give an algorithm for a restricted variant of  \OBRP{},
+where the communication graph can be partitioned into $\ell$ clusters without any inter-cluster edges.
+\maciek{This is not sufficient! We also disallow serving requests remotely.}
+
+Moreover, we assume an optimal offline algorithm (OPT) moves to this \emph{perfect partition}
+at the beginning and stays there permanently.
+\maciek{Why do we have to assume that?
+This is just what OPT does, i.e., we can prove that this strategy (a) lower bounds the OPT's cost and (b) is sufficent for our analysis.
+But now is sounds like we restrict the model even further.}
+
+We refer this variant as \PPOBRP{}.
+The task of an online algorithm for \PPOBRP{} is to ``recover" (or learn) the perfect partition while not paying too much more than OPT.\maciek{No quotation marks should be in technical text.}
+
+\maciek{This is not a valid description of the model.
+We need the assumption that collocation is needed in our model.
+Otherwise OPT would just serve some requests remotely, and our analysis would not hold.}
+This implies that on a request between any two non-collocated nodes,
+any online algorithm must eventually collocate them,
+as otherwise it not competitive under a sequence that repeats the same request arbitrarily many times.
+Hence,
+any competitive algorithm,
+must at some point collocate all nodes of a  connected sub-graph of the communication graph.
+We refer to this connected sub-graph as a \emph{component}.
+An algorithm that  keeps all components collocated is \emph{component-respecting}.
+
+\subsection{PPL (working title)}
 

### 2020-02-16 paper/paper.tex
Preliminaries and consistency check
Fixed one of the crucial paragraphs in k=3

@@
 
+\section{Preliminaries}
+
+\maciek{if preliminaries are short, we should merge them with the previous section}
+
+An algorithm that  keeps all components collocated is \emph{component-respecting}.
+
+\maciek{external requests}
+
 \section{Lower bound} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
@@
 
-  \maciek{Rewrite the paragraph to emphasize in the definition that requests that did not increase the counter are not included}
-  We distinguish two disjoint subsequences of input $\sigma$ that arrived during the phase: $\sigma_I$ and $\sigma_E$, called \emph{internal} and \emph{external}, respectively.
-  We only account for the requests to nodes that were not collocated by \TAlg at the moment of their arrival, i.e., the requests that increased a counter.
-  The subsequence $\sigma_E$ contains requests to pairs that belong to the same component of \TAlg{} at the end of the phase, and $\sigma_I$ denotes the requests to pairs that did not belong to the same component of \TAlg.
+  In our analysis, we use two disjoint subsequences of input $\sigma$ that arrived during the phase: $\sigma_I$ and $\sigma_E$.
+  Let $\sigma'$ be the requests that at the moment of arrival were external requests for \TAlg.
+  The sequence $\sigma_I$ (inter-component requests) are the requests from $\sigma'$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
+  The sequence $\sigma_E$ (extra-component requests) are the requests from $\sigma'$ issued to pairs that did not belong to the same component of \TAlg at the end of the phase.
 

### 2020-02-16 paper/paper.tex
improving the PP model definition, addressed some comments

@@
   \label{lem:1req}
-  The cost of migrations during a single repartition of \TAlg is at most $3\cdot\alpha$.
+  The cost of migrations during a single re-partition of \TAlg is at most $3\cdot\alpha$.
 \end{lemma}
@@
 \begin{proof}
-  If no partition of \TAlg's components exists, then it simply ends the phase and performs no repartition, thus the cost is $0$.
+  If no partition of \TAlg's components exists, then it simply ends the phase and performs no re-partition, thus the cost is $0$.
   Hence, in the following we bound the cost of moving to a component-respecting partition.
@@
 
-    Consider a request $(u, v)$ that triggered the repartitiong and let $U, V$ be clusters of these nodes.
+    Consider a request $(u, v)$ that triggered the r-epartitiong and let $U, V$ be clusters of these nodes.
     Note that $U\neq V$, as otherwise such a request would not trigger a reconfiguration.
@@
 
-%\subsection{The Perfect Partition Variant}	\label{sec:PP}
+A \emph{perfect partition} is a partition of nodes into $\ell$ clusters
+such that it incurs no inter-cluster communication throughout the input sequence.
+Equivalently, a perfect partition it an $\ell$-way cut of the communication graph of the input sequence.
 In this section,
 we give an algorithm for a restricted variant of  \OBRP{},
-where the communication graph can be partitioned into $\ell$ clusters without any inter-cluster edges.
-\maciek{This is not sufficient! We also disallow serving requests remotely.}
-
-Moreover, we assume an optimal offline algorithm (OPT) moves to this \emph{perfect partition}
-at the beginning and stays there permanently.
-\maciek{Why do we have to assume that?
+%\maciek{This is not sufficient! We also disallow serving requests remotely.}
+where an optimal offline algorithm (OPT) moves to this perfect partition
+at the beginning and stays there perpetually.
+%\maciek{Why do we have to assume that?
 This is just what OPT does, i.e., we can prove that this strategy (a) lower bounds the OPT's cost and (b) is sufficent for our analysis.
 But now is sounds like we restrict the model even further.}
-
 We refer this variant as \PPOBRP{}.
-The task of an online algorithm for \PPOBRP{} is to ``recover" (or learn) the perfect partition while not paying too much more than OPT.\maciek{No quotation marks should be in technical text.}
-
-\maciek{This is not a valid description of the model.
+The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect partition while not paying too much relative to OPT.
+%
+%\maciek{This is not a valid description of the model.
 We need the assumption that collocation is needed in our model.
 Otherwise OPT would just serve some requests remotely, and our analysis would not hold.}
-This implies that on a request between any two non-collocated nodes,
-any online algorithm must eventually collocate them,
-as otherwise it not competitive under a sequence that repeats the same request arbitrarily many times.
-Hence,
-any competitive algorithm,
-must at some point collocate all nodes of a  connected sub-graph of the communication graph.
-We refer to this connected sub-graph as a \emph{component}.
-An algorithm that  keeps all components collocated is \emph{component-respecting}.
+The first request between any two non-collocated nodes reveals an edge of the communication graph.
+Our online algorithm collocates them immediately and never separates them since by assumption OPT does the same.
+Revealed edges, up to any point in the sequence,
+induce some connected subgraphs and we refer to them as \emph{components}.
+An algorithm that always  keeps nodes belonging to the same component collocated is \emph{component-respecting}.
 
-\subsection{PPL (working title)}
+\subsection{The Algorithm}
 
@@
 $\OPT \geq \Delta:= dist(P_F, P_I) $.
- With each re-partitioning\maciek{Use consistently: repartition or re-partition. IDK which one is better.},
-  PPL moves to a configuration that minimizes distance to the initial configuration $P_I$.
+ With each re-partitioning,
+  PPL moves to a configuration that minimizes the distance to the initial configuration $P_I$.
 As a result,
-ON never ends up in a configuration that is more than $\Delta$ away from $P_I$.
-This invariant ensures that ON does not pay too much while reaching $P_F$.
+ON never ends up in a configuration that is more than $\Delta$ (single-node) migrations away from $P_I$.
+This invariant ensures that ON does not pay too much while recovering $P_F$.
 
@@
 %      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
-      We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a (perfect) partition closest to $P_I$.
+      We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a perfect partition closest to $P_I$.
 Hence it never moves to a configuration beyond distance $\Delta$.      

### 2020-02-16 paper/paper.tex
Refining k=3

--- 2020-02-16 paper/paper.tex
@@
 
-\maciek{external requests}
+We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
 
@@
 Also, introduce the notation and definitions that are needed.}
-\maciek{We should describe what we mean by  components (that we force alg to keep on the same server).
+\maciek{We should describe what we mean by  components (that we force alg to keep on the same cluster).
 We use components differently in $k=3$, but we denote them by components, too.
@@
 In this section we show that the classic \emph{rent-or-buy} approach\cite{borodin-book}\maciek{any better source? ski rental?} allows to obtain an optimal algorithm for $k=3$.
-Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one server until the end of an epoch (that we define precisely later).
+Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one cluster until the end of an epoch (that we define precisely later).
+
 To this end, we maintain \emph{components of nodes}, in the similar fashion to previous approaches\cite{repartition-disc}.
+We say that \TAlg is \emph{component respecting}, i.e., it maintains the invariant that nodes of each component are located in the same cluster.
+\maciek{Define component-respecting partition and use throughout the paper instead of "have components split"}
 
-%We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different servers.
+\maciek{Consistency: collocated at OR collocated in OR collocated on}
 
@@
 Now we describe the algorithm \TAlg.
-\TAlg partition nodes into components, and
+It partitions nodes into components, and
 intially each node belongs to its own singleton component.
-For each pair of nodes \TAlg maintains a counter, initiated to $0$. 
-Upon receiving a request to a pair that is not collocated at one server, it increases their counter by $1$.
-If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest partition where each component has all of its nodes collocated in one server.
+For each pair of nodes, \TAlg maintains a counter, initiated to $0$. 
+Upon receiving a request to a pair that is not collocated at one cluster, it increases their counter by $1$.
+If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest partition where each component has all of its nodes collocated in one cluster.
 If no such partition exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
@@
 %\maciek{Shouldn't we think about whole component cuts instead of single edge? Hmm. NO!}
-
-We say that \TAlg is \emph{component respecting}, i.e., it maintains the invariant that nodes of each component are located at the same server.
-Observe that this implies that \TAlg never increases any counter above $\alpha$.
 
@@
 \begin{proof}
-  If no partition of \TAlg's components exists, then it simply ends the phase and performs no repartition, thus the cost is $0$.
-  Hence, in the following we bound the cost of moving to a component-respecting partition.
+  Observe that when the repartition is triggered by \TAlg, the resulting partition is component-respecting.
+  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
 
@@
 
-    Consider a request $(u, v)$ that triggered the repartitiong and let $U, V$ be clusters of these nodes.
-    Note that $U\neq V$, as otherwise such a request would not trigger a reconfiguration.
+    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
+    Note that $U\neq V$, as otherwise the request would not trigger a~reconfiguration.
     We consider cases upon types of clusters $U$ and $V$.
-    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we encounter a component of size $4$ and this would contradict the case assumption that the component-respecting partition exists.
+    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component-respecting partition exists, a contradiction.
     If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
@@
   We consider the incomplete phase later in this proof.
+
+  As \TAlg is component respecting, it never increases any counter above $\alpha$.
   We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
@@
   We have $OPT(M) + OPT(\sigma_I) \geq \alpha$, as the components of \TAlg{} could not be partitioned without splitting them, and there exists a pair of nodes with at least $\alpha$ requests between them.
-  $\OPT$ either served them remotely or performed a migration.
+  $\OPT$ either served them remotely or performed a~migration.
 

### 2020-02-16 paper/paper.tex

--- 2020-02-16 paper/paper.tex
@@
 \begin{proof}
-  If no partition of \TAlg's components exists, then it simply ends the phase and performs no re-partition, thus the cost is $0$.
-  Hence, in the following we bound the cost of moving to a component-respecting partition.
+  Observe that when the repartition is triggered by \TAlg, the resulting partition is component-respecting.
+  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
 
@@
 
-    Consider a request $(u, v)$ that triggered the r-epartitiong and let $U, V$ be clusters of these nodes.
-    Note that $U\neq V$, as otherwise such a request would not trigger a reconfiguration.
+    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
+    Note that $U\neq V$, as otherwise the request would not trigger a~reconfiguration.
     We consider cases upon types of clusters $U$ and $V$.
-    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we encounter a component of size $4$ and this would contradict the case assumption that the component-respecting partition exists.
+    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component-respecting partition exists, a contradiction.
     If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.

### 2020-02-16 paper/paper.tex
Fix in inequalities

@@
 
-  In our analysis, we use two disjoint subsequences of input $\sigma$ that arrived during the phase: $\sigma_I$ and $\sigma_E$.
-  Let $\sigma'$ be the requests that at the moment of arrival were external requests for \TAlg.
-  The sequence $\sigma_I$ (inter-component requests) are the requests from $\sigma'$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
-  The sequence $\sigma_E$ (extra-component requests) are the requests from $\sigma'$ issued to pairs that did not belong to the same component of \TAlg at the end of the phase.
+  Let $\sigma$ be the input sequence that arrived during the phase.
+  Let $\sigma_{cost}$ be the requests that at the moment of arrival were external requests for \TAlg (these are the only requests that incurred a cost for \TAlg).
+  In our analysis, we partition $\sigma_{cost}$ into subsequences $\sigma_I$ and $\sigma_E$.
+  The sequence $\sigma_I$ (inter-component requests) are the requests from $\sigma_{cost}$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
+  The sequence $\sigma_E$ (extra-component requests) denotes the requests from $\sigma_{cost}$ that do not appear in $\sigma_I$.
 
@@
   Let $\TAlg(M)$ be the cost of migrations performed by \TAlg in this phase.
-  During a single phase, it performs at most $2\cdot \ell$ component merge operations, as
-  exceeding this number means that a component of size $4$ exists.
-  Combining this with Lemma~\ref{lem:1req} gives us $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
+  During a single phase, it performs at most $2 \ell$ component merge operations, as
+  exceeding this number means that a component of size $4$ exists, and the phase should have ended already.
+  Combining this with Lemma~\ref{lem:1req} gives us $\TAlg(M) \leq 6\alpha\cdot\ell$.
   %Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
@@
   Now we bound $\TAlg(\sigma_I)$.
-  Cluster of type $C_3$ contributes at most $(3\cdot \alpha - 1)$ to $\TAlg(\sigma_I)$, as at most $2$ of pairs of the component are saturated and contribute $\alpha$ each, and the remaining unsaturated pair contributes at most $\alpha-1$.
+  Cluster of type $C_3$ contributes at most $(3 \alpha - 1)$ to $\TAlg(\sigma_I)$, as at most $2$ of pairs of nodes from the component are saturated and contribute $\alpha$ each, and the third, unsaturated pair contributes at most $\alpha-1$.
   Other cluster types contribute less: $C_1$ contributes $0$ and $C_2$ contributes $\alpha$.
-  Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq \ell \cdot (3\cdot \alpha-1)$.
+  Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq \ell \cdot (3 \alpha-1) \leq 4\alpha\cdot\ell$.
 
@@
 
-  In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 9\alpha\cdot \ell + \TAlg(\sigma_E)$ during this phase.
+  In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 10\alpha\cdot \ell + \TAlg(\sigma_E)$ during this phase.
 
@@
   \begin{equation*}
-    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{9\alpha \cdot \ell + \TAlg(\sigma_E)}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{50\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 50 \ell = O(\ell).
+    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{10\alpha \cdot \ell + \TAlg(\sigma_E)}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{52\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 52 \ell = O(\ell).
   \end{equation*}

### 2020-02-16 paper/paper.tex
ALG3's cost

@@
   Let $\TAlg(M)$ be the cost of migrations performed by \TAlg in this phase.
-  During a single phase, it performs at most $2 \ell$ component merge operations, as
+  \TAlg performs at most $2 \ell$ component merge operations, as
   exceeding this number means that a component of size $4$ exists, and the phase should have ended already.
@@
   Now we bound $\TAlg(\sigma_I)$.
-  Cluster of type $C_3$ contributes at most $(3 \alpha - 1)$ to $\TAlg(\sigma_I)$, as at most $2$ of pairs of nodes from the component are saturated and contribute $\alpha$ each, and the third, unsaturated pair contributes at most $\alpha-1$.
+  Cluster of type $C_3$ contributes at most $3 \alpha - 1$ to $\TAlg(\sigma_I)$, as $2$ of pairs of nodes from the component are saturated and contribute $\alpha$ each, and the third, unsaturated pair contributes at most $\alpha-1$.
   Other cluster types contribute less: $C_1$ contributes $0$ and $C_2$ contributes $\alpha$.
-  Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq \ell \cdot (3 \alpha-1) \leq 4\alpha\cdot\ell$.
+  Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq (3 \alpha-1)\cdot \ell \leq 4\alpha\cdot\ell$.
 

### 2020-02-16 paper/paper.tex
Bounds on OPT

@@
   Now we lower-bound the cost of $\OPT$.
-  By $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$ we denote the cost of $\OPT$ on these input sequences ($\sigma_I$ and $\sigma_E$ are defined with respect to components of \TAlg in this phase).
+  By $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$ we denote the cost of $\OPT$ on these input sequences (defined with respect to components of \TAlg in this phase).
   By $\OPT(M)$ we denote the cost of migrations performed by $\OPT$ in this phase.
@@
   %First, we estimate the cost related to $\sigma_I$.
-  We have $OPT(M) + OPT(\sigma_I) \geq \alpha$, as the components of \TAlg{} could not be partitioned without splitting them, and there exists a pair of nodes with at least $\alpha$ requests between them.
+  We have $OPT(M) + OPT(\sigma_I) \geq \alpha$, as the phase ended when the components of \TAlg{} could not be partitioned without splitting them.
+  Hence, for every possible configuration of $\OPT$, there exists a non-collocated pair of nodes with at least $\alpha$ requests between them, and
   $\OPT$ either served them remotely or performed a~migration.

### 2020-02-16 paper/paper.tex
An important detail of alg's cost and start to rewrite of the weakest paragraph

@@
 
-  In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 10\alpha\cdot \ell + \TAlg(\sigma_E)$ during this phase.
+  Moreover, \TAlg paid for all requests from $\sigma_E$, and we bound it $\TAlg(\sigma_E) = |\sigma_E|$.
+  In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 10\alpha\cdot \ell + |\sigma_E|$ during this phase.
 
@@
   \maciek{x must be a sequence of requests, not the number; better name for x}
-  Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to $\sigma_E$.
-  We begin by lower-bounding the number of requests from $\sigma_E$ that $\OPT$ must account (by either serving remotely or migrating), and we denote their number by $x$.
+  Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
+
+  In a fixed configuration of $\OPT$, it might mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
+  Recall that $\sigma_E$ consist of requests between unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ might mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
+  Faced with
+
+  We begin by lower-bounding the number of requests from $\sigma_E$ that would incur a cost for $\OPT$ if it .
+
+  may  account (by either serving remotely or migrating), and we denote their number by $x$.
   It holds that $x \geq \TAlg(\sigma_E) - 3\cdot\ell\cdot(\alpha - 1)$, as it received all the requests that \TAlg did, but it could ignore the cost of serving at most $3\cdot\ell$ pairs of nodes, as it might have them collocated in its clusters.

### 2020-02-16 paper/paper.tex
Pesky paragraph rewritten!

@@
   \medskip
-  \maciek{The weakest paragraph}
-  \maciek{x must be a sequence of requests, not the number; better name for x}
   Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
+  In a fixed configuration of $\OPT$, it might mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
+  Recall that $\sigma_E$ consists of requests between unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ might mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
+  Faced with $W := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests $\sigma_E$ that it could not mitigate, $\OPT$ served some of them remotely and possibly performed some migrations to decrease its cost.
 
-  In a fixed configuration of $\OPT$, it might mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
-  Recall that $\sigma_E$ consist of requests between unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ might mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
-  Faced with
-
-  We begin by lower-bounding the number of requests from $\sigma_E$ that would incur a cost for $\OPT$ if it .
-
-  may  account (by either serving remotely or migrating), and we denote their number by $x$.
-  It holds that $x \geq \TAlg(\sigma_E) - 3\cdot\ell\cdot(\alpha - 1)$, as it received all the requests that \TAlg did, but it could ignore the cost of serving at most $3\cdot\ell$ pairs of nodes, as it might have them collocated in its clusters.
-  Each such pair of nodes account for at most $\alpha-1$ requests (the pairs from $\sigma_E$ are unsaturated) in the cost of $\TAlg(\sigma_E)$, hence collocating them would decrease their number by at most $3\cdot\ell\cdot(\alpha - 1)$.
-
-  Now we bound the cost incurred by $\OPT$ while serving requests from $x$.
-  Instead of serving all of them remotely, $\OPT$ can perform migrations to decrease its cost.
+  Now we estimate the cost of $\OPT(\sigma_E)$ by accounting savings from migrations.
   By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
   This allows to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
-  (express that between a pair of nodes we accounted $\alpha-1$ requests in $x$)
-  Hence, by performing a migration for the cost $\alpha$, $\OPT$ does not pay the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $x$ ($x$ is not a sequence but its length!).
-  Finally, for serving $x$, $\OPT$ pays at least $x \cdot \frac{\alpha}{4\cdot (\alpha-1)}$.
+  As $\sigma_E$ consists of requests between unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
+  Thus, by performing a migration for the cost $\alpha$, $\OPT$ does not pay the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $\sigma_E$.
+  Finally, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4\cdot (\alpha-1)}$.
   This gives us

### 2020-02-16 paper/paper.tex
Pesky paragraph

@@
   By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
-  This allows to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
+  This might allow to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
   As $\sigma_E$ consists of requests between unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
-  Thus, by performing a migration for the cost $\alpha$, $\OPT$ does not pay the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $\sigma_E$.
+  Thus, by performing a single migration for the cost $\alpha$, $\OPT$ may not pay the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $\sigma_E$.
   Finally, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4\cdot (\alpha-1)}$.

### 2020-02-16 paper/paper.tex
A pass on k=3
Fix
Moving stuff around

@@
 
+To this end, we maintain \emph{components of nodes}, in the similar fashion to previous approaches\cite{repartition-disc}.
+We say that \TAlg is \emph{component respecting}, i.e., it maintains the invariant that nodes of each component are located in the same cluster.
+
 We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
@@
 
-In this section we show that the classic \emph{rent-or-buy} approach\cite{borodin-book}\maciek{any better source? ski rental?} allows to obtain an optimal algorithm for $k=3$.
-Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one cluster until the end of an epoch (that we define precisely later).
+\maciek{TODO: find better source for rent-or-buy, e.g. ski rental paper}
 
-To this end, we maintain \emph{components of nodes}, in the similar fashion to previous approaches\cite{repartition-disc}.
-We say that \TAlg is \emph{component respecting}, i.e., it maintains the invariant that nodes of each component are located in the same cluster.
-\maciek{Define component-respecting partition and use throughout the paper instead of "have components split"}
-
-\maciek{Consistency: collocated at OR collocated in OR collocated on}
+In this section we show that the classic \emph{rent-or-buy} approach\cite{borodin-book} allows to obtain an optimal algorithm for $k=3$.
+Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one cluster until the end of a phase (that we define precisely later).
 
@@
   \begin{equation*}
-    \OPT(\sigma_E) + \OPT(M)  \geq (\TAlg(\sigma_E)-3\cdot\ell\cdot(\alpha - 1)) \cdot \frac{\alpha}{4\cdot (\alpha-1)}
-    \geq \TAlg(\sigma_E) / 4 - 4\cdot \ell \cdot \alpha.
+    \OPT(\sigma_E) + \OPT(M)  \geq (|\sigma_E|-3\cdot\ell\cdot(\alpha - 1)) \cdot \frac{\alpha}{4\cdot (\alpha-1)}
+    \geq |\sigma_E| / 4 - 4\cdot \ell \cdot \alpha.
   \end{equation*}
-  From that, we obtain $\TAlg(\sigma_E) \geq 4\cdot(\OPT(\sigma_E)+\OPT(M)) + 16\ell \cdot \alpha$.
-  With that, we are ready to bound the competitive ratio.
-  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, using above bounds,
+  From that, we obtain $|\sigma_E| \geq 4\cdot(\OPT(\sigma_E)+\OPT(M)) + 16\ell \cdot \alpha$.
+  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, we are ready to bound the competitive ratio
   \begin{equation*}
-    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{10\alpha \cdot \ell + \TAlg(\sigma_E)}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{52\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 52 \ell = O(\ell).
+    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{10\alpha \cdot \ell + |\sigma_E|}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{52\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 52 \ell = O(\ell).
   \end{equation*}
@@
 
-  Now we consider the last, unfinished phase.
+  It remains to consider the last, unfinished phase.
   Consider the case, where the unfinished phase is also the first one.
@@
 \todo{Line endings (add tildes where necessary)}
+
+\maciek{Consistency: collocated at OR collocated in OR collocated on}
+
 \end{appendix}

### 2020-02-16 paper/paper.tex
Grammar questions

@@
 Upon receiving a request to a pair that is not collocated at one cluster, it increases their counter by $1$.
-If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest partition where each component has all of its nodes collocated in one cluster.
+If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partition.
 If no such partition exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
+\maciek{Serving remotely: say explicitly}
 
@@
 
+\maciek{Component-respecting vs component respecting. In the same way "a lower bound" and "to lower-bound"}
+
 \end{appendix}

### 2020-02-16 paper/paper.tex
Optimizing costs

@@
     Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component-respecting partition exists, a contradiction.
-    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the rebalance is local within $U$ and $V$, for the cost of at most $3$ migrations.
+    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the rebalance is local within $U$ and $V$, and costs at most $3\alpha$.
+    \maciek{at most $2\alpha$ is sufficient}
 
@@
     Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component-respecting partition exists.
-In this case \TAlg exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost of $2\cdot \alpha$.
+In this case \TAlg exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost $2\alpha$.
+    \maciek{1 exchange costs $\alpha$}
 \end{proof}

### 2020-02-16 paper/paper.tex
Starting addressing TODOs

@@
 
-\newcommand{\OPT}{\mathit{OPT}}
+\newcommand{\OPT}{\textsf{OPT}}
 \newcommand{\ALG}{ALG}
@@
 \newcommand{\PPOBRP}{PP-BRP}
+\newcommand{\dist}{\textsf{dist}}
 \newcommand{\TAlg}{{\ensuremath{\textsf{ALG}_{3}}}\xspace} % we should change this name
@@
 
-  Moreover, \TAlg paid for all requests from $\sigma_E$, and we bound it $\TAlg(\sigma_E) = |\sigma_E|$.
+  Moreover, \TAlg paid for all requests from $\sigma_E$, and thus $\TAlg(\sigma_E) = |\sigma_E|$.
   In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 10\alpha\cdot \ell + |\sigma_E|$ during this phase.
@@
     That is,
-    $\mathit{dist}(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$. 
+    $\dist(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$. 
 In other words,
-at least $\mathit{dist}(P, P_I)/2$ node swaps are required in order to reach the configuration $P$ from $P_I$, and thus
-$\OPT \geq \Delta:= dist(P_F, P_I) $.
+at least $\dist(P, P_I)/2$ node swaps are required in order to reach the configuration $P$ from $P_I$, and thus
+$\OPT \geq \Delta:= \dist(P_F, P_I) $.
  With each re-partitioning,
@@
     Let $P$ be any configuration chosen by Algorithm \ref{alg:ppl} at Line $\ref{line:rebalance}$.
-    Then, $\mathit{dist}(P,P_I) \leq \Delta$.
+    Then, $\dist(P,P_I) \leq \Delta$.
 \end{property}
@@
     Since $|M^-|$ nodes are not in their original cluster before the re-partitioning (i.e., in $P_{t-1}$),
-    the distance before the re-partitioning is $\mathit{dist}(P_{t-1},P_I) \geq | M^-|$.
+    the distance before the re-partitioning is $\dist(P_{t-1},P_I) \geq | M^-|$.
     Analogously,
-     the distance afterwards is $\mathit{dist}(P_{t},P_I) \geq | M^+|$.
+     the distance afterwards is $\dist(P_{t},P_I) \geq | M^+|$.
     Thus,
-    $|M| \leq \mathit{dist}(P_{t-1},P_I) + \mathit{dist}(P_{t},P_I)$.
+    $|M| \leq \dist(P_{t-1},P_I) + \dist(P_{t},P_I)$.
     By Property \ref{prop:dist<OPT},
-    $\mathit{dist}(P_{t-1},P_I) , \mathit{dist}(P_{t},P_I) \leq \Delta \leq \OPT$
+    $\dist(P_{t-1},P_I) , \dist(P_{t},P_I) \leq \Delta \leq \OPT$
     and thereby we have	
@@
      That is, 
-     once it reaches a configuration $P$ at distance $\Delta = \mathit{dist} (P, P_I)$,
+     once it reaches a configuration $P$ at distance $\Delta = \dist(P, P_I)$,
      it does not move to a configuration
-     $P', \mathit{dist} (P', P_I) > \Delta$,
+     $P', \dist(P', P_I) > \Delta$,
      before it enumerates all configurations at distance $\Delta$.
@@
 
-\todo{Font for OPT, ALG, DET, dist etc}
-
 \todo{Line endings (add tildes where necessary)}

### 2020-02-16 paper/paper.tex
Addressing TODOs

@@
   %First, we estimate the cost related to $\sigma_I$.
-  We have $OPT(M) + OPT(\sigma_I) \geq \alpha$, as the phase ended when the components of \TAlg{} could not be partitioned without splitting them.
+  We have $\OPT(M) + \OPT(\sigma_I) \geq \alpha$, as the phase ended when the components of \TAlg{} could not be partitioned without splitting them.
   Hence, for every possible configuration of $\OPT$, there exists a non-collocated pair of nodes with at least $\alpha$ requests between them, and
@@
 %\maciek{This is not sufficient! We also disallow serving requests remotely.}
-where an optimal offline algorithm (OPT) moves to this perfect partition
+where an optimal offline algorithm ($\OPT$) moves to this perfect partition
 at the beginning and stays there perpetually.
@@
 We refer this variant as \PPOBRP{}.
-The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect partition while not paying too much relative to OPT.
+The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect partition while not paying too much relative to \OPT.
 %
@@
 The first request between any two non-collocated nodes reveals an edge of the communication graph.
-Our online algorithm collocates them immediately and never separates them since by assumption OPT does the same.
+Our online algorithm collocates them immediately and never separates them since by assumption \OPT does the same.
 Revealed edges, up to any point in the sequence,
@@
 
-We assume OPT begins with the initial configuration
+We assume \OPT begins with the initial configuration
 $P_I = I_1, \dots, I_{\ell}$ and moves to the final partition
@@
  In Section~\ref{sec:lowerbound} we constructed a $\Omega(k \cdot \ell)$ for \OBRP{}.
- Note that the lower bound holds also in the perfect partition model, as the constructed input sequence allows OPT to move to a perfect partition.
+ Note that the lower bound holds also in the perfect partition model, as the constructed input sequence allows \OPT to move to a perfect partition.
  The corollary is that PPL is optimal.

### 2020-02-16 paper/paper.tex
Eliminating past tense

@@
   Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
-  In a fixed configuration of $\OPT$, it might mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
-  Recall that $\sigma_E$ consists of requests between unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ might mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
+  In a fixed configuration of $\OPT$, it may mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
+  Recall that $\sigma_E$ consists of requests between unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
   Faced with $W := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests $\sigma_E$ that it could not mitigate, $\OPT$ served some of them remotely and possibly performed some migrations to decrease its cost.
@@
   By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
-  This might allow to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
+  This may allow to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
   As $\sigma_E$ consists of requests between unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.

### 2020-02-16 paper/paper.tex
wip

@@
   In a fixed configuration of $\OPT$, it may mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
-  Recall that $\sigma_E$ consists of requests between unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
+  Recall that $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
   Faced with $W := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests $\sigma_E$ that it could not mitigate, $\OPT$ served some of them remotely and possibly performed some migrations to decrease its cost.
@@
   This may allow to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
-  As $\sigma_E$ consists of requests between unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
-  Thus, by performing a single migration for the cost $\alpha$, $\OPT$ may not pay the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $\sigma_E$.
+  As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
+  Thus, by performing a single migration that costs $\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $\sigma_E$.
   Finally, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4\cdot (\alpha-1)}$.

### 2020-02-16 paper/paper.tex
wip

@@
   As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
-  Thus, by performing a single migration that costs $\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $\sigma_E$.
-  Finally, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4\cdot (\alpha-1)}$.
+  By performing a single migration that costs $\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $\sigma_E$.
+  Thus, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4\cdot (\alpha-1)}$.
   This gives us

### 2020-02-16 paper/paper.tex
Styling equations

@@
   As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
-  By performing a single migration that costs $\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $\sigma_E$.
-  Thus, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4\cdot (\alpha-1)}$.
+  By performing a single migration that costs $\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
+  Thus, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4 (\alpha-1)}$.
   This gives us
   \begin{equation*}
-    \OPT(\sigma_E) + \OPT(M)  \geq (|\sigma_E|-3\cdot\ell\cdot(\alpha - 1)) \cdot \frac{\alpha}{4\cdot (\alpha-1)}
-    \geq |\sigma_E| / 4 - 4\cdot \ell \cdot \alpha.
+    \OPT(\sigma_E) + \OPT(M)  \geq (|\sigma_E|-3(\alpha - 1)\cdot \ell) \cdot \frac{\alpha}{4 (\alpha-1)}
+    \geq |\sigma_E| / 4 - 4 \alpha \cdot \ell.
   \end{equation*}
-  From that, we obtain $|\sigma_E| \geq 4\cdot(\OPT(\sigma_E)+\OPT(M)) + 16\ell \cdot \alpha$.
+  From that, we obtain $|\sigma_E| \geq 4(\OPT(\sigma_E)+\OPT(M)) + 16\alpha \cdot \ell$.
   Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, we are ready to bound the competitive ratio

### 2020-02-16 paper/paper.tex
Styling PPL

@@
 
-\newcommand{\OPT}{\textsf{OPT}}
-\newcommand{\ALG}{ALG}
+\newcommand{\OPT}{\textsf{OPT}\xspace}
+\newcommand{\ALG}{\textsf{ALG}\xspace}
+\newcommand{\PPL}{\textsf{PPL}\xspace}
 \newcommand{\OBRP}{BRP}
@@
 
-\subsection{The Algorithm}
+\subsection{Perfect Partition Learner (\PPL) Algorithm}
 
@@
  With each re-partitioning,
-  PPL moves to a configuration that minimizes the distance to the initial configuration $P_I$.
+  \PPL moves to a configuration that minimizes the distance to the initial configuration $P_I$.
 As a result,
-ON never ends up in a configuration that is more than $\Delta$ (single-node) migrations away from $P_I$.
-This invariant ensures that ON does not pay too much while recovering $P_F$.
+\PPL never ends up in a configuration that is more than $\Delta$ (single-node) migrations away from $P_I$.
+This invariant ensures that \PPL does not pay too much while recovering $P_F$.
 
@@
     \end{algorithmic}
-    \caption{Perfect Partition Learner (PPL)}
+    \caption{Perfect Partition Learner (\PPL)}
     \label{alg:ppl}
@@
 \begin{theorem}	\label{thm:upperbound}
-    PPL reaches the final configuration $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
+    \PPL reaches the final configuration $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
 \end{theorem}
@@
      Therefore,
-     PPL eventually reaches $\Delta=\OPT$ and the configuration $P_F$.
+     \PPL eventually reaches $\Delta=\OPT$ and the configuration $P_F$.
 %    including the request that completes revealing of all components that are collocated in $P_F$.

### 2020-02-16 paper/paper.tex
Styling PPL
ACM keywords done

@@
 
-\keywords{online algorithms, competitive analysis}
+\keywords{online algorithms, competitive analysis, graph partitioning, clustering}
 
@@
 \begin{property} \label{prop:dist<OPT}
-    Let $P$ be any configuration chosen by Algorithm \ref{alg:ppl} at Line $\ref{line:rebalance}$.
+    Let $P$ be any configuration chosen by \PPL at Line $\ref{line:rebalance}$.
     Then, $\dist(P,P_I) \leq \Delta$.
@@
 
-\todo{ACM keywords are mandatory}
-
 \todo{CCS concepts are mandatory}

### 2020-02-17 paper/paper.tex
Keeping abstract simple

@@
 \begin{abstract}
-  In the online Graph RePartitioning problem,
-  a sequence of communication requests arrives online,
-  between pairs of $n$ nodes that are partitioned into $\ell$ clusters
-  each of size $k$ (i.e., $n = \ell \cdot k$).
-  The objective is to serve these requests while paying a cost that is reasonably close to the optimal cost of serving them given  the whole sequence in advance. 
-  This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
-  such that frequently communicating nodes (from different clusters)
-  are relocated to the same cluster before they inflict too much communication cost.
-  A \emph{reconfiguration} (i.e., re-partitioning) is performed by
-  moving (i.e., migrating) nodes between clusters at a cost.
-  The goal is to jointly minimize communication and migration costs while serving the sequence.
+  In the online Graph RePartitioning problem, we are given a sequence of pairwise communication requests between n nodes, revealed online.
+  The objective is to service these requests by partitioning the nodes into $\ell$ clusters, each of size $k$, such that frequently communicating nodes are located in the same cluster.
+  The partitioning can be updated by migrating nodes between clusters for a fixed cost.
+  The goal is to jointly minimize the amount of inter-cluster communication and migration cost.
 
-	The best known deterministic algorithm known so far is $O(k^2\cdot \ell^2)$-competitive \cite{repartition-disc},
+	The best known deterministic algorithm known so far was $O(k^2\cdot \ell^2)$-competitive \cite{repartition-disc},
 	and the best known lower bound for the competitive ratio of
-	deterministic algorithms is $\Omega(k)$,
- by the reduction from online paging.
+	deterministic algorithms was $\Omega(k)$,
+ by reduction from online paging.
 	  In this paper, we provide the lower bound of $\Omega(k\cdot \ell)$.
@@
 We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
+
+This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
+such that frequently communicating nodes (from different clusters)
+are relocated to the same cluster before they inflict too much communication cost.
+A \emph{reconfiguration} (i.e., re-partitioning) is performed by
+moving (i.e., migrating) nodes between clusters at a cost.
+The goal is to jointly minimize communication and migration costs while serving the sequence.
 

### 2020-02-17 paper/paper.tex
Keeping abstract simple

@@
 
-	The best known deterministic algorithm known so far was $O(k^2\cdot \ell^2)$-competitive \cite{repartition-disc},
-	and the best known lower bound for the competitive ratio of
+  So far, the best known lower bound for the competitive ratio of
 	deterministic algorithms was $\Omega(k)$,
  by reduction from online paging.
-	  In this paper, we provide the lower bound of $\Omega(k\cdot \ell)$.
-    We also present an optimal algorithm for the case of $k=3$.
+  We improve this result significantly by providing the lower bound of $\Omega(k\cdot \ell)$.
+  We follow up with an optimal algorithm for $k=3$.
   Finally,

### 2020-02-17 paper/paper.tex
Intro

@@
 
-Main technical contribution of this paper is the lower bound $\Omega(k\cdot\ell)$ for the competitive ratio of any deterministic algorithm for Online Balanced Partition problem (cf.
-Section~\ref{sec:lowerbound}).
+Main technical contribution of this paper is the lower bound $\Omega(k\cdot\ell)$ (Section~\ref{sec:lowerbound}) for the competitive ratio of any deterministic algorithm for online Balanced RePartition problem..
 The best known lower bound so far was $\Omega(k)$ \cite{repartition-disc}
@@
 So far, the competitive ratio of the best known algorithm was the component-based algorithm with the ratio $O(k^2\cdot \ell^2)$ \cite{repartition-disc}.
-We present an optimal algorithm for $k=3$ in Section~\ref{sec:k3}.
-In Section~\ref{sec:ppl}, we present a $O(k\cdot \ell)$-competitive algorithm for the restricted scenario
-where the communication graph (i.e., request sequence) admits a \emph{perfect partition} (i.e., a multi-way cut of cost zero) that is also an optimal configuration for the offline problem.
-For this setting, the algorithm is optimal, as the lower bound utilized an input sequence that has a perfect partition.
+By applying classic rent-or-buy technique, we obtained an optimal algorithm for $k=3$, that we present in Section~\ref{sec:k3}.
+
+In Section~\ref{sec:ppl}, we study a restricted model of Balanced RePartition, called Perfect Partition.
+For that, we present a $O(k\cdot \ell)$-competitive algorithm that is optimal.
+We believe that this result is an important building block on the way to an optimal algorithm for the general model.
 

### 2020-02-17 paper/paper.tex
Distributed bits in the intro

@@
 \section{Introduction}
+Distributed   applications,  including  batch  processing, streaming, and scale-out databases generate a significant amount of network traffic and a~considerable fraction of their runtime is due to network activity.
+In this  paper, we study  algorithms  for collocating frequently communicating nodes in a distributed networked systems in an online fashion.
 

### 2020-02-17 paper/paper.tex
Distributed bits in the intro

@@
 
-Main technical contribution of this paper is the lower bound $\Omega(k\cdot\ell)$ (Section~\ref{sec:lowerbound}) for the competitive ratio of any deterministic algorithm for online Balanced RePartition problem..
+Main technical contribution of this paper is the lower bound $\Omega(k\cdot\ell)$ for the competitive ratio of any deterministic algorithm for online Balanced RePartition problem (see Section~\ref{sec:lowerbound}).
 The best known lower bound so far was $\Omega(k)$ \cite{repartition-disc}
-which involves only two cluster.
+which involves only two clusters.
 
-So far, the competitive ratio of the best known algorithm was the component-based algorithm with the ratio $O(k^2\cdot \ell^2)$ \cite{repartition-disc}.
+So far, the competitive ratio of the best known algorithm was the component-based algorithm with the ratio $O(k^2\cdot \ell^2)$~\cite{repartition-disc}.
 By applying classic rent-or-buy technique, we obtained an optimal algorithm for $k=3$, that we present in Section~\ref{sec:k3}.

### 2020-02-17 paper/paper.tex
Distributed bits are better in abstracte

@@
 \begin{abstract}
+  Distributed   applications,  including  batch  processing, streaming, and scale-out databases generate a significant amount of network traffic and a~considerable fraction of their runtime is due to network activity.
+In this  paper, we study  algorithms  for collocating frequently communicating nodes in a distributed networked systems in an online fashion.
+
   In the online Graph RePartitioning problem, we are given a sequence of pairwise communication requests between n nodes, revealed online.
@@
 \section{Introduction}
-Distributed   applications,  including  batch  processing, streaming, and scale-out databases generate a significant amount of network traffic and a~considerable fraction of their runtime is due to network activity.
-In this  paper, we study  algorithms  for collocating frequently communicating nodes in a distributed networked systems in an online fashion.
 

### 2020-02-17 paper/paper.tex
Bypassing is an important word

@@
 In Section~\ref{sec:ppl}, we study a restricted model of Balanced RePartition, called Perfect Partition.
-For that, we present a $O(k\cdot \ell)$-competitive algorithm that is optimal.
+For that, we present a $O(k\cdot \ell)$-competitive algorithm, which is optimal.
 We believe that this result is an important building block on the way to an optimal algorithm for the general model.
@@
 
-%The model is related to online
-%paging~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}.
-In our
-setting,
-each  requests between two nodes  not in the same cluster
-(i.e., inter-cluster requests) inflicts a cost until the nodes are placed in the same cluster.
-This is a reminiscent of caching models \emph{with
-rejection} penalty~\cite{EpImLN11,EpImLN15,Irani02}.
+The model is related to online
+caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}. In our
+setting, requests can be served remotely, which is
+reminiscent of caching models \emph{with
+bypassing}~\cite{EpImLN11,EpImLN15,Irani02}.
 

### 2020-02-17 paper/paper.tex
wip
preliminaries

@@
 
-\section{Problem Definition}
+\section{Preliminaries}
 \label{sec:problem-definition}
 
-Formally, the online \emph{Balanced RePartitioning} problem (\OBRP{}) is defined as
+The online \emph{Balanced Re-Partitioning} problem (\OBRP{}) is defined as
 follows. We are given a set of $n$ nodes,
@@
 each of size~$k$.
-We say that nodes in $C \subset V$ are \emph{collocated}
-% in a partition $P$
-if they are assigned to the same cluster.
-%\maciek{We can ommit the formal definition: "I.e.,
-%$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". %However, if we use the notion of cluster later, then TODO: define cluster}
-%\maciek{We can ommit the formal definition: "I.e.,
-%$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". However, if we use the notion of cluster later, then TODO: define cluster}
 The input consists of a sequence of pairwise communication requests
@@
 where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a fixed amount of data.
-Before serving the request,
-the online algorithm can re-partition,
+Nodes in $C \subset V$ are \emph{collocated}
+if they reside in the same cluster.
+An algorithm serves a communication request between two nodes
+either \emph{locally} at cost~0
+if they are collocated,
+ or \emph{remotely} at cost~1
+ if they are located in different clusters.
+ We refer to the two types of requests as \emph{internal}
+ and \emph{external} requests, respectively.
+Before serving a request,
+an online algorithm can \emph{re-partition},
 i.e.,
-it may (optionally) move some of the nodes into clusters different than their current assigned clusters.
-Once the re-partition is executed,
-we serve
+it may (optionally) move some nodes into clusters different than their current assigneclusters while respecting the capacity of every cluster. he re-partition is executed,
+the algorithm serves the  request
 the communication request between two  nodes at cost~0 if they are collocated,
@@
 
-\section{Preliminaries}
-
-\maciek{if preliminaries are short, we should merge them with the previous section}
-
+ useful sentences. This section is TODO}
 An algorithm that  keeps all components collocated is \emph{component-respecting}.
-
-To this end, we maintain \emph{components of nodes}, in the similar fashion to previous approaches\cite{repartition-disc}.
-We say that \TAlg is \emph{component respecting}, i.e., it maintains the invariant that nodes of each component are located in the same cluster.
-
-We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
-
+%
+A \emph{component} is a subset of frequently communicating nodes.   ly,                                                   ,every node is trivially a \emph{singleton} component.
+We grow components similar fashion to previous approaches~\cite{repartition-disc}.
+which proved to be an effective techique for mimi                                                                it always keep.
+That is,
+if the algorithm needs to move a node,
+it moves the whole component containing the node to the sameemph{component -respecting}, i.e., it maintains the invariant that nodes of each component are collocated in the same cluster.
+%We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
+%
 This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
@@
 are relocated to the same cluster before they inflict too much communication cost.
-A \emph{reconfiguration} (i.e., re-partitioning) is performed by
-moving (i.e., migrating) nodes between clusters at a cost.
-The goal is to jointly minimize communication and migration costs while serving the sequence.
+A \emph{reconfiguration} (i.e.,\re-partitioning) is performed by
+ migrating) nodes betwen clusters at a cost.   ,s to jointly minimize communication and migration costs while serving the sequence.
 
@@
 	we issue a request between $x_0$ and $y_0$.
-	Since this is an external \maciek{external request is not defined at this point} request,
+	Since this is an external request,
 	\ALG{} joins them into one component and collocates them in some cluster other than $S_1$.
@@
 Upon receiving a request to a pair that is not collocated at one cluster, it increases their counter by $1$.
-If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partition.
+If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component-respecting partition.
 If no such partition exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
@@
     Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
-    Note that $U\neq V$, as otherwise the request would not trigger a~reconfiguration.
+    Note that $U\neq V$,
+	 as otherwise the request would not trigger a~re-partitioning.
 	 We consider cases upon types of clusters $U$ and $V$.
@@
 
-  As \TAlg is component respecting, it never increases any counter above $\alpha$.
+  As \TAlg is component-respecting, it never increases any counter above $\alpha$.
   We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.

### 2020-02-17 paper/paper.tex
Fix

@@
 
- useful sentences. This section is TODO}
 An algorithm that  keeps all components collocated is \emph{component-respecting}.
@@
 are relocated to the same cluster before they inflict too much communication cost.
-A \emph{reconfiguration} (i.e.,\re-partitioning) is performed by
+A \emph{reconfiguration} (i.e., re-partitioning) is performed by
  migrating) nodes betwen clusters at a cost.   ,s to jointly minimize communication and migration costs while serving the sequence.

### 2020-02-17 paper/paper.tex
preliminaries (1)

@@
 
-The online \emph{Balanced Re-Partitioning} problem (\OBRP{}) is defined as
+The online \emph{Balanced RePartitioning} problem (\OBRP{}) is defined as
 follows. We are given a set of $n$ nodes,
@@
 Before serving a request,
-an online algorithm can \emph{re-partition},
+an online algorithm may \emph{re-partition},
+%(i.e., \emph{reconfigure}).
 i.e.,
-it may (optionally) move some nodes into clusters different than their current assigneclusters while respecting the capacity of every cluster. he re-partition is executed,
-the algorithm serves the  request
-the communication request between two  nodes at cost~0 if they are collocated,
-or at cost~1 if they are located in different clusters.
+it may (optionally) move some nodes into clusters different than their current clusters while respecting the capacity of every cluster. 
+Only after a re-partition is completed,
+the algorithm serves the  request.
 The cost of migrating a node from one cluster to another
@@
 
-An algorithm that  keeps all components collocated is \emph{component-respecting}.
+%\maciek{I just copied useful sentences. This section is TODO}
+%An algorithm that  keeps all components collocated is \emph{component-respecting}.
 %
-A \emph{component} is a subset of frequently communicating nodes.   ly,                                                   ,every node is trivially a \emph{singleton} component.
-We grow components similar fashion to previous approaches~\cite{repartition-disc}.
-which proved to be an effective techique for mimi                                                                it always keep.
+A \emph{component} is a subset of frequently communicating nodes.
+Initially,
+every node is trivially a \emph{singleton} component.
+We grow components similar to previous approaches~\cite{repartition-disc},
+which proved to be an effective technique for mimicking an optimal offline algorithm.
+An algorithm is \emph{component-respecting}
+if it always keeps nodes that  belong to the same component collocated.
 That is,
 if the algorithm needs to move a node,
-it moves the whole component containing the node to the sameemph{component -respecting}, i.e., it maintains the invariant that nodes of each component are collocated in the same cluster.
+it moves the whole component containing the node.
 %We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
-%
-This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
-such that frequently communicating nodes (from different clusters)
-are relocated to the same cluster before they inflict too much communication cost.
-A \emph{reconfiguration} (i.e., re-partitioning) is performed by
- migrating) nodes betwen clusters at a cost.   ,s to jointly minimize communication and migration costs while serving the sequence.
+%This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
+%such that frequently communicating nodes (from different clusters)
+%are relocated to the same cluster before they inflict too much communication cost.
+%A \emph{reconfiguration} (i.e.,\emph{ re-partitioning})
+%is performed by migrating nodes between clusters,
+%at the total cost of individual node migrations.
+The objective is to jointly minimize communication and migration costs while serving the sequence.
 

### 2020-02-17 paper/paper.tex
partition VS partitioning

@@
 
-  In the online Graph RePartitioning problem, we are given a sequence of pairwise communication requests between n nodes, revealed online.
+  In the online graph re-partitioning problem, we are given a sequence of pairwise communication requests between n nodes, revealed online.
   The objective is to service these requests by partitioning the nodes into $\ell$ clusters, each of size $k$, such that frequently communicating nodes are located in the same cluster.
@@
   we provide a $O(k\cdot \ell)$-competitive algorithm for the variant
-  where the optimal configuration is always a fixed partition
+  where the optimal configuration is always a fixed partitioning (a.k.a~a \emph{configuration})
   that incurs zero communication cost through the whole request sequence.
@@
 
-Main technical contribution of this paper is the lower bound $\Omega(k\cdot\ell)$ for the competitive ratio of any deterministic algorithm for online Balanced RePartition problem (see Section~\ref{sec:lowerbound}).
+Main technical contribution of this paper is the lower bound $\Omega(k\cdot\ell)$ for the competitive ratio of any deterministic algorithm for online balanced re-partition problem (see Section~\ref{sec:lowerbound}).
 The best known lower bound so far was $\Omega(k)$ \cite{repartition-disc}
@@
 
-In Section~\ref{sec:ppl}, we study a restricted model of Balanced RePartition, called Perfect Partition.
+In Section~\ref{sec:ppl}, we study a restricted model of balanced re-partition, called perfect partition.
 For that, we present a $O(k\cdot \ell)$-competitive algorithm, which is optimal.
@@
 
-The online \emph{Balanced RePartitioning} problem (\OBRP{}) is defined as
+The online \emph{balanced re-partitioning} problem (\OBRP{}) is defined as
 follows. We are given a set of $n$ nodes,
@@
 it may (optionally) move some nodes into clusters different than their current clusters while respecting the capacity of every cluster. 
-Only after a re-partition is completed,
+Only after a re-partitioning is completed,
 the algorithm serves the  request.
@@
 It partitions nodes into components, and
-intially each node belongs to its own singleton component.
+initially each node belongs to its own singleton component.
 For each pair of nodes, \TAlg maintains a counter, initiated to $0$. 
 Upon receiving a request to a pair that is not collocated at one cluster, it increases their counter by $1$.
-If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component-respecting partition.
-If no such partition exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
+If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component-respecting partitioning.
+If no such partitioning exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
 \maciek{Serving remotely: say explicitly}
@@
 \begin{proof}
-  Observe that when the repartition is triggered by \TAlg, the resulting partition is component-respecting.
-  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
+  Observe that when the re-partition is triggered by \TAlg, the resulting partitioning is component-respecting.
+  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no re-partition.
 
@@
 
-    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
+    Consider a request between $u$ and $v$ that triggered the re-partition and let $U$ and $V$ be their respective clusters.
     Note that $U\neq V$,
@@
 	 We consider cases upon types of clusters $U$ and $V$.
-    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component-respecting partition exists, a contradiction.
+    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component-respecting partitioning exists, a contradiction.
     If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the rebalance is local within $U$ and $V$, and costs at most $3\alpha$.
@@
     Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component-respecting partition exists.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component-respecting partitioning exists.
     Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
     Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component-respecting partition exists.
+    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component-respecting partitioning exists.
 In this case \TAlg exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost $2\alpha$.
@@
 
-A \emph{perfect partition} is a partition of nodes into $\ell$ clusters
+A \emph{perfect partition} is a partitioning of nodes into $\ell$ clusters
 such that it incurs no inter-cluster communication throughout the input sequence.
-Equivalently, a perfect partition it an $\ell$-way cut of the communication graph of the input sequence.
+Equivalently, a perfect partitioning it an $\ell$-way cut of the communication graph of the input sequence.
 In this section,
@@
 We assume \OPT begins with the initial configuration
-$P_I = I_1, \dots, I_{\ell}$ and moves to the final partition
+$P_I = I_1, \dots, I_{\ell}$ and moves to the final partitioning
 $P_F = F_1, \dots, F_{\ell}$.

### 2020-02-17 paper/paper.tex
perfect-partition

@@
 
-In Section~\ref{sec:ppl}, we study a restricted model of balanced re-partition, called perfect partition.
+In Section~\ref{sec:ppl}, we study a restricted model of balanced re-partition, called perfect-partition.
 For that, we present a $O(k\cdot \ell)$-competitive algorithm, which is optimal.
@@
 
-\section{An Optimal Algorithm for Perfect Partition}	\label{sec:ppl}
+\section{An Optimal Algorithm for Perfect-Partition}	\label{sec:ppl}
 %\subsection{Perfect Partition Model}\label{sec:perfectPartition}
@@
 
-A \emph{perfect partition} is a partitioning of nodes into $\ell$ clusters
+A \emph{perfect-partition} is a partitioning of nodes into $\ell$ clusters
 such that it incurs no inter-cluster communication throughout the input sequence.
-Equivalently, a perfect partitioning it an $\ell$-way cut of the communication graph of the input sequence.
+Equivalently, a perfect-partitioning it an $\ell$-way cut of the communication graph of the input sequence.
 In this section,
@@
 %\maciek{This is not sufficient! We also disallow serving requests remotely.}
-where an optimal offline algorithm ($\OPT$) moves to this perfect partition
+where an optimal offline algorithm ($\OPT$) moves to this perfect-partition
 at the beginning and stays there perpetually.
@@
 We refer this variant as \PPOBRP{}.
-The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect partition while not paying too much relative to \OPT.
+The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect-partition while not paying too much relative to \OPT.
 %
@@
 
-\subsection{Perfect Partition Learner (\PPL) Algorithm}
+\subsection{Perfect-Partition Learner (\PPL) Algorithm}
 
@@
     \end{algorithmic}
-    \caption{Perfect Partition Learner (\PPL)}
+    \caption{Perfect-Partition Learner (\PPL)}
     \label{alg:ppl}
@@
 %      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
-      We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a perfect partition closest to $P_I$.
+      We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a perfect-partition closest to $P_I$.
 Hence it never moves to a configuration beyond distance $\Delta$.      
@@
  In Section~\ref{sec:lowerbound} we constructed a $\Omega(k \cdot \ell)$ for \OBRP{}.
- Note that the lower bound holds also in the perfect partition model, as the constructed input sequence allows \OPT to move to a perfect partition.
+ Note that the lower bound holds also in the perfect-partition model, as the constructed input sequence allows \OPT to move to a perfect-partition.
  The corollary is that PPL is optimal.

### 2020-02-17 paper/paper.tex
lem:1req optimization

--- 2020-02-17 paper/paper.tex
@@
 
-\section{An Optimal Algorithm for $k=3$}
+\section{An Optimal Algorithm for Clusters of Size 3}
 \label{sec:k3}
@@
   \label{lem:1req}
-  The cost of migrations during a single re-partition of \TAlg is at most $3\cdot\alpha$.
+  A single re-partition of \TAlg consists of at most $2$ migrations.
+  \maciek{todo:propagate the optimization of cost}
 \end{lemma}
@@
     Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component-respecting partition exists, a contradiction.
-    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the rebalance is local within $U$ and $V$, and costs at most $3\alpha$.
-    \maciek{at most $2\alpha$ is sufficient}
+    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
+    For two clusters, any configuration can be reached within two migrations, due to the fact that clusters are indistiguishable.
 
@@
     Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component-respecting partition exists.
-In this case \TAlg exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost $2\alpha$.
-    \maciek{1 exchange costs $\alpha$}
+In this case \TAlg performs one migration, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
 \end{proof}

### 2020-02-17 paper/paper.tex

--- 2020-02-17 paper/paper.tex
@@
 	 We consider cases upon types of clusters $U$ and $V$.
-    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component-respecting partitioning exists, a contradiction.
-    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the rebalance is local within $U$ and $V$, and costs at most $3\alpha$.
-    \maciek{at most $2\alpha$ is sufficient}
+    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component-respecting partition exists, a contradiction.
+    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
+    For two clusters, any configuration can be reached within two migrations, due to the fact that clusters are indistiguishable.
 
@@
     Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component-respecting partitioning exists.
-In this case \TAlg exchanges the nodes $u$ and $v$ with any two nodes of $W$, for the total cost $2\alpha$.
-    \maciek{1 exchange costs $\alpha$}
+    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component-respecting partition exists.
+In this case \TAlg performs one migration, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
 \end{proof}

### 2020-02-17 paper/paper.tex
No private e-mail
Propagating optimization

@@
 \author{Maciej Pacut}
-\email{maciek.pacut@gmail.com}
+\email{maciej.pacut@univie.ac.at}
 \orcid{0000-0002-6379-1490}
@@
 If no such partitioning exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
-\maciek{Serving remotely: say explicitly}
 
@@
   A single re-partition of \TAlg consists of at most $2$ migrations.
-  \maciek{todo:propagate the optimization of cost}
 \end{lemma}
@@
   exceeding this number means that a component of size $4$ exists, and the phase should have ended already.
-  Combining this with Lemma~\ref{lem:1req} gives us $\TAlg(M) \leq 6\alpha\cdot\ell$.
+  Combining this with Lemma~\ref{lem:1req} gives us $\TAlg(M) \leq 4\alpha\cdot\ell$.
   %Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
@@
   Other cluster types contribute less: $C_1$ contributes $0$ and $C_2$ contributes $\alpha$.
-  Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq (3 \alpha-1)\cdot \ell \leq 4\alpha\cdot\ell$.
+  Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq (3 \alpha-1)\cdot \ell \leq 3\alpha\cdot\ell$.
 
@@
   Moreover, \TAlg paid for all requests from $\sigma_E$, and thus $\TAlg(\sigma_E) = |\sigma_E|$.
-  In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 10\alpha\cdot \ell + |\sigma_E|$ during this phase.
+  In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 7\alpha\cdot \ell + |\sigma_E|$ during this phase.
 
@@
   \begin{equation*}
-    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{10\alpha \cdot \ell + |\sigma_E|}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{52\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 52 \ell = O(\ell).
+    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{7\alpha \cdot \ell + |\sigma_E|}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{46\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 46 \ell = O(\ell).
   \end{equation*}

### 2020-02-17 paper/paper.tex
Font consistency

@@
 is~$\alpha \in \mathbb{Z}^+$.
-For any algorithm $ALG$,
+For any algorithm $\ALG$,
 its cost,
-denoted by $ALG(\sigma)$,
+denoted by $\ALG(\sigma)$,
 is the total cost of communications and
-the cost of migrations performed by $ALG$ while serving the sequence $\sigma$.
+the cost of migrations performed by $\ALG$ while serving the sequence $\sigma$.
 
@@
 Also, introduce the notation and definitions that are needed.}
-\maciek{We should describe what we mean by  components (that we force alg to keep on the same cluster).
-We use components differently in $k=3$, but we denote them by components, too.
-This is understandable, but must be emphasized.}
 
@@
 
-  \maciek{Components are not introduced. Thus, the reader does not know what a single node means.}
 	Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.

### 2020-02-17 paper/paper.tex
Preliminaries pass

@@
 the cost of migrations performed by $\ALG$ while serving the sequence $\sigma$.
+The objective is to minimize the total incurred cost.
 
-%\maciek{I just copied useful sentences. This section is TODO}
-%An algorithm that  keeps all components collocated is \emph{component-respecting}.
-%
+In this paper, we use the techinique of maintaining \emph{components} of nodes, similar to previous approaches~\cite{repartition-disc}.
 A \emph{component} is a subset of frequently communicating nodes.
@@
 every node is trivially a \emph{singleton} component.
-We grow components similar to previous approaches~\cite{repartition-disc},
-which proved to be an effective technique for mimicking an optimal offline algorithm.
-An algorithm is \emph{component-respecting}
+An algorithm is \emph{component respecting}
 if it always keeps nodes that  belong to the same component collocated.
@@
 %at the total cost of individual node migrations.
-The objective is to jointly minimize communication and migration costs while serving the sequence.
+%The objective is to jointly minimize communication and migration costs while serving the sequence.
 

### 2020-02-17 paper/paper.tex
Ski rental cite

@@
 
-\maciek{Here, we would like to introduce insights about the proof, emphasize the conditions for the bound to work (such as $k\geq 3$), and maybe compare to previously known lower bounds.
+%\maciek{Here, we would like to introduce insights about the proof, emphasize the conditions for the bound to work (such as $k\geq 3$), and maybe compare to previously known lower bounds.
 Also, introduce the notation and definitions that are needed.}
@@
 	$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
-	Let $I(C)$ denote the cluster where nodes in a component $C$ are located initially.
+	Let $I(C)$ denote the cluster where nodes of a component $C$ are located initially.
 	%	\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
@@
 
-\maciek{TODO: find better source for rent-or-buy, e.g. ski rental paper}
-
-In this section we show that the classic \emph{rent-or-buy} approach\cite{borodin-book} allows to obtain an optimal algorithm for $k=3$.
+In this section we show that the classic \emph{rent-or-buy} approach\cite{borodin-book, karlin-ski-rental} allows to obtain an optimal algorithm for $k=3$.
 Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one cluster until the end of a phase (that we define precisely later).

### 2020-02-17 paper/paper.tex
unneeded hyphens

@@
 
-  In the online graph re-partitioning problem, we are given a sequence of pairwise communication requests between n nodes, revealed online.
+  In the online graph repartitioning problem, we are given a sequence of pairwise communication requests between n nodes, revealed online.
   The objective is to service these requests by partitioning the nodes into $\ell$ clusters, each of size $k$, such that frequently communicating nodes are located in the same cluster.
@@
 
-Main technical contribution of this paper is the lower bound $\Omega(k\cdot\ell)$ for the competitive ratio of any deterministic algorithm for online balanced re-partition problem (see Section~\ref{sec:lowerbound}).
+Main technical contribution of this paper is the lower bound $\Omega(k\cdot\ell)$ for the competitive ratio of any deterministic algorithm for online balanced repartition problem (see Section~\ref{sec:lowerbound}).
 The best known lower bound so far was $\Omega(k)$ \cite{repartition-disc}
@@
 
-In Section~\ref{sec:ppl}, we study a restricted model of balanced re-partition, called perfect-partition.
+In Section~\ref{sec:ppl}, we study a restricted model of balanced repartition, called perfect partition.
 For that, we present a $O(k\cdot \ell)$-competitive algorithm, which is optimal.
@@
 
-The online \emph{balanced re-partitioning} problem (\OBRP{}) is defined as
+The online \emph{balanced repartitioning} problem (\OBRP{}) is defined as
 follows. We are given a set of $n$ nodes,
@@
 Before serving a request,
-an online algorithm may \emph{re-partition},
+an online algorithm may \emph{repartition},
 %(i.e., \emph{reconfigure}).
@@
 it may (optionally) move some nodes into clusters different than their current clusters while respecting the capacity of every cluster. 
-Only after a re-partitioning is completed,
+Only after a repartitioning is completed,
 the algorithm serves the  request.
@@
 Upon receiving a request to a pair that is not collocated at one cluster, it increases their counter by $1$.
-If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component-respecting partitioning.
+If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
 If no such partitioning exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
@@
   \label{lem:1req}
-  A single re-partition of \TAlg consists of at most $2$ migrations.
+  A single repartition of \TAlg consists of at most $2$ migrations.
 \end{lemma}
@@
 \begin{proof}
-  Observe that when the re-partition is triggered by \TAlg, the resulting partitioning is component-respecting.
-  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no re-partition.
+  Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
+  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
 
@@
 
-    Consider a request between $u$ and $v$ that triggered the re-partition and let $U$ and $V$ be their respective clusters.
+    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
     Note that $U\neq V$,
-	 as otherwise the request would not trigger a~re-partitioning.
+	 as otherwise the request would not trigger a~repartitioning.
 	 We consider cases upon types of clusters $U$ and $V$.
-    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component-respecting partition exists, a contradiction.
+    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
     If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
@@
     Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component-respecting partitioning exists.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
     Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
     Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component-respecting partition exists.
+    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
 In this case \TAlg performs one migration, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
@@
 
-  As \TAlg is component-respecting, it never increases any counter above $\alpha$.
+  As \TAlg is component respecting, it never increases any counter above $\alpha$.
   We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
@@
 
-\section{An Optimal Algorithm for Perfect-Partition}	\label{sec:ppl}
+\section{An Optimal Algorithm for Perfect Partition}	\label{sec:ppl}
 %\subsection{Perfect Partition Model}\label{sec:perfectPartition}
@@
 
-A \emph{perfect-partition} is a partitioning of nodes into $\ell$ clusters
+A \emph{perfect partition} is a partitioning of nodes into $\ell$ clusters
 such that it incurs no inter-cluster communication throughout the input sequence.
-Equivalently, a perfect-partitioning it an $\ell$-way cut of the communication graph of the input sequence.
+Equivalently, a perfect partitioning it an $\ell$-way cut of the communication graph of the input sequence.
 In this section,
@@
 %\maciek{This is not sufficient! We also disallow serving requests remotely.}
-where an optimal offline algorithm ($\OPT$) moves to this perfect-partition
+where an optimal offline algorithm ($\OPT$) moves to this perfect partition
 at the beginning and stays there perpetually.
@@
 We refer this variant as \PPOBRP{}.
-The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect-partition while not paying too much relative to \OPT.
+The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect partition while not paying too much relative to \OPT.
 %
@@
 induce some connected subgraphs and we refer to them as \emph{components}.
-An algorithm that always  keeps nodes belonging to the same component collocated is \emph{component-respecting}.
+An algorithm that always  keeps nodes belonging to the same component collocated is \emph{component respecting}.
 
-\subsection{Perfect-Partition Learner (\PPL) Algorithm}
+\subsection{Perfect Partition Learner (\PPL) Algorithm}
 
@@
 $\OPT \geq \Delta:= \dist(P_F, P_I) $.
- With each re-partitioning,
+ With each repartitioning,
   \PPL moves to a configuration that minimizes the distance to the initial configuration $P_I$.
@@
     \end{algorithmic}
-    \caption{Perfect-Partition Learner (\PPL)}
+    \caption{Perfect Partition Learner (\PPL)}
     \label{alg:ppl}
@@
 %      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
-      We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a perfect-partition closest to $P_I$.
+      We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a perfect partition closest to $P_I$.
 Hence it never moves to a configuration beyond distance $\Delta$.      
@@
 \begin{lemma}	\label{lemma:rebalancecost}
-    The cost of re-partitioning at Line \ref{line:rebalance} is at most $2\cdot\OPT$.
+    The cost of repartitioning at Line \ref{line:rebalance} is at most $2\cdot\OPT$.
 \end{lemma}
 \begin{proof}
-    Consider the re-partitioning that transforms $P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
+    Consider the repartitioning that transforms $P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
     Let $M \subset V$ denote the set of nodes that migrate during this process.
 	Let $M^-$ and $M^+$ denote the subset of nodes that (respectively)
-    enter or leave their original cluster during the re-partitioning.    
+    enter or leave their original cluster during the repartitioning.    
     Then,
     $M = M^+ \cup M^-$.
-    Since $|M^-|$ nodes are not in their original cluster before the re-partitioning (i.e., in $P_{t-1}$),
-    the distance before the re-partitioning is $\dist(P_{t-1},P_I) \geq | M^-|$.
+    Since $|M^-|$ nodes are not in their original cluster before the repartitioning (i.e., in $P_{t-1}$),
+    the distance before the repartitioning is $\dist(P_{t-1},P_I) \geq | M^-|$.
     Analogously,
@@
     By Lemma \ref{lemma:rebalancecost},
-    each re-partition costs at most $2\cdot\OPT$.
+    each repartition costs at most $2\cdot\OPT$.
     The total cost is therefore at most $2\cdot\OPT\cdot k\cdot\ell$, which implies the competitive ratio.
@@
  In Section~\ref{sec:lowerbound} we constructed a $\Omega(k \cdot \ell)$ for \OBRP{}.
- Note that the lower bound holds also in the perfect-partition model, as the constructed input sequence allows \OPT to move to a perfect-partition.
+ Note that the lower bound holds also in the perfect partition model, as the constructed input sequence allows \OPT to move to a perfect partition.
  The corollary is that PPL is optimal.
@@
 
-\maciek{Component-respecting vs component respecting. In the same way "a lower bound" and "to lower-bound"}
+\maciek{Component respecting vs component respecting. In the same way "a lower bound" and "to lower-bound"}
 

### 2020-02-17 paper/paper.tex
Related work

--- 2020-02-17 paper/paper.tex
@@
 
-The problem was studied under resource augmentation \cite{repartition-disc}, and under stochastic assumptions of input\maciek{cite}.\maciek{cite other papers that cite DISC?}
+The problem was studied under resource augmentation \cite{repartition-disc}, and under stochastic assumptions for the input\cite{stochastic-ring}.
+For clusters of size $2$ (i.e., $k=2$), a constant competitive algorithm exists~\cite{repartition-disc}.
 
@@
 \label{sec:lowerbound}
+
+The adversary constructs components of nodes, and $\OPT$ never splits them.
+The adversary observes the actions of an online algorithm, and if it splits the components at any moment, the adversary issues
 

### 2020-02-17 paper/paper.tex
Cleanup

@@
 
-\todo{Check naming consistency inside sections, and across them. Possibly contain some definitions in preliminaries.}
-
 \todo{ACM format for review? Is there any specific format like this? With line numbers etc.}
-
-\todo{Include e-mails? Or shall we skip it for now.}
-
-\todo{Check ACM package whitelist}
-
-\todo{Replace package algorithmpseudocode, it is not whitelisted https://www.acm.org/publications/taps/whitelist-of-latex-packages}
-
-\todo{Make an ACM account for the submission}
 
@@
 
-\todo{Spellcheck, grammar (grammarly?)}
-
-\todo{Make sure our manuscript is autor anonymous}
-
 \todo{Line endings (add tildes where necessary)}
 
-\maciek{Consistency: collocated at OR collocated in OR collocated on}
-
-\maciek{Component respecting vs component respecting. In the same way "a lower bound" and "to lower-bound"}
+\todo{Consistency: collocated at OR collocated in OR collocated on}
 

### 2020-02-17 paper/paper.tex
No definitions in abstract

@@
   we provide a $O(k\cdot \ell)$-competitive algorithm for the variant
-  where the optimal configuration is always a fixed partitioning (a.k.a~a \emph{configuration})
+  where the optimal strategy is to move to a fixed configuration
   that incurs zero communication cost through the whole request sequence.
@@
 if they are located in different clusters.
-We refer to the two types of requests as \emph{internal}
+We refer to these two types of requests as \emph{internal}
 and \emph{external} requests, respectively.
@@
 
-\todo{ACM format for review? Is there any specific format like this? With line numbers etc.}
+\todo{ACM format for review? Is there any specific format like this? Author anonymous, with line numbers etc.}
 

### 2020-02-17 paper/paper.tex
Tildes

@@
   The objective is to service these requests by partitioning the nodes into $\ell$ clusters, each of size $k$, such that frequently communicating nodes are located in the same cluster.
-  The partitioning can be updated by migrating nodes between clusters for a fixed cost.
+  The partitioning can be updated by migrating nodes between clusters for a~fixed cost.
   The goal is to jointly minimize the amount of inter-cluster communication and migration cost.
@@
   we provide a $O(k\cdot \ell)$-competitive algorithm for the variant
-  where the optimal strategy is to move to a fixed configuration
+  where the optimal strategy is to move to a~fixed configuration
   that incurs zero communication cost through the whole request sequence.
@@
 In Section~\ref{sec:ppl}, we study a restricted model of balanced repartition, called perfect partition.
-For that, we present a $O(k\cdot \ell)$-competitive algorithm, which is optimal.
+For that, we present an~$O(k\cdot \ell)$-competitive algorithm, which is optimal.
 We believe that this result is an important building block on the way to an optimal algorithm for the general model.
@@
 $\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
-where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a fixed amount of data.
+where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a~fixed amount of data.
 Nodes in $C \subset V$ are \emph{collocated}
@@
 In this paper, we use the techinique of maintaining \emph{components} of nodes, similar to previous approaches~\cite{repartition-disc}.
-A \emph{component} is a subset of frequently communicating nodes.
+A~\emph{component} is a subset of frequently communicating nodes.
 Initially,
 every node is trivially a \emph{singleton} component.
-An algorithm is \emph{component respecting}
+An~algorithm is \emph{component respecting}
 if it always keeps nodes that  belong to the same component collocated.
@@
   Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
-  In a fixed configuration of $\OPT$, it may mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
+  In a~fixed configuration of $\OPT$, it may mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
   Recall that $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
@@
 
-\todo{Line endings (add tildes where necessary)}
-
 \todo{Consistency: collocated at OR collocated in OR collocated on}

### 2020-02-17 paper/paper.tex
one pass

@@
 By applying classic rent-or-buy technique, we obtained an optimal algorithm for $k=3$, that we present in Section~\ref{sec:k3}.
-
 In Section~\ref{sec:ppl}, we study a restricted model of balanced repartition, called perfect partition.
@@
 We believe that this result is an important building block on the way to an optimal algorithm for the general model.
+\mahmoud{The mention of distributed alg, very much out of context!} 
+A distributed adaptation of a similar algorithm has been analyzed in \cite{sigmetrics19_partitioning}.
 
@@
 
-In this paper, we use the techinique of maintaining \emph{components} of nodes, similar to previous approaches~\cite{repartition-disc}.
-A~\emph{component} is a subset of frequently communicating nodes.
+In this paper, we use the technique of maintaining \emph{components} of nodes, similar to previous approaches~\cite{repartition-disc}. A~\emph{component} is a subset of frequently communicating nodes.
 Initially,
@@
 For each pair of nodes, \TAlg maintains a counter, initiated to $0$. 
-Upon receiving a request to a pair that is not collocated at one cluster, it increases their counter by $1$.
+Upon receiving a request to a pair that is not collocated in one cluster, it increases their counter by $1$.
 If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
@@
 
-A \emph{perfect partition} is a partitioning of nodes into $\ell$ clusters
-such that it incurs no inter-cluster communication throughout the input sequence.
-Equivalently, a perfect partitioning it an $\ell$-way cut of the communication graph of the input sequence.
+%A \emph{perfect partition} is a partitioning of nodes into $\ell$ clusters
+%such that it incurs no inter-cluster communication throughout the input sequence.
+%Equivalently, a perfect partitioning it an $\ell$-way cut of the communication graph of the input sequence.
 In this section,
-we give an algorithm for a restricted variant of  \OBRP{},
+we give an algorithm for a restricted variant of  \OBRP{}, named \PPOBRP{},
 %\maciek{This is not sufficient! We also disallow serving requests remotely.}
-where an optimal offline algorithm ($\OPT$) moves to this perfect partition
+where an optimal offline algorithm ($\OPT$) moves to a perfect partition
 at the beginning and stays there perpetually.
-%\maciek{Why do we have to assume that?
-This is just what OPT does, i.e., we can prove that this strategy (a) lower bounds the OPT's cost and (b) is sufficent for our analysis.
-But now is sounds like we restrict the model even further.}
-We refer this variant as \PPOBRP{}.
+%We refer this variant as \PPOBRP{}.
 The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect partition while not paying too much relative to \OPT.
@@
 Revealed edges, up to any point in the sequence,
-induce some connected subgraphs and we refer to them as \emph{components}.
-An algorithm that always  keeps nodes belonging to the same component collocated is \emph{component respecting}.
+induce some connected subgraphs which grow as components.
+%An algorithm that always  keeps nodes belonging to the same component collocated is \emph{component respecting}.
 
@@
     	}       
-        \STATE {$P_{t} = \mathit{re\mhyphen partition}(P_{t-1}, P_I, \mathcal{C})$} 
+        \STATE {$P_{t} = \mathit{repartition}(P_{t-1}, P_I, \mathcal{C})$} 
          \COMMENT{move to $P$ closest to $P_I$}
@@
 %      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
-      We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a perfect partition closest to $P_I$.
+      We note that the $\mathit{repartition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a perfect partition closest to $P_I$.
 Hence it never moves to a configuration beyond distance $\Delta$.      
@@
 %    including the request that completes revealing of all components that are collocated in $P_F$.
-    There are at most $(k-1)\cdot\ell < k\cdot\ell $ calls   to $\mathit{re\mhyphen partition}$
+    There are at most $(k-1)\cdot\ell < k\cdot\ell $ calls   to $\mathit{repartition}$
      (i.e., the number of internal edges in $P_F$).

### 2020-02-17 paper/paper.tex
Intro to LB

--- 2020-02-17 paper/paper.tex
@@
 
-The adversary constructs components of nodes, and $\OPT$ never splits them.
-The adversary observes the actions of an online algorithm, and if it splits the components at any moment, the adversary issues
+In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic algorithm.
+The minimum cluster capacity for the construction to remain correct is $3$.
+
+The adversary constructs components of nodes, and issues only requests among nodes in the same component.
+If an online algorithm splits the components at any moment, the adversary issues requests to split pairs of nodes.
+We utilize the fact that the online algorithm does not know the components of the adversary, and it may split some components while evicting nodes.
+On the other hand, $\OPT$ never splits the components.
 

### 2020-02-17 paper/paper.tex
hiding authors
LB intro
wip

@@
 
+\begin{comment}
 \author{Maciej Pacut}
@@
 }
+\end{comment}
 
@@
 In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic algorithm.
-The minimum cluster capacity for the construction to remain correct is $3$.
+It is sufficient to have clusters of capacity $k\geq 3$ for this lower bound.
+For $k=2$ a constant competitive algorithm exists.
 
@@
 
+\pagebreak
+
 \begin{theorem}
@@
 
-  Now we estimate the cost of $\OPT(\sigma_E)$ by accounting savings from migrations.
+  Now we estimate the cost of $\OPT(\sigma_E)$ while accounting savings from migrations.
   By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.

### 2020-02-17 paper/paper.tex
Consistency

@@
 	we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
-	\ALG{} must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
+	\ALG{} must collocate $x_1$ and $C$ in some cluster other than $S_1$ and
 	consequently replaces $x_1$ with some other (single) node $x_2$.
@@
 
-\todo{ACM format for review? Is there any specific format like this? Author anonymous, with line numbers etc.}
+\todo{Author anonymous conference!}
+
+\todo{ACM format for review? Is there any specific format like this? With line numbers etc.}
 
@@
 
-\todo{Consistency: collocated at OR collocated in OR collocated on}
+\todo{ACM Reference format starts with a "." Possibly we did not provide all info}
 

### 2020-02-17 paper/podc20ba.tex
stefan iteration

--- 2020-02-17 paper/paper.tex
@@
 
-\title{Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning}
+\title{Brief Announcement: Tight Bounds for Online Balanced Re-Partitioning}
 
@@
 \begin{abstract}
-  Distributed   applications,  including  batch  processing, streaming, and scale-out databases generate a significant amount of network traffic and a~considerable fraction of their runtime is due to network activity.
-In this  paper, we study  algorithms  for collocating frequently communicating nodes in a distributed networked systems in an online fashion.
-
-  In the online graph repartitioning problem, we are given a sequence of pairwise communication requests between n nodes, revealed online.
-  The objective is to service these requests by partitioning the nodes into $\ell$ clusters, each of size $k$, such that frequently communicating nodes are located in the same cluster.
-  The partitioning can be updated by migrating nodes between clusters for a~fixed cost.
-  The goal is to jointly minimize the amount of inter-cluster communication and migration cost.
-
-  So far, the best known lower bound for the competitive ratio of
-	deterministic algorithms was $\Omega(k)$,
- by reduction from online paging.
-  We improve this result significantly by providing the lower bound of $\Omega(k\cdot \ell)$.
-  We follow up with an optimal algorithm for $k=3$.
-  Finally,
-  we provide a $O(k\cdot \ell)$-competitive algorithm for the variant
-  where the optimal strategy is to move to a~fixed configuration
-  that incurs zero communication cost through the whole request sequence.
+Distributed   applications,  including  batch  processing, streaming, scale-out databases,
+or machine learning, generate a significant amount of network traffic.
+By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), the network load can be reduced and application performance improved.
+However, as the communication pattern is a priori unknown and may change over time, it need to be learned efficiently, in an online manner.
+%
+This paper revisits the online 
+balanced re-partitioning problem 
+(introduced by Avin et al.~at DISC 2016)
+which asks for an algorithm which strikes
+an optimal tradeoff between the benefits
+of collocation (i.e., lower network load) 
+and its costs (i.e., migrations). 
+%
+Our main contribution is a significantly improved
+lower bound of $\Omega(k\cdot \ell)$ on the
+competitive ratio, where $\ell$ is the number
+of clusters and $k$ is the cluster size,
+even for a scenario in which the communication
+pattern is static and can be perfectly partitioned;
+we also provide a tight upper bound 
+of $O(k\cdot \ell)$ for this scenario.
+In addition, we present a tight upper bound
+of $\Theta(\ell)$ for $k=3$,
+for the general model in which the
+communication pattern can change arbitrarily
+over time. 
 \end{abstract}
@@
 
-Main technical contribution of this paper is the lower bound $\Omega(k\cdot\ell)$ for the competitive ratio of any deterministic algorithm for online balanced repartition problem (see Section~\ref{sec:lowerbound}).
-The best known lower bound so far was $\Omega(k)$ \cite{repartition-disc}
-which involves only two clusters.
-
-So far, the competitive ratio of the best known algorithm was the component-based algorithm with the ratio $O(k^2\cdot \ell^2)$~\cite{repartition-disc}.
-By applying classic rent-or-buy technique, we obtained an optimal algorithm for $k=3$, that we present in Section~\ref{sec:k3}.
-In Section~\ref{sec:ppl}, we study a restricted model of balanced repartition, called perfect partition.
-For that, we present an~$O(k\cdot \ell)$-competitive algorithm, which is optimal.
-We believe that this result is an important building block on the way to an optimal algorithm for the general model.
-\mahmoud{The mention of distributed alg, very much out of context!} 
-A distributed adaptation of a similar algorithm has been analyzed in \cite{sigmetrics19_partitioning}.
-
-\section{Related Work}
-
-The problem was studied under resource augmentation \cite{repartition-disc}, and under stochastic assumptions for the input\cite{stochastic-ring}.
-For clusters of size $2$ (i.e., $k=2$), a constant competitive algorithm exists~\cite{repartition-disc}.
-
-The model is related to online
-caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}. In our
-setting, requests can be served remotely, which is
-reminiscent of caching models \emph{with
-bypassing}~\cite{EpImLN11,EpImLN15,Irani02}.
-
-The static offline version of~the~problem, called the
-\emph{$\ell$-balanced graph partitioning problem} is 
-NP-complete, and cannot even be approximated within any finite factor unless P
-= NP~\cite{AndRae06}. The static
-variant where $\ell = 2$ corresponds to the minimum bisection problem, which
-is already NP-hard~\cite{GaJoSt76}, and 
-the currently best approximation ratio is $O(\log n)$~\cite{SarVaz95,ArKaKa99,FeKrNi00,FeiKra02,KraFei06,Raec08}.
-
-\section{Preliminaries}
-\label{sec:problem-definition}
-
-The online \emph{balanced repartitioning} problem (\OBRP{}) is defined as
-follows. We are given a set of $n$ nodes,
-initially arbitrarily partitioned into $\ell$~clusters,
+\noindent \textbf{Motivation and model.}
+The \emph{balanced repartitioning} problem (\OBRP{})
+is a fundamental learning problem
+which finds applications in the context
+distributed systems optimization~\cite{repartition-disc}. We are given a set $V$ of $n$ nodes 
+(e.g., virtual machines or processes),
+initially arbitrarily partitioned into $\ell$~clusters
+(e.g., servers or entire racks),
 each of size~$k$.
-The input consists of a sequence of pairwise communication requests
+The nodes interact using
+a sequence of pairwise communication requests
 $\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
-where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a~fixed amount of data.
+where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a~certain amount of data.
 Nodes in $C \subset V$ are \emph{collocated}
 if they reside in the same cluster.
+
 An algorithm serves a communication request between two nodes
@@
 Before serving a request,
-an online algorithm may \emph{repartition},
+an online algorithm may perform a \emph{repartition},
 %(i.e., \emph{reconfigure}).
 i.e.,
-it may (optionally) move some nodes into clusters different than their current clusters while respecting the capacity of every cluster. 
-Only after a repartitioning is completed,
+it may move (``migrate'') some nodes into clusters different from their current clusters, while respecting the capacity of every cluster. 
+Afterwards, 
 the algorithm serves the  request.
@@
 the cost of migrations performed by $\ALG$ while serving the sequence $\sigma$.
-The objective is to minimize the total incurred cost.
 
-In this paper, we use the technique of maintaining \emph{components} of nodes, similar to previous approaches~\cite{repartition-disc}. A~\emph{component} is a subset of frequently communicating nodes.
+We consider two flavors of the problem
+in this paper. In the first one, which
+was also studied by Henzinger et al.~\cite{sigmetrics19_partitioning}
+(SIGMETRICS 2019) and which
+we here mainly consider for the sake of a stronger
+lower bound, we assume that $\sigma$
+simply reveals the edges of a static graph
+that can be perfectly partitioned
+(i.e., in principle, no external requests
+are required). 
+In the second one, we consider a general
+model where $\sigma$ can be arbitrary;
+this was the original problem introduced
+by Avin et al.~\cite{repartition-disc} at DISC 2016.
+For simplicity, we will refer to the former
+model as the \emph{learning} model (as
+one has to learn the static communication graph) 
+and to the latter as the \emph{online} model.
+
+\noindent \textbf{Contributions.}
+Our main technical contribution 
+is an improved lower bound 
+of $\Omega(k\cdot\ell)$ on the competitive ratio of any deterministic online algorithm 
+even in the learning model;
+the best known lower bound so far was $\Omega(k)$,
+even for a more general model~\cite{repartition-disc}.
+We also present an asymptotically optimal, 
+$O(k\cdot \ell)$-competitive algorithm
+for the restricted model.
+For the online model, we present  
+an asymptotically optimal,
+$\Theta(\ell)$ algorithm for $k=3$;
+the best known upper bound 
+so far was $\O(\ell^2)$~\cite{repartition-disc}.
+%
+Our algorithms can be distributed
+similarly to the approach in~\cite{sigmetrics19_partitioning}.
+
+\noindent \textbf{Related work.}
+The existing results in~\cite{repartition-disc}
+and~\cite{sigmetrics19_partitioning}
+primarily focus on a model where 
+the online algorithm can use augmentation,
+i.e., has slightly larger clusters than the offline 
+algorithm. In contrast, in our paper, we focus
+on a scenario where the nodes need to be
+perfectly balanced among the clusters.
+The problem has also been studied in a weaker
+model where the adversary can only sample
+requests from a fixed distribution~\cite{stochastic-ring}.
+For clusters of size $2$ (i.e., $k=2$), 
+it is known that a constant competitive algorithm exists~\cite{repartition-disc}.
+
+%More generally, the model is related to online
+%caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00},
+%see~\cite{repartition-disc} for a discussion.
+%The static offline version of~the~problem, called the
+%\emph{$\ell$-balanced graph partitioning problem} is 
+%NP-complete, and cannot even be approximated within %any finite factor unless P
+%= NP~\cite{AndRae06}.
+
+\noindent \textbf{Preliminaries.}
+In this paper, we use the technique of maintaining \emph{(connected) components} of nodes, similar to previous approaches~\cite{repartition-disc}.
+A~\emph{component} is a subset of frequently communicating nodes.
 Initially,
@@
 
-\section{Lower bound} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
+\section{The Learning Problem} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
 \label{sec:lowerbound}
 
-In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic algorithm.
-It is sufficient to have clusters of capacity $k\geq 3$ for this lower bound.
-For $k=2$ a constant competitive algorithm exists.
+%In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
+%It is sufficient to have clusters of capacity $k\geq 3$ for this lower bound.
+%For $k=2$ a constant competitive algorithm exists.
+%At the end of this section, we also
+%discuss an upper bound.
 
+\noindent \textbf{Lower bound.}
 The adversary constructs components of nodes, and issues only requests among nodes in the same component.
@@
 
-\section{An Optimal Algorithm for Clusters of Size 3}
-\label{sec:k3}
-
-In this section we show that the classic \emph{rent-or-buy} approach\cite{borodin-book, karlin-ski-rental} allows to obtain an optimal algorithm for $k=3$.
-Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one cluster until the end of a phase (that we define precisely later).
-
-%We define the algorithm \TAlg in the following way.
-Now we describe the algorithm \TAlg.
-It partitions nodes into components, and
-initially each node belongs to its own singleton component.
-For each pair of nodes, \TAlg maintains a counter, initiated to $0$. 
-Upon receiving a request to a pair that is not collocated in one cluster, it increases their counter by $1$.
-If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
-If no such partitioning exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
-
-%\maciek{Shouldn't we think about whole component cuts instead of single edge? Hmm. NO!}
-
-In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
-
-\begin{lemma}
-  \label{lem:1req}
-  A single repartition of \TAlg consists of at most $2$ migrations.
-\end{lemma}
-
-\begin{proof}
-  Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
-  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
-
-  %We distinguish between three types of clusters: $C_1, C_2, C_3$,
-  %which we define as follows.
-  %A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
-  %A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
-  %Finally, a cluster of type $C_3$  contains one component of size $3$.
-
-    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
-    Note that $U\neq V$,
-	 as otherwise the request would not trigger a~repartitioning.
-	 We consider cases upon types of clusters $U$ and $V$.
-    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
-    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
-    For two clusters, any configuration can be reached within two migrations, due to the fact that clusters are indistiguishable.
-
-    Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
-    Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
-    Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
-In this case \TAlg performs one migration, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
-\end{proof}
-
-\pagebreak
-
-\begin{theorem}
-  \TAlg is $O(\ell)$-competitive.
-\end{theorem}
-\begin{proof}
-  Fix a completed phase, and consider the state of \TAlg's counters at the end of it.
-  We consider the incomplete phase later in this proof.
-
-  As \TAlg is component respecting, it never increases any counter above $\alpha$.
-  We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
-
-  Let $\sigma$ be the input sequence that arrived during the phase.
-  Let $\sigma_{cost}$ be the requests that at the moment of arrival were external requests for \TAlg (these are the only requests that incurred a cost for \TAlg).
-  In our analysis, we partition $\sigma_{cost}$ into subsequences $\sigma_I$ and $\sigma_E$.
-  The sequence $\sigma_I$ (inter-component requests) are the requests from $\sigma_{cost}$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
-  The sequence $\sigma_E$ (extra-component requests) denotes the requests from $\sigma_{cost}$ that do not appear in $\sigma_I$.
-
-  %Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
-  %Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
-  Let $\TAlg(M)$ be the cost of migrations performed by \TAlg in this phase.
-  \TAlg performs at most $2 \ell$ component merge operations, as
-  exceeding this number means that a component of size $4$ exists, and the phase should have ended already.
-  Combining this with Lemma~\ref{lem:1req} gives us $\TAlg(M) \leq 4\alpha\cdot\ell$.
-  %Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
-
-  Now we bound $\TAlg(\sigma_I)$.
-  Cluster of type $C_3$ contributes at most $3 \alpha - 1$ to $\TAlg(\sigma_I)$, as $2$ of pairs of nodes from the component are saturated and contribute $\alpha$ each, and the third, unsaturated pair contributes at most $\alpha-1$.
-  Other cluster types contribute less: $C_1$ contributes $0$ and $C_2$ contributes $\alpha$.
-  Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq (3 \alpha-1)\cdot \ell \leq 3\alpha\cdot\ell$.
-
-  %We bound $A^E$ by $k^2 \cdot (\alpha - 1)$, as no more than $k^2$ pairs are unsaturated, and each of them contributes at most $\alpha -1$.
-  %\maciek{not needed most likely}
-
-  Moreover, \TAlg paid for all requests from $\sigma_E$, and thus $\TAlg(\sigma_E) = |\sigma_E|$.
-  In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 7\alpha\cdot \ell + |\sigma_E|$ during this phase.
-
-  \medskip
-
-  Now we lower-bound the cost of $\OPT$.
-  By $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$ we denote the cost of $\OPT$ on these input sequences (defined with respect to components of \TAlg in this phase).
-  By $\OPT(M)$ we denote the cost of migrations performed by $\OPT$ in this phase.
-
-  We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
-  While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts: we separately bound $\OPT$ by $\OPT(\sigma_I) + \OPT(M)$ and $\OPT(\sigma_E) + \OPT(M)$.
-  Combining those bounds gives us $\OPT \geq \max\{\OPT(\sigma_I) + \OPT(M), \OPT(\sigma_E) + \OPT(M)\} \geq (\OPT(\sigma_I) + \OPT(M)) / 2 + (\OPT(\sigma_E) + \OPT(M)) / 2$.
-
-  %Let $O^M$ be the cost of migrations performed by $\OPT$ during the phase.
-  %Let $O^I$ be the cost of serving requests between nodes that were put in one component by \TAlg during this phase.
-  %Let $O^E$ be the cost serving requests between nodes that \TAlg did not put in the same component during that phase.
-
-  %First, we estimate the cost related to $\sigma_I$.
-  We have $\OPT(M) + \OPT(\sigma_I) \geq \alpha$, as the phase ended when the components of \TAlg{} could not be partitioned without splitting them.
-  Hence, for every possible configuration of $\OPT$, there exists a non-collocated pair of nodes with at least $\alpha$ requests between them, and
-  $\OPT$ either served them remotely or performed a~migration.
-
-  \medskip
-  Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
-  In a~fixed configuration of $\OPT$, it may mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
-  Recall that $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
-  Faced with $W := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests $\sigma_E$ that it could not mitigate, $\OPT$ served some of them remotely and possibly performed some migrations to decrease its cost.
-
-  Now we estimate the cost of $\OPT(\sigma_E)$ while accounting savings from migrations.
-  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
-  This may allow to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
-  As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
-  By performing a single migration that costs $\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
-  Thus, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4 (\alpha-1)}$.
-  This gives us
-  \begin{equation*}
-    \OPT(\sigma_E) + \OPT(M)  \geq (|\sigma_E|-3(\alpha - 1)\cdot \ell) \cdot \frac{\alpha}{4 (\alpha-1)}
-    \geq |\sigma_E| / 4 - 4 \alpha \cdot \ell.
-  \end{equation*}
-  From that, we obtain $|\sigma_E| \geq 4(\OPT(\sigma_E)+\OPT(M)) + 16\alpha \cdot \ell$.
-  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, we are ready to bound the competitive ratio
-  \begin{equation*}
-    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{7\alpha \cdot \ell + |\sigma_E|}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{46\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 46 \ell = O(\ell).
-  \end{equation*}
-
-  \medskip
-
-  It remains to consider the last, unfinished phase.
-  Consider the case, where the unfinished phase is also the first one.
-  Then, we cannot charge $\OPT$ due to unability to partition the components.
-  Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
-  We charge $\OPT$ $\alpha$ for the first external $\alpha$ requests or a migration,
-  and we follow the analysis regarding the unsaturated requests.
-  If the input finished before first $\alpha$ external requests, then \TAlg is optimal.
-  Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
-  This way, the competitive ratio increases at most twofold.
-\end{proof}
-
-\section{An Optimal Algorithm for Perfect Partition}	\label{sec:ppl}
-%\subsection{Perfect Partition Model}\label{sec:perfectPartition}
-%\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
-%\maciek{ every algorithm must colocate every communicating pair.
-%	\mahmoud{updated!}}
-%\maciek{Note that the algorithm from Section~\ref{sec:k3} is $O(\ell)$-competitive for the Perfect Partition Model, as during the first phase, a perfect partition of input exists.}
-
-%\subsection{An $O(k\cdot \ell)$-competitive Algorithm for Perfect Partition}
-
-%A \emph{perfect partition} is a partitioning of nodes into $\ell$ clusters
-%such that it incurs no inter-cluster communication throughout the input sequence.
-%Equivalently, a perfect partitioning it an $\ell$-way cut of the communication graph of the input sequence.
-In this section,
-we give an algorithm for a restricted variant of  \OBRP{}, named \PPOBRP{},
+\noindent \textbf{Upper bound.}
+Now we give an algorithm for a restricted variant of  \OBRP{}, named \PPOBRP{},
 %\maciek{This is not sufficient! We also disallow serving requests remotely.}
@@
 %An algorithm that always  keeps nodes belonging to the same component collocated is \emph{component respecting}.
-
-\subsection{Perfect Partition Learner (\PPL) Algorithm}
 
@@
 
-% \section{Corollary and Future Directions}
+\section{The Online Problem}
+\label{sec:k3}
 
-% The gap between upper and lower bound is still substantial, as the best known algorithm is the component-respecting algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\cdot\ell^2)$.
+Let us now discuss the general online
+model where the request sequence
+can be arbitrary. We show that the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental} allows to obtain an optimal algorithm for $k=3$.
+Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one cluster until the end of a phase (that we define precisely later).
+
+%We define the algorithm \TAlg in the following way.
+Now we describe the algorithm \TAlg.
+It partitions nodes into components, and
+initially each node belongs to its own singleton component.
+For each pair of nodes, \TAlg maintains a counter, initiated to $0$. 
+Upon receiving a request to a pair that is not collocated in one cluster, it increases their counter by $1$.
+If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
+If no such partitioning exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
+
+%\maciek{Shouldn't we think about whole component cuts instead of single edge? Hmm. NO!}
+
+In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
+
+\begin{lemma}
+  \label{lem:1req}
+  A single repartition of \TAlg consists of at most $2$ migrations.
+\end{lemma}
+
+\begin{proof}
+  Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
+  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
+
+  %We distinguish between three types of clusters: $C_1, C_2, C_3$,
+  %which we define as follows.
+  %A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+  %A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
+  %Finally, a cluster of type $C_3$  contains one component of size $3$.
+
+    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
+    Note that $U\neq V$,
+	 as otherwise the request would not trigger a~repartitioning.
+	 We consider cases upon types of clusters $U$ and $V$.
+    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
+    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
+    For two clusters, any configuration can be reached within two migrations, due to the fact that clusters are indistiguishable.
+
+    Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
+    Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
+    Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
+In this case \TAlg performs one migration, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
+\end{proof}
+
+\begin{theorem}
+  \TAlg is $O(\ell)$-competitive.
+\end{theorem}
+\begin{proof}
+  Fix a completed phase, and consider the state of \TAlg's counters at the end of it.
+  We consider the incomplete phase later in this proof.
+
+  As \TAlg is component respecting, it never increases any counter above $\alpha$.
+  We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
+
+  Let $\sigma$ be the input sequence that arrived during the phase.
+  Let $\sigma_{cost}$ be the requests that at the moment of arrival were external requests for \TAlg (these are the only requests that incurred a cost for \TAlg).
+  In our analysis, we partition $\sigma_{cost}$ into subsequences $\sigma_I$ and $\sigma_E$.
+  The sequence $\sigma_I$ (inter-component requests) are the requests from $\sigma_{cost}$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
+  The sequence $\sigma_E$ (extra-component requests) denotes the requests from $\sigma_{cost}$ that do not appear in $\sigma_I$.
+
+  %Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
+  %Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
+  Let $\TAlg(M)$ be the cost of migrations performed by \TAlg in this phase.
+  \TAlg performs at most $2 \ell$ component merge operations, as
+  exceeding this number means that a component of size $4$ exists, and the phase should have ended already.
+  Combining this with Lemma~\ref{lem:1req} gives us $\TAlg(M) \leq 4\alpha\cdot\ell$.
+  %Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
+
+  Now we bound $\TAlg(\sigma_I)$.
+  Cluster of type $C_3$ contributes at most $3 \alpha - 1$ to $\TAlg(\sigma_I)$, as $2$ of pairs of nodes from the component are saturated and contribute $\alpha$ each, and the third, unsaturated pair contributes at most $\alpha-1$.
+  Other cluster types contribute less: $C_1$ contributes $0$ and $C_2$ contributes $\alpha$.
+  Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq (3 \alpha-1)\cdot \ell \leq 3\alpha\cdot\ell$.
+
+  %We bound $A^E$ by $k^2 \cdot (\alpha - 1)$, as no more than $k^2$ pairs are unsaturated, and each of them contributes at most $\alpha -1$.
+  %\maciek{not needed most likely}
+
+  Moreover, \TAlg paid for all requests from $\sigma_E$, and thus $\TAlg(\sigma_E) = |\sigma_E|$.
+  In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 7\alpha\cdot \ell + |\sigma_E|$ during this phase.
+
+  \medskip
+
+  Now we lower-bound the cost of $\OPT$.
+  By $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$ we denote the cost of $\OPT$ on these input sequences (defined with respect to components of \TAlg in this phase).
+  By $\OPT(M)$ we denote the cost of migrations performed by $\OPT$ in this phase.
+
+  We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
+  While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts: we separately bound $\OPT$ by $\OPT(\sigma_I) + \OPT(M)$ and $\OPT(\sigma_E) + \OPT(M)$.
+  Combining those bounds gives us $\OPT \geq \max\{\OPT(\sigma_I) + \OPT(M), \OPT(\sigma_E) + \OPT(M)\} \geq (\OPT(\sigma_I) + \OPT(M)) / 2 + (\OPT(\sigma_E) + \OPT(M)) / 2$.
+
+  %Let $O^M$ be the cost of migrations performed by $\OPT$ during the phase.
+  %Let $O^I$ be the cost of serving requests between nodes that were put in one component by \TAlg during this phase.
+  %Let $O^E$ be the cost serving requests between nodes that \TAlg did not put in the same component during that phase.
+
+  %First, we estimate the cost related to $\sigma_I$.
+  We have $\OPT(M) + \OPT(\sigma_I) \geq \alpha$, as the phase ended when the components of \TAlg{} could not be partitioned without splitting them.
+  Hence, for every possible configuration of $\OPT$, there exists a non-collocated pair of nodes with at least $\alpha$ requests between them, and
+  $\OPT$ either served them remotely or performed a~migration.
+
+  \medskip
+  Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
+  In a~fixed configuration of $\OPT$, it may mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
+  Recall that $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
+  Faced with $W := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests $\sigma_E$ that it could not mitigate, $\OPT$ served some of them remotely and possibly performed some migrations to decrease its cost.
+
+  Now we estimate the cost of $\OPT(\sigma_E)$ while accounting savings from migrations.
+  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
+  This may allow to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
+  As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
+  By performing a single migration that costs $\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
+  Thus, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4 (\alpha-1)}$.
+  This gives us
+  \begin{equation*}
+    \OPT(\sigma_E) + \OPT(M)  \geq (|\sigma_E|-3(\alpha - 1)\cdot \ell) \cdot \frac{\alpha}{4 (\alpha-1)}
+    \geq |\sigma_E| / 4 - 4 \alpha \cdot \ell.
+  \end{equation*}
+  From that, we obtain $|\sigma_E| \geq 4(\OPT(\sigma_E)+\OPT(M)) + 16\alpha \cdot \ell$.
+  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, we are ready to bound the competitive ratio
+  \begin{equation*}
+    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{7\alpha \cdot \ell + |\sigma_E|}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{46\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 46 \ell = O(\ell).
+  \end{equation*}
+
+  \medskip
+
+  It remains to consider the last, unfinished phase.
+  Consider the case, where the unfinished phase is also the first one.
+  Then, we cannot charge $\OPT$ due to unability to partition the components.
+  Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
+  We charge $\OPT$ $\alpha$ for the first external $\alpha$ requests or a migration,
+  and we follow the analysis regarding the unsaturated requests.
+  If the input finished before first $\alpha$ external requests, then \TAlg is optimal.
+  Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
+  This way, the competitive ratio increases at most twofold.
+\end{proof}
 

### 2020-02-17 paper/podc20ba.tex
some of Stefans feedbacks

@@
  
-\documentclass[manuscript,screen=true]{acmart}
+\documentclass[manuscript,screen=true, review, anonymous]{acmart}
 
@@
 
-\begin{comment}
 \author{Maciej Pacut}
@@
 }
-\end{comment}
 
@@
 %
-Our main contribution is a significantly improved
+Our first contribution is a significantly improved
 lower bound of $\Omega(k\cdot \ell)$ on the
@@
 is an improved lower bound 
-of $\Omega(k\cdot\ell)$ on the competitive ratio of any deterministic online algorithm 
+of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
 even in the learning model;
@@
 	nodes on the cluster $S_1$.
-	Recall that internal requests does not incur any cost for \ALG{}.
+	Recall that internal requests do not incur any cost for \ALG{}.
 	If \ALG{} at any point splits a component
@@
 	Since this is an external request,
-	\ALG{} joins them into one component and collocates them in some cluster other than $S_1$.
+	\ALG{} joins the endpoints into one component and collocates them in some cluster other than $S_1$.
 	For this,
@@
 
- In Section~\ref{sec:lowerbound} we constructed a $\Omega(k \cdot \ell)$ for \OBRP{}.
- Note that the lower bound holds also in the perfect partition model, as the constructed input sequence allows \OPT to move to a perfect partition.
- The corollary is that PPL is optimal.
+% In Section~\ref{sec:lowerbound} we constructed a $\Omega(k \cdot \ell)$ for \OBRP{}.
+% Note that the lower bound holds also in the perfect partition model, as the constructed input sequence allows \OPT to move to a perfect partition.
+% The corollary is that PPL is optimal.
 
@@
 	 as otherwise the request would not trigger a~repartitioning.
-	 We consider cases upon types of clusters $U$ and $V$.
+	 We consider cases based on the types of clusters $U$ and $V$.
     Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component respecting partitioning exists, a contradiction.

### 2020-02-17 paper/podc20ba.tex
ACM comments, done
Fixing acm-format-ref block
\O not displayed apparently

@@
 \keywords{online algorithms, competitive analysis, graph partitioning, clustering}
+\acmISBN{}\acmPrice{}
+\acmDOI{}
 
@@
 the best known upper bound 
-so far was $\O(\ell^2)$~\cite{repartition-disc}.
+so far was $O(\ell^2)$~\cite{repartition-disc}.
 %
@@
 
-\todo{Author anonymous conference!}
+\todo{Author anonymous conference! Done}
 
-\todo{ACM format for review? Is there any specific format like this? With line numbers etc.}
+\todo{ACM format for review? Is there any specific format like this? With line numbers etc. Done}
 
@@
 
-\todo{ACM Reference format starts with a "." Possibly we did not provide all info}
+\todo{ACM Reference format starts with a "." Possibly we did not provide all info. Done}
 

### 2020-02-18 paper/podc20ba.tex
moving PPL alg to the Appendix
feedback

@@
 \noindent \textbf{Contributions.}
-Our main technical contribution 
-is an improved lower bound 
+We provide an improved lower bound 
 of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
@@
 A~\emph{component} is a subset of frequently communicating nodes.
-Initially,
-every node is trivially a \emph{singleton} component.
+We use this concept in both algorithms in this paper, but also in the lower bound, to incur high cost for any online algorithm that splits them.
 An~algorithm is \emph{component respecting}
@@
 We utilize the fact that the online algorithm does not know the components of the adversary, and it may split some components while evicting nodes.
-On the other hand, $\OPT$ never splits the components.
+On the other hand, $\OPT$ initially collocates all components and never splits them again.
 

### 2020-02-18 paper/podc20ba.tex
feedback

@@
 \noindent \textbf{Lower bound.}
-The adversary constructs components of nodes, and issues only requests among nodes in the same component.
+The adversary constructs components of nodes, whose contents are based on actions of the online algorithm.
+All requests arrive to pairs of nodes within the same component.
 If an online algorithm splits the components at any moment, the adversary issues requests to split pairs of nodes.
@@
 \begin{theorem}
-	The competitive ratio of any deterministic algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$ for any $k\geq 3$.
+	The competitive ratio of any deterministic online algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$ for any $k\geq 3$.
 \end{theorem}

### 2020-02-18 paper/podc20ba.tex
3.1 to appendix

@@
 
-\begin{proof}
-  Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
-  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
-
-  %We distinguish between three types of clusters: $C_1, C_2, C_3$,
-  %which we define as follows.
-  %A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
-  %A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
-  %Finally, a cluster of type $C_3$  contains one component of size $3$.
-
-    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
-    Note that $U\neq V$,
-	 as otherwise the request would not trigger a~repartitioning.
-	 We consider cases based on the types of clusters $U$ and $V$.
-    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
-    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
-    For two clusters, any configuration can be reached within two migrations, due to the fact that clusters are indistiguishable.
-
-    Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
-    Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
-    Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
-In this case \TAlg performs one migration, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
-\end{proof}
-
 \begin{theorem}
@@
 
-\section{Todos}
+\begin{proof}[Proof of Lemma~\ref{lem:1req}]
+  Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
+  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
 
-\todo{Author anonymous conference! Done}
+  %We distinguish between three types of clusters: $C_1, C_2, C_3$,
+  %which we define as follows.
+  %A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+  %A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
+  %Finally, a cluster of type $C_3$  contains one component of size $3$.
 
-\todo{ACM format for review? Is there any specific format like this? With line numbers etc. Done}
+    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
+    Note that $U\neq V$,
+	 as otherwise the request would not trigger a~repartitioning.
+	 We consider cases based on the types of clusters $U$ and $V$.
+    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
+    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
+    For two clusters, any configuration can be reached within two migrations, due to the fact that clusters are indistiguishable.
 
-\todo{CCS concepts are mandatory}
-
-\todo{ACM Reference format starts with a "." Possibly we did not provide all info. Done}
+    Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
+    Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
+    Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
+In this case \TAlg performs one migration, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
+\end{proof}
 

### 2020-02-18 paper/podc20ba.tex
Shortening

@@
 In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
+Before bounding the competitive ratio of \TAlg, we introduce the lemma that estimates the cost of a single repartition of \TAlg. We defer its proof to Appendix~\ref{apx:omitted}.
 
@@
   By performing a single migration that costs $\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
-  Thus, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4 (\alpha-1)}$.
-  This gives us
-  \begin{equation*}
-    \OPT(\sigma_E) + \OPT(M)  \geq (|\sigma_E|-3(\alpha - 1)\cdot \ell) \cdot \frac{\alpha}{4 (\alpha-1)}
-    \geq |\sigma_E| / 4 - 4 \alpha \cdot \ell.
-  \end{equation*}
+  Thus, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 4 - 4 \alpha \cdot \ell$.
   From that, we obtain $|\sigma_E| \geq 4(\OPT(\sigma_E)+\OPT(M)) + 16\alpha \cdot \ell$.
@@
   \begin{equation*}
-    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{7\alpha \cdot \ell + |\sigma_E|}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{46\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 46 \ell = O(\ell).
+    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{7\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + E/2} \leq \frac{46\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 46 \ell = O(\ell).
   \end{equation*}
@@
 \section{Ommited proofs}
+\label{apx:omitted}
 

### 2020-02-18 paper/podc20ba.tex
Shortening
grammar

@@
 By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), the network load can be reduced and application performance improved.
-However, as the communication pattern is a priori unknown and may change over time, it need to be learned efficiently, in an online manner.
+However, as the communication pattern is a priori unknown and may change over time, it needs to be learned efficiently, in an online manner.
 %
@@
 (introduced by Avin et al.~at DISC 2016)
-which asks for an algorithm which strikes
+which asks for an algorithm that strikes
 an optimal tradeoff between the benefits
@@
 is a fundamental learning problem
-which finds applications in the context
+that finds applications in the context of
 distributed systems optimization~\cite{repartition-disc}. We are given a set $V$ of $n$ nodes 
@@
 	$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
-	Let $I(C)$ denote the cluster where nodes of a component $C$ are located initially.
+	Let $I(C)$ denote the cluster where nodes of component $C$ are located initially.
 	%	\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
@@
 	Since this is an external request,
-	\ALG{} joins the endpoints into one component and collocates them in some cluster other than $S_1$.
+	\ALG{} joins the endpoints into one component and collocates them in a~cluster other than $S_1$.
 	For this,
@@
 	we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
-	\ALG{} must collocate $x_1$ and $C$ in some cluster other than $S_1$ and
-	consequently replaces $x_1$ with some other (single) node $x_2$.
+	\ALG{} must collocate $x_1$ and $C$ in a~cluster other than $S_1$ and
+	consequently replace $x_1$ with some other (single) node $x_2$.
 	We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
@@
 	At this point there are at most $\ell+1$ single nodes left,
-	otherwise there would be more pairs of single nodes that were initially in the same cluster.
+	otherwise, there would be more pairs of single nodes that were initially in the same cluster.
 
@@
 model where the request sequence
-can be arbitrary. We show that the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental} allows to obtain an optimal algorithm for $k=3$.
+can be arbitrary. We show that the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental} allows obtaining an optimal algorithm for $k=3$.
 Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one cluster until the end of a phase (that we define precisely later).
@@
 It partitions nodes into components, and
-initially each node belongs to its own singleton component.
+initially, each node belongs to its own singleton component.
 For each pair of nodes, \TAlg maintains a counter, initiated to $0$. 
@@
   We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
-
   Let $\sigma$ be the input sequence that arrived during the phase.
@@
     Note that $U\neq V$,
-	 as otherwise the request would not trigger a~repartitioning.
+	 as otherwise, the request would not trigger a~repartitioning.
 	 We consider cases based on the types of clusters $U$ and $V$.
-    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
+    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
     If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.

### 2020-02-18 paper/podc20ba.tex
grammar

@@
   As \TAlg is component respecting, it never increases any counter above $\alpha$.
-  We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
+  We say that the pair $(u, v)$ is \emph{saturated} if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
   Let $\sigma$ be the input sequence that arrived during the phase.
@@
   Now we bound $\TAlg(\sigma_I)$.
-  Cluster of type $C_3$ contributes at most $3 \alpha - 1$ to $\TAlg(\sigma_I)$, as $2$ of pairs of nodes from the component are saturated and contribute $\alpha$ each, and the third, unsaturated pair contributes at most $\alpha-1$.
+  A~cluster of type $C_3$ contributes at most $3 \alpha - 1$ to $\TAlg(\sigma_I)$, as $2$ of pairs of nodes from the component are saturated and contribute $\alpha$ each, and the third, unsaturated pair contributes at most $\alpha-1$.
   Other cluster types contribute less: $C_1$ contributes $0$ and $C_2$ contributes $\alpha$.
@@
   By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
-  This may allow to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
+  This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
   As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
@@
   Consider the case, where the unfinished phase is also the first one.
-  Then, we cannot charge $\OPT$ due to unability to partition the components.
+  Then, we cannot charge $\OPT$ due to the unability to partition the components.
   Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
@@
   and we follow the analysis regarding the unsaturated requests.
-  If the input finished before first $\alpha$ external requests, then \TAlg is optimal.
+  If the input finished before the first $\alpha$ external requests, then \TAlg is optimal.
   Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.

### 2020-02-18 paper/podc20ba.tex
Change the title: as in DISC

@@
 
-\title{Brief Announcement: Tight Bounds for Online Balanced Re-Partitioning}
+\title{Brief Announcement: Tight Bounds for Online Balanced Repartitioning}
 
@@
 This paper revisits the online 
-balanced re-partitioning problem 
+balanced repartitioning problem 
 (introduced by Avin et al.~at DISC 2016)

### 2020-02-18 paper/podc20ba.tex
LB

@@
 We utilize the fact that the online algorithm does not know the components of the adversary, and it may split some components while evicting nodes.
-On the other hand, $\OPT$ initially collocates all components and never splits them again.
+On the other hand, $\OPT$ initially collocates all nodes of each component and never splits them again.
 
@@
 \begin{proof}
-	We construct an instance of the problem with $\ell$ clusters 
+	Let $\alpha=1$. We construct an instance of the problem with $\ell$ clusters 
 	$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
@@
 	Fix any online algorithm \ALG{}.
-	We construct the input sequence for \ALG{} as follows.
-	First,
-	we issue $k-2$ (internal) requests so that \ALG{} can form a component of $k-1$
-	nodes on the cluster $S_1$.
-	Recall that internal requests do not incur any cost for \ALG{}.
+  We join some nodes into components, and produce the input sequence by observing actions of \ALG{} with respect to nodes of these components.
 	If \ALG{} at any point splits a component
@@
 	then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
-	Hence, if \ALG{} never collocates them then it is not competitive.
+	If \ALG{} never collocates them, then it is not competitive.
+
+  Initially each node belongs to its own component of size $1$.
+  We say that a node is single if it belongs to a singleton component.
+  We begin by forming a component of single $k-1$ nodes on the cluster $S_1$.
 
@@
 	Next,
-	we issue a request between $x_0$ and $y_0$.
-	Since this is an external request,
-	\ALG{} joins the endpoints into one component and collocates them in a~cluster other than $S_1$.
-	For this,
+	we merge components of $x_0$~and~$y_0$.
+  Since these reside on different clusters of \ALG{}, it produces an external request.
+  \ALG{} must collocate them in a~cluster other than $S_1$.
+	To this end,
 	\ALG{} moves to a new configuration
@@
 	Next,
-	we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
+	we merge the component of $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
 	\ALG{} must collocate $x_1$ and $C$ in a~cluster other than $S_1$ and
 	consequently replace $x_1$ with some other (single) node $x_2$.
-	We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
+	We repeat merging the component of single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
   until there are only two single nodes left that  originate from the same cluster.

### 2020-02-18 paper/podc20ba.tex
Rabbit out of the hat

@@
 \begin{proof}
-	Let $\alpha=1$. We construct an instance of the problem with $\ell$ clusters 
+	We construct an instance of the problem with $\ell$ clusters 
 	$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
@@
 	Hence,
-	OPT pays for $4$ node migrations and
-	\ALG{} incurs at least one migration for each node in the sequence $X := x_0, x_1,\dots$.
+	OPT pays for $2$ node swaps and
+	\ALG{} incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
 	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
-	therefore $|X| \geq k \cdot \ell - k - \ell \in \Omega(k\cdot\ell)$.
+  therefore $|X| \geq k \cdot \ell - k - \ell$.
+  Finally, $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
 \end{proof}

### 2020-02-18 paper/podc20ba.tex
\alpha and 2\alpha confusion fixed

@@
   \label{lem:1req}
-  A single repartition of \TAlg consists of at most $2$ migrations.
+  During a single repartition, \TAlg exchanges at most $2$ pairs of nodes.
 \end{lemma}
@@
   exceeding this number means that a component of size $4$ exists, and the phase should have ended already.
-  Combining this with Lemma~\ref{lem:1req} gives us $\TAlg(M) \leq 4\alpha\cdot\ell$.
+  Combining this with Lemma~\ref{lem:1req} gives us $\TAlg(M) \leq 8\alpha\cdot\ell$ (recall that an exchange costs $2\alpha$).
   %Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
@@
   Moreover, \TAlg paid for all requests from $\sigma_E$, and thus $\TAlg(\sigma_E) = |\sigma_E|$.
-  In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 7\alpha\cdot \ell + |\sigma_E|$ during this phase.
+  In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 11\alpha\cdot \ell + |\sigma_E|$ during this phase.
 
@@
   As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
-  By performing a single migration that costs $\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
-  Thus, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 4 - 4 \alpha \cdot \ell$.
-  From that, we obtain $|\sigma_E| \geq 4(\OPT(\sigma_E)+\OPT(M)) + 16\alpha \cdot \ell$.
+  By performing a single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
+  Thus, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{2\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 2 - 2 \alpha \cdot \ell$.
+  From that, we obtain $|\sigma_E| \geq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
   Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, we are ready to bound the competitive ratio
   \begin{equation*}
-    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{7\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + E/2} \leq \frac{46\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 46 \ell = O(\ell).
+    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + E/2} \leq \frac{27\alpha\cdot\ell + 2\cdot E}{\alpha + E} \leq 27 \ell = O(\ell).
   \end{equation*}

### 2020-02-18 paper/podc20ba.tex
Words: migration vs swap

@@
 This invariant ensures that \PPL does not pay too much while recovering $P_F$.
-
-\begin{algorithm}
-    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
-    \renewcommand{\algorithmicensure}{\textbf{Output:}}
-    \begin{algorithmic}
-%        \Require 
-%        $k, \ell$,
-%        initial configuration $P_I$,
-%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
-%        \Ensure A final configuration $P_F$ 
-        \STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$}
-        \STATE{$P_0 := P_I$}
-         \label{line:initcomponents}
-        \FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
-        \STATE Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
-        \IF{$C_1 \neq C_2$}
-        \STATE {Unite the two components into a single component $C'$ and
-        $\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$} \label{line:mergecomponents}
-        \IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
-	     \COMMENT{i.e.~if not in the same cluster}    
-    	}       
-        \STATE {$P_{t} = \mathit{repartition}(P_{t-1}, P_I, \mathcal{C})$} 
-         \COMMENT{move to $P$ closest to $P_I$}
-        \label{line:rebalance} 
-        \ENDIF
-        \ENDIF
-        \ENDFOR
-    \end{algorithmic}
-    \caption{Perfect Partition Learner (\PPL)}
-    \label{alg:ppl}
-      \end{algorithm}
-
 %      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
@@
 Hence it never moves to a configuration beyond distance $\Delta$.
+The scheme of the algorithm can be found in the Appendix \ref{apx:PPL}.  
 \begin{property} \label{prop:dist<OPT}
@@
 
-\begin{appendix}
+\appendix
+\section{PPL Algorithm} \label{apx:PPL}
+\begin{algorithm}
+	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+	\renewcommand{\algorithmicensure}{\textbf{Output:}}
+	\begin{algorithmic}
+		%        \Require 
+		%        $k, \ell$,
+		%        initial configuration $P_I$,
+		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
+		%        \Ensure A final configuration $P_F$ 
+		\STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$}
+		\STATE{$P_0 := P_I$}
+		\label{line:initcomponents}
+		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
+		\STATE Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
+		\IF{$C_1 \neq C_2$}
+		\STATE {Unite the two components into a single component $C'$ and
+			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$} \label{line:mergecomponents}
+		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
+			\COMMENT{i.e.~if not in the same cluster}    
+		}       
+		\STATE {$P_{t} = \mathit{repartition}(P_{t-1}, P_I, \mathcal{C})$} 
+		\COMMENT{move to $P$ closest to $P_I$}
+		\label{line:rebalance} 
+		\ENDIF
+		\ENDIF
+		\ENDFOR
+	\end{algorithmic}
+	\caption{Perfect Partition Learner (\PPL)}
+	\label{alg:ppl}
+\end{algorithm}
 
@@
     If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
-    For two clusters, any configuration can be reached within two migrations, due to the fact that clusters are indistiguishable.
+    For two clusters, any configuration can be reached within two swaps, due to the fact that clusters are indistinguishable.
 
@@
     Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
-In this case \TAlg performs one migration, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
+In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
 \end{proof}
 
-\end{appendix}
-
 \end{document}

### 2020-02-18 paper/podc20ba.tex
Line endings

@@
 %      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
-      We note that the $\mathit{repartition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a perfect partition closest to $P_I$.
+      We note that the $\mathit{repartition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a~perfect partition closest to $P_I$.
 Hence it never moves to a configuration beyond distance $\Delta$.

### 2020-02-18 paper/podc20ba.tex
last edit

@@
 \noindent \textbf{Lower bound.}
-The adversary constructs components of nodes, whose contents are based on actions of the online algorithm.
-All requests arrive to pairs of nodes within the same component.
-If an online algorithm splits the components at any moment, the adversary issues requests to split pairs of nodes.
+The adversary constructs components of nodes based on the partitioning choices of a given online algorithm.
+Adversary issues requests between pairs of nodes that belong to the same component.
+If an online algorithm splits a component at any moment,
+the adversary issues requests to the split pair until the algorithm collocates them.
 We utilize the fact that the online algorithm does not know the components of the adversary, and it may split some components while evicting nodes.

### 2020-02-18 paper/podc20ba.tex
Typos
CCS concepts

@@
 \acmDOI{}
+
+\ccsdesc[300]{Networks~Data center networks}
+\ccsdesc[300]{Theory of computation~Unsupervised learning and clustering}
 
@@
   Consider the case, where the unfinished phase is also the first one.
-  Then, we cannot charge $\OPT$ due to the unability to partition the components.
+  Then, we cannot charge $\OPT$ due to the inability to partition the components.
   Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
@@
 
-\section{Ommited proofs}
+\section{Omitted proofs}
 \label{apx:omitted}
@@
 
+\begin{CCSXML}
+	<ccs2012>
+	<concept>
+	<concept_id>10003033.10003106.10003110</concept_id>
+	<concept_desc>Networks~Data center networks</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10011007.10010940.10010971.10011120.10003100</concept_id>
+	<concept_desc>Software and its engineering~Cloud computing</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10003752.10010070.10010071.10010074</concept_id>
+	<concept_desc>Theory of computation~Unsupervised learning and clustering</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[300]{Networks~Data center networks}
+\ccsdesc[300]{Software and its engineering~Cloud computing}
+\ccsdesc[300]{Theory of computation~Unsupervised learning and clustering}
 \end{document}

### 2020-02-18 paper/podc20ba.tex
CCS concepts (from Stefan)

--- 2020-02-18 paper/podc20ba.tex
@@
 \acmDOI{}
+
+\ccsdesc[500]{Networks~Network algorithms}
+\ccsdesc[300]{Computer systems organization~Cloud computing}
+\ccsdesc[300]{Computer systems organization~Distributed architectures}
 
@@
 
+\pagebreak
 \appendix
@@
 
+\begin{CCSXML}
+	<ccs2012>
+	<concept>
+	<concept_id>10003033.10003068</concept_id>
+	<concept_desc>Networks~Network algorithms</concept_desc>
+	<concept_significance>500</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10010520.10010521.10010537.10003100</concept_id>
+	<concept_desc>Computer systems organization~Cloud computing</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10010520.10010521.10010537</concept_id>
+	<concept_desc>Computer systems organization~Distributed architectures</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	</ccs2012>
+\end{CCSXML}
+
 \end{document}

### 2020-03-03 paper/paper.tex
Fork of the PODC submission

--- 2020-02-17 paper/paper.tex
@@
 
-\title{Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning}
+\title{Brief Announcement: Tight Bounds for Online Balanced Repartitioning}
 
-\begin{comment}
 \author{Maciej Pacut}
@@
 \keywords{online algorithms, competitive analysis, graph partitioning, clustering}
+\acmISBN{}\acmPrice{}
+\acmDOI{}
+
+\ccsdesc[500]{Networks~Network algorithms}
+\ccsdesc[300]{Computer systems organization~Cloud computing}
+\ccsdesc[300]{Computer systems organization~Distributed architectures}
 
@@
 \begin{abstract}
-  Distributed   applications,  including  batch  processing, streaming, and scale-out databases generate a significant amount of network traffic and a~considerable fraction of their runtime is due to network activity.
-In this  paper, we study  algorithms  for collocating frequently communicating nodes in a distributed networked systems in an online fashion.
-
-  In the online graph repartitioning problem, we are given a sequence of pairwise communication requests between n nodes, revealed online.
-  The objective is to service these requests by partitioning the nodes into $\ell$ clusters, each of size $k$, such that frequently communicating nodes are located in the same cluster.
-  The partitioning can be updated by migrating nodes between clusters for a~fixed cost.
-  The goal is to jointly minimize the amount of inter-cluster communication and migration cost.
-
-  So far, the best known lower bound for the competitive ratio of
-	deterministic algorithms was $\Omega(k)$,
- by reduction from online paging.
-  We improve this result significantly by providing the lower bound of $\Omega(k\cdot \ell)$.
-  We follow up with an optimal algorithm for $k=3$.
-  Finally,
-  we provide a $O(k\cdot \ell)$-competitive algorithm for the variant
-  where the optimal strategy is to move to a~fixed configuration
-  that incurs zero communication cost through the whole request sequence.
+Distributed   applications,  including  batch  processing, streaming, scale-out databases,
+or machine learning, generate a significant amount of network traffic.
+By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), the network load can be reduced and application performance improved.
+However, as the communication pattern is a priori unknown and may change over time, it needs to be learned efficiently, in an online manner.
+%
+This paper revisits the online 
+balanced repartitioning problem 
+(introduced by Avin et al.~at DISC 2016)
+which asks for an algorithm that strikes
+an optimal tradeoff between the benefits
+of collocation (i.e., lower network load) 
+and its costs (i.e., migrations). 
+%
+Our first contribution is a significantly improved
+lower bound of $\Omega(k\cdot \ell)$ on the
+competitive ratio, where $\ell$ is the number
+of clusters and $k$ is the cluster size,
+even for a scenario in which the communication
+pattern is static and can be perfectly partitioned;
+we also provide a tight upper bound 
+of $O(k\cdot \ell)$ for this scenario.
+In addition, we present a tight upper bound
+of $\Theta(\ell)$ for $k=3$,
+for the general model in which the
+communication pattern can change arbitrarily
+over time. 
 \end{abstract}
@@
 
-Main technical contribution of this paper is the lower bound $\Omega(k\cdot\ell)$ for the competitive ratio of any deterministic algorithm for online balanced repartition problem (see Section~\ref{sec:lowerbound}).
-The best known lower bound so far was $\Omega(k)$ \cite{repartition-disc}
-which involves only two clusters.
-
-So far, the competitive ratio of the best known algorithm was the component-based algorithm with the ratio $O(k^2\cdot \ell^2)$~\cite{repartition-disc}.
-By applying classic rent-or-buy technique, we obtained an optimal algorithm for $k=3$, that we present in Section~\ref{sec:k3}.
-In Section~\ref{sec:ppl}, we study a restricted model of balanced repartition, called perfect partition.
-For that, we present an~$O(k\cdot \ell)$-competitive algorithm, which is optimal.
-We believe that this result is an important building block on the way to an optimal algorithm for the general model.
-\mahmoud{The mention of distributed alg, very much out of context!} 
-A distributed adaptation of a similar algorithm has been analyzed in \cite{sigmetrics19_partitioning}.
-
-\section{Related Work}
-
-The problem was studied under resource augmentation \cite{repartition-disc}, and under stochastic assumptions for the input\cite{stochastic-ring}.
-For clusters of size $2$ (i.e., $k=2$), a constant competitive algorithm exists~\cite{repartition-disc}.
-
-The model is related to online
-caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}. In our
-setting, requests can be served remotely, which is
-reminiscent of caching models \emph{with
-bypassing}~\cite{EpImLN11,EpImLN15,Irani02}.
-
-The static offline version of~the~problem, called the
-\emph{$\ell$-balanced graph partitioning problem} is 
-NP-complete, and cannot even be approximated within any finite factor unless P
-= NP~\cite{AndRae06}. The static
-variant where $\ell = 2$ corresponds to the minimum bisection problem, which
-is already NP-hard~\cite{GaJoSt76}, and 
-the currently best approximation ratio is $O(\log n)$~\cite{SarVaz95,ArKaKa99,FeKrNi00,FeiKra02,KraFei06,Raec08}.
-
-\section{Preliminaries}
-\label{sec:problem-definition}
-
-The online \emph{balanced repartitioning} problem (\OBRP{}) is defined as
-follows. We are given a set of $n$ nodes,
-initially arbitrarily partitioned into $\ell$~clusters,
+\noindent \textbf{Motivation and model.}
+The \emph{balanced repartitioning} problem (\OBRP{})
+is a fundamental learning problem
+that finds applications in the context of
+distributed systems optimization~\cite{repartition-disc}. We are given a set $V$ of $n$ nodes 
+(e.g., virtual machines or processes),
+initially arbitrarily partitioned into $\ell$~clusters
+(e.g., servers or entire racks),
 each of size~$k$.
-The input consists of a sequence of pairwise communication requests
+The nodes interact using
+a sequence of pairwise communication requests
 $\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
-where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a~fixed amount of data.
+where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a~certain amount of data.
 Nodes in $C \subset V$ are \emph{collocated}
 if they reside in the same cluster.
+
 An algorithm serves a communication request between two nodes
@@
 the cost of migrations performed by $\ALG$ while serving the sequence $\sigma$.
-The objective is to minimize the total incurred cost.
 
-In this paper, we use the technique of maintaining \emph{components} of nodes, similar to previous approaches~\cite{repartition-disc}. A~\emph{component} is a subset of frequently communicating nodes.
-Initially,
-every node is trivially a \emph{singleton} component.
+We consider two flavors of the problem
+in this paper. In the first one, which
+was also studied by Henzinger et al.~\cite{sigmetrics19_partitioning}
+(SIGMETRICS 2019) and which
+we here mainly consider for the sake of a stronger
+lower bound, we assume that $\sigma$
+simply reveals the edges of a static graph
+that can be perfectly partitioned
+(i.e., in principle, no external requests
+are required). 
+In the second one, we consider a general
+model where $\sigma$ can be arbitrary;
+this was the original problem introduced
+by Avin et al.~\cite{repartition-disc} at DISC 2016.
+For simplicity, we will refer to the former
+model as the \emph{learning} model (as
+one has to learn the static communication graph) 
+and to the latter as the \emph{online} model.
+
+\noindent \textbf{Contributions.}
+We provide an improved lower bound 
+of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
+even in the learning model;
+the best known lower bound so far was $\Omega(k)$,
+even for a more general model~\cite{repartition-disc}.
+We also present an asymptotically optimal, 
+$O(k\cdot \ell)$-competitive algorithm
+for the restricted model.
+For the online model, we present  
+an asymptotically optimal,
+$\Theta(\ell)$ algorithm for $k=3$;
+the best known upper bound 
+so far was $O(\ell^2)$~\cite{repartition-disc}.
+%
+Our algorithms can be distributed
+similarly to the approach in~\cite{sigmetrics19_partitioning}.
+
+\noindent \textbf{Related work.}
+The existing results in~\cite{repartition-disc}
+and~\cite{sigmetrics19_partitioning}
+primarily focus on a model where 
+the online algorithm can use augmentation,
+i.e., has slightly larger clusters than the offline 
+algorithm. In contrast, in our paper, we focus
+on a scenario where the nodes need to be
+perfectly balanced among the clusters.
+The problem has also been studied in a weaker
+model where the adversary can only sample
+requests from a fixed distribution~\cite{stochastic-ring}.
+For clusters of size $2$ (i.e., $k=2$), 
+it is known that a constant competitive algorithm exists~\cite{repartition-disc}.
+
+%More generally, the model is related to online
+%caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00},
+%see~\cite{repartition-disc} for a discussion.
+%The static offline version of~the~problem, called the
+%\emph{$\ell$-balanced graph partitioning problem} is 
+%NP-complete, and cannot even be approximated within %any finite factor unless P
+%= NP~\cite{AndRae06}.
+
+\noindent \textbf{Preliminaries.}
+In this paper, we use the technique of maintaining \emph{(connected) components} of nodes, similar to previous approaches~\cite{repartition-disc}.
+A~\emph{component} is a subset of frequently communicating nodes.
+We use this concept in both algorithms in this paper, but also in the lower bound, to incur high cost for any online algorithm that splits them.
 An~algorithm is \emph{component respecting}
@@
 
-\section{Lower bound} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
+\section{The Learning Problem} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
 \label{sec:lowerbound}
 
-In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic algorithm.
-It is sufficient to have clusters of capacity $k\geq 3$ for this lower bound.
-For $k=2$ a constant competitive algorithm exists.
+%In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
+%It is sufficient to have clusters of capacity $k\geq 3$ for this lower bound.
+%For $k=2$ a constant competitive algorithm exists.
+%At the end of this section, we also
+%discuss an upper bound.
 
-The adversary constructs components of nodes, and issues only requests among nodes in the same component.
-If an online algorithm splits the components at any moment, the adversary issues requests to split pairs of nodes.
+\noindent \textbf{Lower bound.}
+The adversary constructs components of nodes based on the partitioning choices of a given online algorithm.
+Adversary issues requests between pairs of nodes that belong to the same component.
+If an online algorithm splits a component at any moment,
+the adversary issues requests to the split pair until the algorithm collocates them.
 We utilize the fact that the online algorithm does not know the components of the adversary, and it may split some components while evicting nodes.
-On the other hand, $\OPT$ never splits the components.
+On the other hand, $\OPT$ initially collocates all nodes of each component and never splits them again.
 
@@
 	$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
-	Let $I(C)$ denote the cluster where nodes of a component $C$ are located initially.
+	Let $I(C)$ denote the cluster where nodes of component $C$ are located initially.
 	%	\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
 	Fix any online algorithm \ALG{}.
-	We construct the input sequence for \ALG{} as follows.
-	First,
-	we issue $k-2$ (internal) requests so that \ALG{} can form a component of $k-1$
-	nodes on the cluster $S_1$.
-	Recall that internal requests does not incur any cost for \ALG{}.
+  We join some nodes into components, and produce the input sequence by observing actions of \ALG{} with respect to nodes of these components.
 	If \ALG{} at any point splits a component
@@
 	Next,
-	we issue a request between $x_0$ and $y_0$.
-	Since this is an external request,
-	\ALG{} joins them into one component and collocates them in some cluster other than $S_1$.
-	For this,
+	we merge components of $x_0$~and~$y_0$.
+  Since these reside on different clusters of \ALG{}, it produces an external request.
+  \ALG{} must collocate them in a~cluster other than $S_1$.
+	To this end,
 	\ALG{} moves to a new configuration
@@
 	Next,
-	we issue a request between $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
-	\ALG{} must collocate $x_1$ and $C$ on some cluster other than $S_1$ and
-	consequently replaces $x_1$ with some other (single) node $x_2$.
-	We repeat issuing requests between the single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
+	we merge the component of $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
+	\ALG{} must collocate $x_1$ and $C$ in a~cluster other than $S_1$ and
+	consequently replace $x_1$ with some other (single) node $x_2$.
+	We repeat merging the component of single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
   until there are only two single nodes left that  originate from the same cluster.
@@
 	Hence,
-	OPT pays for $4$ node migrations and
-	\ALG{} incurs at least one migration for each node in the sequence $X := x_0, x_1,\dots$.
+	OPT pays for $2$ node swaps and
+	\ALG{} incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
 	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
-	therefore $|X| \geq k \cdot \ell - k - \ell \in \Omega(k\cdot\ell)$.
+  therefore $|X| \geq k \cdot \ell - k - \ell$.
+  Finally, $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
 \end{proof}
 
-\section{An Optimal Algorithm for Clusters of Size 3}
-\label{sec:k3}
-
-In this section we show that the classic \emph{rent-or-buy} approach\cite{borodin-book, karlin-ski-rental} allows to obtain an optimal algorithm for $k=3$.
-Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one cluster until the end of a phase (that we define precisely later).
-
-%We define the algorithm \TAlg in the following way.
-Now we describe the algorithm \TAlg.
-It partitions nodes into components, and
-initially each node belongs to its own singleton component.
-For each pair of nodes, \TAlg maintains a counter, initiated to $0$. 
-Upon receiving a request to a pair that is not collocated in one cluster, it increases their counter by $1$.
-If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
-If no such partitioning exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
-
-%\maciek{Shouldn't we think about whole component cuts instead of single edge? Hmm. NO!}
-
-In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
-
-\begin{lemma}
-  \label{lem:1req}
-  A single repartition of \TAlg consists of at most $2$ migrations.
-\end{lemma}
-
-\begin{proof}
-  Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
-  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
-
-  %We distinguish between three types of clusters: $C_1, C_2, C_3$,
-  %which we define as follows.
-  %A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
-  %A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
-  %Finally, a cluster of type $C_3$  contains one component of size $3$.
-
-    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
-    Note that $U\neq V$,
-	 as otherwise the request would not trigger a~repartitioning.
-	 We consider cases upon types of clusters $U$ and $V$.
-    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
-    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
-    For two clusters, any configuration can be reached within two migrations, due to the fact that clusters are indistiguishable.
-
-    Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
-    Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
-    Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
-In this case \TAlg performs one migration, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
-\end{proof}
-
-\pagebreak
-
-\begin{theorem}
-  \TAlg is $O(\ell)$-competitive.
-\end{theorem}
-\begin{proof}
-  Fix a completed phase, and consider the state of \TAlg's counters at the end of it.
-  We consider the incomplete phase later in this proof.
-
-  As \TAlg is component respecting, it never increases any counter above $\alpha$.
-  We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
-
-  Let $\sigma$ be the input sequence that arrived during the phase.
-  Let $\sigma_{cost}$ be the requests that at the moment of arrival were external requests for \TAlg (these are the only requests that incurred a cost for \TAlg).
-  In our analysis, we partition $\sigma_{cost}$ into subsequences $\sigma_I$ and $\sigma_E$.
-  The sequence $\sigma_I$ (inter-component requests) are the requests from $\sigma_{cost}$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
-  The sequence $\sigma_E$ (extra-component requests) denotes the requests from $\sigma_{cost}$ that do not appear in $\sigma_I$.
-
-  %Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
-  %Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
-  Let $\TAlg(M)$ be the cost of migrations performed by \TAlg in this phase.
-  \TAlg performs at most $2 \ell$ component merge operations, as
-  exceeding this number means that a component of size $4$ exists, and the phase should have ended already.
-  Combining this with Lemma~\ref{lem:1req} gives us $\TAlg(M) \leq 4\alpha\cdot\ell$.
-  %Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
-
-  Now we bound $\TAlg(\sigma_I)$.
-  Cluster of type $C_3$ contributes at most $3 \alpha - 1$ to $\TAlg(\sigma_I)$, as $2$ of pairs of nodes from the component are saturated and contribute $\alpha$ each, and the third, unsaturated pair contributes at most $\alpha-1$.
-  Other cluster types contribute less: $C_1$ contributes $0$ and $C_2$ contributes $\alpha$.
-  Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq (3 \alpha-1)\cdot \ell \leq 3\alpha\cdot\ell$.
-
-  %We bound $A^E$ by $k^2 \cdot (\alpha - 1)$, as no more than $k^2$ pairs are unsaturated, and each of them contributes at most $\alpha -1$.
-  %\maciek{not needed most likely}
-
-  Moreover, \TAlg paid for all requests from $\sigma_E$, and thus $\TAlg(\sigma_E) = |\sigma_E|$.
-  In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 7\alpha\cdot \ell + |\sigma_E|$ during this phase.
-
-  \medskip
-
-  Now we lower-bound the cost of $\OPT$.
-  By $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$ we denote the cost of $\OPT$ on these input sequences (defined with respect to components of \TAlg in this phase).
-  By $\OPT(M)$ we denote the cost of migrations performed by $\OPT$ in this phase.
-
-  We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
-  While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts: we separately bound $\OPT$ by $\OPT(\sigma_I) + \OPT(M)$ and $\OPT(\sigma_E) + \OPT(M)$.
-  Combining those bounds gives us $\OPT \geq \max\{\OPT(\sigma_I) + \OPT(M), \OPT(\sigma_E) + \OPT(M)\} \geq (\OPT(\sigma_I) + \OPT(M)) / 2 + (\OPT(\sigma_E) + \OPT(M)) / 2$.
-
-  %Let $O^M$ be the cost of migrations performed by $\OPT$ during the phase.
-  %Let $O^I$ be the cost of serving requests between nodes that were put in one component by \TAlg during this phase.
-  %Let $O^E$ be the cost serving requests between nodes that \TAlg did not put in the same component during that phase.
-
-  %First, we estimate the cost related to $\sigma_I$.
-  We have $\OPT(M) + \OPT(\sigma_I) \geq \alpha$, as the phase ended when the components of \TAlg{} could not be partitioned without splitting them.
-  Hence, for every possible configuration of $\OPT$, there exists a non-collocated pair of nodes with at least $\alpha$ requests between them, and
-  $\OPT$ either served them remotely or performed a~migration.
-
-  \medskip
-  Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
-  In a~fixed configuration of $\OPT$, it may mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
-  Recall that $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
-  Faced with $W := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests $\sigma_E$ that it could not mitigate, $\OPT$ served some of them remotely and possibly performed some migrations to decrease its cost.
-
-  Now we estimate the cost of $\OPT(\sigma_E)$ while accounting savings from migrations.
-  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
-  This may allow to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
-  As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
-  By performing a single migration that costs $\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
-  Thus, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4 (\alpha-1)}$.
-  This gives us
-  \begin{equation*}
-    \OPT(\sigma_E) + \OPT(M)  \geq (|\sigma_E|-3(\alpha - 1)\cdot \ell) \cdot \frac{\alpha}{4 (\alpha-1)}
-    \geq |\sigma_E| / 4 - 4 \alpha \cdot \ell.
-  \end{equation*}
-  From that, we obtain $|\sigma_E| \geq 4(\OPT(\sigma_E)+\OPT(M)) + 16\alpha \cdot \ell$.
-  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, we are ready to bound the competitive ratio
-  \begin{equation*}
-    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{7\alpha \cdot \ell + |\sigma_E|}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{46\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 46 \ell = O(\ell).
-  \end{equation*}
-
-  \medskip
-
-  It remains to consider the last, unfinished phase.
-  Consider the case, where the unfinished phase is also the first one.
-  Then, we cannot charge $\OPT$ due to unability to partition the components.
-  Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
-  We charge $\OPT$ $\alpha$ for the first external $\alpha$ requests or a migration,
-  and we follow the analysis regarding the unsaturated requests.
-  If the input finished before first $\alpha$ external requests, then \TAlg is optimal.
-  Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
-  This way, the competitive ratio increases at most twofold.
-\end{proof}
-
-\section{An Optimal Algorithm for Perfect Partition}	\label{sec:ppl}
-%\subsection{Perfect Partition Model}\label{sec:perfectPartition}
-%\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
-%\maciek{ every algorithm must colocate every communicating pair.
-%	\mahmoud{updated!}}
-%\maciek{Note that the algorithm from Section~\ref{sec:k3} is $O(\ell)$-competitive for the Perfect Partition Model, as during the first phase, a perfect partition of input exists.}
-
-%\subsection{An $O(k\cdot \ell)$-competitive Algorithm for Perfect Partition}
-
-%A \emph{perfect partition} is a partitioning of nodes into $\ell$ clusters
-%such that it incurs no inter-cluster communication throughout the input sequence.
-%Equivalently, a perfect partitioning it an $\ell$-way cut of the communication graph of the input sequence.
-In this section,
-we give an algorithm for a restricted variant of  \OBRP{}, named \PPOBRP{},
+\noindent \textbf{Upper bound.}
+Now we give an algorithm for a restricted variant of  \OBRP{}, named \PPOBRP{},
 %\maciek{This is not sufficient! We also disallow serving requests remotely.}
@@
 
-\subsection{Perfect Partition Learner (\PPL) Algorithm}
-
 We assume \OPT begins with the initial configuration
@@
 This invariant ensures that \PPL does not pay too much while recovering $P_F$.
-
-\begin{algorithm}
-    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
-    \renewcommand{\algorithmicensure}{\textbf{Output:}}
-    \begin{algorithmic}
-%        \Require 
-%        $k, \ell$,
-%        initial configuration $P_I$,
-%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
-%        \Ensure A final configuration $P_F$ 
-        \STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$}
-        \STATE{$P_0 := P_I$}
-         \label{line:initcomponents}
-        \FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
-        \STATE Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
-        \IF{$C_1 \neq C_2$}
-        \STATE {Unite the two components into a single component $C'$ and
-        $\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$} \label{line:mergecomponents}
-        \IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
-	     \COMMENT{i.e.~if not in the same cluster}    
-    	}       
-        \STATE {$P_{t} = \mathit{repartition}(P_{t-1}, P_I, \mathcal{C})$} 
-         \COMMENT{move to $P$ closest to $P_I$}
-        \label{line:rebalance} 
-        \ENDIF
-        \ENDIF
-        \ENDFOR
-    \end{algorithmic}
-    \caption{Perfect Partition Learner (\PPL)}
-    \label{alg:ppl}
-      \end{algorithm}
-
 %      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
-      We note that the $\mathit{repartition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a perfect partition closest to $P_I$.
+      We note that the $\mathit{repartition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a~perfect partition closest to $P_I$.
 Hence it never moves to a configuration beyond distance $\Delta$.
+The scheme of the algorithm can be found in the Appendix \ref{apx:PPL}.  
 \begin{property} \label{prop:dist<OPT}
@@
 
- In Section~\ref{sec:lowerbound} we constructed a $\Omega(k \cdot \ell)$ for \OBRP{}.
- Note that the lower bound holds also in the perfect partition model, as the constructed input sequence allows \OPT to move to a perfect partition.
- The corollary is that PPL is optimal.
+% In Section~\ref{sec:lowerbound} we constructed a $\Omega(k \cdot \ell)$ for \OBRP{}.
+% Note that the lower bound holds also in the perfect partition model, as the constructed input sequence allows \OPT to move to a perfect partition.
+% The corollary is that PPL is optimal.
 
-% \section{Corollary and Future Directions}
+\section{The Online Problem}
+\label{sec:k3}
 
-% The gap between upper and lower bound is still substantial, as the best known algorithm is the component-respecting algorithm of   \cite{repartition-disc} with competitive ratio $O(k^2\cdot\ell^2)$.
+Let us now discuss the general online
+model where the request sequence
+can be arbitrary. We show that the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental} allows obtaining an optimal algorithm for $k=3$.
+Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one cluster until the end of a phase (that we define precisely later).
+
+%We define the algorithm \TAlg in the following way.
+Now we describe the algorithm \TAlg.
+It partitions nodes into components, and
+initially, each node belongs to its own singleton component.
+For each pair of nodes, \TAlg maintains a counter, initiated to $0$. 
+Upon receiving a request to a pair that is not collocated in one cluster, it increases their counter by $1$.
+If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
+If no such partitioning exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
+
+%\maciek{Shouldn't we think about whole component cuts instead of single edge? Hmm. NO!}
+
+In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
+Before bounding the competitive ratio of \TAlg, we introduce the lemma that estimates the cost of a single repartition of \TAlg. We defer its proof to Appendix~\ref{apx:omitted}.
+
+\begin{lemma}
+  \label{lem:1req}
+  During a single repartition, \TAlg exchanges at most $2$ pairs of nodes.
+\end{lemma}
+
+\begin{theorem}
+  \TAlg is $O(\ell)$-competitive.
+\end{theorem}
+\begin{proof}
+  Fix a completed phase, and consider the state of \TAlg's counters at the end of it.
+  We consider the incomplete phase later in this proof.
+
+  As \TAlg is component respecting, it never increases any counter above $\alpha$.
+  We say that the pair $(u, v)$ is \emph{saturated} if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
+  Let $\sigma$ be the input sequence that arrived during the phase.
+  Let $\sigma_{cost}$ be the requests that at the moment of arrival were external requests for \TAlg (these are the only requests that incurred a cost for \TAlg).
+  In our analysis, we partition $\sigma_{cost}$ into subsequences $\sigma_I$ and $\sigma_E$.
+  The sequence $\sigma_I$ (inter-component requests) are the requests from $\sigma_{cost}$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
+  The sequence $\sigma_E$ (extra-component requests) denotes the requests from $\sigma_{cost}$ that do not appear in $\sigma_I$.
+
+  %Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
+  %Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
+  Let $\TAlg(M)$ be the cost of migrations performed by \TAlg in this phase.
+  \TAlg performs at most $2 \ell$ component merge operations, as
+  exceeding this number means that a component of size $4$ exists, and the phase should have ended already.
+  Combining this with Lemma~\ref{lem:1req} gives us $\TAlg(M) \leq 8\alpha\cdot\ell$ (recall that an exchange costs $2\alpha$).
+  %Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
+
+  Now we bound $\TAlg(\sigma_I)$.
+  A~cluster of type $C_3$ contributes at most $3 \alpha - 1$ to $\TAlg(\sigma_I)$, as $2$ of pairs of nodes from the component are saturated and contribute $\alpha$ each, and the third, unsaturated pair contributes at most $\alpha-1$.
+  Other cluster types contribute less: $C_1$ contributes $0$ and $C_2$ contributes $\alpha$.
+  Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq (3 \alpha-1)\cdot \ell \leq 3\alpha\cdot\ell$.
+
+  %We bound $A^E$ by $k^2 \cdot (\alpha - 1)$, as no more than $k^2$ pairs are unsaturated, and each of them contributes at most $\alpha -1$.
+  %\maciek{not needed most likely}
+
+  Moreover, \TAlg paid for all requests from $\sigma_E$, and thus $\TAlg(\sigma_E) = |\sigma_E|$.
+  In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 11\alpha\cdot \ell + |\sigma_E|$ during this phase.
+
+  \medskip
+
+  Now we lower-bound the cost of $\OPT$.
+  By $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$ we denote the cost of $\OPT$ on these input sequences (defined with respect to components of \TAlg in this phase).
+  By $\OPT(M)$ we denote the cost of migrations performed by $\OPT$ in this phase.
+
+  We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
+  While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts: we separately bound $\OPT$ by $\OPT(\sigma_I) + \OPT(M)$ and $\OPT(\sigma_E) + \OPT(M)$.
+  Combining those bounds gives us $\OPT \geq \max\{\OPT(\sigma_I) + \OPT(M), \OPT(\sigma_E) + \OPT(M)\} \geq (\OPT(\sigma_I) + \OPT(M)) / 2 + (\OPT(\sigma_E) + \OPT(M)) / 2$.
+
+  %Let $O^M$ be the cost of migrations performed by $\OPT$ during the phase.
+  %Let $O^I$ be the cost of serving requests between nodes that were put in one component by \TAlg during this phase.
+  %Let $O^E$ be the cost serving requests between nodes that \TAlg did not put in the same component during that phase.
+
+  %First, we estimate the cost related to $\sigma_I$.
+  We have $\OPT(M) + \OPT(\sigma_I) \geq \alpha$, as the phase ended when the components of \TAlg{} could not be partitioned without splitting them.
+  Hence, for every possible configuration of $\OPT$, there exists a non-collocated pair of nodes with at least $\alpha$ requests between them, and
+  $\OPT$ either served them remotely or performed a~migration.
+
+  \medskip
+  Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
+  In a~fixed configuration of $\OPT$, it may mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
+  Recall that $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
+  Faced with $W := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests $\sigma_E$ that it could not mitigate, $\OPT$ served some of them remotely and possibly performed some migrations to decrease its cost.
+
+  Now we estimate the cost of $\OPT(\sigma_E)$ while accounting savings from migrations.
+  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
+  This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
+  As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
+  By performing a single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
+  Thus, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{2\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 2 - 2 \alpha \cdot \ell$.
+  From that, we obtain $|\sigma_E| \geq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
+  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, we are ready to bound the competitive ratio
+  \begin{equation*}
+    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + E/2} \leq \frac{27\alpha\cdot\ell + 2\cdot E}{\alpha + E} \leq 27 \ell = O(\ell).
+  \end{equation*}
+
+  \medskip
+
+  It remains to consider the last, unfinished phase.
+  Consider the case, where the unfinished phase is also the first one.
+  Then, we cannot charge $\OPT$ due to the inability to partition the components.
+  Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
+  We charge $\OPT$ $\alpha$ for the first external $\alpha$ requests or a migration,
+  and we follow the analysis regarding the unsaturated requests.
+  If the input finished before the first $\alpha$ external requests, then \TAlg is optimal.
+  Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
+  This way, the competitive ratio increases at most twofold.
+\end{proof}
 
@@
 
-\begin{appendix}
+\pagebreak
+\appendix
+\section{PPL Algorithm} \label{apx:PPL}
+\begin{algorithm}
+	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+	\renewcommand{\algorithmicensure}{\textbf{Output:}}
+	\begin{algorithmic}
+		%        \Require 
+		%        $k, \ell$,
+		%        initial configuration $P_I$,
+		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
+		%        \Ensure A final configuration $P_F$ 
+		\STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$}
+		\STATE{$P_0 := P_I$}
+		\label{line:initcomponents}
+		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
+		\STATE Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
+		\IF{$C_1 \neq C_2$}
+		\STATE {Unite the two components into a single component $C'$ and
+			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$} \label{line:mergecomponents}
+		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
+			\COMMENT{i.e.~if not in the same cluster}    
+		}       
+		\STATE {$P_{t} = \mathit{repartition}(P_{t-1}, P_I, \mathcal{C})$} 
+		\COMMENT{move to $P$ closest to $P_I$}
+		\label{line:rebalance} 
+		\ENDIF
+		\ENDIF
+		\ENDFOR
+	\end{algorithmic}
+	\caption{Perfect Partition Learner (\PPL)}
+	\label{alg:ppl}
+\end{algorithm}
 
-\section{Ommited proofs}
+\section{Omitted proofs}
+\label{apx:omitted}
 
-\section{Todos}
+\begin{proof}[Proof of Lemma~\ref{lem:1req}]
+  Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
+  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
 
-\todo{ACM format for review? Is there any specific format like this? Author anonymous, with line numbers etc.}
+  %We distinguish between three types of clusters: $C_1, C_2, C_3$,
+  %which we define as follows.
+  %A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+  %A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
+  %Finally, a cluster of type $C_3$  contains one component of size $3$.
 
-\todo{CCS concepts are mandatory}
+    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
+    Note that $U\neq V$,
+	 as otherwise, the request would not trigger a~repartitioning.
+	 We consider cases based on the types of clusters $U$ and $V$.
+    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
+    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
+    For two clusters, any configuration can be reached within two swaps, due to the fact that clusters are indistinguishable.
 
-\todo{Consistency: collocated at OR collocated in OR collocated on}
+    Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
+    Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
+    Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
+In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
+\end{proof}
 
-\end{appendix}
+\section{Omitted results}
+
+\begin{enumerate}
+  		 \item The lower bound of $\Omega(\ell)$ for the scenario with at most $(1/3-1/k)$-augmentation
+
+  		 \item Costly cascade example (it is easy to present $k^2$ and $k\cdot \sqrt{k}$ (with different approach), and then we can combine these to obtain $k^3$)
+
+  		 \item $k$-competitive algorithm for the ring communication pattern
+
+  		 \item $O(\ell)$-competitive algorithm for $k=4$ and $k=5$
+
+  \end{enumerate}
+
+\begin{CCSXML}
+	<ccs2012>
+	<concept>
+	<concept_id>10003033.10003068</concept_id>
+	<concept_desc>Networks~Network algorithms</concept_desc>
+	<concept_significance>500</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10010520.10010521.10010537.10003100</concept_id>
+	<concept_desc>Computer systems organization~Cloud computing</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10010520.10010521.10010537</concept_id>
+	<concept_desc>Computer systems organization~Distributed architectures</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	</ccs2012>
+\end{CCSXML}
 

### 2020-03-26 paper/paper.tex
Future
Table sketch

@@
 \usepackage{algorithm}
+
+\usepackage{balance}
+\usepackage[english]{babel}
+\usepackage{blindtext}
+\usepackage{amsthm,amsmath,amssymb,array,colortbl,graphicx,multirow}
+\usepackage{comment}
+\usepackage{balance}
+\usepackage{tikz}
+\usepackage{amsmath}
+\usetikzlibrary{patterns} %
+\usepackage{algorithm}
+\usepackage[font={footnotesize}]{subcaption}
+\usepackage[font={footnotesize}]{caption}
+\usepackage{breakcites}
+\usepackage{booktabs}
+\usepackage{diagbox}
+\usepackage{xcolor}
+\usepackage{colortbl}
+\usepackage{cleveref}
+\usepackage{enumitem}
 
@@
 
+  \section{Final}
+
+  \begin{table*}
+    \centering
+    \renewcommand{\arraystretch}{1.5}
+      \begin{tabular}{>{\centering\arraybackslash}p{4.5cm}|>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{4.5cm}}
+      \rowcolor{gray!50}
+      \textbf{Clusters/capacity} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
+        \textbf{$k=2$}& 3~\cite{repartition-disc} & 3~(\S 9.3) \\ 
+        \rowcolor{gray!25}
+        \textbf{$k=3$}&  $1357\cdot \ell - 245 = \Omega(\ell)$ (\S 9.3)& $O(k \cdot \ell)$~\cite{repartition-disc}\\
+        \textbf{$\ell=2$}&  $2k-1$ (\S 9.3)& $O(k^2)$~\cite{repartition-disc}\\
+        \rowcolor{gray!25}
+        general $k$ and $\ell$& $\Omega(k\cdot \ell)$ (\S 9.3)&$O(k^2 \cdot \ell^2)$~\cite{repartition-disc} \\
+      \end{tabular}
+                  \caption{Overview of the results. Most specific cases are tight. The general case still has a significant gap between the lower bound and the upper bound.
+                    }
+    \label{tab:overview}
+    \vspace{-7mm}
+  \end{table*}
+
+  Table with summary of all results. Tight for k=2, asymptotically tight for k=3, and a difference in general lk.
+
+  No bypassing model: reserach.
+
 \begin{CCSXML}

### 2020-03-26 paper/paper.tex
Glorious table ;)
Moving stuff around

@@
 similarly to the approach in~\cite{sigmetrics19_partitioning}.
+The overview of results is in Table~\ref{tab:overview}.
+
+\begin{table*}
+  \centering
+  \renewcommand{\arraystretch}{1.5}
+    \begin{tabular}{>{\centering\arraybackslash}p{4.5cm}|>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{4.5cm}}
+    \rowcolor{gray!50}
+    \textbf{Clusters/capacity} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
+      \textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & 3\hspace{0.3cm}(\S 9.3) \\ 
+      \rowcolor{gray!25}
+      \textbf{$k=3$}&  $1357\cdot \ell - 245 = \Omega(\ell)$\hspace{0.3cm}(\S 9.3)& $O(\ell)$\hspace{0.3cm}(\S 9.3)\\
+      \textbf{$\ell=2$}&  $2k-1$\hspace{0.3cm}(\S 9.3)&\hspace{0.3cm}$O(k^2)$\hspace{0.3cm}\cite{repartition-disc}\\
+      \rowcolor{gray!25}
+      general $k$ and $\ell$& $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S 9.3)&$O(k^2 \cdot \ell^2)$\hspace{0.3cm}\cite{repartition-disc} \\
+    \end{tabular}
+                \caption{Overview of the results. Most specific cases are tight. The general case still has a significant gap between the lower bound and the upper bound.
+                  }
+  \label{tab:overview}
+  \vspace{-7mm}
+\end{table*}
 
@@
 
-  \section{Final}
-
-  \begin{table*}
-    \centering
-    \renewcommand{\arraystretch}{1.5}
-      \begin{tabular}{>{\centering\arraybackslash}p{4.5cm}|>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{4.5cm}}
-      \rowcolor{gray!50}
-      \textbf{Clusters/capacity} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
-        \textbf{$k=2$}& 3~\cite{repartition-disc} & 3~(\S 9.3) \\ 
-        \rowcolor{gray!25}
-        \textbf{$k=3$}&  $1357\cdot \ell - 245 = \Omega(\ell)$ (\S 9.3)& $O(k \cdot \ell)$~\cite{repartition-disc}\\
-        \textbf{$\ell=2$}&  $2k-1$ (\S 9.3)& $O(k^2)$~\cite{repartition-disc}\\
-        \rowcolor{gray!25}
-        general $k$ and $\ell$& $\Omega(k\cdot \ell)$ (\S 9.3)&$O(k^2 \cdot \ell^2)$~\cite{repartition-disc} \\
-      \end{tabular}
-                  \caption{Overview of the results. Most specific cases are tight. The general case still has a significant gap between the lower bound and the upper bound.
-                    }
-    \label{tab:overview}
-    \vspace{-7mm}
-  \end{table*}
-
-  Table with summary of all results. Tight for k=2, asymptotically tight for k=3, and a difference in general lk.
-
-  No bypassing model: reserach.
-
 \begin{CCSXML}

### 2020-03-27 paper/paper.tex
On the learning problem

@@
 \section{The Learning Problem} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
+
+\maciek{TODO: Intro to the learning problem is missing. What is the problem definition. How is it different from the general problem. Emphasize the fact that OPT is constrained not to migrate.}
+
+\subsection{Lower bound.}
+
 \label{sec:lowerbound}
 
-%In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
-%It is sufficient to have clusters of capacity $k\geq 3$ for this lower bound.
-%For $k=2$ a constant competitive algorithm exists.
-%At the end of this section, we also
-%discuss an upper bound.
+In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
+It is sufficient to have clusters of capacity $k\geq 3$ for this lower bound.
+For $k=2$ a constant competitive algorithm exists.
+At the end of this section, we also
+discuss an upper bound.
 
-\noindent \textbf{Lower bound.}
+\maciek{TODO: better introduction of the concept of components IN THE LOWER BOUND. Perhaps other name than components to avoid confusion. Plus Figure.}
+
 The adversary constructs components of nodes based on the partitioning choices of a given online algorithm.
@@
 
-%\maciek{Here, we would like to introduce insights about the proof, emphasize the conditions for the bound to work (such as $k\geq 3$), and maybe compare to previously known lower bounds.
+\maciek{Here, we would like to introduce insights about the proof, emphasize the conditions for the bound to work (such as $k\geq 3$), and maybe compare to previously known lower bounds.
 Also, introduce the notation and definitions that are needed.}
@@
 
-\noindent \textbf{Upper bound.}
+\maciek{Remark on resource augmentation. The lower bound for k=3 works even with significant augmentation (linear).}
+
+\subsection{Upper bound.}
 Now we give an algorithm for a restricted variant of  \OBRP{}, named \PPOBRP{},

### 2020-03-27 paper/paper.tex
On the last epoch

@@
 \subsection{Upper bound.}
-Now we give an algorithm for a restricted variant of  \OBRP{}, named \PPOBRP{},
+Now we give an algorithm for a restricthed variant of  \OBRP{}, named \PPOBRP{},
 %\maciek{This is not sufficient! We also disallow serving requests remotely.}
@@
 
+\maciek{REALLY bad name. Both the learning problem and the general one is online.}
+
 Let us now discuss the general online
@@
 If no such partitioning exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
-
-%\maciek{Shouldn't we think about whole component cuts instead of single edge? Hmm. NO!}
 
@@
 
+  \maciek{Big techincal TODO. Formalize the sketched proof for the last phase.}
   It remains to consider the last, unfinished phase.

### 2020-03-27 paper/paper.tex
Figures description
Sketch LB

@@
 \newcommand{\todo}[1]{\noindent\color{brown}{todo: #1}\color{black}}
+
+\begin{CCSXML}
+	<ccs2012>
+	<concept>
+	<concept_id>10003033.10003068</concept_id>
+	<concept_desc>Networks~Network algorithms</concept_desc>
+	<concept_significance>500</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10010520.10010521.10010537.10003100</concept_id>
+	<concept_desc>Computer systems organization~Cloud computing</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10010520.10010521.10010537</concept_id>
+	<concept_desc>Computer systems organization~Distributed architectures</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	</ccs2012>
+\end{CCSXML}
 
@@
 
+\maciek{Figure: cluster types}
+
+\maciek{Figure: an unsaturated edge that is inside a component}
+
 In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
@@
 
-\begin{CCSXML}
-	<ccs2012>
-	<concept>
-	<concept_id>10003033.10003068</concept_id>
-	<concept_desc>Networks~Network algorithms</concept_desc>
-	<concept_significance>500</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10010520.10010521.10010537.10003100</concept_id>
-	<concept_desc>Computer systems organization~Cloud computing</concept_desc>
-	<concept_significance>300</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10010520.10010521.10010537</concept_id>
-	<concept_desc>Computer systems organization~Distributed architectures</concept_desc>
-	<concept_significance>300</concept_significance>
-	</concept>
-	</ccs2012>
-\end{CCSXML}
+\section{Lower bound $2k-1$ for two clusters}
+
+\maciek{Sketch only}
+
+We reduce the paging problem with the universe $U$ of size $2k$ and the cache of size $2k-1$.
+Recall that the lower bound for the competitive ratio for such paging problem is $2k-1$.
+
+We partition nodes into $3$ sets: $A$, $B$ and $C$.
+Sets has the following cardinality: $|A| = k-1$, $|B| = k-1$, and $|C| = 2$.
+We arbitratry assign all elements of the universe $U$ to nodes from $A\cup B\cup C$.
+Note that in every configuration, at least one of these sets is split, i.e., is not completely contained in one cluster.
 

### 2020-03-27 paper/paper.tex
Sketch of 2k-1 lb

@@
 
+\begin{theorem}
+  Every deterministic algorithm for BRP with $\ell=2$ is at least $c$-competitive, where $c=2k-1$.
+\end{theorem}
+
+\begin{proof}
 We partition nodes into $3$ sets: $A$, $B$ and $C$.
 Sets has the following cardinality: $|A| = k-1$, $|B| = k-1$, and $|C| = 2$.
-We arbitratry assign all elements of the universe $U$ to nodes from $A\cup B\cup C$.
+We arbitrarily assign all elements of the universe $U$ to nodes from $A\cup B\cup C$.
+
 Note that in every configuration, at least one of these sets is split, i.e., is not completely contained in one cluster.
+Intuitively, if a node is split from other nodes of any of sets $A,B,C$, then it is outside of the cache.
+
+Similarly to the lower bound TODO ref $\Omega(kl)$, we construct components of nodes, and issue the requests if the algorithm splits the component.
+(In the current proof, the components produced are oblivious to the configuration of the algorithm, yet the requests issued depend on if the algorithm splits the components.
+In $\Omega(kl)$, the components were constructed upon the actions of the algorithm, too.)
+
+Now, we show how to simulate an algorithm for the paging problem using the constructed BRP instance.
+Fix any algorithm for the paging problem.
+If a request to element $e$ comes, we identify its corresponding node $v_e$, and the set $S$ from $A,B,C$ that contains it.
+Then, the adversary produces a component $S$, and removes it right after the algorithm collocates the whole $S$ on one cluster.
+
+Now we bound the costs of the algorithm for any sequence of paging requests transformed the way described above.
+For any request, the algorithm pays $0$ if it has $S$ collocated, and otherwise it pays at least $2\alpha$ for at least one migration needed to achieve state where $S$ is collocated.
+
+Now we construct an offline algorithm OPT.
+OPT only resides in three types of configurations:
+\begin{enumerate}
+  \item With set $A$ collocated in one cluster, set $B$ collocated in the other, and set $B$ having one node in one cluster and the other node in the second
+  \item With set $A$ collocated in one cluster, with a single node from set $B$; on the second cluster, we have both nodes from set $C$, and $k-2$ nodes from set $B$
+  \item the symmetric configuration to the previous one, with set $A$ split, and sets $B$ and $C$ collocated
+\end{enumerate}
+To transition between any of these configuration types, a single exchange is needed, hence OPT pays $2\alpha$ for it.
+Within each type of configuration, OPT may exchange the split node with any node from his set for $2\alpha$.
+
+OPT observes the actions of the algorithm LFD (Least Forward Distance) for the caching problem, and performs adequate reconfigurations within feasible configurations described in the previous paragraph.
+Precisely, if LFD chooses to evict the element $e$ from the cache, OPT chooses the configuration with the corresponding node $v_e$ split, and pays $2\alpha$ for it.
+
+The well-known property of LFD is that it pays $1$ only after $\mathbb{C}$ different pages are revealed to the algorithm, where $\mathbb{C}$ is the size of its cache.
+In our case, $\mathbb{C} = 2k-1$, and hence the algorithm pays at least $2\alpha$ for each (transformed to BRP) paging request, and OPT pays $2\alpha$ every $2k-1$ paging requests.
+Hence, the claim holds.
+\end{proof}
 

### 2020-03-27 paper/paper.tex
Fig description
Rewrite

@@
 
-\maciek{Sketch only}
+\maciek{TODO: splitting nodes explain in the preliminaries}
 
-We reduce the paging problem with the universe $U$ of size $2k$ and the cache of size $2k-1$.
+We show the reduction from the paging problem with the universe $U$ of size $2k$ and the cache of size $2k-1$.
 Recall that the lower bound for the competitive ratio for such paging problem is $2k-1$.
 
+Similarly to the lower bound TODO ref $\Omega(kl)$, we construct components of nodes, and issue the requests if the algorithm splits the component.
+In the proof in this section, the components produced are oblivious to the configuration of the algorithm, yet the requests issued depend on if the algorithm splits the components.
+In $\Omega(kl)$ LB, the components were constructed upon the actions of the algorithm, too.
+
+\maciek{TODO: a figure of 3 sets split.}
+
 \begin{theorem}
-  Every deterministic algorithm for BRP with $\ell=2$ is at least $c$-competitive, where $c=2k-1$.
+  No deterministic algorithm for BRP with two clusters can achieve a competitive ratio smaller than $2k-1$.
 \end{theorem}
@@
 \begin{proof}
-We partition nodes into $3$ sets: $A$, $B$ and $C$.
-Sets has the following cardinality: $|A| = k-1$, $|B| = k-1$, and $|C| = 2$.
-We arbitrarily assign all elements of the universe $U$ to nodes from $A\cup B\cup C$.
+We partition nodes into $3$ sets $A$, $B$, and $C$.
+Sets have cardinalities $|A| = k-1$, $|B| = k-1$, and $|C| = 2$.
+Each element $e$ of the universe $U$ have a corresponding node $v_e \in A\cup B\cup C$.
 
-Note that in every configuration, at least one of these sets is split, i.e., is not completely contained in one cluster.
 Intuitively, if a node is split from other nodes of any of sets $A,B,C$, then it is outside of the cache.
+Note that in every configuration, at least one of these sets is split.
 
-Similarly to the lower bound TODO ref $\Omega(kl)$, we construct components of nodes, and issue the requests if the algorithm splits the component.
-(In the current proof, the components produced are oblivious to the configuration of the algorithm, yet the requests issued depend on if the algorithm splits the components.
-In $\Omega(kl)$, the components were constructed upon the actions of the algorithm, too.)
-
-Now, we show how to simulate an algorithm for the paging problem using the constructed BRP instance.
-Fix any algorithm for the paging problem.
-If a request to element $e$ comes, we identify its corresponding node $v_e$, and the set $S$ from $A,B,C$ that contains it.
-Then, the adversary produces a component $S$, and removes it right after the algorithm collocates the whole $S$ on one cluster.
+For any sequence of paging requests, we produce a corresponding BRP pairwise request sequence.
+To this end, fix an algorithm ALG for the paging problem.
+When a request to element $e$ comes, we identify the set $S \in \{ A,B,C \}$ that contains the corresponding node $v_e$.
+Then, the adversary forms a component $S$, and dismisses it right after the algorithm collocates the whole $S$ on one cluster.
+Recall that forming a component means that the adversary continuously issues requests to any two nodes of a component that are split.
 

### 2020-03-27 paper/paper.tex
Rewrite

@@
 
-\maciek{TODO: a figure of 3 sets split.}
+\maciek{TODO: example figure of 3 sets split. Example figure of three states of OPT}
 
@@
 Each element $e$ of the universe $U$ have a corresponding node $v_e \in A\cup B\cup C$.
+\maciek{define that A,B,C is collocated/split}
 
@@
 
-Now we bound the costs of the algorithm for any sequence of paging requests transformed the way described above.
-For any request, the algorithm pays $0$ if it has $S$ collocated, and otherwise it pays at least $2\alpha$ for at least one migration needed to achieve state where $S$ is collocated.
+Now we lower-bound the cost of the algorithm for any sequence of pairwise requests that were produced by the described transformation of paging requests.
+For any such request, the algorithm pays $0$ if it has $S$ collocated, and otherwise it pays at least $2\alpha$ for at least one migration needed to reach the state with $S$ collocated.
 
-Now we construct an offline algorithm OPT.
-OPT only resides in three types of configurations:
+We construct the following offline algorithm OPT.
+OPT only resides in three types of configurations with respect to sets $A,B,C$ (cf. Figure TODO).
+Without loss of generality, we describe the contents of clusters $X$ and $Y$:
 \begin{enumerate}
-  \item With set $A$ collocated in one cluster, set $B$ collocated in the other, and set $B$ having one node in one cluster and the other node in the second
-  \item With set $A$ collocated in one cluster, with a single node from set $B$; on the second cluster, we have both nodes from set $C$, and $k-2$ nodes from set $B$
-  \item the symmetric configuration to the previous one, with set $A$ split, and sets $B$ and $C$ collocated
+  \item In $X$ we have $A$ and one node from $C$; in $Y$ we have $B$ and the other node from $C$.
+  \item In $X$ we have $A$ and one node from $B$; in $Y$ we have $C$ and $k-2$ nodes from set $B$.
+  \item In $X$ we have $B$ and one node from $A$; in $Y$ we have $C$ and $k-2$ nodes from set $A$.
 \end{enumerate}
-To transition between any of these configuration types, a single exchange is needed, hence OPT pays $2\alpha$ for it.
-Within each type of configuration, OPT may exchange the split node with any node from his set for $2\alpha$.
+
+Now we describe the possible actions performed by OPT.
+Within each type of configuration with $S \in \{ A, B, C\}$ split, OPT may exchange the split node with any node from $S$.
+To transition between any of three configuration types, it suffices to perform a single exchange.
+Hence, for every action OPT pays $2\alpha$.
 

### 2020-03-27 paper/paper.tex
Rewrite

@@
   No deterministic algorithm for BRP with two clusters can achieve a competitive ratio smaller than $2k-1$.
+
+  \maciek{better formulation below}
+  Fix any $k$. If there exists a $\gamma$-competitive \ldots (SIDMA)
 \end{theorem}
@@
 
-For any sequence of paging requests, we produce a corresponding BRP pairwise request sequence.
+For any sequence of paging requests $I_{PAGE}$, we produce a corresponding BRP pairwise request sequence $I_{BRP}$.
 To this end, fix an algorithm ALG for the paging problem.
@@
 
-OPT observes the actions of the algorithm LFD (Least Forward Distance) for the caching problem, and performs adequate reconfigurations within feasible configurations described in the previous paragraph.
-Precisely, if LFD chooses to evict the element $e$ from the cache, OPT chooses the configuration with the corresponding node $v_e$ split, and pays $2\alpha$ for it.
+Finally, we describe the actions of OPT on the input sequence $I_{BRP}$.
+OPT observes the actions of the algorithm LFD (Least Forward Distance for the caching problem) on $I_{PAGE}$, and performs corresponding actions within its feasible configurations.
+Precisely, if LFD chooses to evict the element $e$ from the cache, OPT chooses the configuration with the corresponding node $v_e$ split.
+The cost of OPT is then $2\alpha \cdot LFD$.
 
-The well-known property of LFD is that it pays $1$ only after $\mathbb{C}$ different pages are revealed to the algorithm, where $\mathbb{C}$ is the size of its cache.
-In our case, $\mathbb{C} = 2k-1$, and hence the algorithm pays at least $2\alpha$ for each (transformed to BRP) paging request, and OPT pays $2\alpha$ every $2k-1$ paging requests.
-Hence, the claim holds.
+It is well-known that LFD is an optimal algorithm for paging, and any deterministic algorithm for paging with cache size $2k-1$ is at least $2k-1$-competitive.
+
+TODO: rewrite the conclusion in the similar way to sidma paper
 \end{proof}

### 2020-03-28 paper/paper.tex
Figure description
Reduction not a LB

@@
 
-\title{Brief Announcement: Tight Bounds for Online Balanced Repartitioning}
+\title{Improved Bounds for Online Balanced Repartitioning}
 
@@
 
-\section{Lower bound $2k-1$ for two clusters}
+\section{Dynamic Graph Bisection (two clusters case)}
+
+\subsection{Lower bound $2k-1$}
 
@@
 
-\maciek{TODO: example figure of 3 sets split. Example figure of three states of OPT}
+\maciek{TODO: figure of three states of OPT. It should fit next to the other lower bound figure?}
 
 \begin{theorem}
-  No deterministic algorithm for BRP with two clusters can achieve a competitive ratio smaller than $2k-1$.
+  %No deterministic algorithm for BRP with two clusters can achieve a competitive ratio smaller than $2k-1$.
 
-  \maciek{better formulation below}
-  Fix any $k$. If there exists a $\gamma$-competitive \ldots (SIDMA)
+  Fix any $k$.
+If there exists a $\gamma$-competitive deterministic algorithm $B$ for Dynamic Bisection Problem with cluster size $k$, then there exists a $\gamma$-competitive deterministic algorithm $P$ for the paging problem with cache size $2k-1$ and where the number of different pages is $2k$.
 \end{theorem}

### 2020-03-28 paper/paper.tex
Reduction not a LB

@@
   Fix any $k$.
-If there exists a $\gamma$-competitive deterministic algorithm $B$ for Dynamic Bisection Problem with cluster size $k$, then there exists a $\gamma$-competitive deterministic algorithm $P$ for the paging problem with cache size $2k-1$ and where the number of different pages is $2k$.
+If there exists a $\gamma$-competitive deterministic algorithm $B$ for Dynamic Bisection Problem with cluster size $k$, then there exists a $\gamma$-competitive deterministic algorithm $P$ for the paging problem with cache size $2k-2$ and where the number of different pages is $2k-1$.
 \end{theorem}
@@
 \begin{proof}
+  The pages are denoted by $p_1,p_2,\ldots,p_{2k-1}$. Without loss of generality, we assume that the initial cache is equal to $p_1,p_2,\ldots,p_{k-2}$.
+  We fix any input sequence $\sigma^P = (\sigma^P_1, \sigma^P_2, \sigma^P_3, \ldots)$ for the caching problem, where $\sigma^P_t$ denotes the $t$-th accessed page.
+  We show how to construct an online algorithm $P$ for the caching problem that proceeds in the following online manner.
+  The algorithm internally runs the algorithm~$B$, starting on the initial assignment of nodes to clusters that will be defined below.
+  For a requested page $\sigma^P_t$, it creates a subsequence of~communication requests for the BRP problem, runs $B$ on them, and serves $\sigma^P_t$ on the basis of $B$'s responses.
+
+  We use $2k$ nodes for the BRP problem: $2k-1$ caching nodes denoted $p_1,p_2, \ldots, p_{2k-1}$, and an auxiliary node $a$.
+  We partition nodes into sets $A = \{ a, p_1 \}$, $B = \{ p_2, p_3, \ldots, p_{k+1}\}$, and $C = \{ p_{k+2}, p_{k+3} \ldots, p_{2k-1}\}$.
+  Their cardinalities are $|A| = 2$ and $|B| = |C| = k-1$.
+
+  We say that the node configuration is \emph{well-aligned} if any of the following conditions hold:
+  \begin{enumerate}
+    \item One cluster contains $B$ and one node from $A$ and the other cluster contains $C$ and the other node from $A$.
+    \item One cluster contains $B$ and one node from $C$ and the other cluster contains $A$ and $k-2$ nodes from set $C$.
+    \item (symmetric to the previous condition) One cluster contains $C$ and one node from $B$ and the other cluster contains $A$ and $k-2$ nodes from set $B$.
+  \end{enumerate}
+  We show a bijection between cache contents and well-aligned configurations.
+  We say that a caching node $p_i \in S$, where $S \in \{ A, B, C\}$ belongs to the cache, if the entire set $S$ is collocated in a cluster.
+  Otherwise, if $S$ is split
+
+  one cluster
+contains the node $s$ and $k-1$ caching nodes, and the other cluster contains
+one caching node and all auxiliary nodes. There is a natural bijection between
+possible cache contents and well-aligned configurations: the cache consists of
+the $k-1$ caching nodes that are in the same cluster as node $s$. (Without loss
+of generality, we may assume that the cache of any caching algorithm is always
+full, i.e., consists of $k-1$ pages.) If the configuration $c$ of a~BRP
+algorithm is well-aligned, $\textsc{cache}(c)$ denotes the corresponding cache
+contents.
+The initial configuration for the BRP problem is the well-aligned
+configuration corresponding to the initial cache (pages
+$p_1,p_2,\ldots,p_{k-1}$ in the cache).
+
 We partition nodes into $3$ sets $A$, $B$, and $C$.

### 2020-03-28 paper/paper.tex
Reduction not a LB

@@
   We show a bijection between cache contents and well-aligned configurations.
-  We say that a caching node $p_i \in S$, where $S \in \{ A, B, C\}$ belongs to the cache, if the entire set $S$ is collocated in a cluster.
-  Otherwise, if $S$ is split
-
-  one cluster
-contains the node $s$ and $k-1$ caching nodes, and the other cluster contains
-one caching node and all auxiliary nodes. There is a natural bijection between
-possible cache contents and well-aligned configurations: the cache consists of
-the $k-1$ caching nodes that are in the same cluster as node $s$. (Without loss
-of generality, we may assume that the cache of any caching algorithm is always
-full, i.e., consists of $k-1$ pages.) If the configuration $c$ of a~BRP
-algorithm is well-aligned, $\textsc{cache}(c)$ denotes the corresponding cache
-contents.
-The initial configuration for the BRP problem is the well-aligned
-configuration corresponding to the initial cache (pages
-$p_1,p_2,\ldots,p_{k-1}$ in the cache).
-
-We partition nodes into $3$ sets $A$, $B$, and $C$.
-Sets have cardinalities $|A| = k-1$, $|B| = k-1$, and $|C| = 2$.
-Each element $e$ of the universe $U$ have a corresponding node $v_e \in A\cup B\cup C$.
-\maciek{define that A,B,C is collocated/split}
-
-Intuitively, if a node is split from other nodes of any of sets $A,B,C$, then it is outside of the cache.
-Note that in every configuration, at least one of these sets is split.
+  We say that a caching node $p_i \in S$, where $S \in \{ A, B, C\}$ belongs to the cache if the entire set $S$ is collocated in a cluster.
+  Otherwise, if $p_i$ is located on a different cluster than the other nodes from $S$, we say that it is outside of the cache.
+  (Without loss of generality, we may assume that the cache of any caching algorithm is always full, i.e., consists of $k-1$ pages.)
+  If the configuration $c$ of a~BRP algorithm is well-aligned, $\textsc{cache}(c)$ denotes the corresponding cache contents.
+  The initial configuration for the BRP problem is the well-aligned configuration corresponding to the initial cache (pages $p_1,p_2,\ldots,p_{2k-2}$ in the cache), with sets $A$ and $B$ non-split and the node $p_{2k-1}$ from the set $C$ split.
 

### 2020-03-28 paper/paper.tex
LB

@@
 \newcommand{\dist}{\textsf{dist}}
-\newcommand{\TAlg}{{\ensuremath{\textsf{ALG}_{3}}}\xspace} % we should change this name
+\newcommand{\TAlg}{{\ensuremath{\textsf{ALG}_{3}}}\xspace}
+
+\newcommand{\comm}{\textsc{comm}}
 
@@
 
+When a request $p$ to paging arrives, we create a subsequence $\comm(p)$ of communication requests for the BRP problem.
+We identify the corresponding caching node $p$ and the set $S \in \{ A,B,C \}$ that contains it.
+The construction of $\comm(p)$ depends on the actions of $B$.
+To abstract from the actions of $B$ in the construction, we use the notion of the adversary components.
+Forming an adversary component means that the adversary continuously issues requests to any two nodes of a component that are split.
+The request sequence $\comm(p)$ consists of the requests the adversary issues for $B$ after forming an adversary component $S$.
+When $B$ collocates all nodes of $S$, the adversary dismisses the adversary component $S$, i.e., no longer issues requests if $B$ splits the nodes of $S$.
+
+Note that $B$ must eventually achieve such a node configuration: otherwise its cost would be arbitrarily large while a sequence of repeated $\comm(\sigma^P_t)$ subsequences can be served at~a~constant cost.
+Therefore, the competitive ratio of~$B$ would then be unbounded.
+We denote the resulting sequence of $\comm(\sigma^P_t)$ subsequences by $\comm_t(\sigma^P_t)$.
+
+  \medskip
+
 For any sequence of paging requests $I_{PAGE}$, we produce a corresponding BRP pairwise request sequence $I_{BRP}$.

### 2020-03-28 paper/paper.tex
LB

@@
 \newcommand{\comm}{\textsc{comm}}
+\newcommand{\OFF}{\textsc{Off}\xspace}
 
@@
 
+\medskip
+
 When a request $p$ to paging arrives, we create a subsequence $\comm(p)$ of communication requests for the BRP problem.
@@
 
+To construct the response to the caching request $\sigma^P_t$, the algorithm $P$ runs $B$ on $\comm_t(\sigma^P_t)$.
+Right after processing $\comm_t(\sigma^P_t)$, the node configuration $c$ of $B$ is well-aligned and $\sigma^P_t$ is collocated with other nodes from $S_t$ where $S_t \in \{A,B, C\}$ and $\sigma^P_t \in S_t$.
+Hence, $P$ may change its cache configuration to $\textsc{cache}(c)$: such a response is feasible, because since $S_t$ is collocated, it is included by $P$ in the cache.
+Furthermore, we may relate the cost of $P$ to the cost of~$B$: If $P$ modifies the cache contents, the corresponding cost is $1$, as exactly one page has to
+be fetched.
+Such a change occurs only if $B$ changed the configuration (at a~cost of at least $2 \cdot \alpha$).
+Therefore, $2\cdot \alpha \cdot P(\sigma^P_t) \leq B(\comm_t(\sigma^P_t))$, which, summed over all requests from sequence $\sigma^P$, yields $2 \cdot
+\alpha \cdot P(\sigma^P) \leq B(\sigma^B)$.
+
+\medskip
+
+Now we show that there exists an (offline) solution \OFF to $\sigma^B$, whose cost is exactly ${2 \cdot \alpha \cdot \OPT(\sigma^P)}$.
+Recall that, for a caching request $\sigma^P_t$, $\sigma^B$ contains the~corresponding sequence $\comm_t(\sigma^P_t)$.
+Before serving the first request of~$\comm_t(\sigma^P_t)$, \OFF changes its state to a well-aligned configuration corresponding to the cache of $\OPT$ right after serving caching request $\sigma^P_t$.
+This ensures that the subsequence $\comm_t(\sigma^P_t)$ is free for \OFF.
+
+Now, we claim that the cost of node migration of \OFF is $2 \cdot \alpha$ (two caching nodes are swapped) if $\OPT$ performs a fetch, and 0 if
+$\OPT$ does not change its cache contents.
+Within each type of configuration with $S \in \{ B, C\}$ split, $\OPT$ may achieve a configuration with a different node from $S$ outside of the cache by exchanging the split node with any node from $S$.
+\maciek{Finished here}
+To transition between any of three configuration types ((1) with $A$ (2) with $B$ split and (3) with C split), it suffices to perform a single exchange.
+Therefore, $\OFF(\comm_t(\sigma^P_t)) = 2 \cdot \alpha \cdot \OPT(\sigma^P_t)$, which summed over the entire sequence $\sigma^P$ yields $\OFF(\sigma^B) = 2 \cdot \alpha \cdot \OPT(\sigma^P)$.
+
+As $B$ is $\gamma$-competitive for the BRP problem, there exists a constant $\beta$, such that for any sequence $\sigma^P$ and the corresponding sequence $\sigma^B$, it holds that $B(\sigma^B) \leq \gamma \cdot \OFF(\sigma^B) +
+\beta$.
+Combining this inequality with the inequalities between $P$ and $B$ and between \OFF and \OPT yields
+\[
+	2 \cdot \alpha \cdot P(\sigma^P) \leq 
+	B(\sigma^B) \leq \gamma \cdot \OFF(\sigma^B) + \beta
+	 =\gamma \cdot 2 \cdot \alpha \cdot \OPT(\sigma^P) + \beta,
+\]
+and therefore $P$ is $\gamma$-competitive.
   \medskip

### 2020-03-28 paper/paper.tex
LB comments

@@
 
-\maciek{TODO: better introduction of the concept of components IN THE LOWER BOUND. Perhaps other name than components to avoid confusion. Plus Figure.}
+\maciek{TODO: better introduction of the concept of adversary's components in the lower bound. Perhaps other name than adversary components to avoid confusion.}
+
+\maciek{Rename components in this section to adversary's components. Until the better name is figured out.}
+
+\maciek{Example figure of the construction and growing components.}
 

### 2020-03-29 paper/paper.tex
PPL reading
Reading and reviewing

@@
   The competitive ratio of any deterministic online algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$ for any $k\geq 3$.
+  \maciek{TODO: the exact bound, not asymptotic}
 \end{theorem}
@@
     $|M| \leq 2\cdot\OPT$.
+    \maciek{the comma between $\dist$ is confusing. BOTH are bounded?}
 \end{proof}
@@
     \PPL reaches the final configuration $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
+    \maciek{Consider not hiding the $2(k-1)\ell$ in the theorem. This is a better bound.}
 \end{theorem}
@@
      (i.e., the number of internal edges in $P_F$).
+     \maciek{Clarify why the number of internal edges bound the number of repartitions.}
     By Lemma \ref{lemma:rebalancecost},
@@
 
+\subsection{Perfect Partition reachable within one swap}
+
+(The version for two clusters.)
+
+Results: LB = UB = k+1.
+
 \subsection{Lower bound $2k-1$}

### 2020-03-29 paper/paper.tex
Sketches of 1PPL

@@
 
+\section{Perfect Partition reachable within one swap}
+
+  (The version for two clusters, but extensions possible.)
+  Results: LB = UB = k+1.
+
+  In this section we show that the competitive ratio of the BRP variant, where the perfect partition is reachable within one swap is exactly k+1.
+
+  Previously the best upper bound was $O(k^2)$ of DET \cite{repartition-disc}, and the best lower bound was $k-1$, utilizing the reduction from paging from \cite{repartition-disc}.
+  \maciek{It is possible to improve the reduction, i.e., to reduce from paging with cache $k$, but it is impossible to have reduction from paging with cache $k+1$.}
+
+  \begin{theorem}
+    If the perfect partition is reachable within one swap, the cost of any deterministic algorithm is at least k+1.
+  \end{theorem}
+
+  \begin{proof}
+    \maciek{pre-sketch}
+    A request to any pair. The algorithm puts it on the wrong cluster, evicting some node and paying $1$.
+    We continue to issue the requests to the evicted node (it changes every time), alg pays $k-2$.
+    We issue the final request, and alg must move the initial pair, paying $2$.
+  \end{proof}
+
+  \begin{theorem}
+    There exists a (k+1)-competitive algorithm for the BRP variant, where the perfect partition is reachable within one swap.
+  \end{theorem}
+
+  \begin{proof}
+    \maciek{early sketch}
+    The algorithm maintains components of nodes, and never splits them.
+    Intially all nodes are in singleton components.
+    The only choice is where to place the component, and possibly which components to evict.
+
+  The algorithm places the first component, denoted $Z$, on an arbitrary cluster, and evicts an arbitrary node (it is a singleton).
+  Consider a request $(a,b)$.
+  If $a \in Z$ (also symmetry $b \in Z$), then we move the entire component $Z$ to the cluster $I(b)$, evicting a singleton node that is missplaced (i.e., is not on its initial cluster), and an arbitrary node (we later show that there exists a node that is misplaced).
+  Otherwise, $I(a) = I(b)$, and we place the component there, evicting an arbitrary node (non-missplaced?).
+
+  If a component collocated with $Z$ grows to $k-1$, we evict $Z$ and terminate.
+  \end{proof}
+
 \section{Dynamic Graph Bisection (two clusters case)}
-
-\subsection{Perfect Partition reachable within one swap}
-
-(The version for two clusters.)
-
-Results: LB = UB = k+1.
 

### 2020-03-29 paper/paper.tex
notes

@@
 \subsection{Upper bound.}
-Now we give an algorithm for a restricthed variant of  \OBRP{}, named \PPOBRP{},
+Now we give an algorithm for a restricted variant of  \OBRP{}, named \PPOBRP{},
 %\maciek{This is not sufficient! We also disallow serving requests remotely.}
@@
 at the beginning and stays there perpetually.
+\maciek{Already mention that the LB is for that scenario, too?}
 %We refer this variant as \PPOBRP{}.
@@
 $P_F = F_1, \dots, F_{\ell}$.
- The \emph{distance} of a configuration $P = C_1, \dots, C_{\ell}$ from the initial configuration is the number of nodes in $P$ that do not reside in their initial cluster.
+ The \emph{distance} of a configuration $P = C_1, \dots, C_{\ell}$ from the initial configuration \maciek{, denoted $\dist(P)$} is the number of nodes in $P$ that do not reside in their initial cluster.
+ \maciek{It is OK to use single-argument $\dist$ (two-arg is a must in a definition, and it is already there).}
+ \maciek{We should denote distance $\Delta(P)$, and the current $\Delta$ should be R (like radius).}
     That is,
@@
   \begin{proof}
+    \maciek{Maybe PPL can be used instead?}
+
     \maciek{early sketch}

### 2020-03-29 paper/paper.tex
todos

@@
 
+\maciek{TODO: fill the table: results and references to sections.}
+
 \begin{table*}
@@
     \rowcolor{gray!50}
-    \textbf{Clusters/capacity} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
+    \textbf{Variant} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
       \textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & 3\hspace{0.3cm}(\S 9.3) \\ 
@@
       \rowcolor{gray!25}
-      general $k$ and $\ell$& $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S 9.3)&$O(k^2 \cdot \ell^2)$\hspace{0.3cm}\cite{repartition-disc} \\
+      $\ell \geq 3$& $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S 9.3)&$O(k^2 \cdot \ell^2)$\hspace{0.3cm}\cite{repartition-disc} \\
+      Perfect partition & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S 9.3)&$O(k \cdot \ell)$\hspace{0.3cm} (\S 9) \\
     \end{tabular}
-                \caption{Overview of the results. Most specific cases are tight. The general case still has a significant gap between the lower bound and the upper bound.
+                \caption{Overview of the results.
+Most specific cases are tight.
+The general case still has a significant gap between the lower bound and the upper bound.
+If either $k$ or $\ell$ is missing from a variant description, it is assumed to be arbitrary.
                   }
@@
 \noindent \textbf{Preliminaries.}
+
+\maciek{Here, we need a good and broad description of what a component is.}
+
 In this paper, we use the technique of maintaining \emph{(connected) components} of nodes, similar to previous approaches~\cite{repartition-disc}.
-A~\emph{component} is a subset of frequently communicating nodes.
+A~\emph{component} is a subset of frequently communicating nodes \maciek{This is not a definition, just some intuitions.
+More formal needed.}.
 We use this concept in both algorithms in this paper, but also in the lower bound, to incur high cost for any online algorithm that splits them.
@@
 
+\maciek{The text before the proof should have two parts: a description of adversary components and then general intro to the proof.}
+
 In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
 It is sufficient to have clusters of capacity $k\geq 3$ for this lower bound.
-For $k=2$ a constant competitive algorithm exists.
+For $k=2$, a constant competitive algorithm exists.
 At the end of this section, we also
@@
 
-\maciek{TODO: better introduction of the concept of adversary's components in the lower bound. Perhaps other name than adversary components to avoid confusion.}
+\maciek{TODO: rename everywhere: "adversary's components", i.e., all components in lower bounds should be called "ground sets"}
 
@@
 
+\maciek{Here: we abstract from the value of $\alpha$, but works for any alpha.}
+
 \begin{theorem}
@@
   Fix any online algorithm \ALG{}.
+
+  \maciek{This paragraph should be before the proof, merged with section about adversary components.}
   We join some nodes into components, and produce the input sequence by observing actions of \ALG{} with respect to nodes of these components.
@@
 	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
-  therefore $|X| \geq k \cdot \ell - k - \ell$.
+  therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate on exclusion?}.
   Finally, $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
@@
 at the beginning and stays there perpetually.
-\maciek{Already mention that the LB is for that scenario, too?}
+\maciek{Already mention that the LB is for that scenario}
 %We refer this variant as \PPOBRP{}.
@@
 Revealed edges, up to any point in the sequence,
-induce some connected subgraphs which grow as components.
+induce some connected subgraphs which grow as components \maciek{rewrite sentence, expand}.
 %An algorithm that always  keeps nodes belonging to the same component collocated is \emph{component respecting}.
@@
 Hence it never moves to a configuration beyond distance $\Delta$.
-The scheme of the algorithm can be found in the Appendix \ref{apx:PPL}.  
+The scheme of the algorithm can be found in the Algorithm \ref{alg:PPL}.
+
+\section{PPL Algorithm} \label{alg:PPL}
+\begin{algorithm}
+	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+	\renewcommand{\algorithmicensure}{\textbf{Output:}}
+	\begin{algorithmic}
+		%        \Require 
+		%        $k, \ell$,
+		%        initial configuration $P_I$,
+		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
+		%        \Ensure A final configuration $P_F$ 
+		\STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$}
+		\STATE{$P_0 := P_I$}
+		\label{line:initcomponents}
+		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
+		\STATE Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
+		\IF{$C_1 \neq C_2$}
+		\STATE {Unite the two components into a single component $C'$ and
+			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$} \label{line:mergecomponents}
+		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
+			\COMMENT{i.e.~if not in the same cluster}    
+		}       
+		\STATE {$P_{t} = \mathit{repartition}(P_{t-1}, P_I, \mathcal{C})$} 
+		\COMMENT{move to $P$ closest to $P_I$}
+		\label{line:rebalance} 
+		\ENDIF
+		\ENDIF
+		\ENDFOR
+	\end{algorithmic}
+	\caption{Perfect Partition Learner (\PPL)}
+	\label{alg:ppl}
+\end{algorithm}
+
 \begin{property} \label{prop:dist<OPT}
@@
     \PPL reaches the final configuration $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
-    \maciek{Consider not hiding the $2(k-1)\ell$ in the theorem. This is a better bound.}
+    \maciek{Do not hide that we have $2(k-1)\ell$ in the theorem. This is a better bound.}
 \end{theorem}
@@
 
+\begin{proof}
+    Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
+  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
+
+  %We distinguish between three types of clusters: $C_1, C_2, C_3$,
+  %which we define as follows.
+  %A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+  %A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
+  %Finally, a cluster of type $C_3$  contains one component of size $3$.
+
+    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
+    Note that $U\neq V$,
+	 as otherwise, the request would not trigger a~repartitioning.
+	 We consider cases based on the types of clusters $U$ and $V$.
+    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
+    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
+    For two clusters, any configuration can be reached within two swaps, due to the fact that clusters are indistinguishable.
+
+    Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
+    Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
+    Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
+In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
+\end{proof}
+
 \begin{theorem}
@@
   As \TAlg is component respecting, it never increases any counter above $\alpha$.
-  We say that the pair $(u, v)$ is \emph{saturated} if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
+  We say that the pair $(u, v)$ is \emph{saturated} if the counter has value $\alpha$, and \emph{unsaturated} otherwise \maciek{Rename: saturated/unsaturated edge instead of a pair}.
   Let $\sigma$ be the input sequence that arrived during the phase.
@@
 \appendix
-\section{PPL Algorithm} \label{apx:PPL}
-\begin{algorithm}
-	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-	\renewcommand{\algorithmicensure}{\textbf{Output:}}
-	\begin{algorithmic}
-		%        \Require 
-		%        $k, \ell$,
-		%        initial configuration $P_I$,
-		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
-		%        \Ensure A final configuration $P_F$ 
-		\STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$}
-		\STATE{$P_0 := P_I$}
-		\label{line:initcomponents}
-		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
-		\STATE Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
-		\IF{$C_1 \neq C_2$}
-		\STATE {Unite the two components into a single component $C'$ and
-			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$} \label{line:mergecomponents}
-		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
-			\COMMENT{i.e.~if not in the same cluster}    
-		}       
-		\STATE {$P_{t} = \mathit{repartition}(P_{t-1}, P_I, \mathcal{C})$} 
-		\COMMENT{move to $P$ closest to $P_I$}
-		\label{line:rebalance} 
-		\ENDIF
-		\ENDIF
-		\ENDFOR
-	\end{algorithmic}
-	\caption{Perfect Partition Learner (\PPL)}
-	\label{alg:ppl}
-\end{algorithm}
-
-\section{Omitted proofs}
-\label{apx:omitted}
-
-\begin{proof}[Proof of Lemma~\ref{lem:1req}]
-  Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
-  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
-
-  %We distinguish between three types of clusters: $C_1, C_2, C_3$,
-  %which we define as follows.
-  %A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
-  %A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
-  %Finally, a cluster of type $C_3$  contains one component of size $3$.
-
-    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
-    Note that $U\neq V$,
-	 as otherwise, the request would not trigger a~repartitioning.
-	 We consider cases based on the types of clusters $U$ and $V$.
-    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
-    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
-    For two clusters, any configuration can be reached within two swaps, due to the fact that clusters are indistinguishable.
-
-    Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
-    Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
-    Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
-In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
-\end{proof}
 
@@
 
-\maciek{TODO: splitting nodes explain in the preliminaries}
+\maciek{Dont include in WAOA}
 
@@
 Within each type of configuration with $S \in \{ B, C\}$ split, $\OPT$ may achieve a configuration with a different node from $S$ outside of the cache by exchanging the split node with any node from $S$.
-\maciek{Finished here}
 To transition between any of three configuration types ((1) with $A$ (2) with $B$ split and (3) with C split), it suffices to perform a single exchange.
@@
 and therefore $P$ is $\gamma$-competitive.
-  \medskip
-
-For any sequence of paging requests $I_{PAGE}$, we produce a corresponding BRP pairwise request sequence $I_{BRP}$.
-To this end, fix an algorithm ALG for the paging problem.
-When a request to element $e$ comes, we identify the set $S \in \{ A,B,C \}$ that contains the corresponding node $v_e$.
-Then, the adversary forms a component $S$, and dismisses it right after the algorithm collocates the whole $S$ on one cluster.
-Recall that forming a component means that the adversary continuously issues requests to any two nodes of a component that are split.
-
-Now we lower-bound the cost of the algorithm for any sequence of pairwise requests that were produced by the described transformation of paging requests.
-For any such request, the algorithm pays $0$ if it has $S$ collocated, and otherwise it pays at least $2\alpha$ for at least one migration needed to reach the state with $S$ collocated.
-
-We construct the following offline algorithm OPT.
-OPT only resides in three types of configurations with respect to sets $A,B,C$ (cf. Figure TODO).
-Without loss of generality, we describe the contents of clusters $X$ and $Y$:
-\begin{enumerate}
-  \item In $X$ we have $A$ and one node from $C$; in $Y$ we have $B$ and the other node from $C$.
-  \item In $X$ we have $A$ and one node from $B$; in $Y$ we have $C$ and $k-2$ nodes from set $B$.
-  \item In $X$ we have $B$ and one node from $A$; in $Y$ we have $C$ and $k-2$ nodes from set $A$.
-\end{enumerate}
-
-Now we describe the possible actions performed by OPT.
-Within each type of configuration with $S \in \{ A, B, C\}$ split, OPT may exchange the split node with any node from $S$.
-To transition between any of three configuration types, it suffices to perform a single exchange.
-Hence, for every action OPT pays $2\alpha$.
-
-Finally, we describe the actions of OPT on the input sequence $I_{BRP}$.
-OPT observes the actions of the algorithm LFD (Least Forward Distance for the caching problem) on $I_{PAGE}$, and performs corresponding actions within its feasible configurations.
-Precisely, if LFD chooses to evict the element $e$ from the cache, OPT chooses the configuration with the corresponding node $v_e$ split.
-The cost of OPT is then $2\alpha \cdot LFD$.
-
-It is well-known that LFD is an optimal algorithm for paging, and any deterministic algorithm for paging with cache size $2k-1$ is at least $2k-1$-competitive.
-
-TODO: rewrite the conclusion in the similar way to sidma paper
 \end{proof}

### 2020-04-02 paper/paper.tex
todos
LB improvement by k-1

@@
 similarly to the approach in~\cite{sigmetrics19_partitioning}.
-The overview of results is in Table~\ref{tab:overview}.
+The overview of known results is in Table~\ref{tab:overview}.
 
@@
 
-\maciek{Rename components in this section to adversary's components. Until the better name is figured out.}
-
 \maciek{Example figure of the construction and growing components.}
@@
   \maciek{TODO: the exact bound, not asymptotic}
+
+  \maciek{The bound $(k \cdot \ell - k - \ell) / 2$ can be easily improved to $(k \cdot \ell - k - 1) / 2$ by playing a simple caching game in the first cluster, instead of just giving away the $k-1$ "blocker" (the caching game would build it gradually)}
 \end{theorem}
@@
 In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
-Before bounding the competitive ratio of \TAlg, we introduce the lemma that estimates the cost of a single repartition of \TAlg. We defer its proof to Appendix~\ref{apx:omitted}.
+Before bounding the competitive ratio of \TAlg, we introduce the lemma that estimates the cost of a single repartition of \TAlg.
 

### 2020-04-02 paper/paper.tex
On cascade

@@
 
-\section{Perfect Partition reachable within one swap}
+\section{Maciek's sketches}
+
+\subsection{Cascade}
+
+For now, assume $\alpha=1$ (not to deal with inter-component communication).
+Consider an algorithm DET2, a modified version of DET from \cite{repartition-disc}.
+Modification is the following: when repartition is necessary, move to \emph{the closest} perfect partition of components.
+
+\begin{theorem}
+  A single repartition of DET2 migrates nodes among at most $p(k)^2+k$ clusters, where $p(k)$ is the number of integer partitions $k$.
+  The corollary is, the cost of the repartition is bounded by $O(k \cdot \alpha \cdot e^{c\cdot \sqrt{k}})$ for a constant $c$.
+\end{theorem}
+
+\subsection{Perfect Partition reachable within one swap}
+
+\maciek{Best to include it in the improved LB omega(lk) to help increase the lb by k-1}
 
@@
   \end{proof}
-
-\section{Dynamic Graph Bisection (two clusters case)}
 

### 2020-04-02 paper/paper.tex
On cascade

@@
       \rowcolor{gray!25}
-      \textbf{$k=3$}&  $1357\cdot \ell - 245 = \Omega(\ell)$\hspace{0.3cm}(\S 9.3)& $O(\ell)$\hspace{0.3cm}(\S 9.3)\\
+      \textbf{$k=3$}&  $27\cdot \ell $\hspace{0.3cm}(\S 9.3)& $O(\ell)$\hspace{0.3cm}(\S 9.3)\\
       \textbf{$\ell=2$}&  $2k-1$\hspace{0.3cm}(\S 9.3)&\hspace{0.3cm}$O(k^2)$\hspace{0.3cm}\cite{repartition-disc}\\
@@
 
-\maciek{Figure: cluster types}
+\maciek{Figure: cluster signatures}
 
@@
 
+\maciek{Rewrite: cluster types should be cluster signatures}
 In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
@@
 
+\begin{proof}
+  \maciek{Danger zone, nothing is finished here}
+
+  A \emph{signature} of a cluster $C$, denoted $S(C)$ is a multiset of sizes of components located in this cluster.
+  Note that there are $p(k)$ possible signatures.
+
+  Consider a repartition from a perspective of a single cluster.
+  Prior to the repartition, it had a signature $S^-(C)$, and after the repartition it has a signature $S^+(C)$.
+\end{proof}
+
 \subsection{Perfect Partition reachable within one swap}

### 2020-04-02 paper/paper.tex
On cascade

@@
     \textbf{Variant} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
-      \textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & 3\hspace{0.3cm}(\S 9.3) \\ 
+      \textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & TBD\hspace{0.3cm}(\S 9.3) \\ 
       \rowcolor{gray!25}
       \textbf{$k=3$}&  $27\cdot \ell $\hspace{0.3cm}(\S 9.3)& $O(\ell)$\hspace{0.3cm}(\S 9.3)\\
-      \textbf{$\ell=2$}&  $2k-1$\hspace{0.3cm}(\S 9.3)&\hspace{0.3cm}$O(k^2)$\hspace{0.3cm}\cite{repartition-disc}\\
+      \textbf{$\ell=2$}&  $k+1 = \Omega(k)$\hspace{0.3cm}(\S 9.3)&\hspace{0.3cm}$O(k^2)$\hspace{0.3cm}\cite{repartition-disc}\\
       \rowcolor{gray!25}
@@
 \begin{theorem}
-  A single repartition of DET2 migrates nodes among at most $p(k)^2+k$ clusters, where $p(k)$ is the number of integer partitions $k$.
+  A single repartition of DET2 migrates nodes among at most $p(k)^2+(k+1)$ clusters, where $p(k)$ is the number of integer partitions $k$.
   The corollary is, the cost of the repartition is bounded by $O(k \cdot \alpha \cdot e^{c\cdot \sqrt{k}})$ for a constant $c$.
@@
 
-  Consider a repartition from a perspective of a single cluster.
+  Fix a repartition.
+
+  Consider the repartition from the perspective of a single cluster.
   Prior to the repartition, it had a signature $S^-(C)$, and after the repartition it has a signature $S^+(C)$.
+
+  Now we construct a graph $G$ of component exchange among clusters.
+  The vertices are the clusters.
+  If a pair of clusters $C_1, C_2$ exchange components during the repartition, we produce an edge and we label it with the \emph{signature of exchange}, i.e., a pair of cluster signatures (both ways), limited to the exchanged clusters between $C_1,C_2$ only (i.e., not necessarily summing to $k$, and  note that also not necessarily having the same size!).
+
+  Proof by contradiction.
+  Assume that there exists a component in $G$ of size greater than $p(k)^2+(k+1)$.
+
 \end{proof}

### 2020-04-06 paper/waoa20clustering.tex
Cascade sketch

--- 2020-04-02 paper/paper.tex
@@
 
-  		 \item $O(\ell)$-competitive algorithm for $k=4$ and $k=5$
-
   \end{enumerate}
@@
 
-\subsection{Cascade}
+\subsection{Cascade bounds and an improved algorithm for general k, l}
 
@@
 
+Now we bound the cost of a single repartition of DET2 independent of $\ell$.
+First, we bound the cycle length in one of optimal solutions.
+
 \begin{theorem}
-  A single repartition of DET2 migrates nodes among at most $p(k)^2+k$ clusters, where $p(k)$ is the number of integer partitions $k$.
-  The corollary is, the cost of the repartition is bounded by $O(k \cdot \alpha \cdot e^{c\cdot \sqrt{k}})$ for a constant $c$.
+  \label{th:cascade-cycles}
+  Consider a single repartition of DET2 caused by node merge.
+  There exists a sequence of node migrations that realizes this repartition, where each node-disjoint cycle has length at most $k$.
+\end{theorem}
+Describe the graph. Edges and labels.
+
+Proof by contradiction.
+We call a sequence of node migrations that realize the repartition \emph{a solution to repartition}.
+Assume that each cost-optimal solution to repartition contains a cycle of length at least $k+1$.
+
+Fix a solution to repartition that has the shortest length of the longest cycle, and has the least edge-disjoint cycles of such length.
+
+\begin{proof}
+\end{proof}
+
+\begin{theorem}
+  Consider a single repartition of DET2.
+  The number of clusters it involves is bounded by $k^k$.
+  \label{th:cascade}
 \end{theorem}
@@
 \begin{proof}
-  \maciek{Danger zone, nothing is finished here}
+  Each node has degree at most $k$. Combine with Theorem \ref{th:cascade-cycles}.
 
-  A \emph{signature} of a cluster $C$, denoted $S(C)$ is a multiset of sizes of components located in this cluster.
-  Note that there are $p(k)$ possible signatures.
+Important technical details:
+\begin{itemize}
+  \item There are no cycles attached to another cycles by just one vertex (by equivalent signature skipping).
+This is important, as otherwise we just have a bound on cycles, but the graph contains a chain of cycles.
+  Another formulation (Stefan's): each edge is a part of a cycle that goes through the "origin" part of the graph.
+\end{itemize}
+\end{proof}
 
-  Consider a repartition from a perspective of a single cluster.
-  Prior to the repartition, it had a signature $S^-(C)$, and after the repartition it has a signature $S^+(C)$.
+\begin{theorem}
+  Algorithm DET2 is $O(\ell \cdot k\cdot k^k)$-competitive.
+\end{theorem}
+
+\begin{proof}
+  Use Theorem~\ref{th:cascade}. Then, linearize the inner-component communication using the technique from SIDMA.
 \end{proof}

### 2020-04-20 paper/waoa20clustering.tex
Comments
k=2 proof sketch

@@
 \usepackage{blindtext}
-\usepackage{amsthm,amsmath,amssymb,array,colortbl,graphicx,multirow}
+\usepackage{amsthm,amsmath,array,colortbl,graphicx,multirow}
 \usepackage{comment}
@@
 \end{CCSXML}
+
+% for my small screen
+\setlength{\textwidth}{8cm}
 
@@
 %The objective is to jointly minimize communication and migration costs while serving the sequence.
+
+\maciek{Let's follow the SIDMA notation here. For example, a "single" node (the one that belongs to singleton component) is called "trivial" there.}
 
@@
 To transition between any of three configuration types ((1) with $A$ (2) with $B$ split and (3) with C split), it suffices to perform a single exchange.
-Therefore, $\OFF(\comm_t(\sigma^P_t)) = 2 \cdot \alpha \cdot \OPT(\sigma^P_t)$, which summed over the entire sequence $\sigma^P$ yields $\OFF(\sigma^B) = 2 \cdot \alpha \cdot \OPT(\sigma^P)$.
+Therefore, $\OFF(\comm_t(\sigma^P_t)) = 2 \cdot \alpha \cdot \OPT(\sigma^P_t)$, which summed over the entire sequence $\sigma^P$ 5yields $\OFF(\sigma^B) = 2 \cdot \alpha \cdot \OPT(\sigma^P)$.
 
@@
 
+Algorithm Alg:
+maintain a counter $C_{\set{u,v}}$ for each pair of nodes $\set{u,v}$ and increment it on every remote request between $u$ and $v$.
+Once $C_{\set{u,v}} = \lambda:=2\alpha$,
+reset the counter and collocate the pair arbitrarily in one of the two clusters where $u$ or $v$ resides.
+
+\begin{theorem}
+	The algorithm is 4-competitive.
+\end{theorem}
+\begin{proof}
+\mahmoud{still proof sketch}
+We charge both OPT and Alg on each request that triggers a collocation.
+Let $C^{t}_{\set{u,v}}$ denote the  value of $C_{\set{u,v}}$ immediately after arrival of $\sigma_t$.
+Assume on request $\sigma_{t}=\set{u,v}$,
+Alg collocates the pair,
+which means $ C^t_{\set{u,v}} = \lambda$ immediately before Alg resets the counter.
+Let $t''<t$ be the latest time (i.e.~request) when Alg collocates $\set{u,v}$ before $t$.
+Let $t', t'' < t' < t$ be the  time when Alg separates $\set{u,v}$
+in order to collocate some other pair sharing a node,
+e.g., the pair $\set{u,v'}, v' \neq v$.
+%e.g., the request $\sigma_{t'} = \set{u,v'}$.
+%By Alg's definition,
+%$C^{t'}_{\set{u,v'}}  = C^t_{\set{u,v}} = \lambda$.
+We denote the set of all requests to $\set{u,v}$
+between times $t'$ and $t$ by $\sigma_{\set{u,v}}[t',t]$.
+Thus,
+$| \sigma_{\set{u,v}}[t',t] | = \lambda$.
+The overall cost of requests that increases the counter,
+i.e.~requests in $\sigma_{\set{u,v}}[t',t]$,
+and the moving cost incurred by Alg,
+ is $\lambda + 2\alpha$.
+%That is, the cost of all requests in $\sigma_{\set{u,v}}[t',t]$.
+We distinguish three cases for the cost incurred by OPT with regards to $\set{u,v}$.
+i)
+OPT has $\set{u,v}$ separated through all requests in $\sigma_{\set{u,v}}[t',t]$
+and therefore it pays $\lambda$.
+ii)  
+OPT collocates $\set{u,v}$ some time during $[t'',t]$.
+Recall that whenever Alg collocates a pair,
+we charge OPT only for requests up to that point (i.e., not for later requests).
+Hence,
+we charge the cost of  this collocation to OPT for the first time.
+In particular,
+we charge OPT the cost of moving one node to the cluster of the other,
+e.g.~$u$ to $v$, w.l.o.g.
+Therefore in this case OPT pays $\alpha$.
+iii)
+OPT has $\set{u,v}$ collocated before $t''$ and keeps them collocated during $[t'',t]$.
+Obviously in this case OPT doesn't pays any cost for requests in $\sigma_{\set{u,v}}[t'',t]$.
+However,
+OPT must pay $\lambda$ for requests to $\set{u,v'}$.
+By induction,
+at time $t'$,
+we charge OPT $\alpha$ (out of $\lambda=2\alpha$) for this pair.
+Therefore,
+we can charge OPT another $\alpha$ to account for requests $\sigma_{\set{u,v}}[t'',t]$.
+It can be easily verified that any deviation from these cases only increases the cost to OPT.
+Thus,
+OPT pays at least $\alpha$ for each collocation by Alg
+which gives the competitive ratio 
+$(\lambda + 2\alpha)/ \alpha = 4$.
+
+\end{proof}
+
 \end{document}

### 2020-04-22 paper/disc20clustering.tex
Disc not waoa
Remove: perfect partition reachable within one swap

--- 2020-04-20 paper/waoa20clustering.tex
@@
 % for my small screen
-\setlength{\textwidth}{8cm}
+%\setlength{\textwidth}{8cm}
 
@@
 
+       \item lower and upper bound $k+1$ for two clusters, perfect partition and the optimal configuration reachable within one swap
+
   \end{enumerate}
@@
   Use Theorem~\ref{th:cascade}. Then, linearize the inner-component communication using the technique from SIDMA.
-\end{proof}
-
-\subsection{Perfect Partition reachable within one swap}
-
-\maciek{Best to include it in the improved LB omega(lk) to help increase the lb by k-1}
-
-  (The version for two clusters, but extensions possible.)
-  Results: LB = UB = k+1.
-
-  In this section we show that the competitive ratio of the BRP variant, where the perfect partition is reachable within one swap is exactly k+1.
-
-  Previously the best upper bound was $O(k^2)$ of DET \cite{repartition-disc}, and the best lower bound was $k-1$, utilizing the reduction from paging from \cite{repartition-disc}.
-  \maciek{It is possible to improve the reduction, i.e., to reduce from paging with cache $k$, but it is impossible to have reduction from paging with cache $k+1$.}
-
-  \begin{theorem}
-    If the perfect partition is reachable within one swap, the cost of any deterministic algorithm is at least k+1.
-  \end{theorem}
-
-  \begin{proof}
-    \maciek{pre-sketch}
-    A request to any pair. The algorithm puts it on the wrong cluster, evicting some node and paying $1$.
-    We continue to issue the requests to the evicted node (it changes every time), alg pays $k-2$.
-    We issue the final request, and alg must move the initial pair, paying $2$.
-  \end{proof}
-
-  \begin{theorem}
-    There exists a (k+1)-competitive algorithm for the BRP variant, where the perfect partition is reachable within one swap.
-  \end{theorem}
-
-  \begin{proof}
-    \maciek{Maybe PPL can be used instead?}
-
-    \maciek{early sketch}
-    The algorithm maintains components of nodes, and never splits them.
-    Intially all nodes are in singleton components.
-    The only choice is where to place the component, and possibly which components to evict.
-
-  The algorithm places the first component, denoted $Z$, on an arbitrary cluster, and evicts an arbitrary node (it is a singleton).
-  Consider a request $(a,b)$.
-  If $a \in Z$ (also symmetry $b \in Z$), then we move the entire component $Z$ to the cluster $I(b)$, evicting a singleton node that is missplaced (i.e., is not on its initial cluster), and an arbitrary node (we later show that there exists a node that is misplaced).
-  Otherwise, $I(a) = I(b)$, and we place the component there, evicting an arbitrary node (non-missplaced?).
-
-  If a component collocated with $Z$ grows to $k-1$, we evict $Z$ and terminate.
-  \end{proof}
-
-\subsection{Lower bound $2k-1$}
-
-\maciek{Dont include in WAOA}
-
-We show the reduction from the paging problem with the universe $U$ of size $2k$ and the cache of size $2k-1$.
-Recall that the lower bound for the competitive ratio for such paging problem is $2k-1$.
-
-Similarly to the lower bound TODO ref $\Omega(kl)$, we construct components of nodes, and issue the requests if the algorithm splits the component.
-In the proof in this section, the components produced are oblivious to the configuration of the algorithm, yet the requests issued depend on if the algorithm splits the components.
-In $\Omega(kl)$ LB, the components were constructed upon the actions of the algorithm, too.
-
-\maciek{TODO: figure of three states of OPT. It should fit next to the other lower bound figure?}
-
-\begin{theorem}
-  %No deterministic algorithm for BRP with two clusters can achieve a competitive ratio smaller than $2k-1$.
-
-  Fix any $k$.
-If there exists a $\gamma$-competitive deterministic algorithm $B$ for Dynamic Bisection Problem with cluster size $k$, then there exists a $\gamma$-competitive deterministic algorithm $P$ for the paging problem with cache size $2k-2$ and where the number of different pages is $2k-1$.
-\end{theorem}
-
-\begin{proof}
-  The pages are denoted by $p_1,p_2,\ldots,p_{2k-1}$. Without loss of generality, we assume that the initial cache is equal to $p_1,p_2,\ldots,p_{k-2}$.
-  We fix any input sequence $\sigma^P = (\sigma^P_1, \sigma^P_2, \sigma^P_3, \ldots)$ for the caching problem, where $\sigma^P_t$ denotes the $t$-th accessed page.
-  We show how to construct an online algorithm $P$ for the caching problem that proceeds in the following online manner.
-  The algorithm internally runs the algorithm~$B$, starting on the initial assignment of nodes to clusters that will be defined below.
-  For a requested page $\sigma^P_t$, it creates a subsequence of~communication requests for the BRP problem, runs $B$ on them, and serves $\sigma^P_t$ on the basis of $B$'s responses.
-
-  We use $2k$ nodes for the BRP problem: $2k-1$ caching nodes denoted $p_1,p_2, \ldots, p_{2k-1}$, and an auxiliary node $a$.
-  We partition nodes into sets $A = \{ a, p_1 \}$, $B = \{ p_2, p_3, \ldots, p_{k+1}\}$, and $C = \{ p_{k+2}, p_{k+3} \ldots, p_{2k-1}\}$.
-  Their cardinalities are $|A| = 2$ and $|B| = |C| = k-1$.
-
-  We say that the node configuration is \emph{well-aligned} if any of the following conditions hold:
-  \begin{enumerate}
-    \item One cluster contains $B$ and one node from $A$ and the other cluster contains $C$ and the other node from $A$.
-    \item One cluster contains $B$ and one node from $C$ and the other cluster contains $A$ and $k-2$ nodes from set $C$.
-    \item (symmetric to the previous condition) One cluster contains $C$ and one node from $B$ and the other cluster contains $A$ and $k-2$ nodes from set $B$.
-  \end{enumerate}
-  We show a bijection between cache contents and well-aligned configurations.
-  We say that a caching node $p_i \in S$, where $S \in \{ A, B, C\}$ belongs to the cache if the entire set $S$ is collocated in a cluster.
-  Otherwise, if $p_i$ is located on a different cluster than the other nodes from $S$, we say that it is outside of the cache.
-  (Without loss of generality, we may assume that the cache of any caching algorithm is always full, i.e., consists of $k-1$ pages.)
-  If the configuration $c$ of a~BRP algorithm is well-aligned, $\textsc{cache}(c)$ denotes the corresponding cache contents.
-  The initial configuration for the BRP problem is the well-aligned configuration corresponding to the initial cache (pages $p_1,p_2,\ldots,p_{2k-2}$ in the cache), with sets $A$ and $B$ non-split and the node $p_{2k-1}$ from the set $C$ split.
-
-\medskip
-
-When a request $p$ to paging arrives, we create a subsequence $\comm(p)$ of communication requests for the BRP problem.
-We identify the corresponding caching node $p$ and the set $S \in \{ A,B,C \}$ that contains it.
-The construction of $\comm(p)$ depends on the actions of $B$.
-To abstract from the actions of $B$ in the construction, we use the notion of the adversary components.
-Forming an adversary component means that the adversary continuously issues requests to any two nodes of a component that are split.
-The request sequence $\comm(p)$ consists of the requests the adversary issues for $B$ after forming an adversary component $S$.
-When $B$ collocates all nodes of $S$, the adversary dismisses the adversary component $S$, i.e., no longer issues requests if $B$ splits the nodes of $S$.
-
-Note that $B$ must eventually achieve such a node configuration: otherwise its cost would be arbitrarily large while a sequence of repeated $\comm(\sigma^P_t)$ subsequences can be served at~a~constant cost.
-Therefore, the competitive ratio of~$B$ would then be unbounded.
-We denote the resulting sequence of $\comm(\sigma^P_t)$ subsequences by $\comm_t(\sigma^P_t)$.
-
-To construct the response to the caching request $\sigma^P_t$, the algorithm $P$ runs $B$ on $\comm_t(\sigma^P_t)$.
-Right after processing $\comm_t(\sigma^P_t)$, the node configuration $c$ of $B$ is well-aligned and $\sigma^P_t$ is collocated with other nodes from $S_t$ where $S_t \in \{A,B, C\}$ and $\sigma^P_t \in S_t$.
-Hence, $P$ may change its cache configuration to $\textsc{cache}(c)$: such a response is feasible, because since $S_t$ is collocated, it is included by $P$ in the cache.
-Furthermore, we may relate the cost of $P$ to the cost of~$B$: If $P$ modifies the cache contents, the corresponding cost is $1$, as exactly one page has to
-be fetched.
-Such a change occurs only if $B$ changed the configuration (at a~cost of at least $2 \cdot \alpha$).
-Therefore, $2\cdot \alpha \cdot P(\sigma^P_t) \leq B(\comm_t(\sigma^P_t))$, which, summed over all requests from sequence $\sigma^P$, yields $2 \cdot
-\alpha \cdot P(\sigma^P) \leq B(\sigma^B)$.
-
-\medskip
-
-Now we show that there exists an (offline) solution \OFF to $\sigma^B$, whose cost is exactly ${2 \cdot \alpha \cdot \OPT(\sigma^P)}$.
-Recall that, for a caching request $\sigma^P_t$, $\sigma^B$ contains the~corresponding sequence $\comm_t(\sigma^P_t)$.
-Before serving the first request of~$\comm_t(\sigma^P_t)$, \OFF changes its state to a well-aligned configuration corresponding to the cache of $\OPT$ right after serving caching request $\sigma^P_t$.
-This ensures that the subsequence $\comm_t(\sigma^P_t)$ is free for \OFF.
-
-Now, we claim that the cost of node migration of \OFF is $2 \cdot \alpha$ (two caching nodes are swapped) if $\OPT$ performs a fetch, and 0 if
-$\OPT$ does not change its cache contents.
-Within each type of configuration with $S \in \{ B, C\}$ split, $\OPT$ may achieve a configuration with a different node from $S$ outside of the cache by exchanging the split node with any node from $S$.
-To transition between any of three configuration types ((1) with $A$ (2) with $B$ split and (3) with C split), it suffices to perform a single exchange.
-Therefore, $\OFF(\comm_t(\sigma^P_t)) = 2 \cdot \alpha \cdot \OPT(\sigma^P_t)$, which summed over the entire sequence $\sigma^P$ 5yields $\OFF(\sigma^B) = 2 \cdot \alpha \cdot \OPT(\sigma^P)$.
-
-As $B$ is $\gamma$-competitive for the BRP problem, there exists a constant $\beta$, such that for any sequence $\sigma^P$ and the corresponding sequence $\sigma^B$, it holds that $B(\sigma^B) \leq \gamma \cdot \OFF(\sigma^B) +
-\beta$.
-Combining this inequality with the inequalities between $P$ and $B$ and between \OFF and \OPT yields
-\[
-	2 \cdot \alpha \cdot P(\sigma^P) \leq 
-	B(\sigma^B) \leq \gamma \cdot \OFF(\sigma^B) + \beta
-	 =\gamma \cdot 2 \cdot \alpha \cdot \OPT(\sigma^P) + \beta,
-\]
-and therefore $P$ is $\gamma$-competitive.
 \end{proof}

### 2020-04-22 paper/disc20clustering.tex
Paper restructure

@@
 \newcommand\maciek[1]{\color{brown}\textbf{\\ Maciek: #1}\color{black}}
+%\newcommand\maciek[1]{}
 
@@
 
-\maciek{Example figure of the construction and growing components.}
+\begin{figure}[H]
+  \centering
+    \includegraphics[width=0.3\textwidth]{figs/substitute}
+    \caption{Example figure of the construction of the LB and growing components.}
+    \label{fig:nptree-construction}
+  \end{figure}
 
@@
 
+\maciek{Rewrite above paragraph to say that our algorithm is essentially DET from DISC16, with concretised reconfiguration strategy (i.e., the cheapest)}
+
+\subsection{Improved analysis of the component-based algorithm}
+
 %We define the algorithm \TAlg in the following way.
@@
 
-\maciek{Figure: cluster signatures}
-
-\maciek{Figure: an unsaturated edge that is inside a component}
-
-\maciek{Rewrite: cluster types should be cluster signatures}
-In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
-Before bounding the competitive ratio of \TAlg, we introduce the lemma that estimates the cost of a single repartition of \TAlg.
-
-\begin{lemma}
-  \label{lem:1req}
-  During a single repartition, \TAlg exchanges at most $2$ pairs of nodes.
-\end{lemma}
-
-\begin{proof}
-    Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
-  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
-
-  %We distinguish between three types of clusters: $C_1, C_2, C_3$,
-  %which we define as follows.
-  %A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
-  %A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
-  %Finally, a cluster of type $C_3$  contains one component of size $3$.
-
-    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
-    Note that $U\neq V$,
-	 as otherwise, the request would not trigger a~repartitioning.
-	 We consider cases based on the types of clusters $U$ and $V$.
-    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
-    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
-    For two clusters, any configuration can be reached within two swaps, due to the fact that clusters are indistinguishable.
-
-    Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
-    Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
-    Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
-In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
-\end{proof}
+  \begin{figure}[H]
+    \centering
+      \includegraphics[width=0.3\textwidth]{figs/substitute}
+      \caption{an unsaturated edge that is inside a component}
+    \end{figure}
 
@@
 
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{references}
+\section{Bounding the cost of a single reconfiguration}
 
-\pagebreak
-\appendix
+\subsection{Warm-up: bounding the reconfiguration for $k=3$}
 
-\section{Omitted results}
+\begin{figure}[H]
+  \centering
+    \includegraphics[width=0.3\textwidth]{figs/substitute}
+    \caption{Cluster signatures in warm-up k=3 proof}
+  \end{figure}
 
-\begin{enumerate}
-  		 \item The lower bound of $\Omega(\ell)$ for the scenario with at most $(1/3-1/k)$-augmentation
+\maciek{Rewrite: cluster types should be cluster signatures}
+In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
+Before bounding the competitive ratio of \TAlg, we introduce the lemma that estimates the cost of a single repartition of \TAlg.
 
-  		 \item Costly cascade example (it is easy to present $k^2$ and $k\cdot \sqrt{k}$ (with different approach), and then we can combine these to obtain $k^3$)
+\begin{lemma}
+  \label{lem:1req}
+  During a single repartition, \TAlg exchanges at most $2$ pairs of nodes.
+\end{lemma}
 
-  		 \item $k$-competitive algorithm for the ring communication pattern
+\begin{proof}
+    Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
+  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
 
-       \item lower and upper bound $k+1$ for two clusters, perfect partition and the optimal configuration reachable within one swap
+  %We distinguish between three types of clusters: $C_1, C_2, C_3$,
+  %which we define as follows.
+  %A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+  %A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
+  %Finally, a cluster of type $C_3$  contains one component of size $3$.
 
-  \end{enumerate}
+    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
+    Note that $U\neq V$,
+	 as otherwise, the request would not trigger a~repartitioning.
+	 We consider cases based on the types of clusters $U$ and $V$.
+    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
+    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
+    For two clusters, any configuration can be reached within two swaps, due to the fact that clusters are indistinguishable.
 
-\section{Maciek's sketches}
+    Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
+    Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
+    Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+    Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
+In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
+\end{proof}
+
+\subsection{Characterization of the reconfiguration graph}
 
@@
 First, we bound the cycle length in one of optimal solutions.
+
+\begin{figure}[H]
+  \centering
+    \includegraphics[width=0.3\textwidth]{figs/substitute}
+    \caption{Illustration to bounding the path length.}
+    \label{fig:nptree-construction}
+  \end{figure}
 
@@
 
-\begin{theorem}
-  Algorithm DET2 is $O(\ell \cdot k\cdot k^k)$-competitive.
-\end{theorem}
+Corollary: Algorithm DET2 is $O(\ell \cdot k\cdot k^k)$-competitive.
 
-\begin{proof}
-  Use Theorem~\ref{th:cascade}. Then, linearize the inner-component communication using the technique from SIDMA.
-\end{proof}
+\section{Improved algorithm for $k=2$}
 
@@
 
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{references}
+
+\pagebreak
+\appendix
+
+\section{Omitted results}
+
+\begin{enumerate}
+  		 \item The lower bound of $\Omega(\ell)$ for the scenario with at most $(1/3-1/k)$-augmentation
+
+  		 \item Costly cascade example (it is easy to present $k^2$ and $k\cdot \sqrt{k}$ (with different approach), and then we can combine these to obtain $k^3$)
+
+  		 \item $k$-competitive algorithm for the ring communication pattern
+
+       \item lower and upper bound $k+1$ for two clusters, perfect partition and the optimal configuration reachable within one swap
+
+  \end{enumerate}
+
 \end{document}

### 2020-04-22 paper/disc20clustering.tex
Paper restructure
Comments

@@
 \section{Introduction}
+
+\maciek{TODO: discuss the title. We should not claim tight bounds with an exponential gap. Reviewers will point this out, as this is an easy thing to criticize in the review.}
 
@@
 
-\maciek{REALLY bad name. Both the learning problem and the general one is online.}
+\maciek{Not precise name? Both the learning problem and the general one is online.}
+
+\maciek{Start with referencing the table of results, saying that some results are known already.
+And really emphasize here that we now have an improved lower bound from the previous section, as that model was a subproblem.}
 
@@
 \maciek{Rewrite above paragraph to say that our algorithm is essentially DET from DISC16, with concretised reconfiguration strategy (i.e., the cheapest)}
+
+\maciek{Better name than DET?}
 
@@
 
+\maciek{I'll write it some day. Maciek.}
+
+\subsection{Costly reconfiguration example}
+
+\maciek{A linear path in reconfiguration graph.}
+
 \subsection{Cascade bounds and an improved algorithm for general k, l}
@@
 \section{Improved algorithm for $k=2$}
+
+\maciek{Intro: compare to previous algorithm. Note strict competitive ratio vs large additive term.}
 
@@
 
+\section{Note on distributed algorithms}
+
+\maciek{How to distribute our algorithms: PPL, k=2 and DET}
+
 \bibliographystyle{ACM-Reference-Format}

### 2020-04-22 paper/disc20clustering.tex
On cascade

@@
 
+\maciek{Emphasize the difference to DET: "Modification is the following: when repartition is necessary, move to \emph{the closest} perfect partition of components."}
+
   \begin{figure}[H]
@@
 
+Describe the graph. Edges and labels.
+
 \subsection{Costly reconfiguration example}
@@
 
-For now, assume $\alpha=1$ (not to deal with inter-component communication).
-Consider an algorithm DET2, a modified version of DET from \cite{repartition-disc}.
-Modification is the following: when repartition is necessary, move to \emph{the closest} perfect partition of components.
-
 Now we bound the cost of a single repartition of DET2 independent of $\ell$.
-First, we bound the cycle length in one of optimal solutions.
+First, we bound the length of each path in one of optimal solutions.
 
@@
 \begin{theorem}
-  \label{th:cascade-cycles}
+  \label{th:cascade-paths}
   Consider a single repartition of DET2 caused by node merge.
@@
 \end{theorem}
-Describe the graph. Edges and labels.
+
+\begin{proof}
 
@@
 
-\begin{proof}
 \end{proof}
@@
 \begin{proof}
-  Each node has degree at most $k$. Combine with Theorem \ref{th:cascade-cycles}.
+  Each node has degree at most $k$. Combine with Theorem \ref{th:cascade-paths}.
 
-Important technical details:
-\begin{itemize}
-  \item There are no cycles attached to another cycles by just one vertex (by equivalent signature skipping).
-This is important, as otherwise we just have a bound on cycles, but the graph contains a chain of cycles.
-  Another formulation (Stefan's): each edge is a part of a cycle that goes through the "origin" part of the graph.
-\end{itemize}
 \end{proof}

### 2020-04-22 paper/disc20clustering.tex
On cascade
Generic algorithm

@@
  The \emph{distance} of a configuration $P = C_1, \dots, C_{\ell}$ from the initial configuration \maciek{, denoted $\dist(P)$} is the number of nodes in $P$ that do not reside in their initial cluster.
- \maciek{It is OK to use single-argument $\dist$ (two-arg is a must in a definition, and it is already there).}
- \maciek{We should denote distance $\Delta(P)$, and the current $\Delta$ should be R (like radius).}
+ \maciek{We might denote distance $\Delta(P)$, and the current $\Delta$ might be R (like radius).}
     That is,
@@
     $|M| \leq 2\cdot\OPT$.
-    \maciek{the comma between $\dist$ is confusing. BOTH are bounded?}
+    \maciek{the comma between $\dist$ is confusing. BOTH are bounded? If so, expand the sentence and write both individual bounds.}
 \end{proof}
@@
 
-\maciek{Better name than DET?}
+\maciek{Better name than DET? Currently it is \TAlg, hence it is wrong}
 
@@
   \TAlg is $O(\ell)$-competitive.
+  \maciek{Let Reconf(k,l) be a bound on the cost of the reconfiguration. Then \TAlg is $O(Reconf(k,l)\cdot k \cdot \ell)$-competitive.}
 \end{theorem}
@@
   As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
-  By performing a single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
+  By performing a single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$
+  \maciek{This changes in the general case, and introduces the factor $~1/k$ in opt's cost, increasing the competitive ratio by $O(k)$}.
   Thus, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{2\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 2 - 2 \alpha \cdot \ell$.
@@
 
-  \maciek{Big techincal TODO. Formalize the sketched proof for the last phase.}
   It remains to consider the last, unfinished phase.
@@
   If the input finished before the first $\alpha$ external requests, then \TAlg is optimal.
+  \maciek{Missing second case: then we just charge opt alpha for this one edge and this is sufficient, as the cost of alg only decreases from the full phase}
   Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
@@
 
-\section{Bounding the cost of a single reconfiguration}
+\section{Bounding the cost of a single reconfiguration of components}
+
+\maciek{Overview of the section, intro. Defer introducing the graph notation, the warm-up does not need it}
 
@@
 \begin{proof}
-  Each node has degree at most $k$. Combine with Theorem \ref{th:cascade-paths}.
+  Each node has degree at most $k$.
+  Every subset has flow conservation property, hence each path traverses one of the nodes of graph origin/core. (no subsets with just incoming edges to the origin part)
+  Combine with Theorem \ref{th:cascade-paths}.
 

### 2020-04-22 paper/disc20clustering.tex
Generic algorithm rewrite

@@
   \TAlg is $O(\ell)$-competitive.
-  \maciek{Let Reconf(k,l) be a bound on the cost of the reconfiguration. Then \TAlg is $O(Reconf(k,l)\cdot k \cdot \ell)$-competitive.}
+  \maciek{Proposed new formulation (current version is all about k=3 still): Let Reconf(k,l) be a bound on the cost of a single reconfiguration.
+Then \TAlg is $O(Reconf(k,l)\cdot k \cdot \ell)$-competitive.}
 \end{theorem}
@@
 
+Corollary: DET2 is $O(\ell)$-competitive for $k=3$.
+
 \subsection{Characterization of the reconfiguration graph}
@@
 
-\maciek{Intro: compare to previous algorithm. Note strict competitive ratio vs large additive term.}
+\maciek{Intro: compare to previous 7-competitive algorithm. Note strict competitive ratio vs large additive term. Also, emphasize that in this section the algorithm is different than DET2}
 

### 2020-04-22 paper/disc20clustering.tex
Comments

@@
 
-\maciek{TODO: discuss the title. We should not claim tight bounds with an exponential gap. Reviewers will point this out, as this is an easy thing to criticize in the review.}
-
 \noindent \textbf{Motivation and model.}
@@
 
-\maciek{The text before the proof should have two parts: a description of adversary components and then general intro to the proof.}
+\maciek{TODO: rename everywhere: "adversary's components", i.e., all components in lower bounds should be called "ground sets"}
+
+\maciek{The text before the proof should have two parts: a description of ground sets and then general intro to the proof.}
 
@@
 discuss an upper bound.
-
-\maciek{TODO: rename everywhere: "adversary's components", i.e., all components in lower bounds should be called "ground sets"}
 
@@
 
-\maciek{Here: we abstract from the value of $\alpha$, but works for any alpha.}
+\maciek{Note that the lower bound works for any value of $\alpha$, and we just use the number of migrations}
 
@@
 
-  \maciek{This paragraph should be before the proof, merged with section about adversary components.}
+  \maciek{This paragraph should be before the proof, merged with section about ground sets (i.e., adversary components.)}
   We join some nodes into components, and produce the input sequence by observing actions of \ALG{} with respect to nodes of these components.
@@
 
-\maciek{Start with referencing the table of results, saying that some results are known already.
-And really emphasize here that we now have an improved lower bound from the previous section, as that model was a subproblem.}
+\maciek{Start with referencing the table of results, saying that some results are known already in the literature.
+And emphasize here that we now have an improved lower bound from the previous section, as that model was a subproblem.}
 
@@
 
-Corollary: Algorithm DET2 is $O(\ell \cdot k\cdot k^k)$-competitive.
+Corollary: Algorithm DET2 is $O(\ell \cdot k \cdot \min \{ \ell, k^k \})$-competitive.
 
@@
        \item lower and upper bound $k+1$ for two clusters, perfect partition and the optimal configuration reachable within one swap
+       \item 
+       The bound $(k \cdot \ell - k - \ell) / 2$ can be easily improved to $(k \cdot \ell - k - 1) / 2$ by playing a simple caching game in the first cluster, instead of just giving away the $k-1$ "blocker" (the caching game would build it gradually)
 

### 2020-04-22 paper/disc20clustering.tex
Migration graph characterization

@@
 
-\subsection{Characterization of the reconfiguration graph}
+\subsection{Characterization of the migration graph}
 
-\maciek{I'll write it some day. Maciek.}
+Consider a component merge action performed by \TAlg that triggers a reconfiguration from a configuration $C_I$ to a configuration $C_F$.
+To execute the reconfiguration, the nodes of two non-collocated components migrate to a common cluster.
+Each cluster contains exactly $k$ nodes, thus other nodes migrate to take the place of nodes of merged components.
+Their place must be occupied, too, and this way a single merge action may trigger migrations in clusters not directly involved in a merge action (for an example, see Section~\ref{ssec:example}). 
+Recall that both $C_I$ and $C_F$ are component-respecting configurations, hence a migration of a single node entails migrations of all nodes of its cluster.
 
-Describe the graph. Edges and labels.
+In Section~\ref{ssec:cascade}, we upper bound the cost incurred by a reconfiguration that follows a single component merge.
+To this end, we introduce a \emph{migration graph} that models the reconfiguration.
+In the following, we characterize the structure of the migration graph.
 
-\subsection{Costly reconfiguration example}
+\noindent
+\textbf{Vertices.}
+The vertices of the migration graph are all $\ell$ clusters of the instance.
+We distinguish the \emph{core vertices} that correspond to clusters directly involved in the merge operation: the clusters containing the to-be-merged components in $C_I$ and the cluster containing the merged component in $C_F$.
+There are at most $3$ core vertices in each migration graph (at most two components participate in the merge, and these may migrate to a third cluster).
+
+\noindent
+\textbf{Edges and their labels.}
+
+Cost of cluster reconfiguration is the number of exchanged nodes, and from the standpoind of a migration graph, it is the sum of labels on edges of the graph.
+It is bounded by $2k$ times the size of the connected component.
+
+Core of the graph and the component merge.
+
+Multiple migration graphs for a component merge exist, and multiple of them might have the optimal cost.
+In Section \ref{ssec:cascade}, we show that among all migration graphs for given component merge, there exists an optimal graph with the bounded path length.
+
+\subsection{Costly cluster reconfiguration example}
+\label{ssec:example}
 
@@
 
-\subsection{Cascade bounds and an improved algorithm for general k, l}
+\subsection{Bounds on the size of a cluster reconfiguration graph}
+\label{ssec:cascade}
 
@@
 \begin{proof}
+  Assume that there exists a cluster reconfiguration
 

### 2020-04-22 paper/disc20clustering.tex
k=2 complete proof

@@
 
-\begin{theorem}
-	The algorithm is 4-competitive.
+\begin{theorem} \label{thm:k=2}
+	The algorithm is (strictly) 4-competitive.
 \end{theorem}
-\begin{proof}
+
+We charge both OPT and Alg whenever Alg collocates a pair (i.e., on a collocation event).
+%More precisely,we charge them only the cost of the collocation, per such event.
+%Hence,we charge the cost of  this collocation to OPT for the first time.
+Alg collocates a pair always with a swap (i.e., two simultaneous migrations),
+which  costs $2\alpha$.
+%We charge OPT only for requests up to the one that triggers a collocation (i.e., not for later requests on the same pair).
+OPT may save some costs by collocating multiple pairs at once (before they incur too much cost),
+paying the price of only one migration per pair.
+This observation suggests that we can't charge OPT more than $\alpha$ per collocation,
+even when OPT actually pays more that that.
+E.g.,
+OPT can collocate two costly pairs with a single swap if it has both pairs on the cut of same two clusters.
+Later, Alg collocates those pairs with two swaps (see Figure \ref{fig:TBD}).
+Hence,
+whenever OPT collocates a pair,
+we charge OPT only the cost of moving a single node to the cluster of the other,
+i.e., $\alpha$ (in contrast to $2\alpha$ for Alg).
+
+\begin{proof}[Proof of Theorem \ref{thm:k=2}]
 	\mahmoud{still proof sketch}
-We charge both OPT and Alg on each request that triggers a collocation.
 	Let $C^{t}_{\set{u,v}}$ denote the  value of $C_{\set{u,v}}$ immediately after arrival of $\sigma_t$.
@@
 	which means $ C^t_{\set{u,v}} = \lambda$ immediately before Alg resets the counter.
-Let $t''<t$ be the latest time (i.e.~request) when Alg collocates $\set{u,v}$ before $t$.
-Let $t', t'' < t' < t$ be the  time when Alg separates $\set{u,v}$
-in order to collocate some other pair sharing a node,
-e.g., the pair $\set{u,v'}, v' \neq v$.
+	Let $\sigma_{t''}, t''<t$ be the last request (before $\sigma_{t}$)
+	on which Alg collocates $\set{u,v}$.
+	Let $t'$ s.t.~$t'' < t' < t$ be the  time when Alg separates $\set{u,v}$
+	in order to collocate some other pair that shares a node with $\set{u,v}$,
+	denoted by $\set{u,v'}, v' \neq v$.
 	%e.g., the request $\sigma_{t'} = \set{u,v'}$.
@@
 	$| \sigma_{\set{u,v}}[t',t] | = \lambda$.
-The overall cost of requests that increases the counter,
+	The overall cost of requests that increase the counter,
 	i.e.~requests in $\sigma_{\set{u,v}}[t',t]$,
-and the moving cost incurred by Alg,
- is $\lambda + 2\alpha$.
+	and the moving cost,
+	incurred by Alg is $\lambda + 2\alpha$.
 	%That is, the cost of all requests in $\sigma_{\set{u,v}}[t',t]$.
-We distinguish three cases for the cost incurred by OPT with regards to $\set{u,v}$.
-i)
-OPT has $\set{u,v}$ separated through all requests in $\sigma_{\set{u,v}}[t',t]$
+	We distinguish three cases for the cost incurred by OPT due to $\set{u,v}$.
+	Case i)
+	OPT has $\set{u,v}$ separated (in OPT's configuration) during $[t',t]$
 	and therefore it pays $\lambda$.
-ii)  
+	We charge OPT $\lambda/2 = \alpha$.
+	The remaining cost will be used in a latter case.
+	Case ii)  
 	OPT collocates $\set{u,v}$ some time during $[t'',t]$.
-Recall that whenever Alg collocates a pair,
-we charge OPT only for requests up to that point (i.e., not for later requests).
-Hence,
-we charge the cost of  this collocation to OPT for the first time.
-In particular,
-we charge OPT the cost of moving one node to the cluster of the other,
-e.g.~$u$ to $v$, w.l.o.g.
-Therefore in this case OPT pays $\alpha$.
-iii)
+	Any cost that we charge to OPT upon $\sigma_{t''}$,
+	accounts for OPT's actions before $t''$.
+	Therefore in this case,
+	we charge OPT the (unaccounted) cost $\alpha$.
+	Case iii)
 	OPT has $\set{u,v}$ collocated before $t''$ and keeps them collocated during $[t'',t]$.
-Obviously in this case OPT doesn't pays any cost for requests in $\sigma_{\set{u,v}}[t'',t]$.
+	Obviously in this case OPT doesn't pays any cost for requests in $\sigma_{\set{u,v}}[t'',t]$ and
+	we can't charge it directly.
 	However,
-OPT must pay $\lambda$ for requests to $\set{u,v'}$.
-By induction,
-at time $t'$,
-we charge OPT $\alpha$ (out of $\lambda=2\alpha$) for this pair.
-Therefore,
-we can charge OPT another $\alpha$ to account for requests $\sigma_{\set{u,v}}[t'',t]$.
+	OPT must pay $\lambda$ for all requests in $\sigma_{\set{u,v'}}[t'',t']$,
+	and we charge it partially to OPT at $t'$ (as in Case i).
+	Since we always charge OPT no more than $\alpha$,
+	we can charge OPT another $\alpha$,
+	the remaining cost of requests $\sigma_{\set{u,v'}}[t'',t']$.
+	Intuitively,
+	at $t$,
+	we charge OPT $\alpha$ for separating $\set{u,v}$ at $t'$.
 	It can be easily verified that any deviation from these cases only increases the cost to OPT.
 	Thus,
-OPT pays at least $\alpha$ for each collocation by Alg
+	OPT pays at least $\alpha$ for each collocation performed by Alg,
 	which gives the competitive ratio 
@@
 
+	It remains to charge both Alg and OPT for requests that do not lead to collocations when the sequence ends.
+	Consider  pairs $\set{u,v}$ and $\set{u,v'}$ and
+	times $t'' < t' < t = t_m$ as defined previously.
+	Assume $\sigma_{\set{u,v}}[t'', t_m] \leq \lambda - 1$.
+	This means Alg has the pair separated at $t_m$
+	and paid the (communication) cost $r_{\set{u,v}} := | \sigma_{\set{u,v}}[t'', t_m]|$.
+	If OPT has the pair separated during $[t'',t_m]$ then it pays $r_{\set{u,v}}$.
+	If OPT collocates the pair some time during $[t'', t]$ then analogous to Case (ii),
+	we charge it $\alpha$.
+	Otherwise,
+	if OPT collocates the pair before $t''$,
+	then we charge it $\alpha$ analogously to Case (iii).
+	Hence, 
+	$Cost(OPT)_{\set{u,v}} / Cost(Alg)_{\set{u,v}} = r_{\set{u,v}}/ \alpha \leq (\lambda-1) / \alpha < 2$.
+	The claim follows from here.
 \end{proof}

### 2020-04-22 paper/disc20clustering.tex
Migration graph characterization

--- 2020-04-22 paper/disc20clustering.tex
@@
 
-\maciek{Rewrite above paragraph to say that our algorithm is essentially DET from DISC16, with concretised reconfiguration strategy (i.e., the cheapest)}
+\maciek{Rewrite above paragraph to say that our algorithm is essentially DET from DISC16, with concretised reconfiguration strategy (i.e., an arbitrary one among the minimum cost reconfigurations)}
 
@@
 \noindent
-\textbf{Vertices.}
+\textbf{Vertices of the migration graph and the core vertices.}
 The vertices of the migration graph are all $\ell$ clusters of the instance.
@@
 \textbf{Edges and their labels.}
+The edges of the migration graph denote the migration of nodes from $C_I$ to $C_F$.
+Each edge is labeled with a number $m$ that denotes the number of migrated nodes between clusters.
+Each edge is directed from the cluster that contained the nodes to the cluster they migrated to.
+Both edges might exists between a pair of nodes, as whole components of nodes migrate, and the back-and-forth exchange may be needed for the configuration $C_F$ to be component-respecting.
 
-Cost of cluster reconfiguration is the number of exchanged nodes, and from the standpoind of a migration graph, it is the sum of labels on edges of the graph.
-It is bounded by $2k$ times the size of the connected component.
+\noindent
+\textbf{Vertex degree.}
+At most $k$ nodes may migrate from a cluster and at most $k$ nodes may migrate into a cluster.
+Thus, the sum of the labels for both ingoing and outgoing edges of each vertex is at most $k$.
+This implies that the number of ingoing and outgoing edges is also bounded by $k$.
 
-Core of the graph and the component merge.
+\noindent
+\textbf{Flow preservation.}
+In any feasible configuration, each cluster contains exactly $k$ nodes.
+Thus, for any vertex, the sum of labels of ingoing edges must equal the sum of labels of outgoing edges.
 
-Multiple migration graphs for a component merge exist, and multiple of them might have the optimal cost.
+\noindent
+\textbf{Migration graph and the cost of reconfiguration.}
+The cost of cluster reconfiguration equals the number of exchanged nodes multiplied by $\alpha$.
+From the standpoint of a migration graph, this corresponds to the sum of labels on edges of the graph.
+A trivial upper bound on the cost is the size of the connected component in the migration graph multiplied by $2\alpha k$.
+
+\medskip
+
+Note that for a single component merge, multiple migration graphs may exist, and multiple of them might have the optimal cost.
 In Section \ref{ssec:cascade}, we show that among all migration graphs for given component merge, there exists an optimal graph with the bounded path length.
+Combined with other properties, this allows to bound the size of the migration graph, and consequently bound the cost of each repartition, that results in a bound of the competitive ratio for \TAlg.
 

### 2020-04-22 paper/disc20clustering.tex
wip
Migration graph path bound

@@
 
-\maciek{A linear path in reconfiguration graph.}
+\begin{figure}[H]
+  \centering
+    \includegraphics[width=0.3\textwidth]{figs/substitute}
+    \caption{A costly cascade example.
+On the left: the configuration prior to the merge and the merge of two components (denoted solid).
+On the right: corresponding migration graph for the optimal (unique) reconfiguration.}
+  \end{figure}
 
-\subsection{Bounds on the size of a cluster reconfiguration graph}
+In this section we present an example of a merge action that results in a $k/3$(?) length path in an optimal migration graph.
+
+\subsection{Bounds on the size of a migration graph}
 \label{ssec:cascade}
 
-Now we bound the cost of a single repartition of DET2 independent of $\ell$.
-First, we bound the length of each path in one of optimal solutions.
+In \cite{repartition-disc}, authors trivially bounded the reconfiguration cost for each merge action by $k \cdot \ell$.
+This roughly corresponds to migrating every node in every cluster.
+Now, we bound the cost of a single repartition of \TAlg as a function of $k$.
+Note that the bound of $k \cdot \ell$ is still valid, and the resulting bound is the minimum of these two.
+
+We begin, by bounding the length of each path in one of optimal solutions.
 

### 2020-04-22 paper/disc20clustering.tex
Migration graph path bound

@@
   \label{th:cascade-paths}
-  Consider a single repartition of DET2 caused by node merge.
-  There exists a sequence of node migrations that realizes this repartition, where each node-disjoint cycle has length at most $k$.
+  Consider a single merge action of \TAlg.
+  There exists an optimal cost reconfiguration with the migration graph, where each path have length at most $k$.
 \end{theorem}
@@
 \begin{proof}
-  Assume that there exists a cluster reconfiguration
+  Assume the contrary, i.e., for each optimal cluster reconfiguration, its migration graph contains a path at least $k+1$.
 
-Proof by contradiction.
-We call a sequence of node migrations that realize the repartition \emph{a solution to repartition}.
-Assume that each cost-optimal solution to repartition contains a cycle of length at least $k+1$.
-
-Fix a solution to repartition that has the shortest length of the longest cycle, and has the least edge-disjoint cycles of such length.
+Fix a reconfiguration and its migration graph that has the shortest length $L$ of the longest path, and has the least number of paths of length $L$.
+Let $p = \lbrace v_1, v_2, \ldots, v_L \rbrace$ be any path of length $L$.
 

### 2020-04-22 paper/disc20clustering.tex
Migration graph CYCLE bound

@@
 Thus, for any vertex, the sum of labels of ingoing edges must equal the sum of labels of outgoing edges.
+\maciek{Can we say that this implies that the graph is a circulation?}
 
@@
 Note that for a single component merge, multiple migration graphs may exist, and multiple of them might have the optimal cost.
-In Section \ref{ssec:cascade}, we show that among all migration graphs for given component merge, there exists an optimal graph with the bounded path length.
+In Section \ref{ssec:cascade}, we show that among all migration graphs for given component merge, there exists an optimal graph with the bounded cycle length.
 Combined with other properties, this allows to bound the size of the migration graph, and consequently bound the cost of each repartition, that results in a bound of the competitive ratio for \TAlg.
@@
 
-In this section we present an example of a merge action that results in a $k/3$(?) length path in an optimal migration graph.
+In this section we present an example of a merge action that results in a $k/3$(?) length cycle in an optimal migration graph.
 
@@
 
-We begin, by bounding the length of each path in one of optimal solutions.
+We begin, by bounding the length of each cycle in one of optimal solutions.
 
@@
     \includegraphics[width=0.3\textwidth]{figs/substitute}
-    \caption{Illustration to bounding the path length.}
-    \label{fig:nptree-construction}
+    \caption{Illustration to bounding the cycle length.}
+    \label{fig:cascade-illustration}
   \end{figure}
@@
 \begin{theorem}
-  \label{th:cascade-paths}
+  \label{th:cascade-cycles}
   Consider a single merge action of \TAlg.
-  There exists an optimal cost reconfiguration with the migration graph, where each path have length at most $k$.
+  There exists an optimal cost reconfiguration with the migration graph, where each cycle have length at most $k$.
 \end{theorem}
@@
 \begin{proof}
-  Assume the contrary, i.e., for each optimal cluster reconfiguration, its migration graph contains a path at least $k+1$.
+  Assume the contrary, i.e., for each optimal cluster reconfiguration, its migration graph contains a cycle at least $k+1$.
 
-Fix a reconfiguration and its migration graph that has the shortest length $L$ of the longest path, and has the least number of paths of length $L$.
-Let $p = \lbrace v_1, v_2, \ldots, v_L \rbrace$ be any path of length $L$.
+Fix a reconfiguration and its migration graph $G$ that has the shortest length $L$ of the longest cycle, and has the least number of cycles of length $L$.
+Let $p = \lbrace v_1, v_2, \ldots, v_L \rbrace$ be any cycle of length $L$.
+As edge labels are from $\{ 1, 2, \ldots, k \}$, there exists a pair of edges $e, f$ with equal labels.
+
+We show that in such case, we can modify the solution without affecting its cost and feasibility to decrease the number of cycles of length $L$ by $1$.
+This would contradict the assumption that we consider a solution with the minimal number of cycles of length $L$.
+
+We construct a new migration graph $G'$ of equal cost by swapping the (destination) endpoints of $e$ and $f$ (cf. Figure~\ref{fig:cascade-illustration})
+Formally, let $\lbrace a, b \rbrace = e$ and $\lbrace c, d \rbrace = f$.
+We remove $e$ and $f$ and add $e' = \lbrace a, d \rbrace$ and $f' = \lbrace c, d \rbrace$.
+This way we produced a feasible solution.
+
+It remains to show that while decreasing the number of cycles of length $L$, we did not created any new cycles of length at least $L$.
+Assume that adding $e'$ created a new cycle of length at least $L$.
+However, $a$ and $d$ were already connected in $G$ by a directed path of length at least $2$, and a cycle strictly longer than $L$ existed in $G$.
+
+Assume that adding $f'$ created a new cycle $s^*$ of length $L' \geq L$.
+Contrary to the previous case, the directed cycle from $e$ to $b$ might not exist in $G$.
+However, this would mean that $b$ and $e$ were connected by a path of length $L'-1$.
+This means that we might have followed this alternative path from $b$ to $e$ to obtain a cycle strictly longer than $L$ (cf. Figure~\ref{fig:cascade-illustration}).
 
@@
   Each node has degree at most $k$.
-  Every subset has flow conservation property, hence each path traverses one of the nodes of graph origin/core. (no subsets with just incoming edges to the origin part)
-  Combine with Theorem \ref{th:cascade-paths}.
+  Every subset has flow conservation property, hence each cycle traverses one of the nodes of graph origin/core. (no subsets with just incoming edges to the origin part)
+  Combine with Theorem \ref{th:cascade-cycles}.
 

### 2020-04-22 paper/disc20clustering.tex
Comments

@@
 
+\maciek{The remainder of the paper used "configuration C" to denote the partition P. Easiest to adjust here.}
+
 We assume \OPT begins with the initial configuration
@@
 \end{property}
+\maciek{The property is explained in the previous paragraph. Here, reader thinks that this is obvious and wonders why. Move closer to the exlaination text.}
 
@@
     $M = M^+ \cup M^-$.
-    Since $|M^-|$ nodes are not in their original cluster before the repartitioning (i.e., in $P_{t-1}$),
+    Since $|M^-|$ nodes are not in their original cluster before the repartitioning \maciek{This is non-obvious to me. Why does it hold?} (i.e., in $P_{t-1}$),
     the distance before the repartitioning is $\dist(P_{t-1},P_I) \geq | M^-|$.

### 2020-04-22 paper/disc20clustering.tex
Comments

@@
     $M = M^+ \cup M^-$.
-    Since $|M^-|$ nodes are not in their original cluster before the repartitioning \maciek{This is non-obvious to me. Why does it hold?} (i.e., in $P_{t-1}$),
+    Since \maciek{at least?} $|M^-|$ nodes are not in their original cluster before the repartitioning (i.e., in $P_{t-1}$),
     the distance before the repartitioning is $\dist(P_{t-1},P_I) \geq | M^-|$.
     Analogously,
-     the distance afterwards is $\dist(P_{t},P_I) \geq | M^+|$.
+     the distance afterwards is \maciek{at least?} $\dist(P_{t},P_I) \geq | M^+|$.
     Thus,

### 2020-04-22 paper/disc20clustering.tex
Comments

@@
     Analogously,
-     the distance afterwards is \maciek{at least?} $\dist(P_{t},P_I) \geq | M^+|$.
+     the distance afterwards is $\dist(P_{t},P_I) \geq | M^+|$.
     Thus,
@@
   %That is, the cost of all requests in $\sigma_{\set{u,v}}[t',t]$.
+
 	We distinguish three cases for the cost incurred by OPT due to $\set{u,v}$.
-	Case i)
+
+  \begin{enumerate}
+  \item Case i)
 	OPT has $\set{u,v}$ separated (in OPT's configuration) during $[t',t]$
@@
 	The remaining cost will be used in a latter case.
-	Case ii)  
+
+  \item Case ii)  
 	OPT collocates $\set{u,v}$ some time during $[t'',t]$.
@@
 	we charge OPT the (unaccounted) cost $\alpha$.
-	Case iii)
+
+  \item Case iii)
 	OPT has $\set{u,v}$ collocated before $t''$ and keeps them collocated during $[t'',t]$.
@@
   we charge OPT $\alpha$ for separating $\set{u,v}$ at $t'$.
+  \end{enumerate}
+
 	It can be easily verified that any deviation from these cases only increases the cost to OPT.
@@
 \end{proof}
+
+\maciek{Missing the case of the first-time collocation. What if t' does not exist? What if t'' does not exist?}
+\maciek{Define $C^t$ to use once}
+\maciek{"It can be easily verified"... True, but avoid such formulation. Something along these lines: if opt migrates earlier than t', then additionally pays for X, if later than Y.}
+\maciek{"Thus, OPT pays ... which gives the competitive ratio" It is too early to give the competitive ratio if in the next paragraph we consider yet another case}
 

### 2020-04-22 paper/disc20clustering.tex
Comments

@@
 
+\maciek{$\sigma_t$ undefined (I know what it is, though).}
 \maciek{Missing the case of the first-time collocation. What if t' does not exist? What if t'' does not exist?}
@@
 \maciek{"It can be easily verified"... True, but avoid such formulation. Something along these lines: if opt migrates earlier than t', then additionally pays for X, if later than Y.}
-\maciek{"Thus, OPT pays ... which gives the competitive ratio" It is too early to give the competitive ratio if in the next paragraph we consider yet another case}
+\maciek{"Thus, OPT pays ... which gives the competitive ratio" It is too early to give the competitive ratio if in the next paragraph we consider yet another case. }
+\maciek{Reason that the cost sharing between cases C1 and C3 is possible. Unclear to me, why every time C3 happens there must be a corresponding C1.}
 

### 2020-04-24 paper/disc20clustering.tex
Comments
did some of the comments on k=2 proof

@@
 E.g.,
-OPT can collocate two costly pairs with a single swap if it has both pairs on the cut of same two clusters.
+OPT can collocate two costly pairs with a single swap if
+it has both pairs on the cut of  the same two clusters.
 Later, Alg collocates those pairs with two swaps (see Figure \ref{fig:TBD}).
@@
 	\mahmoud{still proof sketch}
-	Let $C^{t}_{\set{u,v}}$ denote the  value of $C_{\set{u,v}}$ immediately after arrival of $\sigma_t$.
+	Let $C^{t}_{\set{u,v}}$ denote the  value of $C_{\set{u,v}}$
+	immediately after arrival of the $t$th request,
+	$\sigma_t$.
 	Assume on request $\sigma_{t}=\set{u,v}$, 
+	i.e., at (time) $t \geq 1$,
 	Alg collocates the pair,
@@
 	Let $\sigma_{t''}, t''<t$ be the last request (before $\sigma_{t}$)
-	on which Alg collocates $\set{u,v}$.
+	upon which Alg collocates $\set{u,v}$.
 	Let $t'$ s.t.~$t'' < t' < t$ be the  time when Alg separates $\set{u,v}$
@@
 	denoted by $\set{u,v'}, v' \neq v$.
+	In case the first collocation of $\set{u,v}$ occurs at $t$,
+	if the pair is collocated in the initial configuration then
+	we set $t''=0$,
+	else,
+	we set $ t'' = -\infty, t'=0$.
 	%e.g., the request $\sigma_{t'} = \set{u,v'}$.
@@
 	We charge OPT $\lambda/2 = \alpha$.
-	The remaining cost will be used in a latter case.
+%	The remaining cost will be used in a latter case.
 
   \item Case ii)  
-	OPT collocates $\set{u,v}$ some time during $[t'',t]$.
+	OPT collocates $\set{u,v}$ some time during $[t',t]$.
 	Any cost that we charge to OPT upon $\sigma_{t''}$,
@@
   \item Case iii)
-	OPT has $\set{u,v}$ collocated before $t''$ and keeps them collocated during $[t'',t]$.
-	Obviously in this case OPT doesn't pays any cost for requests in $\sigma_{\set{u,v}}[t'',t]$ and
+	$t'' \geq 0$ and OPT has $\set{u,v}$ collocated before $t''$ and keeps them collocated during $[t'',t]$.
+	In this case OPT doesn't pays any cost for requests in $\sigma_{\set{u,v}}[t'',t]$ and
 	we can't charge it directly.
@@
 	OPT must pay $\lambda$ for all requests in $\sigma_{\set{u,v'}}[t'',t']$,
-	and we charge it partially to OPT at $t'$ (as in Case i).
-	Since we always charge OPT no more than $\alpha$,
-	we can charge OPT another $\alpha$,
-	the remaining cost of requests $\sigma_{\set{u,v'}}[t'',t']$.
-	Intuitively,
-	at $t$,
-  we charge OPT $\alpha$ for separating $\set{u,v}$ at $t'$.
+	 i.e.,
+	 Case (i) applies to $\set{u,v'}$  at $t'$ and
+ 	we charge only $\lambda/2$ to OPT (at $t'$).
+	This means,
+	we can charge OPT an additional $\alpha$,
+	i.e.~the remaining cost of requests $\sigma_{\set{u,v'}}[t'',t']$.
+%	Intuitively, at $t$,  we charge $\alpha$ to OPT for separating $\set{u,v}$ (at $t'$) by Alg.
   \end{enumerate}
@@
 	Thus,
-	OPT pays at least $\alpha$ for each collocation performed by Alg,
-	which gives the competitive ratio 
-	$(\lambda + 2\alpha)/ \alpha = 4$.
+	OPT pays at least $\alpha$ for each collocation performed by Alg.
+%	which gives the competitive ratio 
+%	$(\lambda + 2\alpha)/ \alpha = 4$.
 
@@
 	Hence, 
-	$Cost(OPT)_{\set{u,v}} / Cost(Alg)_{\set{u,v}} = r_{\set{u,v}}/ \alpha \leq (\lambda-1) / \alpha < 2$.
-	The claim follows from here.
+	$Cost(OPT)_{\set{u,v}} / Cost(Alg)_{\set{u,v}} =
+	\max \set{ r_{\set{u,v}}/ \alpha \leq (\lambda-1) / \alpha , (\lambda + 2\alpha)/ \alpha} = 4$.
+%	The claim follows from here.
 \end{proof}
 
-\maciek{$\sigma_t$ undefined (I know what it is, though).}
-\maciek{Missing the case of the first-time collocation. What if t' does not exist? What if t'' does not exist?}
-\maciek{Define $C^t$ to use once}
+Missing:
+\maciek{Reason that the cost sharing between cases C1 and C3 is possible. Unclear to me, why every time C3 happens there must be a corresponding C1. \mahmoud{done}}
+\maciek{Missing the case of the first-time collocation. What if t' does not exist? What if t'' does not exist? 
+\mahmoud{added remark for this case: $t''=0$}}
+
+Minor:
 \maciek{"It can be easily verified"... True, but avoid such formulation. Something along these lines: if opt migrates earlier than t', then additionally pays for X, if later than Y.}
-\maciek{"Thus, OPT pays ... which gives the competitive ratio" It is too early to give the competitive ratio if in the next paragraph we consider yet another case. }
-\maciek{Reason that the cost sharing between cases C1 and C3 is possible. Unclear to me, why every time C3 happens there must be a corresponding C1.}
+\maciek{"Thus, OPT pays ... which gives the competitive ratio" It is too early to give the competitive ratio if in the next paragraph we consider yet another case. \mahmoud{done!}}
+
+Tiny:
+\maciek{$\sigma_t$ undefined (I know what it is, though). \mahmoud{added inline def. Do we have a global def for this?}}
+\maciek{Define $C^t$ to use once \mahmoud{already defined! I don't get it.}}
+\maciek{A word "Obviously" is unnecessary, and most people, including me are allergic to it in writing
+\mahmoud{obviously that's true!}}
 

### 2020-04-24 paper/disc20clustering.tex
Comments

@@
 
-\newcommand\mahmoud[1]{\color{green}\textbf{\\ Mahmoud: #1}\\\color{black}}
+\newcommand\mahmoud[1]{\color{orange}\textbf{\\ Mahmoud: #1}\\\color{black}}
 \newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}
@@
 Missing:
-\maciek{Reason that the cost sharing between cases C1 and C3 is possible. Unclear to me, why every time C3 happens there must be a corresponding C1. \mahmoud{done}}
 \maciek{Missing the case of the first-time collocation. What if t' does not exist? What if t'' does not exist? 
 \mahmoud{added remark for this case: $t''=0$}}
+\maciek{But t' might not exist (as it might be the first collocation).}
 
@@
 \maciek{$\sigma_t$ undefined (I know what it is, though). \mahmoud{added inline def. Do we have a global def for this?}}
+\maciek{Not yet, but we still must write the "Problem definition" section (yes, we don't have one!), and we can put it there.}
+
 \maciek{Define $C^t$ to use once \mahmoud{already defined! I don't get it.}}
-\maciek{A word "Obviously" is unnecessary, and most people, including me are allergic to it in writing
-\mahmoud{obviously that's true!}}
+\maciek{discussed}
 

### 2020-04-24 paper/disc20clustering.tex
Comments

@@
 \maciek{Missing the case of the first-time collocation. What if t' does not exist? What if t'' does not exist? 
-\mahmoud{added remark for this case: $t''=0$}}
-\maciek{But t' might not exist (as it might be the first collocation).}
+\mahmoud{added remark for this case: $t''=0$
+\maciek{But t' might not exist (as it might be the first collocation).}}}
 
@@
 \maciek{"It can be easily verified"... True, but avoid such formulation. Something along these lines: if opt migrates earlier than t', then additionally pays for X, if later than Y.}
-\maciek{"Thus, OPT pays ... which gives the competitive ratio" It is too early to give the competitive ratio if in the next paragraph we consider yet another case. \mahmoud{done!}}
 
 Tiny:
-\maciek{$\sigma_t$ undefined (I know what it is, though). \mahmoud{added inline def. Do we have a global def for this?}}
-\maciek{Not yet, but we still must write the "Problem definition" section (yes, we don't have one!), and we can put it there.}
-
-\maciek{Define $C^t$ to use once \mahmoud{already defined! I don't get it.}}
-\maciek{discussed}
+\maciek{$\sigma_t$ undefined (I know what it is, though). \mahmoud{added inline def. Do we have a global def for this?}
+\maciek{Not yet, but we still must write the "Problem definition" section (yes, we don't have one!), and we can put it there.}}
+
+\maciek{Define $C^t$ to use once \mahmoud{already defined! I don't get it.
+\maciek{discussed}}}
 

### 2020-04-29 paper/disc20clustering.tex
Title suggestion
backup k=2 draft

@@
 
+\maciek{Title discussion: "Online Balanced Graph Partitioning: Understanding the role of clusters". Reason: all our results most notably pertain to the number of clusters.}
+
 \noindent \textbf{Motivation and model.}
@@
 and \emph{external} requests, respectively.
+\mahmoud{"Before serving a request": does this mean that only after a request arrives and before serving it we migrate?
+In our algorithms we migrate after serving $\lambda$ requests.
+Do we have discrepancy with the model?}
 Before serving a request,
@@
 maintain a counter $C_{\set{u,v}}$ for each pair of nodes $\set{u,v}$ and increment it on every remote request between $u$ and $v$.
-Once $C_{\set{u,v}} = \lambda:=2\alpha$,
+Once $C_{\set{u,v}} = \lambda$,
 reset the counter and collocate the pair arbitrarily in one of the two clusters where $u$ or $v$ resides.
@@
 \begin{theorem} \label{thm:k=2}
-	The algorithm is (strictly) 4-competitive.
+	The algorithm is (strictly) 6-competitive.
 \end{theorem}
@@
 This observation suggests that we can't charge OPT more than $\alpha$ per collocation,
-even when OPT actually pays more that that.
+even when OPT actually pays more than that.
+E.g.,
+OPT can collocate two costly pairs with a single swap if
+it has both pairs on the cut of  the same two clusters  (see Figure \ref{fig:TBD}).
+Later, Alg collocates those pairs with two swaps.
+Hence,
+whenever OPT collocates a pair,
+we charge OPT only the cost of moving a single node to the cluster of the other,
+i.e., $\alpha$ (in contrast to $2\alpha$ on Alg).
+
+\begin{proof}[Proof of Theorem \ref{thm:k=2}]
+	Assume on request $\sigma_{t}=\set{u,v}$, 
+	i.e., at (time) $t$,
+	Alg collocates the pair $\set{u,v}$.
+	This means,
+	the value of $C_{\set{u,v}}$ at $t$,
+	denoted by $C^t_{\set{u,v}}$, 
+	reaches $\lambda$ immediately before Alg resets the counter.
+	I.e.,
+	$ C^t_{\set{u,v}} = \lambda$.
+%	In case the first collocation of $\set{u,v}$ occurs at $t$,
+%	if the pair is collocated in the initial configuration then
+%	we set $t''=0$,
+%	else,
+%	we set $ t'' = -\infty, t'=0$.
+	We denote the set of all requests to $\set{x,y}$
+	during $(t_1,t_2]$ by $\sigma_{\set{x,y}}(t_1,t_2]$.
+	If $t$ is the first time Alg collocates $\{u,v\}$ then we can charge
+	OPT $\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
+	 or $\alpha$ for collocating the pair and serving them with no additional cost.
+	Therefore in this case ,
+	OPT pays $\min{\{ \lambda,\alpha \}}$.
+	Else,
+	let $t'$, $t' < t$ be the  time when Alg separates $\set{u,v}$
+	in order to collocate some (intersecting) pair
+	$\{x,y\} \neq \{u,v\}, x \in \{u,v\}$, 
+	e.g.,
+	$\{x,y\}=\{u,v'\}$.
+%	I.e., $ C^{t'}_{\set{u,v'}} = \lambda$.
+	Similarly,
+	if $t'$ is not the first time that Alg has $\{u,v'\}$ collocated
+	then let $t''<t'$ be the last time before $t'$ when Alg separates $\set{u,v'}$.
+	Else,
+	$\{u,v'\}$ is not collocated in the initial configuration and let $t''=0$.
+	Then by these definitions,
+	we have
+	$| \sigma_{\set{u,v'}}(t'',t'] | = | \sigma_{\set{u,v}}(t',t] |  =  \lambda$.
+	The overall cost of requests in $\sigma_{\set{u,v}}(t',t]$,
+	and the moving cost,
+	incurred by Alg is
+	$\lambda + 2\alpha$.
+	We charge OPT the cost of serving requests in
+	$\sigma_{\{u,v'\}}[t'',t']$ ($\sigma_{\{u,v\}}(t',t]$ ) 
+	that it pays only within the interval $(t'', t']$ ($(t', t]$).
+	Observe that for requests in $r_t := \sigma_{\{u,v'\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t]$,
+	OPT must pay at least
+	$\min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$
+	within $(0,m]$,
+	regardless of its choices.
+	Based on this observation,
+	we distinguish three cases of OPT's choices
+	while it serves requests in	$r_t $.
+	We denote its cost by
+	$\mathit{Cost}_{\mathit{OPT}} (r_t) := \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v'\}}(t'',t']) 
+	+ \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) $.
+	\begin{enumerate}[label=\roman*.]
+		\item
+		OPT collocates $\{u,v'\}$ and $\{u,v\}$  at times  $t''$ and $t'$ respectively
+		and keeps them so during $(t'',t']$ and $(t',t]$, respectively.
+		Therefore it serves all requests in $r_t$ with no additional cost and
+		$\mathit{Cost}_{\mathit{OPT}} (r_t) = 2\alpha$.
+		\item
+		OPT has $\{u,v\}$ ($\{u,v'\}$) already collocated at $t'$ ($t''$) as well as during $(t',t]$ ($(t'',t]$),
+		and has $\{u,v'\}$ ($\{u,v\}$) separated during $(t'',t']$ ($(t',t]$).
+		Hence,
+		it pays only $\lambda$ for serving all requests in $r_t$ and
+		$\mathit{Cost}_{\mathit{OPT}} (r_t) = \lambda$
+		\item
+		OPT has $\{u,v'\}$ ($\{u,v\}$) already  collocated at $t''$ ($t'$)  and also during $(t'',t']$ ($(t', t])$,
+		 and it collocates $\{u,v\}$ ($\{u,v'\}$) at $t'+1$ ($t''+1$) paying $\alpha$.
+%		Therefore it pays only $\alpha$ during $[t'',t]$,
+%		for collocating $\{u,v'\}$
+%	    and for the one request to $\{u,v\}$ at $t'$.
+%		I.e., it pays zero for every other  request in $r_t$.
+Then,
+$\mathit{Cost}_{\mathit{OPT}} (r_t) = \alpha$.
+	\end{enumerate}
+	 In every of these cases,
+   	we have
+   	$\mathit{Cost}_{\mathit{OPT}} (r_t) \geq  \min{ \{ \lambda, \alpha \}}$.
+  	OPT may deviate from these cases,
+	e.g.~by collocating a pair after partially paying for requests to this pair.
+	 However,
+	 any deviation only increases the cost to OPT.
+	 \mahmoud{It would be better to describe and bound the "deviation" formally.}
+	Therefore in any case,
+ 	$\mathit{Cost}_{\mathit{OPT}} (r_t) \geq  \min{ \{ \lambda, \alpha \}}$.
+
+	It remains to charge both Alg and OPT for requests that do not lead to collocations until the  end of the sequence at $m$.
+	Assume $\{u,v\}$ is not collocated at $m$
+	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
+	which means Alg pays $C^{m}_{ \{u,v\} }$
+	for requests to $\{u,v\}$ not accounted previously.
+	If $\{u,v\}$ is not collocated in the initial configuration
+	and Alg never collocates it through the sequence,
+	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[0,m] |$.
+	We charge OPT the cost
+	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[0,t]) 
+	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
+	for collocating this pair or (and) paying (partially) for requests in $\sigma_{\{u,v\}}[0,t]$.
+	Else,
+	either $\{u,v\}$ is collocated in the initial configuration
+	or Alg collocates it at some point.
+	Then there exists an intersecting pair $\set{u,v'}$
+	that is collocated by Alg at $t' < m$ (separating $\{u,v\}$).
+	We define times $t'' < t' < m$ similarly to the case where Alg collocates $\{u,v\}$.
+	Let $r_{\{u,v\}} := \sigma_{\{u,v'\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',t]$.
+	Then,
+	we apply cases (i) to (iii) for $t=m$ in an analogous way,
+	after replacing $\lambda$ with $C^{m}_{ \{u,v\}}$.
+	Thus, 
+	we obtain
+	 $\mathit{Cost}_{\mathit{OPT}} (r_{\{u,v\}}) 
+	 \geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.
+
+	 Next, we sum up all costs charged to OPT.
+	 Observe that by definition,
+	 $r_{t'} \cap r_t = \sigma_{\{u,v'\}}[t'',t']$.
+	 This means,
+	 the cost of r.h.s.~to OPT is counted at most two times in the expression
+	 $\mathit{Cost}_{\mathit{OPT}} (r_{t'}) + \mathit{Cost}_{\mathit{OPT}} (r_t)$.
+	 Hence,
+	 for all collocations performed by Alg and for $t=m$,
+	 OPT pays at least 
+	 $\sum_{ 0 < t \leq |\sigma|} \mathit{Cost}_{\mathit{OPT}} (r_t) / 2$.
+	Let
+	$T := \{ 0 < t \leq m ~\vert~ \exists \{u,v\}: C^{t}_{\{u,v\}} = \lambda \}$.
+	By setting $\lambda= \alpha$
+	we obtain
+\begin{align*}
+	\mathit{Cost}_{\mathit{ALG}} (\sigma)	/
+	 \mathit{Cost}_{\mathit{OPT}} (\sigma)
+	 &=
+	 \Big(
+	 \sum_{ t \in T}(\lambda + 2\alpha) +
+	 \sum_{\{u,v\}} C^{m}_{\{u,v\}}
+	 \Big)	 \big /
+	 \Big(
+	 \sum_{ t \in T} \mathit{Cost}_{\mathit{OPT}} (r_t) / 2
+	 + \sum_{\{u,v\}}\mathit{Cost}_{\mathit{OPT}} (r_{\{u,v\}})  / 2 
+	 \Big)	\\
+	 &\leq
+	 \Big(
+	 \sum_{ t \in T}(\lambda + 2\alpha) +
+	 \sum_{\{u,v\}} C^{m}_{\{u,v\}}
+	 \Big)	 \big /
+	 \Big(
+	 \sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  / 2
+	 + \sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \}  / 2 } 
+	 \Big)	\\
+	 &=
+	 \Big(
+	\sum_{ t \in T} 3\alpha +
+	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
+	\Big)	 \big /
+	\Big(
+	\sum_{ t \in T} \alpha  / 2
+	+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}  / 2 
+	\Big)	\leq 6
+\end{align*}.
+	%	The claim follows from here.
+\end{proof}
+
+\begin{comment}
+We charge both OPT and Alg whenever Alg collocates a pair (i.e., on a collocation event).
+%More precisely,we charge them only the cost of the collocation, per such event.
+%Hence,we charge the cost of  this collocation to OPT for the first time.
+Alg collocates a pair always with a swap (i.e., two simultaneous migrations),
+which  costs $2\alpha$.
+%We charge OPT only for requests up to the one that triggers a collocation (i.e., not for later requests on the same pair).
+OPT may save some costs by collocating multiple pairs at once (before they incur too much cost),
+paying the price of only one migration per pair.
+This observation suggests that we can't charge OPT more than $\alpha$ per collocation,
+even when OPT actually pays more than that.
 E.g.,
@@
 \begin{proof}[Proof of Theorem \ref{thm:k=2}]
-	\mahmoud{still proof sketch}
-	Let $C^{t}_{\set{u,v}}$ denote the  value of $C_{\set{u,v}}$
-	immediately after arrival of the $t$th request,
-	$\sigma_t$.
+	\mahmoud{not correct}
+%	Let $C^{t}_{\set{u,v}}$ denote the  value of $C_{\set{u,v}}$
+%	immediately after arrival of the $t$th request,
+%	$\sigma_t$.
 	Assume on request $\sigma_{t}=\set{u,v}$, 
 	i.e., at (time) $t \geq 1$,
-	Alg collocates the pair,
-	which means $ C^t_{\set{u,v}} = \lambda$ immediately before Alg resets the counter.
+	Alg collocates the pair $\set{u,v}$.
+	This means,
+	the value of $C_{\set{u,v}}$ at $t$,
+	immediately before Alg resets the counter is $\lambda$.
+	That is, $ C^t_{\set{u,v}} = \lambda$.
 	Let $\sigma_{t''}, t''<t$ be the last request (before $\sigma_{t}$)
 	upon which Alg collocates $\set{u,v}$.
-	Let $t'$ s.t.~$t'' < t' < t$ be the  time when Alg separates $\set{u,v}$
+	Let $t'$, $t'' < t' < t$ be the  time when Alg separates $\set{u,v}$
 	in order to collocate some other pair that shares a node with $\set{u,v}$,
-	denoted by $\set{u,v'}, v' \neq v$.
+	 e.g.~$\set{u,v'}, v' \neq v$.
 	In case the first collocation of $\set{u,v}$ occurs at $t$,
@@
 	we set $ t'' = -\infty, t'=0$.
-	%e.g., the request $\sigma_{t'} = \set{u,v'}$.
-	%By Alg's definition,
-	%$C^{t'}_{\set{u,v'}}  = C^t_{\set{u,v}} = \lambda$.
 	We denote the set of all requests to $\set{u,v}$
-	between times $t'$ and $t$ by $\sigma_{\set{u,v}}[t',t]$.
+	during $[t_1,t_2]$ by $\sigma_{\set{u,v}}[t_1,t_2]$.
 	Thus,
 	$| \sigma_{\set{u,v}}[t',t] | = \lambda$.
-	The overall cost of requests that increase the counter,
-	i.e.~requests in $\sigma_{\set{u,v}}[t',t]$,
+	The overall cost of requests that increase the counters of intersecting pairs,
+%	i.e.~requests in $\sigma_{\set{u,v}}[t'',t]$,
 	and the moving cost,
-	incurred by Alg is $\lambda + 2\alpha$.
+	incurred by Alg is
+	$\lambda + 2\alpha$.
   %That is, the cost of all requests in $\sigma_{\set{u,v}}[t',t]$.
 
-	We distinguish three cases for the cost incurred by OPT due to $\set{u,v}$.
+	We distinguish three cases for the cost incurred by OPT with respect to the collocation of $\set{u,v}$ at $t$.
 
@@
 	and therefore it pays $\lambda$.
-	We charge OPT $\lambda/2 = \alpha$.
-%	The remaining cost will be used in a latter case.
+	We charge OPT $\lambda/2$, i.e.,
+	$\mathit{Cost}^{t}_{\{u,v\}} (\mathit{OPT})= \lambda / 2$.
 
   \item Case ii)  
-	OPT collocates $\set{u,v}$ some time during $[t',t]$.
-	Any cost that we charge to OPT upon $\sigma_{t''}$,
-	accounts for OPT's actions before $t''$.
-	Therefore in this case,
-	we charge OPT the (unaccounted) cost $\alpha$.
+	OPT collocates $\set{u,v}$ some time during $[t'',t]$ and pays $\alpha$.
+	We charge it $\alpha/2$, i.e.,
+	$\mathit{Cost}^{t}_{\{u,v\}} (\mathit{OPT}) = \alpha / 2$.
 
   \item Case iii)
-	$t'' \geq 0$ and OPT has $\set{u,v}$ collocated before $t''$ and keeps them collocated during $[t'',t]$.
-	In this case OPT doesn't pays any cost for requests in $\sigma_{\set{u,v}}[t'',t]$ and
-	we can't charge it directly.
+	$t'' \geq 0$ and OPT has $\set{u,v}$ already collocated before $t''$.
+%	 and keeps them collocated during $[t'',t]$.
+	In this case,
+	OPT possibly pays zero for requests in $\sigma_{\set{u,v}}[t'',t]$.
 	However,
-	OPT must pay $\lambda$ for all requests in $\sigma_{\set{u,v'}}[t'',t']$,
-	 i.e.,
-	 Case (i) applies to $\set{u,v'}$  at $t'$ and
- 	we charge only $\lambda/2$ to OPT (at $t'$).
-	This means,
-	we can charge OPT an additional $\alpha$,
-	i.e.~the remaining cost of requests $\sigma_{\set{u,v'}}[t'',t']$.
+	by assumption,
+	 Alg collocates $\{u,v'\}$ at $t'$ and
+	OPT doesn't have $\{u,v'\}$ collocated through the entire
+	$[\tau^{t''}_{\{u,v'\}}, t']$.
+	This implies that only cases (i) or (ii) applies to $\{u,v'\}$ at $t'$.
+	That is,
+	OPT pays $\min \{ \lambda, \alpha \}$,
+	and it is charged $\min \{ \lambda/2, \alpha/2 \}$
+	w.r.t.~the collocation of $\{u, v'\}$ at $t'$.
+	Hence,
+	we can charge it  $\min \{ \lambda/2, \alpha/2 \}$ in addition to the  cost already charged.
+%	i.e.~the remaining (unaccounted) cost of requests in $\sigma_{\set{u,v'}}[t'',t']$.
 %	Intuitively, at $t$,  we charge $\alpha$ to OPT for separating $\set{u,v}$ (at $t'$) by Alg.
@@
 \end{proof}
+\end{comment}
 
@@
 
+\maciek{$\sigma_t$ undefined (I know what it is, though). \mahmoud{added inline def. Do we have a global def for this?}}
+\maciek{Define $C^t$ to use once \mahmoud{already defined! I don't get it.}}
+\maciek{A word "Obviously" is unnecessary, and most people, including me are allergic to it in writing
+\mahmoud{obviously that's true! connect the two sentences}}
+
 \section{Note on distributed algorithms}

### 2020-04-30 paper/disc20clustering.tex
minor
k=2 proof, clean up

@@
 OPT may save some costs by collocating multiple pairs at once (before they incur too much cost),
-paying the price of only one migration per pair.
-This observation suggests that we can't charge OPT more than $\alpha$ per collocation,
-even when OPT actually pays more than that.
-E.g.,
-OPT can collocate two costly pairs with a single swap if
-it has both pairs on the cut of  the same two clusters  (see Figure \ref{fig:TBD}).
-Later, Alg collocates those pairs with two swaps.
+paying the price of only one migration per pair  (see Figure \ref{fig:TBD}).
+%This observation suggests that we can't charge OPT more than $\alpha$ per collocation.
+%E.g.,
+%OPT can collocate two costly pairs with a single swap if
+%it has both pairs on the cut of  the same two clusters.
+%Later, Alg collocates those pairs with two swaps.
 Hence,
 whenever OPT collocates a pair,
-we charge OPT only the cost of moving a single node to the cluster of the other,
+we charge it only the cost of moving a single node to the cluster of the other,
 i.e., $\alpha$ (in contrast to $2\alpha$ on Alg).
@@
 %	we set $ t'' = -\infty, t'=0$.
-	We denote the set of all requests to $\set{x,y}$
+	We denote the set of all requests to a pair $\set{x,y}$ that arrive
 	during $(t_1,t_2]$ by $\sigma_{\set{x,y}}(t_1,t_2]$.
-	If $t$ is the first time Alg collocates $\{u,v\}$ then we can charge
+	If $t$ is the first time that Alg collocates $\{u,v\}$ then we  charge
 	OPT $\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
@@
 	in order to collocate some (intersecting) pair
-	$\{x,y\} \neq \{u,v\}, x \in \{u,v\}$, 
+	$\{x,y\} \neq \{u,v\}, \{x,y\} \cap \{u,v\} \neq \emptyset$, 
 	e.g.,
@@
 	Similarly,
-	if $t'$ is not the first time that Alg has $\{u,v'\}$ collocated
+	if $t'$ is not the first time that Alg  collocates $\{u,v'\}$ 
 	then let $t''<t'$ be the last time before $t'$ when Alg separates $\set{u,v'}$.
 	Else,
-	$\{u,v'\}$ is not collocated in the initial configuration and let $t''=0$.
-	Then by these definitions,
+	$\{u,v'\}$ is not collocated in the initial configuration and we let $t''=0$.
+	By these definitions,
 	we have
@@
 	$\lambda + 2\alpha$.
-	We charge OPT the cost of serving requests in
-	$\sigma_{\{u,v'\}}[t'',t']$ ($\sigma_{\{u,v\}}(t',t]$ ) 
-	that it pays only within the interval $(t'', t']$ ($(t', t]$).
 	Observe that for requests in $r_t := \sigma_{\{u,v'\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t]$,
-	OPT must pay at least
+	OPT must pay a total cost of serving and (or) migration at least
 	$\min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$
@@
 	regardless of its choices.
+	%% 
+	However,
+	we cannot charge this cost to OPT repeatedly,
+	i.e.,
+	when Alg collocates the two pairs a second time.
+	In particular,
+	we charge OPT a cost incurred by requests in
+	$\sigma_{\{u,v'\}}(t'',t']$ and $\sigma_{\{u,v\}}(t',t]$,
+	if and only if  OPT incurs the cost within  intervals $(t'', t']$ and $(t', t]$,
+	respectively.	
+	Latter,
+	this ensures us that we do not charge OPT for a cost more than twice.
+	%%
 	Based on this observation,
@@
 		\item
-		OPT collocates $\{u,v'\}$ and $\{u,v\}$  at times  $t''$ and $t'$ respectively
+		OPT collocates $\{u,v'\}$ at  $t''$, $\{u,v\}$  at $t'$,
 		and keeps them so during $(t'',t']$ and $(t',t]$, respectively.
@@
 	 However,
-	 any deviation only increases the cost to OPT.
+	 no deviation can decrease the cost to OPT.
 	 \mahmoud{It would be better to describe and bound the "deviation" formally.}
@@
 
-	It remains to charge both Alg and OPT for requests that do not lead to collocations until the  end of the sequence at $m$.
-	Assume $\{u,v\}$ is not collocated at $m$
+	It remains to charge both Alg and OPT for requests that do not lead to collocations until the  end of the sequence at $t=m$.
+	Assume $\{u,v\}$ is not collocated at $t=m$
 	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
 	which means Alg pays $C^{m}_{ \{u,v\} }$
-	for requests to $\{u,v\}$ not accounted previously.
+	for requests to $\{u,v\}$ so far unaccounted.
 	If $\{u,v\}$ is not collocated in the initial configuration
 	and Alg never collocates it through the sequence,
-	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[0,m] |$.
+	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[1,m] |$.
 	We charge OPT the cost
-	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[0,t]) 
+	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[1,t]) 
 	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
-	for collocating this pair or (and) paying (partially) for requests in $\sigma_{\{u,v\}}[0,t]$.
+	for collocating this pair or (and) paying (partially) for requests in $\sigma_{\{u,v\}}[1,t]$.
 	Else,
@@
 	We define times $t'' < t' < m$ similarly to the case where Alg collocates $\{u,v\}$.
-	Let $r_{\{u,v\}} := \sigma_{\{u,v'\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',t]$.
+	Let $r_{\{u,v\}} := \sigma_{\{u,v'\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
 	Then,
-	we apply cases (i) to (iii) for $t=m$ in an analogous way,
-	after replacing $\lambda$ with $C^{m}_{ \{u,v\}}$.
+	we apply cases (i) to (iii) in an analogous way,
+	 replacing $\lambda$ with $C^{m}_{ \{u,v\}}$ in the analysis.
 	Thus, 
@@
 	 This means,
-	 the cost of r.h.s.~to OPT is counted at most two times in the expression
+	 the cost of r.h.s.~to OPT is counted at most twice in the expression
 	 $\mathit{Cost}_{\mathit{OPT}} (r_{t'}) + \mathit{Cost}_{\mathit{OPT}} (r_t)$.
 	 Hence,
-	 for all collocations performed by Alg and for $t=m$,
+	 for all collocations performed by Alg,
+	 and for final costs at $t=m$,
 	 OPT pays at least 
-	 $\sum_{ 0 < t \leq |\sigma|} \mathit{Cost}_{\mathit{OPT}} (r_t) / 2$.
+	 $(
+	 \sum_{ 0 < t \leq |\sigma|} \mathit{Cost}_{\mathit{OPT}} (r_t) +
+	 \sum_{\{u,v\}} \mathit{Cost}_{\mathit{OPT}} (r_{\{u,v\}})
+	 ) / 2$.
 	Let
 	$T := \{ 0 < t \leq m ~\vert~ \exists \{u,v\}: C^{t}_{\{u,v\}} = \lambda \}$.
-	By setting $\lambda= \alpha$
+	By setting $\lambda= \alpha$,
 	we obtain
@@
 	 \Big)	 \big /
+	 \frac{1}{2}
 	 \Big(
-	 \sum_{ t \in T} \mathit{Cost}_{\mathit{OPT}} (r_t) / 2
-	 + \sum_{\{u,v\}}\mathit{Cost}_{\mathit{OPT}} (r_{\{u,v\}})  / 2 
+	 \sum_{ t \in T} \mathit{Cost}_{\mathit{OPT}} (r_t) 
+	 + \sum_{\{u,v\}}\mathit{Cost}_{\mathit{OPT}} (r_{\{u,v\}})
 	 \Big)	\\
@@
 	 \Big)	 \big /
+		\frac{1}{2}
 	 \Big(
-	 \sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  / 2
-	 + \sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \}  / 2 } 
+	 \sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  +
+   \sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \} } 
 	 \Big)	\\
 	 &=
-	 \Big(
+	 2\Big(
 	\sum_{ t \in T} 3\alpha +
@@
 	\Big(
-	\sum_{ t \in T} \alpha  / 2
-	+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}  / 2 
-	\Big)	\leq 6
+	\sum_{ t \in T} \alpha  
+	+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}  
+	\Big)	< 6
 \end{align*}.

### 2020-04-30 paper/disc20clustering.tex
Readability

@@
 	during $(t_1,t_2]$ by $\sigma_{\set{x,y}}(t_1,t_2]$.
+
 	If $t$ is the first time that Alg collocates $\{u,v\}$ then we  charge
@@
 	OPT pays $\min{\{ \lambda,\alpha \}}$.
-	Else,
-	let $t'$, $t' < t$ be the  time when Alg separates $\set{u,v}$
+
+	In the remainder of this proof, we consider that Alg collocated $\{u,v\}$ in the past.
+	Let $t'$, $t' < t$ be the  time when Alg separates $\set{u,v}$
 	in order to collocate some (intersecting) pair

### 2020-04-30 paper/disc20clustering.tex
Comments

@@
 
-	In the remainder of this proof, we consider that Alg collocated $\{u,v\}$ in the past.
+	Else, we consider that Alg collocated $\{u,v\}$ in the past.
 	Let $t'$, $t' < t$ be the  time when Alg separates $\set{u,v}$
@@
 
+Major:
+
+\maciek{I feel slightly uneasy here: "In particular,
+we charge OPT a cost incurred by requests in
+$\sigma_{\{u,v'\}}(t'',t']$ and $\sigma_{\{u,v\}}(t',t]$,
+if and only if  OPT incurs the cost within  intervals $(t'', t']$ and $(t', t]$,
+respectively.	"}
+
+\maciek{"Observe that for requests in $r_t$".
+First the separate sentence with the definition of $r_t$ (it is challenging enough to just understand it.
+Then reason why it pays the minimum of three values (I don't see it, and later you show that cost(rt) is at least minimum of other two values).
+Is it a consequence of three cases?}
+
+\maciek{"We charge OPT the cost
+$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[1,t]) 
+\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
+for collocating this pair or (and) paying (partially) for requests in $\sigma_{\{u,v\}}[1,t]$." No issues with charge sharing from "regular collocation" case?
+Maybe I just don't understand "partially" here}
+
+Minor:
+
+\maciek{Remove all "charge both". Just sum up costs. I would say not use "charge" at all, but it's up to you.}
+
+\maciek{Structure in paragraphs.
+3 parts: first collocation, regular collocation and unfinished.
+Do this for ALG and then for OPT, and do not mix (unless I don't see some dependencies).
+Start each paragraph by the sentence that says what the paragraph is about (first/opt, regular/alg etc.).
+Assume that the reader reads only the first sentence of each paragraph (if it's short), and the proof must be clear from this (for a professional)}
+
+\maciek{$r_t$ is the central concept in this proof.
+Possibly even start with the paragraph: charging OPT, regular collocation.
+Make this definition in its separate line to further emphasize?
+It would suffice for a profesional to see the definiton of $t''/t'$ and this line to convince himself that the proof is correct.}
+
+\maciek{I think that "deviate" is OK to use like this.}
+
+Notation:
+
+\maciek{End of the time proposition: $t_e / t^* / T$ instead of $m$? Or capital $T$ (but I am aware of the conflict)}
+
+\maciek{Instead of $cost_{opt}$ just use opt / alg?}
+
+\maciek{Consider naming the adjacent edges $e$ and $f$.}
+
 \begin{comment}
@@
 
-Missing:
-\maciek{Missing the case of the first-time collocation. What if t' does not exist? What if t'' does not exist? 
-\mahmoud{added remark for this case: $t''=0$
-\maciek{But t' might not exist (as it might be the first collocation).}}}
-
-Minor:
-\maciek{"It can be easily verified"... True, but avoid such formulation. Something along these lines: if opt migrates earlier than t', then additionally pays for X, if later than Y.}
-
-Tiny:
-\maciek{$\sigma_t$ undefined (I know what it is, though). \mahmoud{added inline def. Do we have a global def for this?}
-\maciek{Not yet, but we still must write the "Problem definition" section (yes, we don't have one!), and we can put it there.}}
-
-\maciek{Define $C^t$ to use once \mahmoud{already defined! I don't get it.
-\maciek{discussed}}}
-
-\maciek{$\sigma_t$ undefined (I know what it is, though). \mahmoud{added inline def. Do we have a global def for this?}}
-\maciek{Define $C^t$ to use once \mahmoud{already defined! I don't get it.}}
-\maciek{A word "Obviously" is unnecessary, and most people, including me are allergic to it in writing
-\mahmoud{obviously that's true! connect the two sentences}}
-
 \section{Note on distributed algorithms}

### 2020-05-01 paper/disc20clustering.tex
Comments
Distributed k=2 sketch

@@
 for collocating this pair or (and) paying (partially) for requests in $\sigma_{\{u,v\}}[1,t]$." No issues with charge sharing from "regular collocation" case?
-Maybe I just don't understand "partially" here}
+Maybe I just don't understand "partially" here.
+Not necessarily rewrite!
+Just maybe check if it's true.}
 
@@
 Start each paragraph by the sentence that says what the paragraph is about (first/opt, regular/alg etc.).
-Assume that the reader reads only the first sentence of each paragraph (if it's short), and the proof must be clear from this (for a professional)}
+Assume that the reader reads only the first sentence of each paragraph (if it's short), and the proof structure must be clear from this.
+Some other important things go to their own paragraph too, and the first sentence must describe what we do there.}
 
@@
 
+\subsection{Distributed implementation of $k=2$ algorithm}
+
+For a repartition, it is sufficient to find the cluster with the signature $1,1,1$. 
+We call any such cluster \emph{X (todo name)}.
+Idea of distributed implementation of this procedure:
+\begin{enumerate}
+	\item Assign ids to clusters and order them
+	\item Recursively partition clusters into two groups
+	\item Each group maintains the pointer to a subgroup that contains at least one X cluster; a pointer may be null (then no X there)
+	\item This way, finding a cluster X requires $\log_2(\ell)$ communication
+	\item After removing an X cluster, we need to fix the pointers. This is done by following the path from the removed X cluster to the root. We simply ask the sibling if it contains X.
+\end{enumerate}
+
 \maciek{How to distribute our algorithms: PPL, k=2 and DET}

### 2020-05-01 paper/disc20clustering.tex
k=3

--- 2020-05-01 paper/disc20clustering.tex
@@
 
-\subsection{Distributed implementation of $k=2$ algorithm}
+\subsection{Distributed implementation of $k=3$ algorithm}
 
@@
 	\item After removing an X cluster, we need to fix the pointers. This is done by following the path from the removed X cluster to the root. We simply ask the sibling if it contains X.
+	\item Possible improvement: we can have a fixed number of rounds for finding the X cluster by maintaining a direct pointer to an X cluster (not only to subgroup).
+However, the fix still requires the log number of rounds.
+But we improved the responsiveness.
 \end{enumerate}
 
-\maciek{How to distribute our algorithms: PPL, k=2 and DET}
+\maciek{How to distribute algorithms: PPL and DET}
 

### 2020-05-01 paper/disc20clustering.tex
wip
PPL question
Table updated

@@
 
-\title{Improved Bounds for Online Balanced Repartitioning}
+\title{Online Balanced Graph Partitioning: Optimizations for Large Clusters}
 
@@
 \setcopyright{acmlicensed}
-\acmConference{PODC '20}{July 29-August 2, 2020}{Toronto, Canada}
+\acmConference{DISC '20}{October 12-16, 2020}{Freiburg, Germany}
 
@@
 \newcommand{\OFF}{\textsc{Off}\xspace}
+\newcommand{\Rep}{\textsc{Rep}}
 
@@
 
-\maciek{Title discussion: "Online Balanced Graph Partitioning: Understanding the role of clusters". Reason: all our results most notably pertain to the number of clusters.}
-
 \noindent \textbf{Motivation and model.}
@@
 and \emph{external} requests, respectively.
-\mahmoud{"Before serving a request": does this mean that only after a request arrives and before serving it we migrate?
-In our algorithms we migrate after serving $\lambda$ requests.
-Do we have discrepancy with the model?}
 Before serving a request,
@@
 is~$\alpha \in \mathbb{Z}^+$.
+\maciek{Why restrict $\alpha$? It should just be positive}
 For any algorithm $\ALG$,
@@
     \textbf{Variant} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
-      \textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & TBD\hspace{0.3cm}(\S 9.3) \\ 
+      \textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & 6\hspace{0.3cm}(\S \ref{sec:k2}) \\ 
       \rowcolor{gray!25}
-      \textbf{$k=3$}&  $27\cdot \ell $\hspace{0.3cm}(\S 9.3)& $O(\ell)$\hspace{0.3cm}(\S 9.3)\\
-      \textbf{$\ell=2$}&  $k+1 = \Omega(k)$\hspace{0.3cm}(\S 9.3)&\hspace{0.3cm}$O(k^2)$\hspace{0.3cm}\cite{repartition-disc}\\
+      \textbf{$k=3$}&  $\ell - 3/2$ \hspace{0.3cm}(\S \ref{sec:lowerbound})& $54\cdot \ell $\hspace{0.3cm}(\S \ref{sec:k3})\\
+      $k > 3$ & $(k\cdot \ell - k - \ell) / 2$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell \cdot \min \{ k \cdot \ell, \Rep(k, \ell) \})$\hspace{0.3cm} (\S \ref{sec:k3}) \\
       \rowcolor{gray!25}
-      $\ell \geq 3$& $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S 9.3)&$O(k^2 \cdot \ell^2)$\hspace{0.3cm}\cite{repartition-disc} \\
-      Perfect partition & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S 9.3)&$O(k \cdot \ell)$\hspace{0.3cm} (\S 9) \\
+      Perfect partition & $(k\cdot \ell - k - \ell) / 2$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$2(k -1 )\cdot \ell$\hspace{0.3cm} (\S \ref{sec:ppl}) \\
     \end{tabular}
-                \caption{Overview of the results.
-Most specific cases are tight.
-The general case still has a significant gap between the lower bound and the upper bound.
-If either $k$ or $\ell$ is missing from a variant description, it is assumed to be arbitrary.
+                \caption{Overview of known results and the results provided in this paper.
+By $\Rep$ we denote the cost of a single component repartition, and we elaborate on this in Section~\ref{sec:k3}
                   }
@@
 \subsection{Upper bound.}
+\label{sec:ppl}
 Now we give an algorithm for a restricted variant of  \OBRP{}, named \PPOBRP{},
@@
 \section{Improved algorithm for $k=2$}
+\label{sec:k2}
 
@@
 
-\maciek{How to distribute algorithms: PPL and DET}
+\subsection{Distributed implementation of $k=2$ algorithm}
+
+Trivial? Counters are local. We migrate between the clusters that have just communicated.
+
+\subsection{Distributed implementation of PPL}
+
+\maciek{How to distribute PPL better?}
 

### 2020-05-01 paper/disc20clustering.tex
Pass over intro

@@
 
-We consider two flavors of the problem
-in this paper. In the first one, which
-was also studied by Henzinger et al.~\cite{sigmetrics19_partitioning}
-(SIGMETRICS 2019) and which
-we here mainly consider for the sake of a stronger
-lower bound, we assume that $\sigma$
-simply reveals the edges of a static graph
+In addition to the general model, we study a simplified Perfect Partition model (also studied by Henzinger et al.~\cite{sigmetrics19_partitioning} (SIGMETRICS 2019)) that we consider for a stronger lower bound.
+There, we assume that $\sigma$
+reveals the edges of a static graph
 that can be perfectly partitioned
-(i.e., in principle, no external requests
-are required). 
-In the second one, we consider a general
-model where $\sigma$ can be arbitrary;
-this was the original problem introduced
+(no external requests are required).
+
+The general model, where $\sigma$ can be arbitrary
+was the original problem introduced
 by Avin et al.~\cite{repartition-disc} at DISC 2016.
-For simplicity, we will refer to the former
+For simplicity, we will refer to the Perfect Partition
 model as the \emph{learning} model (as
 one has to learn the static communication graph) 
-and to the latter as the \emph{online} model.
+and to the original as the \emph{repartitioning} model (as one continuously adjusts the partition in response to changing communication graph).
 
@@
 $O(k\cdot \ell)$-competitive algorithm
-for the restricted model.
-For the online model, we present  
+for the learning model.
+For the repartitioning model, we present  
 an asymptotically optimal,
@@
 so far was $O(\ell^2)$~\cite{repartition-disc}.
+Then, we present a strictly $6$-competitive algorithm for $k=2$.
+In addition, we show a new, improved analysis of the algorithm from \cite{repartition-disc} that depends on the solutions to an interesting combinatorial problem (see Section~\ref{ssec:cascade}, The Cascade Problem).
 %
-Our algorithms can be distributed
-similarly to the approach in~\cite{sigmetrics19_partitioning}.
 The overview of known results is in Table~\ref{tab:overview}.
-
-\maciek{TODO: fill the table: results and references to sections.}
 
@@
 For clusters of size $2$ (i.e., $k=2$), 
-it is known that a constant competitive algorithm exists~\cite{repartition-disc}.
+it is known that a $7$-competitive algorithm exists~\cite{repartition-disc} with large ($\Omega(\ell^2)$) additive constant.
 
@@
 
-\noindent \textbf{Preliminaries.}
+\section{Preliminaries.}
 
@@
 
+Our algorithms can be distributed
+similarly to the approach in~\cite{sigmetrics19_partitioning}.
+
 \bibliographystyle{ACM-Reference-Format}

### 2020-05-01 paper/disc20clustering.tex

@@
 
-\maciek{Here, we need a good and broad description of what a component is.}
-
-In this paper, we use the technique of maintaining \emph{(connected) components} of nodes, similar to previous approaches~\cite{repartition-disc}.
-A~\emph{component} is a subset of frequently communicating nodes \maciek{This is not a definition, just some intuitions.
-More formal needed.}.
-We use this concept in both algorithms in this paper, but also in the lower bound, to incur high cost for any online algorithm that splits them.
-An~algorithm is \emph{component respecting}
+Across this paper, we often refer to groups of communicating nodes.
+As this concept is used differently in the lower bound and the upper bounds, we name them differently to avoid confusion.
+
+\noindent
+\textbf{Components.}
+In our algorithms, a~\emph{component} is a subset of frequently communicating nodes.
+We maintain all nodes of each component collocated in one cluster~\cite{repartition-disc}.
+If we need to move a node, we move the whole component that contains it.
+We maintain a balanced partition of our components, a similar concept to partition of integers into sets with equal sum (\maciek{TODO: cite}).
+In contrast, our partition is time-varying: communicating components are merged into one, and we adjust the partition.
+
+\noindent
+\textbf{Ground sets.}
+In the lower bound, a~\emph{ground set} is a group of nodes that start communicating if an algorithm splits them (places at different clusters).
+Their role is to ensure that the algorithm maintains a partition of ground sets.
+We emphasize that ground sets are revealed only upon their nodes are split.
+\maciek{Other names: division, segmentation, alignment, formation}
+
+\maciek{Needed? "An~algorithm is \emph{component respecting}
 if it always keeps nodes that  belong to the same component collocated.
@@
 if the algorithm needs to move a node,
-it moves the whole component containing the node.
+it moves the whole component containing the node."}
 %We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
@@
 
-\maciek{Let's follow the SIDMA notation here. For example, a "single" node (the one that belongs to singleton component) is called "trivial" there.}
+We say that a component or a ground set is \emph{trivial}, if it contains a single node.
 
@@
 
+At the end of this section, we discuss an asymptotically optimal upper bound.
+
 \subsection{Lower bound.}
@@
 
-\maciek{TODO: rename everywhere: "adversary's components", i.e., all components in lower bounds should be called "ground sets"}
-
-\maciek{The text before the proof should have two parts: a description of ground sets and then general intro to the proof.}
-
 In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
-It is sufficient to have clusters of capacity $k\geq 3$ for this lower bound.
-For $k=2$, a constant competitive algorithm exists.
-At the end of this section, we also
-discuss an upper bound.
+To show it, clusters of capacity $k\geq 3$ are sufficient.
+Note that for $k=2$, a lower bound of $3$ is known \cite{repartition-disc}.
 
@@
 
-The adversary constructs components of nodes based on the partitioning choices of a given online algorithm.
-Adversary issues requests between pairs of nodes that belong to the same component.
-If an online algorithm splits a component at any moment,
-the adversary issues requests to the split pair until the algorithm collocates them.
+The adversary constructs ground sets in reaction to choices of a deterministic online algorithm.
+Once we construct a ground set, it lasts until the end of the input sequence.
+If the algorithm splits any pair of nodes of a ground set, the adversary issue requests to any pair of split nodes.
+
+We construct the ground sets in a way that they always form a balanced partition of nodes.
+This way, the optimal algorithm may move to the ground-set-respecting configuration, and serve all requests for free.
+(This implies that the algorithm must eventually move to a ground-set-respecting configuration if it is competitive.)
+
+\maciek{Finished here.}
+
 We utilize the fact that the online algorithm does not know the components of the adversary, and it may split some components while evicting nodes.

### 2020-05-01 paper/disc20clustering.tex
Intro to the lower bound

@@
 	\end{tabular}
-                \caption{Overview of known results and the results provided in this paper.
-By $\Rep$ we denote the cost of a single component repartition, and we elaborate on this in Section~\ref{sec:k3}
+	\caption{Overview of known results and the results provided in this paper. By $\Rep$ we denote the cost of a single component repartition, and we elaborate on this in Section~\ref{sec:k3}
 	}
@@
 
-\maciek{Needed? "An~algorithm is \emph{component respecting}
-if it always keeps nodes that  belong to the same component collocated.
-That is,
-if the algorithm needs to move a node,
-it moves the whole component containing the node."}
 %We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
@@
 We say that a component or a ground set is \emph{trivial}, if it contains a single node.
+We say that a~configuration is \emph{component respecting}
+if the nodes belonging to the same component are collocated.
+Similarly, we define a \emph{ground-set-respecting} configuration.
+
+\maciek{Eviction - introduce.}
 
@@
 The adversary constructs ground sets in reaction to choices of a deterministic online algorithm.
+The algorithm does not know the ground set before they are revealed, and it may split some of them while evicting nodes.
+
 Once we construct a ground set, it lasts until the end of the input sequence.
@@
 We construct the ground sets in a way that they always form a balanced partition of nodes.
-This way, the optimal algorithm may move to the ground-set-respecting configuration, and serve all requests for free.
-(This implies that the algorithm must eventually move to a ground-set-respecting configuration if it is competitive.)
-
-\maciek{Finished here.}
-
-We utilize the fact that the online algorithm does not know the components of the adversary, and it may split some components while evicting nodes.
-On the other hand, $\OPT$ initially collocates all nodes of each component and never splits them again.
-
-\maciek{Here, we would like to introduce insights about the proof, emphasize the conditions for the bound to work (such as $k\geq 3$), and maybe compare to previously known lower bounds.
-Also, introduce the notation and definitions that are needed.}
-
-\maciek{Note that the lower bound works for any value of $\alpha$, and we just use the number of migrations}
+This way, the optimal algorithm moves to a~ground-set-respecting configuration, and serves all requests for free.
+(This implies that the algorithm must eventually move to any ground-set-respecting configuration if it is competitive.)
+
+The lower bound requires $k\geq 3$.
+It is oblivious to the choice of the reconfiguration cost $\alpha$, and we compare the number of performed node exchanges.
 
 \begin{theorem}
-  The competitive ratio of any deterministic online algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$ for any $k\geq 3$.
-  \maciek{TODO: the exact bound, not asymptotic}
-
-  \maciek{The bound $(k \cdot \ell - k - \ell) / 2$ can be easily improved to $(k \cdot \ell - k - 1) / 2$ by playing a simple caching game in the first cluster, instead of just giving away the $k-1$ "blocker" (the caching game would build it gradually)}
+	\label{th:lowerbound}
+	The competitive ratio of any deterministic online algorithm for \OBRP{} is  $(k \cdot \ell - k - \ell) / 2 = \Omega(k\cdot \ell)$ for any $k\geq 3$.
 \end{theorem}

### 2020-05-01 paper/disc20clustering.tex
Non-formal definitions in the lower bound

@@
 
-We say that a component or a ground set is \emph{trivial}, if it contains a single node.
+We say that a component or a ground set is \emph{trivial}, if it contains exactly one node.
+We say that the node that belongs to trivial component or a ground set is single.
 We say that a~configuration is \emph{component respecting}
@@
 	\label{th:lowerbound}
-	The competitive ratio of any deterministic online algorithm for \OBRP{} is  $(k \cdot \ell - k - \ell) / 2 = \Omega(k\cdot \ell)$ for any $k\geq 3$.
+	The competitive ratio of any deterministic online algorithm for \OBRP{} is  $(k \cdot \ell - k - \ell) / 2 = \Omega(k\cdot \ell)$ for any $k\geq 3$ and for any $\ell \geq 2$.
 \end{theorem}
@@
 \begin{proof}
-	We construct an instance of the problem with $\ell$ clusters 
-	$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
-	Let $I(C)$ denote the cluster where nodes of component $C$ are located initially.
-	%	\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
 	Fix any online algorithm \ALG{}.
-
-	\maciek{This paragraph should be before the proof, merged with section about ground sets (i.e., adversary components.)}
-	We join some nodes into components, and produce the input sequence by observing actions of \ALG{} with respect to nodes of these components.
-	If \ALG{} at any point splits a component
-	(i.e., spreads its nodes over two or more clusters),
-	then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
-	If \ALG{} never collocates them, then it is not competitive.
-
-	Initially each node belongs to its own component of size $1$.
-	We say that a node is single if it belongs to a singleton component.
-	We begin by forming a component of single $k-1$ nodes on the cluster $S_1$.
-
-	Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
-	Next,
-	we merge components of $x_0$~and~$y_0$.
-	Since these reside on different clusters of \ALG{}, it produces an external request.
-	\ALG{} must collocate them in a~cluster other than $S_1$.
-	To this end,
-	\ALG{} moves to a new configuration
-	that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
+	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
+	\maciek{The problem with the definition! The first pair does not have the same initial location of its nodes.}
+	Initially all nodes are single (they belong to their own trivial ground sets).
+
+	In our construction we distinguish a cluster $S_1$, where we form a ground set of size $k-1$.
+	Let $x_0$  be the only single node left on $S_1$ and let $y_0$ be an  arbitrary node that does not reside in $S_1$.
+	First,
+	we construct a ground set $\{x_0, y_0\}$.
+	Since \ALG{} has them split, the adversary produces an external request.
+	Since $S_1$ contains a ground set of size $k$, \ALG{} must collocate $\{x_0, y_0\}$ in another cluster.
+	To preserve a feasible partition of nodes into clusters,
+	 replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
+	 \maciek{Not covering all cases. $y_0$ may stay in its place. This sentence should be about $x_0$ replaced by $x_1$ only.}
 

### 2020-05-01 paper/disc20clustering.tex

@@
 We say that a component or a ground set is \emph{trivial}, if it contains exactly one node.
-We say that the node that belongs to trivial component or a ground set is single.
+We say that the node that belongs to trivial component or a ground set is \emph{alone}.
 We say that a~configuration is \emph{component respecting}
@@
 In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
-To show it, clusters of capacity $k\geq 3$ are sufficient.
+It is oblivious to the choice of the reconfiguration cost $\alpha$, and we compare the number of performed node exchanges.
+The lower bound requires $k\geq 3$.
 Note that for $k=2$, a lower bound of $3$ is known \cite{repartition-disc}.
@@
 
-The lower bound requires $k\geq 3$.
-It is oblivious to the choice of the reconfiguration cost $\alpha$, and we compare the number of performed node exchanges.
+First, we construct a ground set of size $k-1$ on one of the clusters.
+In any configuration, there exists a node that accompanies it.
+We only issue requests between such accompanying (time-varying) node, and any node that was initially collocated with it.
+Repeating such requests, almost every node must visit this cluster.
+In comparison, \OPT performs one migration.
 
@@
 	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
-	\maciek{The problem with the definition! The first pair does not have the same initial location of its nodes.}
+	\maciek{A problem with the definition! The first pair does not have the same initial location of its nodes.}
 	Initially all nodes are single (they belong to their own trivial ground sets).
 
-	In our construction we distinguish a cluster $S_1$, where we form a ground set of size $k-1$.
-	Let $x_0$  be the only single node left on $S_1$ and let $y_0$ be an  arbitrary node that does not reside in $S_1$.
+	We distinguish a cluster $S_1$, where we form a ground set of size $k-1$.
+	We never dismiss any constructed ground set.
+	Each cluster host exactly $k$ nodes, hence in any configuration of \ALG, there exists a single alone node in $S_1$.
+
+	We construct a ground set containing this alone node (and some other node).
+	Precisely, let $x_0$  be the alone node from $S_1$ and let $y_0$ be an arbitrary node that does not reside in $S_1$.
 	First,
@@
 	To preserve a feasible partition of nodes into clusters,
-	 replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
-	 \maciek{Not covering all cases. $y_0$ may stay in its place. This sentence should be about $x_0$ replaced by $x_1$ only.}
-
-	Next,
-	we merge the component of $x_1$ and the largest component $C$ s.t.~$I(C) = I(x_1)$.
+	\ALG must replace $x_0$ with another single node $x_1$.
+
+	We continue to construct ground sets containing the current alone node from $S_1$.
+	Precisely, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t.~$I(C) = I(x_1)$.
+	To preserve a feasible partition of nodes into clusters,
 	\ALG{} must collocate $x_1$ and $C$ in a~cluster other than $S_1$ and
-	consequently replace $x_1$ with some other (single) node $x_2$.
+	consequently replace $x_1$ with another alone node $x_2$.
+
+	We end this procedure prematurely (before all nodes visit $S_1$) to keep \OPT's cost low.
 	We repeat merging the component of single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
@@
 	we have $I(x') \neq I(y')$.
-	At this point there are at most $\ell+1$ single nodes left,
+	Note that at this point there are at most $\ell+1$ single nodes left,
 	otherwise, there would be more pairs of single nodes that were initially in the same cluster.
 
+	Finally, we bound the cost of \OPT.
 	Given this sequence of requests,
-	the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
+	\OPT migrates $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
 	swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.

### 2020-05-01 paper/disc20clustering.tex
Remark on resource augmentation

@@
 
-\maciek{Remark on resource augmentation. The lower bound for k=3 works even with significant augmentation (linear).}
+\noindent
+\textbf{Remark on resource augmentation.}
+The majority of work on Online Balanced Partioning so far (todo: cite) focuses on the scenerio with resource augmentation.
+We note that this lower bound provides insights into this scenario as well.
+Consider an instance with $k$ divisible by $3$, and construct $3$ ground sets of size $k/3$ in each cluster.
+Then, apply the construction from the lower bound for $k=3$, now treating these ground sets in the way we used individual nodes.
+For the resource augmentation as large as $1+1/3-\epsilon$, no more than $3$ such ground sets fit in one cluster.
+Therefore, the lower bound $\Omega(\ell)$ holds.
 

### 2020-05-01 paper/disc20clustering.tex
No ground set respecting

@@
 We say that the node that belongs to trivial component or a ground set is \emph{alone}.
-We say that a~configuration is \emph{component respecting}
+We say that a~configuration is \emph{component-respecting}
 if the nodes belonging to the same component are collocated.
-Similarly, we define a \emph{ground-set-respecting} configuration.
 
@@
 We construct the ground sets in a way that they always form a balanced partition of nodes.
-This way, the optimal algorithm moves to a~ground-set-respecting configuration, and serves all requests for free.
-(This implies that the algorithm must eventually move to any ground-set-respecting configuration if it is competitive.)
+This way, the optimal algorithm moves to such a configuration, and serves all requests for free.\maciek{i can do better: such a?}
+(This implies that the algorithm must eventually collocate nodes of a ground set to be competitive.)
 

### 2020-05-01 paper/disc20clustering.tex
Comments

@@
 \begin{proof}
+	\maciek{Better name than $S_1$?}
 	Fix any online algorithm \ALG{}.
@@
 	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
-	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate on exclusion?}.
+	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate. Don't use the word "exclude" unless very obvious. Just sum up the real cost.}.
 	Finally, $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.

### 2020-05-01 paper/disc20clustering.tex
Remark on resource augmentation
Typos

@@
 \textbf{Remark on resource augmentation.}
-The majority of work on Online Balanced Partioning so far (todo: cite) focuses on the scenerio with resource augmentation.
+The majority of work on Online Balanced Partitioning so far (todo: cite) focuses on the scenario with resource augmentation.
 We note that this lower bound provides insights into this scenario as well.
@@
 For the resource augmentation as large as $1+1/3-\epsilon$, no more than $3$ such ground sets fit in one cluster.
-Therefore, the lower bound $\Omega(\ell)$ holds.
+The cost of any algorithm (including \OPT) scales up by $k/3$, and the lower bound $\Omega(\ell)$ holds.
+
+Finally, we note the possibility for improvement (a gap). The algorithm CREP~\cite{repartition-disc} requires $(2+\epsilon)$-augmentation to guarantee the competitive ratio independent of $\ell$.
 

### 2020-05-02 paper/disc20clustering.tex
Lower bound & table asymptotic

@@
 		\rowcolor{gray!25}
-		\textbf{$k=3$}&  $\ell - 3/2$ \hspace{0.3cm}(\S \ref{sec:lowerbound})& $54\cdot \ell $\hspace{0.3cm}(\S \ref{sec:k3})\\
-		$k > 3$ & $(k\cdot \ell - k - \ell) / 2$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell \cdot \min \{ k \cdot \ell, \Rep(k, \ell) \})$\hspace{0.3cm} (\S \ref{sec:k3}) \\
+		\textbf{$k=3$}&  $\Omega(\ell)$ \hspace{0.3cm}(\S \ref{sec:lowerbound})& $O(\ell) $\hspace{0.3cm}(\S \ref{sec:k3})\\
+		$k > 3$ & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell \cdot \min \{ k \cdot \ell, \Rep(k, \ell) \})$\hspace{0.1cm} (\S \ref{sec:k3}) \\
 		\rowcolor{gray!25}
-		Perfect partition & $(k\cdot \ell - k - \ell) / 2$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$2(k -1 )\cdot \ell$\hspace{0.3cm} (\S \ref{sec:ppl}) \\
+		Learning model & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell)$\hspace{0.3cm} (\S \ref{sec:ppl}) \\
 	\end{tabular}
-	\caption{Overview of known results and the results provided in this paper. By $\Rep$ we denote the cost of a single component repartition, and we elaborate on this in Section~\ref{sec:k3}
+	\caption{Overview of known results and the results provided in this paper.
+By $\Rep$ we denote the cost of a single component repartition, and we elaborate on this in Section~\ref{sec:k3}.
+The results pertain to the general partitioning model, but the last row, that pertains to the learning model for arbitrary $k$ and $\ell$.
 	}
@@
 
-\maciek{Eviction - introduce.}
+\maciek{Eviction - introduce the concept.}
 
@@
 
-\maciek{TODO: Intro to the learning problem is missing. What is the problem definition. How is it different from the general problem. Emphasize the fact that OPT is constrained not to migrate.}
-
+In this section we consider a variant of Online Balanced Partitioning, where requests are permanent: once a pair of nodes start communicating, they continue forever.
+Each communicating pair must be collocated, as otherwise the cost would be infinite.
+We study the instances with finite cost: where the communication graph can be perfectly partitioned in the clusters.
+
+For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$.
+The lower bound can be easily translated to the partitioning model (Section~\ref{sec:part}) with finite request sequence.
 At the end of this section, we discuss an asymptotically optimal upper bound.
@@
 
+We conclude by emphasizing the generality of the lower bound.
+
 \noindent
-\textbf{Remark on resource augmentation.}
+\textbf{General partitioning problem.}
+Every lower bound for the learning problem constitutes a valid lower bound for the partitioning problem.
+The bound for partitioning converges to the bound for learning as the input sequence grows.
+
+Our construction allows for an exact lower bound for partitioning with finite input sequence.
+To this end, we issue only enough requests until the algorithm collocates the nodes of each ground set.
+If it never does, we repeat until it pays $k\cdot \ell$.
+
+\noindent
+\textbf{Resource augmentation.}
 The majority of work on Online Balanced Partitioning so far (todo: cite) focuses on the scenario with resource augmentation.
 We note that this lower bound provides insights into this scenario as well.
+
 Consider an instance with $k$ divisible by $3$, and construct $3$ ground sets of size $k/3$ in each cluster.
@@
 
-\section{The Online Problem}
-\label{sec:k3}
-
-\maciek{Not precise name? Both the learning problem and the general one is online.}
+\section{The Partitioning Problem}
+\label{sec:part}
 
@@
 model where the request sequence
-can be arbitrary. We show that the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental} allows obtaining an optimal algorithm for $k=3$.
+can be arbitrary.
+We show that the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental} allows obtaining an optimal algorithm for $k=3$.
 Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one cluster until the end of a phase (that we define precisely later).
@@
 \subsection{Improved analysis of the component-based algorithm}
+\label{sec:k3}
 

### 2020-05-02 paper/disc20clustering.tex
PPL issues

@@
 \section{Preliminaries.}
+\label{sec:prelim}
 
@@
 In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
-It is oblivious to the choice of the reconfiguration cost $\alpha$, and we compare the number of performed node exchanges.
+The construction is oblivious to the choice of the reconfiguration cost $\alpha$: we compare the number of node exchanges.
 The lower bound requires $k\geq 3$.
-Note that for $k=2$, a lower bound of $3$ is known \cite{repartition-disc}.
+For $k=2$, a lower bound of $3$ is known \cite{repartition-disc}, and we provide a $6$-competitive algorithm (see Section~\ref{sec:k2}).
 
@@
 The adversary constructs ground sets in reaction to choices of a deterministic online algorithm.
-The algorithm does not know the ground set before they are revealed, and it may split some of them while evicting nodes.
-
+We emphasize that the algorithm does not know the ground set before they are revealed.
 Once we construct a ground set, it lasts until the end of the input sequence.
+
+The constructed ground sets can be partitioned into our clusters (without splitting nodes).
+This way, the optimal algorithm moves to such a partition, and serves all requests for free.
 If the algorithm splits any pair of nodes of a ground set, the adversary issue requests to any pair of split nodes.
-
-We construct the ground sets in a way that they always form a balanced partition of nodes.
-This way, the optimal algorithm moves to such a configuration, and serves all requests for free.\maciek{i can do better: such a?}
-(This implies that the algorithm must eventually collocate nodes of a ground set to be competitive.)
+These two observations imply that the algorithm must eventually collocate nodes of a ground set to be competitive.
 
@@
 \label{sec:ppl}
-Now we give an algorithm for a restricted variant of  \OBRP{}, named \PPOBRP{},
-%\maciek{This is not sufficient! We also disallow serving requests remotely.}
-where an optimal offline algorithm ($\OPT$) moves to a perfect partition
-at the beginning and stays there perpetually.
-\maciek{Already mention that the LB is for that scenario}
-%We refer this variant as \PPOBRP{}.
-The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect partition while not paying too much relative to \OPT.
-%
-%\maciek{This is not a valid description of the model.
-We need the assumption that collocation is needed in our model.
-Otherwise OPT would just serve some requests remotely, and our analysis would not hold.}
-The first request between any two non-collocated nodes reveals an edge of the communication graph.
-Our online algorithm collocates them immediately and never separates them since by assumption \OPT does the same.
-Revealed edges, up to any point in the sequence,
-induce some connected subgraphs which grow as components \maciek{rewrite sentence, expand}.
-%An algorithm that always  keeps nodes belonging to the same component collocated is \emph{component respecting}.
+
+We present an asymptotically optimal algorithm for the learning problem.
+The algorithm immediately collocates every communicating pair and never separates them.
+In other words, it maintains components, introduced in the preliminaries (Section~\ref{sec:prelim}).
+
+Upon receiving any external request, two components merge --- and this requires a reconfiguration.
+Many configurations may be feasible, and we choose the one that is \emph{closest to the initial configuration}.
 
@@
 
-We assume \OPT begins with the initial configuration
-$P_I = I_1, \dots, I_{\ell}$ and moves to the final partitioning
-$P_F = F_1, \dots, F_{\ell}$.
+\noindent
+\textbf{Perfect Partition Learner algorithm.}
+Fix the initial configuration
+$P_I = I_1, \dots, I_{\ell}$ and \OPT's final configuration
+$P_F = F_1, \dots, F_{\ell}$\maciek{How do we define a configuration?}.
 The \emph{distance} of a configuration $P = C_1, \dots, C_{\ell}$ from the initial configuration \maciek{, denoted $\dist(P)$} is the number of nodes in $P$ that do not reside in their initial cluster.
@@
 That is,
-$\dist(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$. 
+$\dist(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$
+\maciek{Major: what does this definition mean? This is highly unclear, most likely because a configuration was not defined.}. 
 In other words,
@@
 As a result,
-\PPL never ends up in a configuration that is more than $\Delta$ (single-node) migrations away from $P_I$.
-This invariant ensures that \PPL does not pay too much while recovering $P_F$.
+\PPL never ends up \maciek{ends up is informal} in a configuration that is more than $\Delta$ (single-node) migrations away from $P_I$ \maciek{This is not an algorithm definition, that's its property.
+Should not be in this paragraph}.
+This invariant ensures that \PPL does not pay too much while recovering $P_F$ \maciek{Recovering $P_F$?
+It's not the goal.
+At least at this point the reader is confused: it does not know yet that PPL is proven to end up in $P_F$}.
 %      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
-We note that the $\mathit{repartition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a~perfect partition closest to $P_I$.
-Hence it never moves to a configuration beyond distance $\Delta$.
+We note \maciek{Define?} that the $\mathit{repartition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a~perfect partition closest to $P_I$.
+Hence it never moves to a configuration beyond distance $\Delta$ \maciek{This is again mixing the properties of ppl and its definition. Restructure}.
 The scheme of the algorithm can be found in the Algorithm \ref{alg:PPL}.
@@
 
+\maciek{Algorithm pseudocode comments}
+\maciek{This is not good that we need a comment about what $cluster \neq cluster$ means}
+\maciek{In pseudocode we need extremely short sentences. $C := singleton components ... $?; $Merge C, C'$ etc}
+\maciek{No need to maintain $P_i$'s. Just use one $P$? Or without naming: "move to partition (closeToInit(...))", and remove the partition args from cluster().}
+\maciek{Online algorithms should have different pseudocode. Instead of for loop, start with "When a request u,v arrives" (before that it is ok to have init phase just like it is now)}
+
 \begin{property} \label{prop:dist<OPT}
@@
 	Then, $\dist(P,P_I) \leq \Delta$.
+	\maciek{Express the configuration without refering to pseudocode?}
 \end{property}
-\maciek{The property is explained in the previous paragraph. Here, reader thinks that this is obvious and wonders why. Move closer to the exlaination text.}
+\maciek{The property is explained in the previous paragraph.
+Here, reader thinks that this is obvious and wonders why.
+Restructure: Move closer to the exlaination text.
+Or even make this property a lemma (most likely better).}
 
@@
 	The cost of repartitioning at Line \ref{line:rebalance} is at most $2\cdot\OPT$.
+	\maciek{Express the reconfiguration without pseudocode?}
 \end{lemma}

### 2020-05-03 paper/disc20clustering.tex
Comments
Related work update

@@
 so far was $O(\ell^2)$~\cite{repartition-disc}.
-Then, we present a strictly $6$-competitive algorithm for $k=2$.
+Then, we present a strictly $6$-competitive algorithm for $k=2$ that improves upon previous $7$-competitive algorithm with $O(\ell^2)$ additive constant..
 In addition, we show a new, improved analysis of the algorithm from \cite{repartition-disc} that depends on the solutions to an interesting combinatorial problem (see Section~\ref{ssec:cascade}, The Cascade Problem).
@@
 
-\noindent \textbf{Related work.}
+\section{Related work}
+
+In the paper that originated a study on BRP \cite{repartition-disc} (DISC 2016), authors provided a $O(k^2 \cdot \ell^2)$-competitive algorithm, and a lower bound $\Omega(k)$.
+For $k=2$, they presented a $7$-competitive algorithm~\cite{repartition-disc} with substantial ($\Omega(\ell^2)$) additive constant.
+
 The existing results in~\cite{repartition-disc}
@@
 requests from a fixed distribution~\cite{stochastic-ring}.
-For clusters of size $2$ (i.e., $k=2$), 
-it is known that a $7$-competitive algorithm exists~\cite{repartition-disc} with large ($\Omega(\ell^2)$) additive constant.
 
@@
 
+\maciek{The big inequality is impressive, but not a lot change in each row. Instead, bound the cost of ALG separately in one line, and the cost of OPT in another. (possibly far away from eachother)}
+
 \maciek{Remove all "charge both". Just sum up costs. I would say not use "charge" at all, but it's up to you.}

### 2020-05-03 paper/disc20clustering.tex
Rewrite

@@
 Across this paper, we often refer to groups of communicating nodes.
-As this concept is used differently in the lower bound and the upper bounds, we name them differently to avoid confusion.
+We use this concept slightly differently in the lower bound and the upper bounds.
+In our algorithms, we group communicating nodes into \emph{components}, and in the lower bound, we group nodes that may communicate into \emph{ground sets}.
 

### 2020-05-03 paper/disc20clustering.tex
Restructure intro

@@
 
-\noindent \textbf{Motivation and model.}
+\maciek{Practical motivation}
+
+\subsection{Model.}
+
+\maciek{We study two models. One is a subproblem.}
+
+The general model, where $\sigma$ can be arbitrary
+was the original problem introduced
+by Avin et al.~\cite{repartition-disc} at DISC 2016.
+For simplicity, we will refer to the Perfect Partition
+model as the \emph{learning} model (as
+one has to learn the static communication graph) 
+and to the original as the \emph{repartitioning} model (as one continuously adjusts the partition in response to changing communication graph).
+
+\noindent
+\textbf{Partitioning model.}
 The \emph{balanced repartitioning} problem (\OBRP{})
@@
 
+\noindent
+\textbf{Learning model.}
 In addition to the general model, we study a simplified Perfect Partition model (also studied by Henzinger et al.~\cite{sigmetrics19_partitioning} (SIGMETRICS 2019)) that we consider for a stronger lower bound.
@@
 
-The general model, where $\sigma$ can be arbitrary
-was the original problem introduced
-by Avin et al.~\cite{repartition-disc} at DISC 2016.
-For simplicity, we will refer to the Perfect Partition
-model as the \emph{learning} model (as
-one has to learn the static communication graph) 
-and to the original as the \emph{repartitioning} model (as one continuously adjusts the partition in response to changing communication graph).
-
-\noindent \textbf{Contributions.}
+\subsection{Related work}
+
+In the paper that originated a study on BRP \cite{repartition-disc} (DISC 2016), authors provided a $O(k^2 \cdot \ell^2)$-competitive algorithm, and a lower bound $\Omega(k)$.
+For $k=2$, they presented a $7$-competitive algorithm~\cite{repartition-disc} with substantial ($\Omega(\ell^2)$) additive constant.
+
+The existing results in~\cite{repartition-disc}
+and~\cite{sigmetrics19_partitioning}
+primarily focus on a model where 
+the online algorithm can use augmentation,
+i.e., has slightly larger clusters than the offline 
+algorithm. In contrast, in our paper, we focus
+on a scenario where the nodes need to be
+perfectly balanced among the clusters.
+The problem has also been studied in a weaker
+model where the adversary can only sample
+requests from a fixed distribution~\cite{stochastic-ring}.
+
+\maciek{Offline results}
+\maciek{Links to paging}
+
+%More generally, the model is related to online
+%caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00},
+%see~\cite{repartition-disc} for a discussion.
+%The static offline version of~the~problem, called the
+%\emph{$\ell$-balanced graph partitioning problem} is 
+%NP-complete, and cannot even be approximated within %any finite factor unless P
+%= NP~\cite{AndRae06}.
+
+\subsection{Our Contributions.}
 We provide an improved lower bound 
@@
 
-\section{Related work}
-
-In the paper that originated a study on BRP \cite{repartition-disc} (DISC 2016), authors provided a $O(k^2 \cdot \ell^2)$-competitive algorithm, and a lower bound $\Omega(k)$.
-For $k=2$, they presented a $7$-competitive algorithm~\cite{repartition-disc} with substantial ($\Omega(\ell^2)$) additive constant.
-
-The existing results in~\cite{repartition-disc}
-and~\cite{sigmetrics19_partitioning}
-primarily focus on a model where 
-the online algorithm can use augmentation,
-i.e., has slightly larger clusters than the offline 
-algorithm. In contrast, in our paper, we focus
-on a scenario where the nodes need to be
-perfectly balanced among the clusters.
-The problem has also been studied in a weaker
-model where the adversary can only sample
-requests from a fixed distribution~\cite{stochastic-ring}.
-
-%More generally, the model is related to online
-%caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00},
-%see~\cite{repartition-disc} for a discussion.
-%The static offline version of~the~problem, called the
-%\emph{$\ell$-balanced graph partitioning problem} is 
-%NP-complete, and cannot even be approximated within %any finite factor unless P
-%= NP~\cite{AndRae06}.
-
 \section{Preliminaries.}
@@
 We use this concept slightly differently in the lower bound and the upper bounds.
-In our algorithms, we group communicating nodes into \emph{components}, and in the lower bound, we group nodes that may communicate into \emph{ground sets}.
+In our algorithms, we group nodes with a communication history into \emph{components}, and in the lower bound, we group nodes that may communicate into \emph{ground sets}.
 
@@
 \textbf{Ground sets.}
-In the lower bound, a~\emph{ground set} is a group of nodes that start communicating if an algorithm splits them (places at different clusters).
-Their role is to ensure that the algorithm maintains a partition of ground sets.
+In the lower bound, a~\emph{ground set} is a group of nodes that repeatedly communicate when an algorithm splits them (places at different clusters).
+Their role is to ensure that the algorithm maintains a partition of ground sets into clusters.
 We emphasize that ground sets are revealed only upon their nodes are split.
@@
 
-In this section we consider a variant of Online Balanced Partitioning, where requests are permanent: once a pair of nodes start communicating, they continue forever.
-Each communicating pair must be collocated, as otherwise the cost would be infinite.
-We study the instances with finite cost: where the communication graph can be perfectly partitioned in the clusters.
+In this section we consider a variant of Online Balanced Partitioning, where communication requests are permanent: once a pair of nodes start communicating, they continue forever.
+To avoid infinite cost, each communicating pair must be collocated.
+We study the instances with finite cost: we assume that the communication graph can be perfectly partitioned in the clusters.
+The algorithm's goal is to \emph{learn} the communication graph's partition.
 

### 2020-05-03 paper/disc20clustering.tex
The dead end lower bound (but other interesting comments, too)

@@
 
-\noindent
-\textbf{Partitioning model.}
+\subsubsection{Partitioning model.}
 The \emph{balanced repartitioning} problem (\OBRP{})
@@
 
-\noindent
-\textbf{Learning model.}
+\maciek{The competitive ratio definition and the worst case inputs}
+
+\maciek{Ilustrate the challenges. Rent-or-buy. Decision what to evict}
+
+\subsubsection{Learning model.}
 In addition to the general model, we study a simplified Perfect Partition model (also studied by Henzinger et al.~\cite{sigmetrics19_partitioning} (SIGMETRICS 2019)) that we consider for a stronger lower bound.
@@
 
+The existing results in~\cite{repartition-disc} and~\cite{sigmetrics19_partitioning} primarily focus on a model with resource augmentation, where  the online algorithm has slightly larger clusters than the offline algorithm.
+They showed that their lower bound $\Omega(k)$ holds even for a significant resource augmentation.
+They provided an algorithm with the competitive ratio independent of $\ell$ using the $(2+\epsilon)$-augmented cluster capacity.
+
+In contrast, we study the non-augmented setting, where the nodes need to be perfectly balanced  among the clusters.
 In the paper that originated a study on BRP \cite{repartition-disc} (DISC 2016), authors provided a $O(k^2 \cdot \ell^2)$-competitive algorithm, and a lower bound $\Omega(k)$.
-For $k=2$, they presented a $7$-competitive algorithm~\cite{repartition-disc} with substantial ($\Omega(\ell^2)$) additive constant.
-
-The existing results in~\cite{repartition-disc}
-and~\cite{sigmetrics19_partitioning}
-primarily focus on a model where 
-the online algorithm can use augmentation,
-i.e., has slightly larger clusters than the offline 
-algorithm. In contrast, in our paper, we focus
-on a scenario where the nodes need to be
-perfectly balanced among the clusters.
+For $k=2$, they presented a $7$-competitive algorithm~\cite{repartition-disc} with a~substantial ($\Omega(\ell^2)$) additive constant.
+
 The problem has also been studied in a weaker
@@
 \maciek{Offline results}
-\maciek{Links to paging}
+\maciek{Links other problems: for $k=2$ a flavor of online matching, for $\ell = 2$ it generalizes paging}
 
@@
 We emphasize that ground sets are revealed only upon their nodes are split.
-\maciek{Other names: division, segmentation, alignment, formation}
+%\maciek{Other names: division, segmentation, alignment, formation}
 
@@
 
-\maciek{Eviction - introduce the concept.}
-
 \section{The Learning Problem} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
@@
 
-First, we construct a ground set of size $k-1$ on one of the clusters.
+\maciek{Better name than $S_1$? Here and in the proof}
+First, we construct a ground set of size $k-1$ on one of the clusters $S_1$.
 In any configuration, there exists a node that accompanies it.
-We only issue requests between such accompanying (time-varying) node, and any node that was initially collocated with it.
-Repeating such requests, almost every node must visit this cluster.
+We only issue requests between such accompanying (time-varying) node, and any alone node that was initially collocated with it.
+Repeating such requests, almost half of all node must visit the cluster $S_1$.
 In comparison, \OPT performs one migration.
@@
 	\label{th:lowerbound}
-	The competitive ratio of any deterministic online algorithm for \OBRP{} is  $(k \cdot \ell - k - \ell) / 2 = \Omega(k\cdot \ell)$ for any $k\geq 3$ and for any $\ell \geq 2$.
+	The competitive ratio of any deterministic online algorithm for \OBRP{} is  $\Omega(k\cdot \ell)$ for any $k\geq 3$ and for any $\ell \geq 2$.
 \end{theorem}
@@
 \begin{proof}
-	\maciek{Better name than $S_1$?}
 	Fix any online algorithm \ALG{}.
-	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
-	\maciek{A problem with the definition! The first pair does not have the same initial location of its nodes.}
-	Initially all nodes are single (they belong to their own trivial ground sets).
+	Assume that $k$ is odd.
+	We show how to proceed with even $k$ later in this proof.
+	Let $I(z)$ denote the cluster where the node $z$ was located initially.
+	Throughout the construction, we maintain ground sets for each node.
+	Initially all nodes are \emph{alone} (they belong to their own trivial ground sets).
 
@@
 
-	We construct a ground set containing this alone node (and some other node).
+	We construct a ground set containing this alone node from $S_1$ to force its eviction.
 	Precisely, let $x_0$  be the alone node from $S_1$ and let $y_0$ be an arbitrary node that does not reside in $S_1$.
-	First,
-	we construct a ground set $\{x_0, y_0\}$.
-	Since \ALG{} has them split, the adversary produces an external request.
-	Since $S_1$ contains a ground set of size $k$, \ALG{} must collocate $\{x_0, y_0\}$ in another cluster.
-	To preserve a feasible partition of nodes into clusters,
-	\ALG must replace $x_0$ with another single node $x_1$.
+	We construct a ground set $\{x_0, y_0\}$.
+	\ALG has them split, and the adversary produces an external request.
+	$S_1$ already contains a ground set of size $k-1$, hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.
 
 	We continue to construct ground sets containing the current alone node from $S_1$.
-	Precisely, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t.~$I(C) = I(x_1)$.
-	To preserve a feasible partition of nodes into clusters,
-	\ALG{} must collocate $x_1$ and $C$ in a~cluster other than $S_1$ and
-	consequently replace $x_1$ with another alone node $x_2$.
-
-	We end this procedure prematurely (before all nodes visit $S_1$) to keep \OPT's cost low.
-	We repeat merging the component of single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
-	until there are only two single nodes left that  originate from the same cluster.
-	Formally, we denote these two single nodes by $x^*, y^*$, and we have $I(x^*) = I(y^*)$, and~for any other pair of single nodes
-	$x'$ and $y'$,
-	we have $I(x') \neq I(y')$.
-	Note that at this point there are at most $\ell+1$ single nodes left,
-	otherwise, there would be more pairs of single nodes that were initially in the same cluster.
-
-	Finally, we bound the cost of \OPT.
-	Given this sequence of requests,
-	\OPT migrates $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
-	swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
-	Hence,
-	OPT pays for $2$ node swaps and
-	\ALG{} incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
+	To preserve a feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
+	\ALG must replace $x_0$ with another single node (that we name $x_1$).
+	Then, we merge the ground sets of $x_1$ and an arbitrary node $y_1$ that was initially collocated with $x_1$.
+	Continuing this, we observe a sequence $X := x_0, x_1, x_2\dots$ of alone nodes that visit $S_1$.
+	(Note that subsequent ground sets are different than the first ground set $\{x_0, y_0\}$, whose nodes were initially in different clusters.)
+
+	We claim that this procedure can continue as long as a perfect partition of ground sets still exists.
+	%and we construct a sequence of ground sets $X := \langle \{x_0, y_0\}, \{x_1, y_2\}, \{x_2, y_2\}, \ldots \rangle$ of length $\Omega(k \cdot \ell)$.
+	Consider the moment of creating a ground set $\{x_i, y_i\}$.
+	If a perfect partition of ground sets still exists, then
+	\ALG{} must collocate $\{x_i,y_i\}$ in a~cluster other than $S_1$.
+	To this end, it replaces $x_i$ with another alone node $x_{i+1}$.
+	To continue, we must guarantee the existence of an alone node that was initially collocated with $x_{i+1}$.
+
+	We end this procedure prematurely to keep \OPT's cost low.
+	The sequence must end when no perfect partition of ground sets exists ---
+	however, we end it when \emph{two more} ground sets are still possible to give.
+	Precisely, we repeat constructing ground sets,
+	until there are only two alone nodes left that originate from the same cluster.
+	We denote these two nodes by $x^*, y^*$.
+	Observe that at this point there are at most $\ell+1$ alone nodes left:
+	otherwise, by pigeon hole principle, there would be more pairs of alone nodes that were initially in the same cluster.
+
+	Now we lower bound the length of the sequence $X := x_0, x_1,\dots$.
+
+	The optimal strategy is to perform two node swaps.
+	Precisely, \OPT collocates $\set{x_0,y_0}$ by swapping them with $x^*$ and $y^*$.
+	In this configuration, no ground set is split.
+	We never issue requests between nodes $x^*, y^*$ that are split in this configuration.
+
+	Finally, we bound the cost of \ALG.
+	\ALG incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
+
 	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
@@
 
-\section{Improved algorithm for $k=2$}
+\section{Improved algorithm for Online Rematching}
 \label{sec:k2}
@@
 	The algorithm is (strictly) 6-competitive.
+	\maciek{for what lambda?}
 \end{theorem}

### 2020-05-03 paper/disc20clustering.tex
Bringing back the old LB

@@
 \begin{proof}
+	\maciek{Better name than $S_1$?}
+	Fix any online algorithm \ALG{}.
+	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
+	\maciek{A problem with the definition! The first pair does not have the same initial location of its nodes.}
+	Initially all nodes are single (they belong to their own trivial ground sets).
+
+	We distinguish a cluster $S_1$, where we form a ground set of size $k-1$.
+	We never dismiss any constructed ground set.
+	Each cluster host exactly $k$ nodes, hence in any configuration of \ALG, there exists a single alone node in $S_1$.
+
+	We construct a ground set containing this alone node (and some other node).
+	Precisely, let $x_0$  be the alone node from $S_1$ and let $y_0$ be an arbitrary node that does not reside in $S_1$.
+	First,
+	we construct a ground set $\{x_0, y_0\}$.
+	Since \ALG{} has them split, the adversary produces an external request.
+	Since $S_1$ contains a ground set of size $k$, \ALG{} must collocate $\{x_0, y_0\}$ in another cluster.
+	To preserve a feasible partition of nodes into clusters,
+	\ALG must replace $x_0$ with another single node $x_1$.
+
+	We continue to construct ground sets containing the current alone node from $S_1$.
+	Precisely, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t.~$I(C) = I(x_1)$.
+	To preserve a feasible partition of nodes into clusters,
+	\ALG{} must collocate $x_1$ and $C$ in a~cluster other than $S_1$ and
+	consequently replace $x_1$ with another alone node $x_2$.
+
+	We end this procedure prematurely (before all nodes visit $S_1$) to keep \OPT's cost low.
+	We repeat merging the component of single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
+	until there are only two single nodes left that  originate from the same cluster.
+	Formally, we denote these two single nodes by $x^*, y^*$, and we have $I(x^*) = I(y^*)$, and~for any other pair of single nodes
+	$x'$ and $y'$,
+	we have $I(x') \neq I(y')$.
+	Note that at this point there are at most $\ell+1$ single nodes left,
+	otherwise, there would be more pairs of single nodes that were initially in the same cluster.
+
+	Finally, we bound the cost of \OPT.
+	Given this sequence of requests,
+	\OPT migrates $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
+	swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
+	Hence,
+	OPT pays for $2$ node swaps and
+	\ALG{} incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
+	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
+	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate. Don't use the word "exclude" unless very obvious. Just sum up the real cost.}.
+	Finally, $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
+
+	\maciek{BELOW OLD VERSION. TODO: INCORPORATE BETTER SENTENCES}
+
 	Fix any online algorithm \ALG{}.

### 2020-05-03 paper/disc20clustering.tex
A great fix!
Definition of I(C) as a partial function

@@
 	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
-	\maciek{A problem with the definition! The first pair does not have the same initial location of its nodes.}
+	If $C$ contains nodes that originate in different clusters, $I(C)$ is undefined.
 	Initially all nodes are single (they belong to their own trivial ground sets).
@@
 	We continue to construct ground sets containing the current alone node from $S_1$.
-	Precisely, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t.~$I(C) = I(x_1)$.
+	Precisely, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t. and $C \neq \{x_0,y_0\}$~$I(C) = I(x_1)$.
 	To preserve a feasible partition of nodes into clusters,
@@
 	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
-	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate. Don't use the word "exclude" unless very obvious. Just sum up the real cost.}.
+	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate on exclude. Just sum up the real cost.}.
 	Finally, $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
@@
 	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
-	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate. Don't use the word "exclude" unless very obvious. Just sum up the real cost.}.
+	therefore $|X| \geq k \cdot \ell - k - \ell$.
 	Finally, $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.

### 2020-05-04 paper/disc20clustering.tex
Lower bound writing improvements

--- 2020-05-03 paper/disc20clustering.tex
@@
 \begin{proof}
+	\maciek{Better name than $S_1$?}
 	Fix any online algorithm \ALG{}.
-	Assume that $k$ is odd.
-	We show how to proceed with even $k$ later in this proof.
-	Let $I(z)$ denote the cluster where the node $z$ was located initially.
+	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
+	If $C$ contains nodes that originate from different clusters, $I(C)$ is undefined.
 	Throughout the construction, we maintain ground sets for each node.
@@
 
-	We construct a ground set containing this alone node from $S_1$ to force its eviction.
+	We construct a ground set containing this alone node (and some other node).
 	Precisely, let $x_0$  be the alone node from $S_1$ and let $y_0$ be an arbitrary node that does not reside in $S_1$.
 	We construct a ground set $\{x_0, y_0\}$.
-	\ALG has them split, and the adversary produces an external request.
-	$S_1$ already contains a ground set of size $k-1$, hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.
+	\ALG has them split, and the adversary produces external requests until these are collocated.
+	$S_1$ already contains a ground set of size $k-1$ (that it cannot split), hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.
 
@@
 	\ALG must replace $x_0$ with another single node (that we name $x_1$).
-	Then, we merge the ground sets of $x_1$ and an arbitrary node $y_1$ that was initially collocated with $x_1$.
+	Then, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t. and $C \neq \{x_0,y_0\}$~$I(C) = I(x_1)$.
 	Continuing this, we observe a sequence $X := x_0, x_1, x_2\dots$ of alone nodes that visit $S_1$.
@@
 
-	We claim that this procedure can continue as long as a perfect partition of ground sets still exists.
+	We claim that this procedure (of constructing ground sets) can continue as long as a perfect partition of ground sets still exists.
 	%and we construct a sequence of ground sets $X := \langle \{x_0, y_0\}, \{x_1, y_2\}, \{x_2, y_2\}, \ldots \rangle$ of length $\Omega(k \cdot \ell)$.
@@
 	To continue, we must guarantee the existence of an alone node that was initially collocated with $x_{i+1}$.
-
-	We end this procedure prematurely to keep \OPT's cost low.
+	\maciek{elaborate more}
+
+	We end this procedure prematurely (before all nodes visit $S_1$) to keep \OPT's cost low.
 	The sequence must end when no perfect partition of ground sets exists ---
@@
 
-	Now we lower bound the length of the sequence $X := x_0, x_1,\dots$.
+	\ALG incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
+	We lower bound the length of the sequence $X := x_0, x_1,\dots$.
+	\maciek{TODO}
+	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
+	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate on exclude. Just sum up the real cost.}.
 
@@
 
-	Finally, we bound the cost of \ALG.
-	\ALG incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
-
-	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
-	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate. Don't use the word "exclude" unless very obvious. Just sum up the real cost.}.
-	Finally, $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
+	The competitive ratio is then: $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
 \end{proof}

### 2020-05-04 paper/disc20clustering.tex
Lower bound todos

@@
 
-\maciek{Better name than $S_1$? Here and in the proof}
-First, we construct a ground set of size $k-1$ on one of the clusters $S_1$.
+First, we construct a ground set of size $k-1$ on one of the clusters $S$.
 In any configuration, there exists a node that accompanies it.
@@
 \begin{proof}
-	\maciek{Better name than $S_1$?}
 	Fix any online algorithm \ALG{}.
@@
 
-	We distinguish a cluster $S_1$, where we form a ground set of size $k-1$.
+	We distinguish a cluster $S$, where we form a ground set of size $k-1$.
 	We never dismiss any constructed ground set.
-	Each cluster host exactly $k$ nodes, hence in any configuration of \ALG, there exists a single alone node in $S_1$.
+	Each cluster host exactly $k$ nodes, hence in any configuration of \ALG, there exists a single alone node in $S$.
 
 	We construct a ground set containing this alone node (and some other node).
-	Precisely, let $x_0$  be the alone node from $S_1$ and let $y_0$ be an arbitrary node that does not reside in $S_1$.
+	Precisely, let $x_0$  be the alone node from $S_1$ and let $y_0$ be an arbitrary node that does not reside in $S$.
 	We construct a ground set $\{x_0, y_0\}$.
@@
 
-	We continue to construct ground sets containing the current alone node from $S_1$.
+	We continue to construct ground sets containing the current alone node from $S$.
 	To preserve a feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
@@
 	Then, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t. and $C \neq \{x_0,y_0\}$~$I(C) = I(x_1)$.
-	Continuing this, we observe a sequence $X := x_0, x_1, x_2\dots$ of alone nodes that visit $S_1$.
+	Continuing this, we observe a sequence $X := x_0, x_1, x_2\dots$ of alone nodes that visit $S$.
 	(Note that subsequent ground sets are different than the first ground set $\{x_0, y_0\}$, whose nodes were initially in different clusters.)
@@
 	If a perfect partition of ground sets still exists, then
-	\ALG{} must collocate $\{x_i,y_i\}$ in a~cluster other than $S_1$.
+	\ALG{} must collocate $\{x_i,y_i\}$ in a~cluster other than $S$.
 	To this end, it replaces $x_i$ with another alone node $x_{i+1}$.
-	To continue, we must guarantee the existence of an alone node that was initially collocated with $x_{i+1}$.
-	\maciek{elaborate more}
-
-	We end this procedure prematurely (before all nodes visit $S_1$) to keep \OPT's cost low.
+	To continue, we must guarantee that
+	\maciek{elaborate}
+
+	We end this procedure prematurely (before all nodes visit $S$) to keep \OPT's cost low.
 	The sequence must end when no perfect partition of ground sets exists ---

### 2020-05-04 paper/disc20clustering.tex
Comments

@@
 \maciek{Algorithm pseudocode comments}
+\maciek{Major thing to discuss: do we even need this?
+The algorithm is very simple, and the description in the paragraph would be much better.
+The only way to save it would be to have a general algorithm pseudocode (equivalent to DET), and parametrize it with the repartition procedure.
+Here, we would say that we choose the closest to initial, and it ALG3 we say that we choose the closest to current.}
 \maciek{This is not good that we need a comment about what $cluster \neq cluster$ means}

### 2020-05-05 paper/disc20clustering.tex
Missing parts of LB

@@
 	\label{th:lowerbound}
-	The competitive ratio of any deterministic online algorithm for \OBRP{} is  $\Omega(k\cdot \ell)$ for any $k\geq 3$ and for any $\ell \geq 2$.
+	The competitive ratio of any deterministic online algorithm for \OBRP{} is  $(k\cdot \ell - k - \ell)/2$ for any $k\geq 3$ and for any $\ell \geq 2$.
 \end{theorem}
@@
 
-	We distinguish a cluster $S$, where we form a ground set of size $k-1$.
-	We never dismiss any constructed ground set.
+	First, where we form a ground set of size $k-1$ on an arbitrarily chosen cluster $S$.
+	%We never dismiss any constructed ground set.
 	Each cluster host exactly $k$ nodes, hence in any configuration of \ALG, there exists a single alone node in $S$.
 
-	We construct a ground set containing this alone node (and some other node).
-	Precisely, let $x_0$  be the alone node from $S_1$ and let $y_0$ be an arbitrary node that does not reside in $S$.
-	We construct a ground set $\{x_0, y_0\}$.
+	Then, we construct a ground set containing this alone node (and some other node).
+	Precisely,
+	we construct a ground set $\{x_0, y_0\}$, 
+	where $x_0$ is the alone node from $S$ and $y_0$ is an arbitrary node that does not reside in $S$.
 	\ALG has them split, and the adversary produces external requests until these are collocated.
-	$S_1$ already contains a ground set of size $k-1$ (that it cannot split), hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.
-
-	We continue to construct ground sets containing the current alone node from $S$.
+	$S$ already contains a ground set of size $k-1$ (that it cannot split), hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.
+
+	We continue to construct ground sets containing the node that is currently alone in $S$.
 	To preserve a feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
 	\ALG must replace $x_0$ with another single node (that we name $x_1$).
-	Then, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t. and $C \neq \{x_0,y_0\}$~$I(C) = I(x_1)$.
-	Continuing this, we observe a sequence $X := x_0, x_1, x_2\dots$ of alone nodes that visit $S$.
+	Consequently, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x_1)$.
+	Repeating this, we observe a sequence $X := x_0, x_1, x_2\dots$ of alone nodes that visit $S$.
 	(Note that subsequent ground sets are different than the first ground set $\{x_0, y_0\}$, whose nodes were initially in different clusters.)
 
-	We claim that this procedure (of constructing ground sets) can continue as long as a perfect partition of ground sets still exists.
-	%and we construct a sequence of ground sets $X := \langle \{x_0, y_0\}, \{x_1, y_2\}, \{x_2, y_2\}, \ldots \rangle$ of length $\Omega(k \cdot \ell)$.
-	Consider the moment of creating a ground set $\{x_i, y_i\}$.
-	If a perfect partition of ground sets still exists, then
-	\ALG{} must collocate $\{x_i,y_i\}$ in a~cluster other than $S$.
-	To this end, it replaces $x_i$ with another alone node $x_{i+1}$.
-	To continue, we must guarantee that
-	\maciek{elaborate}
-
-	We end this procedure prematurely (before all nodes visit $S$) to keep \OPT's cost low.
-	The sequence must end when no perfect partition of ground sets exists ---
-	however, we end it when \emph{two more} ground sets are still possible to give.
-	Precisely, we repeat constructing ground sets,
-	until there are only two alone nodes left that originate from the same cluster.
+	We claim that constructing such ground sets can continue as long as at least $\ell$ alone node exists.
+	Fix the moment just before the creation of a ground set with the node $x_i$.
+	We show that if at least $\ell$ alone nodes exists (including $x_i, y_i$), then there exists a perfect partition of ground sets.
+
+	To show this claim, we consider a configuration, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
+	In this configuration, each cluster contains:  at most one ground set larger than $1$, possibly some alone nodes, and possibly the ground set $\{x_0, y_0\}$.
+	If $I(x_i)$ contains an alone node, then we simply obtain a perfect partition by swapping it with $x_i$.
+	Otherwise, $\{x_0, y_0\}$ is present on $I(x_i)$ (a single ground set cannot occupy $I(x_i)$, as without $x_i$ its size is at most $k-1$).
+	In this case, we first swap $\{x_0, y_0\}$ with two alone nodes that occupy the same cluster.
+	To see that such two nodes must exist, observe that we have at most $\ell-1$ clusters with alone nodes, and by the assumption we have $\ell$ alone nodes.
+	After this swap, we migrate $x_i$ to $I(x_i)$, evicting a lone node to $S$, and we conclude that the perfect partition exists.
+
+	To keep \OPT's cost low, we end constructing ground sets prematurely --- while there are still $\ell+1$ alone nodes (i.e., not performing two last constructions).
+	Then, there two alone nodes exists that originate from the same cluster (we have $\ell+1$ alone nodes and $\ell$ clusters).
 	We denote these two nodes by $x^*, y^*$.
-	Observe that at this point there are at most $\ell+1$ alone nodes left:
-	otherwise, by pigeon hole principle, there would be more pairs of alone nodes that were initially in the same cluster.
-
-	\ALG incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
-	We lower bound the length of the sequence $X := x_0, x_1,\dots$.
-	\maciek{TODO}
-	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
-	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate on exclude. Just sum up the real cost.}.
 
@@
 
-	The competitive ratio is then: $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
+	\ALG performs at least one swap for each ground set $\langle \{x_0, y_0\}, \{x_1, y_1\}, \{x_2, y_2\}, \ldots \rangle$.
+	We start with $k \cdot \ell - (k-1)$ single nodes (a single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with $\ell+1$ nodes.
+	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - l) / 2$.
 \end{proof}

### 2020-05-05 paper/disc20clustering.tex

@@
 
-	First, where we form a ground set of size $k-1$ on an arbitrarily chosen cluster $S$.
+	First, we create a ground set of size $k-1$ on an arbitrarily chosen cluster $S$.
 	%We never dismiss any constructed ground set.
-	Each cluster host exactly $k$ nodes, hence in any configuration of \ALG, there exists a single alone node in $S$.
-
-	Then, we construct a ground set containing this alone node (and some other node).
+	Each cluster hosts exactly $k$ nodes, and in any configuration a single alone node resides in $S$.
+
+	Then, we create a ground set containing this alone node.
 	Precisely,
-	we construct a ground set $\{x_0, y_0\}$, 
+	we create a ground set $\{x_0, y_0\}$, 
 	where $x_0$ is the alone node from $S$ and $y_0$ is an arbitrary node that does not reside in $S$.
-	\ALG has them split, and the adversary produces external requests until these are collocated.
-	$S$ already contains a ground set of size $k-1$ (that it cannot split), hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.
-
-	We continue to construct ground sets containing the node that is currently alone in $S$.
+	\ALG has them split, and the adversary produces external requests until they are collocated.
+	The cluster $S$ already contains a ground set of size $k-1$, hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.
 	To preserve a feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
-	\ALG must replace $x_0$ with another single node (that we name $x_1$).
-	Consequently, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x_1)$.
-	Repeating this, we observe a sequence $X := x_0, x_1, x_2\dots$ of alone nodes that visit $S$.
-	(Note that subsequent ground sets are different than the first ground set $\{x_0, y_0\}$, whose nodes were initially in different clusters.)
-
-	We claim that constructing such ground sets can continue as long as at least $\ell$ alone node exists.
+	\ALG must replace $x_0$ with another single node.
+
+	We continue adding the node that is currently alone in $S$ to a larger ground set.
+	Precisely, for an alone node $x$ from $S$, we merge the ground sets of $x$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x)$.
+	This way, $x$ is evicted similarly to $x_0$, and we continue creating such ground sets (we describe the termination condition shortly).
+	Note that subsequent ground sets are different than the first ground set $\{x_0, y_0\}$: now the nodes originate in the same cluster, whereas nodes $x_0$ and $y_0$ were initially in different clusters.
+
+	We claim that expanding ground sets in this way can continue as long as at least $\ell$ alone node exists.
 	Fix the moment just before the creation of a ground set with the node $x_i$.
@@
 	If $I(x_i)$ contains an alone node, then we simply obtain a perfect partition by swapping it with $x_i$.
-	Otherwise, $\{x_0, y_0\}$ is present on $I(x_i)$ (a single ground set cannot occupy $I(x_i)$, as without $x_i$ its size is at most $k-1$).
+	Otherwise, $\{x_0, y_0\}$ is present on $I(x_i)$ (a single ground set cannot occupy the whole cluster $I(x_i)$, as without $x_i$ its size is at most $k-1$).
 	In this case, we first swap $\{x_0, y_0\}$ with two alone nodes that occupy the same cluster.
@@
 
-	To keep \OPT's cost low, we end constructing ground sets prematurely --- while there are still $\ell+1$ alone nodes (i.e., not performing two last constructions).
-	Then, there two alone nodes exists that originate from the same cluster (we have $\ell+1$ alone nodes and $\ell$ clusters).
-	We denote these two nodes by $x^*, y^*$.
+	To keep \OPT's cost low, we end extending ground sets prematurely --- while there are still $\ell+1$ alone nodes (i.e., neglecting two last ground set expansions).
+	At the moment of termination, there exist two alone nodes that originate from the same cluster (as we have $\ell+1$ alone nodes and $\ell$ clusters).
+	We denote these two nodes by $x^*$ and $y^*$.
 
@@
 
-	\ALG performs at least one swap for each ground set $\langle \{x_0, y_0\}, \{x_1, y_1\}, \{x_2, y_2\}, \ldots \rangle$.
-	We start with $k \cdot \ell - (k-1)$ single nodes (a single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with $\ell+1$ nodes.
+	\ALG performs at least one swap for every expansion of a ground set.
+	Each expansion reduces the number of alone nodes by $1$.
+	We start with $k \cdot \ell - (k-1)$ alone nodes (a single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with alone $\ell+1$ nodes.
 	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - l) / 2$.

### 2020-05-05 paper/disc20clustering.tex
Intro to LB

@@
 First, we construct a ground set of size $k-1$ on one of the clusters $S$.
-In any configuration, there exists a node that accompanies it.
-We only issue requests between such accompanying (time-varying) node, and any alone node that was initially collocated with it.
-Repeating such requests, almost half of all node must visit the cluster $S_1$.
+In any configuration, there exists an alone node that accompanies it.
+We only issue requests between the current alone node from $S$, and some node that was initially collocated with it.
+Repeating such requests, almost all node must visit the cluster $S_1$.
 In comparison, \OPT performs one migration.

### 2020-05-05 paper/disc20clustering.tex
Intro to LB

@@
 First, we construct a ground set of size $k-1$ on one of the clusters $S$.
-In any configuration, there exists an alone node that accompanies it.
-We only issue requests between the current alone node from $S$, and some node that was initially collocated with it.
-Repeating such requests, almost all node must visit the cluster $S_1$.
+This, way, in any configuration, there exists an alone node that accompanies it.
+In such setting, we only issue requests between the current alone node from $S$, and some node that was initially collocated with it.
+Repeating such requests, almost all node must visit the cluster $S$.
 In comparison, \OPT performs one migration.

### 2020-05-05 paper/disc20clustering.tex
Intro to LB
wip

--- 2020-05-05 paper/disc20clustering.tex
@@
 To avoid infinite cost, each communicating pair must be collocated.
-We study the instances with finite cost: we assume that the communication graph can be perfectly partitioned in the clusters.
-The algorithm's goal is to \emph{learn} the communication graph's partition.
+We study instances with finite cost: we assume that the communication graph can be perfectly partitioned in the clusters.
+The algorithm's goal is to \emph{learn} the communication graph, and partition it without too many reconfigurations.
 
 For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$.
-The lower bound can be easily translated to the partitioning model (Section~\ref{sec:part}) with finite request sequence.
+The lower bound can be easily translated to the partitioning model (studied in Section~\ref{sec:part}) with finite request sequence.
 At the end of this section, we discuss an asymptotically optimal upper bound.
@@
 
-In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
-The construction is oblivious to the choice of the reconfiguration cost $\alpha$: we compare the number of node exchanges.
+We provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm for the learning problem.
+Later we elaborate on how to efficiently transform it to a lower bound for the partitioning problem.
 The lower bound requires $k\geq 3$.
-For $k=2$, a lower bound of $3$ is known \cite{repartition-disc}, and we provide a $6$-competitive algorithm (see Section~\ref{sec:k2}).
+For $k=2$, the learning problem is trivial: immediate collocation of a communicating pair is $1$-competitive.
+(A partitioning problem for $k=2$ is non-trivial, a lower bound of $3$ is known \cite{repartition-disc}, and we provide a $6$-competitive algorithm (see Section~\ref{sec:k2}).)
 
@@
 
-The adversary constructs ground sets in reaction to choices of a deterministic online algorithm.
-We emphasize that the algorithm does not know the ground set before they are revealed.
+The adversary constructs ground sets (a set of nodes that may communicate, see Section~\ref{sec:prelim}) depending on choices of a deterministic online algorithm.
+We emphasize that the algorithm does not know ground sets before they are revealed.
 Once we construct a ground set, it lasts until the end of the input sequence.
 
-The constructed ground sets can be partitioned into our clusters (without splitting nodes).
-This way, the optimal algorithm moves to such a partition, and serves all requests for free.
+The constructed ground sets can be perfectly partitioned into the clusters (i.e., nodes of each ground set collocated).
+This way, the optimal algorithm moves to a perfect partition at the beginning (where requests incur the cost $0$), thus its cost is bounded.
 If the algorithm splits any pair of nodes of a ground set, the adversary issue requests to any pair of split nodes.
-These two observations imply that the algorithm must eventually collocate nodes of a ground set to be competitive.
+These two observations imply that the algorithm must eventually collocate all nodes of a ground set to be competitive.
 
@@
 
-We conclude by emphasizing the generality of the lower bound.
-
 \noindent
@@
 The bound for partitioning converges to the bound for learning as the input sequence grows.
-
-Our construction allows for an exact lower bound for partitioning with finite input sequence.
-To this end, we issue only enough requests until the algorithm collocates the nodes of each ground set.
-If it never does, we repeat until it pays $k\cdot \ell$.
+However, our construction allows for an exact lower bound for partitioning with finite input sequence.
+To this end, we issue only enough requests until the algorithm collocates the nodes of each ground set (if it never does, we terminate after it pays $k\cdot \ell$).
+The construction is oblivious to the choice of the reconfiguration cost $\alpha$: we compare the 	number of node exchanges of \ALG and \OPT.
 

### 2020-05-05 paper/disc20clustering.tex

@@
 First, we construct a ground set of size $k-1$ on one of the clusters $S$.
-In any configuration, there exists a node that accompanies it.
-We only issue requests between such accompanying (time-varying) node, and any alone node that was initially collocated with it.
-Repeating such requests, almost half of all node must visit the cluster $S_1$.
+This, way, in any configuration, there exists an alone node that accompanies it.
+In such setting, we only issue requests between the current alone node from $S$, and some node that was initially collocated with it.
+Repeating such requests, almost all the nodes must visit the cluster $S$.
 In comparison, \OPT performs one migration.

### 2020-05-05 paper/disc20clustering.tex
wip

@@
 First, we construct a ground set of size $k-1$ on one of the clusters $S$.
-This, way, in any configuration, there exists an alone node that accompanies it.
+Then, in any configuration, there exists an alone node on $S$.
 In such setting, we only issue requests between the current alone node from $S$, and some node that was initially collocated with it.
@@
 	We claim that expanding ground sets in this way can continue as long as at least $\ell$ alone node exists.
-	Fix the moment just before the creation of a ground set with the node $x_i$.
+	Fix the moment just before the creation of a ground set with the node $x_i$.\maciek{no $x_i$ def any more}
 	We show that if at least $\ell$ alone nodes exists (including $x_i, y_i$), then there exists a perfect partition of ground sets.

### 2020-05-06 paper/disc20clustering.tex
wip

@@
 
-\section{Note on distributed algorithms}
-
-\subsection{Distributed implementation of $k=3$ algorithm}
-
-For a repartition, it is sufficient to find the cluster with the signature $1,1,1$. 
-We call any such cluster \emph{X (todo name)}.
-Idea of distributed implementation of this procedure:
+\section{Distributed implementation.}
+The algorithm PPL from Section~\ref{sec:ppl} can be distributed
+similarly to the approach in~\cite{sigmetrics19_partitioning}.
+The algorithm for $k=2$ from Section~\ref{sec:k2} performs only local communication for each request: counters are kept on the clusters and updated locally, and each migration concerns only two clusters.
+
+The algorithm for $k=3$ from Section~\ref{sec:k3} can be efficiently distributed.
+There are two non-local operations:
 \begin{enumerate}
-	\item Assign ids to clusters and order them
-	\item Recursively partition clusters into two groups
-	\item Each group maintains the pointer to a subgroup that contains at least one X cluster; a pointer may be null (then no X there)
-	\item This way, finding a cluster X requires $\log_2(\ell)$ communication
-	\item After removing an X cluster, we need to fix the pointers. This is done by following the path from the removed X cluster to the root. We simply ask the sibling if it contains X.
-	\item Possible improvement: we can have a fixed number of rounds for finding the X cluster by maintaining a direct pointer to an X cluster (not only to subgroup).
+	\item We broadcast the end of the phase to reset all counters.
+	\item Merge of two  may require finding a third cluster that contains $3$ alone nodes (For the details, see the proof of Lemma~\ref{lem:1req})).
+\end{enumerate}
+
+If a component grows to size $4$ or fresh cluster is found, we conclude that no perfect partition exists, and we broadcast (\maciek{TODO: efficient broadcast cite}) the message to end the phase and reset the counters in all clusters.
+
+Let's call the cluster containing $3$ alone nodes \emph{fresh}.
+Now we show how to efficiently find a fresh cluster.
+We organize the clusters into an arbitrary balanced binary tree (\maciek{TODO: how to do it efficiently? Does not matter much, as this is performed once at the beginning}), and broadcast the root.
+Each cluster counts the number of fresh clusters in its subtree.
+To find a fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
+Upon encountering a fresh cluster, we stop the traversal and decrease the counters on the followed path by $1$.
+
+The operation (ref1) requires a single broadcast, and the operation (ref2) requires $O(\log(\ell))$ communication.
+Each of these may occur on every request.
+
+\maciek{Possible improvement: we can have a fixed number of rounds for finding the X cluster by maintaining a direct pointer to an X cluster (not only to subgroup).
 However, the fix still requires the log number of rounds.
-But we improved the responsiveness.
-\end{enumerate}
-
-\subsection{Distributed implementation of $k=2$ algorithm}
-
-Trivial? Counters are local. We migrate between the clusters that have just communicated.
-
-\subsection{Distributed implementation of PPL}
-
-\maciek{How to distribute PPL better?}
-
-Our algorithms can be distributed
-similarly to the approach in~\cite{sigmetrics19_partitioning}.
+But we improved the responsiveness.}
 

### 2020-05-06 paper/disc20clustering.tex
wip

@@
  \newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}
- \newcommand\maciek[1]{\color{brown}\textbf{\\ Maciek: #1}\color{black}}
+ \newcommand\maciek[1]{\color{brown}\textbf{(Maciek: #1)}\color{black}}
  %\newcommand\maciek[1]{}
@@
 \begin{enumerate}
-	\item We broadcast the end of the phase to reset all counters.
-	\item Merge of two  may require finding a third cluster that contains $3$ alone nodes (For the details, see the proof of Lemma~\ref{lem:1req})).
+	\item Broadcasting the end of the phase.\label{it:broadcast}
+	\item Component merge.
+	\label{it:merge}
 \end{enumerate}
 
-If a component grows to size $4$ or fresh cluster is found, we conclude that no perfect partition exists, and we broadcast (\maciek{TODO: efficient broadcast cite}) the message to end the phase and reset the counters in all clusters.
-
-Let's call the cluster containing $3$ alone nodes \emph{fresh}.
-Now we show how to efficiently find a fresh cluster.
-We organize the clusters into an arbitrary balanced binary tree (\maciek{TODO: how to do it efficiently? Does not matter much, as this is performed once at the beginning}), and broadcast the root.
-Each cluster counts the number of fresh clusters in its subtree.
+If a component reaches size $4$ or fresh cluster is not found, we conclude that no perfect partition exists.
+Then, we end the phase and broadcast (\maciek{TODO: efficient broadcast cite}) the message to  reset the counters in all clusters.
+
+A merge of two components may require finding a third cluster that contains $3$ alone nodes (for details see the proof of Lemma~\ref{lem:1req})).
+We say that a cluster containing $3$ alone nodes is \emph{fresh}.
+In the following, we show how to efficiently find a fresh cluster in a distributed manner.
+We organize the clusters into an arbitrary rooted balanced binary tree (\maciek{TODO: how to do it efficiently?
+Does not matter much, as this is performed once at the beginning}), and we broadcast the root to each cluster.
+Each cluster maintains the counter of fresh clusters in its subtree.
 To find a fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
@@
 
-The operation (ref1) requires a single broadcast, and the operation (ref2) requires $O(\log(\ell))$ communication.
-Each of these may occur on every request.
-
-\maciek{Possible improvement: we can have a fixed number of rounds for finding the X cluster by maintaining a direct pointer to an X cluster (not only to subgroup).
-However, the fix still requires the log number of rounds.
-But we improved the responsiveness.}
+The operation (\ref{it:broadcast}) requires a single broadcast, and the operation (\ref{it:merge}) requires $O(\log(\ell))$ communication.
+Each of these may occur, on average, every $\alpha$ external requests.
 
@@
 
-\pagebreak
 \appendix
 
-\section{Omitted results}
+\section{Omitted results (for our future papers)}
 
 \begin{enumerate}
-	\item The lower bound of $\Omega(\ell)$ for the scenario with at most $(1/3-1/k)$-augmentation
-
 	\item Costly cascade example (it is easy to present $k^2$ and $k\cdot \sqrt{k}$ (with different approach), and then we can combine these to obtain $k^3$)
@@
 	\item lower and upper bound $k+1$ for two clusters, perfect partition and the optimal configuration reachable within one swap
-	\item 
-	The bound $(k \cdot \ell - k - \ell) / 2$ can be easily improved to $(k \cdot \ell - k - 1) / 2$ by playing a simple caching game in the first cluster, instead of just giving away the $k-1$ "blocker" (the caching game would build it gradually)
+
+	\item The lower bound $(k \cdot \ell - k - \ell) / 2$ can be easily improved to $(k \cdot \ell - k - 1) / 2$ by playing a simple caching game in the first cluster, instead of just giving away the $k-1$ "blocker" (the caching game would build it gradually)
 

### 2020-05-06 paper/disc20clustering.tex
Rewrite of the crucial part of LB

@@
 
-The constructed ground sets can be perfectly partitioned into the clusters (i.e., nodes of each ground set collocated).
-This way, the optimal algorithm moves to a perfect partition at the beginning (where requests incur the cost $0$), thus its cost is bounded.
-If the algorithm splits any pair of nodes of a ground set, the adversary issue requests to any pair of split nodes.
-These two observations imply that the algorithm must eventually collocate all nodes of a ground set to be competitive.
-
-First, we construct a ground set of size $k-1$ on one of the clusters $S$.
-Then, in any configuration, there exists an alone node on $S$.
-In such setting, we only issue requests between the current alone node from $S$, and some node that was initially collocated with it.
+The constructed ground sets can be perfectly partitioned into the clusters (i.e., with nodes of each ground set collocated).
+This way, the optimal algorithm moves to a perfect partition at the beginning (where requests incur the cost~$0$), thus its cost is bounded.
+If the algorithm splits any pair of nodes of a ground set, the adversary issues requests to any pair of split nodes.
+These two observations imply that the algorithm must eventually collocate all nodes of a~ground set to be competitive.
+
+First, we construct a ground set of size $k-1$ on an arbitrary cluster $S$.
+In any configuration, there exists an alone node on $S$.
+We issue requests between the current alone node from $S$, and some node that was initially collocated with it.
 Repeating such requests, almost all the nodes must visit the cluster $S$.
@@
 	\label{th:lowerbound}
-	The competitive ratio of any deterministic online algorithm for \OBRP{} is  $(k\cdot \ell - k - \ell)/2$ for any $k\geq 3$ and for any $\ell \geq 2$.
+	The competitive ratio of any deterministic online algorithm for \OBRP is at least $(k\cdot \ell - k - \ell)/2$ for any $k\geq 3$ and $\ell \geq 2$.
 \end{theorem}
@@
 
-	Then, we create a ground set containing this alone node.
+	Then, we add the alone node from $S$ to a larger ground set to force its eviction.
 	Precisely,
@@
 	We continue adding the node that is currently alone in $S$ to a larger ground set.
-	Precisely, for an alone node $x$ from $S$, we merge the ground sets of $x$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x)$.
-	This way, $x$ is evicted similarly to $x_0$, and we continue creating such ground sets (we describe the termination condition shortly).
-	Note that subsequent ground sets are different than the first ground set $\{x_0, y_0\}$: now the nodes originate in the same cluster, whereas nodes $x_0$ and $y_0$ were initially in different clusters.
+	Precisely, for an alone node $x$ from~$S$, we merge the ground sets of $x$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x)$.
+	This way, $x$ is evicted similarly to $x_0$, and we continue creating such ground sets.
+	We describe the termination condition shortly.
+	(Note that these ground sets contain nodes that originate from the same cluster, whereas the nodes $x_0$ and $y_0$ originate from different clusters.)
 
 	We claim that expanding ground sets in this way can continue as long as at least $\ell$ alone node exists.
-	Fix the moment just before the creation of a ground set with the node $x_i$.\maciek{no $x_i$ def any more}
-	We show that if at least $\ell$ alone nodes exists (including $x_i, y_i$), then there exists a perfect partition of ground sets.
+	Fix the moment just before the creation of a ground set with the alone node $z$ that currently resides in $S$.
+	We show that if at least $\ell$ alone nodes exists (including $z$), then there exists a perfect partition of ground sets after expanding a ground set with $z$.
 
@@
 	In this configuration, each cluster contains:  at most one ground set larger than $1$, possibly some alone nodes, and possibly the ground set $\{x_0, y_0\}$.
-	If $I(x_i)$ contains an alone node, then we simply obtain a perfect partition by swapping it with $x_i$.
-	Otherwise, $\{x_0, y_0\}$ is present on $I(x_i)$ (a single ground set cannot occupy the whole cluster $I(x_i)$, as without $x_i$ its size is at most $k-1$).
-	In this case, we first swap $\{x_0, y_0\}$ with two alone nodes that occupy the same cluster.
-	To see that such two nodes must exist, observe that we have at most $\ell-1$ clusters with alone nodes, and by the assumption we have $\ell$ alone nodes.
-	After this swap, we migrate $x_i$ to $I(x_i)$, evicting a lone node to $S$, and we conclude that the perfect partition exists.
+	If $I(z)$ contains an alone node, then we simply obtain a~perfect partition by swapping it with $z$.
+	Otherwise, if no alone nodes exist on $I(z)$, then its largest ground set is at most $k-1$, because $z$ is on $S$.
+	Hence, $\{x_0, y_0\}$ is present on $I(z)$, and we must evict it to make space for $z$.
+	To find the cluster for $\{x_0, y_0\}$, observe that we have at most $\ell-1$ clusters with alone nodes, and we have at least $\ell$ alone nodes --- hence a cluster with two alone nodes exists.
+	To find a feasible partition, we swap $\{x_0, y_0\}$ with these two collocated alone nodes, and then we swap an alone node from $I(z)$ with $z$.
 
@@
 The algorithm for $k=2$ from Section~\ref{sec:k2} performs only local communication for each request: counters are kept on the clusters and updated locally, and each migration concerns only two clusters.
-
-The algorithm for $k=3$ from Section~\ref{sec:k3} can be efficiently distributed.
-There are two non-local operations:
+\medskip
+
+In the rest of this section, we propose the distributed implementation of the algorithm for $k=3$ from Section~\ref{sec:k3}.
+The algorithm performs two non-local operations:
 \begin{enumerate}

### 2020-05-06 paper/disc20clustering.tex

@@
 	If $I(z)$ contains an alone node, then we simply obtain a~perfect partition by swapping it with $z$.
-	Otherwise, if no alone nodes exist on $I(z)$, then its largest ground set is at most $k-1$, because $z$ is on $S$.
-	Hence, $\{x_0, y_0\}$ is present on $I(z)$, and we must evict it to make space for $z$.
+	In the following we assume that no alone nodes exist on $I(z)$.
+	The largest ground set of $I(z)$ is at most $k-1$, because $z$ is on $S$.
+	Therefore $\{x_0, y_0\}$ is present on $I(z)$, and we must evict it to make space for $z$.
 	To find the cluster for $\{x_0, y_0\}$, observe that we have at most $\ell-1$ clusters with alone nodes, and we have at least $\ell$ alone nodes --- hence a cluster with two alone nodes exists.

### 2020-05-06 paper/disc20clustering.tex

@@
 	To keep \OPT's cost low, we end extending ground sets prematurely --- while there are still $\ell+1$ alone nodes (i.e., neglecting two last ground set expansions).
-	At the moment of termination, there exist two alone nodes that originate from the same cluster (as we have $\ell+1$ alone nodes and $\ell$ clusters).
+	Consider \ALG's configuration after the last expansion.
+	Among $\ell+1$ alone nodes there must exist a pair that originates from the same cluster (we have $\ell$ clusters). 
 	We denote these two nodes by $x^*$ and $y^*$.

### 2020-05-06 paper/disc20clustering.tex
wip

@@
 
+\maciek{TODOs: "requests between nodes" to "requests to pair of nodes"}
+
 \section{Introduction}
@@
 Repeating such requests, almost all the nodes must visit the cluster $S$.
-In comparison, \OPT performs one migration.
+In comparison, \OPT performs only two migrations.
 
@@
 \textbf{Resource augmentation.}
-The majority of work on Online Balanced Partitioning so far (todo: cite) focuses on the scenario with resource augmentation.
-We note that this lower bound provides insights into this scenario as well.
-
-Consider an instance with $k$ divisible by $3$, and construct $3$ ground sets of size $k/3$ in each cluster.
-Then, apply the construction from the lower bound for $k=3$, now treating these ground sets in the way we used individual nodes.
-For the resource augmentation as large as $1+1/3-\epsilon$, no more than $3$ such ground sets fit in one cluster.
+The majority of work on Online Balanced Partitioning so far \cite{repartition-disc,sigmetrics19_partitioning} focuses on the scenario with resource augmentation, where the clusters of \ALG are larger than the clusters of \OPT.
+We can adjust our construction to show a lower bound of $\Omega(\ell)$ for resource augmentation.
+
+Consider a partitioning problem with resource augmentation $1+1/3-\epsilon$.
+Fix $k$ divisible by $3$, and construct $3$ ground sets of size $k/3$ in each cluster.
+Note that no more than $3$ such ground sets fit in one cluster.
+Then, apply the construction from the lower bound for $k=3$, using these ground sets in the way we used individual nodes.
 The cost of any algorithm (including \OPT) scales up by $k/3$, and the lower bound $\Omega(\ell)$ holds.
 
-Finally, we note the possibility for improvement (a gap). The algorithm CREP~\cite{repartition-disc} requires $(2+\epsilon)$-augmentation to guarantee the competitive ratio independent of $\ell$.
+Finally, we note the possibility for improvement. The algorithm CREP~\cite{repartition-disc} requires $(2+\epsilon)$-augmentation to guarantee the competitive ratio independent of $\ell$.
 

### 2020-05-06 paper/disc20clustering.tex

@@
 \textbf{General partitioning problem.}
-Every lower bound for the learning problem constitutes a valid lower bound for the partitioning problem.
-The bound for partitioning converges to the bound for learning as the input sequence grows.
+Every lower bound for the learning problem constitutes a converging lower bound for the partitioning problem as the input sequence grows to infinity.
 However, our construction allows for an exact lower bound for partitioning with finite input sequence.

### 2020-05-06 paper/disc20clustering.tex
major k=2 comments

@@
 
- \newcommand\mahmoud[1]{\color{orange}\textbf{\\ Mahmoud: #1}\\\color{black}}
+ \newcommand\mahmoud[1]{\color{orange}\textbf{Mahmoud: #1}\\\color{black}}
  \newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}
@@
 Algorithm Alg:
+\mahmoud{find a better name for Alg.}
 maintain a counter $C_{\set{u,v}}$ for each pair of nodes $\set{u,v}$ and increment it on every remote request between $u$ and $v$.
@@
 \begin{theorem} \label{thm:k=2}
-	The algorithm is (strictly) 6-competitive.
-	\maciek{for what lambda?}
+	The algorithm, for $\lambda=\alpha$, is (strictly) 6-competitive.
+	\maciek{for what lambda? \mahmoud{done}}
 \end{theorem}
@@
 OPT may save some costs by collocating multiple pairs at once (before they incur too much cost),
-paying the price of only one migration per pair  (see Figure \ref{fig:TBD}).
+paying the price of only one migration per pair  (\mahmoud{see Figure \ref{fig:TBD}}).
 %This observation suggests that we can't charge OPT more than $\alpha$ per collocation.
@@
 
-\begin{proof}[Proof of Theorem \ref{thm:k=2}]
 Assume on request $\sigma_{t}=\set{u,v}$, 
@@
 We denote the set of all requests to a pair $\set{x,y}$ that arrive
-	during $(t_1,t_2]$ by $\sigma_{\set{x,y}}(t_1,t_2]$.
-
-	If $t$ is the first time that Alg collocates $\{u,v\}$ then we  charge
-	OPT $\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
-	or $\alpha$ for collocating the pair and serving them with no additional cost.
-	Therefore in this case ,
-	OPT pays $\min{\{ \lambda,\alpha \}}$.
-
-	Else, we consider that Alg collocated $\{u,v\}$ in the past.
-	Let $t'$, $t' < t$ be the  time when Alg separates $\set{u,v}$
+during $(t_1,t_2]$ by $\sigma_{\set{x,y}}(t_1,t_2]$,
+and its overall cost to OPT by
+$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) $.
+%
+Observe that for requests in
+$R_t := \sigma_{\{u,v'\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t]$ 
+and the entire duration of the input sequence (i.e., $(0,m]$),
+OPT must pay a total cost
+%We denote the cost w.r.t~$r_t$ by
+$\mathit{Cost}_{\mathit{OPT}} (R_t) \geq 
+	c_{R_t} := \min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$,
+for serving (some of) these requests remotely or (and) collocating (one of) the pairs,
+regardless of its choices.
+Then the cost per pair, in particular the cost w.r.t.~$\{u,v\}$, is
+$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) \geq c_{R_t} / 2$.
+%
+However,
+we can charge this cost to OPT only at $t=t_0$,
+i.e.~the first time Alg collocates $\{u,v\}$,
+and not at any consequent time $\tilde{t} > t_0$ that Alg collocates $\{u,v\}$.
+%	i.e.,
+%	when Alg collocates the two pairs a second time.
+That is,
+we charge OPT a cost incurred by requests in
+$\sigma_{\{u,v'\}}(t'',t']$ or $\sigma_{\{u,v\}}(t',t]$,
+if and only if  OPT incurs the cost during  intervals $(t'', t']$ or $(t', t]$,
+respectively.	
+Latter,
+this ensures us that we do not charge OPT for a cost more than twice.
+%%
+Based on this observation,
+in the proof that follows,
+we distinguish three cases of OPT's choices
+while it serves requests in	$R_t $.
+
+\todo{what is causing this vertical space before proof block??}
+
+\begin{proof}[Proof of Theorem \ref{thm:k=2}]
+	If $t$ is the first time that Alg collocates $\{u,v\}$ then  OPT pays
+	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
+	or $\alpha$ for collocating the pair and
+	serving (some of) the requests with cost 0.
+	Therefore in this case,
+	OPT pays
+	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
+	%
+	Else, this is not the first time and
+	Alg collocates $\{u,v\}$ also some time prior to $t$.
+	Let $t' < t$ be the  time when Alg separates $\set{u,v}$
 	in order to collocate some (intersecting) pair
@@
 	if $t'$ is not the first time that Alg  collocates $\{u,v'\}$ 
-	then let $t''<t'$ be the last time before $t'$ when Alg separates $\set{u,v'}$.
+	then let $t''<t'$ be the latest time before $t'$ when Alg separates $\set{u,v'}$.
 	Else,
@@
 	$| \sigma_{\set{u,v'}}(t'',t'] | = | \sigma_{\set{u,v}}(t',t] |  =  \lambda$.
+	%
 	The overall cost of requests in $\sigma_{\set{u,v}}(t',t]$,
@@
 	$\lambda + 2\alpha$.
-	Observe that for requests in $r_t := \sigma_{\{u,v'\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t]$,
-	OPT must pay a total cost of serving and (or) migration at least
-	$\min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$
-	within $(0,m]$,
-	regardless of its choices.
-	%% 
-	However,
-	we cannot charge this cost to OPT repeatedly,
-	i.e.,
-	when Alg collocates the two pairs a second time.
-	In particular,
-	we charge OPT a cost incurred by requests in
-	$\sigma_{\{u,v'\}}(t'',t']$ and $\sigma_{\{u,v\}}(t',t]$,
-	if and only if  OPT incurs the cost within  intervals $(t'', t']$ and $(t', t]$,
-	respectively.	
+	OPT pays,
+	$\mathit{Cost}_{\mathit{OPT}} (R_t) := \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v'\}}(t'',t']) 
+	+ \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) $.
+	We distinguish three (non-exhaustive) cases for the cost to OPT.
 	Latter,
-	this ensures us that we do not charge OPT for a cost more than twice.
-	%%
-	Based on this observation,
-	we distinguish three cases of OPT's choices
-	while it serves requests in	$r_t $.
-	We denote its cost by
-	$\mathit{Cost}_{\mathit{OPT}} (r_t) := \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v'\}}(t'',t']) 
-	+ \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) $.
+	 we argue that OPT cannot save cost by deviating from these cases.
 	\begin{enumerate}[label=\roman*.]
@@
 		and keeps them so during $(t'',t']$ and $(t',t]$, respectively.
-		Therefore it serves all requests in $r_t$ with no additional cost and
-		$\mathit{Cost}_{\mathit{OPT}} (r_t) = 2\alpha$.
+		Therefore it serves all requests in $R_t$ with no additional cost and
+		$ c_1 (R_t) := 2\alpha$.
 		\item
@@
 		Hence,
-		it pays only $\lambda$ for serving all requests in $r_t$ and
-		$\mathit{Cost}_{\mathit{OPT}} (r_t) = \lambda$
+		it pays only $\lambda$ for serving all requests in $R_t$ and
+		$c_2 (R_t) := \lambda$
 		\item
-		OPT has $\{u,v'\}$ ($\{u,v\}$) already  collocated at $t''$ ($t'$)  and also during $(t'',t']$ ($(t', t])$,
-		and it collocates $\{u,v\}$ ($\{u,v'\}$) at $t'+1$ ($t''+1$) paying $\alpha$.
-		%		Therefore it pays only $\alpha$ during $[t'',t]$,
-		%		for collocating $\{u,v'\}$
-		%	    and for the one request to $\{u,v\}$ at $t'$.
-		%		I.e., it pays zero for every other  request in $r_t$.
-		Then,
-		$\mathit{Cost}_{\mathit{OPT}} (r_t) = \alpha$.
+		OPT has $\{u,v'\}$ ($\{u,v\}$) already collocated at $t''$ ($t'$)  and also during $(t'',t']$ ($(t', t])$.
+		Then it pays 0 for $\sigma_{\{u,v'\}}[t'',t']$ ($(\sigma_{\{u,v\}}[t',t]$), and
+		$c_3 (R_t) := \alpha$.
 	\end{enumerate}
-	In every of these cases,
-	we have
-	$\mathit{Cost}_{\mathit{OPT}} (r_t) \geq  \min{ \{ \lambda, \alpha \}}$.
-	OPT may deviate from these cases,
-	e.g.~by collocating a pair after partially paying for requests to this pair.
+	If OPT's choice is one of these cases,
+	then
+	$\mathit{Cost}_{\mathit{OPT}} (R_t) \geq
+		\min{\{ c_1(R_t), c_2(R_t), c_3(R_t) \}}  =
+		\min{ \{ \lambda, \alpha \}}$.
+	Otherwise,
+	OPT deviates from these cases,
+	e.g.~by collocating a pair after paying for some of requests to this pair.
 	However,
 	no deviation can decrease the cost to OPT.
-	\mahmoud{It would be better to describe and bound the "deviation" formally.}
 	Therefore in any case,
-	$\mathit{Cost}_{\mathit{OPT}} (r_t) \geq  \min{ \{ \lambda, \alpha \}}$.
-
-	It remains to charge both Alg and OPT for requests that do not lead to collocations until the  end of the sequence at $t=m$.
+	$\mathit{Cost}_{\mathit{OPT}} (R_t) \geq  \min{ \{ \lambda, \alpha \}}$.
+
+	It remains to bound the cost incurred by Alg and OPT for requests that do not lead to collocations until the  end of the sequence at $t=m$.
 	Assume $\{u,v\}$ is not collocated at $t=m$
@@
 	If $\{u,v\}$ is not collocated in the initial configuration
-	and Alg never collocates it through the sequence,
+	and Alg never collocates it,
 	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[1,m] |$.
-	We charge OPT the cost
+	OPT pays
 	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[1,t]) 
 	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
-	for collocating this pair or (and) paying (partially) for requests in $\sigma_{\{u,v\}}[1,t]$.
+	for collocating this pair or (and) paying for (some of) requests in $\sigma_{\{u,v\}}[1,t]$.
 	Else,
@@
 	We define times $t'' < t' < m$ similarly to the case where Alg collocates $\{u,v\}$.
-	Let $r_{\{u,v\}} := \sigma_{\{u,v'\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
+	Let $R_{\{u,v\}} := \sigma_{\{u,v'\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
 	Then,
@@
 	we obtain
-	$\mathit{Cost}_{\mathit{OPT}} (r_{\{u,v\}}) 
+	$\mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}}) 
 	\geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.
-
-	Next, we sum up all costs charged to OPT.
+	The overall cost to Alg is
+
+\begin{align*}	%\label{eq:costAlg}
+	\mathit{Cost}_{\mathit{ALG}} (\sigma)
+	=
+	\sum_{ t \in T}(\lambda + 2\alpha) +
+	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
+	\leq
+	\sum_{ t \in T}(\lambda + 2\alpha) +
+	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
+	=
+	\sum_{ t \in T} 3\alpha +
+	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
+	.
+	\end{align*}
+
+	Next, we sum up all costs incurred by OPT.
 	Observe that by definition,
-	$r_{t'} \cap r_t = \sigma_{\{u,v'\}}[t'',t']$.
+	$R_{t'} \cap R_t = \sigma_{\{u,v'\}}[t'',t']$.
 	This means,
-	the cost of r.h.s.~to OPT is counted at most twice in the expression
-	$\mathit{Cost}_{\mathit{OPT}} (r_{t'}) + \mathit{Cost}_{\mathit{OPT}} (r_t)$.
+	the cost of r.h.s.~to OPT is counted at most twice in each of  the expressions
+	$\mathit{Cost}_{\mathit{OPT}} (R_{t'}) + \mathit{Cost}_{\mathit{OPT}} (R_t)$
+	and  
+	$\mathit{Cost}_{\mathit{OPT}} (R_{t'}) + \mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}})$.
 	Hence,
@@
 	OPT pays at least 
-	$(
-	\sum_{ 0 < t \leq |\sigma|} \mathit{Cost}_{\mathit{OPT}} (r_t) +
-	\sum_{\{u,v\}} \mathit{Cost}_{\mathit{OPT}} (r_{\{u,v\}})
-	) / 2$.
+	$\frac{1}{2}(
+	\sum_{ t \in [1,m] } \mathit{Cost}_{\mathit{OPT}} (R_t) +
+	\sum_{\{u,v\}} \mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}})
+	) $.
 	Let
-	$T := \{ 0 < t \leq m ~\vert~ \exists \{u,v\}: C^{t}_{\{u,v\}} = \lambda \}$.
-	By setting $\lambda= \alpha$,
-	we obtain
+	$T := \{ t \in [1,m] ~\vert~ \exists \{u,v\}: C^{t}_{\{u,v\}} = \lambda \}$.
+	Then,
+	the total cost to OPT is
+
+	\begin{align*} 	%\label{eq:costOPT}
+		\mathit{Cost}_{\mathit{OPT}} (\sigma)
+		&=
+		\frac{1}{2}
+		\Big(
+		\sum_{ t \in T} \mathit{Cost}_{\mathit{OPT}} (R_t) 
+		+ \sum_{\{u,v\}}\mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}})
+		\Big)	\\
+		&\leq
+		\Big(
+		\sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  +
+		\sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \} } 
+		\Big)		
+		=
+		\frac{1}{2}		
+		\Big(
+		\sum_{ t \in T} \alpha  
+		+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}
+		\Big).
+	\end{align*}
+
+Thus, we have
+
 \begin{align*}
@@
 	\mathit{Cost}_{\mathit{OPT}} (\sigma)
-		&=
-		\Big(
-		\sum_{ t \in T}(\lambda + 2\alpha) +
-		\sum_{\{u,v\}} C^{m}_{\{u,v\}}
-		\Big)	 \big /
-		\frac{1}{2}
-		\Big(
-		\sum_{ t \in T} \mathit{Cost}_{\mathit{OPT}} (r_t) 
-		+ \sum_{\{u,v\}}\mathit{Cost}_{\mathit{OPT}} (r_{\{u,v\}})
-		\Big)	\\
-		&\leq
-		\Big(
-		\sum_{ t \in T}(\lambda + 2\alpha) +
-		\sum_{\{u,v\}} C^{m}_{\{u,v\}}
-		\Big)	 \big /
-		\frac{1}{2}
-		\Big(
-		\sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  +
-		\sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \} } 
-		\Big)	\\
-		&=
+	\leq
 	2\Big(
@@
 	+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}  
-		\Big)	< 6
-	\end{align*}.
+	\Big)	< 6.
+\end{align*}
 %	The claim follows from here.
+
 \end{proof}
@@
 if and only if  OPT incurs the cost within  intervals $(t'', t']$ and $(t', t]$,
-respectively.	"}
-
-\maciek{"Observe that for requests in $r_t$".
-First the separate sentence with the definition of $r_t$ (it is challenging enough to just understand it.
+respectively.
+\mahmoud{this is only for intuition but a key statement.
+It is meant to say: OPT may pay cost for the same two pairs outside the intervals.
+But we don't charge OPT  those costs at $t$.
+This way we avoid over charging OPT as only intervals $r_{t'}$ and $r_t$ can overlap which leads to double charging and we take care of it.}
+"}
+
+\maciek{"Observe that for requests in $R_t$".
+First the separate sentence with the definition of $R_t$ (it is challenging enough to just understand it.
 Then reason why it pays the minimum of three values (I don't see it, and later you show that cost(rt) is at least minimum of other two values).
-Is it a consequence of three cases?}
+Is it a consequence of three cases?
+\mahmoud{the duration is the key and the latter duration is different.
+I reworded the whole thing.
+And yes, the same three cases can be used to explain this observation.
+Now the "some of" in the parenthesis might be enough for a clue.
+This observation is  used implicitly in the proof when it bounds the cost of first time collocation.}}
 
@@
 Not necessarily rewrite!
-Just maybe check if it's true.}
+Just maybe check if it's true.
+\mahmoud{this is the cost of first collocation of $\{u,v\}$. "partially" is confusing and now it says "some of" instead.}}
 
@@
 
-\maciek{The big inequality is impressive, but not a lot change in each row. Instead, bound the cost of ALG separately in one line, and the cost of OPT in another. (possibly far away from eachother)}
-
-\maciek{Remove all "charge both". Just sum up costs. I would say not use "charge" at all, but it's up to you.}
+\maciek{The big inequality is impressive, but not a lot change in each row.
+Instead, bound the cost of ALG separately in one line, and the cost of OPT in another.
+(possibly far away from eachother) \mahmoud{done}}
+
+\maciek{Remove all "charge both". Just sum up costs. I would say not use "charge" at all, but it's up to you. \mahmoud{done}}
 
@@
 
-\maciek{$r_t$ is the central concept in this proof.
+\maciek{$R_t$ is the central concept in this proof.
 Possibly even start with the paragraph: charging OPT, regular collocation.

### 2020-05-06 paper/disc20clustering.tex
restructuring the proof

@@
 	The algorithm, for $\lambda=\alpha$, is (strictly) 6-competitive.
-	\maciek{for what lambda? \mahmoud{done}}
 \end{theorem}
@@
 We denote the set of all requests to a pair $\set{x,y}$ that arrive
-during $(t_1,t_2]$ by $\sigma_{\set{x,y}}(t_1,t_2]$,
+during $[t_1,t_2]$ by $\sigma_{\set{x,y}}[t_1,t_2]$,
 and its overall cost to OPT by
-$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) $.
+$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[t_1,t_2])$.
 %
+%Else, this is not the first time and
+%Alg collocates $\{u,v\}$ also some time prior to $t$.
+
+If $t$ is not the first time that Alg collocates $\{u,v\}$ then
+let $0 < t' < t$ be the latest time when Alg separates $\set{u,v}$
+in order to collocate some (intersecting) pair
+$\{x,y\} \neq \{u,v\}, \{x,y\} \cap \{u,v\} \neq \emptyset$, 
+e.g.,
+$\{x,y\}=\{u,v'\}$.
+Else,
+$t$ is the first time that Alg collocates $\{u,v\}$ and let $t' := 0$.
+Similarly,
+if $t' > 0$ is not the first time that Alg  collocates $\{u,v'\}$ 
+then let $0 < t'' < t'$ be the latest time before $t'$ when Alg separates $\set{u,v'}$.
+Else,
+$t'$ is the first time that Alg collocates $\{u,v'\}$ and we let $t''=0$.
+%
+
 Observe that for requests in
@@
 
-\todo{what is causing this vertical space before proof block??}
-
 \begin{proof}[Proof of Theorem \ref{thm:k=2}]
+	First,
+	we bound the costs incurred by Alg for requests that
+	lead to the collocation of $\{u,v\}$ at time $t \in T$, where
+	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$.
+	By definitions of $t$ and $t'$,
+	the overall cost of requests in $\sigma_{\set{u,v}}(t',t]$ incurred by Alg,
+	i.e., the total cost of remote serving
+	and the moving cost, is
+	$\lambda + 2\alpha$.	
+	Next,
+	we bound the cost incurred by Alg
+	for requests that do not lead to collocations until the  end of the sequence at $t=m$.
+	Assume $\{u,v\}$ is not collocated at $t=m$
+	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
+	which means Alg pays $C^{m}_{ \{u,v\} }$
+	for these requests $\sigma_{\set{u,v}}(t',m]$.
+    Then the overall cost to Alg is
+
+	\begin{align*}	%\label{eq:costAlg}
+	\mathit{Cost}_{\mathit{ALG}} (\sigma)
+	=
+	\sum_{ t \in T}(\lambda + 2\alpha) +
+	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
+	\leq
+	\sum_{ t \in T}(\lambda + 2\alpha) +
+	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
+	=
+	\sum_{ t \in T} 3\alpha +
+	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
+	.
+	\end{align*}
+
+	Next,
+	we bound  costs incurred by OPT	for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
 	If $t$ is the first time that Alg collocates $\{u,v\}$ then  OPT pays
@@
 	%
-	Else, this is not the first time and
-	Alg collocates $\{u,v\}$ also some time prior to $t$.
-	Let $t' < t$ be the  time when Alg separates $\set{u,v}$
-	in order to collocate some (intersecting) pair
-	$\{x,y\} \neq \{u,v\}, \{x,y\} \cap \{u,v\} \neq \emptyset$, 
-	e.g.,
-	$\{x,y\}=\{u,v'\}$.
-	%	I.e., $ C^{t'}_{\set{u,v'}} = \lambda$.
-	Similarly,
-	if $t'$ is not the first time that Alg  collocates $\{u,v'\}$ 
-	then let $t''<t'$ be the latest time before $t'$ when Alg separates $\set{u,v'}$.
-	Else,
-	$\{u,v'\}$ is not collocated in the initial configuration and we let $t''=0$.
 	By these definitions,
@@
 	$| \sigma_{\set{u,v'}}(t'',t'] | = | \sigma_{\set{u,v}}(t',t] |  =  \lambda$.
-	%
-	The overall cost of requests in $\sigma_{\set{u,v}}(t',t]$,
-	and the moving cost,
-	incurred by Alg is
-	$\lambda + 2\alpha$.
-	OPT pays,
+	For requests $R_t := \sigma_{\set{u,v'}}(t'',t']  \cup \sigma_{\set{u,v}}(t',t]$,
+	OPT pays
 	$\mathit{Cost}_{\mathit{OPT}} (R_t) := \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v'\}}(t'',t']) 
@@
 
-	It remains to bound the cost incurred by Alg and OPT for requests that do not lead to collocations until the  end of the sequence at $t=m$.
-	Assume $\{u,v\}$ is not collocated at $t=m$
-	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
-	which means Alg pays $C^{m}_{ \{u,v\} }$
-	for requests to $\{u,v\}$ so far unaccounted.
+	It remains to bound the cost  incurred by OPT for requests to $\{u,v\}$ that do not lead to its collocation until the end of the sequence at $t=m$.
 	If $\{u,v\}$ is not collocated in the initial configuration
@@
 	OPT pays
-	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[1,t]) 
+	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[1,m]) 
 	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
-	for collocating this pair or (and) paying for (some of) requests in $\sigma_{\{u,v\}}[1,t]$.
+	for collocating this pair or (and) paying for (some of) requests in $\sigma_{\{u,v\}}[1,m]$.
 	Else,
@@
 	\geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.
-	The overall cost to Alg is
-
-\begin{align*}	%\label{eq:costAlg}
-	\mathit{Cost}_{\mathit{ALG}} (\sigma)
-	=
-	\sum_{ t \in T}(\lambda + 2\alpha) +
-	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
-	\leq
-	\sum_{ t \in T}(\lambda + 2\alpha) +
-	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
-	=
-	\sum_{ t \in T} 3\alpha +
-	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
-	.
-	\end{align*}
 
@@
 	) $.
-	Let
-	$T := \{ t \in [1,m] ~\vert~ \exists \{u,v\}: C^{t}_{\{u,v\}} = \lambda \}$.
 	Then,
@@
 
-Thus, we have
+Finally, we have
 

### 2020-05-07 paper/disc20clustering.tex
restructuring the proof
comments

@@
 
- \newcommand\mahmoud[1]{\color{orange}\textbf{Mahmoud: #1}\\\color{black}}
+ \newcommand\mahmoud[1]{\color{orange}\textbf{Mahmoud: #1~}\color{black}}
  \newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}
@@
 	Fix any online algorithm \ALG{}.
-	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
+	Let $I(C)$ denote the cluster where nodes of the ground set $C$ are located initially.
 	If $C$ contains nodes that originate from different clusters, $I(C)$ is undefined.
+	\mahmoud{i.e., $I(C):=...$formalized with notations}.
 	Throughout the construction, we maintain ground sets for each node.
 	Initially all nodes are \emph{alone} (they belong to their own trivial ground sets).
-
+\mahmoud{isolated would be a better word for "alone"}
 	First, we create a ground set of size $k-1$ on an arbitrarily chosen cluster $S$.
+	\mahmoud{$\mathcal{S}$ is better for a set of sets.}
 	%We never dismiss any constructed ground set.
 	Each cluster hosts exactly $k$ nodes, and in any configuration a single alone node resides in $S$.
+\mahmoud{an isolated node is collocated with  this GS in $\mathcal{S}$.
+We denote the first such node by $x_0$.}
 
@@
 	where $x_0$ is the alone node from $S$ and $y_0$ is an arbitrary node that does not reside in $S$.
-	\ALG has them split, and the adversary produces external requests until they are collocated.
+	\ALG has them split, and the adversary issues external requests until \ALG collocates them.
 	The cluster $S$ already contains a ground set of size $k-1$, hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.
@@
 	\ALG must replace $x_0$ with another single node.
+	\mahmoud{In any new partition, \ALG must replace $x_0$..., otherwise it is not a feasible partition.}
 
@@
 	Precisely, for an alone node $x$ from~$S$, we merge the ground sets of $x$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x)$.
+	\mahmoud{$x_i$?}
 	This way, $x$ is evicted similarly to $x_0$, and we continue creating such ground sets.
 	We describe the termination condition shortly.
-	(Note that these ground sets contain nodes that originate from the same cluster, whereas the nodes $x_0$ and $y_0$ originate from different clusters.)
+	(Note that these ground sets contain nodes that originate from the same cluster, whereas the nodes $x_0$ and $y_0$ originate from distinct clusters.)
 
 	We claim that expanding ground sets in this way can continue as long as at least $\ell$ alone node exists.
-	Fix the moment just before the creation of a ground set with the alone node $z$ that currently resides in $S$.
+	Fix the moment just before the creation of a ground set with the alone node \mahmoud{$x_{t^*}$?} that currently resides in $S$.
+	\mahmoud{ALG might change $S$. We can say, "currently collocated with the...".}
 	We show that if at least $\ell$ alone nodes exists (including $z$), then there exists a perfect partition of ground sets after expanding a ground set with $z$.
-
+	\mahmoud{the termination condition is needed before the claim statement.}
 	To show this claim, we consider a configuration, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
@@
 %%
-Based on this observation,
-in the proof that follows,
-we distinguish three cases of OPT's choices
-while it serves requests in	$R_t $.
+We use notations defined so far in the proof that follows.
 
@@
 	Next, we sum up all costs incurred by OPT.
-	Observe that by definition,
-	$R_{t'} \cap R_t = \sigma_{\{u,v'\}}[t'',t']$.
+	By definitions of $R_t$ and $R_{\{u,v\}}$, we have either
+	$R_{t'} \cap R_t = \sigma_{\{u,v'\}}[t'',t']$ or
+	$R_{t'} \cap R_{\{u,v\}} = \sigma_{\{u,v'\}}[t'',t']$
+	(\mahmoud{See figure \ref{fig:doubleCount}}). 
 	This means,
-	the cost of r.h.s.~to OPT is counted at most twice in each of  the expressions
+	$\mathit{Cost}_{OPT} ( \sigma_{\{u,v'\}}[t'',t'])$
+	is counted at most twice in each of  the expressions
 	$\mathit{Cost}_{\mathit{OPT}} (R_{t'}) + \mathit{Cost}_{\mathit{OPT}} (R_t)$

### 2020-05-07 paper/disc20clustering.tex
Removing Cascade

@@
  \newcommand\mahmoud[1]{\color{orange}\textbf{Mahmoud: #1~}\color{black}}
- \newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}
+ \newcommand\stefan[1]{\color{blue}\textbf{Stefan: #1}\color{black}}
  \newcommand\maciek[1]{\color{brown}\textbf{(Maciek: #1)}\color{black}}
+ %\newcommand\mahmoud[1]{}
+ %\newcommand\stefan[1]{}
  %\newcommand\maciek[1]{}
@@
 \end{proof}
-
-\section{Bounding the cost of a single reconfiguration of components}
-
-\maciek{Overview of the section, intro. Defer introducing the graph notation, the warm-up does not need it}
-
-\subsection{Warm-up: bounding the reconfiguration for $k=3$}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.3\textwidth]{figs/substitute}
-	\caption{Cluster signatures in warm-up k=3 proof}
-\end{figure}
-
-\maciek{Rewrite: cluster types should be cluster signatures}
-In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
-Before bounding the competitive ratio of \TAlg, we introduce the lemma that estimates the cost of a single repartition of \TAlg.
-
-\begin{lemma}
-	\label{lem:1req}
-	During a single repartition, \TAlg exchanges at most $2$ pairs of nodes.
-\end{lemma}
-
-\begin{proof}
-	Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
-	Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
-
-	%We distinguish between three types of clusters: $C_1, C_2, C_3$,
-	%which we define as follows.
-	%A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
-	%A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
-	%Finally, a cluster of type $C_3$  contains one component of size $3$.
-
-	Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
-	Note that $U\neq V$,
-	as otherwise, the request would not trigger a~repartitioning.
-	We consider cases based on the types of clusters $U$ and $V$.
-	Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
-	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
-	For two clusters, any configuration can be reached within two swaps, due to the fact that clusters are indistinguishable.
-
-	Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
-	Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
-	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-	Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
-In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
-\end{proof}
-
-Corollary: DET2 is $O(\ell)$-competitive for $k=3$.
-
-\subsection{Characterization of the migration graph}
-
-Consider a component merge action performed by \TAlg that triggers a reconfiguration from a configuration $C_I$ to a configuration $C_F$.
-To execute the reconfiguration, the nodes of two non-collocated components migrate to a common cluster.
-Each cluster contains exactly $k$ nodes, thus other nodes migrate to take the place of nodes of merged components.
-Their place must be occupied, too, and this way a single merge action may trigger migrations in clusters not directly involved in a merge action (for an example, see Section~\ref{ssec:example}). 
-Recall that both $C_I$ and $C_F$ are component-respecting configurations, hence a migration of a single node entails migrations of all nodes of its cluster.
-
-In Section~\ref{ssec:cascade}, we upper bound the cost incurred by a reconfiguration that follows a single component merge.
-To this end, we introduce a \emph{migration graph} that models the reconfiguration.
-In the following, we characterize the structure of the migration graph.
-
-\noindent
-\textbf{Vertices of the migration graph and the core vertices.}
-The vertices of the migration graph are all $\ell$ clusters of the instance.
-We distinguish the \emph{core vertices} that correspond to clusters directly involved in the merge operation: the clusters containing the to-be-merged components in $C_I$ and the cluster containing the merged component in $C_F$.
-There are at most $3$ core vertices in each migration graph (at most two components participate in the merge, and these may migrate to a third cluster).
-
-\noindent
-\textbf{Edges and their labels.}
-The edges of the migration graph denote the migration of nodes from $C_I$ to $C_F$.
-Each edge is labeled with a number $m$ that denotes the number of migrated nodes between clusters.
-Each edge is directed from the cluster that contained the nodes to the cluster they migrated to.
-Both edges might exists between a pair of nodes, as whole components of nodes migrate, and the back-and-forth exchange may be needed for the configuration $C_F$ to be component-respecting.
-
-\noindent
-\textbf{Vertex degree.}
-At most $k$ nodes may migrate from a cluster and at most $k$ nodes may migrate into a cluster.
-Thus, the sum of the labels for both ingoing and outgoing edges of each vertex is at most $k$.
-This implies that the number of ingoing and outgoing edges is also bounded by $k$.
-
-\noindent
-\textbf{Flow preservation.}
-In any feasible configuration, each cluster contains exactly $k$ nodes.
-Thus, for any vertex, the sum of labels of ingoing edges must equal the sum of labels of outgoing edges.
-\maciek{Can we say that this implies that the graph is a circulation?}
-
-\noindent
-\textbf{Migration graph and the cost of reconfiguration.}
-The cost of cluster reconfiguration equals the number of exchanged nodes multiplied by $\alpha$.
-From the standpoint of a migration graph, this corresponds to the sum of labels on edges of the graph.
-A trivial upper bound on the cost is the size of the connected component in the migration graph multiplied by $2\alpha k$.
-
-\medskip
-
-Note that for a single component merge, multiple migration graphs may exist, and multiple of them might have the optimal cost.
-In Section \ref{ssec:cascade}, we show that among all migration graphs for given component merge, there exists an optimal graph with the bounded cycle length.
-Combined with other properties, this allows to bound the size of the migration graph, and consequently bound the cost of each repartition, that results in a bound of the competitive ratio for \TAlg.
-
-\subsection{Costly cluster reconfiguration example}
-\label{ssec:example}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.3\textwidth]{figs/substitute}
-	\caption{A costly cascade example.
-On the left: the configuration prior to the merge and the merge of two components (denoted solid).
-On the right: corresponding migration graph for the optimal (unique) reconfiguration.}
-\end{figure}
-
-In this section we present an example of a merge action that results in a $k/3$(?) length cycle in an optimal migration graph.
-
-\subsection{Bounds on the size of a migration graph}
-\label{ssec:cascade}
-
-In \cite{repartition-disc}, authors trivially bounded the reconfiguration cost for each merge action by $k \cdot \ell$.
-This roughly corresponds to migrating every node in every cluster.
-Now, we bound the cost of a single repartition of \TAlg as a function of $k$.
-Note that the bound of $k \cdot \ell$ is still valid, and the resulting bound is the minimum of these two.
-
-We begin, by bounding the length of each cycle in one of optimal solutions.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.3\textwidth]{figs/substitute}
-	\caption{Illustration to bounding the cycle length.}
-	\label{fig:cascade-illustration}
-\end{figure}
-
-\begin{theorem}
-	\label{th:cascade-cycles}
-	Consider a single merge action of \TAlg.
-	There exists an optimal cost reconfiguration with the migration graph, where each cycle have length at most $k$.
-\end{theorem}
-
-\begin{proof}
-	Assume the contrary, i.e., for each optimal cluster reconfiguration, its migration graph contains a cycle at least $k+1$.
-
-	Fix a reconfiguration and its migration graph $G$ that has the shortest length $L$ of the longest cycle, and has the least number of cycles of length $L$.
-	Let $p = \lbrace v_1, v_2, \ldots, v_L \rbrace$ be any cycle of length $L$.
-	As edge labels are from $\{ 1, 2, \ldots, k \}$, there exists a pair of edges $e, f$ with equal labels.
-
-	We show that in such case, we can modify the solution without affecting its cost and feasibility to decrease the number of cycles of length $L$ by $1$.
-	This would contradict the assumption that we consider a solution with the minimal number of cycles of length $L$.
-
-	We construct a new migration graph $G'$ of equal cost by swapping the (destination) endpoints of $e$ and $f$ (cf. Figure~\ref{fig:cascade-illustration})
-	Formally, let $\lbrace a, b \rbrace = e$ and $\lbrace c, d \rbrace = f$.
-	We remove $e$ and $f$ and add $e' = \lbrace a, d \rbrace$ and $f' = \lbrace c, d \rbrace$.
-	This way we produced a feasible solution.
-
-	It remains to show that while decreasing the number of cycles of length $L$, we did not created any new cycles of length at least $L$.
-	Assume that adding $e'$ created a new cycle of length at least $L$.
-	However, $a$ and $d$ were already connected in $G$ by a directed path of length at least $2$, and a cycle strictly longer than $L$ existed in $G$.
-
-	Assume that adding $f'$ created a new cycle $s^*$ of length $L' \geq L$.
-	Contrary to the previous case, the directed cycle from $e$ to $b$ might not exist in $G$.
-	However, this would mean that $b$ and $e$ were connected by a path of length $L'-1$.
-	This means that we might have followed this alternative path from $b$ to $e$ to obtain a cycle strictly longer than $L$ (cf. Figure~\ref{fig:cascade-illustration}).
-
-\end{proof}
-
-\begin{theorem}
-	Consider a single repartition of DET2.
-	The number of clusters it involves is bounded by $k^k$.
-	\label{th:cascade}
-\end{theorem}
-
-\begin{proof}
-	Each node has degree at most $k$.
-	Every subset has flow conservation property, hence each cycle traverses one of the nodes of graph origin/core. (no subsets with just incoming edges to the origin part)
-	Combine with Theorem \ref{th:cascade-cycles}.
-
-\end{proof}
-
-Corollary: Algorithm DET2 is $O(\ell \cdot k \cdot \min \{ \ell, k^k \})$-competitive.
 

### 2020-05-07 paper/disc20clustering.tex
Cascade commented
notations, defs etc.

--- 2020-05-07 paper/disc20clustering.tex
@@
  \newcommand{\TAlg}{{\ensuremath{\textsf{ALG}_{3}}}\xspace}
+  \newcommand{\RM}{\textsf{RM}\xspace} % rematching alg
 
@@
 
-\section{Bounding the cost of a single reconfiguration of components}
-
-\maciek{Overview of the section, intro. Defer introducing the graph notation, the warm-up does not need it}
-
 \subsection{Warm-up: bounding the reconfiguration for $k=3$}
@@
 
+\begin{comment}
+\section{Bounding the cost of a single reconfiguration of components}
+
+\maciek{Overview of the section, intro. Defer introducing the graph notation, the warm-up does not need it}
+
 \subsection{Characterization of the migration graph}
@@
 Corollary: Algorithm DET2 is $O(\ell \cdot k \cdot \min \{ \ell, k^k \})$-competitive.
+\end{comment}
 
@@
 
-Algorithm Alg:
-\mahmoud{find a better name for Alg.}
-maintain a counter $C_{\set{u,v}}$ for each pair of nodes $\set{u,v}$ and increment it on every remote request between $u$ and $v$.
-Once $C_{\set{u,v}} = \lambda$,
-reset the counter and collocate the pair arbitrarily in one of the two clusters where $u$ or $v$ resides.
+We describe the algorithm \RM for the online rematching problem as follows.
+The input to \RM is a sequence of  requests
+$\sigma:=\{\sigma_1,\dots, \sigma_m\}, \sigma_t:=\{x,y\} \in V^2, x \neq y$.
+We maintain a counter $C_{\set{x,y}}$ for each pair of nodes $\set{x,y}$ and increment it on every remote request between $x$ and $y$.
+Once $C_{\set{x,y}} = \lambda$,
+reset the counter and collocate the pair arbitrarily in one of the two clusters where $x$ or $y$ resides.
 
 \begin{theorem} \label{thm:k=2}
-	The algorithm, for $\lambda=\alpha$, is (strictly) 6-competitive.
+	The algorithm \RM, for $\lambda=\alpha$, is (strictly) 6-competitive.
 \end{theorem}
 
-We charge both OPT and Alg whenever Alg collocates a pair (i.e., on a collocation event).
+We charge both \OPT and \RM whenever \RM collocates a pair (i.e., on a collocation event).
 %More precisely,we charge them only the cost of the collocation, per such event.
-%Hence,we charge the cost of  this collocation to OPT for the first time.
+%Hence,we charge the cost of  this collocation to \OPT for the first time.
 Alg collocates a pair always with a swap (i.e., two simultaneous migrations),
 which  costs $2\alpha$.
-%We charge OPT only for requests up to the one that triggers a collocation (i.e., not for later requests on the same pair).
+%We charge \OPT only for requests up to the one that triggers a collocation (i.e., not for later requests on the same pair).
 OPT may save some costs by collocating multiple pairs at once (before they incur too much cost),
 paying the price of only one migration per pair  (\mahmoud{see Figure \ref{fig:TBD}}).
-%This observation suggests that we can't charge OPT more than $\alpha$ per collocation.
+%This observation suggests that we can't charge \OPT more than $\alpha$ per collocation.
 %E.g.,
@@
 %it has both pairs on the cut of  the same two clusters.
-%Later, Alg collocates those pairs with two swaps.
+%Later, \RM collocates those pairs with two swaps.
 Hence,
-whenever OPT collocates a pair,
+whenever \OPT collocates a pair,
 we charge it only the cost of moving a single node to the cluster of the other,
@@
 
+Observe that for requests in
+$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $ 
+and the entire duration of the input sequence (i.e., $(0,m]$),
+OPT must pay a total cost
+%We denote the cost w.r.t~$r_t$ by
+$\mathit{OPT} (R_t) \geq 
+	c_{R_t} := \min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$,
+for serving (some of) these requests remotely or (and) collocating (one of) the pairs,
+regardless of its choices.
+Then the cost per pair, in particular the cost w.r.t.~$\{u,v\}$, is
+$\mathit{OPT} (\sigma_{\{u,v\}}) \geq c_{R_t} / 2$.
+%
+However,
+we can charge this cost to \OPT only at $t=t_0$,
+i.e.~the first time \RM collocates $\{u,v\}$,
+and not at any consequent time that \RM collocates $\{u,v\}$.
+That is,
+we charge \OPT a cost incurred by requests in
+$\sigma_{\{u,w\}}$ or $\sigma_{\{u,v\}}$,
+if and only if  \OPT incurs the cost during  intervals $(t'', t']$ or $(t', t]$,
+respectively.	
+Latter,
+this ensures us that we do not charge \OPT for a cost more than twice.
+%%
+We use notations defined so far in the proof that follows.
+
+\begin{proof}[Proof of Theorem \ref{thm:k=2}]
 	Assume on request $\sigma_{t}=\set{u,v}$, 
@@
 	denoted by $C^t_{\set{u,v}}$, 
-reaches $\lambda$ immediately before Alg resets the counter.
+	reaches $\lambda$ immediately before \RM resets the counter.
 	I.e.,
@@
 	during $[t_1,t_2]$ by $\sigma_{\set{x,y}}[t_1,t_2]$,
-and its overall cost to OPT by
-$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[t_1,t_2])$.
+	and its overall cost to \OPT by
+	$\mathit{OPT} (\sigma_{\{u,v\}}[t_1,t_2])$.
+	We may use $\sigma_{\set{x,y}}$ whenever
+	 the interval $[t_1,t_2]$ is clear from the context.
 	%
@@
 
-If $t$ is not the first time that Alg collocates $\{u,v\}$ then
-let $0 < t' < t$ be the latest time when Alg separates $\set{u,v}$
+	If $t$ is not the first time that \RM collocates $\{u,v\}$ then
+	let $0 < t' < t$ be the latest time when \RM separates $\set{u,v}$
 	in order to collocate some (intersecting) pair
@@
 	e.g.,
-$\{x,y\}=\{u,v'\}$.
+	$\{x,y\}=\{u,w\}$.
 	Else,
-$t$ is the first time that Alg collocates $\{u,v\}$ and let $t' := 0$.
+	$t$ is the first time that \RM collocates $\{u,v\}$ and let $t' := 0$.
 	Similarly,
-if $t' > 0$ is not the first time that Alg  collocates $\{u,v'\}$ 
-then let $0 < t'' < t'$ be the latest time before $t'$ when Alg separates $\set{u,v'}$.
+	if $t' > 0$ is not the first time that \RM  collocates $\{u,w\}$ 
+	then let $0 < t'' < t'$ be the latest time before $t'$ when \RM separates $\set{u,w}$.
 	Else,
-$t'$ is the first time that Alg collocates $\{u,v'\}$ and we let $t''=0$.
+	$t'$ is the first time that \RM collocates $\{u,w\}$ and we let $t''=0$.
 	%
 
-Observe that for requests in
-$R_t := \sigma_{\{u,v'\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t]$ 
-and the entire duration of the input sequence (i.e., $(0,m]$),
-OPT must pay a total cost
-%We denote the cost w.r.t~$r_t$ by
-$\mathit{Cost}_{\mathit{OPT}} (R_t) \geq 
-	c_{R_t} := \min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$,
-for serving (some of) these requests remotely or (and) collocating (one of) the pairs,
-regardless of its choices.
-Then the cost per pair, in particular the cost w.r.t.~$\{u,v\}$, is
-$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) \geq c_{R_t} / 2$.
-%
-However,
-we can charge this cost to OPT only at $t=t_0$,
-i.e.~the first time Alg collocates $\{u,v\}$,
-and not at any consequent time $\tilde{t} > t_0$ that Alg collocates $\{u,v\}$.
-%	i.e.,
-%	when Alg collocates the two pairs a second time.
-That is,
-we charge OPT a cost incurred by requests in
-$\sigma_{\{u,v'\}}(t'',t']$ or $\sigma_{\{u,v\}}(t',t]$,
-if and only if  OPT incurs the cost during  intervals $(t'', t']$ or $(t', t]$,
-respectively.	
-Latter,
-this ensures us that we do not charge OPT for a cost more than twice.
-%%
-We use notations defined so far in the proof that follows.
-
-\begin{proof}[Proof of Theorem \ref{thm:k=2}]
 	First,
-	we bound the costs incurred by Alg for requests that
+	we bound the costs incurred by \RM for requests that
 	lead to the collocation of $\{u,v\}$ at time $t \in T$, where
-	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$.
+	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$
+	is the set of times when \RM performs a collocation.
 	By definitions of $t$ and $t'$,
-	the overall cost of requests in $\sigma_{\set{u,v}}(t',t]$ incurred by Alg,
+	the overall cost of requests in $\sigma_{\set{u,v}}$ incurred by Alg,
 	i.e., the total cost of remote serving
@@
 	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
-	which means Alg pays $C^{m}_{ \{u,v\} }$
+	which means \RM pays $C^{m}_{ \{u,v\} }$
 	for these requests $\sigma_{\set{u,v}}(t',m]$.
-    Then the overall cost to Alg is
+    Then the overall cost to \RM is
 
 	\begin{align*}	%\label{eq:costAlg}
-	\mathit{Cost}_{\mathit{ALG}} (\sigma)
+	\mathit{ALG} (\sigma)
 	=
@@
 	Next,
-	we bound  costs incurred by OPT	for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
-	If $t$ is the first time that Alg collocates $\{u,v\}$ then  OPT pays
+	we bound  costs incurred by \OPT for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
+	If $t$ is the first time that \RM collocates $\{u,v\}$ then  \OPT pays
 	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
@@
 	Therefore in this case,
-	OPT pays
-	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
+	$\mathit{OPT} (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
 	%
-	By these definitions,
-	we have
-	$| \sigma_{\set{u,v'}}(t'',t'] | = | \sigma_{\set{u,v}}(t',t] |  =  \lambda$.
-	For requests $R_t := \sigma_{\set{u,v'}}(t'',t']  \cup \sigma_{\set{u,v}}(t',t]$,
-	OPT pays
-	$\mathit{Cost}_{\mathit{OPT}} (R_t) := \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v'\}}(t'',t']) 
-	+ \mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}(t',t]) $.
-	We distinguish three (non-exhaustive) cases for the cost to OPT.
+	Otherwise,
+	it is not the first collocation and
+	consider times $t'$ and $t''$ as define previously,
+	 and let 
+	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
+	We define $R_{t'}$ for the collocation at $t'$ in an analogous way.
+	Then,
+	$\mathit{OPT} (R_t) = \mathit{OPT} (\sigma_{\{u,w\}}) 
+	+ \mathit{OPT} (\sigma_{\{u,v\}}) $.
+	We distinguish three (non-exhaustive) cases for the cost to \OPT w.r.t.~$R_t$.
 	Latter,
-	 we argue that OPT cannot save cost by deviating from these cases.
+	 we argue that \OPT cannot save cost by deviating from these cases.
 	\begin{enumerate}[label=\roman*.]
 		\item
-		OPT collocates $\{u,v'\}$ at  $t''$, $\{u,v\}$  at $t'$,
+		\OPT collocates $\{u,w\}$ at  $t''$, $\{u,v\}$  at $t'$,
 		and keeps them so during $(t'',t']$ and $(t',t]$, respectively.
 		Therefore it serves all requests in $R_t$ with no additional cost and
-		$ c_1 (R_t) := 2\alpha$.
+		let $\OPT_1 (R_t) := 2\alpha$.
 		\item
-		OPT has $\{u,v\}$ ($\{u,v'\}$) already collocated at $t'$ ($t''$) as well as during $(t',t]$ ($(t'',t]$),
-		and has $\{u,v'\}$ ($\{u,v\}$) separated during $(t'',t']$ ($(t',t]$).
+		\OPT has one of the pairs  already collocated and the other pair separated,
+			  prior to and during their respective intervals.
 		Hence,
-		it pays only $\lambda$ for serving all requests in $R_t$ and
-		$c_2 (R_t) := \lambda$
+		it pays only $\lambda$ for serving all requests in $R_t$
+		and let $\OPT_2 (R_t) := \lambda$.
 		\item
-		OPT has $\{u,v'\}$ ($\{u,v\}$) already collocated at $t''$ ($t'$)  and also during $(t'',t']$ ($(t', t])$.
-		Then it pays 0 for $\sigma_{\{u,v'\}}[t'',t']$ ($(\sigma_{\{u,v\}}[t',t]$), and
-		$c_3 (R_t) := \alpha$.
+		\OPT has one of the pairs already collocated prior to its interval,
+		and it collocates the other pair during the respective interval.
+		Then it pays 0 for one pair and $\alpha$ for the other and 
+		let $\OPT_3 (R_t) := \alpha$.
 	\end{enumerate}
-	If OPT's choice is one of these cases,
+%	We consider cases (ii) and (iii) also for the pair $\{u,w\}$, symmetrically.
+	If \OPT behaves in one of these cases,
 	then
-	$\mathit{Cost}_{\mathit{OPT}} (R_t) \geq
-		\min{\{ c_1(R_t), c_2(R_t), c_3(R_t) \}}  =
+	$\mathit{OPT} (R_t) \geq
+		\min{\{ \OPT_1(R_t), \OPT_2(R_t), \OPT_3(R_t) \}}  =
 		\min{ \{ \lambda, \alpha \}}$.
 	Otherwise,
-	OPT deviates from these cases,
+	\OPT deviates from these cases,
 	e.g.~by collocating a pair after paying for some of requests to this pair.
 	However,
-	no deviation can decrease the cost to OPT.
+	no deviation can decrease the cost to \OPT. \mahmoud{prove this!}
 	Therefore in any case,
-	$\mathit{Cost}_{\mathit{OPT}} (R_t) \geq  \min{ \{ \lambda, \alpha \}}$.
-
-	It remains to bound the cost  incurred by OPT for requests to $\{u,v\}$ that do not lead to its collocation until the end of the sequence at $t=m$.
-	If $\{u,v\}$ is not collocated in the initial configuration
-	and Alg never collocates it,
+	$\mathit{OPT} (R_t) \geq  \min{ \{ \lambda, \alpha \}}$.
+
+	It remains to bound the cost  incurred by \OPT for requests to $\{u,v\}$ that do not lead to its collocation until the end of the sequence at $t=m$.
+	If $\{u,v\}$ is not collocated in the initial matching
+	and \RM never collocates it,
 	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[1,m] |$.
-	OPT pays
-	$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[1,m]) 
+	\OPT pays
+	$\mathit{OPT} (\sigma_{\{u,v\}}[1,m]) 
 	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
@@
 	Else,
-	either $\{u,v\}$ is collocated in the initial configuration
-	or Alg collocates it at some point.
-	Then there exists an intersecting pair $\set{u,v'}$
-	that is collocated by Alg at $t' < m$ (separating $\{u,v\}$).
-	We define times $t'' < t' < m$ similarly to the case where Alg collocates $\{u,v\}$.
-	Let $R_{\{u,v\}} := \sigma_{\{u,v'\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
+	either $\{u,v\}$ is collocated in the initial matching
+	or \RM collocates it at some point.
+	Then there exists an intersecting pair $\set{u,w}$
+	that is collocated by \RM at $t' < m$ (separating $\{u,v\}$).
+	We define times $t'' < t' < m$ similarly to the case where \RM collocates $\{u,v\}$.
+	Let $R^*_{\{u,v\}} := \sigma_{\{u,w\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
 	Then,
@@
 	we obtain
-	$\mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}}) 
+	$\mathit{OPT} (R^*_{\{u,v\}}) 
 	\geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.
 
-	Next, we sum up all costs incurred by OPT.
-	By definitions of $R_t$ and $R_{\{u,v\}}$, we have either
-	$R_{t'} \cap R_t = \sigma_{\{u,v'\}}[t'',t']$ or
-	$R_{t'} \cap R_{\{u,v\}} = \sigma_{\{u,v'\}}[t'',t']$
+	Next, we sum up all costs incurred by \OPT.
+	By definitions of $R_t$ and $R^*_{\{u,v\}}$, we have either
+	$R_{t'} \cap R_t = \sigma_{\{u,w\}}$ or
+	$R_{t'} \cap R^*_{\{u,v\}} = \sigma_{\{u,w\}}$
 	(\mahmoud{See figure \ref{fig:doubleCount}}). 
 	This means,
-	$\mathit{Cost}_{OPT} ( \sigma_{\{u,v'\}}[t'',t'])$
+	$\mathit{Cost}_{OPT} ( \sigma_{\{u,w\}})$
 	is counted at most twice in each of  the expressions
-	$\mathit{Cost}_{\mathit{OPT}} (R_{t'}) + \mathit{Cost}_{\mathit{OPT}} (R_t)$
+	$\mathit{OPT} (R_{t'}) + \mathit{OPT} (R_t)$
 	and  
-	$\mathit{Cost}_{\mathit{OPT}} (R_{t'}) + \mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}})$.
+	$\mathit{OPT} (R_{t'}) + \mathit{OPT} (R^*_{\{u,v\}})$.
 	Hence,
@@
 	and for final costs at $t=m$,
-	OPT pays at least 
+	\OPT pays at least 
 	$\frac{1}{2}(
-	\sum_{ t \in [1,m] } \mathit{Cost}_{\mathit{OPT}} (R_t) +
-	\sum_{\{u,v\}} \mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}})
+	\sum_{ t \in [1,m] } \mathit{OPT} (R_t) +
+	\sum_{\{u,v\}} \mathit{OPT} (R^*_{\{u,v\}})
 	) $.
 	Then,
-	the total cost to OPT is
+	the total cost to \OPT is
 
 	\begin{align*} 	%\label{eq:costOPT}
-		\mathit{Cost}_{\mathit{OPT}} (\sigma)
+		\mathit{OPT} (\sigma)
 		&=
@@
 		\Big(
-		\sum_{ t \in T} \mathit{Cost}_{\mathit{OPT}} (R_t) 
-		+ \sum_{\{u,v\}}\mathit{Cost}_{\mathit{OPT}} (R_{\{u,v\}})
+		\sum_{ t \in T} \mathit{OPT} (R_t) 
+		+ \sum_{\{u,v\}}\mathit{OPT} (R^*_{\{u,v\}})
 		\Big)	\\
@@
 \begin{align*}
-	\mathit{Cost}_{\mathit{ALG}} (\sigma)	/
-	\mathit{Cost}_{\mathit{OPT}} (\sigma)
+	\mathit{ALG} (\sigma)	/
+	\mathit{OPT} (\sigma)
 	\leq
@@
 \maciek{I feel slightly uneasy here: "In particular,
-we charge OPT a cost incurred by requests in
-$\sigma_{\{u,v'\}}(t'',t']$ and $\sigma_{\{u,v\}}(t',t]$,
-if and only if  OPT incurs the cost within  intervals $(t'', t']$ and $(t', t]$,
+we charge \OPT a cost incurred by requests in
+$\sigma_{\{u,w\}}$ and $\sigma_{\{u,v\}}$,
+if and only if  \OPT incurs the cost within  intervals $(t'', t']$ and $(t', t]$,
 respectively.
 \mahmoud{this is only for intuition but a key statement.
-It is meant to say: OPT may pay cost for the same two pairs outside the intervals.
-But we don't charge OPT  those costs at $t$.
-This way we avoid over charging OPT as only intervals $r_{t'}$ and $r_t$ can overlap which leads to double charging and we take care of it.}
+It is meant to say: \OPT may pay cost for the same two pairs outside the intervals.
+But we don't charge \OPT  those costs at $t$.
+This way we avoid over charging \OPT as only intervals $r_{t'}$ and $r_t$ can overlap which leads to double charging and we take care of it.}
 "}
@@
 
-\maciek{"We charge OPT the cost
-$\mathit{Cost}_{\mathit{OPT}} (\sigma_{\{u,v\}}[1,t]) 
+\maciek{"We charge \OPT the cost
+$\mathit{\OPT} (\sigma_{\{u,v\}}[1,t]) 
 \geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
@@
 \maciek{The big inequality is impressive, but not a lot change in each row.
-Instead, bound the cost of ALG separately in one line, and the cost of OPT in another.
+Instead, bound the cost of ALG separately in one line, and the cost of \OPT in another.
 (possibly far away from eachother) \mahmoud{done}}
@@
 3 parts: first collocation, regular collocation and unfinished.
-Do this for ALG and then for OPT, and do not mix (unless I don't see some dependencies).
+Do this for ALG and then for \OPT, and do not mix (unless I don't see some dependencies).
 Start each paragraph by the sentence that says what the paragraph is about (first/opt, regular/alg etc.).
@@
 \maciek{$R_t$ is the central concept in this proof.
-Possibly even start with the paragraph: charging OPT, regular collocation.
+Possibly even start with the paragraph: charging \OPT, regular collocation.
 Make this definition in its separate line to further emphasize?
@@
 \maciek{Consider naming the adjacent edges $e$ and $f$.}
-
-\begin{comment}
-	We charge both OPT and Alg whenever Alg collocates a pair (i.e., on a collocation event).
-	%More precisely,we charge them only the cost of the collocation, per such event.
-	%Hence,we charge the cost of  this collocation to OPT for the first time.
-	Alg collocates a pair always with a swap (i.e., two simultaneous migrations),
-	which  costs $2\alpha$.
-	%We charge OPT only for requests up to the one that triggers a collocation (i.e., not for later requests on the same pair).
-	OPT may save some costs by collocating multiple pairs at once (before they incur too much cost),
-	paying the price of only one migration per pair.
-	This observation suggests that we can't charge OPT more than $\alpha$ per collocation,
-	even when OPT actually pays more than that.
-	E.g.,
-	OPT can collocate two costly pairs with a single swap if
-	it has both pairs on the cut of  the same two clusters.
-	Later, Alg collocates those pairs with two swaps (see Figure \ref{fig:TBD}).
-	Hence,
-	whenever OPT collocates a pair,
-	we charge OPT only the cost of moving a single node to the cluster of the other,
-	i.e., $\alpha$ (in contrast to $2\alpha$ for Alg).
-
-	\begin{proof}[Proof of Theorem \ref{thm:k=2}]
-		\mahmoud{not correct}
-		%	Let $C^{t}_{\set{u,v}}$ denote the  value of $C_{\set{u,v}}$
-		%	immediately after arrival of the $t$th request,
-		%	$\sigma_t$.
-		Assume on request $\sigma_{t}=\set{u,v}$, 
-		i.e., at (time) $t \geq 1$,
-		Alg collocates the pair $\set{u,v}$.
-		This means,
-		the value of $C_{\set{u,v}}$ at $t$,
-		immediately before Alg resets the counter is $\lambda$.
-		That is, $ C^t_{\set{u,v}} = \lambda$.
-		Let $\sigma_{t''}, t''<t$ be the last request (before $\sigma_{t}$)
-		upon which Alg collocates $\set{u,v}$.
-		Let $t'$, $t'' < t' < t$ be the  time when Alg separates $\set{u,v}$
-		in order to collocate some other pair that shares a node with $\set{u,v}$,
-		e.g.~$\set{u,v'}, v' \neq v$.
-		In case the first collocation of $\set{u,v}$ occurs at $t$,
-		if the pair is collocated in the initial configuration then
-		we set $t''=0$,
-		else,
-		we set $ t'' = -\infty, t'=0$.
-		We denote the set of all requests to $\set{u,v}$
-		during $[t_1,t_2]$ by $\sigma_{\set{u,v}}[t_1,t_2]$.
-		Thus,
-		$| \sigma_{\set{u,v}}[t',t] | = \lambda$.
-		The overall cost of requests that increase the counters of intersecting pairs,
-		%	i.e.~requests in $\sigma_{\set{u,v}}[t'',t]$,
-		and the moving cost,
-		incurred by Alg is
-		$\lambda + 2\alpha$.
-		%That is, the cost of all requests in $\sigma_{\set{u,v}}[t',t]$.
-
-		We distinguish three cases for the cost incurred by OPT with respect to the collocation of $\set{u,v}$ at $t$.
-
-		\begin{enumerate}
-			\item Case i)
-			OPT has $\set{u,v}$ separated (in OPT's configuration) during $[t',t]$
-			and therefore it pays $\lambda$.
-			We charge OPT $\lambda/2$, i.e.,
-			$\mathit{Cost}^{t}_{\{u,v\}} (\mathit{OPT})= \lambda / 2$.
-
-			\item Case ii)  
-			OPT collocates $\set{u,v}$ some time during $[t'',t]$ and pays $\alpha$.
-			We charge it $\alpha/2$, i.e.,
-			$\mathit{Cost}^{t}_{\{u,v\}} (\mathit{OPT}) = \alpha / 2$.
-
-			\item Case iii)
-			$t'' \geq 0$ and OPT has $\set{u,v}$ already collocated before $t''$.
-			%	 and keeps them collocated during $[t'',t]$.
-			In this case,
-			OPT possibly pays zero for requests in $\sigma_{\set{u,v}}[t'',t]$.
-			However,
-			by assumption,
-			Alg collocates $\{u,v'\}$ at $t'$ and
-			OPT doesn't have $\{u,v'\}$ collocated through the entire
-			$[\tau^{t''}_{\{u,v'\}}, t']$.
-			This implies that only cases (i) or (ii) applies to $\{u,v'\}$ at $t'$.
-			That is,
-			OPT pays $\min \{ \lambda, \alpha \}$,
-			and it is charged $\min \{ \lambda/2, \alpha/2 \}$
-			w.r.t.~the collocation of $\{u, v'\}$ at $t'$.
-			Hence,
-			we can charge it  $\min \{ \lambda/2, \alpha/2 \}$ in addition to the  cost already charged.
-			%	i.e.~the remaining (unaccounted) cost of requests in $\sigma_{\set{u,v'}}[t'',t']$.
-			%	Intuitively, at $t$,  we charge $\alpha$ to OPT for separating $\set{u,v}$ (at $t'$) by Alg.
-		\end{enumerate}
-
-		It can be easily verified that any deviation from these cases only increases the cost to OPT.
-		Thus,
-		OPT pays at least $\alpha$ for each collocation performed by Alg.
-		%	which gives the competitive ratio 
-		%	$(\lambda + 2\alpha)/ \alpha = 4$.
-
-		It remains to charge both Alg and OPT for requests that do not lead to collocations when the sequence ends.
-		Consider  pairs $\set{u,v}$ and $\set{u,v'}$ and
-		times $t'' < t' < t = t_m$ as defined previously.
-		Assume $\sigma_{\set{u,v}}[t'', t_m] \leq \lambda - 1$.
-		This means Alg has the pair separated at $t_m$
-		and paid the (communication) cost $r_{\set{u,v}} := | \sigma_{\set{u,v}}[t'', t_m]|$.
-		If OPT has the pair separated during $[t'',t_m]$ then it pays $r_{\set{u,v}}$.
-		If OPT collocates the pair some time during $[t'', t]$ then analogous to Case (ii),
-		we charge it $\alpha$.
-		Otherwise,
-		if OPT collocates the pair before $t''$,
-		then we charge it $\alpha$ analogously to Case (iii).
-		Hence, 
-		$Cost(OPT)_{\set{u,v}} / Cost(Alg)_{\set{u,v}} =
-		\max \set{ r_{\set{u,v}}/ \alpha \leq (\lambda-1) / \alpha , (\lambda + 2\alpha)/ \alpha} = 4$.
-		%	The claim follows from here.
-	\end{proof}
-\end{comment}
 

### 2020-05-07 paper/disc20clustering.tex
the intuition paragragh and more

@@
  \newcommand{\OPT}{\textsf{OPT}\xspace}
+  \newcommand{\OPTM}{\mathit{OPT}}
  \newcommand{\ALG}{\textsf{ALG}\xspace}
@@
 %Hence,we charge the cost of  this collocation to \OPT for the first time.
-Alg collocates a pair always with a swap (i.e., two simultaneous migrations),
-which  costs $2\alpha$.
-%We charge \OPT only for requests up to the one that triggers a collocation (i.e., not for later requests on the same pair).
-OPT may save some costs by collocating multiple pairs at once (before they incur too much cost),
-paying the price of only one migration per pair  (\mahmoud{see Figure \ref{fig:TBD}}).
-%This observation suggests that we can't charge \OPT more than $\alpha$ per collocation.
-%E.g.,
-%OPT can collocate two costly pairs with a single swap if
-%it has both pairs on the cut of  the same two clusters.
-%Later, \RM collocates those pairs with two swaps.
+\RM collocates a pair always with a swap (i.e., two simultaneous migrations),
+which  costs $2\alpha$,
+while OPT may save some costs by collocating multiple pairs at once, 
+ before they inflict too much communication cost.
+ Therefore it pays the price of only one migration per pair  (\mahmoud{see Figure \ref{fig:TBD}}).
 Hence,
@@
 we charge it only the cost of moving a single node to the cluster of the other,
-i.e., $\alpha$ (in contrast to $2\alpha$ on Alg).
-
-Observe that for requests in
-$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $ 
-and the entire duration of the input sequence (i.e., $(0,m]$),
-OPT must pay a total cost
-%We denote the cost w.r.t~$r_t$ by
-$\mathit{OPT} (R_t) \geq 
-	c_{R_t} := \min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$,
-for serving (some of) these requests remotely or (and) collocating (one of) the pairs,
-regardless of its choices.
-Then the cost per pair, in particular the cost w.r.t.~$\{u,v\}$, is
-$\mathit{OPT} (\sigma_{\{u,v\}}) \geq c_{R_t} / 2$.
+i.e., $\alpha$ (in contrast to $2\alpha$ on \RM).
+
+\textbf{The charging Scheme.}
+Consider two  pairs that share a same node, 
+i.e.~\emph{intersecting pairs},
+and the set of requests that causes the collocation of both pairs,
+at some  times during  $[1,m]$.
+%i.e.~the entire duration of the sequence.
+Observe that \OPT must pay a non-zero cost
+for these requests over the entire $\sigma$,
+since it cannot have both pairs initially collocated.
+%More precisely,
+%OPT  pays a total cost at least
+%	$\min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$ for, respectively,
+%	serving requests to both pairs remotely,
+%	serving requests to one pair remotely and collocating the other,
+%	or collocating both pairs. 
+%
+%for serving (some of) these requests remotely or (and) collocating (one of) the pairs,
+%regardless of its choices.
+%Then the cost per pair, in particular the cost w.r.t.~$\{u,v\}$, is
+%$\mathit{OPT} (\sigma_{\{u,v\}}) \geq c_{R_t} / 2$.
 %
 However,
-we can charge this cost to \OPT only at $t=t_0$,
-i.e.~the first time \RM collocates $\{u,v\}$,
-and not at any consequent time that \RM collocates $\{u,v\}$.
-That is,
-we charge \OPT a cost incurred by requests in
-$\sigma_{\{u,w\}}$ or $\sigma_{\{u,v\}}$,
-if and only if  \OPT incurs the cost during  intervals $(t'', t']$ or $(t', t]$,
-respectively.	
-Latter,
-this ensures us that we do not charge \OPT for a cost more than twice.
+we can charge the whole cost to \OPT only the first time \RM collocates a pair,
+and not at any consequent time when \RM collocates it a second time.
+Otherwise,
+ \OPT is possibly charged for a same cost repeatedly.
+For this reason,
+we charge \OPT a cost inflicted by a pair,
+if and only if  \OPT incurs the cost some time since the last time \RM separated the pair.
 %%
-We use notations defined so far in the proof that follows.
 
@@
 	let $0 < t' < t$ be the latest time when \RM separates $\set{u,v}$
-	in order to collocate some (intersecting) pair
+	in order to collocate some intersecting pair
 	$\{x,y\} \neq \{u,v\}, \{x,y\} \cap \{u,v\} \neq \emptyset$, 
@@
 	First,
-	we bound the costs incurred by \RM for requests that
+	we bound  costs incurred by \RM for requests that
 	lead to the collocation of $\{u,v\}$ at time $t \in T$, where
@@
 	By definitions of $t$ and $t'$,
-	the overall cost of requests in $\sigma_{\set{u,v}}$ incurred by Alg,
+	the overall cost of requests in $\sigma_{\set{u,v}}$ incurred by \RM,
 	i.e., the total cost of remote serving
@@
 	Next,
-	we bound the cost incurred by Alg
+	we bound costs incurred by \RM
 	for requests that do not lead to collocations until the  end of the sequence at $t=m$.
@@
 	which means \RM pays $C^{m}_{ \{u,v\} }$
-	for these requests $\sigma_{\set{u,v}}(t',m]$.
+	for  requests in $\sigma_{\set{u,v}}(t',m]$.
     Then the overall cost to \RM is
-
-	\begin{align*}	%\label{eq:costAlg}
-	\mathit{ALG} (\sigma)
+$	\mathit{RM} (\sigma)
 =
@@
 \sum_{\{u,v\}} C^{m}_{\{u,v\}}	
-	\leq
-	\sum_{ t \in T}(\lambda + 2\alpha) +
-	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
-	=
-	\sum_{ t \in T} 3\alpha +
-	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
-	.
-	\end{align*}
+$.
+%	\begin{align*}	%\label{eq:costAlg}
+%	\mathit{RM} (\sigma)
+%	=
+%	\sum_{ t \in T}(\lambda + 2\alpha) +
+%	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
+%	\leq
+%	\sum_{ t \in T}(\lambda + 2\alpha) +
+%	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
+%	=
+%	\sum_{ t \in T} 3\alpha +
+%	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
+%	.
+%	\end{align*}
 
@@
 	Therefore in this case,
-	$\mathit{OPT} (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
+	$\OPTM (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
 	%
@@
 	Then,
-	$\mathit{OPT} (R_t) = \mathit{OPT} (\sigma_{\{u,w\}}) 
-	+ \mathit{OPT} (\sigma_{\{u,v\}}) $.
+	$\OPTM(R_t) = \mathit{OPT} (\sigma_{\{u,w\}}) 
+	+ \OPTM(\sigma_{\{u,v\}}) $.
 	We distinguish three (non-exhaustive) cases for the cost to \OPT w.r.t.~$R_t$.
@@
 		Therefore it serves all requests in $R_t$ with no additional cost and
-		let $\OPT_1 (R_t) := 2\alpha$.
+		let $\OPTM_1 (R_t) := 2\alpha$.
 		\item
@@
 		it pays only $\lambda$ for serving all requests in $R_t$
-		and let $\OPT_2 (R_t) := \lambda$.
+		and let $\OPTM_2 (R_t) := \lambda$.
 		\item
@@
 		Then it pays 0 for one pair and $\alpha$ for the other and 
-		let $\OPT_3 (R_t) := \alpha$.
+		let $\OPTM_3 (R_t) := \alpha$.
 	\end{enumerate}
@@
 	$\mathit{OPT} (R_t) \geq
-		\min{\{ \OPT_1(R_t), \OPT_2(R_t), \OPT_3(R_t) \}}  =
+		\min{\{ \OPTM_1(R_t), \OPTM_2(R_t), \OPTM_3(R_t) \}}  =
 		\min{ \{ \lambda, \alpha \}}$.
@@
 	This means,
-	$\mathit{Cost}_{OPT} ( \sigma_{\{u,w\}})$
+	$\OPTM ( \sigma_{\{u,w\}})$
 	is counted at most twice in each of  the expressions
@@
 	Hence,
-	for all collocations performed by Alg,
+	for all collocations performed by \RM,
 	and for final costs at $t=m$,
@@
 	$\frac{1}{2}(
-	\sum_{ t \in [1,m] } \mathit{OPT} (R_t) +
+	\sum_{ t \in T } \mathit{OPT} (R_t) +
 	\sum_{\{u,v\}} \mathit{OPT} (R^*_{\{u,v\}})
@@
 		\Big)	\\
-		&\leq
+		&\geq
 		\Big(

### 2020-05-08 paper/disc20clustering.tex
Intro

--- 2020-05-07 paper/disc20clustering.tex
@@
 
- \title{Online Balanced Graph Partitioning: Optimizations for Large Clusters}
+ \title{Tigth Bounds for Online Balanced Graph Partitioning}
 
@@
 	or machine learning, generate a significant amount of network traffic.
-By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), the network load can be reduced and application performance improved.
-	However, as the communication pattern is a priori unknown and may change over time, it needs to be learned efficiently, in an online manner.
+	By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), we can reduce the network load and  improve application performance. 
+	However, the communication pattern is a priori unknown and may change over time, hence it needs to be learned in an online manner.
 	%
@@
 	of $O(k\cdot \ell)$ for this scenario.
-	In addition, we present a tight upper bound
+	We contribute a tight upper bound
 	of $\Theta(\ell)$ for $k=3$,
@@
 	over time.
+	We improve the result for $k=2$ by providing a strictly $6$-competitive upper bound for the general model.
+
 \end{abstract}
@@
 
-\maciek{We study two models. One is a subproblem.}
-
-The general model, where $\sigma$ can be arbitrary
-was the original problem introduced
-by Avin et al.~\cite{repartition-disc} at DISC 2016.
-For simplicity, we will refer to the Perfect Partition
-model as the \emph{learning} model (as
-one has to learn the static communication graph) 
-and to the original as the \emph{repartitioning} model (as one continuously adjusts the partition in response to changing communication graph).
-
-\subsubsection{Partitioning model.}
+We study two models in this paper: a general \emph{partitioning} model, and its subproblem, a \emph{learning} variant.
+The general partitioning model was introduced by Avin et al.~\cite{repartition-disc} at DISC 2016.
+The learning model was introduced by Henzinger et al.~\cite{sigmetrics19_partitioning} at SIGMETRICS 2019.
+
+\noindent
+\textbf{Partitioning model.}
 The \emph{balanced repartitioning} problem (\OBRP{})
@@
 
+\subsubsection{Learning model.}
+
+We consider a variant of Online Balanced Partitioning, where communication requests are permanent: once a pair of nodes start communicating, they continue forever.
+To avoid infinite cost, each communicating pair must be collocated.
+We study instances with finite cost: we assume that the communication graph can be perfectly partitioned in the clusters (no external requests are required).
+The algorithm's goal is to \emph{learn} the communication graph, and partition it without too many reconfigurations.
+
+\noindent
+\textbf{Adversarial input and the competitive analysis.}
 \maciek{The competitive ratio definition and the worst case inputs}
 
-\maciek{Ilustrate the challenges. Rent-or-buy. Decision what to evict}
-
-\subsubsection{Learning model.}
-In addition to the general model, we study a simplified Perfect Partition model (also studied by Henzinger et al.~\cite{sigmetrics19_partitioning} (SIGMETRICS 2019)) that we consider for a stronger lower bound.
-There, we assume that $\sigma$
-reveals the edges of a static graph
-that can be perfectly partitioned
-(no external requests are required).
+\noindent
+\textbf{Challenges.}
+The partitioning model has a rent-or-buy dilemas. The decision, what to evict is present in both models.
 
@@
 
-\maciek{Offline results}
-\maciek{Links other problems: for $k=2$ a flavor of online matching, for $\ell = 2$ it generalizes paging}
+\maciek{Below is similar to SIDMA, possibly too similar}
+
+The static offline version of~the~partitioning~problem, i.e., a problem variant where
+migration is not allowed, where all requests are known in advance, and where
+the goal is to find an assignment of $n$ nodes to $\ell$~physical machines, each of~capacity $n/\ell$, is known as the
+\emph{$\ell$-balanced graph partitioning problem}. The problem is 
+NP-complete, and cannot even be approximated within any finite factor unless P
+= NP~\cite{AndRae06}.  The static
+variant where $\ell = 2$ corresponds to the minimum bisection problem, which
+is already NP-hard~\cite{GaJoSt76}, and 
+the currently best approximation ratio is $O(\log n)$~\cite{SarVaz95,ArKaKa99,FeKrNi00,FeiKra02,KraFei06,Raec08}.
+The inapproximability of the static variant for general values of $\ell$
+motivated research on the bicriteria variant, which can be seen as the offline
+counterpart of our capacity augmentation approach. Here, the~goal
+is~to~compute a~graph partitioning into $\ell$ components of~size at most~$k$ (where $k > n/\ell$) and the cost of the cut is compared to the optimal (non-augmented)
+solution where all components are of a~size at most $n/k$. The variant where
+$n \geq 2 \cdot k \cdot \ell$ was considered in
+\cite{LeMaTr90,SimTen97,EvNaRS00,EvNaRS99,KrNaSc09}. So far, the~best result~is~an~$O(\!\sqrt{\log n \cdot \log \ell})$-approximation algorithm~\cite{KrNaSc09}.
+
+The studied model is related to online
+caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}, sometimes also referred to
+as online caching, where requests for data items (nodes) arrive over time and
+need to be served from a cache of finite capacity, and where the number of
+cache misses must be minimized. Classic problem variants usually boil down to
+finding a smart eviction strategy, such as Least Recently Used (LRU)~\cite{SleTar85}. In our
+setting, requests can be served remotely (i.e.,~without fetching the
+corresponding nodes to a single physical machine). In this light, our model is more
+reminiscent of caching models \emph{with
+bypassing}~\cite{EpImLN11,EpImLN15,Irani02}. As a~side result, we show that our problem is
+capable of emulating online caching.
+A major difference between  these problems is that in the caching problems, each request involves a~single element of the universe, while in our model \emph{both} endpoints of a communication request are subject to~optimization.
 
@@
 an asymptotically optimal,
-$\Theta(\ell)$ algorithm for $k=3$;
-the best known upper bound 
-so far was $O(\ell^2)$~\cite{repartition-disc}.
-Then, we present a strictly $6$-competitive algorithm for $k=2$ that improves upon previous $7$-competitive algorithm with $O(\ell^2)$ additive constant..
-In addition, we show a new, improved analysis of the algorithm from \cite{repartition-disc} that depends on the solutions to an interesting combinatorial problem (see Section~\ref{ssec:cascade}, The Cascade Problem).
+$\Theta(\ell)$ algorithm for $k=3$, improving the best known upper bound 
+so far $O(\ell^2)$~\cite{repartition-disc}.
+Then, we present a strictly $6$-competitive algorithm for $k=2$ that improves upon previous $7$-competitive algorithm with $O(\ell^2)$ additive constant.
 %
@@
 		\textbf{$k=3$}&  $\Omega(\ell)$ \hspace{0.3cm}(\S \ref{sec:lowerbound})& $O(\ell) $\hspace{0.3cm}(\S \ref{sec:k3})\\
-		$k > 3$ & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell \cdot \min \{ k \cdot \ell, \Rep(k, \ell) \})$\hspace{0.1cm} (\S \ref{sec:k3}) \\
+		$k > 3$ & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k^2 \cdot \ell^2)$\hspace{0.1cm} \cite{repartition-disc} \\
 		\rowcolor{gray!25}
@@
 \end{proof}
+
+\subsection{Warm-up: bounding the reconfiguration for $k=3$}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.3\textwidth]{figs/substitute}
+	\caption{Cluster signatures in warm-up k=3 proof}
+\end{figure}
+
+\maciek{Rewrite: cluster types should be cluster signatures}
+In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
+Before bounding the competitive ratio of \TAlg, we introduce the lemma that estimates the cost of a single repartition of \TAlg.
+
+\begin{lemma}
+	\label{lem:1req}
+	During a single repartition, \TAlg exchanges at most $2$ pairs of nodes.
+\end{lemma}
+
+\begin{proof}
+	Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
+	Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
+
+	%We distinguish between three types of clusters: $C_1, C_2, C_3$,
+	%which we define as follows.
+	%A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+	%A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
+	%Finally, a cluster of type $C_3$  contains one component of size $3$.
+
+	Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
+	Note that $U\neq V$,
+	as otherwise, the request would not trigger a~repartitioning.
+	We consider cases based on the types of clusters $U$ and $V$.
+	Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
+	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
+	For two clusters, any configuration can be reached within two swaps, due to the fact that clusters are indistinguishable.
+
+	Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
+	Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
+	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+	Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
+In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
+\end{proof}
+
+Corollary: DET2 is $O(\ell)$-competitive for $k=3$.
+
+\begin{comment}
+\section{Bounding the cost of a single reconfiguration of components}
+
+\maciek{Overview of the section, intro. Defer introducing the graph notation, the warm-up does not need it}
+
+\subsection{Characterization of the migration graph}
+
+Consider a component merge action performed by \TAlg that triggers a reconfiguration from a configuration $C_I$ to a configuration $C_F$.
+To execute the reconfiguration, the nodes of two non-collocated components migrate to a common cluster.
+Each cluster contains exactly $k$ nodes, thus other nodes migrate to take the place of nodes of merged components.
+Their place must be occupied, too, and this way a single merge action may trigger migrations in clusters not directly involved in a merge action (for an example, see Section~\ref{ssec:example}). 
+Recall that both $C_I$ and $C_F$ are component-respecting configurations, hence a migration of a single node entails migrations of all nodes of its cluster.
+
+In Section~\ref{ssec:cascade}, we upper bound the cost incurred by a reconfiguration that follows a single component merge.
+To this end, we introduce a \emph{migration graph} that models the reconfiguration.
+In the following, we characterize the structure of the migration graph.
+
+\noindent
+\textbf{Vertices of the migration graph and the core vertices.}
+The vertices of the migration graph are all $\ell$ clusters of the instance.
+We distinguish the \emph{core vertices} that correspond to clusters directly involved in the merge operation: the clusters containing the to-be-merged components in $C_I$ and the cluster containing the merged component in $C_F$.
+There are at most $3$ core vertices in each migration graph (at most two components participate in the merge, and these may migrate to a third cluster).
+
+\noindent
+\textbf{Edges and their labels.}
+The edges of the migration graph denote the migration of nodes from $C_I$ to $C_F$.
+Each edge is labeled with a number $m$ that denotes the number of migrated nodes between clusters.
+Each edge is directed from the cluster that contained the nodes to the cluster they migrated to.
+Both edges might exists between a pair of nodes, as whole components of nodes migrate, and the back-and-forth exchange may be needed for the configuration $C_F$ to be component-respecting.
+
+\noindent
+\textbf{Vertex degree.}
+At most $k$ nodes may migrate from a cluster and at most $k$ nodes may migrate into a cluster.
+Thus, the sum of the labels for both ingoing and outgoing edges of each vertex is at most $k$.
+This implies that the number of ingoing and outgoing edges is also bounded by $k$.
+
+\noindent
+\textbf{Flow preservation.}
+In any feasible configuration, each cluster contains exactly $k$ nodes.
+Thus, for any vertex, the sum of labels of ingoing edges must equal the sum of labels of outgoing edges.
+\maciek{Can we say that this implies that the graph is a circulation?}
+
+\noindent
+\textbf{Migration graph and the cost of reconfiguration.}
+The cost of cluster reconfiguration equals the number of exchanged nodes multiplied by $\alpha$.
+From the standpoint of a migration graph, this corresponds to the sum of labels on edges of the graph.
+A trivial upper bound on the cost is the size of the connected component in the migration graph multiplied by $2\alpha k$.
+
+\medskip
+
+Note that for a single component merge, multiple migration graphs may exist, and multiple of them might have the optimal cost.
+In Section \ref{ssec:cascade}, we show that among all migration graphs for given component merge, there exists an optimal graph with the bounded cycle length.
+Combined with other properties, this allows to bound the size of the migration graph, and consequently bound the cost of each repartition, that results in a bound of the competitive ratio for \TAlg.
+
+\subsection{Costly cluster reconfiguration example}
+\label{ssec:example}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.3\textwidth]{figs/substitute}
+	\caption{A costly cascade example.
+On the left: the configuration prior to the merge and the merge of two components (denoted solid).
+On the right: corresponding migration graph for the optimal (unique) reconfiguration.}
+\end{figure}
+
+In this section we present an example of a merge action that results in a $k/3$(?) length cycle in an optimal migration graph.
+
+\subsection{Bounds on the size of a migration graph}
+\label{ssec:cascade}
+
+In \cite{repartition-disc}, authors trivially bounded the reconfiguration cost for each merge action by $k \cdot \ell$.
+This roughly corresponds to migrating every node in every cluster.
+Now, we bound the cost of a single repartition of \TAlg as a function of $k$.
+Note that the bound of $k \cdot \ell$ is still valid, and the resulting bound is the minimum of these two.
+
+We begin, by bounding the length of each cycle in one of optimal solutions.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.3\textwidth]{figs/substitute}
+	\caption{Illustration to bounding the cycle length.}
+	\label{fig:cascade-illustration}
+\end{figure}
+
+\begin{theorem}
+	\label{th:cascade-cycles}
+	Consider a single merge action of \TAlg.
+	There exists an optimal cost reconfiguration with the migration graph, where each cycle have length at most $k$.
+\end{theorem}
+
+\begin{proof}
+	Assume the contrary, i.e., for each optimal cluster reconfiguration, its migration graph contains a cycle at least $k+1$.
+
+	Fix a reconfiguration and its migration graph $G$ that has the shortest length $L$ of the longest cycle, and has the least number of cycles of length $L$.
+	Let $p = \lbrace v_1, v_2, \ldots, v_L \rbrace$ be any cycle of length $L$.
+	As edge labels are from $\{ 1, 2, \ldots, k \}$, there exists a pair of edges $e, f$ with equal labels.
+
+	We show that in such case, we can modify the solution without affecting its cost and feasibility to decrease the number of cycles of length $L$ by $1$.
+	This would contradict the assumption that we consider a solution with the minimal number of cycles of length $L$.
+
+	We construct a new migration graph $G'$ of equal cost by swapping the (destination) endpoints of $e$ and $f$ (cf. Figure~\ref{fig:cascade-illustration})
+	Formally, let $\lbrace a, b \rbrace = e$ and $\lbrace c, d \rbrace = f$.
+	We remove $e$ and $f$ and add $e' = \lbrace a, d \rbrace$ and $f' = \lbrace c, d \rbrace$.
+	This way we produced a feasible solution.
+
+	It remains to show that while decreasing the number of cycles of length $L$, we did not created any new cycles of length at least $L$.
+	Assume that adding $e'$ created a new cycle of length at least $L$.
+	However, $a$ and $d$ were already connected in $G$ by a directed path of length at least $2$, and a cycle strictly longer than $L$ existed in $G$.
+
+	Assume that adding $f'$ created a new cycle $s^*$ of length $L' \geq L$.
+	Contrary to the previous case, the directed cycle from $e$ to $b$ might not exist in $G$.
+	However, this would mean that $b$ and $e$ were connected by a path of length $L'-1$.
+	This means that we might have followed this alternative path from $b$ to $e$ to obtain a cycle strictly longer than $L$ (cf. Figure~\ref{fig:cascade-illustration}).
+
+\end{proof}
+
+\begin{theorem}
+	Consider a single repartition of DET2.
+	The number of clusters it involves is bounded by $k^k$.
+	\label{th:cascade}
+\end{theorem}
+
+\begin{proof}
+	Each node has degree at most $k$.
+	Every subset has flow conservation property, hence each cycle traverses one of the nodes of graph origin/core. (no subsets with just incoming edges to the origin part)
+	Combine with Theorem \ref{th:cascade-cycles}.
+
+\end{proof}
+
+Corollary: Algorithm DET2 is $O(\ell \cdot k \cdot \min \{ \ell, k^k \})$-competitive.
+\end{comment}
 

### 2020-05-08 paper/disc20clustering.tex
More restructuring

@@
 
-\section{PPL Algorithm} \label{alg:PPL}
+\label{alg:PPL}
 \begin{algorithm}
@@
 
-\section{The Partitioning Problem}
+\section{General Partitioning Model}
 \label{sec:part}
+
+Now  we discuss the general online
+model where the request sequence
+can be arbitrary.
+In Section~\ref{sec:k3}, we show that the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental} provides an optimal algorithm for $k=3$.
+Then, in Section~\ref{sec:k2}, we show a strictly $6$-competitive algorithm for $k=2$.
+
+\subsection{An Optimal Algorithm for Clusters of Size $3$}
+\label{sec:k3}
 
@@
 \maciek{Better name than DET? Currently it is \TAlg, hence it is wrong}
-
-\subsection{Improved analysis of the component-based algorithm}
-\label{sec:k3}
 
@@
 \maciek{Emphasize the difference to DET: "Modification is the following: when repartition is necessary, move to \emph{the closest} perfect partition of components."}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.3\textwidth]{figs/substitute}
+	\caption{Cluster signatures in warm-up k=3 proof}
+\end{figure}
+
+\maciek{Rewrite: cluster types should be cluster signatures}
+In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
+Before bounding the competitive ratio of \TAlg, we introduce the lemma that estimates the cost of a single repartition of \TAlg.
+
+\begin{lemma}
+	\label{lem:1req}
+	During a single repartition, \TAlg exchanges at most $2$ pairs of nodes.
+\end{lemma}
+
+\begin{proof}
+	Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
+	Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
+
+	%We distinguish between three types of clusters: $C_1, C_2, C_3$,
+	%which we define as follows.
+	%A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+	%A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
+	%Finally, a cluster of type $C_3$  contains one component of size $3$.
+
+	Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
+	Note that $U\neq V$,
+	as otherwise, the request would not trigger a~repartitioning.
+	We consider cases based on the types of clusters $U$ and $V$.
+	Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
+	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
+	For two clusters, any configuration can be reached within two swaps, due to the fact that clusters are indistinguishable.
+
+	Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
+	Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
+	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+	Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
+In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
+\end{proof}
 
@@
 
-\subsection{Warm-up: bounding the reconfiguration for $k=3$}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.3\textwidth]{figs/substitute}
-	\caption{Cluster signatures in warm-up k=3 proof}
-\end{figure}
-
-\maciek{Rewrite: cluster types should be cluster signatures}
-In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
-Before bounding the competitive ratio of \TAlg, we introduce the lemma that estimates the cost of a single repartition of \TAlg.
-
-\begin{lemma}
-	\label{lem:1req}
-	During a single repartition, \TAlg exchanges at most $2$ pairs of nodes.
-\end{lemma}
-
-\begin{proof}
-	Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
-	Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
-
-	%We distinguish between three types of clusters: $C_1, C_2, C_3$,
-	%which we define as follows.
-	%A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
-	%A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
-	%Finally, a cluster of type $C_3$  contains one component of size $3$.
-
-	Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
-	Note that $U\neq V$,
-	as otherwise, the request would not trigger a~repartitioning.
-	We consider cases based on the types of clusters $U$ and $V$.
-	Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
-	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
-	For two clusters, any configuration can be reached within two swaps, due to the fact that clusters are indistinguishable.
-
-	Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
-	Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
-	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-	Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
-In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
-\end{proof}
-
-Corollary: DET2 is $O(\ell)$-competitive for $k=3$.
-
-\begin{comment}
-\section{Bounding the cost of a single reconfiguration of components}
-
-\maciek{Overview of the section, intro. Defer introducing the graph notation, the warm-up does not need it}
-
-\subsection{Characterization of the migration graph}
-
-Consider a component merge action performed by \TAlg that triggers a reconfiguration from a configuration $C_I$ to a configuration $C_F$.
-To execute the reconfiguration, the nodes of two non-collocated components migrate to a common cluster.
-Each cluster contains exactly $k$ nodes, thus other nodes migrate to take the place of nodes of merged components.
-Their place must be occupied, too, and this way a single merge action may trigger migrations in clusters not directly involved in a merge action (for an example, see Section~\ref{ssec:example}). 
-Recall that both $C_I$ and $C_F$ are component-respecting configurations, hence a migration of a single node entails migrations of all nodes of its cluster.
-
-In Section~\ref{ssec:cascade}, we upper bound the cost incurred by a reconfiguration that follows a single component merge.
-To this end, we introduce a \emph{migration graph} that models the reconfiguration.
-In the following, we characterize the structure of the migration graph.
-
-\noindent
-\textbf{Vertices of the migration graph and the core vertices.}
-The vertices of the migration graph are all $\ell$ clusters of the instance.
-We distinguish the \emph{core vertices} that correspond to clusters directly involved in the merge operation: the clusters containing the to-be-merged components in $C_I$ and the cluster containing the merged component in $C_F$.
-There are at most $3$ core vertices in each migration graph (at most two components participate in the merge, and these may migrate to a third cluster).
-
-\noindent
-\textbf{Edges and their labels.}
-The edges of the migration graph denote the migration of nodes from $C_I$ to $C_F$.
-Each edge is labeled with a number $m$ that denotes the number of migrated nodes between clusters.
-Each edge is directed from the cluster that contained the nodes to the cluster they migrated to.
-Both edges might exists between a pair of nodes, as whole components of nodes migrate, and the back-and-forth exchange may be needed for the configuration $C_F$ to be component-respecting.
-
-\noindent
-\textbf{Vertex degree.}
-At most $k$ nodes may migrate from a cluster and at most $k$ nodes may migrate into a cluster.
-Thus, the sum of the labels for both ingoing and outgoing edges of each vertex is at most $k$.
-This implies that the number of ingoing and outgoing edges is also bounded by $k$.
-
-\noindent
-\textbf{Flow preservation.}
-In any feasible configuration, each cluster contains exactly $k$ nodes.
-Thus, for any vertex, the sum of labels of ingoing edges must equal the sum of labels of outgoing edges.
-\maciek{Can we say that this implies that the graph is a circulation?}
-
-\noindent
-\textbf{Migration graph and the cost of reconfiguration.}
-The cost of cluster reconfiguration equals the number of exchanged nodes multiplied by $\alpha$.
-From the standpoint of a migration graph, this corresponds to the sum of labels on edges of the graph.
-A trivial upper bound on the cost is the size of the connected component in the migration graph multiplied by $2\alpha k$.
-
-\medskip
-
-Note that for a single component merge, multiple migration graphs may exist, and multiple of them might have the optimal cost.
-In Section \ref{ssec:cascade}, we show that among all migration graphs for given component merge, there exists an optimal graph with the bounded cycle length.
-Combined with other properties, this allows to bound the size of the migration graph, and consequently bound the cost of each repartition, that results in a bound of the competitive ratio for \TAlg.
-
-\subsection{Costly cluster reconfiguration example}
-\label{ssec:example}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.3\textwidth]{figs/substitute}
-	\caption{A costly cascade example.
-On the left: the configuration prior to the merge and the merge of two components (denoted solid).
-On the right: corresponding migration graph for the optimal (unique) reconfiguration.}
-\end{figure}
-
-In this section we present an example of a merge action that results in a $k/3$(?) length cycle in an optimal migration graph.
-
-\subsection{Bounds on the size of a migration graph}
-\label{ssec:cascade}
-
-In \cite{repartition-disc}, authors trivially bounded the reconfiguration cost for each merge action by $k \cdot \ell$.
-This roughly corresponds to migrating every node in every cluster.
-Now, we bound the cost of a single repartition of \TAlg as a function of $k$.
-Note that the bound of $k \cdot \ell$ is still valid, and the resulting bound is the minimum of these two.
-
-We begin, by bounding the length of each cycle in one of optimal solutions.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.3\textwidth]{figs/substitute}
-	\caption{Illustration to bounding the cycle length.}
-	\label{fig:cascade-illustration}
-\end{figure}
-
-\begin{theorem}
-	\label{th:cascade-cycles}
-	Consider a single merge action of \TAlg.
-	There exists an optimal cost reconfiguration with the migration graph, where each cycle have length at most $k$.
-\end{theorem}
-
-\begin{proof}
-	Assume the contrary, i.e., for each optimal cluster reconfiguration, its migration graph contains a cycle at least $k+1$.
-
-	Fix a reconfiguration and its migration graph $G$ that has the shortest length $L$ of the longest cycle, and has the least number of cycles of length $L$.
-	Let $p = \lbrace v_1, v_2, \ldots, v_L \rbrace$ be any cycle of length $L$.
-	As edge labels are from $\{ 1, 2, \ldots, k \}$, there exists a pair of edges $e, f$ with equal labels.
-
-	We show that in such case, we can modify the solution without affecting its cost and feasibility to decrease the number of cycles of length $L$ by $1$.
-	This would contradict the assumption that we consider a solution with the minimal number of cycles of length $L$.
-
-	We construct a new migration graph $G'$ of equal cost by swapping the (destination) endpoints of $e$ and $f$ (cf. Figure~\ref{fig:cascade-illustration})
-	Formally, let $\lbrace a, b \rbrace = e$ and $\lbrace c, d \rbrace = f$.
-	We remove $e$ and $f$ and add $e' = \lbrace a, d \rbrace$ and $f' = \lbrace c, d \rbrace$.
-	This way we produced a feasible solution.
-
-	It remains to show that while decreasing the number of cycles of length $L$, we did not created any new cycles of length at least $L$.
-	Assume that adding $e'$ created a new cycle of length at least $L$.
-	However, $a$ and $d$ were already connected in $G$ by a directed path of length at least $2$, and a cycle strictly longer than $L$ existed in $G$.
-
-	Assume that adding $f'$ created a new cycle $s^*$ of length $L' \geq L$.
-	Contrary to the previous case, the directed cycle from $e$ to $b$ might not exist in $G$.
-	However, this would mean that $b$ and $e$ were connected by a path of length $L'-1$.
-	This means that we might have followed this alternative path from $b$ to $e$ to obtain a cycle strictly longer than $L$ (cf. Figure~\ref{fig:cascade-illustration}).
-
-\end{proof}
-
-\begin{theorem}
-	Consider a single repartition of DET2.
-	The number of clusters it involves is bounded by $k^k$.
-	\label{th:cascade}
-\end{theorem}
-
-\begin{proof}
-	Each node has degree at most $k$.
-	Every subset has flow conservation property, hence each cycle traverses one of the nodes of graph origin/core. (no subsets with just incoming edges to the origin part)
-	Combine with Theorem \ref{th:cascade-cycles}.
-
-\end{proof}
-
-Corollary: Algorithm DET2 is $O(\ell \cdot k \cdot \min \{ \ell, k^k \})$-competitive.
-\end{comment}
-
-\section{Improved algorithm for Online Rematching}
+\subsection{Improved algorithm for Online Rematching}
 \label{sec:k2}
@@
 
+\begin{comment}
+	\section{Bounding the cost of a single reconfiguration of components}
+
+	\maciek{Overview of the section, intro. Defer introducing the graph notation, the warm-up does not need it}
+
+	\subsection{Characterization of the migration graph}
+
+	Consider a component merge action performed by \TAlg that triggers a reconfiguration from a configuration $C_I$ to a configuration $C_F$.
+	To execute the reconfiguration, the nodes of two non-collocated components migrate to a common cluster.
+	Each cluster contains exactly $k$ nodes, thus other nodes migrate to take the place of nodes of merged components.
+	Their place must be occupied, too, and this way a single merge action may trigger migrations in clusters not directly involved in a merge action (for an example, see Section~\ref{ssec:example}). 
+	Recall that both $C_I$ and $C_F$ are component-respecting configurations, hence a migration of a single node entails migrations of all nodes of its cluster.
+
+	In Section~\ref{ssec:cascade}, we upper bound the cost incurred by a reconfiguration that follows a single component merge.
+	To this end, we introduce a \emph{migration graph} that models the reconfiguration.
+	In the following, we characterize the structure of the migration graph.
+
+	\noindent
+	\textbf{Vertices of the migration graph and the core vertices.}
+	The vertices of the migration graph are all $\ell$ clusters of the instance.
+	We distinguish the \emph{core vertices} that correspond to clusters directly involved in the merge operation: the clusters containing the to-be-merged components in $C_I$ and the cluster containing the merged component in $C_F$.
+	There are at most $3$ core vertices in each migration graph (at most two components participate in the merge, and these may migrate to a third cluster).
+
+	\noindent
+	\textbf{Edges and their labels.}
+	The edges of the migration graph denote the migration of nodes from $C_I$ to $C_F$.
+	Each edge is labeled with a number $m$ that denotes the number of migrated nodes between clusters.
+	Each edge is directed from the cluster that contained the nodes to the cluster they migrated to.
+	Both edges might exists between a pair of nodes, as whole components of nodes migrate, and the back-and-forth exchange may be needed for the configuration $C_F$ to be component-respecting.
+
+	\noindent
+	\textbf{Vertex degree.}
+	At most $k$ nodes may migrate from a cluster and at most $k$ nodes may migrate into a cluster.
+	Thus, the sum of the labels for both ingoing and outgoing edges of each vertex is at most $k$.
+	This implies that the number of ingoing and outgoing edges is also bounded by $k$.
+
+	\noindent
+	\textbf{Flow preservation.}
+	In any feasible configuration, each cluster contains exactly $k$ nodes.
+	Thus, for any vertex, the sum of labels of ingoing edges must equal the sum of labels of outgoing edges.
+	\maciek{Can we say that this implies that the graph is a circulation?}
+
+	\noindent
+	\textbf{Migration graph and the cost of reconfiguration.}
+	The cost of cluster reconfiguration equals the number of exchanged nodes multiplied by $\alpha$.
+	From the standpoint of a migration graph, this corresponds to the sum of labels on edges of the graph.
+	A trivial upper bound on the cost is the size of the connected component in the migration graph multiplied by $2\alpha k$.
+
+	\medskip
+
+	Note that for a single component merge, multiple migration graphs may exist, and multiple of them might have the optimal cost.
+	In Section \ref{ssec:cascade}, we show that among all migration graphs for given component merge, there exists an optimal graph with the bounded cycle length.
+	Combined with other properties, this allows to bound the size of the migration graph, and consequently bound the cost of each repartition, that results in a bound of the competitive ratio for \TAlg.
+
+	\subsection{Costly cluster reconfiguration example}
+	\label{ssec:example}
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.3\textwidth]{figs/substitute}
+		\caption{A costly cascade example.
+On the left: the configuration prior to the merge and the merge of two components (denoted solid).
+On the right: corresponding migration graph for the optimal (unique) reconfiguration.}
+	\end{figure}
+
+	In this section we present an example of a merge action that results in a $k/3$(?) length cycle in an optimal migration graph.
+
+	\subsection{Bounds on the size of a migration graph}
+	\label{ssec:cascade}
+
+	In \cite{repartition-disc}, authors trivially bounded the reconfiguration cost for each merge action by $k \cdot \ell$.
+	This roughly corresponds to migrating every node in every cluster.
+	Now, we bound the cost of a single repartition of \TAlg as a function of $k$.
+	Note that the bound of $k \cdot \ell$ is still valid, and the resulting bound is the minimum of these two.
+
+	We begin, by bounding the length of each cycle in one of optimal solutions.
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.3\textwidth]{figs/substitute}
+		\caption{Illustration to bounding the cycle length.}
+		\label{fig:cascade-illustration}
+	\end{figure}
+
+	\begin{theorem}
+		\label{th:cascade-cycles}
+		Consider a single merge action of \TAlg.
+		There exists an optimal cost reconfiguration with the migration graph, where each cycle have length at most $k$.
+	\end{theorem}
+
+	\begin{proof}
+		Assume the contrary, i.e., for each optimal cluster reconfiguration, its migration graph contains a cycle at least $k+1$.
+
+		Fix a reconfiguration and its migration graph $G$ that has the shortest length $L$ of the longest cycle, and has the least number of cycles of length $L$.
+		Let $p = \lbrace v_1, v_2, \ldots, v_L \rbrace$ be any cycle of length $L$.
+		As edge labels are from $\{ 1, 2, \ldots, k \}$, there exists a pair of edges $e, f$ with equal labels.
+
+		We show that in such case, we can modify the solution without affecting its cost and feasibility to decrease the number of cycles of length $L$ by $1$.
+		This would contradict the assumption that we consider a solution with the minimal number of cycles of length $L$.
+
+		We construct a new migration graph $G'$ of equal cost by swapping the (destination) endpoints of $e$ and $f$ (cf. Figure~\ref{fig:cascade-illustration})
+		Formally, let $\lbrace a, b \rbrace = e$ and $\lbrace c, d \rbrace = f$.
+		We remove $e$ and $f$ and add $e' = \lbrace a, d \rbrace$ and $f' = \lbrace c, d \rbrace$.
+		This way we produced a feasible solution.
+
+		It remains to show that while decreasing the number of cycles of length $L$, we did not created any new cycles of length at least $L$.
+		Assume that adding $e'$ created a new cycle of length at least $L$.
+		However, $a$ and $d$ were already connected in $G$ by a directed path of length at least $2$, and a cycle strictly longer than $L$ existed in $G$.
+
+		Assume that adding $f'$ created a new cycle $s^*$ of length $L' \geq L$.
+		Contrary to the previous case, the directed cycle from $e$ to $b$ might not exist in $G$.
+		However, this would mean that $b$ and $e$ were connected by a path of length $L'-1$.
+		This means that we might have followed this alternative path from $b$ to $e$ to obtain a cycle strictly longer than $L$ (cf. Figure~\ref{fig:cascade-illustration}).
+
+	\end{proof}
+
+	\begin{theorem}
+		Consider a single repartition of DET2.
+		The number of clusters it involves is bounded by $k^k$.
+		\label{th:cascade}
+	\end{theorem}
+
+	\begin{proof}
+		Each node has degree at most $k$.
+		Every subset has flow conservation property, hence each cycle traverses one of the nodes of graph origin/core. (no subsets with just incoming edges to the origin part)
+		Combine with Theorem \ref{th:cascade-cycles}.
+
+	\end{proof}
+
+	Corollary: Algorithm DET2 is $O(\ell \cdot k \cdot \min \{ \ell, k^k \})$-competitive.
+	\end{comment}
+
 \end{document}

### 2020-05-08 paper/disc20clustering.tex
Intro continuation

@@
 
-\maciek{Practical motivation}
+We study a problem of managing the virtual machines in a data center.
+We focus on the scenario where virtual machines can be migrated during runtime to another physical machine.
+The possibility of migration allows reacting to unpredictable communication patterns.
+For example, if some distant virtual machines communicate often, it is vital to reduce their distance to save network bandwidth.
+The objective is to~minimize the total network bandwidth used for communication and for migration.
+
+Intuitively, we would like to minimize inter-machine
+interactions by mapping frequently communicating virtual machines to the same physical machine
+Since communication patterns change over time, the~placement of virtual machines should be managed in an online manner, by migrating them between physical machines.
+The~objective is to minimize the weighted sum of inter-machine communication and repartitioning costs.
+The former is defined as the number of communication requests between virtual machines placed at distinct physical machines, and the latter as the number of migrations.
 
@@
 
-\subsubsection{Learning model.}
-
-We consider a variant of Online Balanced Partitioning, where communication requests are permanent: once a pair of nodes start communicating, they continue forever.
+\noindent
+\textbf{Learning model.}
+We study a variant of Online Balanced Partitioning, where communication requests are permanent: once a pair of nodes start communicating, they continue forever.
 To avoid infinite cost, each communicating pair must be collocated.
@@
 The algorithm's goal is to \emph{learn} the communication graph, and partition it without too many reconfigurations.
-
-\noindent
-\textbf{Adversarial input and the competitive analysis.}
-\maciek{The competitive ratio definition and the worst case inputs}
-
-\noindent
-\textbf{Challenges.}
-The partitioning model has a rent-or-buy dilemas. The decision, what to evict is present in both models.
+We refer to this problem variant as the \emph{learning} problem.
+
+\medskip 
+In both models we assume that the communication patterns are not known in advance to our algorithm.
+We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.
+In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.
 

### 2020-05-08 paper/disc20clustering.tex
k=3 addressing comments

@@
 if the nodes belonging to the same component are collocated.
+
+\noindent
+\textbf{Note on our algorithms.}
+\maciek{Should not be here probably, but where?}
+Two of our algorithms (for the learning model and for $k=3$ in the partitioning model) are modified versions of the algorithm DET from \cite{repartition-disc}.
+The difference is in the the choice of partition after a component merge.
+In DET, the partition was arbitrary.
+In PPL, we choose the partition closest to the initial configuration.
+In \TAlg, we choose the partition closest to the previous configuration (a repartition of minimum cost).
 
@@
 
-\maciek{Start with referencing the table of results, saying that some results are known already in the literature.
-And emphasize here that we now have an improved lower bound from the previous section, as that model was a subproblem.}
-
-Let us now discuss the general online
-model where the request sequence
-can be arbitrary.
-We show that the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental} allows obtaining an optimal algorithm for $k=3$.
-Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one cluster until the end of a phase (that we define precisely later).
-
-\maciek{Rewrite above paragraph to say that our algorithm is essentially DET from DISC16, with concretised reconfiguration strategy (i.e., an arbitrary one among the minimum cost reconfigurations)}
-
-\maciek{Better name than DET? Currently it is \TAlg, hence it is wrong}
-
-%We define the algorithm \TAlg in the following way.
-Now we describe the algorithm \TAlg.
+\noindent
+\textbf{Component-based algorithm.}
+Now we describe the algorithm \TAlg that we use throughout this section.
 It partitions nodes into components, and
 initially, each node belongs to its own singleton component.
-For each pair of nodes, \TAlg maintains a counter, initiated to $0$. 
-Upon receiving a request to a pair that is not collocated in one cluster, it increases their counter by $1$.
-If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
+For each pair of nodes $\set{x,y}$, \TAlg maintains a counter $C_{\set{x,y}}$ and increments it on every external request between $x$ and $y$.
+Once $C_{\set{x,y}} = \alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
 If no such partitioning exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
-
-\maciek{Emphasize the difference to DET: "Modification is the following: when repartition is necessary, move to \emph{the closest} perfect partition of components."}
 
@@
 
-\maciek{Rewrite: cluster types should be cluster signatures}
 In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
@@
 	\label{lem:1req}
-	During a single repartition, \TAlg exchanges at most $2$ pairs of nodes.
+	During a single reconfiguration of nodes, \TAlg exchanges at most $2$ pairs of nodes.
 \end{lemma}
@@
 \begin{theorem}
-	\TAlg is $O(\ell)$-competitive.
-	\maciek{Proposed new formulation (current version is all about k=3 still): Let Reconf(k,l) be a bound on the cost of a single reconfiguration.
-Then \TAlg is $O(Reconf(k,l)\cdot k \cdot \ell)$-competitive.}
+	\TAlg is $O(\ell)$-competitive for $k=3$.
 \end{theorem}
@@
 	As \TAlg is component respecting, it never increases any counter above $\alpha$.
-	We say that the pair $(u, v)$ is \emph{saturated} if the counter has value $\alpha$, and \emph{unsaturated} otherwise \maciek{Rename: saturated/unsaturated edge instead of a pair}.
+	We say that the pair $(u, v)$ is \emph{saturated} if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
 	Let $\sigma$ be the input sequence that arrived during the phase.
@@
 	By performing a single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$
-	\maciek{This changes in the general case, and introduces the factor $~1/k$ in opt's cost, increasing the competitive ratio by $O(k)$}.
 	Thus, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{2\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 2 - 2 \alpha \cdot \ell$.
@@
 	Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
-	We charge $\OPT$ $\alpha$ for the first external $\alpha$ requests or a migration,
-	and we follow the analysis regarding the unsaturated requests.
-	If the input finished before the first $\alpha$ external requests, then \TAlg is optimal.
-	\maciek{Missing second case: then we just charge opt alpha for this one edge and this is sufficient, as the cost of alg only decreases from the full phase}
+	If the input finished before the first $\alpha$ external requests, then \TAlg is $1$-competitive.
+	If there were at least $\alpha$ external requests, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a migration; for the cost of \TAlg we follow the analysis of the unsaturated pairs.
 	Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.

### 2020-05-08 paper/disc20clustering.tex
k=3 addressing comments

@@
 
- \title{Tigth Bounds for Online Balanced Graph Partitioning}
+ \title{Tight Bounds for Online Balanced Graph Partitioning}
 
@@
 In Section~\ref{sec:k3}, we show that the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental} provides an optimal algorithm for $k=3$.
+We showed a lower bound of $\Omega(k \cdot \ell)$ that holds for $k=3$ (cf. Section~\ref{sec:lowerbound}), hence the result from this section is asymptotically optimal.
+
 Then, in Section~\ref{sec:k2}, we show a strictly $6$-competitive algorithm for $k=2$.
+\maciek{Intro: compare to previous 7-competitive algorithm. Note strict competitive ratio vs large additive term. Also, emphasize that the algorithm is different than 7-competitive}
 
@@
 Now we describe the algorithm \TAlg that we use throughout this section.
-It partitions nodes into components, and
+The algorithm partitions nodes into components, and
 initially, each node belongs to its own singleton component.
@@
 	\includegraphics[width=0.3\textwidth]{figs/substitute}
-	\caption{Cluster signatures in warm-up k=3 proof}
+	\caption{Cluster types used in the analysis of \TAlg.}
 \end{figure}
@@
 In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
-Before bounding the competitive ratio of \TAlg, we introduce the lemma that estimates the cost of a single repartition of \TAlg.
+Before bounding the competitive ratio of \TAlg, we introduce the lemma that upper bounds the cost of a single repartition of \TAlg.
 
@@
 	Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
-	Note that $U\neq V$,
-	as otherwise, the request would not trigger a~repartitioning.
+	The request triggered the repartition, hence it was external and $U\neq V$.
 	We consider cases based on the types of clusters $U$ and $V$.
-	Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
+
 	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
-	For two clusters, any configuration can be reached within two swaps, due to the fact that clusters are indistinguishable.
-
-	Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
+	In this case, a local repartition is possible (within $U$ and $V$), for a cost of at most $2$ swaps.
+	Note that it is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, a components of size $3$ participates in a merge, which results in a component of size at least $4$, and in this case no component respecting partitioning exists, a contradiction.
+
+	It remains to consider the case where both $U$ and $V$ are of type $C_2$.
 Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.

### 2020-05-08 paper/disc20clustering.tex
bullet proof cost to OPT

--- 2020-05-08 paper/disc20clustering.tex
@@
 
- \title{Tigth Bounds for Online Balanced Graph Partitioning}
+ \title{Online Balanced Graph Partitioning: Optimizations for Large Clusters}
 
@@
 	or machine learning, generate a significant amount of network traffic.
-	By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), we can reduce the network load and  improve application performance. 
-	However, the communication pattern is a priori unknown and may change over time, hence it needs to be learned in an online manner.
+By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), the network load can be reduced and application performance improved.
+	However, as the communication pattern is a priori unknown and may change over time, it needs to be learned efficiently, in an online manner.
 	%
@@
 	of $O(k\cdot \ell)$ for this scenario.
-	We contribute a tight upper bound
+	In addition, we present a tight upper bound
 	of $\Theta(\ell)$ for $k=3$,
@@
 	over time. 
-	We improve the result for $k=2$ by providing a strictly $6$-competitive upper bound for the general model.
-
 \end{abstract}
@@
 
-We study a problem of managing the virtual machines in a data center.
-We focus on the scenario where virtual machines can be migrated during runtime to another physical machine.
-The possibility of migration allows reacting to unpredictable communication patterns.
-For example, if some distant virtual machines communicate often, it is vital to reduce their distance to save network bandwidth.
-The objective is to~minimize the total network bandwidth used for communication and for migration.
-
-Intuitively, we would like to minimize inter-machine
-interactions by mapping frequently communicating virtual machines to the same physical machine
-Since communication patterns change over time, the~placement of virtual machines should be managed in an online manner, by migrating them between physical machines.
-The~objective is to minimize the weighted sum of inter-machine communication and repartitioning costs.
-The former is defined as the number of communication requests between virtual machines placed at distinct physical machines, and the latter as the number of migrations.
+\maciek{Practical motivation}
 
@@
 
-We study two models in this paper: a general \emph{partitioning} model, and its subproblem, a \emph{learning} variant.
-The general partitioning model was introduced by Avin et al.~\cite{repartition-disc} at DISC 2016.
-The learning model was introduced by Henzinger et al.~\cite{sigmetrics19_partitioning} at SIGMETRICS 2019.
-
-\noindent
-\textbf{Partitioning model.}
+\maciek{We study two models. One is a subproblem.}
+
+The general model, where $\sigma$ can be arbitrary
+was the original problem introduced
+by Avin et al.~\cite{repartition-disc} at DISC 2016.
+For simplicity, we will refer to the Perfect Partition
+model as the \emph{learning} model (as
+one has to learn the static communication graph) 
+and to the original as the \emph{repartitioning} model (as one continuously adjusts the partition in response to changing communication graph).
+
+\subsubsection{Partitioning model.}
 The \emph{balanced repartitioning} problem (\OBRP{})
@@
 
-\noindent
-\textbf{Learning model.}
-We study a variant of Online Balanced Partitioning, where communication requests are permanent: once a pair of nodes start communicating, they continue forever.
-To avoid infinite cost, each communicating pair must be collocated.
-We study instances with finite cost: we assume that the communication graph can be perfectly partitioned in the clusters (no external requests are required).
-The algorithm's goal is to \emph{learn} the communication graph, and partition it without too many reconfigurations.
-We refer to this problem variant as the \emph{learning} problem.
-
-\medskip 
-In both models we assume that the communication patterns are not known in advance to our algorithm.
-We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.
-In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.
+\maciek{The competitive ratio definition and the worst case inputs}
+
+\maciek{Ilustrate the challenges. Rent-or-buy. Decision what to evict}
+
+\subsubsection{Learning model.}
+In addition to the general model, we study a simplified Perfect Partition model (also studied by Henzinger et al.~\cite{sigmetrics19_partitioning} (SIGMETRICS 2019)) that we consider for a stronger lower bound.
+There, we assume that $\sigma$
+reveals the edges of a static graph
+that can be perfectly partitioned
+(no external requests are required).
 
@@
 
-\maciek{Below is similar to SIDMA, possibly too similar}
-
-The static offline version of~the~partitioning~problem, i.e., a problem variant where
-migration is not allowed, where all requests are known in advance, and where
-the goal is to find an assignment of $n$ nodes to $\ell$~physical machines, each of~capacity $n/\ell$, is known as the
-\emph{$\ell$-balanced graph partitioning problem}. The problem is 
-NP-complete, and cannot even be approximated within any finite factor unless P
-= NP~\cite{AndRae06}.  The static
-variant where $\ell = 2$ corresponds to the minimum bisection problem, which
-is already NP-hard~\cite{GaJoSt76}, and 
-the currently best approximation ratio is $O(\log n)$~\cite{SarVaz95,ArKaKa99,FeKrNi00,FeiKra02,KraFei06,Raec08}.
-The inapproximability of the static variant for general values of $\ell$
-motivated research on the bicriteria variant, which can be seen as the offline
-counterpart of our capacity augmentation approach. Here, the~goal
-is~to~compute a~graph partitioning into $\ell$ components of~size at most~$k$ (where $k > n/\ell$) and the cost of the cut is compared to the optimal (non-augmented)
-solution where all components are of a~size at most $n/k$. The variant where
-$n \geq 2 \cdot k \cdot \ell$ was considered in
-\cite{LeMaTr90,SimTen97,EvNaRS00,EvNaRS99,KrNaSc09}. So far, the~best result~is~an~$O(\!\sqrt{\log n \cdot \log \ell})$-approximation algorithm~\cite{KrNaSc09}.
-
-The studied model is related to online
-caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}, sometimes also referred to
-as online caching, where requests for data items (nodes) arrive over time and
-need to be served from a cache of finite capacity, and where the number of
-cache misses must be minimized. Classic problem variants usually boil down to
-finding a smart eviction strategy, such as Least Recently Used (LRU)~\cite{SleTar85}. In our
-setting, requests can be served remotely (i.e.,~without fetching the
-corresponding nodes to a single physical machine). In this light, our model is more
-reminiscent of caching models \emph{with
-bypassing}~\cite{EpImLN11,EpImLN15,Irani02}. As a~side result, we show that our problem is
-capable of emulating online caching.
-A major difference between  these problems is that in the caching problems, each request involves a~single element of the universe, while in our model \emph{both} endpoints of a communication request are subject to~optimization.
+\maciek{Offline results}
+\maciek{Links other problems: for $k=2$ a flavor of online matching, for $\ell = 2$ it generalizes paging}
 
@@
 an asymptotically optimal,
-$\Theta(\ell)$ algorithm for $k=3$, improving the best known upper bound 
-so far $O(\ell^2)$~\cite{repartition-disc}.
-Then, we present a strictly $6$-competitive algorithm for $k=2$ that improves upon previous $7$-competitive algorithm with $O(\ell^2)$ additive constant.
+$\Theta(\ell)$ algorithm for $k=3$;
+the best known upper bound 
+so far was $O(\ell^2)$~\cite{repartition-disc}.
+Then, we present a strictly $6$-competitive algorithm for $k=2$ that improves upon previous $7$-competitive algorithm with $O(\ell^2)$ additive constant..
+In addition, we show a new, improved analysis of the algorithm from \cite{repartition-disc} that depends on the solutions to an interesting combinatorial problem (see Section~\ref{ssec:cascade}, The Cascade Problem).
 %
@@
 		\textbf{$k=3$}&  $\Omega(\ell)$ \hspace{0.3cm}(\S \ref{sec:lowerbound})& $O(\ell) $\hspace{0.3cm}(\S \ref{sec:k3})\\
-		$k > 3$ & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k^2 \cdot \ell^2)$\hspace{0.1cm} \cite{repartition-disc} \\
+		$k > 3$ & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell \cdot \min \{ k \cdot \ell, \Rep(k, \ell) \})$\hspace{0.1cm} (\S \ref{sec:k3}) \\
 		\rowcolor{gray!25}
@@
 
-\label{alg:PPL}
+\section{PPL Algorithm} \label{alg:PPL}
 \begin{algorithm}
@@
 
-\section{General Partitioning Model}
+\section{The Partitioning Problem}
 \label{sec:part}
-
-Now  we discuss the general online
-model where the request sequence
-can be arbitrary.
-In Section~\ref{sec:k3}, we show that the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental} provides an optimal algorithm for $k=3$.
-Then, in Section~\ref{sec:k2}, we show a strictly $6$-competitive algorithm for $k=2$.
-
-\subsection{An Optimal Algorithm for Clusters of Size $3$}
-\label{sec:k3}
 
@@
 \maciek{Better name than DET? Currently it is \TAlg, hence it is wrong}
+
+\subsection{Improved analysis of the component-based algorithm}
+\label{sec:k3}
 
@@
 \maciek{Emphasize the difference to DET: "Modification is the following: when repartition is necessary, move to \emph{the closest} perfect partition of components."}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.3\textwidth]{figs/substitute}
-	\caption{Cluster signatures in warm-up k=3 proof}
-\end{figure}
-
-\maciek{Rewrite: cluster types should be cluster signatures}
-In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
-Before bounding the competitive ratio of \TAlg, we introduce the lemma that estimates the cost of a single repartition of \TAlg.
-
-\begin{lemma}
-	\label{lem:1req}
-	During a single repartition, \TAlg exchanges at most $2$ pairs of nodes.
-\end{lemma}
-
-\begin{proof}
-	Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
-	Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
-
-	%We distinguish between three types of clusters: $C_1, C_2, C_3$,
-	%which we define as follows.
-	%A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
-	%A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
-	%Finally, a cluster of type $C_3$  contains one component of size $3$.
-
-	Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
-	Note that $U\neq V$,
-	as otherwise, the request would not trigger a~repartitioning.
-	We consider cases based on the types of clusters $U$ and $V$.
-	Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
-	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
-	For two clusters, any configuration can be reached within two swaps, due to the fact that clusters are indistinguishable.
-
-	Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
-	Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
-	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-	Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
-In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
-\end{proof}
 
@@
 
-\subsection{Improved algorithm for Online Rematching}
+\subsection{Warm-up: bounding the reconfiguration for $k=3$}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.3\textwidth]{figs/substitute}
+	\caption{Cluster signatures in warm-up k=3 proof}
+\end{figure}
+
+\maciek{Rewrite: cluster types should be cluster signatures}
+In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
+Before bounding the competitive ratio of \TAlg, we introduce the lemma that estimates the cost of a single repartition of \TAlg.
+
+\begin{lemma}
+	\label{lem:1req}
+	During a single repartition, \TAlg exchanges at most $2$ pairs of nodes.
+\end{lemma}
+
+\begin{proof}
+	Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
+	Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
+
+	%We distinguish between three types of clusters: $C_1, C_2, C_3$,
+	%which we define as follows.
+	%A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+	%A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
+	%Finally, a cluster of type $C_3$  contains one component of size $3$.
+
+	Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
+	Note that $U\neq V$,
+	as otherwise, the request would not trigger a~repartitioning.
+	We consider cases based on the types of clusters $U$ and $V$.
+	Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
+	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
+	For two clusters, any configuration can be reached within two swaps, due to the fact that clusters are indistinguishable.
+
+	Finally, we consider the case where both $U$ and $V$ are of type $C_2$.
+Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
+	Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
+	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+	Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
+In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
+\end{proof}
+
+Corollary: DET2 is $O(\ell)$-competitive for $k=3$.
+
+\begin{comment}
+\section{Bounding the cost of a single reconfiguration of components}
+
+\maciek{Overview of the section, intro. Defer introducing the graph notation, the warm-up does not need it}
+
+\subsection{Characterization of the migration graph}
+
+Consider a component merge action performed by \TAlg that triggers a reconfiguration from a configuration $C_I$ to a configuration $C_F$.
+To execute the reconfiguration, the nodes of two non-collocated components migrate to a common cluster.
+Each cluster contains exactly $k$ nodes, thus other nodes migrate to take the place of nodes of merged components.
+Their place must be occupied, too, and this way a single merge action may trigger migrations in clusters not directly involved in a merge action (for an example, see Section~\ref{ssec:example}). 
+Recall that both $C_I$ and $C_F$ are component-respecting configurations, hence a migration of a single node entails migrations of all nodes of its cluster.
+
+In Section~\ref{ssec:cascade}, we upper bound the cost incurred by a reconfiguration that follows a single component merge.
+To this end, we introduce a \emph{migration graph} that models the reconfiguration.
+In the following, we characterize the structure of the migration graph.
+
+\noindent
+\textbf{Vertices of the migration graph and the core vertices.}
+The vertices of the migration graph are all $\ell$ clusters of the instance.
+We distinguish the \emph{core vertices} that correspond to clusters directly involved in the merge operation: the clusters containing the to-be-merged components in $C_I$ and the cluster containing the merged component in $C_F$.
+There are at most $3$ core vertices in each migration graph (at most two components participate in the merge, and these may migrate to a third cluster).
+
+\noindent
+\textbf{Edges and their labels.}
+The edges of the migration graph denote the migration of nodes from $C_I$ to $C_F$.
+Each edge is labeled with a number $m$ that denotes the number of migrated nodes between clusters.
+Each edge is directed from the cluster that contained the nodes to the cluster they migrated to.
+Both edges might exists between a pair of nodes, as whole components of nodes migrate, and the back-and-forth exchange may be needed for the configuration $C_F$ to be component-respecting.
+
+\noindent
+\textbf{Vertex degree.}
+At most $k$ nodes may migrate from a cluster and at most $k$ nodes may migrate into a cluster.
+Thus, the sum of the labels for both ingoing and outgoing edges of each vertex is at most $k$.
+This implies that the number of ingoing and outgoing edges is also bounded by $k$.
+
+\noindent
+\textbf{Flow preservation.}
+In any feasible configuration, each cluster contains exactly $k$ nodes.
+Thus, for any vertex, the sum of labels of ingoing edges must equal the sum of labels of outgoing edges.
+\maciek{Can we say that this implies that the graph is a circulation?}
+
+\noindent
+\textbf{Migration graph and the cost of reconfiguration.}
+The cost of cluster reconfiguration equals the number of exchanged nodes multiplied by $\alpha$.
+From the standpoint of a migration graph, this corresponds to the sum of labels on edges of the graph.
+A trivial upper bound on the cost is the size of the connected component in the migration graph multiplied by $2\alpha k$.
+
+\medskip
+
+Note that for a single component merge, multiple migration graphs may exist, and multiple of them might have the optimal cost.
+In Section \ref{ssec:cascade}, we show that among all migration graphs for given component merge, there exists an optimal graph with the bounded cycle length.
+Combined with other properties, this allows to bound the size of the migration graph, and consequently bound the cost of each repartition, that results in a bound of the competitive ratio for \TAlg.
+
+\subsection{Costly cluster reconfiguration example}
+\label{ssec:example}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.3\textwidth]{figs/substitute}
+	\caption{A costly cascade example.
+On the left: the configuration prior to the merge and the merge of two components (denoted solid).
+On the right: corresponding migration graph for the optimal (unique) reconfiguration.}
+\end{figure}
+
+In this section we present an example of a merge action that results in a $k/3$(?) length cycle in an optimal migration graph.
+
+\subsection{Bounds on the size of a migration graph}
+\label{ssec:cascade}
+
+In \cite{repartition-disc}, authors trivially bounded the reconfiguration cost for each merge action by $k \cdot \ell$.
+This roughly corresponds to migrating every node in every cluster.
+Now, we bound the cost of a single repartition of \TAlg as a function of $k$.
+Note that the bound of $k \cdot \ell$ is still valid, and the resulting bound is the minimum of these two.
+
+We begin, by bounding the length of each cycle in one of optimal solutions.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.3\textwidth]{figs/substitute}
+	\caption{Illustration to bounding the cycle length.}
+	\label{fig:cascade-illustration}
+\end{figure}
+
+\begin{theorem}
+	\label{th:cascade-cycles}
+	Consider a single merge action of \TAlg.
+	There exists an optimal cost reconfiguration with the migration graph, where each cycle have length at most $k$.
+\end{theorem}
+
+\begin{proof}
+	Assume the contrary, i.e., for each optimal cluster reconfiguration, its migration graph contains a cycle at least $k+1$.
+
+	Fix a reconfiguration and its migration graph $G$ that has the shortest length $L$ of the longest cycle, and has the least number of cycles of length $L$.
+	Let $p = \lbrace v_1, v_2, \ldots, v_L \rbrace$ be any cycle of length $L$.
+	As edge labels are from $\{ 1, 2, \ldots, k \}$, there exists a pair of edges $e, f$ with equal labels.
+
+	We show that in such case, we can modify the solution without affecting its cost and feasibility to decrease the number of cycles of length $L$ by $1$.
+	This would contradict the assumption that we consider a solution with the minimal number of cycles of length $L$.
+
+	We construct a new migration graph $G'$ of equal cost by swapping the (destination) endpoints of $e$ and $f$ (cf. Figure~\ref{fig:cascade-illustration})
+	Formally, let $\lbrace a, b \rbrace = e$ and $\lbrace c, d \rbrace = f$.
+	We remove $e$ and $f$ and add $e' = \lbrace a, d \rbrace$ and $f' = \lbrace c, d \rbrace$.
+	This way we produced a feasible solution.
+
+	It remains to show that while decreasing the number of cycles of length $L$, we did not created any new cycles of length at least $L$.
+	Assume that adding $e'$ created a new cycle of length at least $L$.
+	However, $a$ and $d$ were already connected in $G$ by a directed path of length at least $2$, and a cycle strictly longer than $L$ existed in $G$.
+
+	Assume that adding $f'$ created a new cycle $s^*$ of length $L' \geq L$.
+	Contrary to the previous case, the directed cycle from $e$ to $b$ might not exist in $G$.
+	However, this would mean that $b$ and $e$ were connected by a path of length $L'-1$.
+	This means that we might have followed this alternative path from $b$ to $e$ to obtain a cycle strictly longer than $L$ (cf. Figure~\ref{fig:cascade-illustration}).
+
+\end{proof}
+
+\begin{theorem}
+	Consider a single repartition of DET2.
+	The number of clusters it involves is bounded by $k^k$.
+	\label{th:cascade}
+\end{theorem}
+
+\begin{proof}
+	Each node has degree at most $k$.
+	Every subset has flow conservation property, hence each cycle traverses one of the nodes of graph origin/core. (no subsets with just incoming edges to the origin part)
+	Combine with Theorem \ref{th:cascade-cycles}.
+
+\end{proof}
+
+Corollary: Algorithm DET2 is $O(\ell \cdot k \cdot \min \{ \ell, k^k \})$-competitive.
+\end{comment}
+
+\section{Improved algorithm for Online Rematching}
 \label{sec:k2}
@@
 	or $\alpha$ for collocating the pair and
-	serving (some of) the requests with cost 0.
+	serving (some of) the requests with  cost zero.
 	Therefore in this case,
@@
 	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
-	We define $R_{t'}$ for the collocation at $t'$ in an analogous way.
+	We define $R_{t'}$ for the collocation at $t'$  analogously.
 	Then,
@@
 	+ \OPTM(\sigma_{\{u,v\}}) $.
+	Note that \OPT cannot have both pairs collocated at the same time.
+	However, it can have one of the pairs already collocated prior to its respective interval.
+	Let us assume  \OPT has one pair,
+	e.g.~$\{u,v\}$ collocated already prior its respective interval, i.e.~$(t',t]$,
+	 and keeps it so during the interval.
+	 Then it pays zero while serving $\sigma_{\{u,v\}}$.
+	Hence,
+	it must pay $\alpha$ for collocating the other pair, in this case $\{u,w\}$,
+	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$. 
+	Therefore in any case $\OPTM(R_t) \geq  \min{\{ \lambda,\alpha \}}$.
+	\begin{comment}
 	We distinguish three (non-exhaustive) cases for the cost to \OPT w.r.t.~$R_t$.
@@
 	$\mathit{OPT} (R_t) \geq  \min{ \{ \lambda, \alpha \}}$.
-
-	It remains to bound the cost  incurred by \OPT for requests to $\{u,v\}$ that do not lead to its collocation until the end of the sequence at $t=m$.
+	\end{comment}
+
+	It remains to bound the cost  incurred by \OPT due to requests to $\{u,v\}$ that do not lead to its collocation until the end of the sequence at $t=m$.
+	We bound the cost analogously to the case where \RM collocates $\{u,v\}$.
 	If $\{u,v\}$ is not collocated in the initial matching
@@
 	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
-	for collocating this pair or (and) paying for (some of) requests in $\sigma_{\{u,v\}}[1,m]$.
+	for collocating this pair or (and) paying for (resp.~some of) requests in $\sigma_{\{u,v\}}[1,m]$.
 	Else,
@@
 	that is collocated by \RM at $t' < m$ (separating $\{u,v\}$).
-	We define times $t'' < t' < m$ similarly to the case where \RM collocates $\{u,v\}$.
+	We define times $t'' < t' < m$ similarly as in the former case.
 	Let $R^*_{\{u,v\}} := \sigma_{\{u,w\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
 	Then,
-	we apply cases (i) to (iii) in an analogous way,
-	replacing $\lambda$ with $C^{m}_{ \{u,v\}}$ in the analysis.
+	\OPT must pay for collocating at least one pair or (and) serving requests 
+	to the other pair remotely.
 	Thus,
-	we obtain
 	$\mathit{OPT} (R^*_{\{u,v\}}) 
@@
 
-\begin{comment}
-	\section{Bounding the cost of a single reconfiguration of components}
-
-	\maciek{Overview of the section, intro. Defer introducing the graph notation, the warm-up does not need it}
-
-	\subsection{Characterization of the migration graph}
-
-	Consider a component merge action performed by \TAlg that triggers a reconfiguration from a configuration $C_I$ to a configuration $C_F$.
-	To execute the reconfiguration, the nodes of two non-collocated components migrate to a common cluster.
-	Each cluster contains exactly $k$ nodes, thus other nodes migrate to take the place of nodes of merged components.
-	Their place must be occupied, too, and this way a single merge action may trigger migrations in clusters not directly involved in a merge action (for an example, see Section~\ref{ssec:example}). 
-	Recall that both $C_I$ and $C_F$ are component-respecting configurations, hence a migration of a single node entails migrations of all nodes of its cluster.
-
-	In Section~\ref{ssec:cascade}, we upper bound the cost incurred by a reconfiguration that follows a single component merge.
-	To this end, we introduce a \emph{migration graph} that models the reconfiguration.
-	In the following, we characterize the structure of the migration graph.
-
-	\noindent
-	\textbf{Vertices of the migration graph and the core vertices.}
-	The vertices of the migration graph are all $\ell$ clusters of the instance.
-	We distinguish the \emph{core vertices} that correspond to clusters directly involved in the merge operation: the clusters containing the to-be-merged components in $C_I$ and the cluster containing the merged component in $C_F$.
-	There are at most $3$ core vertices in each migration graph (at most two components participate in the merge, and these may migrate to a third cluster).
-
-	\noindent
-	\textbf{Edges and their labels.}
-	The edges of the migration graph denote the migration of nodes from $C_I$ to $C_F$.
-	Each edge is labeled with a number $m$ that denotes the number of migrated nodes between clusters.
-	Each edge is directed from the cluster that contained the nodes to the cluster they migrated to.
-	Both edges might exists between a pair of nodes, as whole components of nodes migrate, and the back-and-forth exchange may be needed for the configuration $C_F$ to be component-respecting.
-
-	\noindent
-	\textbf{Vertex degree.}
-	At most $k$ nodes may migrate from a cluster and at most $k$ nodes may migrate into a cluster.
-	Thus, the sum of the labels for both ingoing and outgoing edges of each vertex is at most $k$.
-	This implies that the number of ingoing and outgoing edges is also bounded by $k$.
-
-	\noindent
-	\textbf{Flow preservation.}
-	In any feasible configuration, each cluster contains exactly $k$ nodes.
-	Thus, for any vertex, the sum of labels of ingoing edges must equal the sum of labels of outgoing edges.
-	\maciek{Can we say that this implies that the graph is a circulation?}
-
-	\noindent
-	\textbf{Migration graph and the cost of reconfiguration.}
-	The cost of cluster reconfiguration equals the number of exchanged nodes multiplied by $\alpha$.
-	From the standpoint of a migration graph, this corresponds to the sum of labels on edges of the graph.
-	A trivial upper bound on the cost is the size of the connected component in the migration graph multiplied by $2\alpha k$.
-
-	\medskip
-
-	Note that for a single component merge, multiple migration graphs may exist, and multiple of them might have the optimal cost.
-	In Section \ref{ssec:cascade}, we show that among all migration graphs for given component merge, there exists an optimal graph with the bounded cycle length.
-	Combined with other properties, this allows to bound the size of the migration graph, and consequently bound the cost of each repartition, that results in a bound of the competitive ratio for \TAlg.
-
-	\subsection{Costly cluster reconfiguration example}
-	\label{ssec:example}
-
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=0.3\textwidth]{figs/substitute}
-		\caption{A costly cascade example.
-On the left: the configuration prior to the merge and the merge of two components (denoted solid).
-On the right: corresponding migration graph for the optimal (unique) reconfiguration.}
-	\end{figure}
-
-	In this section we present an example of a merge action that results in a $k/3$(?) length cycle in an optimal migration graph.
-
-	\subsection{Bounds on the size of a migration graph}
-	\label{ssec:cascade}
-
-	In \cite{repartition-disc}, authors trivially bounded the reconfiguration cost for each merge action by $k \cdot \ell$.
-	This roughly corresponds to migrating every node in every cluster.
-	Now, we bound the cost of a single repartition of \TAlg as a function of $k$.
-	Note that the bound of $k \cdot \ell$ is still valid, and the resulting bound is the minimum of these two.
-
-	We begin, by bounding the length of each cycle in one of optimal solutions.
-
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=0.3\textwidth]{figs/substitute}
-		\caption{Illustration to bounding the cycle length.}
-		\label{fig:cascade-illustration}
-	\end{figure}
-
-	\begin{theorem}
-		\label{th:cascade-cycles}
-		Consider a single merge action of \TAlg.
-		There exists an optimal cost reconfiguration with the migration graph, where each cycle have length at most $k$.
-	\end{theorem}
-
-	\begin{proof}
-		Assume the contrary, i.e., for each optimal cluster reconfiguration, its migration graph contains a cycle at least $k+1$.
-
-		Fix a reconfiguration and its migration graph $G$ that has the shortest length $L$ of the longest cycle, and has the least number of cycles of length $L$.
-		Let $p = \lbrace v_1, v_2, \ldots, v_L \rbrace$ be any cycle of length $L$.
-		As edge labels are from $\{ 1, 2, \ldots, k \}$, there exists a pair of edges $e, f$ with equal labels.
-
-		We show that in such case, we can modify the solution without affecting its cost and feasibility to decrease the number of cycles of length $L$ by $1$.
-		This would contradict the assumption that we consider a solution with the minimal number of cycles of length $L$.
-
-		We construct a new migration graph $G'$ of equal cost by swapping the (destination) endpoints of $e$ and $f$ (cf. Figure~\ref{fig:cascade-illustration})
-		Formally, let $\lbrace a, b \rbrace = e$ and $\lbrace c, d \rbrace = f$.
-		We remove $e$ and $f$ and add $e' = \lbrace a, d \rbrace$ and $f' = \lbrace c, d \rbrace$.
-		This way we produced a feasible solution.
-
-		It remains to show that while decreasing the number of cycles of length $L$, we did not created any new cycles of length at least $L$.
-		Assume that adding $e'$ created a new cycle of length at least $L$.
-		However, $a$ and $d$ were already connected in $G$ by a directed path of length at least $2$, and a cycle strictly longer than $L$ existed in $G$.
-
-		Assume that adding $f'$ created a new cycle $s^*$ of length $L' \geq L$.
-		Contrary to the previous case, the directed cycle from $e$ to $b$ might not exist in $G$.
-		However, this would mean that $b$ and $e$ were connected by a path of length $L'-1$.
-		This means that we might have followed this alternative path from $b$ to $e$ to obtain a cycle strictly longer than $L$ (cf. Figure~\ref{fig:cascade-illustration}).
-
-	\end{proof}
-
-	\begin{theorem}
-		Consider a single repartition of DET2.
-		The number of clusters it involves is bounded by $k^k$.
-		\label{th:cascade}
-	\end{theorem}
-
-	\begin{proof}
-		Each node has degree at most $k$.
-		Every subset has flow conservation property, hence each cycle traverses one of the nodes of graph origin/core. (no subsets with just incoming edges to the origin part)
-		Combine with Theorem \ref{th:cascade-cycles}.
-
-	\end{proof}
-
-	Corollary: Algorithm DET2 is $O(\ell \cdot k \cdot \min \{ \ell, k^k \})$-competitive.
-	\end{comment}
-
 \end{document}

### 2020-05-08 paper/disc20clustering.tex
minor
wip

@@
 We study instances with finite cost: we assume that the communication graph can be perfectly partitioned in the clusters (no external requests are required).
+In the partitioning model, $\alpha$ denoted the ratio between the cost of a single communication request and the cost of migration; in the learning model, we set $\alpha = 1$.
 The algorithm's goal is to \emph{learn} the communication graph, and partition it without too many reconfigurations.
-We refer to this problem variant as the \emph{learning} problem.
 
@@
 	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$. 
-	Therefore in any case $\OPTM(R_t) \geq  \min{\{ \lambda,\alpha \}}$.
+	Therefore in any case,
+	$\OPTM(R_t) \geq  \min{\{ \lambda,\alpha \}}$.
 	\begin{comment}
@@
 	Then there exists an intersecting pair $\set{u,w}$
-	that is collocated by \RM at $t' < m$ (separating $\{u,v\}$).
+	that is collocated by \RM at $t' < m$,
+	separating $\{u,v\}$.
 	We define times $t'' < t' < m$ similarly as in the former case.

### 2020-05-08 paper/disc20clustering.tex
wip

@@
 
-\maciek{Below is similar to SIDMA, possibly too similar}
-
 The static offline version of~the~partitioning~problem, i.e., a problem variant where
@@
 the currently best approximation ratio is $O(\log n)$~\cite{SarVaz95,ArKaKa99,FeKrNi00,FeiKra02,KraFei06,Raec08}.
-The inapproximability of the static variant for general values of $\ell$
-motivated research on the bicriteria variant, which can be seen as the offline
-counterpart of our capacity augmentation approach. Here, the~goal
-is~to~compute a~graph partitioning into $\ell$ components of~size at most~$k$ (where $k > n/\ell$) and the cost of the cut is compared to the optimal (non-augmented)
-solution where all components are of a~size at most $n/k$. The variant where
-$n \geq 2 \cdot k \cdot \ell$ was considered in
-\cite{LeMaTr90,SimTen97,EvNaRS00,EvNaRS99,KrNaSc09}. So far, the~best result~is~an~$O(\!\sqrt{\log n \cdot \log \ell})$-approximation algorithm~\cite{KrNaSc09}.
 
 The studied model is related to online
-caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}, sometimes also referred to
+paging~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}, sometimes also referred to
 as online caching, where requests for data items (nodes) arrive over time and
@@
 reminiscent of caching models \emph{with
-bypassing}~\cite{EpImLN11,EpImLN15,Irani02}. As a~side result, we show that our problem is
-capable of emulating online caching.
+bypassing}~\cite{EpImLN11,EpImLN15,Irani02}.
 A major difference between  these problems is that in the caching problems, each request involves a~single element of the universe, while in our model \emph{both} endpoints of a communication request are subject to~optimization.

### 2020-05-08 paper/disc20clustering.tex

@@
 
-In this section we consider a variant of Online Balanced Partitioning, where communication requests are permanent: once a pair of nodes start communicating, they continue forever.
-To avoid infinite cost, each communicating pair must be collocated.
-We study instances with finite cost: we assume that the communication graph can be perfectly partitioned in the clusters.
-The algorithm's goal is to \emph{learn} the communication graph, and partition it without too many reconfigurations.
-
+In this section we consider the learning variant of Online Balanced Partitioning.
 For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$.
@@
 	We say that the pair $(u, v)$ is \emph{saturated} if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
-	Let $\sigma$ be the input sequence that arrived during the phase.
-	Let $\sigma_{cost}$ be the requests that at the moment of arrival were external requests for \TAlg (these are the only requests that incurred a cost for \TAlg).
+	By $\sigma$ we denote the input sequence that arrived during the phase.
+	By $\sigma_{cost}$ we denote the requests that at the moment of arrival were external requests for \TAlg (these are the only requests that incur cost for \TAlg).
 	In our analysis, we partition $\sigma_{cost}$ into subsequences $\sigma_I$ and $\sigma_E$.
@@
 	Now we lower-bound the cost of $\OPT$.
-	By $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$ we denote the cost of $\OPT$ on these input sequences (defined with respect to components of \TAlg in this phase).
+	By $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$ we denote the cost of $\OPT$ on requests from sequences $\sigma_I$ and $\sigma_E$, respectively.
+	Note that these costs are defined with respect to components of \TAlg in this phase.
 	By $\OPT(M)$ we denote the cost of migrations performed by $\OPT$ in this phase.

### 2020-05-08 paper/disc20clustering.tex
Wip

@@
 	Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
-	In a~fixed configuration of $\OPT$, it may mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
+	By entering some fixed configuration before serving $\sigma_E$, \OPT may mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
 	Recall that $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.

### 2020-05-08 paper/disc20clustering.tex
removing comments

@@
 
-Major:
-
-\maciek{I feel slightly uneasy here: "In particular,
-we charge \OPT a cost incurred by requests in
-$\sigma_{\{u,w\}}$ and $\sigma_{\{u,v\}}$,
-if and only if  \OPT incurs the cost within  intervals $(t'', t']$ and $(t', t]$,
-respectively.
-\mahmoud{this is only for intuition but a key statement.
-It is meant to say: \OPT may pay cost for the same two pairs outside the intervals.
-But we don't charge \OPT  those costs at $t$.
-This way we avoid over charging \OPT as only intervals $r_{t'}$ and $r_t$ can overlap which leads to double charging and we take care of it.}
-"}
-
-\maciek{"Observe that for requests in $R_t$".
-First the separate sentence with the definition of $R_t$ (it is challenging enough to just understand it.
-Then reason why it pays the minimum of three values (I don't see it, and later you show that cost(rt) is at least minimum of other two values).
-Is it a consequence of three cases?
-\mahmoud{the duration is the key and the latter duration is different.
-I reworded the whole thing.
-And yes, the same three cases can be used to explain this observation.
-Now the "some of" in the parenthesis might be enough for a clue.
-This observation is  used implicitly in the proof when it bounds the cost of first time collocation.}}
-
-\maciek{"We charge \OPT the cost
-$\mathit{\OPT} (\sigma_{\{u,v\}}[1,t]) 
-\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
-for collocating this pair or (and) paying (partially) for requests in $\sigma_{\{u,v\}}[1,t]$." No issues with charge sharing from "regular collocation" case?
-Maybe I just don't understand "partially" here.
-Not necessarily rewrite!
-Just maybe check if it's true.
-\mahmoud{this is the cost of first collocation of $\{u,v\}$. "partially" is confusing and now it says "some of" instead.}}
-
-Minor:
-
-\maciek{The big inequality is impressive, but not a lot change in each row.
-Instead, bound the cost of ALG separately in one line, and the cost of \OPT in another.
-(possibly far away from eachother) \mahmoud{done}}
-
-\maciek{Remove all "charge both". Just sum up costs. I would say not use "charge" at all, but it's up to you. \mahmoud{done}}
-
-\maciek{Structure in paragraphs.
-3 parts: first collocation, regular collocation and unfinished.
-Do this for ALG and then for \OPT, and do not mix (unless I don't see some dependencies).
-Start each paragraph by the sentence that says what the paragraph is about (first/opt, regular/alg etc.).
-Assume that the reader reads only the first sentence of each paragraph (if it's short), and the proof structure must be clear from this.
-Some other important things go to their own paragraph too, and the first sentence must describe what we do there.}
-
-\maciek{$R_t$ is the central concept in this proof.
-Possibly even start with the paragraph: charging \OPT, regular collocation.
-Make this definition in its separate line to further emphasize?
-It would suffice for a profesional to see the definiton of $t''/t'$ and this line to convince himself that the proof is correct.}
-
-\maciek{I think that "deviate" is OK to use like this.}
-
-Notation:
-
-\maciek{End of the time proposition: $t_e / t^* / T$ instead of $m$? Or capital $T$ (but I am aware of the conflict)}
-
-\maciek{Instead of $cost_{opt}$ just use opt / alg?}
-
-\maciek{Consider naming the adjacent edges $e$ and $f$.}
-
 \section{Distributed implementation.}

### 2020-05-08 paper/disc20clustering.tex
Lower bound issues addressed

--- 2020-05-08 paper/disc20clustering.tex
@@
 is~$\alpha \in \mathbb{Z}^+$.
-\maciek{Why restrict $\alpha$? It should just be positive}
 For any algorithm $\ALG$,
@@
 First, we construct a ground set of size $k-1$ on an arbitrary cluster $S$.
-In any configuration, there exists an alone node on $S$.
-We issue requests between the current alone node from $S$, and some node that was initially collocated with it.
+In any configuration, there exists an isolated node on $S$.
+We issue requests between the current isolated node from $S$, and some node that was initially collocated with it.
 Repeating such requests, almost all the nodes must visit the cluster $S$.
@@
 	Fix any online algorithm \ALG{}.
-	Let $I(C)$ denote the cluster where nodes of the ground set $C$ are located initially.
+	Let $I(C)$ denote the cluster where nodes of the ground set $C$ are located in the initial configuration.
 	If $C$ contains nodes that originate from different clusters, $I(C)$ is undefined.
-	\mahmoud{i.e., $I(C):=...$formalized with notations}.
 	Throughout the construction, we maintain ground sets for each node.
-	Initially all nodes are \emph{alone} (they belong to their own trivial ground sets).
-\mahmoud{isolated would be a better word for "alone"}
-	First, we create a ground set of size $k-1$ on an arbitrarily chosen cluster $S$.
-	\mahmoud{$\mathcal{S}$ is better for a set of sets.}
+	Initially all nodes are \emph{isolated}.
+	First, we create a ground set $B$ of size $k-1$ on an arbitrarily.
 	%We never dismiss any constructed ground set.
-	Each cluster hosts exactly $k$ nodes, and in any configuration a single alone node resides in $S$.
-\mahmoud{an isolated node is collocated with  this GS in $\mathcal{S}$.
-We denote the first such node by $x_0$.}
-
-	Then, we add the alone node from $S$ to a larger ground set to force its eviction.
+	Each cluster hosts exactly $k$ nodes, and in any configuration a single isolated node resides in $S$.
+	We say that this isolated node that accompany the ground set $B$ is \emph{special}.
+	We denote the first special node by $x_0$.
+
+	Then, we add the special node to a larger ground set to force its eviction.
 	Precisely,
 	we create a ground set $\{x_0, y_0\}$, 
-	where $x_0$ is the alone node from $S$ and $y_0$ is an arbitrary node that does not reside in $S$.
+	where $x_0$ is the current special node and $y_0$ is an arbitrary node that does not reside in $S$.
 	\ALG has them split, and the adversary issues external requests until \ALG collocates them.
-	The cluster $S$ already contains a ground set of size $k-1$, hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.
+	There is no space to accommodate $\{x_0, y_0\}$ in the cluster of $x_0$ (as $B$'s size is $k-1$), hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.
 	To preserve a feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
-	\ALG must replace $x_0$ with another single node.
-	\mahmoud{In any new partition, \ALG must replace $x_0$..., otherwise it is not a feasible partition.}
-
-	We continue adding the node that is currently alone in $S$ to a larger ground set.
-	Precisely, for an alone node $x$ from~$S$, we merge the ground sets of $x$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x)$.
-	\mahmoud{$x_i$?}
-	This way, $x$ is evicted similarly to $x_0$, and we continue creating such ground sets.
+	\ALG must replace $x_0$ with another single node that we call $x_1$.
+
+	We continue adding the current special node to a larger ground set.
+	Precisely, for an isolated node $x_i$ from~$S$, we merge the ground sets of $x_i$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x_i)$.
+	This way, $x_i$ is evicted similarly to $x_0$, and we continue creating such ground sets.
 	We describe the termination condition shortly.
@@
 
-	We claim that expanding ground sets in this way can continue as long as at least $\ell$ alone node exists.
-	Fix the moment just before the creation of a ground set with the alone node \mahmoud{$x_{t^*}$?} that currently resides in $S$.
-	\mahmoud{ALG might change $S$. We can say, "currently collocated with the...".}
-	We show that if at least $\ell$ alone nodes exists (including $z$), then there exists a perfect partition of ground sets after expanding a ground set with $z$.
-	\mahmoud{the termination condition is needed before the claim statement.}
+	We claim that expanding ground sets in this way can continue as long as at least $\ell$ isolated nodes exists.
+	Fix the configuration just before the creation of a ground set with $x_i$ for some $i$.
+	We show that if at least $\ell$ isolated nodes exists (including $x_i$), then there exists a perfect partition of ground sets after expanding a ground set with $x_i$.
 	To show this claim, we consider a configuration, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
-	In this configuration, each cluster contains:  at most one ground set larger than $1$, possibly some alone nodes, and possibly the ground set $\{x_0, y_0\}$.
-	If $I(z)$ contains an alone node, then we simply obtain a~perfect partition by swapping it with $z$.
-	In the following we assume that no alone nodes exist on $I(z)$.
-	The largest ground set of $I(z)$ is at most $k-1$, because $z$ is on $S$.
-	Therefore $\{x_0, y_0\}$ is present on $I(z)$, and we must evict it to make space for $z$.
-	To find the cluster for $\{x_0, y_0\}$, observe that we have at most $\ell-1$ clusters with alone nodes, and we have at least $\ell$ alone nodes --- hence a cluster with two alone nodes exists.
-	To find a feasible partition, we swap $\{x_0, y_0\}$ with these two collocated alone nodes, and then we swap an alone node from $I(z)$ with $z$.
-
-	To keep \OPT's cost low, we end extending ground sets prematurely --- while there are still $\ell+1$ alone nodes (i.e., neglecting two last ground set expansions).
+	In this configuration, each cluster contains:  at most one ground set larger than $1$, possibly some isolated nodes, and possibly the ground set $\{x_0, y_0\}$.
+	If $I(x_i)$ contains an isolated node, then we simply obtain a~perfect partition by swapping it with $x_i$.
+	In the following we assume that no isolated nodes exist on $I(x_i)$.
+	The largest ground set of $I(x_i)$ is at most $k-1$, because $x_i$ is on $S$.
+	Therefore $\{x_0, y_0\}$ is present on $I(x_i)$, and we must evict it to make space for $x_i$.
+	To find the cluster for $\{x_0, y_0\}$, observe that we have at most $\ell-1$ clusters with isolated nodes, and we have at least $\ell$ isolated nodes --- hence a cluster with two isolated nodes exists.
+	To find a feasible partition, we swap $\{x_0, y_0\}$ with these two collocated isolated nodes, and then we swap an isolated node from $I(x_i)$ with $x_i$.
+
+	To keep \OPT's cost low, we end extending ground sets prematurely --- while there are still $\ell+1$ isolated nodes (i.e., neglecting two last ground set expansions).
 	Consider \ALG's configuration after the last expansion.
-	Among $\ell+1$ alone nodes there must exist a pair that originates from the same cluster (we have $\ell$ clusters). 
+	Among $\ell+1$ isolated nodes there must exist a pair that originates from the same cluster (we have $\ell$ clusters). 
 	We denote these two nodes by $x^*$ and $y^*$.
@@
 	\ALG performs at least one swap for every expansion of a ground set.
-	Each expansion reduces the number of alone nodes by $1$.
-	We start with $k \cdot \ell - (k-1)$ alone nodes (a single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with alone $\ell+1$ nodes.
+	Each expansion reduces the number of isolated nodes by $1$.
+	We start with $k \cdot \ell - (k-1)$ isolated nodes (a single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
 	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - l) / 2$.
@@
 	From that, we obtain $|\sigma_E| \geq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
-	Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, we are ready to bound the competitive ratio
+	Let $E := \OPT(\sigma_E) + \OPT(M)$.
+
+	Finally, we are ready to bound the competitive ratio
 	\begin{equation*}

### 2020-05-08 paper/disc20clustering.tex
Consistency

@@
 
-\maciek{TODOs: "requests between nodes" to "requests to pair of nodes"}
-
 \section{Introduction}
@@
 
-\subsection{Model.}
+\subsection{Model}
 
@@
 
-\subsection{Our Contributions.}
+\subsection{Our Contributions}
 We provide an improved lower bound 
@@
 
-\section{Preliminaries.}
+\section{Preliminaries}
 \label{sec:prelim}
@@
 
-\subsection{Lower bound.}
+\subsection{Lower Bound}
 
@@
 
-\subsection{Upper bound.}
+\subsection{Upper Bound}
 \label{sec:ppl}
@@
 
-\subsection{Improved algorithm for Online Rematching}
+\subsection{Improved Algorithm for Online Rematching}
 \label{sec:k2}
@@
 
-\textbf{The charging Scheme.}
+\textbf{The charging scheme.}
 Consider two  pairs that share a same node, 
@@
 
-\section{Distributed implementation.}
+\section{Distributed Implementation}
 The algorithm PPL from Section~\ref{sec:ppl} can be distributed
@@
 
+\begin{comment}
+
 \section{Omitted results (for our future papers)}
@@
 
-\begin{comment}
 	\section{Bounding the cost of a single reconfiguration of components}

### 2020-05-08 paper/disc20clustering.tex
wip

@@
 
- \newcommand\mahmoud[1]{\color{orange}\textbf{Mahmoud: #1~}\color{black}}
+ %\newcommand\mahmoud[1]{\color{orange}\textbf{Mahmoud: #1~}\color{black}}
  \newcommand\stefan[1]{\color{blue}\textbf{Stefan: #1}\color{black}}
- \newcommand\maciek[1]{\color{brown}\textbf{(Maciek: #1)}\color{black}}
- %\newcommand\mahmoud[1]{}
+ %\newcommand\maciek[1]{\color{brown}\textbf{(Maciek: #1)}\color{black}}
+ \newcommand\mahmoud[1]{}
  %\newcommand\stefan[1]{}
- %\newcommand\maciek[1]{}
+ \newcommand\maciek[1]{}
 
@@
 
-	Finally, we are ready to bound the competitive ratio
+	Finally, we bound the competitive ratio
 	\begin{equation*}
@@
 	It remains to consider the last, unfinished phase.
-	Consider the case, where the unfinished phase is also the first one.
+	First, consider the case, where the unfinished phase is also the first one.
 	Then, we cannot charge $\OPT$ due to the inability to partition the components.
@@
 	If there were at least $\alpha$ external requests, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a migration; for the cost of \TAlg we follow the analysis of the unsaturated pairs.
-	Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
+	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
 	This way, the competitive ratio increases at most twofold.

### 2020-05-08 paper/disc20clustering.tex
Comments back
Comments
wip

@@
 
- %\newcommand\mahmoud[1]{\color{orange}\textbf{Mahmoud: #1~}\color{black}}
+ \newcommand\mahmoud[1]{\color{orange}\textbf{Mahmoud: #1~}\color{black}}
  \newcommand\stefan[1]{\color{blue}\textbf{Stefan: #1}\color{black}}
- %\newcommand\maciek[1]{\color{brown}\textbf{(Maciek: #1)}\color{black}}
- \newcommand\mahmoud[1]{}
+ \newcommand\maciek[1]{\color{brown}\textbf{(Maciek: #1)}\color{black}}
+ %\newcommand\mahmoud[1]{}
  %\newcommand\stefan[1]{}
- \newcommand\maciek[1]{}
+ %\newcommand\maciek[1]{}
 
@@
 
+\maciek{TODO style: documentclass[a4paper,anonymous,USenglish] lipics-v2019.}
+\maciek{Keywords: distributed computing?}
+
 We study a problem of managing the virtual machines in a data center.
@@
 By $\Rep$ we denote the cost of a single component repartition, and we elaborate on this in Section~\ref{sec:k3}.
-The results pertain to the general partitioning model, but the last row, that pertains to the learning model for arbitrary $k$ and $\ell$.
+The first three rows show the results for the general partitioning model, and last row shows the results for the learning model for arbitrary $k$ and $\ell$.
 	}
@@
 For $k=2$, the learning problem is trivial: immediate collocation of a communicating pair is $1$-competitive.
-(A partitioning problem for $k=2$ is non-trivial, a lower bound of $3$ is known \cite{repartition-disc}, and we provide a $6$-competitive algorithm (see Section~\ref{sec:k2}).)
+(A partitioning problem for $k=2$ is non-trivial, a lower bound of $3$ is known \cite{repartition-disc}, and we provide a~$6$-competitive algorithm (see Section~\ref{sec:k2}).)
 
@@
 	\label{lem:1req}
-	During a single reconfiguration of nodes, \TAlg exchanges at most $2$ pairs of nodes.
+	In a single reconfiguration of nodes (after a merge of components), \TAlg exchanges at most $2$ pairs of nodes.
 \end{lemma}

### 2020-05-08 paper/disc20clustering.tex
Wip
intro for rematching

@@
  \newcommand{\PPL}{\textsf{PPL}\xspace}
- \newcommand{\OBRP}{BRP}
+ \newcommand{\OBRP}{BRP\xspace}
  \newcommand{\PPOBRP}{PP-BRP}
@@
 
-\subsection{Improved Algorithm for Online Rematching}
-\label{sec:k2}
-
-\maciek{Intro: compare to previous 7-competitive algorithm. Note strict competitive ratio vs large additive term. Also, emphasize that in this section the algorithm is different than DET2}
+\subsection{Improved algorithm for Online Rematching} \label{sec:k2}
+In this section,
+we present \RM,
+ an algorithm for the \OBRP restricted to clusters of capacity 2,
+i.e.~$k=2$.
+We interpret a pair of nodes collocated in one cluster as a matched pair.
+Hence,
+the problem is an online variant of the classic matching problem where
+a matched pair can separate from each other in order to ``rematch'' with two other nodes.
+This is known as the  \emph{Online Rematching} 
+problem and a (non-strict) 7-competitive algorithm is already given by \cite{repartition-disc},
+in which the ratio comes with an additive factor in $O(\alpha\ell^2)$.
+We do not only improve upon their competitive ratio,
+but also show that our ratio holds \emph{strictly}
+(i.e., with no additive factor).
+In addition,
+our algorithm is similar
+and slightly simpler than the one in \cite{repartition-disc}, 
+while our analysis is significantly more simple and concise,
+thanks to the charging scheme we devise here.
 
@@
 since it cannot have both pairs initially collocated.
-%More precisely,
-%OPT  pays a total cost at least
-%	$\min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$ for, respectively,
-%	serving requests to both pairs remotely,
-%	serving requests to one pair remotely and collocating the other,
-%	or collocating both pairs. 
-%
-%for serving (some of) these requests remotely or (and) collocating (one of) the pairs,
-%regardless of its choices.
-%Then the cost per pair, in particular the cost w.r.t.~$\{u,v\}$, is
-%$\mathit{OPT} (\sigma_{\{u,v\}}) \geq c_{R_t} / 2$.
-%
 However,
@@
 	$ C^t_{\set{u,v}} = \lambda$.
-	%	In case the first collocation of $\set{u,v}$ occurs at $t$,
-	%	if the pair is collocated in the initial configuration then
-	%	we set $t''=0$,
-	%	else,
-	%	we set $ t'' = -\infty, t'=0$.
 	We denote the set of all requests to a pair $\set{x,y}$ that arrive
@@
 $.
-%	\begin{align*}	%\label{eq:costAlg}
-%	\mathit{RM} (\sigma)
-%	=
-%	\sum_{ t \in T}(\lambda + 2\alpha) +
-%	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
-%	\leq
-%	\sum_{ t \in T}(\lambda + 2\alpha) +
-%	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
-%	=
-%	\sum_{ t \in T} 3\alpha +
-%	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
-%	.
-%	\end{align*}
 
@@
 	$\OPTM(R_t) \geq  \min{\{ \lambda,\alpha \}}$.
-	\begin{comment}
-	We distinguish three (non-exhaustive) cases for the cost to \OPT w.r.t.~$R_t$.
-	Latter,
-	 we argue that \OPT cannot save cost by deviating from these cases.
-	\begin{enumerate}[label=\roman*.]
-		\item
-		\OPT collocates $\{u,w\}$ at  $t''$, $\{u,v\}$  at $t'$,
-		and keeps them so during $(t'',t']$ and $(t',t]$, respectively.
-		Therefore it serves all requests in $R_t$ with no additional cost and
-		let $\OPTM_1 (R_t) := 2\alpha$.
-		\item
-		\OPT has one of the pairs  already collocated and the other pair separated,
-			  prior to and during their respective intervals.
-		Hence,
-		it pays only $\lambda$ for serving all requests in $R_t$
-		and let $\OPTM_2 (R_t) := \lambda$.
-		\item
-		\OPT has one of the pairs already collocated prior to its interval,
-		and it collocates the other pair during the respective interval.
-		Then it pays 0 for one pair and $\alpha$ for the other and 
-		let $\OPTM_3 (R_t) := \alpha$.
-	\end{enumerate}
-%	We consider cases (ii) and (iii) also for the pair $\{u,w\}$, symmetrically.
-	If \OPT behaves in one of these cases,
-	then
-	$\mathit{OPT} (R_t) \geq
-		\min{\{ \OPTM_1(R_t), \OPTM_2(R_t), \OPTM_3(R_t) \}}  =
-		\min{ \{ \lambda, \alpha \}}$.
-	Otherwise,
-	\OPT deviates from these cases,
-	e.g.~by collocating a pair after paying for some of requests to this pair.
-	However,
-	no deviation can decrease the cost to \OPT. \mahmoud{prove this!}
-	Therefore in any case,
-	$\mathit{OPT} (R_t) \geq  \min{ \{ \lambda, \alpha \}}$.
-	\end{comment}
 

### 2020-05-09 paper/disc20clustering.tex
Related Work

@@
 
- \title{Tight Bounds for Online Balanced Graph Partitioning}
+ \title{Tight Bounds for Dynamic Balanced Graph Partitioning}
 
@@
 	By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), we can reduce the network load and  improve application performance. 
-	However, the communication pattern is a priori unknown and may change over time, hence it needs to be learned in an online manner.
+	However, the communication pattern of different application is often unknown a priori, and may change over time, hence it needs to be learned in an online manner.
 	%
@@
 	We improve the result for $k=2$ by providing a strictly $6$-competitive upper bound for the general model.
+	In contrast to most prior work, our algorithms respect all capacity constraints, and do not require resource augmentation.
 
@@
 
-We study a problem of managing the virtual machines in a data center.
-We focus on the scenario where virtual machines can be migrated during runtime to another physical machine.
-The possibility of migration allows reacting to unpredictable communication patterns.
-For example, if some distant virtual machines communicate often, it is vital to reduce their distance to save network bandwidth.
-The objective is to~minimize the total network bandwidth used for communication and for migration.
-
-Intuitively, we would like to minimize inter-machine
-interactions by mapping frequently communicating virtual machines to the same physical machine
-Since communication patterns change over time, the~placement of virtual machines should be managed in an online manner, by migrating them between physical machines.
-The~objective is to minimize the weighted sum of inter-machine communication and repartitioning costs.
-The former is defined as the number of communication requests between virtual machines placed at distinct physical machines, and the latter as the number of migrations.
+The popularity of data-centric, distributed applications has led to an explosive growth of network traffic, especially in data centers \cite{roy2015inside,singh2015jupiter}.
+As the performance of these distributed applications often critically depends on the underlying network \cite{mogul2012we}, an efficient operation of these networks is important.
+At the same time, distributed systems are often highly virtualized today, and provide interesting new opportunities for resource optimization.
+In particular, it has become possible to operate data centers in a more demand-aware manner: 
+by dynamically migrating nodes (e.g., virtual machines) which communicate frequently topologically closer to each other, network traffic can be reduced significantly.  
+However, migrations should be used moderately, as migrations also entail overheads.
+
+This paper studies the algorithmic problem underlying such demand-aware
+optimizations, aiming to strike a balance between the benefits of migrations (e.g., reduced network load) and their costs.
+In particular, we are interested in an online variant of the problem: since communication patterns can change over time, an online algorithm needs to react dynamically to new traffic patterns, and migrate nodes  accordingly.
+Ideally, this algorithm should perform close to an optimal offline algorithm, without requiring any information about future traffic demands.
+
+This problem is known as the dynamic balanced graph partitioning problem introduced by Avin et al.
+\cite{repartition-disc, sidma-arxiv} at DISC 2016.
+A special variant of the general problem has later been studied by Henzinger et al.
+\cite{sigmetrics19_partitioning} at SIGMETRICS 2019.
+We will refer to the latter as the learning model.
 
@@
 
-We study two models in this paper: a general \emph{partitioning} model, and its subproblem, a \emph{learning} variant.
-The general partitioning model was introduced by Avin et al.~\cite{repartition-disc} at DISC 2016.
-The learning model was introduced by Henzinger et al.~\cite{sigmetrics19_partitioning} at SIGMETRICS 2019.
-
-\noindent
-\textbf{Partitioning model.}
-The \emph{balanced repartitioning} problem (\OBRP{})
-is a fundamental learning problem
-that finds applications in the context of
-distributed systems optimization~\cite{repartition-disc}. We are given a set $V$ of $n$ nodes 
+We study two models in this paper: the \emph{general partitioning} model, and its subproblem, the \emph{learning} model.
+
+\noindent
+\textbf{General partitioning model.}
+In the \emph{dynamic balanced graph partitioning} problem, we are given a set $V$ of $n$ nodes 
 (e.g., virtual machines or processes),
@@
 To avoid infinite cost, each communicating pair must be collocated.
-We study instances with finite cost: we assume that the communication graph can be perfectly partitioned in the clusters (no external requests are required).
-In the partitioning model, $\alpha$ denoted the ratio between the cost of a single communication request and the cost of migration; in the learning model, we set $\alpha = 1$.
+As in Henzinger et al. \cite{sigmetrics19_partitioning}, we assume that the communication graph can be perfectly partitioned in the clusters (no external requests are required).
+For the learning model, we will assume that the migration cost is $\alpha=1$.
 The algorithm's goal is to \emph{learn} the communication graph, and partition it without too many reconfigurations.
@@
 
+\noindent
+\textbf{Algorithmic challenges.}
+\maciek{Integer partition in non-augmented problem. Also: what to evict, rent or buy etc.}
+
 \subsection{Related work}
 
-The existing results in~\cite{repartition-disc} and~\cite{sigmetrics19_partitioning} primarily focus on a model with resource augmentation, where  the online algorithm has slightly larger clusters than the offline algorithm.
-They showed that their lower bound $\Omega(k)$ holds even for a significant resource augmentation.
-They provided an algorithm with the competitive ratio independent of $\ell$ using the $(2+\epsilon)$-augmented cluster capacity.
+The two works closest to ours are by Avin et al.
+(on the general partitioning model) \cite{repartition-disc, sidma-arxiv} and by Henzinger et al.
+(on the learning model) \cite{sigmetrics19_partitioning}.
+The focus of these papers however is primarily on models with resource augmentation: the online algorithm can use slightly larger clusters than the offline algorithm.  
+Avin et al.
+actually showed that their lower bound $\Omega(k)$ holds even for a significant resource augmentation, and they provided an algorithm with the competitive ratio $O(k \log k)$ using the $(2+\epsilon)$-augmented cluster capacity.
+Their ratio is independent of $\ell$, which is impossible without significant resource augmentation.
 
 In contrast, we study the non-augmented setting, where the nodes need to be perfectly balanced  among the clusters.
-In the paper that originated a study on BRP \cite{repartition-disc} (DISC 2016), authors provided a $O(k^2 \cdot \ell^2)$-competitive algorithm, and a lower bound $\Omega(k)$.
-For $k=2$, they presented a $7$-competitive algorithm~\cite{repartition-disc} with a~substantial ($\Omega(\ell^2)$) additive constant.
+This assumption is not only more realistic but also significantly more challenging, as it is related to hard problems such as integer partitioning \cite{integer-partitions-book}.
+In terms of results without augmentation, so far, it is only known that there exists a $O(k^2 \cdot \ell^2)$-competitive algorithm \cite{repartition-disc}; the best known lower bound is significantly lower, namely $\Omega(k)$.
+For $k=2$, Avin et al.\cite{repartition-disc} presented a $7$-competitive algorithm with a~substantial ($\Omega(\ell^2)$) additive constant.
 

### 2020-05-09 paper/disc20clustering.tex
Related work finished

@@
 
-\noindent
-\textbf{Algorithmic challenges.}
-\maciek{Integer partition in non-augmented problem. Also: what to evict, rent or buy etc.}
-
 \subsection{Related work}
@@
 
-The studied model is related to online
-paging~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}, sometimes also referred to
+Our problem is further related to some classic online problems.
+In particular, it is related to online paging~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}, sometimes also referred to
 as online caching, where requests for data items (nodes) arrive over time and
@@
 A major difference between  these problems is that in the caching problems, each request involves a~single element of the universe, while in our model \emph{both} endpoints of a communication request are subject to~optimization.
+In this light, we can see our model as a "symmetric" version of online paging.
+
+Dynamic graph partitioning problems are generally fundamental in computer science, and arise in many different contexts \cite{streaming-soda,streaming1}.
 
@@
 By $\Rep$ we denote the cost of a single component repartition, and we elaborate on this in Section~\ref{sec:k3}.
-The first three rows show the results for the general partitioning model, and last row shows the results for the learning model for arbitrary $k$ and $\ell$.
+The first three rows show the results for the general partitioning model, except for the last row that shows the results for the learning model for arbitrary $k$ and $\ell$.
 	}

### 2020-05-09 paper/disc20clustering.tex
Contributions

@@
 \subsection{Our Contributions}
-We provide an improved lower bound 
-of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
-even in the learning model;
-the best known lower bound so far was $\Omega(k)$,
+
+This paper presents several new results on the dynamic graph partitioning problem  without augmentation.
+Our main result is a tight bound on the general case: we present a lower bound of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
+even in the learning model.
+The best known lower bound so far was $\Omega(k)$,
 even for a more general model~\cite{repartition-disc}.
@@
 so far $O(\ell^2)$~\cite{repartition-disc}.
-Then, we present a strictly $6$-competitive algorithm for $k=2$ that improves upon previous $7$-competitive algorithm with $O(\ell^2)$ additive constant.
+We further present a strictly $6$-competitive algorithm for $k=2$ that improves upon previous $7$-competitive algorithm with $O(\ell^2)$ additive constant.
 %
-The overview of known results is in Table~\ref{tab:overview}.
+Table \ref{tab:overview} provides an overview of our contributions compared to prior work.
 

### 2020-05-09 paper/disc20clustering.tex
PPL comments from Stefan

@@
 
-\section{Preliminaries}
-\label{sec:prelim}
-
-Across this paper, we often refer to groups of communicating nodes.
-We use this concept slightly differently in the lower bound and the upper bounds.
-In our algorithms, we group nodes with a communication history into \emph{components}, and in the lower bound, we group nodes that may communicate into \emph{ground sets}.
-
-\noindent
-\textbf{Components.}
-In our algorithms, a~\emph{component} is a subset of frequently communicating nodes.
-We maintain all nodes of each component collocated in one cluster~\cite{repartition-disc}.
-If we need to move a node, we move the whole component that contains it.
-We maintain a balanced partition of our components, a similar concept to partition of integers into sets with equal sum (\maciek{TODO: cite}).
-In contrast, our partition is time-varying: communicating components are merged into one, and we adjust the partition.
-
-\noindent
-\textbf{Ground sets.}
-In the lower bound, a~\emph{ground set} is a group of nodes that repeatedly communicate when an algorithm splits them (places at different clusters).
-Their role is to ensure that the algorithm maintains a partition of ground sets into clusters.
-We emphasize that ground sets are revealed only upon their nodes are split.
-%\maciek{Other names: division, segmentation, alignment, formation}
-
-%We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
-%This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
-%such that frequently communicating nodes (from different clusters)
-%are relocated to the same cluster before they inflict too much communication cost.
-%A \emph{reconfiguration} (i.e.,\emph{ re-partitioning})
-%is performed by migrating nodes between clusters,
-%at the total cost of individual node migrations.
-%The objective is to jointly minimize communication and migration costs while serving the sequence.
-
-We say that a component or a ground set is \emph{trivial}, if it contains exactly one node.
-We say that the node that belongs to trivial component or a ground set is \emph{alone}.
-We say that a~configuration is \emph{component-respecting}
-if the nodes belonging to the same component are collocated.
-
-\noindent
-\textbf{Note on our algorithms.}
-\maciek{Should not be here probably, but where?}
-Two of our algorithms (for the learning model and for $k=3$ in the partitioning model) are modified versions of the algorithm DET from \cite{repartition-disc}.
-The difference is in the the choice of partition after a component merge.
-In DET, the partition was arbitrary.
-In PPL, we choose the partition closest to the initial configuration.
-In \TAlg, we choose the partition closest to the previous configuration (a repartition of minimum cost).
-
 \section{The Learning Problem} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
 
-In this section we consider the learning variant of Online Balanced Partitioning.
+In this section we consider the learning variant of Dynamic Balanced Graph Partitioning.
 For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$.
@@
 We provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm for the learning problem.
-Later we elaborate on how to efficiently transform it to a lower bound for the partitioning problem.
+Later we elaborate on how to efficiently transform it to a lower bound for the general partitioning problem.
 The lower bound requires $k\geq 3$.
@@
 
+Across this paper, we often refer to groups of communicating nodes.
+We use this concept slightly differently in the lower bound and the upper bounds.
+In our algorithms, we group nodes with a communication history into \emph{components}, and in the lower bound, we group nodes that may communicate into \emph{ground sets}.
+
+A~\emph{ground set} is a group of nodes that repeatedly communicate when an algorithm splits them (places at different clusters).
+Their role is to ensure that the algorithm maintains a partition of ground sets into clusters.
+We emphasize that ground sets are revealed only upon their nodes are split.
+We say that a ground set is \emph{trivial}, if it contains exactly one node.
+We say that the node that belongs to trivial a ground set is \emph{isolated}.
+
 The adversary constructs ground sets (a set of nodes that may communicate, see Section~\ref{sec:prelim}) depending on choices of a deterministic online algorithm.
@@
 	\label{th:lowerbound}
-	The competitive ratio of any deterministic online algorithm for \OBRP is at least $(k\cdot \ell - k - \ell)/2$ for any $k\geq 3$ and $\ell \geq 2$.
+	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least $(k\cdot \ell - k - \ell)/2$ for any $k\geq 3$ and $\ell \geq 2$.
 \end{theorem}
@@
 	Initially all nodes are \emph{isolated}.
-	First, we create a ground set $B$ of size $k-1$ on an arbitrarily.
+	First, we create a ground set $B$ of $k-1$ nodes on an arbitrarily chosen cluster (the nodes  are already collocated).
 	%We never dismiss any constructed ground set.
 	Each cluster hosts exactly $k$ nodes, and in any configuration a single isolated node resides in $S$.
-	We say that this isolated node that accompany the ground set $B$ is \emph{special}.
+	We say that this isolated node that accompanies the ground set $B$ is \emph{special}.
 	We denote the first special node by $x_0$.
@@
 	Fix the configuration just before the creation of a ground set with $x_i$ for some $i$.
-	We show that if at least $\ell$ isolated nodes exists (including $x_i$), then there exists a perfect partition of ground sets after expanding a ground set with $x_i$.
+	We show that if at least $\ell$ isolated nodes exist (including $x_i$), then there exists a perfect partition of ground sets after expanding a ground set with $x_i$.
 	To show this claim, we consider a configuration, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
@@
 The algorithm immediately collocates every communicating pair and never separates them.
-In other words, it maintains components, introduced in the preliminaries (Section~\ref{sec:prelim}).
-
-Upon receiving any external request, two components merge --- and this requires a reconfiguration.
-Many configurations may be feasible, and we choose the one that is \emph{closest to the initial configuration}.
+To help maintain the collocation of communicating pairs, we use the concept of components, introduced by Avin et al. \cite{repartition-disc}.
+
+In our algorithms, a~\emph{component} is a subset of frequently communicating nodes.
+We maintain all nodes of each component collocated in one cluster~\cite{repartition-disc}.
+If we need to move a node, we move the whole component that contains it.
+We maintain a balanced partition of our components, a similar concept to partition of integers into sets with equal sum \cite{integer-partitions-book}.
+In contrast, our partition is time-varying: communicating components are merged into one, and we adjust the partition.
+
+We say that a component is \emph{trivial}, if it contains exactly one node.
+We say that the node that belongs to trivial component \emph{isolated}.
+We say that a~configuration is \emph{component-respecting}
+if the nodes belonging to the same component are collocated.
+
+The algorithm in this section, and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from \cite{repartition-disc}.
+The difference is in the the choice of partition after a component merge.
+In DET, the partition was arbitrary.
+In the algorithm for the learning model, we choose the partition closest to the initial configuration.
+In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the previous configuration (a repartition of minimum cost).
 
@@
 \maciek{Algorithm pseudocode comments}
+\maciek{Stefan's comment: "container" components undefined}
 \maciek{Major thing to discuss: do we even need this?

### 2020-05-09 paper/disc20clustering.tex
k=2 Stefan's comments

@@
 \textbf{Component-based algorithm.}
-Now we describe the algorithm \TAlg that we use throughout this section.
+Now we describe the algorithm \TAlg.
 The algorithm partitions nodes into components, and
@@
 	In this case, a local repartition is possible (within $U$ and $V$), for a cost of at most $2$ swaps.
-	Note that it is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, a components of size $3$ participates in a merge, which results in a component of size at least $4$, and in this case no component respecting partitioning exists, a contradiction.
+	Note that it is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, a components of size $3$ participates in a merge.
+	After merge, we would have a component of size at least $4$, and in this case no component respecting partitioning exists, a contradiction.
 
@@
 	Now we bound $\TAlg(\sigma_I)$.
-	A~cluster of type $C_3$ contributes at most $3 \alpha - 1$ to $\TAlg(\sigma_I)$, as $2$ of pairs of nodes from the component are saturated and contribute $\alpha$ each, and the third, unsaturated pair contributes at most $\alpha-1$.
+	A~cluster of type $C_3$ contributes at most $3 \alpha - 1$ to $\TAlg(\sigma_I)$, as two pairs of nodes from the component are saturated and contribute $\alpha$ each, and the third, unsaturated pair contributes at most $\alpha-1$.
 	Other cluster types contribute less: $C_1$ contributes $0$ and $C_2$ contributes $\alpha$.
@@
 	i.e., at (time) $t$,
-	Alg collocates the pair $\set{u,v}$.
+	\RM collocates the pair $\set{u,v}$.
 	This means,
@@
 
-\section{Distributed Implementation}
+\section{Discussion and Future Work}
 The algorithm PPL from Section~\ref{sec:ppl} can be distributed

### 2020-05-09 paper/disc20clustering.tex
Distributed

@@
 \section{Discussion and Future Work}
+
+While we have described our algorithms globally so far, we note that these allow for efficient distributed implementations. 
 The algorithm PPL from Section~\ref{sec:ppl} can be distributed
@@
 \medskip
-
-In the rest of this section, we propose the distributed implementation of the algorithm for $k=3$ from Section~\ref{sec:k3}.
+Now, we propose the distributed implementation of the algorithm for $k=3$ from Section~\ref{sec:k3}.
 The algorithm performs two non-local operations:
@@
 
-If a component reaches size $4$ or fresh cluster is not found, we conclude that no perfect partition exists.
-Then, we end the phase and broadcast (\maciek{TODO: efficient broadcast cite}) the message to  reset the counters in all clusters.
-
-A merge of two components may require finding a third cluster that contains $3$ alone nodes (for details see the proof of Lemma~\ref{lem:1req})).
-We say that a cluster containing $3$ alone nodes is \emph{fresh}.
+If a component reaches size $4$ or a fresh cluster is not found, we conclude that no perfect partition exists.
+Then, we end the phase and broadcast the message to  reset the counters in all clusters.
+
+A merge of two components may require finding a third cluster that contains $3$ isolated nodes (for details see the proof of Lemma~\ref{lem:1req})).
+We say that a cluster containing $3$ isolated nodes is \emph{fresh}.
 In the following, we show how to efficiently find a fresh cluster in a distributed manner.
-We organize the clusters into an arbitrary rooted balanced binary tree (\maciek{TODO: how to do it efficiently?
-Does not matter much, as this is performed once at the beginning}), and we broadcast the root to each cluster.
+We organize the clusters into an arbitrary rooted balanced binary tree, and we broadcast the root to each cluster.
 Each cluster maintains the counter of fresh clusters in its subtree.
@@
 
-The operation (\ref{it:broadcast}) requires a single broadcast, and the operation (\ref{it:merge}) requires $O(\log(\ell))$ communication.
+The operation (\ref{it:broadcast}) requires a single broadcast, and the operation (\ref{it:merge}) requires $O(\log \ell)$ communication.
 Each of these may occur, on average, every $\alpha$ external requests.
+
+\medskip
+
+This paper revisited the dynamic graph partitioning problem and presented several tight bounds for the important model where capacities cannot be exceeded, both for a general partitioning model and for a special learning model.
+
+While are bounds are tight, there are several interesting avenues for future research.
+In particular, we have so far focused on deterministic algorithms, and it would be interesting to study the power of randomization in this context.
+On the practical side, it would also be interesting to study our algorithms empirically, under realistic workloads.
 

### 2020-05-09 paper/disc20clustering.tex
Cleanup of comments (for backup, unroll this commit)

@@
 
-\begin{comment}
-
-\section{Omitted results (for our future papers)}
-
-\begin{enumerate}
-	\item Costly cascade example (it is easy to present $k^2$ and $k\cdot \sqrt{k}$ (with different approach), and then we can combine these to obtain $k^3$)
-
-	\item $k$-competitive algorithm for the ring communication pattern
-
-	\item lower and upper bound $k+1$ for two clusters, perfect partition and the optimal configuration reachable within one swap
-
-	\item The lower bound $(k \cdot \ell - k - \ell) / 2$ can be easily improved to $(k \cdot \ell - k - 1) / 2$ by playing a simple caching game in the first cluster, instead of just giving away the $k-1$ "blocker" (the caching game would build it gradually)
-
-\end{enumerate}
-
-	\section{Bounding the cost of a single reconfiguration of components}
-
-	\maciek{Overview of the section, intro. Defer introducing the graph notation, the warm-up does not need it}
-
-	\subsection{Characterization of the migration graph}
-
-	Consider a component merge action performed by \TAlg that triggers a reconfiguration from a configuration $C_I$ to a configuration $C_F$.
-	To execute the reconfiguration, the nodes of two non-collocated components migrate to a common cluster.
-	Each cluster contains exactly $k$ nodes, thus other nodes migrate to take the place of nodes of merged components.
-	Their place must be occupied, too, and this way a single merge action may trigger migrations in clusters not directly involved in a merge action (for an example, see Section~\ref{ssec:example}). 
-	Recall that both $C_I$ and $C_F$ are component-respecting configurations, hence a migration of a single node entails migrations of all nodes of its cluster.
-
-	In Section~\ref{ssec:cascade}, we upper bound the cost incurred by a reconfiguration that follows a single component merge.
-	To this end, we introduce a \emph{migration graph} that models the reconfiguration.
-	In the following, we characterize the structure of the migration graph.
-
-	\noindent
-	\textbf{Vertices of the migration graph and the core vertices.}
-	The vertices of the migration graph are all $\ell$ clusters of the instance.
-	We distinguish the \emph{core vertices} that correspond to clusters directly involved in the merge operation: the clusters containing the to-be-merged components in $C_I$ and the cluster containing the merged component in $C_F$.
-	There are at most $3$ core vertices in each migration graph (at most two components participate in the merge, and these may migrate to a third cluster).
-
-	\noindent
-	\textbf{Edges and their labels.}
-	The edges of the migration graph denote the migration of nodes from $C_I$ to $C_F$.
-	Each edge is labeled with a number $m$ that denotes the number of migrated nodes between clusters.
-	Each edge is directed from the cluster that contained the nodes to the cluster they migrated to.
-	Both edges might exists between a pair of nodes, as whole components of nodes migrate, and the back-and-forth exchange may be needed for the configuration $C_F$ to be component-respecting.
-
-	\noindent
-	\textbf{Vertex degree.}
-	At most $k$ nodes may migrate from a cluster and at most $k$ nodes may migrate into a cluster.
-	Thus, the sum of the labels for both ingoing and outgoing edges of each vertex is at most $k$.
-	This implies that the number of ingoing and outgoing edges is also bounded by $k$.
-
-	\noindent
-	\textbf{Flow preservation.}
-	In any feasible configuration, each cluster contains exactly $k$ nodes.
-	Thus, for any vertex, the sum of labels of ingoing edges must equal the sum of labels of outgoing edges.
-	\maciek{Can we say that this implies that the graph is a circulation?}
-
-	\noindent
-	\textbf{Migration graph and the cost of reconfiguration.}
-	The cost of cluster reconfiguration equals the number of exchanged nodes multiplied by $\alpha$.
-	From the standpoint of a migration graph, this corresponds to the sum of labels on edges of the graph.
-	A trivial upper bound on the cost is the size of the connected component in the migration graph multiplied by $2\alpha k$.
-
-	\medskip
-
-	Note that for a single component merge, multiple migration graphs may exist, and multiple of them might have the optimal cost.
-	In Section \ref{ssec:cascade}, we show that among all migration graphs for given component merge, there exists an optimal graph with the bounded cycle length.
-	Combined with other properties, this allows to bound the size of the migration graph, and consequently bound the cost of each repartition, that results in a bound of the competitive ratio for \TAlg.
-
-	\subsection{Costly cluster reconfiguration example}
-	\label{ssec:example}
-
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=0.3\textwidth]{figs/substitute}
-		\caption{A costly cascade example.
-On the left: the configuration prior to the merge and the merge of two components (denoted solid).
-On the right: corresponding migration graph for the optimal (unique) reconfiguration.}
-	\end{figure}
-
-	In this section we present an example of a merge action that results in a $k/3$(?) length cycle in an optimal migration graph.
-
-	\subsection{Bounds on the size of a migration graph}
-	\label{ssec:cascade}
-
-	In \cite{repartition-disc}, authors trivially bounded the reconfiguration cost for each merge action by $k \cdot \ell$.
-	This roughly corresponds to migrating every node in every cluster.
-	Now, we bound the cost of a single repartition of \TAlg as a function of $k$.
-	Note that the bound of $k \cdot \ell$ is still valid, and the resulting bound is the minimum of these two.
-
-	We begin, by bounding the length of each cycle in one of optimal solutions.
-
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=0.3\textwidth]{figs/substitute}
-		\caption{Illustration to bounding the cycle length.}
-		\label{fig:cascade-illustration}
-	\end{figure}
-
-	\begin{theorem}
-		\label{th:cascade-cycles}
-		Consider a single merge action of \TAlg.
-		There exists an optimal cost reconfiguration with the migration graph, where each cycle have length at most $k$.
-	\end{theorem}
-
-	\begin{proof}
-		Assume the contrary, i.e., for each optimal cluster reconfiguration, its migration graph contains a cycle at least $k+1$.
-
-		Fix a reconfiguration and its migration graph $G$ that has the shortest length $L$ of the longest cycle, and has the least number of cycles of length $L$.
-		Let $p = \lbrace v_1, v_2, \ldots, v_L \rbrace$ be any cycle of length $L$.
-		As edge labels are from $\{ 1, 2, \ldots, k \}$, there exists a pair of edges $e, f$ with equal labels.
-
-		We show that in such case, we can modify the solution without affecting its cost and feasibility to decrease the number of cycles of length $L$ by $1$.
-		This would contradict the assumption that we consider a solution with the minimal number of cycles of length $L$.
-
-		We construct a new migration graph $G'$ of equal cost by swapping the (destination) endpoints of $e$ and $f$ (cf. Figure~\ref{fig:cascade-illustration})
-		Formally, let $\lbrace a, b \rbrace = e$ and $\lbrace c, d \rbrace = f$.
-		We remove $e$ and $f$ and add $e' = \lbrace a, d \rbrace$ and $f' = \lbrace c, d \rbrace$.
-		This way we produced a feasible solution.
-
-		It remains to show that while decreasing the number of cycles of length $L$, we did not created any new cycles of length at least $L$.
-		Assume that adding $e'$ created a new cycle of length at least $L$.
-		However, $a$ and $d$ were already connected in $G$ by a directed path of length at least $2$, and a cycle strictly longer than $L$ existed in $G$.
-
-		Assume that adding $f'$ created a new cycle $s^*$ of length $L' \geq L$.
-		Contrary to the previous case, the directed cycle from $e$ to $b$ might not exist in $G$.
-		However, this would mean that $b$ and $e$ were connected by a path of length $L'-1$.
-		This means that we might have followed this alternative path from $b$ to $e$ to obtain a cycle strictly longer than $L$ (cf. Figure~\ref{fig:cascade-illustration}).
-
-	\end{proof}
-
-	\begin{theorem}
-		Consider a single repartition of DET2.
-		The number of clusters it involves is bounded by $k^k$.
-		\label{th:cascade}
-	\end{theorem}
-
-	\begin{proof}
-		Each node has degree at most $k$.
-		Every subset has flow conservation property, hence each cycle traverses one of the nodes of graph origin/core. (no subsets with just incoming edges to the origin part)
-		Combine with Theorem \ref{th:cascade-cycles}.
-
-	\end{proof}
-
-	Corollary: Algorithm DET2 is $O(\ell \cdot k \cdot \min \{ \ell, k^k \})$-competitive.
-	\end{comment}
-
 \end{document}

### 2020-05-09 paper/disc20clustering.tex
LIPICS style

@@
- 
- \documentclass[manuscript,screen=true, review, anonymous]{acmart}
+ \documentclass[a4paper,anonymous,USenglish]{lipics-v2019}
 
@@
  \usepackage{balance}
- \usepackage[english]{babel}
- \usepackage{blindtext}
  \usepackage{amsthm,amsmath,array,colortbl,graphicx,multirow}
@@
 
- \author{Maciej Pacut}
- \email{maciej.pacut@univie.ac.at}
- \orcid{0000-0002-6379-1490}
- \affiliation{%
- \institution{Faculty of Computer Science, University of Vienna}
- \country{Austria}
- }
-
- \author{Mahmoud Parham} 
- \email{mahmoud.parham@univie.ac.at}
- \orcid{0000-0002-6211-077X}
- \affiliation{%
- \institution{Faculty of Computer Science, University of Vienna}
- \country{Austria}
- }
-
- \author{Stefan Schmid} 
- \email{stefan_schmid@univie.ac.at}
- \affiliation{%
- \institution{Faculty of Computer Science, University of Vienna}
- \country{Austria}
- }
-
- \copyrightyear{2020} 
- \acmYear{2020} 
- \setcopyright{acmlicensed}
- \acmConference{DISC '20}{October 12-16, 2020}{Freiburg, Germany}
-
- \keywords{online algorithms, competitive analysis, graph partitioning, clustering}
- \acmISBN{}\acmPrice{}
- \acmDOI{}
+ \author{Maciej Pacut}{maciej.pacut@univie.ac.at}{Faculty of Computer Science, University of Vienna,Austria}{0000-0002-6379-1490}{}
+
+ \author{Mahmoud Parham}{mahmoud.parham@univie.ac.at}{Faculty of Computer Science, University of Vienna, Austria}{0000-0002-6211-077X}{}
+
+ \author{Stefan Schmid}{stefan\_schmid@univie.ac.at}{Faculty of Computer Science, University of Vienna, Austria}{}{}
+
+% \copyrightyear{2020} 
+% \acmYear{2020} 
+% \setcopyright{acmlicensed}
+% \acmConference{DISC '20}{October 12-16, 2020}{Freiburg, Germany}
+
+ \keywords{online algorithms, competitive analysis, distributed computing, graph partitioning, clustering, self-adjusting networks}
+% \acmISBN{}\acmPrice{}
+% \acmDOI{}
 
@@
 
- \newtheorem{claim}{Claim}
+ %\newtheorem{claim}{Claim}
  \newtheorem{fact}{Fact}
@@
 
-\renewcommand{\shortauthors}{M.~Pacut, M.~Parham, S.~Schmid}
+%\renewcommand{\shortauthors}{M.~Pacut, M.~Parham, S.~Schmid}
 
 \section{Introduction}
-
-\maciek{TODO style: documentclass[a4paper,anonymous,USenglish] lipics-v2019.}
-\maciek{Keywords: distributed computing?}
 
@@
 
-\bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{plainurl}
 \bibliography{references}
-
 \appendix

### 2020-05-09 paper/disc20clustering.tex
Reference fix

@@
 
-% \copyrightyear{2020} 
-% \acmYear{2020} 
-% \setcopyright{acmlicensed}
-% \acmConference{DISC '20}{October 12-16, 2020}{Freiburg, Germany}
-
+ \authorrunning{M. Pacut, M. Parham, and S. Schmid}
+ \Copyright{Maciej Pacut, Mahmoud Parham, and Stefan Schmid}
  \keywords{online algorithms, competitive analysis, distributed computing, graph partitioning, clustering, self-adjusting networks}
-% \acmISBN{}\acmPrice{}
-% \acmDOI{}
+
+% \EventEditors{}
+%\EventNoEds{3}
+ \EventLongTitle{34th International Symposium on DIStributed Computing (DISC) 2020)}
+ \EventShortTitle{DISC 2020}
+ \EventAcronym{DISC}
+ \EventYear{2020}
+ \EventDate{October 12--16, 2020}
+ \EventLocation{Freiburg, Germany (virtual conference)}
+ \EventLogo{}
+ \SeriesVolume{}
+ \ArticleNo{}
 
@@
  \ccsdesc[300]{Computer systems organization~Distributed architectures}
+ \ccsdesc[500]{Theory of computation~Online algorithms}
 
@@
 
-The adversary constructs ground sets (a set of nodes that may communicate, see Section~\ref{sec:prelim}) depending on choices of a deterministic online algorithm.
+The adversary constructs ground sets depending on choices of a deterministic online algorithm.
 We emphasize that the algorithm does not know ground sets before they are revealed.

### 2020-05-09 paper/disc20clustering.tex
refining the model section: Learning model

@@
 	pattern is static and can be perfectly partitioned;
-	we also provide a tight upper bound 
+	we also provide a \mahmoud{asymptotically?} tight upper bound 
 	of $O(k\cdot \ell)$ for this scenario.
@@
 \textbf{Learning model.}
-We study a variant of Online Balanced Partitioning, where communication requests are permanent: once a pair of nodes start communicating, they continue forever.
-To avoid infinite cost, each communicating pair must be collocated.
-As in Henzinger et al. \cite{sigmetrics19_partitioning}, we assume that the communication graph can be perfectly partitioned in the clusters (no external requests are required).
-For the learning model, we will assume that the migration cost is $\alpha=1$.
-The algorithm's goal is to \emph{learn} the communication graph, and partition it without too many reconfigurations.
+We study a variant of Online Balanced Partitioning,
+where the communication pattern is \emph{static}:
+an adversary determines the pairs of nodes that ever communicate, a priori,
+and requests to these pairs arrive indefinitely.
+Any algorithm must eventually collocate  pairs of communicating nodes,
+as otherwise it cannot be competitive.
+As in Henzinger et al. \cite{sigmetrics19_partitioning}, we assume that the communication graph admits a \emph{perfect partition},
+i.e., a partition in which no inter-cluster request ever occurs.
+%any algorithm must learn this perfect partition.
+The algorithm's objective is to \emph{learn} the (static) communication graph on-the-fly,
+without executing too many migrations.
+For the learning model, we  assume that the migration cost is $\alpha=1$.
 
 \medskip 
-In both models we assume that the communication patterns are not known in advance to our algorithm.
+In both models, we assume that communication patterns are not known to our algorithms at the beginning.
 We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.

### 2020-05-09 paper/disc20clustering.tex
refining the model section: Learning model (1)
<= 6

@@
 %any algorithm must learn this perfect partition.
-The algorithm's objective is to \emph{learn} the (static) communication graph on-the-fly,
-without executing too many migrations.
+The algorithm's objective is to \emph{learn} the (static) communication graph
+ while serving all requests,
+and without executing too many migrations.
 For the learning model, we  assume that the migration cost is $\alpha=1$.
@@
 	+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}  
-	\Big)	< 6.
+	\Big)	\leq 6.
 \end{align*}
@@
 
+\mahmoud{mention the offline approx.}
 While we have described our algorithms globally so far, we note that these allow for efficient distributed implementations. 

### 2020-05-09 paper/disc20clustering.tex
Cleanup

--- 2020-05-09 paper/disc20clustering.tex
@@
 We showed a lower bound of $\Omega(k \cdot \ell)$ that holds for $k=3$ (cf. Section~\ref{sec:lowerbound}), hence the result from this section is asymptotically optimal.
-
-Then, in Section~\ref{sec:k2}, we show a strictly $6$-competitive algorithm for $k=2$.
-\maciek{Intro: compare to previous 7-competitive algorithm. Note strict competitive ratio vs large additive term. Also, emphasize that the algorithm is different than 7-competitive}
-
-\subsection{An Optimal Algorithm for Clusters of Size $3$}
+Furthermore, in Section~\ref{sec:k2}, we show a strictly $6$-competitive algorithm for $k=2$.
+
+\subsection{An Optimal Algorithm for Clusters of Size 3}
 \label{sec:k3}
@@
 
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.3\textwidth]{figs/substitute}
-	\caption{Cluster types used in the analysis of \TAlg.}
-\end{figure}
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[width=0.3\textwidth]{figs/substitute}
+%	\caption{Cluster types used in the analysis of \TAlg.}
+%\end{figure}
 
@@
 	\includegraphics[width=0.3\textwidth]{figs/substitute}
-	\caption{an unsaturated edge that is inside a component}
+	\caption{an unsaturated edge that is inside a component. If there is a space next to it, maybe cluster types should fit next to it? Not sure.}
 \end{figure}

### 2020-05-09 paper/disc20clustering.tex
Line breaks

--- 2020-05-09 paper/disc20clustering.tex
@@
 The nodes interact using
-a sequence of pairwise communication requests
+a~sequence of pairwise communication requests
 $\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
@@
 where the communication pattern is \emph{static}:
-an adversary determines the pairs of nodes that ever communicate, a priori,
+an adversary determines the pairs of nodes that ever communicate, a~priori,
 and requests to these pairs arrive indefinitely.
@@
 A~\emph{ground set} is a group of nodes that repeatedly communicate when an algorithm splits them (places at different clusters).
-Their role is to ensure that the algorithm maintains a partition of ground sets into clusters.
+Their role is to ensure that the algorithm maintains a~partition of ground sets into clusters.
 We emphasize that ground sets are revealed only upon their nodes are split.
@@
 	Fix the configuration just before the creation of a ground set with $x_i$ for some $i$.
-	We show that if at least $\ell$ isolated nodes exist (including $x_i$), then there exists a perfect partition of ground sets after expanding a ground set with $x_i$.
+	We show that if at least $\ell$ isolated nodes exist (including $x_i$), then there exists a~perfect partition of ground sets after expanding a ground set with $x_i$.
 	To show this claim, we consider a configuration, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
@@
 	Therefore $\{x_0, y_0\}$ is present on $I(x_i)$, and we must evict it to make space for $x_i$.
-	To find the cluster for $\{x_0, y_0\}$, observe that we have at most $\ell-1$ clusters with isolated nodes, and we have at least $\ell$ isolated nodes --- hence a cluster with two isolated nodes exists.
+	To find the cluster for $\{x_0, y_0\}$, observe that we have at most $\ell-1$ clusters with isolated nodes, and we have at least $\ell$ isolated nodes --- hence a~cluster with two isolated nodes exists.
 	To find a feasible partition, we swap $\{x_0, y_0\}$ with these two collocated isolated nodes, and then we swap an isolated node from $I(x_i)$ with $x_i$.
@@
 	Each expansion reduces the number of isolated nodes by $1$.
-	We start with $k \cdot \ell - (k-1)$ isolated nodes (a single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
+	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
 	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - l) / 2$.
@@
 We maintain all nodes of each component collocated in one cluster~\cite{repartition-disc}.
-If we need to move a node, we move the whole component that contains it.
+If we need to move a~node, we move the whole component that contains it.
 We maintain a balanced partition of our components, a similar concept to partition of integers into sets with equal sum \cite{integer-partitions-book}.
@@
 	A~cluster of type $C_3$ contributes at most $3 \alpha - 1$ to $\TAlg(\sigma_I)$, as two pairs of nodes from the component are saturated and contribute $\alpha$ each, and the third, unsaturated pair contributes at most $\alpha-1$.
-	Other cluster types contribute less: $C_1$ contributes $0$ and $C_2$ contributes $\alpha$.
+	Other cluster types contribute less: $C_1$ contributes~$0$ and $C_2$ contributes $\alpha$.
 	Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq (3 \alpha-1)\cdot \ell \leq 3\alpha\cdot\ell$.
@@
 we present \RM,
- an algorithm for the \OBRP restricted to clusters of capacity 2,
-i.e.~$k=2$.
+ an algorithm for the \OBRP restricted to clusters of capacity $k=2$.
 We interpret a pair of nodes collocated in one cluster as a matched pair.

### 2020-05-09 paper/disc20clustering.tex
On k=3

@@
 \begin{theorem}
-	\TAlg is $O(\ell)$-competitive for $k=3$.
+	\TAlg is $(54\ell)$-competitive for $k=3$.
 \end{theorem}
@@
 	Recall that $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
-	Faced with $W := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests $\sigma_E$ that it could not mitigate, $\OPT$ served some of them remotely and possibly performed some migrations to decrease its cost.
+	Faced with at least $W := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests $\sigma_E$ that it could not mitigate, $\OPT$ served some of them remotely and possibly performed some migrations to decrease its cost.
 
@@
 	By performing a single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$
-	Thus, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{2\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 2 - 2 \alpha \cdot \ell$.
-	From that, we obtain $|\sigma_E| \geq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
-	Let $E := \OPT(\sigma_E) + \OPT(M)$.
-
-	Finally, we bound the competitive ratio
-	\begin{equation*}
-		\frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + E/2} \leq \frac{27\alpha\cdot\ell + 2\cdot E}{\alpha + E} \leq 27 \ell = O(\ell).
-	\end{equation*}
+	Hence, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{2\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 2 - 2 \alpha \cdot \ell$.
+
+	Finally, we bound the competitive ratio.
+	From the bound on \OPT's cost we have $|\sigma_E| \geq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
+	Let $\xi := \OPT(\sigma_E) + \OPT(M)$.
+	The competitive ratio is
+
+	\[
+		\frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + \xi/2} \leq \frac{27\alpha\cdot\ell + 2\cdot \xi}{\alpha + \xi} \leq 27 \ell.
+	\]
 
@@
 	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
-	This way, the competitive ratio increases at most twofold.
+	This way, the competitive ratio increases at most twofold in comparison to a finished phase, and the competitive ratio is $\TAlg(\sigma)/\OPT(\sigma) \leq 54\ell$.
 \end{proof}

### 2020-05-09 paper/disc20clustering.tex
Consistency in section names

@@
 
-\subsection{Improved algorithm for Online Rematching} \label{sec:k2}
+\subsection{Improved Algorithm for Online Rematching} \label{sec:k2}
 In this section,

### 2020-05-09 paper/disc20clustering.tex
Comments k=2

@@
 
+	\maciek{Proposition 1: For $\lambda=\alpha$ the algorithm \RM is strictly 6-competitive.	Proposition 2: "The algorithm \RM is strictly 6-competitive for $\lambda=\alpha$.	"?}
+
+\maciek{First the intuitions, then the theorem, and right after the proof?}
+
+\maciek{The first sentence of the proof: first parenthesis does not explain things better, can be omitted.
+The second parenthesis also, I think: we use swap a lot all over the paper, and it should be understandable at the last proof.}
+
+\maciek{by collocating multiple pairs at once. "by collocating \textbf{two} pairs at once"}
+
+\maciek{"before they inflict too much cost" Unnecessary? Just collocating is enough.}
+
+\maciek{"Therefore it pays a cost of 1 migration per pair". I am confused by this sentence. Charging one is ok, but payment is always the same. Maybe it will become clear with the pic}
+
+\maciek{"a cost inflicted by a pair" the cost?}
+
+\maciek{"incurs the cost some time since" Maybe a different word than some? It is unclear if we refer to interval or a specific moment.}
+
 \begin{proof}[Proof of Theorem \ref{thm:k=2}]
@@
 \end{proof}
+\maciek{Fonts of OPT and \OPT differ. The latter is better.}
+
+\maciek{Excessive comma before ", is $\lambda +2\alpha$?}
+
+\maciek{m not introduced as the end of time in the proof, and also used in the intuitions. Consider some other symbol like $\tau$ or $t_F$}
+
+\maciek{"then \OPT pays $\lambda$ for serving" comma before then?}
+
+\maciek{", or $\alpha$ for collocating \textbf{and serving some requests remotely}" remove what's bold? The migration is sufficient}
+
+\maciek{"it can have one edge already collocated" at most one edge?}
+
+\maciek{"at the same time. However" One sentence? joined with "but"}
+
+\maciek{"similarly as in the former" "\textbf{similarly to the former}"}
+
+\maciek{"this means," "\textbf{this means that}"}
+
+\maciek{Suggestion: maybe call "its intervals" "\textbf{the saturating intervals}"?}
 

### 2020-05-09 paper/disc20clustering.tex
Lemma formulation
Intro to general model

@@
 can be arbitrary.
-In Section~\ref{sec:k3}, we show that the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental} provides an optimal algorithm for $k=3$.
-We showed a lower bound of $\Omega(k \cdot \ell)$ that holds for $k=3$ (cf. Section~\ref{sec:lowerbound}), hence the result from this section is asymptotically optimal.
+In Section~\ref{sec:k3}, we show a $O(k \cdot \ell)$-competitive algorithm for $k=3$ using the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental}.
+Prior to this section, we showed a lower bound of $\Omega(k \cdot \ell)$  that holds for the general model, hence the result from this section is asymptotically optimal (cf.
+Section~\ref{sec:lowerbound}).
 Furthermore, in Section~\ref{sec:k2}, we show a strictly $6$-competitive algorithm for $k=2$.
@@
 	\label{lem:1req}
-	In a single reconfiguration of nodes (after a merge of components), \TAlg exchanges at most $2$ pairs of nodes.
+	In a single reconfiguration of nodes (after a merge of components), \TAlg exchanges at most two pairs of nodes.
 \end{lemma}

### 2020-05-09 paper/disc20clustering.tex
Intro to LB

@@
 For $k=2$, the learning problem is trivial: immediate collocation of a communicating pair is $1$-competitive.
-(A partitioning problem for $k=2$ is non-trivial, a lower bound of $3$ is known \cite{repartition-disc}, and we provide a~$6$-competitive algorithm (see Section~\ref{sec:k2}).)
+(A partitioning problem for $k=2$ is non-trivial, a lower bound of $3$ is known \cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)
 

### 2020-05-09 paper/disc20clustering.tex
Reformulation in k=3

@@
 
-	To keep \OPT's cost low, we end extending ground sets prematurely --- while there are still $\ell+1$ isolated nodes (i.e., neglecting two last ground set expansions).
+	To keep \OPT's cost low, we terminate prematurely --- we stop extending ground sets while there are still $\ell+1$ isolated nodes (i.e., neglecting two last ground set expansions).
 	Consider \ALG's configuration after the last expansion.

### 2020-05-09 paper/disc20clustering.tex
Comments

@@
 
-\maciek{"at the same time. However" One sentence? joined with "but"}
-
 \maciek{"similarly as in the former" "\textbf{similarly to the former}"}
@@
 
+\maciek{Suggestion: "Note that \OPT cannot have both..." (till the end of paragraph) You told me this in one sentence during the call.
+Maybe use the same formulation?
+I feel like one sentence can be enough to explain this}
+
 \maciek{Suggestion: maybe call "its intervals" "\textbf{the saturating intervals}"?}
+
+\maciek{OPT's cost formula: make it one line by removing the expression with minimums.
+Move minimums to the paragraph text, by immediatelly expanding minimums (two steps in paragraph, instead of in the formula)}
 

### 2020-05-09 paper/disc20clustering.tex
comments
PPL fixes

@@
 
- \newcommand{\OPT}{\textsf{OPT}\xspace}
+ \newcommand{\OPT}{\textsc{OPT}\xspace}
   \newcommand{\OPTM}{\mathit{OPT}}
- \newcommand{\ALG}{\textsf{ALG}\xspace}
- \newcommand{\PPL}{\textsf{PPL}\xspace}
+ \newcommand{\ALG}{\textsc{ALG}\xspace}
+ \newcommand{\PPL}{\textsc{PPL}\xspace}
  \newcommand{\OBRP}{BRP\xspace}
  \newcommand{\PPOBRP}{PP-BRP}
- \newcommand{\dist}{\textsf{dist}}
+ \newcommand{\dist}{\mathit{dist}}
  \newcommand{\TAlg}{{\ensuremath{\textsf{ALG}_{3}}}\xspace}
-  \newcommand{\RM}{\textsf{RM}\xspace} % rematching alg
-
- \newcommand{\comm}{\textsc{comm}}
- \newcommand{\OFF}{\textsc{Off}\xspace}
+  \newcommand{\RM}{\textsc{RM}\xspace} % rematching alg
+
  \newcommand{\Rep}{\textsc{Rep}}
@@
 where the communication pattern is \emph{static}:
-an adversary determines the pairs of nodes that ever communicate, a~priori,
-and requests to these pairs arrive indefinitely.
+whether a pair of  nodes ever communicate or not, 
+is determined a priori and is unknown to algorithms,
+ and requests to such pairs arrive indefinitely.
 Any algorithm must eventually collocate  pairs of communicating nodes,
@@
 In our algorithms, a~\emph{component} is a subset of frequently communicating nodes.
-We maintain all nodes of each component collocated in one cluster~\cite{repartition-disc}.
-If we need to move a~node, we move the whole component that contains it.
-We maintain a balanced partition of our components, a similar concept to partition of integers into sets with equal sum \cite{integer-partitions-book}.
-In contrast, our partition is time-varying: communicating components are merged into one, and we adjust the partition.
-
-We say that a component is \emph{trivial}, if it contains exactly one node.
-We say that the node that belongs to trivial component \emph{isolated}.
-We say that a~configuration is \emph{component-respecting}
-if the nodes belonging to the same component are collocated.
+We maintain all nodes of a component always collocated in a same cluster~\cite{repartition-disc}.
+When we  move a~node, we move the whole component that contains it.
+We maintain a balanced partition of our components,
+a reminiscent of partitioning given integers into sets of equal sum \cite{integer-partitions-book}.
+In contrast, our partition is time-varying:  two components are merged into one component once they communicate, and we adjust the partition accordingly.
+
+We say that a component is \emph{trivial}, if it contains exactly one node,
+and the node in this component is \emph{isolated}.
+\mahmoud{we also use "singletone component". One is redundant.}
+We say that a~partition is \emph{component-respecting}
+if the nodes belonging to the same component are always collocated.
 
@@
 In DET, the partition was arbitrary.
-In the algorithm for the learning model, we choose the partition closest to the initial configuration.
-In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the previous configuration (a repartition of minimum cost).
-
-\maciek{The remainder of the paper used "configuration C" to denote the partition P. Easiest to adjust here.}
+In the algorithm for the learning model, we choose the partition closest to the initial partition.
+In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).
+
+\maciek{The remainder of the paper used "configuration C" to denote the partition P.
+Easiest to adjust here.
+\mahmoud{are "configuration" and "partition" the same thing everywhere?
+Then let's use "partition" only.}}
 
@@
 \textbf{Perfect Partition Learner algorithm.}
-Fix the initial configuration
-$P_I = I_1, \dots, I_{\ell}$ and \OPT's final configuration
-$P_F = F_1, \dots, F_{\ell}$\maciek{How do we define a configuration?}.
-The \emph{distance} of a configuration $P = C_1, \dots, C_{\ell}$ from the initial configuration \maciek{, denoted $\dist(P)$} is the number of nodes in $P$ that do not reside in their initial cluster.
-\maciek{We might denote distance $\Delta(P)$, and the current $\Delta$ might be R (like radius).}
-That is,
-$\dist(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$
-\maciek{Major: what does this definition mean? This is highly unclear, most likely because a configuration was not defined.}. 
+Fix the initial partition
+$P_I = I_1, \dots, I_{\ell}$ and \OPT's final partition
+$P_F = F_1, \dots, F_{\ell}$.
+The \emph{distance} of a partition $P = C_1, \dots, C_{\ell}$ from the initial partition,
+defined as 
+$\Delta(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$,
+ is the number of nodes in $P$ that do not reside in their initial cluster.
 In other words,
-at least $\dist(P, P_I)/2$ node swaps are required in order to reach the configuration $P$ from $P_I$, and thus
-$\OPT \geq \Delta:= \dist(P_F, P_I) $.
-With each repartitioning,
-\PPL moves to a configuration that minimizes the distance to the initial configuration $P_I$.
-As a result,
-\PPL never ends up \maciek{ends up is informal} in a configuration that is more than $\Delta$ (single-node) migrations away from $P_I$ \maciek{This is not an algorithm definition, that's its property.
-Should not be in this paragraph}.
-This invariant ensures that \PPL does not pay too much while recovering $P_F$ \maciek{Recovering $P_F$?
-It's not the goal.
-At least at this point the reader is confused: it does not know yet that PPL is proven to end up in $P_F$}.
-%      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
-We note \maciek{Define?} that the $\mathit{repartition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a~perfect partition closest to $P_I$.
-Hence it never moves to a configuration beyond distance $\Delta$ \maciek{This is again mixing the properties of ppl and its definition. Restructure}.
+at least $\Delta(P, P_I)$ node migrations are required in order to reach the partition $P$ from $P_I$, and thus
+$\OPT \geq \Delta^*:= \Delta(P_F, P_I) $.
+We  use $ \Delta(P_F) $ wherever the initial partition is clear from the context.
 The scheme of the algorithm can be found in the Algorithm \ref{alg:PPL}.
@@
 		%        \Ensure A final configuration $P_F$ 
-		\STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$}
-		\STATE{$P_0 := P_I$}
+		\STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$.}
+		\STATE{$P := P_I$}
 		\label{line:initcomponents}
 		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
-		\STATE Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
+		\STATE{Let $C_1 \ni u$ and $C_2 \ni v$ be the  components containing $u$ and $v$, respectively.}
 		\IF{$C_1 \neq C_2$}
-		\STATE {Unite the two components into a single component $C'$ and
-		$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$} \label{line:mergecomponents}
-		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
-		\COMMENT{i.e.~if not in the same cluster}    
-		}       
-		\STATE {$P_{t} = \mathit{repartition}(P_{t-1}, P_I, \mathcal{C})$} 
-		\COMMENT{move to $P$ closest to $P_I$}
+		\STATE {Unite $C_1$ and $C_2$ into one component $C'$ and
+		$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$.} \label{line:mergecomponents}
+%		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
+%		\COMMENT{i.e.~if not in the same cluster.}}       
+		\IF{$C_1$ and $C_2$ are not collcoated}
+		\STATE {Move to a partition closest to $P_I$, respecting all components in $\mathcal{C}$.} 
+%		\COMMENT{move to a partition closest to $P_I$}
 		\label{line:rebalance} 
@@
 
-\maciek{Algorithm pseudocode comments}
-\maciek{Stefan's comment: "container" components undefined}
-\maciek{Major thing to discuss: do we even need this?
-The algorithm is very simple, and the description in the paragraph would be much better.
-The only way to save it would be to have a general algorithm pseudocode (equivalent to DET), and parametrize it with the repartition procedure.
-Here, we would say that we choose the closest to initial, and it ALG3 we say that we choose the closest to current.}
-\maciek{This is not good that we need a comment about what $cluster \neq cluster$ means}
-\maciek{In pseudocode we need extremely short sentences. $C := singleton components ... $?; $Merge C, C'$ etc}
-\maciek{No need to maintain $P_i$'s. Just use one $P$? Or without naming: "move to partition (closeToInit(...))", and remove the partition args from cluster().}
-\maciek{Online algorithms should have different pseudocode. Instead of for loop, start with "When a request u,v arrives" (before that it is ok to have init phase just like it is now)}
+With each repartitioning,
+\PPL moves to a partition that minimizes the distance to the initial partition $P_I$.
+As a result,
+\PPL never moves to a~partition that is more than $\Delta^*$  node migrations away from $P_I$.
+This invariant latter ensures us that \PPL does not pay too much while recovering $P_F$.
+We emphasis that a repartitioning by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
+Meaning, it never moves to a partition beyond distance $\Delta^*$.
 
 \begin{property} \label{prop:dist<OPT}
-	Let $P$ be any configuration chosen by \PPL at Line $\ref{line:rebalance}$.
-	Then, $\dist(P,P_I) \leq \Delta$.
-	\maciek{Express the configuration without refering to pseudocode?}
+	Let $P$ be any configuration chosen by \PPL at any time.
+	Then, $\Delta(P,P_I) \leq \Delta^*$.	
 \end{property}
-\maciek{The property is explained in the previous paragraph.
-Here, reader thinks that this is obvious and wonders why.
-Restructure: Move closer to the exlaination text.
-Or even make this property a lemma (most likely better).}
 
 \begin{lemma}	\label{lemma:rebalancecost}
-	The cost of repartitioning at Line \ref{line:rebalance} is at most $2\cdot\OPT$.
-	\maciek{Express the reconfiguration without pseudocode?}
+	The cost of each repartitioning by \PPL is $2\cdot\OPT$.
 \end{lemma}
@@
 	Since \maciek{at least?} $|M^-|$ nodes are not in their original cluster before the repartitioning (i.e., in $P_{t-1}$),
-	the distance before the repartitioning is $\dist(P_{t-1},P_I) \geq | M^-|$.
+	the distance before the repartitioning is $\Delta(P_{t-1},P_I) \geq | M^-|$.
 	Analogously,
-	the distance afterwards is $\dist(P_{t},P_I) \geq | M^+|$.
+	the distance afterwards is $\Delta(P_{t},P_I) \geq | M^+|$.
 	Thus,
-	$|M| \leq \dist(P_{t-1},P_I) + \dist(P_{t},P_I)$.
+	$|M| \leq \Delta(P_{t-1},P_I) + \Delta(P_{t},P_I)$.
 	By Property \ref{prop:dist<OPT},
-	$\dist(P_{t-1},P_I) , \dist(P_{t},P_I) \leq \Delta \leq \OPT$
+	$\Delta(P_{t-1},P_I) , \Delta(P_{t},P_I) \leq \Delta^* \leq \OPT$
 	and thereby we have	
 	$|M| \leq 2\cdot\OPT$.
-	\maciek{the comma between $\dist$ is confusing. BOTH are bounded? If so, expand the sentence and write both individual bounds.}
+	\maciek{the comma between $\Delta$ is confusing. BOTH are bounded? If so, expand the sentence and write both individual bounds.}
 \end{proof}
@@
 	That is, 
-	once it reaches a configuration $P$ at distance $\Delta = \dist(P, P_I)$,
+	once it reaches a configuration $P$ at distance $\Delta^* = \Delta(P, P_I)$,
 	it does not move to a configuration
-	$P', \dist(P', P_I) > \Delta$,
-	before it enumerates all configurations at distance $\Delta$.
+	$P', \Delta(P', P_I) > \Delta^*$,
+	before it enumerates all configurations at distance $\Delta^*$.
 	Therefore,
-	\PPL eventually reaches $\Delta=\OPT$ and the configuration $P_F$.
+	\PPL eventually reaches $\Delta^*=\OPT$ and the configuration $P_F$.
 	%    including the request that completes revealing of all components that are collocated in $P_F$.
@@
 \begin{theorem} \label{thm:k=2}
-	The algorithm \RM, for $\lambda=\alpha$, is (strictly) 6-competitive.
+	For $\lambda=\alpha$, the algorithm \RM  is strictly 6-competitive.
 \end{theorem}
 
-We charge both \OPT and \RM whenever \RM collocates a pair (i.e., on a collocation event).
-%More precisely,we charge them only the cost of the collocation, per such event.
-%Hence,we charge the cost of  this collocation to \OPT for the first time.
-\RM collocates a pair always with a swap (i.e., two simultaneous migrations),
+We charge both \OPT and \RM whenever \RM collocates a pair.%(i.e., on a collocation event).
+\RM collocates a pair always with a swap,%(i.e., two simultaneous migrations),
 which  costs $2\alpha$,
-while OPT may save some costs by collocating multiple pairs at once, 
- before they inflict too much communication cost.
+while OPT may save some costs by collocating multiple pairs at once.
+% before they inflict too much communication cost.
  Therefore it pays the price of only one migration per pair  (\mahmoud{see Figure \ref{fig:TBD}}).
@@
 we charge \OPT a cost inflicted by a pair,
-if and only if  \OPT incurs the cost some time since the last time \RM separated the pair.
+if and only if  \OPT incurs that cost after the latest separation of the pair by \RM.
 %%
-
-	\maciek{Proposition 1: For $\lambda=\alpha$ the algorithm \RM is strictly 6-competitive.	Proposition 2: "The algorithm \RM is strictly 6-competitive for $\lambda=\alpha$.	"?}
-
-\maciek{First the intuitions, then the theorem, and right after the proof?}
-
-\maciek{The first sentence of the proof: first parenthesis does not explain things better, can be omitted.
-The second parenthesis also, I think: we use swap a lot all over the paper, and it should be understandable at the last proof.}
-
-\maciek{by collocating multiple pairs at once. "by collocating \textbf{two} pairs at once"}
-
-\maciek{"before they inflict too much cost" Unnecessary? Just collocating is enough.}
-
-\maciek{"Therefore it pays a cost of 1 migration per pair". I am confused by this sentence. Charging one is ok, but payment is always the same. Maybe it will become clear with the pic}
-
-\maciek{"a cost inflicted by a pair" the cost?}
-
-\maciek{"incurs the cost some time since" Maybe a different word than some? It is unclear if we refer to interval or a specific moment.}
 
@@
 
+\noindent
+\textbf{Approximation for offline rematching.}
+\maciek{Maybe mention the offline approx here?}
+
 \section{Discussion and Future Work}

### 2020-05-09 paper/disc20clustering.tex
isolated double definition fix
PPL proof

@@
 Fix the initial partition
-$P_I = I_1, \dots, I_{\ell}$ and \OPT's final partition
-$P_F = F_1, \dots, F_{\ell}$.
-The \emph{distance} of a partition $P = C_1, \dots, C_{\ell}$ from the initial partition,
+$P_I :=\{ I_1, \dots, I_{\ell}\}$ and \OPT's final partition
+$P_F := \{F_1, \dots, F_{\ell}\}$.
+The \emph{distance} of a partition
+ $P = \{C_1, \dots, C_{\ell}\}$ from the initial partition,
 defined as 
@@
 \begin{property} \label{prop:dist<OPT}
-	Let $P$ be any configuration chosen by \PPL at any time.
-	Then, $\Delta(P,P_I) \leq \Delta^*$.	
+	Let $P$ be any partition chosen by \PPL at any time.
+	Then, $\Delta(P) \leq \Delta^*$.	
 \end{property}
@@
 \begin{proof}
-	Consider the repartitioning that transforms $P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
+	Let $P_i$ denote the partition of \PPL immediately after serving $\sigma_{i}$.
+	Consider the repartitioning that transforms 
+	$P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
 	Let $M \subset V$ denote the set of nodes that migrate during this process.
-	Let $M^-$ and $M^+$ denote the subset of nodes that (respectively)
-	enter or leave their original cluster during the repartitioning.    
+	Let $M^-$ and $M^+$ denote the subset of nodes that, respectively,
+	enter or leave their initial cluster during the repartitioning.    
 	Then,
 	$M = M^+ \cup M^-$.
-	Since \maciek{at least?} $|M^-|$ nodes are not in their original cluster before the repartitioning (i.e., in $P_{t-1}$),
-	the distance before the repartitioning is $\Delta(P_{t-1},P_I) \geq | M^-|$.
+	Since at least $|M^-|$ nodes are not in their initial cluster before the repartitioning (i.e., in $P_{t-1}$),
+	the distance before the repartitioning is $\Delta(P_{t-1}) \geq | M^-|$.
 	Analogously,
-	the distance afterwards is $\Delta(P_{t},P_I) \geq | M^+|$.
+	the distance afterwards is $\Delta(P_{t}) \geq | M^+|$.
 	Thus,
-	$|M| \leq \Delta(P_{t-1},P_I) + \Delta(P_{t},P_I)$.
+	$|M| \leq \Delta(P_{t-1}) + \Delta(P_{t})$.
 	By Property \ref{prop:dist<OPT},
-	$\Delta(P_{t-1},P_I) , \Delta(P_{t},P_I) \leq \Delta^* \leq \OPT$
-	and thereby we have	
-	$|M| \leq 2\cdot\OPT$.
-	\maciek{the comma between $\Delta$ is confusing. BOTH are bounded? If so, expand the sentence and write both individual bounds.}
+	$\Delta(P_{t-1})  \leq \Delta^*$ and
+	$\Delta(P_{t}) \leq \Delta^*$.
+	Since $\Delta^* \leq \OPT$,
+	we obtain	 $|M| \leq 2\cdot\OPT$.
 \end{proof}
@@
 \begin{theorem}	\label{thm:upperbound}
-	\PPL reaches the final configuration $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
-	\maciek{Do not hide that we have $2(k-1)\ell$ in the theorem. This is a better bound.}
+	\PPL reaches the final partition $P_F$
+	 and it is $(2\cdot (k-1)\cdot\ell)$-competitive.
 \end{theorem}
@@
 	the algorithm enumerates all $\ell$-way partitions of components
-	that are in the same (closest) distance of $P_I$.
+	that are in the same (closest) distance to $P_I$.
 	That is, 
-	once it reaches a configuration $P$ at distance $\Delta^* = \Delta(P, P_I)$,
-	it does not move to a configuration
-	$P', \Delta(P', P_I) > \Delta^*$,
-	before it enumerates all configurations at distance $\Delta^*$.
+	once it reaches a partition $P$ at distance $\Delta^* = \Delta(P)$,
+	it does not move to a partition
+	$P', \Delta(P') > \Delta^*$,
+	before it enumerates all partitions at distance $\Delta^*$.
 	Therefore,
-	\PPL eventually reaches $\Delta^*=\OPT$ and the configuration $P_F$.
-	%    including the request that completes revealing of all components that are collocated in $P_F$.
-	There are at most $(k-1)\cdot\ell < k\cdot\ell $ calls   to $\mathit{repartition}$
-	(i.e., the number of internal edges in $P_F$).
-	\maciek{Clarify why the number of internal edges bound the number of repartitions.}
+	\PPL eventually reaches  the partition $P_F$ at distance $\Delta^*=\OPT$.
+	With each distinct request, 
+	the size of some component increases by one.
+	For any cluster $F_i \in P_F$,
+	we have $\sum_{C \in F_i} |C| = k$.
+	A component $C \in F_i$,
+	 initially begins as an isolated node and it grows by gaining $|C|-1$ more nodes.
+	Hence, the total number of times a component in $F_i$ grows is 
+	$\sum_{C \in F_i} (|C|-1) \leq k-1$.
+	Therefore, there are at most  $(k-1)\cdot\ell $ distinct requests
+	for which \PPL performs a repartitioning.
+	I.e.,
+	 \PPL performs at most $(k-1)\cdot\ell $ repartitionings.
 	By Lemma \ref{lemma:rebalancecost},
-	each repartition costs at most $2\cdot\OPT$.
-	The total cost is therefore at most $2\cdot\OPT\cdot k\cdot\ell$, which implies the competitive ratio.
-	%    \mahmoud{This is the cost of moving and the cost of remote comm. is not counted.
-	%    	So it is 4-competitive (?)}
+	each repartitioning costs at most $2\cdot\OPT$.
+	The total cost is therefore at most $2\cdot\OPT\cdot (k-1) \cdot\ell$, which implies the competitive ratio.
 \end{proof}

### 2020-05-09 paper/disc20clustering.tex
a same to the same

@@
 	Throughout the construction, we maintain ground sets for each node.
-	Initially all nodes are \emph{isolated}.
+	Initially all nodes are isolated (i.e., in ground sets of size $1$).
 	First, we create a ground set $B$ of $k-1$ nodes on an arbitrarily chosen cluster (the nodes  are already collocated).
@@
 In our algorithms, a~\emph{component} is a subset of frequently communicating nodes.
-We maintain all nodes of a component always collocated in a same cluster~\cite{repartition-disc}.
+We maintain all nodes of a component always collocated in the same cluster~\cite{repartition-disc}.
 When we  move a~node, we move the whole component that contains it.
@@
 \textbf{The charging scheme.}
-Consider two  pairs that share a same node, 
+Consider two  pairs that share the same node, 
 i.e.~\emph{intersecting pairs},
@@
 Otherwise,
- \OPT is possibly charged for a same cost repeatedly.
+ \OPT is possibly charged for the same cost repeatedly.
 For this reason,

### 2020-05-10 paper/disc20clustering.tex
PPL to appendix
Notation consistency

@@
 and the node in this component is \emph{isolated}.
-\mahmoud{we also use "singletone component". One is redundant.}
 We say that a~partition is \emph{component-respecting}
@@
 We  use $ \Delta(P_F) $ wherever the initial partition is clear from the context.
-The scheme of the algorithm can be found in the Algorithm \ref{alg:PPL}.
-
-\label{alg:PPL}
-\begin{algorithm}
-	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-	\renewcommand{\algorithmicensure}{\textbf{Output:}}
-	\begin{algorithmic}
-		%        \Require 
-		%        $k, \ell$,
-		%        initial configuration $P_I$,
-		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
-		%        \Ensure A final configuration $P_F$ 
-		\STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$.}
-		\STATE{$P := P_I$}
-		\label{line:initcomponents}
-		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
-		\STATE{Let $C_1 \ni u$ and $C_2 \ni v$ be the  components containing $u$ and $v$, respectively.}
-		\IF{$C_1 \neq C_2$}
-		\STATE {Unite $C_1$ and $C_2$ into one component $C'$ and
-		$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$.} \label{line:mergecomponents}
-%		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
-%		\COMMENT{i.e.~if not in the same cluster.}}       
-		\IF{$C_1$ and $C_2$ are not collcoated}
-		\STATE {Move to a partition closest to $P_I$, respecting all components in $\mathcal{C}$.} 
-%		\COMMENT{move to a partition closest to $P_I$}
-		\label{line:rebalance} 
-		\ENDIF
-		\ENDIF
-		\ENDFOR
-	\end{algorithmic}
-	\caption{Perfect Partition Learner (\PPL)}
-	\label{alg:ppl}
-\end{algorithm}
+The scheme of the algorithm can be found in the  appendix (Algorithm \ref{alg:ppl}).
 
@@
 The algorithm partitions nodes into components, and
-initially, each node belongs to its own singleton component.
+initially, each node is isolated (belongs to its own component).
 For each pair of nodes $\set{x,y}$, \TAlg maintains a counter $C_{\set{x,y}}$ and increments it on every external request between $x$ and $y$.
 Once $C_{\set{x,y}} = \alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
-If no such partitioning exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
+If no such partitioning exists, \TAlg resets all components to trivial components, resets all counters to $0$, and ends the phase.
 
@@
 In particular, we have so far focused on deterministic algorithms, and it would be interesting to study the power of randomization in this context.
-On the practical side, it would also be interesting to study our algorithms empirically, under realistic workloads.
+On the practical side, it would also be interesting to study our algorithms empirically, under realistic workloads.h
 
@@
 \bibliography{references}
+
 \appendix
 
+\section{Pseudocode of Perfect Partition Learner}
+
+\begin{algorithm}
+	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+	\renewcommand{\algorithmicensure}{\textbf{Output:}}
+	\begin{algorithmic}
+		%        \Require 
+		%        $k, \ell$,
+		%        initial configuration $P_I$,
+		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
+		%        \Ensure A final configuration $P_F$ 
+		\STATE {For each node $v$ create a component $C_v = \{ v \}$ and add it to $\mathcal{C}$.}
+		%\STATE{$P := P_I$}
+		\label{line:initcomponents}
+		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
+		\STATE{Let $C_1 \ni u$ and $C_2 \ni v$ be the  components containing $u$ and $v$, respectively.}
+		\IF{$C_1 \neq C_2$}
+		\STATE {Merge $C_1$ and $C_2$ into one component $C'$ and
+			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$.} \label{line:mergecomponents}
+		%		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
+		%		\COMMENT{i.e.~if not in the same cluster.}}       
+		\IF{$C_1$ and $C_2$ are not collocated}
+		\STATE {Move to a partition closest to $P_I$, respecting all components in $\mathcal{C}$.} 
+		%		\COMMENT{move to a partition closest to $P_I$}
+		\label{line:rebalance} 
+		\ENDIF
+		\ENDIF
+		\ENDFOR
+	\end{algorithmic}
+	\caption{Perfect Partition Learner (\PPL)}
+	\label{alg:ppl}
+\end{algorithm}
+
 \end{document}

### 2020-05-10 paper/disc20clustering.tex
Self-review

@@
 		\rowcolor{gray!25}
-		Learning model & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell)$\hspace{0.3cm} (\S \ref{sec:ppl}) \\
+		Learning model & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell)$\hspace{0.1cm} (\S \ref{sec:ppl}) \\
 	\end{tabular}
@@
 \begin{theorem}
-	\TAlg is $(54\ell)$-competitive for $k=3$.
+	\TAlg is $(60\ell)$-competitive for $k=3$.
 \end{theorem}
@@
 	Finally, we bound the competitive ratio.
-	From the bound on \OPT's cost we have $|\sigma_E| \geq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
+	From the bound on \OPT's cost we have $|\sigma_E| \leq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
 	Let $\xi := \OPT(\sigma_E) + \OPT(M)$.
@@
 	\[
-		\frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + \xi/2} \leq \frac{27\alpha\cdot\ell + 2\cdot \xi}{\alpha + \xi} \leq 27 \ell.
+		\frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + \xi/2} \leq \frac{30\alpha\cdot\ell + 4\cdot \xi}{\alpha + \xi} \leq 30 \ell.
 	\]
@@
 	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
-	This way, the competitive ratio increases at most twofold in comparison to a finished phase, and the competitive ratio is $\TAlg(\sigma)/\OPT(\sigma) \leq 54\ell$.
+	This way, the competitive ratio increases at most twofold in comparison to a finished phase, and the competitive ratio is $\TAlg(\sigma)/\OPT(\sigma) \leq 60\ell$.
 \end{proof}

### 2020-05-10 paper/disc20clustering.tex
Comments

@@
 In the algorithm for the learning model, we choose the partition closest to the initial partition.
+\maciek{component respecting?}
 In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).
@@
 \mahmoud{are "configuration" and "partition" the same thing everywhere?
-Then let's use "partition" only.}}
+Then let's use "partition" only.\maciek{I thought partition is on component level, and configuration on node level (todo check where to add component-respecting to the partition in this chapter).
+Maybe say in the intro in this chapter that configuration = partition?
+We use both heavily.}}}
 
@@
 With each repartitioning,
-\PPL moves to a partition that minimizes the distance to the initial partition $P_I$.
+\PPL moves to a \maciek{component-respecting} partition that minimizes the distance to the initial partition $P_I$.
 As a result,
@@
 	On each inter-cluster request,
-	the algorithm enumerates all $\ell$-way partitions of components
+	the algorithm enumerates all \maciek{component-respecting} $\ell$-way partitions of components
 	that are in the same (closest) distance to $P_I$.
@@
 	I.e.,
-	 \PPL performs at most $(k-1)\cdot\ell $ repartitionings.
+	 \PPL performs at most $(k-1)\cdot\ell $ repartitions.
 	By Lemma \ref{lemma:rebalancecost},

### 2020-05-10 paper/disc20clustering.tex
Comments
Self-review

@@
 	\label{th:lowerbound}
-	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least $(k\cdot \ell - k - \ell)/2$ for any $k\geq 3$ and $\ell \geq 2$.
+	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least $(k\cdot \ell - k - 2\ell)/2$ for any $k\geq 3$ and $\ell \geq 2$.
 \end{theorem}
@@
 	\ALG performs at least one swap for every expansion of a ground set.
-	Each expansion reduces the number of isolated nodes by $1$.
+	Each expansion reduces the number of isolated nodes by $1$ or $2$.
+	The decrease of two isolated nodes occurs when the largest component on the initial cluster of $x_i$ is still a trivial component; this can occur at most $\ell$ times.
 	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
-	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - l) / 2$.
+	Hence, the total number of swaps of \ALG is at least $k\cdot \ell - k - 2 \ell$.
+	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - 2\ell) / 2$.
 \end{proof}
@@
 		\IF{$C_1$ and $C_2$ are not collocated}
-		\STATE {Move to a partition closest to $P_I$, respecting all components in $\mathcal{C}$.} 
+		\STATE {Move to a partition closest to $P_I$ and respecting all components in $\mathcal{C}$.} 
 		%		\COMMENT{move to a partition closest to $P_I$}

### 2020-05-10 paper/disc20clustering.tex
Self-review

@@
 	Each expansion reduces the number of isolated nodes by $1$ or $2$.
-	The decrease of two isolated nodes occurs when the largest component on the initial cluster of $x_i$ is still a trivial component; this can occur at most $\ell$ times.
+	The decrease of two isolated nodes occurs when the largest component on  $I(x_i)$ is still trivial; this occurs at most $\ell$ times.
 	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
@@
 	Let $\TAlg(M)$ be the cost of migrations performed by \TAlg in this phase.
-	\TAlg performs at most $2 \ell$ component merge operations, as
-	exceeding this number means that a component of size $4$ exists, and the phase should have ended already.
-	Combining this with Lemma~\ref{lem:1req} gives us $\TAlg(M) \leq 8\alpha\cdot\ell$ (recall that an exchange costs $2\alpha$).
+	\TAlg performs at most $2 \ell$ component merge operations ---
+	exceeding this number would mean that a~component of size $4$ exists, and the phase should have ended already.
+	We bound the cost of a each repartition by Lemma~\ref{lem:1req}, to obtain $\TAlg(M) \leq 8\alpha\cdot\ell$.
 	%Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
@@
 	Other cluster types contribute less: $C_1$ contributes~$0$ and $C_2$ contributes $\alpha$.
-	Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq (3 \alpha-1)\cdot \ell \leq 3\alpha\cdot\ell$.
+	Summing over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq (3 \alpha-1)\cdot \ell \leq 3\alpha\cdot\ell$.
 

### 2020-05-10 paper/disc20clustering.tex
Typos
Big grammar check of all paper; several issues fixed

@@
 	By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), we can reduce the network load and  improve application performance. 
-	However, the communication pattern of different application is often unknown a priori, and may change over time, hence it needs to be learned in an online manner.
+	However, the communication pattern of different applications is often unknown a priori and may change over time, hence it needs to be learned in an online manner.
 	%
@@
 	(introduced by Avin et al.~at DISC 2016)
-	which asks for an algorithm that strikes
+	that asks for an algorithm that strikes
 	an optimal tradeoff between the benefits
@@
 	We improve the result for $k=2$ by providing a strictly $6$-competitive upper bound for the general model.
-	In contrast to most prior work, our algorithms respect all capacity constraints, and do not require resource augmentation.
+	In contrast to most prior work, our algorithms respect all capacity constraints and do not require resource augmentation.
 
@@
 The popularity of data-centric, distributed applications has led to an explosive growth of network traffic, especially in data centers \cite{roy2015inside,singh2015jupiter}.
-As the performance of these distributed applications often critically depends on the underlying network \cite{mogul2012we}, an efficient operation of these networks is important.
+The performance of these distributed applications often critically depends on the underlying network \cite{mogul2012we}, and efficient operation of these networks is important.
 At the same time, distributed systems are often highly virtualized today, and provide interesting new opportunities for resource optimization.
@@
 it may move (``migrate'') some nodes into clusters different from their current clusters, while respecting the capacity of every cluster. 
-Afterwards, 
+Afterward, 
 the algorithm serves the  request.
@@
 (on the learning model) \cite{sigmetrics19_partitioning}.
-The focus of these papers however is primarily on models with resource augmentation: the online algorithm can use slightly larger clusters than the offline algorithm.  
+However, the focus of these papers is primarily on models with resource augmentation: the online algorithm can use slightly larger clusters than the offline algorithm.  
 Avin et al.
@@
 
-In this section we consider the learning variant of Dynamic Balanced Graph Partitioning.
+In this section, we consider the learning variant of Dynamic Balanced Graph Partitioning.
 For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$.
-The lower bound can be easily translated to the partitioning model (studied in Section~\ref{sec:part}) with finite request sequence.
+The lower bound can be easily translated to the partitioning model (studied in Section~\ref{sec:part}) with a~finite request sequence.
 At the end of this section, we discuss an asymptotically optimal upper bound.
@@
 We emphasize that ground sets are revealed only upon their nodes are split.
-We say that a ground set is \emph{trivial}, if it contains exactly one node.
+We say that a ground set is \emph{trivial} if it contains exactly one node.
 We say that the node that belongs to trivial a ground set is \emph{isolated}.
 
-The adversary constructs ground sets depending on choices of a deterministic online algorithm.
+The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
 We emphasize that the algorithm does not know ground sets before they are revealed.
@@
 	Throughout the construction, we maintain ground sets for each node.
-	Initially all nodes are isolated (i.e., in ground sets of size $1$).
+	Initially, all nodes are isolated (i.e., in ground sets of size $1$).
 	First, we create a ground set $B$ of $k-1$ nodes on an arbitrarily chosen cluster (the nodes  are already collocated).
 	%We never dismiss any constructed ground set.
-	Each cluster hosts exactly $k$ nodes, and in any configuration a single isolated node resides in $S$.
+	Each cluster hosts exactly $k$ nodes, and in any configuration, a single isolated node resides in $S$.
 	We say that this isolated node that accompanies the ground set $B$ is \emph{special}.
@@
 
-	We claim that expanding ground sets in this way can continue as long as at least $\ell$ isolated nodes exists.
+	We claim that expanding ground sets in this way can continue as long as at least $\ell$ isolated nodes exist.
 	Fix the configuration just before the creation of a ground set with $x_i$ for some $i$.
@@
 	If $I(x_i)$ contains an isolated node, then we simply obtain a~perfect partition by swapping it with $x_i$.
-	In the following we assume that no isolated nodes exist on $I(x_i)$.
+	In the following, we assume that no isolated nodes exist on $I(x_i)$.
 	The largest ground set of $I(x_i)$ is at most $k-1$, because $x_i$ is on $S$.
@@
 Every lower bound for the learning problem constitutes a converging lower bound for the partitioning problem as the input sequence grows to infinity.
-However, our construction allows for an exact lower bound for partitioning with finite input sequence.
+However, our construction allows for an exact lower bound for partitioning with a~finite input sequence.
 To this end, we issue only enough requests until the algorithm collocates the nodes of each ground set (if it never does, we terminate after it pays $k\cdot \ell$).
@@
 
-Finally, we note the possibility for improvement. The algorithm CREP~\cite{repartition-disc} requires $(2+\epsilon)$-augmentation to guarantee the competitive ratio independent of $\ell$.
+Finally, we note the possibility of improvement. The algorithm CREP~\cite{repartition-disc} requires $(2+\epsilon)$-augmentation to guarantee the competitive ratio independent of $\ell$.
 
@@
 
-We say that a component is \emph{trivial}, if it contains exactly one node,
+We say that a component is \emph{trivial} if it contains exactly one node,
 and the node in this component is \emph{isolated}.
@@
 
-The algorithm in this section, and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from \cite{repartition-disc}.
-The difference is in the the choice of partition after a component merge.
+The algorithm in this section and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from \cite{repartition-disc}.
+The difference is in the choice of partition after a component merge.
 In DET, the partition was arbitrary.
@@
 This invariant latter ensures us that \PPL does not pay too much while recovering $P_F$.
-We emphasis that a repartitioning by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
-Meaning, it never moves to a partition beyond distance $\Delta^*$.
+We emphasize that a repartitioning by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
+Meaning, it never moves to a partition beyond the distance $\Delta^*$.
 
@@
 	Analogously,
-	the distance afterwards is $\Delta(P_{t}) \geq | M^+|$.
+	the distance afterward is $\Delta(P_{t}) \geq | M^+|$.
 	Thus,
@@
 can be arbitrary.
-In Section~\ref{sec:k3}, we show a $O(k \cdot \ell)$-competitive algorithm for $k=3$ using the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental}.
+In Section~\ref{sec:k3}, we show an $O(k \cdot \ell)$-competitive algorithm for $k=3$ using the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental}.
 Prior to this section, we showed a lower bound of $\Omega(k \cdot \ell)$  that holds for the general model, hence the result from this section is asymptotically optimal (cf.
@@
 	In this case, a local repartition is possible (within $U$ and $V$), for a cost of at most $2$ swaps.
-	Note that it is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, a components of size $3$ participates in a merge.
-	After merge, we would have a component of size at least $4$, and in this case no component respecting partitioning exists, a contradiction.
+	Note that it is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, a component of size $3$ participates in a merge.
+	After the merge, we would have a component of size at least $4$, and in this case, no component respecting partitioning exists, a contradiction.
 
@@
 	Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
-In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
+In this case, \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes~of~$W$.
 \end{proof}
@@
 	exceeding this number would mean that a~component of size $4$ exists, and the phase should have ended already.
-	We bound the cost of a each repartition by Lemma~\ref{lem:1req}, to obtain $\TAlg(M) \leq 8\alpha\cdot\ell$.
+	We bound the cost of each repartition by Lemma~\ref{lem:1req}, to obtain $\TAlg(M) \leq 8\alpha\cdot\ell$.
 	%Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
@@
 	By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
-	This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
+	This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterward.
 	As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
@@
 
-While are bounds are tight, there are several interesting avenues for future research.
+While our bounds are tight, there are several interesting avenues for future research.
 In particular, we have so far focused on deterministic algorithms, and it would be interesting to study the power of randomization in this context.
-On the practical side, it would also be interesting to study our algorithms empirically, under realistic workloads.h
+On the practical side, it would also be interesting to study our algorithms empirically, under realistic workloads.
 

### 2020-05-10 paper/disc20clustering.tex
OPT undefined
Comments

@@
 
+\maciek{Above definition is the definition of strict competitive ratio. In comparison to non-strict with additive constant. Emphasize? Maybe already use the notion of \OPT}
+
 \subsection{Related work}
@@
 Table \ref{tab:overview} provides an overview of our contributions compared to prior work.
+
+\maciek{emphasize that all our algorithms have strict competitive ratio}
 
@@
 Repeating such requests, almost all the nodes must visit the cluster $S$.
-In comparison, \OPT performs only two migrations.
+In comparison, we show that there exists an optimal offline algorithm \OPT that performs only two migrations.
 

### 2020-05-10 paper/disc20clustering.tex
Commentsg

@@
 the problem is an online variant of the classic matching problem where
-a matched pair can separate from each other in order to ``rematch'' with two other nodes.
+a matched pair can separate from each other in order to ``rematch'' with two other nodes
+\maciek{Maybe we use SIDMA's connection to matching instead of this local view.
+SIDMA: "This can be viewed as an online maximal (re)matching problem: clusters of size two contain (“match”) exactly one pair of nodes, and maximizing pairwise communication within each cluster is equivalent to minimizing inter-cluster communication."}.
 This is known as the  \emph{Online Rematching} 

### 2020-05-10 paper/disc20clustering.tex
wip
Improvement of lem:1req, now regular proof instead of by contradiction

@@
 \textbf{Component-based algorithm.}
-Now we describe the algorithm \TAlg.
+We introduce the algorithm \TAlg that we analyze in this section.
 The algorithm partitions nodes into components, and
@@
 \begin{proof}
-	Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
-	Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.
+	If no component respecting partition exists after the merge of components, \TAlg simply ends the phase and performs no reconfiguration.
+	It suffices to show that the merged component has size at least $4$ to conclude that \TAlg incurs no cost.
 
@@
 	In this case, a local repartition is possible (within $U$ and $V$), for a cost of at most $2$ swaps.
-	Note that it is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, a component of size $3$ participates in a merge.
-	After the merge, we would have a component of size at least $4$, and in this case, no component respecting partitioning exists, a contradiction.
+	If either $U$ or $V$ is of type $C_3$, a component of size $3$ participates in a merge, and we have a component of size at least $4$, and \TAlg ends the phase with no reconfiguration.
 
 	It remains to consider the case where both $U$ and $V$ are of type $C_2$.
-Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists.
-	Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
+	If $(u,v)$ both belong to components of size $2$, then the merged component has size $4$, and \TAlg incurs no cost. 
+	Otherwise, if one of $u,v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
+
 	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
-	Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists.
-In this case, \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes~of~$W$.
+	Note that if $C_1$-type cluster does not exist, then no component respecting partitioning exists.
+	Otherwise, \TAlg performs one swap --- it exchanges the nodes $u$ and $v$ with any two nodes~of~any cluster of type $C_1$.
+
+	In each case, we showed that a partition is reachable in at most two swaps.
 \end{proof}

### 2020-05-10 paper/disc20clustering.tex
Better explaination (we have a ton of space left)
Expanding and rewriting the proof of k=3

@@
 	By $\sigma$ we denote the input sequence that arrived during the phase.
-	By $\sigma_{cost}$ we denote the requests that at the moment of arrival were external requests for \TAlg (these are the only requests that incur cost for \TAlg).
-	In our analysis, we partition $\sigma_{cost}$ into subsequences $\sigma_I$ and $\sigma_E$.
-	The sequence $\sigma_I$ (inter-component requests) are the requests from $\sigma_{cost}$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
+	In our analysis, we focus on the requests that were external to \TAlg at the moment of their arrival; these are the only requests that incur cost for \TAlg.
+	We denote these external requests by $\sigma_{cost}$.
+	We partition the sequence $\sigma_{cost}$ into subsequences $\sigma_I$ and $\sigma_E$.
+	The sequence $\sigma_I$ (inter-component requests) denotes the requests from $\sigma_{cost}$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
 	The sequence $\sigma_E$ (extra-component requests) denotes the requests from $\sigma_{cost}$ that do not appear in $\sigma_I$.
@@
 	%Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
-	Let $\TAlg(M)$ be the cost of migrations performed by \TAlg in this phase.
-	\TAlg performs at most $2 \ell$ component merge operations ---
+	Let $\TAlg(M)$ denote the cost of migrations performed by \TAlg in this phase.
+	During the phase, \TAlg performs at most $2 \ell$ component merge operations ---
 	exceeding this number would mean that a~component of size $4$ exists, and the phase should have ended already.
-	We bound the cost of each repartition by Lemma~\ref{lem:1req}, to obtain $\TAlg(M) \leq 8\alpha\cdot\ell$.
+	We bound the cost of each reconfiguration after a merge by Lemma~\ref{lem:1req}, obtaining $\TAlg(M) \leq 8\alpha\cdot\ell$.
 	%Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
@@
 	Now we bound $\TAlg(\sigma_I)$.
-	A~cluster of type $C_3$ contributes at most $3 \alpha - 1$ to $\TAlg(\sigma_I)$, as two pairs of nodes from the component are saturated and contribute $\alpha$ each, and the third, unsaturated pair contributes at most $\alpha-1$.
+	A~cluster of type $C_3$ contributes at most $3 \alpha - 1$ to $\TAlg(\sigma_I)$, as two pairs of nodes from the component are saturated and incurred the cost $\alpha$ each, and the third, unsaturated pair incurs the cost at most $\alpha-1$.
 	Other cluster types contribute less: $C_1$ contributes~$0$ and $C_2$ contributes $\alpha$.
@@
 
-	We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
+	The cost of \OPT is lower-bounded by the cost of serving $\sigma_I$ and the cost of serving $\sigma_E$.
 	While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts: we separately bound $\OPT$ by $\OPT(\sigma_I) + \OPT(M)$ and $\OPT(\sigma_E) + \OPT(M)$.
-	Combining those bounds gives us $\OPT \geq \max\{\OPT(\sigma_I) + \OPT(M), \OPT(\sigma_E) + \OPT(M)\} \geq (\OPT(\sigma_I) + \OPT(M)) / 2 + (\OPT(\sigma_E) + \OPT(M)) / 2$.
+	Combining those bounds and using the relation between the maximum and the average gives us
+	\begin{align*}
+		\OPT&\geq \max\{\OPT(\sigma_I) + \OPT(M), \OPT(\sigma_E) + \OPT(M)\}\\
+		&\geq (\OPT(\sigma_I) + \OPT(M)) / 2 + (\OPT(\sigma_E) + \OPT(M)) / 2.
+	\end{align*}
 
@@
 	Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
-	By entering some fixed configuration before serving $\sigma_E$, \OPT may mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
-	Recall that $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
-	Faced with at least $W := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests $\sigma_E$ that it could not mitigate, $\OPT$ served some of them remotely and possibly performed some migrations to decrease its cost.
-
-	Now we estimate the cost of $\OPT(\sigma_E)$ while accounting savings from migrations.
-	By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
+	Fix a configuration of \OPT prior serving the sequence $\sigma_E$.
+	Even without migrations, \OPT may mitigate paying for requests between at most $3\ell$ pairs of nodes that can be collocated in its clusters.
+	Recall that $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), and the counter of an unsaturated pair is at most $\alpha - 1$.
+	Hence, $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
+	Finally, during the phase, \OPT is faced with at least $\chi := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests from $\sigma_E$ that it cannot mitigate.
+
+	By migrating nodes, \OPT may decrease its cost, and pay less than $\chi$ for requests from $\sigma_E$.
+	By swapping a pair of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
 	This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterward.
@@
 	By performing a single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$
-	Hence, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{2\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 2 - 2 \alpha \cdot \ell$.
-
-	Finally, we bound the competitive ratio.
-	From the bound on \OPT's cost we have $|\sigma_E| \leq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
-	Let $\xi := \OPT(\sigma_E) + \OPT(M)$.
-	The competitive ratio is
-
+	Hence, for serving at least $\chi$ unavoidable requests from $\sigma_E$, $\OPT$ pays at least
+	%
+	\[
+		\OPT(\sigma_E) + \OPT(M) \geq \chi \cdot \frac{2\alpha}{4 (\alpha-1)}\geq \frac{|\sigma_E|} {2} - 2 \alpha \cdot \ell.
+	\]
+	%	
+	To bound the competitive ratio, we transform the above inequality in the following way: $|\sigma_E| \leq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
+	For succintness, let $\xi := \OPT(\sigma_E) + \OPT(M)$.
+	Combining the bounds on the cost of \TAlg and \OPT during each finished phase, the competitive ratio is
+%
 	\[
@@
 	\]
-
+%	
 	\medskip

### 2020-05-10 paper/disc20clustering.tex
Expanding and rewriting the proof of k=3

@@
 \begin{proof}
-	If no component respecting partition exists after the merge of components, \TAlg simply ends the phase and performs no reconfiguration.
+	If no component respecting partition exists after the merge of components, then \TAlg resets all components, ends the phase and performs no reconfiguration.
 	It suffices to show that the merged component has size at least $4$ to conclude that \TAlg incurs no cost.
@@
 
-	Now we lower-bound the cost of $\OPT$.
+	Now we lower-bound the cost of the optimal offline solution.
+	To this end, we fix any optimal offline algorithm $\OPT$.
 	By $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$ we denote the cost of $\OPT$ on requests from sequences $\sigma_I$ and $\sigma_E$, respectively.

### 2020-05-10 paper/disc20clustering.tex
Wip

@@
 In Section~\ref{sec:k3}, we show an $O(k \cdot \ell)$-competitive algorithm for $k=3$ using the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental}.
-Prior to this section, we showed a lower bound of $\Omega(k \cdot \ell)$  that holds for the general model, hence the result from this section is asymptotically optimal (cf.
-Section~\ref{sec:lowerbound}).
+Prior to this section, we showed a lower bound of $\Omega(k \cdot \ell)$  that holds for the general model (cf.
+Section~\ref{sec:lowerbound}), hence the result from this section is asymptotically optimal.
 Furthermore, in Section~\ref{sec:k2}, we show a strictly $6$-competitive algorithm for $k=2$.
 
-\subsection{An Optimal Algorithm for Clusters of Size 3}
+\subsection{Optimal Algorithm for Clusters of Size 3}
 \label{sec:k3}

### 2020-05-10 paper/disc20clustering.tex
wip

@@
 
-\section{The Learning Problem} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
+\section{The Learning Model} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
 
@@
 	If the input finished before the first $\alpha$ external requests, then \TAlg is $1$-competitive.
-	If there were at least $\alpha$ external requests, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a migration; for the cost of \TAlg we follow the analysis of the unsaturated pairs.
+	If at least $\alpha$ external requests were issued, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a migration; for the cost of \TAlg we follow the analysis of the unsaturated pairs.
 	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.

### 2020-05-10 paper/disc20clustering.tex
wip

@@
 
-This problem is known as the dynamic balanced graph partitioning problem introduced by Avin et al.
+This problem is known as the dynamic balanced graph partitioning problem and was introduced by Avin et al.
 \cite{repartition-disc, sidma-arxiv} at DISC 2016.
@@
 \cite{sigmetrics19_partitioning} at SIGMETRICS 2019.
-We will refer to the latter as the learning model.
+We refer to the latter as the learning model.
 
@@
 \textbf{Resource augmentation.}
-The majority of work on Online Balanced Partitioning so far \cite{repartition-disc,sigmetrics19_partitioning} focuses on the scenario with resource augmentation, where the clusters of \ALG are larger than the clusters of \OPT.
+The majority of work on Online Balanced Partitioning so far \cite{repartition-disc,sigmetrics19_partitioning} focuses on the scenario with resource augmentation, where the clusters of an online algorithm are larger than the clusters of the offline optimal algorithm that we compare the performance to.
 We can adjust our construction to show a lower bound of $\Omega(\ell)$ for resource augmentation.
@@
 Finally, we note the possibility of improvement. The algorithm CREP~\cite{repartition-disc} requires $(2+\epsilon)$-augmentation to guarantee the competitive ratio independent of $\ell$.
+In contrast, our construction shows that the linear term $\ell$ is inevitable if the augmentation is smaller than $1+1/3$.
 

### 2020-05-10 paper/disc20clustering.tex
wip

@@
 Finally, we note the possibility of improvement. The algorithm CREP~\cite{repartition-disc} requires $(2+\epsilon)$-augmentation to guarantee the competitive ratio independent of $\ell$.
-In contrast, our construction shows that the linear term $\ell$ is inevitable if the augmentation is smaller than $1+1/3$.
+In contrast, our construction shows that the linear term $\ell$ is inevitable if the augmentation is smaller than~$1+1/3$.
 
@@
 	While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts: we separately bound $\OPT$ by $\OPT(\sigma_I) + \OPT(M)$ and $\OPT(\sigma_E) + \OPT(M)$.
-	Combining those bounds and using the relation between the maximum and the average gives us
+	Combining those bounds and using the relation between the maximum and the average, we obtain the bound
 	\begin{align*}
@@
 Move minimums to the paragraph text, by immediatelly expanding minimums (two steps in paragraph, instead of in the formula)}
+
+\maciek{"	I.e., $ C^t_{\set{u,v}} = \lambda$." is too short for a sentence. Merge with the previous sentence?}
 

### 2020-05-10 paper/disc20clustering.tex
Comments
wip

@@
 	\end{tabular}
-	\caption{Overview of known results and the results provided in this paper.
-By $\Rep$ we denote the cost of a single component repartition, and we elaborate on this in Section~\ref{sec:k3}.
-The first three rows show the results for the general partitioning model, except for the last row that shows the results for the learning model for arbitrary $k$ and $\ell$.
+	\caption{Overview of known results and our contributions.
+The first three rows summarizes the results for the general partitioning model, except for the last row that summarizes the results for the learning model for arbitrary $k$ and $\ell$.
 	}
@@
 
+\maciek{"and for final costs at t=m" for final requests}
+
+\maciek{"then the total cost to OPT" cost for OPT}
+
+\maciek{Suggestion: instead of "this means", I saw exclusively the use of "therefore", "hence" and "thus", and maybe "this implies" in papers}
+
 \noindent

### 2020-05-10 paper/disc20clustering.tex
Contributions

@@
 
+\maciek{the model name is still unclear: in the abstract we mention the repartitioning model introduced by Avin et al.}
+
 The popularity of data-centric, distributed applications has led to an explosive growth of network traffic, especially in data centers \cite{roy2015inside,singh2015jupiter}.
@@
 This paper presents several new results on the dynamic graph partitioning problem  without augmentation.
-Our main result is a tight bound on the general case: we present a lower bound of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
-even in the learning model.
-The best known lower bound so far was $\Omega(k)$,
-even for a more general model~\cite{repartition-disc}.
-We also present an asymptotically optimal, 
+For the learning model we present a lower bound of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
+(that holds also in the general partitioning model).
+The best known lower bound so far was $\Omega(k)$~\cite{repartition-disc} that holds only in the general partitioning model.
+We further present an asymptotically optimal, 
 $O(k\cdot \ell)$-competitive algorithm
 for the learning model.
-For the repartitioning model, we present  
+
+For the general partitioning model, we design
 an asymptotically optimal,
-$\Theta(\ell)$ algorithm for $k=3$, improving the best known upper bound 
+$\Theta(\ell)$-competitive algorithm for $k=3$, improving the best known upper bound 
 so far $O(\ell^2)$~\cite{repartition-disc}.
-We further present a strictly $6$-competitive algorithm for $k=2$ that improves upon previous $7$-competitive algorithm with $O(\ell^2)$ additive constant.
-%
+We further present a strictly $6$-competitive algorithm for $k=2$ that improves upon previous $7$-competitive algorithm with $O(\alpha\ell^2)$ additive constant.
+
+All algorithms in this paper have strict competitive ratio (i.e., without an additive term).
 Table \ref{tab:overview} provides an overview of our contributions compared to prior work.
-
-\maciek{emphasize that all our algorithms have strict competitive ratio}
 

### 2020-05-10 paper/disc20clustering.tex
Contributions
Expanding some sentences in the k=3 proof

@@
 
-\begin{table*}
+\begin{table*}[h]
 	\centering
@@
 For $k=2$, the learning problem is trivial: immediate collocation of a communicating pair is $1$-competitive.
-(A partitioning problem for $k=2$ is non-trivial, a lower bound of $3$ is known \cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.3\textwidth]{figs/substitute}
-	\caption{Example figure of the construction of the LB and growing components.}
-	\label{fig:nptree-construction}
-\end{figure}
+(A partitioning problem for $k=2$ is non-trivial, a lower bound of $3$ is known~\cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)
 
@@
 In comparison, we show that there exists an optimal offline algorithm \OPT that performs only two migrations.
+Figure~\ref{fig:lb} illustrates the constructs used in the lower bound.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.3\textwidth]{figs/substitute}
+	\caption{todo.}
+	\label{fig:lb}
+\end{figure}
 
@@
 \textbf{General partitioning problem.}
-Every lower bound for the learning problem constitutes a converging lower bound for the partitioning problem as the input sequence grows to infinity.
+Every lower bound for the learning problem constitutes a converging lower bound for the partitioning problem as the input sequence grows to infinity \maciek{possibly too strong sentence; I need to rethink}.
 However, our construction allows for an exact lower bound for partitioning with a~finite input sequence.
@@
 \begin{proof}
-	Fix a completed phase, and consider the state of \TAlg's counters at the end of it.
+	Fix a completed phase, and consider the state of \TAlg's counters at the end of it (before the reset).
 	We consider the incomplete phase later in this proof.
@@
 	As \TAlg is component respecting, it never increases any counter above $\alpha$.
-	We say that the pair $(u, v)$ is \emph{saturated} if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
+	We say that the pair $(u, v)$ is \emph{saturated} if the counter's value is $\alpha$, and \emph{unsaturated} otherwise (saturation of a pair leads to a merge action).
 	By $\sigma$ we denote the input sequence that arrived during the phase.

### 2020-05-10 paper/disc20clustering.tex
Comments
wip
Distributed

@@
 \begin{theorem}
-	\TAlg is $(60\ell)$-competitive for $k=3$.
+	\TAlg is $60\ell$-competitive for $k=3$.
 \end{theorem}
@@
 \textbf{Approximation for offline rematching.}
-\maciek{Maybe mention the offline approx here?}
+\maciek{Maybe mention the offline approx here instead of the next section? similar to e.g. the remarks after the LB}
+
+\mahmoud{mention the offline approx.}
 
@@
 
-\mahmoud{mention the offline approx.}
+This paper revisited the dynamic graph partitioning problem and presented several tight bounds for the important model where capacities cannot be exceeded, both for a general partitioning model and for a special learning model.
+
+While our bounds are tight, there are several interesting avenues for future research.
+In particular, we have so far focused on deterministic algorithms, and it would be interesting to study the power of randomization in this context.
+On the practical side, it would also be interesting to study our algorithms empirically, under realistic workloads.
+
+\medskip
+\noindent
+\textbf{Distributed implementation.}
 While we have described our algorithms globally so far, we note that these allow for efficient distributed implementations. 
@@
 The algorithm for $k=2$ from Section~\ref{sec:k2} performs only local communication for each request: counters are kept on the clusters and updated locally, and each migration concerns only two clusters.
-\medskip
-Now, we propose the distributed implementation of the algorithm for $k=3$ from Section~\ref{sec:k3}.
-The algorithm performs two non-local operations:
-\begin{enumerate}
-	\item Broadcasting the end of the phase.\label{it:broadcast}
-	\item Component merge.
-	\label{it:merge}
-\end{enumerate}
-
-If a component reaches size $4$ or a fresh cluster is not found, we conclude that no perfect partition exists.
-Then, we end the phase and broadcast the message to  reset the counters in all clusters.
-
-A merge of two components may require finding a third cluster that contains $3$ isolated nodes (for details see the proof of Lemma~\ref{lem:1req})).
+
+Furthermore, we propose an efficient distributed implementation of the algorithm for $k=3$ from Section~\ref{sec:k3}.
+The algorithm performs two types of operations that require communication with other clusters: a~component merge, and a broadcast of the end of the phase.
 We say that a cluster containing $3$ isolated nodes is \emph{fresh}.
+A merge of two components may require finding a fresh cluster (for details see the proof of Lemma~\ref{lem:1req})).
 In the following, we show how to efficiently find a fresh cluster in a distributed manner.
@@
 To find a fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
-Upon encountering a fresh cluster, we stop the traversal and decrease the counters on the followed path by $1$.
-
-The operation (\ref{it:broadcast}) requires a single broadcast, and the operation (\ref{it:merge}) requires $O(\log \ell)$ communication.
-Each of these may occur, on average, every $\alpha$ external requests.
-
-\medskip
-
-This paper revisited the dynamic graph partitioning problem and presented several tight bounds for the important model where capacities cannot be exceeded, both for a general partitioning model and for a special learning model.
-
-While our bounds are tight, there are several interesting avenues for future research.
-In particular, we have so far focused on deterministic algorithms, and it would be interesting to study the power of randomization in this context.
-On the practical side, it would also be interesting to study our algorithms empirically, under realistic workloads.
+Upon encountering a fresh cluster, we end the traversal and decrease the counters on the followed path by $1$.
+Summarizing, ending the phase requires a single broadcast, and merging components requires $O(\log \ell)$ communication.
 
@@
 \section{Pseudocode of Perfect Partition Learner}
+
+\maciek{Make it more general (essentially the pseudocode of DET) to also link to this section from k=3.}
+
+\maciek{todo: intro to the algorithm (mostly about the choice of repartition procedure in \TAlg and PPL)}
 

### 2020-05-10 paper/disc20clustering.tex
Distributed

@@
 
+\medskip
+\noindent
+\textbf{Distributed implementation.}
+While we have described the algorithm globally so far, we note that it allows for an efficient distributed implementations. 
+The algorithm performs two types of operations that require communication with other clusters: a~component merge, and a broadcast of the end of the phase.
+We say that a cluster containing $3$ isolated nodes is \emph{fresh}.
+A merge of two components may require finding a fresh cluster (for details see the proof of Lemma~\ref{lem:1req}).
+In the following, we show how to efficiently find a fresh cluster in a distributed manner.
+We organize the clusters into an arbitrary rooted balanced binary tree, and we broadcast the root to each cluster.
+Each cluster maintains the counter of fresh clusters in its subtree.
+To find a fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
+Upon encountering a fresh cluster, we end the traversal and decrease the counters on the followed path by $1$.
+Summarizing, ending the phase requires a single broadcast, and merging components requires $O(\log \ell)$ communication.
+
 \subsection{Improved Algorithm for Online Rematching} \label{sec:k2}
@@
 
-\medskip
-\noindent
-\textbf{Distributed implementation.}
-While we have described our algorithms globally so far, we note that these allow for efficient distributed implementations. 
+Although we have described our algorithms globally so far, we note that these allow for efficient distributed implementations. 
 The algorithm PPL from Section~\ref{sec:ppl} can be distributed
@@
 The algorithm for $k=2$ from Section~\ref{sec:k2} performs only local communication for each request: counters are kept on the clusters and updated locally, and each migration concerns only two clusters.
-
-Furthermore, we propose an efficient distributed implementation of the algorithm for $k=3$ from Section~\ref{sec:k3}.
-The algorithm performs two types of operations that require communication with other clusters: a~component merge, and a broadcast of the end of the phase.
-We say that a cluster containing $3$ isolated nodes is \emph{fresh}.
-A merge of two components may require finding a fresh cluster (for details see the proof of Lemma~\ref{lem:1req})).
-In the following, we show how to efficiently find a fresh cluster in a distributed manner.
-We organize the clusters into an arbitrary rooted balanced binary tree, and we broadcast the root to each cluster.
-Each cluster maintains the counter of fresh clusters in its subtree.
-To find a fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
-Upon encountering a fresh cluster, we end the traversal and decrease the counters on the followed path by $1$.
-Summarizing, ending the phase requires a single broadcast, and merging components requires $O(\log \ell)$ communication.
+Furthermore, we proposed an efficient distributed implementation of the algorithm for $k=3$ in Section~\ref{sec:k3}.
 

### 2020-05-10 paper/disc20clustering.tex
Comments
Distributed

@@
 
-\maciek{"this means," "\textbf{this means that}"}
-
 \maciek{Suggestion: "Note that \OPT cannot have both..." (till the end of paragraph) You told me this in one sentence during the call.
@@
 similarly to the approach in~\cite{sigmetrics19_partitioning}.
-The algorithm for $k=2$ from Section~\ref{sec:k2} performs only local communication for each request: counters are kept on the clusters and updated locally, and each migration concerns only two clusters.
+The algorithm for $k=2$ from Section~\ref{sec:k2} performs only local communication for each request: counters are kept on the clusters and updated locally, and each migration is local within two clusters that reached the collocation threshold $\lambda$.
 Furthermore, we proposed an efficient distributed implementation of the algorithm for $k=3$ in Section~\ref{sec:k3}.

### 2020-05-10 paper/disc20clustering.tex
wip

@@
 
+The algorithm in this section and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from \cite{repartition-disc}.
+The difference is in the choice of partition after a component merge.
+In DET, the partition was arbitrary.
+In the algorithm for the learning model, we choose the \maciek{component respecting?} partition closest to the initial partition.
+In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).
+
+\medskip
+
 In our algorithms, a~\emph{component} is a subset of frequently communicating nodes.
@@
 if the nodes belonging to the same component are always collocated.
-
-The algorithm in this section and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from \cite{repartition-disc}.
-The difference is in the choice of partition after a component merge.
-In DET, the partition was arbitrary.
-In the algorithm for the learning model, we choose the partition closest to the initial partition.
-\maciek{component respecting?}
-In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).
 
@@
 	Therefore, there are at most  $(k-1)\cdot\ell $ distinct requests
-	for which \PPL performs a repartitioning.
-	I.e.,
+	for which \PPL performs a repartitioning and
 	 \PPL performs at most $(k-1)\cdot\ell $ repartitions.

### 2020-05-10 paper/disc20clustering.tex
Avoiding repetitions
LB

@@
 We emphasize that ground sets are revealed only upon their nodes are split.
+The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
+Once we construct a ground set, it lasts until the end of the input sequence.
 We say that a ground set is \emph{trivial} if it contains exactly one node.
@@
 
-The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
-We emphasize that the algorithm does not know ground sets before they are revealed.
-Once we construct a ground set, it lasts until the end of the input sequence.
-
-The constructed ground sets can be perfectly partitioned into the clusters (i.e., with nodes of each ground set collocated).
+The ground sets constructed in this lower bound can be perfectly partitioned into the clusters (i.e., with nodes of each ground set collocated).
 This way, the optimal algorithm moves to a perfect partition at the beginning (where requests incur the cost~$0$), thus its cost is bounded.

### 2020-05-10 paper/disc20clustering.tex
Finishing
Comments

@@
 	We claim that expanding ground sets in this way can continue as long as at least $\ell$ isolated nodes exist.
-	Fix the configuration just before the creation of a ground set with $x_i$ for some $i$.
+	Fix the configuration just before the creation of a ground set with $x_i$ for some~$i$.
 	We show that if at least $\ell$ isolated nodes exist (including $x_i$), then there exists a~perfect partition of ground sets after expanding a ground set with $x_i$.
@@
 	To find the cluster for $\{x_0, y_0\}$, observe that we have at most $\ell-1$ clusters with isolated nodes, and we have at least $\ell$ isolated nodes --- hence a~cluster with two isolated nodes exists.
-	To find a feasible partition, we swap $\{x_0, y_0\}$ with these two collocated isolated nodes, and then we swap an isolated node from $I(x_i)$ with $x_i$.
+	To find a feasible partition, we swap $\{x_0, y_0\}$ with these two collocated isolated nodes, and then we swap an isolated node from $I(x_i)$ with~$x_i$.
 
@@
 
-\maciek{Suggestion: instead of "this means", I saw exclusively the use of "therefore", "hence" and "thus", and maybe "this implies" in papers}
+\maciek{(strong) Suggestion: instead of "this means", "meaning", I saw exclusively the use of "therefore", "hence" and "thus", and maybe "this implies" in papers. Also in PPl}
 

### 2020-05-10 paper/disc20clustering.tex
Rewording
minors

@@
 In our algorithms, a~\emph{component} is a subset of frequently communicating nodes.
-We maintain all nodes of a component always collocated in the same cluster~\cite{repartition-disc}.
+We keep all nodes of each component collocated in the same cluster, similar to Avin et al.~\cite{repartition-disc}.
 When we  move a~node, we move the whole component that contains it.
@@
 	denoted by $C^t_{\set{u,v}}$, 
-	reaches $\lambda$ immediately before \RM resets the counter.
-	I.e.,
+	reaches $\lambda$ immediately before \RM resets the counter,
+	i.e.,
 	$ C^t_{\set{u,v}} = \lambda$.
@@
 	the overall cost of requests in $\sigma_{\set{u,v}}$ incurred by \RM,
-	i.e., the total cost of remote serving
-	and the moving cost, is
+	i.e., the total cost of remote serving 	and the moving cost is
 	$\lambda + 2\alpha$.	
@@
 	we bound  costs incurred by \OPT for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
-	If $t$ is the first time that \RM collocates $\{u,v\}$ then  \OPT pays
+	If $t$ is the first time that \RM collocates $\{u,v\}$,
+	 then  \OPT pays
 	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
@@
 	+ \OPTM(\sigma_{\{u,v\}}) $.
+	If \OPT has both pairs separated during their respective intervals,
+	then obviously it pays $2\lambda$ during those intervals.
 	Note that \OPT cannot have both pairs collocated at the same time.
-	However, it can have one of the pairs already collocated prior to its respective interval.
-	Let us assume  \OPT has one pair,
-	e.g.~$\{u,v\}$ collocated already prior its respective interval, i.e.~$(t',t]$,
+%	However, it can have one of the pairs already collocated prior to its respective interval.
+	Let us assume \OPT has one of the pairs,
+	e.g.~$\{u,v\}$,
+	 collocated already prior its respective interval, $(t',t]$,
 	 and keeps it so during the interval.
@@
 	separating $\{u,v\}$.
-	We define times $t'' < t' < m$ similarly as in the former case.
+	We define times $t'' < t' < m$ similarly to the former case.
 	Let $R^*_{\{u,v\}} := \sigma_{\{u,w\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
@@
 	for all collocations performed by \RM,
-	and for final costs at $t=m$,
+	and for final requests at $t=m$,
 	\OPT pays at least 
@@
 
-\maciek{"similarly as in the former" "\textbf{similarly to the former}"}
-
 \maciek{Suggestion: "Note that \OPT cannot have both..." (till the end of paragraph) You told me this in one sentence during the call.
@@
 \maciek{OPT's cost formula: make it one line by removing the expression with minimums.
-Move minimums to the paragraph text, by immediatelly expanding minimums (two steps in paragraph, instead of in the formula)}
+Move minimums to the paragraph text, by immediately expanding minimums (two steps in paragraph, instead of in the formula)}
 

### 2020-05-11 paper/disc20clustering.tex
Intro to lb

--- 2020-05-10 paper/disc20clustering.tex
@@
 
-First, we construct a ground set of size $k-1$ on an arbitrary cluster $S$.
-In any configuration, there exists an isolated node on $S$.
-We issue requests between the current isolated node from $S$, and some node that was initially collocated with it.
+First, we construct a ground set of size $k-1$ on an arbitrarily chosen cluster.
+In any configuration, there must exist an isolated node collocated with the ground set of size $k-1$.
+We issue requests between this isolated node, and some node that was initially collocated with it.
 Repeating such requests, almost all the nodes must visit the cluster $S$.

### 2020-05-11 paper/disc20clustering.tex
Table width
Stretch
some changes to PPL model and k=2 comments

@@
 	\renewcommand{\arraystretch}{1.5}
-	\begin{tabular}{>{\centering\arraybackslash}p{4.5cm}|>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{4.5cm}}
+	\begin{tabular}{>{\centering\arraybackslash}p{4.25cm}|>{\centering\arraybackslash}p{4.25cm}>{\centering\arraybackslash}p{4.25cm}}
 		\rowcolor{gray!50}
@@
 We present an asymptotically optimal algorithm for the learning problem.
-The algorithm immediately collocates every communicating pair and never separates them.
-To help maintain the collocation of communicating pairs, we use the concept of components, introduced by Avin et al. \cite{repartition-disc}.
+The algorithm  collocates  a pair as soon as they communicate and it never separates them.
+In order to preserve collocated pairs,
+we employ the concept of components,
+introduced by Avin et al. \cite{repartition-disc}.
+
+%\medskip
+%In our algorithms, 
+A~\emph{component} is a subset of frequently communicating nodes.
+We keep all nodes of a component always collocated in the same cluster,
+i.e., when we  move a~node,
+we move the whole component that contains it.
+A~partition that maintains this invariant is a \emph{component-respecting} partition.
+%We say that a~partition is \emph{component-respecting}
+%if the nodes belonging to the same component are always collocated.
+We say that a component is \emph{trivial} if it contains exactly one node,
+and the node in this component is \emph{isolated}.
+
+In addition,
+we maintain a balanced partition of our components as long as such partition exists,
+a reminiscent of partitioning given integers into sets of equal sum \cite{integer-partitions-book}.
+In contrast, our partition is time-varying:  two components are merged into one component once they communicate, and we adjust the partition accordingly.
 
@@
 In DET, the partition was arbitrary.
-In the algorithm for the learning model, we choose the \maciek{component respecting?} partition closest to the initial partition.
+In the algorithm for the learning model,
+we choose a component-respecting partition closest to the initial partition.
 In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).
-
-\medskip
-
-In our algorithms, a~\emph{component} is a subset of frequently communicating nodes.
-We keep all nodes of each component collocated in the same cluster, similar to Avin et al.~\cite{repartition-disc}.
-When we  move a~node, we move the whole component that contains it.
-We maintain a balanced partition of our components,
-a reminiscent of partitioning given integers into sets of equal sum \cite{integer-partitions-book}.
-In contrast, our partition is time-varying:  two components are merged into one component once they communicate, and we adjust the partition accordingly.
-
-We say that a component is \emph{trivial} if it contains exactly one node,
-and the node in this component is \emph{isolated}.
-We say that a~partition is \emph{component-respecting}
-if the nodes belonging to the same component are always collocated.
+A generic sketch of these two algorithms can be found in the appendix (Algorithm \ref{alg:ppl}).
 
@@
 With each repartitioning,
-\PPL moves to a \maciek{component-respecting} partition that minimizes the distance to the initial partition $P_I$.
+\PPL moves to a component-respecting partition that minimizes the distance to the initial partition $P_I$.
 As a result,
@@
 	On each inter-cluster request,
-	the algorithm enumerates all \maciek{component-respecting} $\ell$-way partitions of components
+	the algorithm enumerates all component-respecting $\ell$-way partitions of components
 	that are in the same (closest) distance to $P_I$.
@@
 which  costs $2\alpha$,
-while OPT may save some costs by collocating multiple pairs at once.
+while \OPT may save some costs by collocating multiple pairs at once.
 % before they inflict too much communication cost.
@@
 \end{proof}
-\maciek{Fonts of OPT and \OPT differ. The latter is better.}
-
-\maciek{Excessive comma before ", is $\lambda +2\alpha$?}
 
 \maciek{m not introduced as the end of time in the proof, and also used in the intuitions. Consider some other symbol like $\tau$ or $t_F$}
-
-\maciek{"then \OPT pays $\lambda$ for serving" comma before then?}
-
-\maciek{", or $\alpha$ for collocating \textbf{and serving some requests remotely}" remove what's bold? The migration is sufficient}
-
-\maciek{"it can have one edge already collocated" at most one edge?}
 
@@
 Maybe use the same formulation?
-I feel like one sentence can be enough to explain this}
-
-\maciek{Suggestion: maybe call "its intervals" "\textbf{the saturating intervals}"?}
+I feel like one sentence can be enough to explain this \mahmoud{I removed one redundant sentence.
+Since this is the core of the proof let's keep it wordy.}}
 
 \maciek{OPT's cost formula: make it one line by removing the expression with minimums.
-Move minimums to the paragraph text, by immediately expanding minimums (two steps in paragraph, instead of in the formula)}
-
-\maciek{"	I.e., $ C^t_{\set{u,v}} = \lambda$." is too short for a sentence. Merge with the previous sentence?}
-
-\maciek{"and for final costs at t=m" for final requests}
-
-\maciek{"then the total cost to OPT" cost for OPT}
-
-\maciek{(strong) Suggestion: instead of "this means", "meaning", I saw exclusively the use of "therefore", "hence" and "thus", and maybe "this implies" in papers. Also in PPl}
+Move minimums to the paragraph text, by immediately expanding minimums (two steps in paragraph, instead of in the formula) \mahmoud{it is useful for reader's intuition to have the min till the end.
+So that they can remember why choice of $\lambda$ is not trivial.}}
 
@@
 \maciek{Maybe mention the offline approx here instead of the next section? similar to e.g. the remarks after the LB}
-
-\mahmoud{mention the offline approx.}
 

### 2020-05-11 paper/disc20clustering.tex
configuration=>partition, trivial=>singleton etc.

@@
 	pattern is static and can be perfectly partitioned;
-	we also provide a \mahmoud{asymptotically?} tight upper bound 
+	we also provide an asymptotically tight upper bound 
 	of $O(k\cdot \ell)$ for this scenario.
-	We contribute a tight upper bound
+	We contribute an asymptotically tight upper bound
 	of $\Theta(\ell)$ for $k=3$,
 	for the general model in which the
-	communication pattern can change arbitrarily
-	over time.
+	communication pattern can change arbitrarily over time.
 	We improve the result for $k=2$ by providing a strictly $6$-competitive upper bound for the general model.
@@
 Once we construct a ground set, it lasts until the end of the input sequence.
-We say that a ground set is \emph{trivial} if it contains exactly one node.
-We say that the node that belongs to trivial a ground set is \emph{isolated}.
+We say that a ground set is \emph{singleton} if it contains exactly one node.
+We say that the node that belongs to a singleton ground set is \emph{isolated}.
 
@@
 First, we construct a ground set of size $k-1$ on an arbitrarily chosen cluster.
-In any configuration, there must exist an isolated node collocated with the ground set of size $k-1$.
+In any partition, there must exist an isolated node collocated with the ground set of size $k-1$.
 We issue requests between this isolated node, and some node that was initially collocated with it.
@@
 	Fix any online algorithm \ALG{}.
-	Let $I(C)$ denote the cluster where nodes of the ground set $C$ are located in the initial configuration.
+	Let $I(C)$ denote the cluster where nodes of the ground set $C$ are located in the initial partition.
 	If $C$ contains nodes that originate from different clusters, $I(C)$ is undefined.
@@
 	%We never dismiss any constructed ground set.
-	Each cluster hosts exactly $k$ nodes, and in any configuration, a single isolated node resides in $S$.
+	Each cluster hosts exactly $k$ nodes, and in any partition, a single isolated node resides in $S$.
 	We say that this isolated node that accompanies the ground set $B$ is \emph{special}.
@@
 	We claim that expanding ground sets in this way can continue as long as at least $\ell$ isolated nodes exist.
-	Fix the configuration just before the creation of a ground set with $x_i$ for some~$i$.
+	Fix the partition just before the creation of a ground set with $x_i$ for some~$i$.
 	We show that if at least $\ell$ isolated nodes exist (including $x_i$), then there exists a~perfect partition of ground sets after expanding a ground set with $x_i$.
-	To show this claim, we consider a configuration, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
-	In this configuration, each cluster contains:  at most one ground set larger than $1$, possibly some isolated nodes, and possibly the ground set $\{x_0, y_0\}$.
+	To show this claim, we consider a partition, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
+	In this partition, each cluster contains:  at most one ground set larger than $1$, possibly some isolated nodes, and possibly the ground set $\{x_0, y_0\}$.
 	If $I(x_i)$ contains an isolated node, then we simply obtain a~perfect partition by swapping it with $x_i$.
@@
 	To keep \OPT's cost low, we terminate prematurely --- we stop extending ground sets while there are still $\ell+1$ isolated nodes (i.e., neglecting two last ground set expansions).
-	Consider \ALG's configuration after the last expansion.
+	Consider \ALG's partition after the last expansion.
 	Among $\ell+1$ isolated nodes there must exist a pair that originates from the same cluster (we have $\ell$ clusters). 
@@
 	Precisely, \OPT collocates $\set{x_0,y_0}$ by swapping them with $x^*$ and $y^*$.
-	In this configuration, no ground set is split.
-	We never issue requests between nodes $x^*, y^*$ that are split in this configuration.
+	In this partition, no ground set is split.
+	We never issue requests between nodes $x^*, y^*$ that are separated in this partition.
 
@@
 	Each expansion reduces the number of isolated nodes by $1$ or $2$.
-	The decrease of two isolated nodes occurs when the largest component on  $I(x_i)$ is still trivial; this occurs at most $\ell$ times.
+	The decrease of two isolated nodes occurs when the largest component on  $I(x_i)$ is still singleton; this occurs at most $\ell$ times.
 	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
@@
 To this end, we issue only enough requests until the algorithm collocates the nodes of each ground set (if it never does, we terminate after it pays $k\cdot \ell$).
-The construction is oblivious to the choice of the reconfiguration cost $\alpha$: we compare the 	number of node exchanges of \ALG and \OPT.
+The construction is oblivious to the choice of the repartitioning cost $\alpha$: we compare the	number of node exchanges of \ALG and \OPT.
 
@@
 we move the whole component that contains it.
-A~partition that maintains this invariant is a \emph{component-respecting} partition.
-%We say that a~partition is \emph{component-respecting}
+A~partition that maintains this invariant is a \emph{component respecting} partition.
+%We say that a~partition is \emph{component respecting}
 %if the nodes belonging to the same component are always collocated.
-We say that a component is \emph{trivial} if it contains exactly one node,
+We say that a component is \emph{singleton} if it contains exactly one node,
 and the node in this component is \emph{isolated}.
@@
 In the algorithm for the learning model,
-we choose a component-respecting partition closest to the initial partition.
+we choose a component respecting partition closest to the initial partition.
 In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).
 A generic sketch of these two algorithms can be found in the appendix (Algorithm \ref{alg:ppl}).
-
-\maciek{The remainder of the paper used "configuration C" to denote the partition P.
-Easiest to adjust here.
-\mahmoud{are "configuration" and "partition" the same thing everywhere?
-Then let's use "partition" only.\maciek{I thought partition is on component level, and configuration on node level (todo check where to add component-respecting to the partition in this chapter).
-Maybe say in the intro in this chapter that configuration = partition?
-We use both heavily.}}}
 
@@
 With each repartitioning,
-\PPL moves to a component-respecting partition that minimizes the distance to the initial partition $P_I$.
+\PPL moves to a component respecting partition that minimizes the distance to the initial partition $P_I$.
 As a result,
@@
 	On each inter-cluster request,
-	the algorithm enumerates all component-respecting $\ell$-way partitions of components
+	the algorithm enumerates all component respecting $\ell$-way partitions of components
 	that are in the same (closest) distance to $P_I$.
@@
 Once $C_{\set{x,y}} = \alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
-If no such partitioning exists, \TAlg resets all components to trivial components, resets all counters to $0$, and ends the phase.
+If no such partitioning exists, \TAlg resets all components to singleton components, resets all counters to $0$, and ends the phase.
 
@@
 	\label{lem:1req}
-	In a single reconfiguration of nodes (after a merge of components), \TAlg exchanges at most two pairs of nodes.
+	In a single repartitioning of nodes (after a merge of components), \TAlg exchanges at most two pairs of nodes.
 \end{lemma}
@@
 \begin{proof}
-	If no component respecting partition exists after the merge of components, then \TAlg resets all components, ends the phase and performs no reconfiguration.
+	If no component respecting partition exists after the merge of components, then \TAlg resets all components, ends the phase and performs no repartitioning.
 	It suffices to show that the merged component has size at least $4$ to conclude that \TAlg incurs no cost.
@@
 	%which we define as follows.
-	%A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial configuration of any cluster).
+	%A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial partition of any cluster).
 	%A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
@@
 
-	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
+	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the repartitioning is local within $U$ and $V$.
 	In this case, a local repartition is possible (within $U$ and $V$), for a cost of at most $2$ swaps.
-	If either $U$ or $V$ is of type $C_3$, a component of size $3$ participates in a merge, and we have a component of size at least $4$, and \TAlg ends the phase with no reconfiguration.
+	If either $U$ or $V$ is of type $C_3$, a component of size $3$ participates in a merge, and we have a component of size at least $4$, and \TAlg ends the phase with no repartitioning.
 
@@
 	exceeding this number would mean that a~component of size $4$ exists, and the phase should have ended already.
-	We bound the cost of each reconfiguration after a merge by Lemma~\ref{lem:1req}, obtaining $\TAlg(M) \leq 8\alpha\cdot\ell$.
+	We bound the cost of each repartitioning after a merge by Lemma~\ref{lem:1req}, obtaining $\TAlg(M) \leq 8\alpha\cdot\ell$.
 	%Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
@@
 	We have $\OPT(M) + \OPT(\sigma_I) \geq \alpha$, as the phase ended when the components of \TAlg{} could not be partitioned without splitting them.
-	Hence, for every possible configuration of $\OPT$, there exists a non-collocated pair of nodes with at least $\alpha$ requests between them, and
+	Hence, for every possible partition of $\OPT$, there exists a non-collocated pair of nodes with at least $\alpha$ requests between them, and
 	$\OPT$ either served them remotely or performed a~migration.
@@
 	Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
-	Fix a configuration of \OPT prior serving the sequence $\sigma_E$.
+	Fix a partition for \OPT prior serving the sequence $\sigma_E$.
 	Even without migrations, \OPT may mitigate paying for requests between at most $3\ell$ pairs of nodes that can be collocated in its clusters.
@@
 	Then, we cannot charge $\OPT$ due to the inability to partition the components.
-	Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
+	Instead, we use the fact that \TAlg and $\OPT$ started with the same initial partition.
 	If the input finished before the first $\alpha$ external requests, then \TAlg is $1$-competitive.
@@
 		+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}
-		\Big).
+		\Big),
 	\end{align*}
-
-Finally, we have
-
-\begin{align*}
+and
+$
 	\mathit{ALG} (\sigma)	/
@@
 	\Big) 	\leq 6.
-\end{align*}
-%	The claim follows from here.
-
+$
 \end{proof}
-
-\maciek{m not introduced as the end of time in the proof, and also used in the intuitions. Consider some other symbol like $\tau$ or $t_F$}
-
-\maciek{Suggestion: "Note that \OPT cannot have both..." (till the end of paragraph) You told me this in one sentence during the call.
-Maybe use the same formulation?
-I feel like one sentence can be enough to explain this \mahmoud{I removed one redundant sentence.
-Since this is the core of the proof let's keep it wordy.}}
-
-\maciek{OPT's cost formula: make it one line by removing the expression with minimums.
-Move minimums to the paragraph text, by immediately expanding minimums (two steps in paragraph, instead of in the formula) \mahmoud{it is useful for reader's intuition to have the min till the end.
-So that they can remember why choice of $\lambda$ is not trivial.}}
 
@@
 \textbf{Approximation for offline rematching.}
-\maciek{Maybe mention the offline approx here instead of the next section? similar to e.g. the remarks after the LB}
+\mahmoud{TBD}
+%We briefly sketch an approximation for the offline problem,
+%that makes use of \RM as a subroutine.
+%In our approximation algorithm,
+%we simulate \RM on the given sequence.
+%Whenever,
+%\RM collocoates a pair
 
@@
 		%        $k, \ell$,
-		%        initial configuration $P_I$,
+		%        initial partition $P_I$,
 		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
-		%        \Ensure A final configuration $P_F$ 
-		\STATE {For each node $v$ create a component $C_v = \{ v \}$ and add it to $\mathcal{C}$.}
+		%        \Ensure A final partition $P_F$ 
+		\STATE {For each node $v$ create a singleton component $C_v = \{ v \}$ and add it to $\mathcal{C}$.}
 		%\STATE{$P := P_I$}

### 2020-05-11 paper/disc20clustering.tex
LB for general problem: too trivial after all to include

@@
 	This paper revisits the online 
-	balanced repartitioning problem 
+	balanced partitioning problem 
 	(introduced by Avin et al.~at DISC 2016)
@@
 	of $O(k\cdot \ell)$ for this scenario.
-	We contribute an asymptotically tight upper bound
-	of $\Theta(\ell)$ for $k=3$,
+	For $k=3$, we contribute an asymptotically tight upper bound
+	of $\Theta(\ell)$
 	for the general model in which the
@@
 \section{Introduction}
-
-\maciek{the model name is still unclear: in the abstract we mention the repartitioning model introduced by Avin et al.}
 
@@
 In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.
-
-\maciek{Above definition is the definition of strict competitive ratio. In comparison to non-strict with additive constant. Emphasize? Maybe already use the notion of \OPT}
+We emphasize that all our algorithms have a \emph{strict} competitive ratio (without an additive term).
 
@@
 \end{proof}
-
-\noindent
-\textbf{General partitioning problem.}
-Every lower bound for the learning problem constitutes a converging lower bound for the partitioning problem as the input sequence grows to infinity \maciek{possibly too strong sentence; I need to rethink}.
-However, our construction allows for an exact lower bound for partitioning with a~finite input sequence.
-To this end, we issue only enough requests until the algorithm collocates the nodes of each ground set (if it never does, we terminate after it pays $k\cdot \ell$).
-The construction is oblivious to the choice of the repartitioning cost $\alpha$: we compare the	number of node exchanges of \ALG and \OPT.
 

### 2020-05-11 paper/disc20clustering.tex
Comments

@@
 We emphasize that a repartitioning by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
-Meaning, it never moves to a partition beyond the distance $\Delta^*$.
+This way \PPL never moves to a partition beyond the distance $\Delta^*$.
 
@@
 
+\maciek{OPT's font differ from other proofs. We don't want italic OPT in mathenv}
+
+\maciek{I really think that minimums should be expanded earlier than in the competitive ratio.
+The ratio should be bounded easily, and if you wish, you can emphasize the choice in the paragraph, or dedicate entire line to the cost expression.
+}
+
 \noindent

### 2020-05-11 paper/disc20clustering.tex
wip

@@
 We study two models in this paper: the \emph{general partitioning} model, and its subproblem, the \emph{learning} model.
+In both models, we assume that communication patterns are not known to our algorithms at the beginning.
+We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.
+In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.
+%We emphasize that all our algorithms have a \emph{strict} competitive ratio (without an additive term).
 
@@
 For the learning model, we  assume that the migration cost is $\alpha=1$.
-
-\medskip 
-In both models, we assume that communication patterns are not known to our algorithms at the beginning.
-We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.
-In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.
-We emphasize that all our algorithms have a \emph{strict} competitive ratio (without an additive term).
 
@@
 Upon encountering a fresh cluster, we end the traversal and decrease the counters on the followed path by $1$.
-Summarizing, ending the phase requires a single broadcast, and merging components requires $O(\log \ell)$ communication.
+Summarizing, ending the phase requires a single broadcast, and merging components has $O(\log \ell)$ communication complexity.
 

### 2020-05-11 paper/disc20clustering.tex
Consistency

@@
 \textbf{Learning model.}
-We study a variant of Online Balanced Partitioning,
+We study a \emph{learning} variant of dynamic balanced graph partitioning,
 where the communication pattern is \emph{static}:

### 2020-05-11 paper/disc20clustering.tex
Consistency

@@
 	\caption{Overview of known results and our contributions.
-The first three rows summarizes the results for the general partitioning model, except for the last row that summarizes the results for the learning model for arbitrary $k$ and $\ell$.
+The table summarizes the results for the general partitioning model, except for the last row that summarizes the results for the learning model for arbitrary $k$ and $\ell$.
 	}

### 2020-05-11 paper/disc20clustering.tex
Consistency

@@
 
-In this section, we consider the learning variant of Dynamic Balanced Graph Partitioning.
+In this section, we consider the learning variant of dynamic balanced graph partitioning problem.
 For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$.
-The lower bound can be easily translated to the partitioning model (studied in Section~\ref{sec:part}) with a~finite request sequence.
-At the end of this section, we discuss an asymptotically optimal upper bound.
+The lower bound holds also in the general partitioning model (studied in Section~\ref{sec:part}).
+At the end of this section, we discuss an asymptotically optimal upper bound for the learning variant.
 

### 2020-05-11 paper/disc20clustering.tex
wip

@@
 In this section, we consider the learning variant of dynamic balanced graph partitioning problem.
-For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$.
+For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$ for $k\geq 3$.
 The lower bound holds also in the general partitioning model (studied in Section~\ref{sec:part}).
@@
 The lower bound requires $k\geq 3$.
-For $k=2$, the learning problem is trivial: immediate collocation of a communicating pair is $1$-competitive.
+In contrast, for $k=2$ the learning problem is trivial: immediate collocation of communicating pairs is $1$-competitive.
 (A partitioning problem for $k=2$ is non-trivial, a lower bound of $3$ is known~\cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)
@@
 Once we construct a ground set, it lasts until the end of the input sequence.
-We say that a ground set is \emph{singleton} if it contains exactly one node.
+We say that a ground set is a \emph{singleton} if it contains exactly one node.
 We say that the node that belongs to a singleton ground set is \emph{isolated}.

### 2020-05-11 paper/disc20clustering.tex
comments

@@
 The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
-Once we construct a ground set, it lasts until the end of the input sequence.
+Once we construct a ground set, it lasts until the end of the algorithm's execution.
 We say that a ground set is a \emph{singleton} if it contains exactly one node.
@@
 The majority of work on Online Balanced Partitioning so far \cite{repartition-disc,sigmetrics19_partitioning} focuses on the scenario with resource augmentation, where the clusters of an online algorithm are larger than the clusters of the offline optimal algorithm that we compare the performance to.
-We can adjust our construction to show a lower bound of $\Omega(\ell)$ for resource augmentation.
+We can adjust our construction to show a lower bound of $\Omega(\ell)$ for a setting with resource augmentation.
 
@@
 we employ the concept of components,
-introduced by Avin et al. \cite{repartition-disc}.
+introduced by Avin~et~al.~\cite{repartition-disc}.
 
@@
 %In our algorithms, 
-A~\emph{component} is a subset of frequently communicating nodes.
+A~\emph{component} is a subset of frequently communicating nodes \maciek{this is not a good definition. intuitions, maybe}.
 We keep all nodes of a component always collocated in the same cluster,
@@
 $\OPT \geq \Delta^*:= \Delta(P_F, P_I) $.
-We  use $ \Delta(P_F) $ wherever the initial partition is clear from the context.
+We  use $ \Delta(P_F) $ wherever the initial partition is clear from the context \maciek{It always is, we fixed it in the beginning.
+Maybe just define it as the distance from initial partition, but don't pass it as an argument}.
 The scheme of the algorithm can be found in the  appendix (Algorithm \ref{alg:ppl}).

### 2020-05-11 paper/disc20clustering.tex
Algorithm formulation

@@
 \textbf{Component-based algorithm.}
-We introduce the algorithm \TAlg that we analyze in this section.
-The algorithm partitions nodes into components, and
+The algorithm \TAlg partitions nodes into components, and
 initially, each node is isolated (belongs to its own component).
@@
 
-We describe the algorithm \RM for the online rematching problem as follows.
-The input to \RM is a sequence of  requests
-$\sigma:=\{\sigma_1,\dots, \sigma_m\}, \sigma_t:=\{x,y\} \in V^2, x \neq y$.
-We maintain a counter $C_{\set{x,y}}$ for each pair of nodes $\set{x,y}$ and increment it on every remote request between $x$ and $y$.
+\medskip
+
+\noindent
+\textbf{Algorithm ReMatch.} 
+The algorithm ReMatch (\RM) maintains a counter $C_{\set{x,y}}$ for each pair of nodes $\set{x,y}$ and increments it on every remote request between $x$ and $y$.
 Once $C_{\set{x,y}} = \lambda$,
-reset the counter and collocate the pair arbitrarily in one of the two clusters where $x$ or $y$ resides.
-
-\begin{theorem} \label{thm:k=2}
-	For $\lambda=\alpha$, the algorithm \RM  is strictly 6-competitive.
-\end{theorem}
-
+it resets the counter $C_{\set{x,y}}:=0$ and collocates the pair by swapping $x$ with the node collocated with $y$.
+
+\medskip
+
+\noindent
+\textbf{The charging scheme.}
 We charge both \OPT and \RM whenever \RM collocates a pair.%(i.e., on a collocation event).
@@
 
-\textbf{The charging scheme.}
 Consider two  pairs that share the same node, 
@@
 
-\begin{proof}[Proof of Theorem \ref{thm:k=2}]
+\begin{theorem} \label{thm:k=2}
+	For $\lambda=\alpha$, the algorithm \RM  is strictly 6-competitive.
+\end{theorem}
+
+\begin{proof}%[Proof of Theorem \ref{thm:k=2}]
+	Fix an input sequence of requests
+$\sigma:=\{\sigma_1,\dots, \sigma_m\}$.
 	Assume on request $\sigma_{t}=\set{u,v}$, 

### 2020-05-11 paper/disc20clustering.tex
New section about algorithmic techniques

@@
 \end{table*}
+
+\medskip
+\noindent
+\textbf{Algorithmic techniques.} The variant for general $k$ is still unresolved, hence it is vital to summarize algorithmic techniques used in this paper.
+A straightforward analysis of the algorithm from Section~\ref{sec:k3} results in the bound $O(\ell^2)$, and the analysis must be improved in two places.
+One must bound the cost of each of $O(\ell)$ reconfigurations in a phase by a~constant, and show that the optimum offline algorithm must pay for a significant portion of inter-cluster requests.
+We bounded the cost of the latter by estimating the capabilities of the optimum offline algorithm to \emph{prepare} for an incoming sequence of requests.
+In the analysis of the algorithm from Section~\ref{sec:k2}, we propose a novel charging scheme for intersecting edges.
+
+Furthermore, we state the implications of the tight result for the learning model.
+If one hopes to improve the lower bound for the general model, the ground sets must be dynamically added and \emph{removed}: if ground set are only added, the instance is solvable by the algorithm from Section~\ref{sec:ppl} within the cost $O(k\cdot \ell)$.
 

### 2020-05-11 paper/disc20clustering.tex
New section about algorithmic techniques

@@
 We bounded the cost of the latter by estimating the capabilities of the optimum offline algorithm to \emph{prepare} for an incoming sequence of requests.
-In the analysis of the algorithm from Section~\ref{sec:k2}, we propose a novel charging scheme for intersecting edges.
-
-Furthermore, we state the implications of the tight result for the learning model.
+Furthermore, in the analysis of the algorithm from Section~\ref{sec:k2}, we proposed a~novel charging scheme for edges that share a vertex.
+
+We emphasize the implications of the tight result for the learning model.
 If one hopes to improve the lower bound for the general model, the ground sets must be dynamically added and \emph{removed}: if ground set are only added, the instance is solvable by the algorithm from Section~\ref{sec:ppl} within the cost $O(k\cdot \ell)$.

### 2020-05-11 paper/disc20clustering.tex
wip

@@
 One must bound the cost of each of $O(\ell)$ reconfigurations in a phase by a~constant, and show that the optimum offline algorithm must pay for a significant portion of inter-cluster requests.
-We bounded the cost of the latter by estimating the capabilities of the optimum offline algorithm to \emph{prepare} for an incoming sequence of requests.
-Furthermore, in the analysis of the algorithm from Section~\ref{sec:k2}, we proposed a~novel charging scheme for edges that share a vertex.
+We bound the cost of the latter by estimating the capabilities of the optimum offline algorithm to \emph{prepare} for an incoming sequence of requests.
+Furthermore, in the analysis of the algorithm from Section~\ref{sec:k2}, we propose a~novel charging scheme for edges that share a vertex.
 
@@
 \label{sec:k3}
+
+\maciek{Intro is missing.}
+\maciek{
+	A straightforward analysis of the algorithm from Section~\ref{sec:k3} results in the bound $O(\ell^2)$, and the analysis must be improved in two places.
+One must bound the cost of each of $O(\ell)$ reconfigurations in a phase by a~constant, and show that the optimum offline algorithm must pay for a significant portion of inter-cluster requests.
+	We bounded the cost of the latter by estimating the capabilities of the optimum offline algorithm to \emph{prepare} for an incoming sequence of requests.}
 

### 2020-05-12 paper/disc20clustering.tex
comments
The right insigths into generalizing LB

@@
 A straightforward analysis of the algorithm from Section~\ref{sec:k3} results in the bound $O(\ell^2)$, and the analysis must be improved in two places.
-One must bound the cost of each of $O(\ell)$ reconfigurations in a phase by a~constant, and show that the optimum offline algorithm must pay for a significant portion of inter-cluster requests.
+One must bound the cost of each of $O(\ell)$ reconfigurations in a phase by a~constant, and show that the optimum offline algorithm must pay for a significant portion of inter-cluster requests \maciek{how to concisely say that there is also a migrate possibilty}.
 We bound the cost of the latter by estimating the capabilities of the optimum offline algorithm to \emph{prepare} for an incoming sequence of requests.
@@
 We say that the node that belongs to a singleton ground set is \emph{isolated}.
-
-The ground sets constructed in this lower bound can be perfectly partitioned into the clusters (i.e., with nodes of each ground set collocated).
-This way, the optimal algorithm moves to a perfect partition at the beginning (where requests incur the cost~$0$), thus its cost is bounded.
-If the algorithm splits any pair of nodes of a ground set, the adversary issues requests to any pair of split nodes.
-These two observations imply that the algorithm must eventually collocate all nodes of a~ground set to be competitive.
 
@@
 \end{proof}
+
+\noindent
+\textbf{Lower bound for the general problem.}
+For a lower bound $\Omega(k\cdot \ell)$ for the general partitioning problem, the adversary continues to issue requests to split nodes of a ground set until the algorithm collocates them.
+The ground sets constructed in our lower bound can be perfectly partitioned into the clusters.
+This way, the optimal algorithm moves to a perfect partition at the beginning (where requests incur the cost~$0$), thus its cost is bounded.
+If the algorithm splits any pair of nodes of a ground set, the adversary issues requests to any pair of split nodes.
+These two observations imply that the algorithm must eventually collocate all nodes of a~ground set to be competitive.
+The construction is oblivious to the choice of the reconfiguration cost $\alpha$: we compare the 	number of node exchanges of \ALG and \OPT.
+
+\medskip
 

### 2020-05-12 paper/disc20clustering.tex
The right insigths into generalizing LB

@@
 First, we construct a ground set of size $k-1$ on an arbitrarily chosen cluster.
-In any partition, there must exist an isolated node collocated with the ground set of size $k-1$.
-We issue requests between this isolated node, and some node that was initially collocated with it.
+In any partition, there must exist an isolated node collocated with the ground set of size $k-1$, and we call it \emph{special}.
+We issue requests between the current special node, and some node that was initially collocated with it.
 Repeating such requests, almost all the nodes must visit the cluster $S$.

### 2020-05-12 paper/disc20clustering.tex
The right insigths into generalizing LB

@@
 
-First, we construct a ground set of size $k-1$ on an arbitrarily chosen cluster.
+We start by constructing a ground set of size $k-1$ on an arbitrarily chosen cluster.
 In any partition, there must exist an isolated node collocated with the ground set of size $k-1$, and we call it \emph{special}.
@@
 	Throughout the construction, we maintain ground sets for each node.
-	Initially, all nodes are isolated (i.e., in ground sets of size $1$).
+	Initially, all nodes are isolated (i.e., in singleton ground sets).
 	First, we create a ground set $B$ of $k-1$ nodes on an arbitrarily chosen cluster (the nodes  are already collocated).
@@
 	There is no space to accommodate $\{x_0, y_0\}$ in the cluster of $x_0$ (as $B$'s size is $k-1$), hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.
-	To preserve a feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
+	To preserve a~feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
 	\ALG must replace $x_0$ with another single node that we call $x_1$.
@@
 For a lower bound $\Omega(k\cdot \ell)$ for the general partitioning problem, the adversary continues to issue requests to split nodes of a ground set until the algorithm collocates them.
-The ground sets constructed in our lower bound can be perfectly partitioned into the clusters.
-This way, the optimal algorithm moves to a perfect partition at the beginning (where requests incur the cost~$0$), thus its cost is bounded.
-If the algorithm splits any pair of nodes of a ground set, the adversary issues requests to any pair of split nodes.
-These two observations imply that the algorithm must eventually collocate all nodes of a~ground set to be competitive.
-The construction is oblivious to the choice of the reconfiguration cost $\alpha$: we compare the 	number of node exchanges of \ALG and \OPT.
+Note that the ground sets constructed in our lower bound can be perfectly partitioned into the clusters.
+Hence, the optimal algorithm moves to a perfect partition at the beginning (where requests incur the cost~$0$), and its cost is bounded.
+This means that the algorithm must eventually collocate all nodes of a~ground set to be competitive.
+We reveal the next ground set only after the collocation, hence we can repeat the analysis of the algorithm for the learing problem.
+Finally, we note that the construction is oblivious to the choice of the reconfiguration cost $\alpha$: we compare the 	number of node exchanges of \ALG and \OPT.
 

### 2020-05-12 paper/disc20clustering.tex
tildes

@@
 	Distributed   applications,  including  batch  processing, streaming, scale-out databases,
-	or machine learning, generate a significant amount of network traffic.
+	or machine learning, generate a~significant amount of network traffic.
 	By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), we can reduce the network load and  improve application performance. 
-	However, the communication pattern of different applications is often unknown a priori and may change over time, hence it needs to be learned in an online manner.
+	However, the communication pattern of different applications is often unknown a~priori and may change over time, hence it needs to be learned in an online manner.
 	%
@@
 	%
-	Our first contribution is a significantly improved
+	Our first contribution is a~significantly improved
 	lower bound of $\Omega(k\cdot \ell)$ on the
@@
 	of clusters and $k$ is the cluster size,
-	even for a scenario in which the communication
+	even for a~scenario in which the communication
 	pattern is static and can be perfectly partitioned;
@@
 	communication pattern can change arbitrarily over time.
-	We improve the result for $k=2$ by providing a strictly $6$-competitive upper bound for the general model.
+	We improve the result for $k=2$ by providing a~strictly $6$-competitive upper bound for the general model.
 	In contrast to most prior work, our algorithms respect all capacity constraints and do not require resource augmentation.
@@
 
-The popularity of data-centric, distributed applications has led to an explosive growth of network traffic, especially in data centers \cite{roy2015inside,singh2015jupiter}.
-The performance of these distributed applications often critically depends on the underlying network \cite{mogul2012we}, and efficient operation of these networks is important.
+The popularity of data-centric, distributed applications has led to an explosive growth of network traffic, especially in data centers~\cite{roy2015inside,singh2015jupiter}.
+The performance of these distributed applications often critically depends on the underlying network~\cite{mogul2012we}, and efficient operation of these networks is important.
 At the same time, distributed systems are often highly virtualized today, and provide interesting new opportunities for resource optimization.
-In particular, it has become possible to operate data centers in a more demand-aware manner: 
+In particular, it has become possible to operate data centers in a~more demand-aware manner: 
 by dynamically migrating nodes (e.g., virtual machines) which communicate frequently topologically closer to each other, network traffic can be reduced significantly.  
@@
 This paper studies the algorithmic problem underlying such demand-aware
-optimizations, aiming to strike a balance between the benefits of migrations (e.g., reduced network load) and their costs.
+optimizations, aiming to strike a~balance between the benefits of migrations (e.g., reduced network load) and their costs.
 In particular, we are interested in an online variant of the problem: since communication patterns can change over time, an online algorithm needs to react dynamically to new traffic patterns, and migrate nodes  accordingly.
@@
 
-This problem is known as the dynamic balanced graph partitioning problem and was introduced by Avin et al.
-\cite{repartition-disc, sidma-arxiv} at DISC 2016.
-A special variant of the general problem has later been studied by Henzinger et al.
-\cite{sigmetrics19_partitioning} at SIGMETRICS 2019.
+This problem is known as the dynamic balanced graph partitioning problem and was introduced by Avin et al.~\cite{repartition-disc, sidma-arxiv} at DISC 2016.
+a~special variant of the general problem has later been studied by Henzinger et al.~\cite{sigmetrics19_partitioning} at SIGMETRICS 2019.
 We refer to the latter as the learning model.
@@
 We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.
-In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.
-%We emphasize that all our algorithms have a \emph{strict} competitive ratio (without an additive term).
+In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a~given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.
+%We emphasize that all our algorithms have a~\emph{strict} competitive ratio (without an additive term).
 
@@
 \textbf{General partitioning model.}
-In the \emph{dynamic balanced graph partitioning} problem, we are given a set $V$ of $n$ nodes 
+In the \emph{dynamic balanced graph partitioning} problem, we are given a~set $V$ of $n$ nodes 
 (e.g., virtual machines or processes),
@@
 $\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
-where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a~certain amount of data.
+where a~pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a~certain amount of data.
 Nodes in $C \subset V$ are \emph{collocated}
@@
 
-An algorithm serves a communication request between two nodes
+An algorithm serves a~communication request between two nodes
 either \emph{locally} at cost~0
@@
 and \emph{external} requests, respectively.
-Before serving a request,
-an online algorithm may perform a \emph{repartition},
+Before serving a~request,
+an online algorithm may perform a~\emph{repartition},
 %(i.e., \emph{reconfigure}).
@@
 the algorithm serves the  request.
-The cost of migrating a node from one cluster to another
+The cost of migrating a~node from one cluster to another
 is~$\alpha \in \mathbb{Z}^+$.
@@
 \textbf{Learning model.}
-We study a \emph{learning} variant of dynamic balanced graph partitioning,
+We study a~\emph{learning} variant of dynamic balanced graph partitioning,
 where the communication pattern is \emph{static}:
-whether a pair of  nodes ever communicate or not, 
-is determined a priori and is unknown to algorithms,
+whether a~pair of  nodes ever communicate or not, 
+is determined a~priori and is unknown to algorithms,
  and requests to such pairs arrive indefinitely.
@@
 as otherwise it cannot be competitive.
-As in Henzinger et al. \cite{sigmetrics19_partitioning}, we assume that the communication graph admits a \emph{perfect partition},
-i.e., a partition in which no inter-cluster request ever occurs.
+As in Henzinger et al.~\cite{sigmetrics19_partitioning}, we assume that the communication graph admits a~\emph{perfect partition},
+i.e., a~partition in which no inter-cluster request ever occurs.
 %any algorithm must learn this perfect partition.
@@
 The two works closest to ours are by Avin et al.
-(on the general partitioning model) \cite{repartition-disc, sidma-arxiv} and by Henzinger et al.
-(on the learning model) \cite{sigmetrics19_partitioning}.
+(on the general partitioning model)~\cite{repartition-disc, sidma-arxiv} and by Henzinger et al.
+(on the learning model)~\cite{sigmetrics19_partitioning}.
 However, the focus of these papers is primarily on models with resource augmentation: the online algorithm can use slightly larger clusters than the offline algorithm.  
 Avin et al.
-actually showed that their lower bound $\Omega(k)$ holds even for a significant resource augmentation, and they provided an algorithm with the competitive ratio $O(k \log k)$ using the $(2+\epsilon)$-augmented cluster capacity.
+actually showed that their lower bound $\Omega(k)$ holds even for a~significant resource augmentation, and they provided an algorithm with the competitive ratio $O(k \log k)$ using the $(2+\epsilon)$-augmented cluster capacity.
 Their ratio is independent of $\ell$, which is impossible without significant resource augmentation.
@@
 In contrast, we study the non-augmented setting, where the nodes need to be perfectly balanced  among the clusters.
-This assumption is not only more realistic but also significantly more challenging, as it is related to hard problems such as integer partitioning \cite{integer-partitions-book}.
-In terms of results without augmentation, so far, it is only known that there exists a $O(k^2 \cdot \ell^2)$-competitive algorithm \cite{repartition-disc}; the best known lower bound is significantly lower, namely $\Omega(k)$.
-For $k=2$, Avin et al.\cite{repartition-disc} presented a $7$-competitive algorithm with a~substantial ($\Omega(\ell^2)$) additive constant.
-
-The problem has also been studied in a weaker
+This assumption is not only more realistic but also significantly more challenging, as it is related to hard problems such as integer partitioning~\cite{integer-partitions-book}.
+In terms of results without augmentation, so far, it is only known that there exists a~$O(k^2 \cdot \ell^2)$-competitive algorithm~\cite{repartition-disc}; the best known lower bound is significantly lower, namely $\Omega(k)$.
+For $k=2$, Avin et al.~\cite{repartition-disc} presented a~$7$-competitive algorithm with a~substantial ($\Omega(\ell^2)$) additive constant.
+
+The problem has also been studied in a~weaker
 model where the adversary can only sample
-requests from a fixed distribution~\cite{stochastic-ring}.
-
-The static offline version of~the~partitioning~problem, i.e., a problem variant where
+requests from a~fixed distribution~\cite{stochastic-ring}.
+
+The static offline version of~the~partitioning~problem, i.e., a~problem variant where
 migration is not allowed, where all requests are known in advance, and where
@@
 as online caching, where requests for data items (nodes) arrive over time and
-need to be served from a cache of finite capacity, and where the number of
+need to be served from a~cache of finite capacity, and where the number of
 cache misses must be minimized. Classic problem variants usually boil down to
-finding a smart eviction strategy, such as Least Recently Used (LRU)~\cite{SleTar85}. In our
+finding a~smart eviction strategy, such as Least Recently Used (LRU)~\cite{SleTar85}. In our
 setting, requests can be served remotely (i.e.,~without fetching the
-corresponding nodes to a single physical machine). In this light, our model is more
+corresponding nodes to a~single physical machine). In this light, our model is more
 reminiscent of caching models \emph{with
 bypassing}~\cite{EpImLN11,EpImLN15,Irani02}.
-A major difference between  these problems is that in the caching problems, each request involves a~single element of the universe, while in our model \emph{both} endpoints of a communication request are subject to~optimization.
-In this light, we can see our model as a "symmetric" version of online paging.
-
-Dynamic graph partitioning problems are generally fundamental in computer science, and arise in many different contexts \cite{streaming-soda,streaming1}.
+A major difference between  these problems is that in the caching problems, each request involves a~single element of the universe, while in our model \emph{both} endpoints of a~communication request are subject to~optimization.
+In this light, we can see our model as a~"symmetric" version of online paging.
+
+Dynamic graph partitioning problems are generally fundamental in computer science, and arise in many different contexts~\cite{streaming-soda,streaming1}.
 
@@
 %caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00},
-%see~\cite{repartition-disc} for a discussion.
+%see~\cite{repartition-disc} for a~discussion.
 %The static offline version of~the~problem, called the
@@
 This paper presents several new results on the dynamic graph partitioning problem  without augmentation.
-For the learning model we present a lower bound of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
+For the learning model we present a~lower bound of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
 (that holds also in the general partitioning model).
@@
 so far $O(\ell^2)$~\cite{repartition-disc}.
-We further present a strictly $6$-competitive algorithm for $k=2$ that improves upon previous $7$-competitive algorithm with $O(\alpha\ell^2)$ additive constant.
+We further present a~strictly $6$-competitive algorithm for $k=2$ that improves upon previous $7$-competitive algorithm with $O(\alpha\ell^2)$ additive constant.
 
@@
 		\textbf{$k=3$}&  $\Omega(\ell)$ \hspace{0.3cm}(\S \ref{sec:lowerbound})& $O(\ell) $\hspace{0.3cm}(\S \ref{sec:k3})\\
-		$k > 3$ & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k^2 \cdot \ell^2)$\hspace{0.1cm} \cite{repartition-disc} \\
+		$k > 3$ & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k^2 \cdot \ell^2)$\hspace{0.1cm}~\cite{repartition-disc} \\
 		\rowcolor{gray!25}
@@
 A straightforward analysis of the algorithm from Section~\ref{sec:k3} results in the bound $O(\ell^2)$, and the analysis must be improved in two places.
-One must bound the cost of each of $O(\ell)$ reconfigurations in a phase by a~constant, and show that the optimum offline algorithm must pay for a significant portion of inter-cluster requests \maciek{how to concisely say that there is also a migrate possibilty}.
+One must bound the cost of each of $O(\ell)$ reconfigurations in~phase by a~constant, and show that the optimum offline algorithm must pay for a~significant portion of inter-cluster requests \maciek{how to concisely say that there is also a~migrate possibilty}.
 We bound the cost of the latter by estimating the capabilities of the optimum offline algorithm to \emph{prepare} for an incoming sequence of requests.
-Furthermore, in the analysis of the algorithm from Section~\ref{sec:k2}, we propose a~novel charging scheme for edges that share a vertex.
+Furthermore, in the analysis of the algorithm from Section~\ref{sec:k2}, we propose a~novel charging scheme for edges that share a~vertex.
 
@@
 In this section, we consider the learning variant of dynamic balanced graph partitioning problem.
-For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$ for $k\geq 3$.
+For this setting, we show a~surprisingly high lower bound of $\Omega(k \cdot \ell)$ for $k\geq 3$.
 The lower bound holds also in the general partitioning model (studied in Section~\ref{sec:part}).
@@
 
-We provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm for the learning problem.
-Later we elaborate on how to efficiently transform it to a lower bound for the general partitioning problem.
+We provide a~lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm for the learning problem.
+Later we elaborate on how to efficiently transform it to a~lower bound for the general partitioning problem.
 The lower bound requires $k\geq 3$.
 In contrast, for $k=2$ the learning problem is trivial: immediate collocation of communicating pairs is $1$-competitive.
-(A partitioning problem for $k=2$ is non-trivial, a lower bound of $3$ is known~\cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)
+(A partitioning problem for $k=2$ is non-trivial, a~lower bound of $3$ is known~\cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)
 
@@
 We use this concept slightly differently in the lower bound and the upper bounds.
-In our algorithms, we group nodes with a communication history into \emph{components}, and in the lower bound, we group nodes that may communicate into \emph{ground sets}.
-
-A~\emph{ground set} is a group of nodes that repeatedly communicate when an algorithm splits them (places at different clusters).
+In our algorithms, we group nodes with a~communication history into \emph{components}, and in the lower bound, we group nodes that may communicate into \emph{ground sets}.
+
+A~\emph{ground set} is a~group of nodes that repeatedly communicate when an algorithm splits them (places at different clusters).
 Their role is to ensure that the algorithm maintains a~partition of ground sets into clusters.
 We emphasize that ground sets are revealed only upon their nodes are split.
-The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
-Once we construct a ground set, it lasts until the end of the algorithm's execution.
-We say that a ground set is a \emph{singleton} if it contains exactly one node.
-We say that the node that belongs to a singleton ground set is \emph{isolated}.
-
-We start by constructing a ground set of size $k-1$ on an arbitrarily chosen cluster.
+The adversary constructs ground sets depending on the choices of a~deterministic online algorithm.
+Once we construct a~ground set, it lasts until the end of the algorithm's execution.
+We say that a~ground set is a~\emph{singleton} if it contains exactly one node.
+We say that the node that belongs to a~singleton ground set is \emph{isolated}.
+
+We start by constructing a~ground set of size $k-1$ on an arbitrarily chosen cluster.
 In any partition, there must exist an isolated node collocated with the ground set of size $k-1$, and we call it \emph{special}.
@@
 	Initially, all nodes are isolated (i.e., in singleton ground sets).
-	First, we create a ground set $B$ of $k-1$ nodes on an arbitrarily chosen cluster (the nodes  are already collocated).
+	First, we create a~ground set $B$ of $k-1$ nodes on an arbitrarily chosen cluster (the nodes  are already collocated).
 	%We never dismiss any constructed ground set.
-	Each cluster hosts exactly $k$ nodes, and in any partition, a single isolated node resides in $S$.
+	Each cluster hosts exactly $k$ nodes, and in any partition, a~single isolated node resides in $S$.
 	We say that this isolated node that accompanies the ground set $B$ is \emph{special}.
@@
 
-	Then, we add the special node to a larger ground set to force its eviction.
+	Then, we add the special node to a~larger ground set to force its eviction.
 	Precisely,
-	we create a ground set $\{x_0, y_0\}$, 
+	we create a~ground set $\{x_0, y_0\}$, 
 	where $x_0$ is the current special node and $y_0$ is an arbitrary node that does not reside in $S$.
@@
 
-	We continue adding the current special node to a larger ground set.
+	We continue adding the current special node to a~larger ground set.
 	Precisely, for an isolated node $x_i$ from~$S$, we merge the ground sets of $x_i$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x_i)$.
@@
 	We claim that expanding ground sets in this way can continue as long as at least $\ell$ isolated nodes exist.
-	Fix the partition just before the creation of a ground set with $x_i$ for some~$i$.
-	We show that if at least $\ell$ isolated nodes exist (including $x_i$), then there exists a~perfect partition of ground sets after expanding a ground set with $x_i$.
-	To show this claim, we consider a partition, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
+	Fix the partition just before the creation of a~ground set with $x_i$ for some~$i$.
+	We show that if at least $\ell$ isolated nodes exist (including $x_i$), then there exists a~perfect partition of ground sets after expanding a~ground set with $x_i$.
+	To show this claim, we consider a~partition, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
 	In this partition, each cluster contains:  at most one ground set larger than $1$, possibly some isolated nodes, and possibly the ground set $\{x_0, y_0\}$.
@@
 	To find the cluster for $\{x_0, y_0\}$, observe that we have at most $\ell-1$ clusters with isolated nodes, and we have at least $\ell$ isolated nodes --- hence a~cluster with two isolated nodes exists.
-	To find a feasible partition, we swap $\{x_0, y_0\}$ with these two collocated isolated nodes, and then we swap an isolated node from $I(x_i)$ with~$x_i$.
+	To find a~feasible partition, we swap $\{x_0, y_0\}$ with these two collocated isolated nodes, and then we swap an isolated node from $I(x_i)$ with~$x_i$.
 
@@
 	Consider \ALG's partition after the last expansion.
-	Among $\ell+1$ isolated nodes there must exist a pair that originates from the same cluster (we have $\ell$ clusters). 
+	Among $\ell+1$ isolated nodes there must exist a~pair that originates from the same cluster (we have $\ell$ clusters). 
 	We denote these two nodes by $x^*$ and $y^*$.
@@
 
-	\ALG performs at least one swap for every expansion of a ground set.
+	\ALG performs at least one swap for every expansion of a~ground set.
 	Each expansion reduces the number of isolated nodes by $1$ or $2$.
@@
 \textbf{Lower bound for the general problem.}
-For a lower bound $\Omega(k\cdot \ell)$ for the general partitioning problem, the adversary continues to issue requests to split nodes of a ground set until the algorithm collocates them.
+For a~lower bound $\Omega(k\cdot \ell)$ for the general partitioning problem, the adversary continues to issue requests to split nodes of a~ground set until the algorithm collocates them.
 Note that the ground sets constructed in our lower bound can be perfectly partitioned into the clusters.
-Hence, the optimal algorithm moves to a perfect partition at the beginning (where requests incur the cost~$0$), and its cost is bounded.
+Hence, the optimal algorithm moves to a~perfect partition at the beginning (where requests incur the cost~$0$), and its cost is bounded.
 This means that the algorithm must eventually collocate all nodes of a~ground set to be competitive.
@@
 \textbf{Resource augmentation.}
-The majority of work on Online Balanced Partitioning so far \cite{repartition-disc,sigmetrics19_partitioning} focuses on the scenario with resource augmentation, where the clusters of an online algorithm are larger than the clusters of the offline optimal algorithm that we compare the performance to.
-We can adjust our construction to show a lower bound of $\Omega(\ell)$ for a setting with resource augmentation.
-
-Consider a partitioning problem with resource augmentation $1+1/3-\epsilon$.
+The majority of work on Online Balanced Partitioning so far~\cite{repartition-disc,sigmetrics19_partitioning} focuses on the scenario with resource augmentation, where the clusters of an online algorithm are larger than the clusters of the offline optimal algorithm that we compare the performance to.
+We can adjust our construction to show a~lower bound of $\Omega(\ell)$ for a~setting with resource augmentation.
+
+Consider a~partitioning problem with resource augmentation $1+1/3-\epsilon$.
 Fix $k$ divisible by $3$, and construct $3$ ground sets of size $k/3$ in each cluster.
@@
 We present an asymptotically optimal algorithm for the learning problem.
-The algorithm  collocates  a pair as soon as they communicate and it never separates them.
+The algorithm  collocates  a~pair as soon as they communicate and it never separates them.
 In order to preserve collocated pairs,
@@
 %In our algorithms, 
-A~\emph{component} is a subset of frequently communicating nodes \maciek{this is not a good definition. intuitions, maybe}.
-We keep all nodes of a component always collocated in the same cluster,
+A~\emph{component} is a~subset of frequently communicating nodes \maciek{this is not a~good definition. intuitions, maybe}.
+We keep all nodes of a~component always collocated in the same cluster,
 i.e., when we  move a~node,
 we move the whole component that contains it.
-A~partition that maintains this invariant is a \emph{component respecting} partition.
+A~partition that maintains this invariant is a~\emph{component respecting} partition.
 %We say that a~partition is \emph{component respecting}
 %if the nodes belonging to the same component are always collocated.
-We say that a component is \emph{singleton} if it contains exactly one node,
+We say that a~component is \emph{singleton} if it contains exactly one node,
 and the node in this component is \emph{isolated}.
@@
 In addition,
-we maintain a balanced partition of our components as long as such partition exists,
-a reminiscent of partitioning given integers into sets of equal sum \cite{integer-partitions-book}.
+we maintain a~balanced partition of our components as long as such partition exists,
+a reminiscent of partitioning given integers into sets of equal sum~\cite{integer-partitions-book}.
 In contrast, our partition is time-varying:  two components are merged into one component once they communicate, and we adjust the partition accordingly.
 
-The algorithm in this section and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from \cite{repartition-disc}.
-The difference is in the choice of partition after a component merge.
+The algorithm in this section and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from~\cite{repartition-disc}.
+The difference is in the choice of partition after a~component merge.
 In DET, the partition was arbitrary.
 In the algorithm for the learning model,
-we choose a component respecting partition closest to the initial partition.
+we choose a~component respecting partition closest to the initial partition.
 In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).
@@
 $P_F := \{F_1, \dots, F_{\ell}\}$.
-The \emph{distance} of a partition
+The \emph{distance} of a~partition
  $P = \{C_1, \dots, C_{\ell}\}$ from the initial partition,
@@
 With each repartitioning,
-\PPL moves to a component respecting partition that minimizes the distance to the initial partition $P_I$.
-As a result,
+\PPL moves to a~component respecting partition that minimizes the distance to the initial partition $P_I$.
+As a~result,
 \PPL never moves to a~partition that is more than $\Delta^*$  node migrations away from $P_I$.
 This invariant latter ensures us that \PPL does not pay too much while recovering $P_F$.
-We emphasize that a repartitioning by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
-This way \PPL never moves to a partition beyond the distance $\Delta^*$.
+We emphasize that a~repartitioning by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
+This way \PPL never moves to a~partition beyond the distance $\Delta^*$.
 
@@
 	That is, 
-	once it reaches a partition $P$ at distance $\Delta^* = \Delta(P)$,
-	it does not move to a partition
+	once it reaches a~partition $P$ at distance $\Delta^* = \Delta(P)$,
+	it does not move to a~partition
 	$P', \Delta(P') > \Delta^*$,
@@
 	 initially begins as an isolated node and it grows by gaining $|C|-1$ more nodes.
-	Hence, the total number of times a component in $F_i$ grows is 
+	Hence, the total number of times a~component in $F_i$ grows is 
 	$\sum_{C \in F_i} (|C|-1) \leq k-1$.
 	Therefore, there are at most  $(k-1)\cdot\ell $ distinct requests
-	for which \PPL performs a repartitioning and
+	for which \PPL performs a~repartitioning and
 	 \PPL performs at most $(k-1)\cdot\ell $ repartitions.
@@
 
-% In Section~\ref{sec:lowerbound} we constructed a $\Omega(k \cdot \ell)$ for \OBRP{}.
-% Note that the lower bound holds also in the perfect partition model, as the constructed input sequence allows \OPT to move to a perfect partition.
+% In Section~\ref{sec:lowerbound} we constructed a~$\Omega(k \cdot \ell)$ for \OBRP{}.
+% Note that the lower bound holds also in the perfect partition model, as the constructed input sequence allows \OPT to move to a~perfect partition.
 % The corollary is that PPL is optimal.
@@
 In Section~\ref{sec:k3}, we show an $O(k \cdot \ell)$-competitive algorithm for $k=3$ using the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental}.
-Prior to this section, we showed a lower bound of $\Omega(k \cdot \ell)$  that holds for the general model (cf.
+Prior to this section, we showed a~lower bound of $\Omega(k \cdot \ell)$  that holds for the general model (cf.
 Section~\ref{sec:lowerbound}), hence the result from this section is asymptotically optimal.
-Furthermore, in Section~\ref{sec:k2}, we show a strictly $6$-competitive algorithm for $k=2$.
+Furthermore, in Section~\ref{sec:k2}, we show a~strictly $6$-competitive algorithm for $k=2$.
 
@@
 	A straightforward analysis of the algorithm from Section~\ref{sec:k3} results in the bound $O(\ell^2)$, and the analysis must be improved in two places.
-One must bound the cost of each of $O(\ell)$ reconfigurations in a phase by a~constant, and show that the optimum offline algorithm must pay for a significant portion of inter-cluster requests.
+One must bound the cost of each of $O(\ell)$ reconfigurations in a~phase by a~constant, and show that the optimum offline algorithm must pay for a~significant portion of inter-cluster requests.
 	We bounded the cost of the latter by estimating the capabilities of the optimum offline algorithm to \emph{prepare} for an incoming sequence of requests.}
@@
 initially, each node is isolated (belongs to its own component).
-For each pair of nodes $\set{x,y}$, \TAlg maintains a counter $C_{\set{x,y}}$ and increments it on every external request between $x$ and $y$.
+For each pair of nodes $\set{x,y}$, \TAlg maintains a~counter $C_{\set{x,y}}$ and increments it on every external request between $x$ and $y$.
 Once $C_{\set{x,y}} = \alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
@@
 
-In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
-Before bounding the competitive ratio of \TAlg, we introduce the lemma that upper bounds the cost of a single repartition of \TAlg.
+In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a~cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
+Before bounding the competitive ratio of \TAlg, we introduce the lemma that upper bounds the cost of a~single repartition of \TAlg.
 
@@
 	\label{lem:1req}
-	In a single repartitioning of nodes (after a merge of components), \TAlg exchanges at most two pairs of nodes.
+	In a~single repartitioning of nodes (after a~merge of components), \TAlg exchanges at most two pairs of nodes.
 \end{lemma}
@@
 	%A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
-	%Finally, a cluster of type $C_3$  contains one component of size $3$.
-
-	Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
+	%Finally, a~cluster of type $C_3$  contains one component of size $3$.
+
+	Consider a~request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
 	The request triggered the repartition, hence it was external and $U\neq V$.
@@
 	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the repartitioning is local within $U$ and $V$.
-	In this case, a local repartition is possible (within $U$ and $V$), for a cost of at most $2$ swaps.
-	If either $U$ or $V$ is of type $C_3$, a component of size $3$ participates in a merge, and we have a component of size at least $4$, and \TAlg ends the phase with no repartitioning.
+	In this case, a~local repartition is possible (within $U$ and $V$), for a~cost of at most $2$ swaps.
+	If either $U$ or $V$ is of type $C_3$, a~component of size $3$ participates in a~merge, and we have a~component of size at least $4$, and \TAlg ends the phase with no repartitioning.
 
@@
 	If $(u,v)$ both belong to components of size $2$, then the merged component has size $4$, and \TAlg incurs no cost. 
-	Otherwise, if one of $u,v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
-
-	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
+	Otherwise, if one of $u,v$ belongs to a~component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
+
+	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a~cluster different from $U$ and $V$.
 	Note that if $C_1$-type cluster does not exist, then no component respecting partitioning exists.
@@
 
-	In each case, we showed that a partition is reachable in at most two swaps.
+	In each case, we showed that a~partition is reachable in at most two swaps.
 \end{proof}
@@
 	\includegraphics[width=0.3\textwidth]{figs/substitute}
-	\caption{an unsaturated edge that is inside a component. If there is a space next to it, maybe cluster types should fit next to it? Not sure.}
+	\caption{an unsaturated edge that is inside a~component. If there is a~space next to it, maybe cluster types should fit next to it? Not sure.}
 \end{figure}
@@
 \begin{proof}
-	Fix a completed phase, and consider the state of \TAlg's counters at the end of it (before the reset).
+	Fix a~completed phase, and consider the state of \TAlg's counters at the end of it (before the reset).
 	We consider the incomplete phase later in this proof.
@@
 	As \TAlg is component respecting, it never increases any counter above $\alpha$.
-	We say that the pair $(u, v)$ is \emph{saturated} if the counter's value is $\alpha$, and \emph{unsaturated} otherwise (saturation of a pair leads to a merge action).
+	We say that the pair $(u, v)$ is \emph{saturated} if the counter's value is $\alpha$, and \emph{unsaturated} otherwise (saturation of a~pair leads to a~merge action).
 	By $\sigma$ we denote the input sequence that arrived during the phase.
@@
 	exceeding this number would mean that a~component of size $4$ exists, and the phase should have ended already.
-	We bound the cost of each repartitioning after a merge by Lemma~\ref{lem:1req}, obtaining $\TAlg(M) \leq 8\alpha\cdot\ell$.
+	We bound the cost of each repartitioning after a~merge by Lemma~\ref{lem:1req}, obtaining $\TAlg(M) \leq 8\alpha\cdot\ell$.
 	%Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
@@
 	We have $\OPT(M) + \OPT(\sigma_I) \geq \alpha$, as the phase ended when the components of \TAlg{} could not be partitioned without splitting them.
-	Hence, for every possible partition of $\OPT$, there exists a non-collocated pair of nodes with at least $\alpha$ requests between them, and
+	Hence, for every possible partition of $\OPT$, there exists a~non-collocated pair of nodes with at least $\alpha$ requests between them, and
 	$\OPT$ either served them remotely or performed a~migration.
@@
 	Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
-	Fix a partition for \OPT prior serving the sequence $\sigma_E$.
+	Fix a~partition for \OPT prior serving the sequence $\sigma_E$.
 	Even without migrations, \OPT may mitigate paying for requests between at most $3\ell$ pairs of nodes that can be collocated in its clusters.
@@
 	By migrating nodes, \OPT may decrease its cost, and pay less than $\chi$ for requests from $\sigma_E$.
-	By swapping a pair of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
+	By swapping a~pair of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
 	This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterward.
 	As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
-	By performing a single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$
+	By performing a~single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$
 	Hence, for serving at least $\chi$ unavoidable requests from $\sigma_E$, $\OPT$ pays at least
@@
 	If the input finished before the first $\alpha$ external requests, then \TAlg is $1$-competitive.
-	If at least $\alpha$ external requests were issued, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a migration; for the cost of \TAlg we follow the analysis of the unsaturated pairs.
+	If at least $\alpha$ external requests were issued, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a~migration; for the cost of \TAlg we follow the analysis of the unsaturated pairs.
 	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
-	This way, the competitive ratio increases at most twofold in comparison to a finished phase, and the competitive ratio is $\TAlg(\sigma)/\OPT(\sigma) \leq 60\ell$.
+	This way, the competitive ratio increases at most twofold in comparison to a~finished phase, and the competitive ratio is $\TAlg(\sigma)/\OPT(\sigma) \leq 60\ell$.
 \end{proof}
@@
 While we have described the algorithm globally so far, we note that it allows for an efficient distributed implementations. 
-The algorithm performs two types of operations that require communication with other clusters: a~component merge, and a broadcast of the end of the phase.
-We say that a cluster containing $3$ isolated nodes is \emph{fresh}.
-A merge of two components may require finding a fresh cluster (for details see the proof of Lemma~\ref{lem:1req}).
-In the following, we show how to efficiently find a fresh cluster in a distributed manner.
+The algorithm performs two types of operations that require communication with other clusters: a~component merge, and a~broadcast of the end of the phase.
+We say that a~cluster containing $3$ isolated nodes is \emph{fresh}.
+A merge of two components may require finding a~fresh cluster (for details see the proof of Lemma~\ref{lem:1req}).
+In the following, we show how to efficiently find a~fresh cluster in a~distributed manner.
 We organize the clusters into an arbitrary rooted balanced binary tree, and we broadcast the root to each cluster.
 Each cluster maintains the counter of fresh clusters in its subtree.
-To find a fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
-Upon encountering a fresh cluster, we end the traversal and decrease the counters on the followed path by $1$.
-Summarizing, ending the phase requires a single broadcast, and merging components has $O(\log \ell)$ communication complexity.
+To find a~fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
+Upon encountering a~fresh cluster, we end the traversal and decrease the counters on the followed path by $1$.
+Summarizing, ending the phase requires a~single broadcast, and merging components has $O(\log \ell)$ communication complexity.
 
@@
  an algorithm for the \OBRP restricted to clusters of capacity $k=2$.
-We interpret a pair of nodes collocated in one cluster as a matched pair.
+We interpret a~pair of nodes collocated in one cluster as a~matched pair.
 Hence,
@@
 This is known as the  \emph{Online Rematching} 
-problem and a (non-strict) 7-competitive algorithm is already given by \cite{repartition-disc},
+problem and a~(non-strict) 7-competitive algorithm is already given by~\cite{repartition-disc},
 in which the ratio comes with an additive factor in $O(\alpha\ell^2)$.
@@
 our algorithm is similar
-and slightly simpler than the one in \cite{repartition-disc}, 
+and slightly simpler than the one in~\cite{repartition-disc}, 
 while our analysis is significantly more simple and concise,
@@
 \textbf{Algorithm ReMatch.} 
-The algorithm ReMatch (\RM) maintains a counter $C_{\set{x,y}}$ for each pair of nodes $\set{x,y}$ and increments it on every remote request between $x$ and $y$.
+The algorithm ReMatch (\RM) maintains a~counter $C_{\set{x,y}}$ for each pair of nodes $\set{x,y}$ and increments it on every remote request between $x$ and $y$.
 Once $C_{\set{x,y}} = \lambda$,
@@
 \textbf{The charging scheme.}
-We charge both \OPT and \RM whenever \RM collocates a pair.%(i.e., on a collocation event).
-\RM collocates a pair always with a swap,%(i.e., two simultaneous migrations),
+We charge both \OPT and \RM whenever \RM collocates a~pair.%(i.e., on a~collocation event).
+\RM collocates a~pair always with a~swap,%(i.e., two simultaneous migrations),
 which  costs $2\alpha$,
@@
 Hence,
-whenever \OPT collocates a pair,
-we charge it only the cost of moving a single node to the cluster of the other,
+whenever \OPT collocates a~pair,
+we charge it only the cost of moving a~single node to the cluster of the other,
 i.e., $\alpha$ (in contrast to $2\alpha$ on \RM).
@@
 %i.e.~the entire duration of the sequence.
-Observe that \OPT must pay a non-zero cost
+Observe that \OPT must pay a~non-zero cost
 for these requests over the entire $\sigma$,
@@
 However,
-we can charge the whole cost to \OPT only the first time \RM collocates a pair,
-and not at any consequent time when \RM collocates it a second time.
+we can charge the whole cost to \OPT only the first time \RM collocates a~pair,
+and not at any consequent time when \RM collocates it a~second time.
 Otherwise,
@@
 For this reason,
-we charge \OPT a cost inflicted by a pair,
+we charge \OPT a~cost inflicted by a~pair,
 if and only if  \OPT incurs that cost after the latest separation of the pair by \RM.
@@
 	$ C^t_{\set{u,v}} = \lambda$.
-	We denote the set of all requests to a pair $\set{x,y}$ that arrive
+	We denote the set of all requests to a~pair $\set{x,y}$ that arrive
 	during $[t_1,t_2]$ by $\sigma_{\set{x,y}}[t_1,t_2]$,
@@
 	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$
-	is the set of times when \RM performs a collocation.
+	is the set of times when \RM performs a~collocation.
 	By definitions of $t$ and $t'$,
@@
 %We briefly sketch an approximation for the offline problem,
-%that makes use of \RM as a subroutine.
+%that makes use of \RM as a~subroutine.
 %In our approximation algorithm,
@@
 %Whenever,
-%\RM collocoates a pair
+%\RM collocoates a~pair
 
@@
 
-This paper revisited the dynamic graph partitioning problem and presented several tight bounds for the important model where capacities cannot be exceeded, both for a general partitioning model and for a special learning model.
+This paper revisited the dynamic graph partitioning problem and presented several tight bounds for the important model where capacities cannot be exceeded, both for a~general partitioning model and for a~special learning model.
 
@@
 		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
-		%        \Ensure A final partition $P_F$ 
-		\STATE {For each node $v$ create a singleton component $C_v = \{ v \}$ and add it to $\mathcal{C}$.}
+		%        \Ensure a~final partition $P_F$ 
+		\STATE {For each node $v$ create a~singleton component $C_v = \{ v \}$ and add it to $\mathcal{C}$.}
 		%\STATE{$P := P_I$}
@@
 		\IF{$C_1$ and $C_2$ are not collocated}
-		\STATE {Move to a partition closest to $P_I$ and respecting all components in $\mathcal{C}$.} 
-		%		\COMMENT{move to a partition closest to $P_I$}
+		\STATE {Move to a~partition closest to $P_I$ and respecting all components in $\mathcal{C}$.} 
+		%		\COMMENT{move to a~partition closest to $P_I$}
 		\label{line:rebalance} 

### 2020-05-12 paper/disc20clustering.tex
Intro to ALG3

@@
 A straightforward analysis of the algorithm from Section~\ref{sec:k3} results in the bound $O(\ell^2)$, and the analysis must be improved in two places.
-One must bound the cost of each of $O(\ell)$ reconfigurations in~phase by a~constant, and show that the optimum offline algorithm must pay for a~significant portion of inter-cluster requests \maciek{how to concisely say that there is also a~migrate possibilty}.
+One must bound the cost of each of $O(\ell)$ reconfigurations in~phase by a~constant, and show that the optimum offline algorithm must pay significant cost for inter-cluster requests.
 We bound the cost of the latter by estimating the capabilities of the optimum offline algorithm to \emph{prepare} for an incoming sequence of requests.
@@
 In contrast, our partition is time-varying:  two components are merged into one component once they communicate, and we adjust the partition accordingly.
+
+\medskip
 
@@
 
-\maciek{Intro is missing.}
-\maciek{
-	A straightforward analysis of the algorithm from Section~\ref{sec:k3} results in the bound $O(\ell^2)$, and the analysis must be improved in two places.
-One must bound the cost of each of $O(\ell)$ reconfigurations in a~phase by a~constant, and show that the optimum offline algorithm must pay for a~significant portion of inter-cluster requests.
-	We bounded the cost of the latter by estimating the capabilities of the optimum offline algorithm to \emph{prepare} for an incoming sequence of requests.}
+The proposed algorithm is a modified version of the algorithm DET proposed by Avin et al.~\cite{repartition-disc}, that for $k=3$ is $O(\ell^2)$-competitive.
+In our algorithm, we choose the closest partition after a component merge instead of an arbitrary one.
+This allows to bound the cost of repartition by a constant (Lemma~\ref{lem:1req}).
+
+This modification alone is insufficient to obtain $O(\ell)$-competitive algorithm, and the analysis must be further improved.
+In particular, pairs of nodes that did not reach the collocation threshold $\alpha$ (called external requests) incur the cost $O(\ell^2)$ for the algorithm in each phase.
+The novel part of the analysis lower-bounds the cost of \OPT on external requests, while considering its savings from migrations and possibly different configuration at the beginning of the phase.
+This way, we show that \OPT paid a significant portion of the algorithm's cost on external requests.
+
+\medskip
 
@@
 
-In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a~cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
+In our analysis, we distinguish among three types of clusters: $C_1, C_2$ and $C_3$. In a~cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
 Before bounding the competitive ratio of \TAlg, we introduce the lemma that upper bounds the cost of a~single repartition of \TAlg.

### 2020-05-12 paper/disc20clustering.tex
Intro to ALG3

@@
 
+\begin{theorem}
+	\TAlg is $60\ell$-competitive for $k=3$.
+	\label{thm:k=3}
+\end{theorem}
+
 %\begin{figure}[H]
@@
 
+Before bounding the competitive ratio of \TAlg, we introduce the lemma that upper bounds the cost of a~single repartition of \TAlg.
 In our analysis, we distinguish among three types of clusters: $C_1, C_2$ and $C_3$. In a~cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
-Before bounding the competitive ratio of \TAlg, we introduce the lemma that upper bounds the cost of a~single repartition of \TAlg.
 
@@
 
-\begin{theorem}
-	\TAlg is $60\ell$-competitive for $k=3$.
-\end{theorem}
-\begin{proof}
+\begin{proof}[Proof of Theorem \ref{thm:k=3}]
 	Fix a~completed phase, and consider the state of \TAlg's counters at the end of it (before the reset).
@@
 
-\medskip
+\begin{theorem} \label{thm:k=2}
+	For $\lambda=\alpha$, the algorithm \RM  is strictly 6-competitive.
+\end{theorem}
 
@@
 
-\begin{theorem} \label{thm:k=2}
-	For $\lambda=\alpha$, the algorithm \RM  is strictly 6-competitive.
-\end{theorem}
-
-\begin{proof}%[Proof of Theorem \ref{thm:k=2}]
+\begin{proof}[Proof of Theorem \ref{thm:k=2}]
 	Fix an input sequence of requests

### 2020-05-12 paper/disc20clustering.tex
wip

@@
 
-Before bounding the competitive ratio of \TAlg, we introduce the lemma that upper bounds the cost of a~single repartition of \TAlg.
-In our analysis, we distinguish among three types of clusters: $C_1, C_2$ and $C_3$. In a~cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
+Before bounding the competitive ratio of \TAlg, we upper-bound the cost of a~single repartition of \TAlg.
+In our analysis, we distinguish among three types of clusters: $C_1, C_2$ and $C_3$. In a~cluster of type $C_i$, the size of the largest component contained in this cluster is~$i$.
 

### 2020-05-12 paper/disc20clustering.tex
Intro to ALG3

@@
 
-The proposed algorithm is a modified version of the algorithm DET proposed by Avin et al.~\cite{repartition-disc}, that for $k=3$ is $O(\ell^2)$-competitive.
+The algorithm analyzed in this section is a modified version of the algorithm DET proposed by Avin et al.~\cite{repartition-disc}, nthat for $k=3$ is $O(\ell^2)$-competitive.
 In our algorithm, we choose the closest partition after a component merge instead of an arbitrary one.

### 2020-05-12 paper/disc20clustering.tex
Intro to ALG3

--- 2020-05-12 paper/disc20clustering.tex
@@
 
-The proposed algorithm is a modified version of the algorithm DET proposed by Avin et al.~\cite{repartition-disc}, that for $k=3$ is $O(\ell^2)$-competitive.
+The algorithm analyzed in this section is a modified version of the algorithm DET proposed by Avin et al.~\cite{repartition-disc}, that for $k=3$ is $O(\ell^2)$-competitive.
 In our algorithm, we choose the closest partition after a component merge instead of an arbitrary one.

### 2020-05-12 paper/disc20clustering.tex
font issues with \OPT \RM

@@
  \newcommand{\OPT}{\textsc{OPT}\xspace}
-  \newcommand{\OPTM}{\mathit{OPT}}
  \newcommand{\ALG}{\textsc{ALG}\xspace}
@@
  \newcommand{\PPOBRP}{PP-BRP}
- \newcommand{\dist}{\mathit{dist}}
+ \newcommand{\dist}{\textsc{dist}}
  \newcommand{\TAlg}{{\ensuremath{\textsf{ALG}_{3}}}\xspace}
@@
 	and its overall cost to \OPT by
-	$\mathit{OPT} (\sigma_{\{u,v\}}[t_1,t_2])$.
+	$\OPT (\sigma_{\{u,v\}}[t_1,t_2])$.
 	We may use $\sigma_{\set{x,y}}$ whenever
@@
     Then the overall cost to \RM is
-$	\mathit{RM} (\sigma)
+$	\RM (\sigma)
 =
@@
 	Therefore in this case,
-	$\OPTM (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
+	$\OPT(\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
 	%
@@
 	Then,
-	$\OPTM(R_t) = \mathit{OPT} (\sigma_{\{u,w\}}) 
-	+ \OPTM(\sigma_{\{u,v\}}) $.
+	$\OPT(R_t) =\OPT (\sigma_{\{u,w\}}) 
+	+ \OPT(\sigma_{\{u,v\}}) $.
 	If \OPT has both pairs separated during their respective intervals,
@@
 	Therefore in any case,
-	$\OPTM(R_t) \geq  \min{\{ \lambda,\alpha \}}$.
+	$\OPT(R_t) \geq  \min{\{ \lambda,\alpha \}}$.
 
@@
 	\OPT pays
-	$\mathit{OPT} (\sigma_{\{u,v\}}[1,m]) 
+	$\OPT(\sigma_{\{u,v\}}[1,m]) 
 	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
@@
 	Thus,
-	$\mathit{OPT} (R^*_{\{u,v\}}) 
+	$\OPT (R^*_{\{u,v\}}) 
 	\geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.
@@
 	This means,
-	$\OPTM ( \sigma_{\{u,w\}})$
+	$\OPT ( \sigma_{\{u,w\}})$
 	is counted at most twice in each of  the expressions
-	$\mathit{OPT} (R_{t'}) + \mathit{OPT} (R_t)$
+	$\OPT (R_{t'}) + \OPT (R_t)$
 	and  
-	$\mathit{OPT} (R_{t'}) + \mathit{OPT} (R^*_{\{u,v\}})$.
+	$\OPT (R_{t'}) + \OPT(R^*_{\{u,v\}})$.
 	Hence,
@@
 	$\frac{1}{2}(
-	\sum_{ t \in T } \mathit{OPT} (R_t) +
-	\sum_{\{u,v\}} \mathit{OPT} (R^*_{\{u,v\}})
+	\sum_{ t \in T } \OPT (R_t) +
+	\sum_{\{u,v\}}\OPT (R^*_{\{u,v\}})
 	) $.
@@
 	\begin{align*} 	%\label{eq:costOPT}
-		\mathit{OPT} (\sigma)
+		\OPT (\sigma)
 		&=
@@
 		\Big(
-		\sum_{ t \in T} \mathit{OPT} (R_t) 
-		+ \sum_{\{u,v\}}\mathit{OPT} (R^*_{\{u,v\}})
+		\sum_{ t \in T} \OPT (R_t) 
+		+ \sum_{\{u,v\}}\OPT (R^*_{\{u,v\}})
 		\Big)	\\
@@
 $
-	\mathit{ALG} (\sigma)	/
-	\mathit{OPT} (\sigma)
+	\RM(\sigma)	/
+	\OPT (\sigma)
 	\leq
@@
 			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$.} \label{line:mergecomponents}
-		%		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
-		%		\COMMENT{i.e.~if not in the same cluster.}}       
 		\IF{$C_1$ and $C_2$ are not collocated}

### 2020-05-12 paper/disc20clustering.tex
shortended formula

@@
 	Therefore in any case,
-	$\OPT(R_t) \geq  \min{\{ \lambda,\alpha \}}$.
+	$\OPT(R_t) \geq  \min{\{ \lambda,\alpha \}} = \alpha$.
 
@@
 	the total cost to \OPT is
-
 	\begin{align*} 	%\label{eq:costOPT}
@@
 		+ \sum_{\{u,v\}}\OPT (R^*_{\{u,v\}})
-		\Big)	\\
-		&\geq
-		\Big(
-		\sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  +
-		\sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \} } 
 		\Big)	
-		=
+		\geq
+%		\Big(
+%		\sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  +
+%		\sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \} } 
+%		\Big)		
+%		=
 		\frac{1}{2}		
@@
 \end{proof}
-
-\maciek{OPT's font differ from other proofs. We don't want italic OPT in mathenv}
-
-\maciek{I really think that minimums should be expanded earlier than in the competitive ratio.
-The ratio should be bounded easily, and if you wish, you can emphasize the choice in the paragraph, or dedicate entire line to the cost expression.
-}
 

### 2020-05-12 paper/disc20clustering.tex
changing the LB

@@
 
-A~\emph{ground set} is a~group of nodes that repeatedly communicate when an algorithm splits them (places at different clusters).
-Their role is to ensure that the algorithm maintains a~partition of ground sets into clusters.
-We emphasize that ground sets are revealed only upon their nodes are split.
-The adversary constructs ground sets depending on the choices of a~deterministic online algorithm.
-Once we construct a~ground set, it lasts until the end of the algorithm's execution.
-We say that a~ground set is a~\emph{singleton} if it contains exactly one node.
-We say that the node that belongs to a~singleton ground set is \emph{isolated}.
+A~\emph{ground set} is a~subset of nodes such that if an algorithm does not have them all collocated,
+then the split parts communicate indefinitely until the algorithm collocates them.
+%A~\emph{ground set} is a~group of nodes that  communicate indefinitely when an algorithm splits them (places at different clusters).
+This ensure us that any competitive algorithm would maintain a~partition of ground sets into clusters.
+As a result,
+ a ground set is revealed to an algorithm only after the algorithm splits it.
+ More precisely,
+the adversary constructs ground sets depending on the choices of a~given deterministic online algorithm.
+%Once we construct a~ground set, it lasts until the algorithm terminates.
+We say that a~ground set is \emph{singleton} if it contains only one node,
+and we refer to such node as an \emph{isolated} node.
 
@@
 We issue requests between the current special node, and some node that was initially collocated with it.
-Repeating such requests, almost all the nodes must visit the cluster $S$.
+By repeating such requests, almost every node is collocated with the first ground set at some point .
 In comparison, we show that there exists an optimal offline algorithm \OPT that performs only two migrations.
+\mahmoud{4 migrations?}
 Figure~\ref{fig:lb} illustrates the constructs used in the lower bound.
@@
 \begin{proof}
-	Fix any online algorithm \ALG{}.
-	Let $I(C)$ denote the cluster where nodes of the ground set $C$ are located in the initial partition.
-	If $C$ contains nodes that originate from different clusters, $I(C)$ is undefined.
+	Fix any online algorithm \ALG.
+	For a ground set $C$ of nodes that are initially collocated in one cluster,
+	let $I(C)$ denote the cluster.
+%	If $C$ contains nodes that originate from different clusters
+%	then $I(C)$ is undefined.
 	Throughout the construction, we maintain ground sets for each node.
 	Initially, all nodes are isolated (i.e., in singleton ground sets).
-	First, we create a~ground set $B$ of $k-1$ nodes on an arbitrarily chosen cluster (the nodes  are already collocated).
+	First, we choose a cluster arbitrarily and
+	 create a~ground set $B$ of $k-1$ nodes in this cluster.
 	%We never dismiss any constructed ground set.
-	Each cluster hosts exactly $k$ nodes, and in any partition, a~single isolated node resides in $S$.
+	Each cluster hosts exactly $k$ nodes, and in any partition, a~single isolated node must be collocated with $B$.
 	We say that this isolated node that accompanies the ground set $B$ is \emph{special}.
@@
 
-	Then, we add the special node to a~larger ground set to force its eviction.
+	Then, we join the special node to a~larger ground set to force its eviction.
 	Precisely,
 	we create a~ground set $\{x_0, y_0\}$, 
-	where $x_0$ is the current special node and $y_0$ is an arbitrary node that does not reside in $S$.
-	\ALG has them split, and the adversary issues external requests until \ALG collocates them.
-	There is no space to accommodate $\{x_0, y_0\}$ in the cluster of $x_0$ (as $B$'s size is $k-1$), hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.
-	To preserve a~feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
-	\ALG must replace $x_0$ with another single node that we call $x_1$.
-
-	We continue adding the current special node to a~larger ground set.
-	Precisely, for an isolated node $x_i$ from~$S$, we merge the ground sets of $x_i$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x_i)$.
-	This way, $x_i$ is evicted similarly to $x_0$, and we continue creating such ground sets.
-	We describe the termination condition shortly.
-	(Note that these ground sets contain nodes that originate from the same cluster, whereas the nodes $x_0$ and $y_0$ originate from distinct clusters.)
+	where $x_0$ is the current special node and $y_0$ is an arbitrary isolated node.
+	Since \ALG does not have $\{x_0,y_0\}$ collocated, 
+	the adversary issues external requests between them until \ALG collocates them.
+	\ALG cannot collocate $\{x_0, y_0\}$ with $B$ (as $B$'s size is $k-1$), hence it collocates them in a different cluster.
+	In order to preserve a~feasible partition of nodes after collocating $\{x_0, y_0\}$,
+	\ALG must replace $x_0$ with another isolated node,
+	denoted by $x_1$.
+
+	We continue joining the current special node to a~larger ground set.
+	Precisely, for the isolated node $x_i$ currently collocated with $B$,
+	 we issue a request between $x_i$ and a node $y_i$
+	 in the largest ground set $C$
+	 s.t.~$I(C) = I(x_i), C \neq \{x_0,y_0\}$.
+
+\mahmoud{
+	If a feasible partition exists 
+	that collocates the new ground set $\{x_i,\} \cup C $,  
+	then $x_i$ is replaced by some isolated node
+	 $x_{i+1} \not\in \{y_0, x_0, \dots, x_i\}$.
+%	 similarly to $x_0$, and we continue creating such ground sets.
+%	We describe the termination condition shortly.
+	Otherwise,
+	the current partition is not feasible,
+	 we terminate the process and we let $x^* := x_i, y^*:=y_i$. 
+	 At this point,
+	 there are at most $\ell$ isolated nodes left.
+	 Assume for contradiction that this is not the case,
+	 i.e., there are at least $\ell+1$ isolated nodes left.
+	 Then there must be two isolated nodes $u$ and $v$ with the same cluster of origin,
+	 i.e., $I(\{u\}) = I(\{v\})$.
+	 Consider the partition $P$,
+	  obtained from the initial partition after swapping $x_0$ and $y_0$
+	 with $u$ and $v$ (respectively).
+	 In this partition,
+	 the ground set $\{x_0,y_0\}$ is collocated in the cluster $I(\{u,v\})$.
+	 Note that after the first request $\{x_0,y_0\}$,
+	   we issue requests only between nodes that have the same cluster of origin and
+	   all these nodes are collocated in $P$.
+	 Therefore $P$  has all ground sets constructed so far collocated, i.e.,
+	 it is a feasible partition, which contradicts our assumption.
+}
 
@@
 	Fix the partition just before the creation of a~ground set with $x_i$ for some~$i$.
-	We show that if at least $\ell$ isolated nodes exist (including $x_i$), then there exists a~perfect partition of ground sets after expanding a~ground set with $x_i$.
-	To show this claim, we consider a~partition, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$).
+	We show that if at least $\ell$ isolated nodes exist (including $x_i$), then there exists a~perfect partition of ground sets after joining $x_i$ to a~ground.
+	To show this claim, we consider a~partition, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$). \mahmoud{this sentence doesn't make sense.}
 	In this partition, each cluster contains:  at most one ground set larger than $1$, possibly some isolated nodes, and possibly the ground set $\{x_0, y_0\}$.

### 2020-05-13 paper/disc20clustering.tex
small

@@
 	that collocates the new ground set $\{x_i,\} \cup C $,  
-	then $x_i$ is replaced by some isolated node
-	 $x_{i+1} \not\in \{y_0, x_0, \dots, x_i\}$.
+	then $x_i$ is replaced by some isolated node $x_{i+1}$.
 %	 similarly to $x_0$, and we continue creating such ground sets.

### 2020-05-13 paper/disc20clustering.tex
Comments

--- 2020-05-12 paper/disc20clustering.tex
@@
 	Our first contribution is a~significantly improved
-	lower bound of $\Omega(k\cdot \ell)$ on the
+	\maciek{How to put here the information that it is for deterministic algorithms only? This sentence is already very long.}lower bound of $\Omega(k\cdot \ell)$ on the
 	competitive ratio, where $\ell$ is the number

### 2020-05-13 paper/disc20clustering.tex
comments

@@
 By repeating such requests, almost every node is collocated with the first ground set at some point .
-In comparison, we show that there exists an optimal offline algorithm \OPT that performs only two migrations.
-\mahmoud{4 migrations?}
+In comparison, we show that there exists an optimal offline algorithm \OPT that performs only two node exchanges ("swaps").
 Figure~\ref{fig:lb} illustrates the constructs used in the lower bound.

### 2020-05-13 paper/disc20clustering.tex
Classification
rewriting the LB

@@
 
+ \ccsdesc[500]{Theory of computation~Online algorithms}
  \ccsdesc[500]{Networks~Network algorithms}
- \ccsdesc[300]{Computer systems organization~Cloud computing}
+ %\ccsdesc[300]{Computer systems organization~Cloud computing}
  \ccsdesc[300]{Computer systems organization~Distributed architectures}
- \ccsdesc[500]{Theory of computation~Online algorithms}
 
@@
 	\label{th:lowerbound}
-	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least $(k\cdot \ell - k - 2\ell)/2$ for any $k\geq 3$ and $\ell \geq 2$.
+	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least
+	 $(k-2)(\ell-1)/2 - 2$ for any $k\geq 3$ and $\ell \geq 2$.
 \end{theorem}
@@
 	Since \ALG does not have $\{x_0,y_0\}$ collocated, 
-	the adversary issues external requests between them until \ALG collocates them.
+	we issue an external request between them so that \ALG collocates them.
 	\ALG cannot collocate $\{x_0, y_0\}$ with $B$ (as $B$'s size is $k-1$), hence it collocates them in a different cluster.
 	In order to preserve a~feasible partition of nodes after collocating $\{x_0, y_0\}$,
-	\ALG must replace $x_0$ with another isolated node,
-	denoted by $x_1$.
-
-	We continue joining the current special node to a~larger ground set.
-	Precisely, for the isolated node $x_i$ currently collocated with $B$,
-	 we issue a request between $x_i$ and a node $y_i$
-	 in the largest ground set $C$
-	 s.t.~$I(C) = I(x_i), C \neq \{x_0,y_0\}$.
-
-\mahmoud{
-	If a feasible partition exists 
-	that collocates the new ground set $\{x_i,\} \cup C $,  
-	then $x_i$ is replaced by some isolated node $x_{i+1}$.
-%	 similarly to $x_0$, and we continue creating such ground sets.
-%	We describe the termination condition shortly.
+	\ALG must replace $x_0$ with another isolated node.
+%	denoted by $x_1$.
+
+	We proceed by joining the current special node to a~larger ground set,
+	as long as a feasible partition exists afterwards.
+	Consider the isolated node $x_i$ currently collocated with $B$.
+	 We issue a request between $x_i$ and some node $y_i \in C_i$,
+	 where $C_i$ is the largest ground set
+	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$,
+	 only if  a feasible partition exists
+	 that collocates all nodes in $\{x_i\} \cup C_i $.
+	Any such feasible partition replaces $x_i$ with some isolated node $x_{i+1}$,
+	as the new ground set $\{x_i\} \cup C_i $  may not be ever split.
 	Otherwise,
-	the current partition is not feasible,
-	 we terminate the process and we let $x^* := x_i, y^*:=y_i$. 
-	 At this point,
-	 there are at most $\ell$ isolated nodes left.
+	if no such feasible partition exists,
+	 then we terminate the process and
+	 we let $x^* := x_i, y^*:=y_i$.
+
+	 Once the process terminates,
+	 there are at most $\ell$ isolated nodes left 
+	 that are not contained in $\{x^*,y^*\}$.
 	 Assume for contradiction that this is not the case,
-	 i.e., there are at least $\ell+1$ isolated nodes left.
-	 Then there must be two isolated nodes $u$ and $v$ with the same cluster of origin,
+	 i.e., there are at least $\ell+1$ isolated nodes left
+	 in addition to $x^*$ and possibly $y^*$.
+	 Then there must be two isolated nodes $u$ and $v$,
+	 $u,v \not\in \{x^*, y^*\}$,
+	  with the same cluster of origin,
 	 i.e., $I(\{u\}) = I(\{v\})$.
-	 Consider the partition $P$,
+	 Consider the partition $P$
 	  obtained from the initial partition after swapping $x_0$ and $y_0$
@@
 	   all these nodes are collocated in $P$.
-	 Therefore $P$  has all ground sets constructed so far collocated, i.e.,
-	 it is a feasible partition, which contradicts our assumption.
-}
-
+	 Therefore all ground sets constructed so far are collocated in  $P$,
+	 i.e., $P$ is a feasible partition,
+	 which contradicts our assumption.
+	 Hence, at the point of termination,
+	 there are at most $\ell$ isolated nodes not in $\{x^*,y^*\}$,
+	 or at most $\ell+2$ isolated nodes overall.
+
+\begin{comment}
 	We claim that expanding ground sets in this way can continue as long as at least $\ell$ isolated nodes exist.
@@
 	Therefore $\{x_0, y_0\}$ is present on $I(x_i)$, and we must evict it to make space for $x_i$.
+	\mahmoud{why  $\{x_0, y_0\}$ must be present there?}
 	To find the cluster for $\{x_0, y_0\}$, observe that we have at most $\ell-1$ clusters with isolated nodes, and we have at least $\ell$ isolated nodes --- hence a~cluster with two isolated nodes exists.
 	To find a~feasible partition, we swap $\{x_0, y_0\}$ with these two collocated isolated nodes, and then we swap an isolated node from $I(x_i)$ with~$x_i$.
+	\mahmoud{but didn't you assume there is no isolated on $I(x_i)$?}
 
 	To keep \OPT's cost low, we terminate prematurely --- we stop extending ground sets while there are still $\ell+1$ isolated nodes (i.e., neglecting two last ground set expansions).
+	\mahmoud{here you are making an exception to the already established process.
+Instead, adapt the termination condition so that it triggers prematurely, by saying something like "issue the last request only if there will be a feasible partition." }
 	Consider \ALG's partition after the last expansion.
@@
 	We denote these two nodes by $x^*$ and $y^*$.
-
+\end{comment}
 	The optimal strategy is to perform two node swaps.
-	Precisely, \OPT collocates $\set{x_0,y_0}$ by swapping them with $x^*$ and $y^*$.
+	Precisely, \OPT collocates $\set{x_0,y_0}$ 
+	by swapping them with $x^*$ and $y^*$.
 	In this partition, no ground set is split.
-	We never issue requests between nodes $x^*, y^*$ that are separated in this partition.
-
-	\ALG performs at least one swap for every expansion of a~ground set.
+	Due to the termination condition,
+	we never issue requests between nodes $x^*, y^*$
+	 which are separated in this partition.
+
+	\ALG performs at least one swap each time a~ground set grows,
+	i.e., each time it joins the isolated node $x_i$ (collocated with $B$), to  the ground set $C_i$ as specified before.
+%	In other words,
+%	some ground set grows by one node at each of these steps.
+	Consider any ground set $C^*$ at the point of termination.
+	This ground set has grown exactly $|C^*|-1$ times until the termination. 
+	Let $\mathcal{C}^*$ be the set of all ground sets at the point of termination.
+	Thus, $\mathcal{C}^*$ includes ground sets $B$,
+	$\{x_0,y_0\}$,
+	and up to $\ell+2$ singleton ground sets.
+%	(including $x^*$ and possibly $y^*$).	
+	Each time we join the special node to the a largest ground set.
+	Hence, every  other ground set in $\mathcal{C}^*$ has grown maximally.
+	There are at most $\ell-1$ such ground sets in addition to the ones 
+	we considered so far,
+	one per cluster except the cluster containing $B$. 
+	Therefore,  $|\mathcal{C}^*| \leq 1 + 1 + (\ell+2)  + (\ell-1) = 2\ell+3$.
+	Note that among all non-singleton ground sets in $\mathcal{C}^*$,
+	 only $B$ does not grow during the process.
+	 Thus,
+    the total number of times that 
+	a ground set in $\mathcal{C}^*$  has grown is
+	\begin{align*}
+	\sum_{C^* \in \mathcal{C}^* } (|C^*|-1) - (k-1) &= 
+	\sum_{C^* \in \mathcal{C}^* } |C^*| - 
+		\sum_{C^* \in \mathcal{C}^* } 1 - (k-1) \\
+	&\geq  k  \ell - (2\ell+3) - (k-1) =  (k-2)(\ell-1) - 4,
+	\end{align*}
+	which is the least number of swaps performed by \ALG.
+	The competitive ratio is then
+	$\ALG / \OPT \geq ((k-2)(\ell-1) - 4) / 2$.
+	%%
+	\begin{comment}
+	\ALG performs at least one swap  every time   a~ground set is constructed.
 	Each expansion reduces the number of isolated nodes by $1$ or $2$.
@@
 	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - 2\ell) / 2$.
+	\end{comment}
+	%%
 \end{proof}

### 2020-05-13 paper/disc20clustering.tex
rewriting the LB (1)

@@
 %	then $I(C)$ is undefined.
-	Throughout the construction, we maintain ground sets for each node.
-	Initially, all nodes are isolated (i.e., in singleton ground sets).
+%	Throughout the construction, we maintain ground sets for each node.
+	Initially, all nodes are isolated, i.e., each node is in a singleton ground set.
 	First, we choose a cluster arbitrarily and
@@
 	%We never dismiss any constructed ground set.
-	Each cluster hosts exactly $k$ nodes, and in any partition, a~single isolated node must be collocated with $B$.
-	We say that this isolated node that accompanies the ground set $B$ is \emph{special}.
-	We denote the first special node by $x_0$.
+	Each cluster hosts exactly $k$ nodes, and in any feasible partition,
+	 a~single isolated node must be collocated with $B$.
+	At any time,
+	we refer to the isolated node currently collocated with $B$ as the \emph{special} node.
+	Let $x_0$ denote the first special node.
 
@@
 	Precisely,
-	we create a~ground set $\{x_0, y_0\}$, 
-	where $x_0$ is the current special node and $y_0$ is an arbitrary isolated node.
+	we create a~ground set $\{x_0, y_0\}$, 	where
+%	 $x_0$ is the current special node and
+	 $y_0$ is an arbitrary isolated node.
 	Since \ALG does not have $\{x_0,y_0\}$ collocated, 
@@
 
-	We proceed by joining the current special node to a~larger ground set,
+	We proceed by joining the current special node to a~larger ground set in a similar way,
 	as long as a feasible partition exists afterwards.
@@
 	Otherwise,
-	if no such feasible partition exists,
-	 then we terminate the process and
+	no such feasible partition exists,
+	we terminate the process and
 	 we let $x^* := x_i, y^*:=y_i$.

### 2020-05-13 paper/disc20clustering.tex
Title

@@
  \usepackage[font={footnotesize}]{subcaption}
- \usepackage[font={footnotesize}]{caption}
+ \usepackage[font={footnotesize}]{caption}https://www.wykop.pl/
  \usepackage{breakcites}
@@
 
- \title{Tight Bounds for Dynamic Balanced Graph Partitioning}
+ \title{Optimal Algorithms for Dynamic Balanced Graph Partitioning}
 
@@
 %\EventNoEds{3}
- \EventLongTitle{34th International Symposium on DIStributed Computing (DISC) 2020)}
+ \EventLongTitle{34th International Symposium on DIStributed Computing (DISC) 2020}
  \EventShortTitle{DISC 2020}

### 2020-05-13 paper/disc20clustering.tex
rewriting the LB (fixed)

@@
  \usepackage[font={footnotesize}]{subcaption}
- \usepackage[font={footnotesize}]{caption}https://www.wykop.pl/
+ \usepackage[font={footnotesize}]{caption}
  \usepackage{breakcites}
@@
 	%
-	Our first contribution is a~significantly improved
-	\maciek{How to put here the information that it is for deterministic algorithms only? This sentence is already very long.}lower bound of $\Omega(k\cdot \ell)$ on the
+	Our first contribution is a~significantly improved deterministic
+	lower bound of $\Omega(k\cdot \ell)$ on the
 	competitive ratio, where $\ell$ is the number
@@
 	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least
-	 $(k-2)(\ell-1)/2 - 2$ for any $k\geq 3$ and $\ell \geq 2$.
+	 $(k-2)(\ell-1)/2 - 3/2$ for any $k\geq 3$ and $\ell \geq 2$.
 \end{theorem}
@@
 	For a ground set $C$ of nodes that are initially collocated in one cluster,
-	let $I(C)$ denote the cluster.
+	let $I(C)$ denote the cluster,
+	i.e., the cluster that is the \emph{origin} of $C$.
 %	If $C$ contains nodes that originate from different clusters
@@
 	Otherwise,
-	no such feasible partition exists,
-	we terminate the process and
-	 we let $x^* := x_i, y^*:=y_i$.
+	no such feasible partition exists and
+	we terminate the process.
 
 	 Once the process terminates,
-	 there are at most $\ell$ isolated nodes left 
-	 that are not contained in $\{x^*,y^*\}$.
+	 there are at most $\ell$ isolated nodes left. 
 	 Assume for contradiction that this is not the case,
-	 i.e., there are at least $\ell+1$ isolated nodes left
-	 in addition to $x^*$ and possibly $y^*$.
+	 i.e., there are at least $\ell+1$ isolated nodes left.
 	 Then there must be two isolated nodes $u$ and $v$,
-	 $u,v \not\in \{x^*, y^*\}$,
 	  with the same cluster of origin,
@@
 	 which contradicts our assumption.
-	 Hence, at the point of termination,
-	 there are at most $\ell$ isolated nodes not in $\{x^*,y^*\}$,
-	 or at most $\ell+2$ isolated nodes overall.
-
-\begin{comment}
-	We claim that expanding ground sets in this way can continue as long as at least $\ell$ isolated nodes exist.
-	Fix the partition just before the creation of a~ground set with $x_i$ for some~$i$.
-	We show that if at least $\ell$ isolated nodes exist (including $x_i$), then there exists a~perfect partition of ground sets after joining $x_i$ to a~ground.
-	To show this claim, we consider a~partition, where the largest ground set of each cluster originates from it (excluding the ground set $\{x_0, y_0\}$). \mahmoud{this sentence doesn't make sense.}
-	In this partition, each cluster contains:  at most one ground set larger than $1$, possibly some isolated nodes, and possibly the ground set $\{x_0, y_0\}$.
-	If $I(x_i)$ contains an isolated node, then we simply obtain a~perfect partition by swapping it with $x_i$.
-	In the following, we assume that no isolated nodes exist on $I(x_i)$.
-	The largest ground set of $I(x_i)$ is at most $k-1$, because $x_i$ is on $S$.
-	Therefore $\{x_0, y_0\}$ is present on $I(x_i)$, and we must evict it to make space for $x_i$.
-	\mahmoud{why  $\{x_0, y_0\}$ must be present there?}
-	To find the cluster for $\{x_0, y_0\}$, observe that we have at most $\ell-1$ clusters with isolated nodes, and we have at least $\ell$ isolated nodes --- hence a~cluster with two isolated nodes exists.
-	To find a~feasible partition, we swap $\{x_0, y_0\}$ with these two collocated isolated nodes, and then we swap an isolated node from $I(x_i)$ with~$x_i$.
-	\mahmoud{but didn't you assume there is no isolated on $I(x_i)$?}
-
-	To keep \OPT's cost low, we terminate prematurely --- we stop extending ground sets while there are still $\ell+1$ isolated nodes (i.e., neglecting two last ground set expansions).
-	\mahmoud{here you are making an exception to the already established process.
-Instead, adapt the termination condition so that it triggers prematurely, by saying something like "issue the last request only if there will be a feasible partition." }
-	Consider \ALG's partition after the last expansion.
-	Among $\ell+1$ isolated nodes there must exist a~pair that originates from the same cluster (we have $\ell$ clusters). 
-	We denote these two nodes by $x^*$ and $y^*$.
-\end{comment}
+	 Hence, there are at most $\ell$ isolated nodes
+	 at the point of termination.
+
+	The number of isolated nodes decrease after each step.
+	Consider the last step, $i^*$, when exactly $\ell+1$ isolated nodes exist before  $x_{i^*}$ joins to another ground set.
+	I.e., there are less than $\ell+1$ isolated nodes after the step $i^*$.
+	Then, at the end of step $i^*-1$,
+	there exist two isolated nodes $x^*$ and $y^*$
+	with the same cluster of origin, i.e., $I(\{x^*\})=I(\{y^*\})$.
+%
+    In what follows,
+    we bound costs incurred by \OPT and \ALG over the  sequence of requests
+    $\sigma := \{ \{x_0,y_0\}, \dots, \{x_{i^*-1},y_{i^*-1}\}  \}$.
+
 	The optimal strategy is to perform two node swaps.
@@
 	by swapping them with $x^*$ and $y^*$.
-	In this partition, no ground set is split.
-	Due to the termination condition,
-	we never issue requests between nodes $x^*, y^*$
-	 which are separated in this partition.
+	In this partition, no ground set is split and
+%	there is no request to $x^*$ or $y^*$ in $\sigma$.
+	therefore \OPT pays only for the two swaps.
 
@@
 %	some ground set grows by one node at each of these steps.
-	Consider any ground set $C^*$ at the point of termination.
-	This ground set has grown exactly $|C^*|-1$ times until the termination. 
-	Let $\mathcal{C}^*$ be the set of all ground sets at the point of termination.
+	Consider any ground set $C^*$ after serving $\sigma$.
+	This ground set has grown exactly $|C^*|-1$ times until the end of $\sigma$. 
+	Let $\mathcal{C}^*$ be the set of all ground sets after serving the last request in $\sigma$.
 	Thus, $\mathcal{C}^*$ includes ground sets $B$,
 	$\{x_0,y_0\}$,
-	and up to $\ell+2$ singleton ground sets.
+	and $\ell+1$ singleton ground sets.
 %	(including $x^*$ and possibly $y^*$).	
-	Each time we join the special node to the a largest ground set.
+	Each time we join the special node to the largest ground set of same origin.
 	Hence, every  other ground set in $\mathcal{C}^*$ has grown maximally.
@@
 	one per cluster except the cluster containing $B$. 
-	Therefore,  $|\mathcal{C}^*| \leq 1 + 1 + (\ell+2)  + (\ell-1) = 2\ell+3$.
+	Therefore,  $|\mathcal{C}^*| \leq 1 + 1 + (\ell+1)  + (\ell-1) = 2\ell+2$.
 	Note that among all non-singleton ground sets in $\mathcal{C}^*$,
@@
 		\sum_{C^* \in \mathcal{C}^* } 1 - (k-1) \\
-	&\geq  k  \ell - (2\ell+3) - (k-1) =  (k-2)(\ell-1) - 4,
+	&\geq  k  \ell - (2\ell+2) - (k-1) =  (k-2)(\ell-1) - 3,
 	\end{align*}
@@
 	The competitive ratio is then
-	$\ALG / \OPT \geq ((k-2)(\ell-1) - 4) / 2$.
+	$\ALG / \OPT \geq ((k-2)(\ell-1) - 3) / 2$.
 	%%
@@
 defined as 
-$\Delta(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$,
+$\Delta(P) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$,
  is the number of nodes in $P$ that do not reside in their initial cluster.
 In other words,
-at least $\Delta(P, P_I)$ node migrations are required in order to reach the partition $P$ from $P_I$, and thus
-$\OPT \geq \Delta^*:= \Delta(P_F, P_I) $.
-We  use $ \Delta(P_F) $ wherever the initial partition is clear from the context \maciek{It always is, we fixed it in the beginning.
-Maybe just define it as the distance from initial partition, but don't pass it as an argument}.
+at least $\Delta(P)$ node migrations are required in order to reach the partition $P$ from $P_I$, and thus
+$\OPT \geq \Delta^*:= \Delta(P_F) $.
+%We  use $ \Delta(P_F) $ wherever the initial partition is clear from the context
 The scheme of the algorithm can be found in the  appendix (Algorithm \ref{alg:ppl}).
@@
 \noindent
-\textbf{Approximation for offline rematching.}
-\mahmoud{TBD}
+%\textbf{Approximation for offline rematching.}
 %We briefly sketch an approximation for the offline problem,

### 2020-05-13 paper/disc20clustering.tex
special => pivot node

@@
 We start by constructing a~ground set of size $k-1$ on an arbitrarily chosen cluster.
-In any partition, there must exist an isolated node collocated with the ground set of size $k-1$, and we call it \emph{special}.
-We issue requests between the current special node, and some node that was initially collocated with it.
-By repeating such requests, almost every node is collocated with the first ground set at some point .
+In any partition, there must exist an isolated node collocated with the ground set of size $k-1$.
+%, and we refer to it as the \emph{pivot} node.
+We issue requests between this node and some node that was initially collocated with it.
+By repeating such requests, almost every node is once collocated with the first ground set.
 In comparison, we show that there exists an optimal offline algorithm \OPT that performs only two node exchanges ("swaps").
@@
 	At any time,
-	we refer to the isolated node currently collocated with $B$ as the \emph{special} node.
-	Let $x_0$ denote the first special node.
-
-	Then, we join the special node to a~larger ground set to force its eviction.
+	we refer to the isolated node currently collocated with $B$ as the \emph{pivot} node.
+	Let $x_0$ denote the first pivot node.
+
+	Then, we join the pivot node to a~larger ground set to force its eviction.
 	Precisely,
 	we create a~ground set $\{x_0, y_0\}$, 	where
-%	 $x_0$ is the current special node and
+%	 $x_0$ is the current pivot node and
 	 $y_0$ is an arbitrary isolated node.
@@
 
-	We proceed by joining the current special node to a~larger ground set in a similar way,
+	We proceed by joining the current pivot node to a~larger ground set in a similar way,
 	as long as a feasible partition exists afterwards.
@@
 %	(including $x^*$ and possibly $y^*$).	
-	Each time we join the special node to the largest ground set of same origin.
+	Each time we join the pivot node to the largest ground set of same origin.
 	Hence, every  other ground set in $\mathcal{C}^*$ has grown maximally.

### 2020-05-13 paper/disc20clustering.tex
improvements

@@
 	We proceed by joining the current pivot node to a~larger ground set in a similar way,
-	as long as a feasible partition exists afterwards.
+	as long as a feasible partition exists afterward.
 	Consider the isolated node $x_i$ currently collocated with $B$.
@@
 	 which contradicts our assumption.
-	 Hence, there are at most $\ell$ isolated nodes
-	 at the point of termination.
+	 Hence, there are at most $\ell$ isolated nodes after the termination.
 
 	The number of isolated nodes decrease after each step.
-	Consider the last step, $i^*$, when exactly $\ell+1$ isolated nodes exist before  $x_{i^*}$ joins to another ground set.
-	I.e., there are less than $\ell+1$ isolated nodes after the step $i^*$.
-	Then, at the end of step $i^*-1$,
+	Consider the last step, $j$, when exactly $\ell+1$ isolated nodes exist before  $x_{j}$ joins to another ground set.
+	That is, there are less than $\ell+1$ isolated nodes after the step $j$.
+	Then, at the end of step $j-1$,
 	there exist two isolated nodes $x^*$ and $y^*$
@@
 %
-    In what follows,
+   Next,
     we bound costs incurred by \OPT and \ALG over the  sequence of requests
-    $\sigma := \{ \{x_0,y_0\}, \dots, \{x_{i^*-1},y_{i^*-1}\}  \}$.
+    $\sigma := \langle \{x_0,y_0\}, \dots, \{x_{j-1},y_{j-1}\}  \rangle$.
 
@@
 	\ALG performs at least one swap each time a~ground set grows,
-	i.e., each time it joins the isolated node $x_i$ (collocated with $B$), to  the ground set $C_i$ as specified before.
+	i.e., each time it joins the isolated node $x_i$ (collocated with $B$) to  the ground set $C_i$ as specified before.
 %	In other words,
@@
 	This ground set has grown exactly $|C^*|-1$ times until the end of $\sigma$. 
-	Let $\mathcal{C}^*$ be the set of all ground sets after serving the last request in $\sigma$.
-	Thus, $\mathcal{C}^*$ includes ground sets $B$,
+	Let $\mathcal{S}$ be the set of all ground sets after serving the last request in $\sigma$.
+	Thus, $\mathcal{S}$ includes ground sets $B$,
 	$\{x_0,y_0\}$,
 	and $\ell+1$ singleton ground sets.
-%	(including $x^*$ and possibly $y^*$).	
-	Each time we join the pivot node to the largest ground set of same origin.
-	Hence, every  other ground set in $\mathcal{C}^*$ has grown maximally.
-	There are at most $\ell-1$ such ground sets in addition to the ones 
-	we considered so far,
-	one per cluster except the cluster containing $B$. 
-	Therefore,  $|\mathcal{C}^*| \leq 1 + 1 + (\ell+1)  + (\ell-1) = 2\ell+2$.
-	Note that among all non-singleton ground sets in $\mathcal{C}^*$,
+	Among the remaining ground sets in $\mathcal{S}$,
+	no two ground sets have the same origin.
+	Otherwise,
+   the smaller ground set is either a singleton,
+   which contradicts the bound $\ell+1$ on the number of singletons,
+	or we have joined nodes to it at some step,
+	contradicting our choice of the largest ground set of the same origin.
+	Hence, every  other ground set in
+	$\mathcal{S}$ has grown \emph{maximally} over $\sigma$.
+	There are at most $\ell-1$ maximally grown ground sets,
+	one per cluster,
+	 excluding the cluster containing $B$. 
+	Therefore,  $|\mathcal{S}| \leq 1 + 1 + (\ell+1)  + (\ell-1) = 2\ell+2$.
+	Note that among all non-singleton ground sets in $\mathcal{S}$,
 	 only $B$ does not grow during the process.
@@
     the total number of times that 
-	a ground set in $\mathcal{C}^*$  has grown is
+	a ground set in $\mathcal{S}$  has grown is
 	\begin{align*}
-	\sum_{C^* \in \mathcal{C}^* } (|C^*|-1) - (k-1) &= 
-	\sum_{C^* \in \mathcal{C}^* } |C^*| - 
-		\sum_{C^* \in \mathcal{C}^* } 1 - (k-1) \\
+	\sum_{C^* \in \mathcal{S} } (|C^*|-1) - (k-1) &= 
+	\sum_{C^* \in \mathcal{S} } |C^*| - 
+		\sum_{C^* \in \mathcal{S} } 1 - (k-1) \\
 	&\geq  k  \ell - (2\ell+2) - (k-1) =  (k-2)(\ell-1) - 3,
 	\end{align*}
-	which is the least number of swaps performed by \ALG.
+	which bounds the number of swaps performed by \ALG.
 	The competitive ratio is then

### 2020-05-13 paper/disc20clustering.tex
defining component

@@
 
-%\medskip
-%In our algorithms, 
-A~\emph{component} is a~subset of frequently communicating nodes \maciek{this is not a~good definition. intuitions, maybe}.
+We maintain subsets of frequently communicating nodes as \emph{components}.
+Components 
+Initially,
+each node constitutes a single-node component which we refer to as a \emph{singleton} components, 
+and the contained node is an \emph{isolated} node.
+We define larger components in terms of smaller components.
+Concretely,
+given a (sub)sequence of requests,
+two components $C_1$ and $C_2$
+ \emph{merge} into one component as soon as
+ for some pair of nodes $v_1 \in C_1$ and $v_2 \in C_2$,
+the frequency of requests $\{v_1,v_2\}$ reaches a certain threshold through the sequence.
 We keep all nodes of a~component always collocated in the same cluster,
@@
 A~partition that maintains this invariant is a~\emph{component respecting} partition.
-%We say that a~partition is \emph{component respecting}
-%if the nodes belonging to the same component are always collocated.
-We say that a~component is \emph{singleton} if it contains exactly one node,
-and the node in this component is \emph{isolated}.
+%Intuitively, a component preserves the history of requests between its nodes.
 

### 2020-05-13 paper/disc20clustering.tex
RM intro

@@
 Initially,
-each node constitutes a single-node component which we refer to as a \emph{singleton} components, 
+each node constitutes a single-node component which we refer to as a \emph{singleton} component, 
 and the contained node is an \emph{isolated} node.
@@
  an algorithm for the \OBRP restricted to clusters of capacity $k=2$.
-We interpret a~pair of nodes collocated in one cluster as a~matched pair.
+We interpret a~pair of nodes collocated in one cluster as a~``matched'' pair.
 Hence,
-the problem is an online variant of the classic matching problem where
-a matched pair can separate from each other in order to ``rematch'' with two other nodes
-\maciek{Maybe we use SIDMA's connection to matching instead of this local view.
-SIDMA: "This can be viewed as an online maximal (re)matching problem: clusters of size two contain (“match”) exactly one pair of nodes, and maximizing pairwise communication within each cluster is equivalent to minimizing inter-cluster communication."}.
-This is known as the  \emph{Online Rematching} 
-problem and a~(non-strict) 7-competitive algorithm is already given by~\cite{repartition-disc},
+the problem is an online variant of the maximal matching problem where
+%a match is retractable.
+%That is, 
+a matched pair can separate in order to ``rematch'' with two other nodes.
+Rematching is necessary for  maximizing intra-cluster communications,
+which is dual to minimizing inter-cluster communications.
+This is known as the  \emph{online rematching} 
+problem and a~non-strict 7-competitive algorithm is already given by~\cite{repartition-disc},
 in which the ratio comes with an additive factor in $O(\alpha\ell^2)$.
@@
 (i.e., with no additive factor).
-In addition,
-our algorithm is similar
-and slightly simpler than the one in~\cite{repartition-disc}, 
-while our analysis is significantly more simple and concise,
+Our algorithm is slightly simpler than the one in~\cite{repartition-disc}, 
+while our analysis is significantly  simpler and more concise,
 thanks to the charging scheme we devise here.

### 2020-05-14 paper/disc20clustering.tex
cleanup
figure 1

@@
 Figure~\ref{fig:lb} illustrates the constructs used in the lower bound.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.3\textwidth]{figs/substitute}
-	\caption{todo.}
-	\label{fig:lb}
-\end{figure}
 
@@
 
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.3\textwidth]{figs/substitute}
-	\caption{an unsaturated edge that is inside a~component. If there is a~space next to it, maybe cluster types should fit next to it? Not sure.}
-\end{figure}
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[width=0.3\textwidth]{figs/substitute}
+%	\caption{an unsaturated edge that is inside a~component. If there is a~space next to it, maybe cluster types should fit next to it? Not sure.}
+%\end{figure}
 
@@
 
-\medskip
-
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.4\textwidth]{figs/cyclic}
+	\caption{
+		\OPT  collocates 5 pairs, paying one migration per pair.
+	Dashed lines represent requests.
+	An arrow, e.g.~from $b_1$ to $a_1$ means  that $b_1$ replaces $a_1$. }
+	\label{fig:cyclic}
+\end{figure}
+
+%\medskip
 \noindent
@@
 % before they inflict too much communication cost.
- Therefore it pays the price of only one migration per pair  (\mahmoud{see Figure \ref{fig:TBD}}).
+ Therefore it pays the price of only one migration per pair  (see Figure \ref{fig:cyclic}).
 Hence,
@@
 	$R_{t'} \cap R_t = \sigma_{\{u,w\}}$ or
-	$R_{t'} \cap R^*_{\{u,v\}} = \sigma_{\{u,w\}}$
-	(\mahmoud{See figure \ref{fig:doubleCount}}). 
+	$R_{t'} \cap R^*_{\{u,v\}} = \sigma_{\{u,w\}}$.
+%	(\mahmoud{See figure \ref{fig:doubleCount}}). 
 	This means,
@@
 
-\maciek{Make it more general (essentially the pseudocode of DET) to also link to this section from k=3.}
-
-\maciek{todo: intro to the algorithm (mostly about the choice of repartition procedure in \TAlg and PPL)}
-
 \begin{algorithm}

### 2020-05-14 paper/disc20clustering.tex
Title / abstract
wip

@@
 \begin{document}
+\maketitle
 
@@
 	By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), we can reduce the network load and  improve application performance. 
-	However, the communication pattern of different applications is often unknown a~priori and may change over time, hence it needs to be learned in an online manner.
+	However, the communication pattern of different applications is often unknown a~priori and may change over time, hence it needs to be learned in an~online manner.
 	%
@@
 
-\maketitle
-
 %\renewcommand{\shortauthors}{M.~Pacut, M.~Parham, S.~Schmid}
@@
 by dynamically migrating nodes (e.g., virtual machines) which communicate frequently topologically closer to each other, network traffic can be reduced significantly.  
-However, migrations should be used moderately, as migrations also entail overheads.
+However, migrations entail overhead, and should be used moderately.
 
@@
 We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.
-In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a~given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the whole input sequence in advance.
+In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a~given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the entire input sequence in advance.
 %We emphasize that all our algorithms have a~\emph{strict} competitive ratio (without an additive term).
@@
 is determined a~priori and is unknown to algorithms,
- and requests to such pairs arrive indefinitely.
+ and such pairs communicate forever.
 Any algorithm must eventually collocate  pairs of communicating nodes,
@@
 This assumption is not only more realistic but also significantly more challenging, as it is related to hard problems such as integer partitioning~\cite{integer-partitions-book}.
-In terms of results without augmentation, so far, it is only known that there exists a~$O(k^2 \cdot \ell^2)$-competitive algorithm~\cite{repartition-disc}; the best known lower bound is significantly lower, namely $\Omega(k)$.
+In terms of results without augmentation, so far, it is only known that there exists an~$O(k^2 \cdot \ell^2)$-competitive algorithm~\cite{repartition-disc}; the best known lower bound is significantly lower, namely $\Omega(k)$.
 For $k=2$, Avin et al.~\cite{repartition-disc} presented a~$7$-competitive algorithm with a~substantial ($\Omega(\ell^2)$) additive constant.
@@
 This paper presents several new results on the dynamic graph partitioning problem  without augmentation.
-For the learning model we present a~lower bound of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
+For the learning model, we present a~lower bound of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
 (that holds also in the general partitioning model).
 The best known lower bound so far was $\Omega(k)$~\cite{repartition-disc} that holds only in the general partitioning model.
-We further present an asymptotically optimal, 
+We complement this result with an asymptotically optimal, 
 $O(k\cdot \ell)$-competitive algorithm
@@
 so far $O(\ell^2)$~\cite{repartition-disc}.
-We further present a~strictly $6$-competitive algorithm for $k=2$ that improves upon previous $7$-competitive algorithm with $O(\alpha\ell^2)$ additive constant.
-
-All algorithms in this paper have strict competitive ratio (i.e., without an additive term).
+We further present a~strictly $6$-competitive algorithm for $k=2$ that improves upon the previous $7$-competitive algorithm with $O(\alpha\ell^2)$ additive constant.
+
+All algorithms in this paper have a strict competitive ratio (i.e., without an additive term).
 Table \ref{tab:overview} provides an overview of our contributions compared to prior work.
@@
 A straightforward analysis of the algorithm from Section~\ref{sec:k3} results in the bound $O(\ell^2)$, and the analysis must be improved in two places.
-One must bound the cost of each of $O(\ell)$ reconfigurations in~phase by a~constant, and show that the optimum offline algorithm must pay significant cost for inter-cluster requests.
+One must bound the cost of each of $O(\ell)$ reconfigurations in~phase by a~constant, and show that the optimum offline algorithm must pay a~significant cost for inter-cluster requests.
 We bound the cost of the latter by estimating the capabilities of the optimum offline algorithm to \emph{prepare} for an incoming sequence of requests.
@@
 
-A~\emph{ground set} is a~subset of nodes such that if an algorithm does not have them all collocated,
-then the split parts communicate indefinitely until the algorithm collocates them.
-%A~\emph{ground set} is a~group of nodes that  communicate indefinitely when an algorithm splits them (places at different clusters).
-This ensure us that any competitive algorithm would maintain a~partition of ground sets into clusters.
-As a result,
- a ground set is revealed to an algorithm only after the algorithm splits it.
- More precisely,
-the adversary constructs ground sets depending on the choices of a~given deterministic online algorithm.
-%Once we construct a~ground set, it lasts until the algorithm terminates.
-We say that a~ground set is \emph{singleton} if it contains only one node,
-and we refer to such node as an \emph{isolated} node.
+A~\emph{ground set} is a group of nodes that repeatedly communicate if they are not collocated.
+If an algorithm splits any pair of nodes from a ground set, then the adversary continues to produce requests to this pair until the algorithm collocates it.
+Any competitive algorithm must maintain a~perfect partition of ground sets into clusters.
+We emphasize that ground sets are revealed only upon their nodes are split.
+The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
+Once we construct a ground set, it lasts until the end of the input sequence.
+We say that a ground set is \emph{singleton} if it contains exactly one node.
+We say that the node that belongs to a singleton ground set is \emph{isolated}.
 
@@
 In comparison, we show that there exists an optimal offline algorithm \OPT that performs only two node exchanges ("swaps").
-Figure~\ref{fig:lb} illustrates the constructs used in the lower bound.
+%Figure~\ref{fig:lb} illustrates the constructs used in the lower bound.
 
@@
 	\label{th:lowerbound}
-	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least
+	The competitive ratio of any deterministic online algorithm for the learning variant of Dynamic Balanced Graph Partitioning is at least
 	 $(k-2)(\ell-1)/2 - 3/2$ for any $k\geq 3$ and $\ell \geq 2$.

### 2020-05-14 paper/disc20clustering.tex
Rewording the LB (still follows the Mahmoud's construction)

@@
 	Since \ALG does not have $\{x_0,y_0\}$ collocated, 
-	we issue an external request between them so that \ALG collocates them.
-	\ALG cannot collocate $\{x_0, y_0\}$ with $B$ (as $B$'s size is $k-1$), hence it collocates them in a different cluster.
+	the adversary issues external requests to this pair until \ALG collocates it.
+	\ALG cannot collocate $\{x_0, y_0\}$ with $B$ (as $B$'s size is $k-1$), hence it collocates them in a~different cluster.
 	In order to preserve a~feasible partition of nodes after collocating $\{x_0, y_0\}$,
-	\ALG must replace $x_0$ with another isolated node.
+	\ALG must replace $x_0$ with another isolated node that becomes the new pivot.
 %	denoted by $x_1$.
 
-	We proceed by joining the current pivot node to a~larger ground set in a similar way,
-	as long as a feasible partition exists afterward.
-	Consider the isolated node $x_i$ currently collocated with $B$.
-	 We issue a request between $x_i$ and some node $y_i \in C_i$,
-	 where $C_i$ is the largest ground set
-	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$,
-	 only if  a feasible partition exists
-	 that collocates all nodes in $\{x_i\} \cup C_i $.
-	Any such feasible partition replaces $x_i$ with some isolated node $x_{i+1}$,
-	as the new ground set $\{x_i\} \cup C_i $  may not be ever split.
-	Otherwise,
-	no such feasible partition exists and
-	we terminate the process.
+	We proceed by joining the current pivot node to a~larger ground set in a similar way, as long as a feasible partition of ground sets exists after the join.
+	In a single step, we merge the ground sets of the current pivot $x_i$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x_i)$.
+	This forces \ALG to collocate $\{x_i\} \cup C_i $ and to replace $x_i$ with some isolated node $x_{i+1}$.
+	(Note that these ground sets contain nodes that originate from the same cluster, whereas the nodes $x_0$ and $y_0$ originate from distinct clusters.)
 
@@
 
-	The number of isolated nodes decrease after each step.
+	The number of isolated nodes decreases after each step.
 	Consider the last step, $j$, when exactly $\ell+1$ isolated nodes exist before  $x_{j}$ joins to another ground set.

### 2020-05-14 paper/disc20clustering.tex
Comment

@@
 A~\emph{ground set} is a group of nodes that repeatedly communicate if they are not collocated.
-If an algorithm splits any pair of nodes from a ground set, then the adversary continues to produce requests to this pair until the algorithm collocates it.
+If an algorithm splits any pair of nodes from a ground set, then the adversary continues to produce requests to this pair until the algorithm collocates it \maciek{learning model: does not need to continue issuing requests.}
 Any competitive algorithm must maintain a~perfect partition of ground sets into clusters.

### 2020-05-14 paper/disc20clustering.tex
Small

@@
 We maintain subsets of frequently communicating nodes as \emph{components}.
-Components 
 Initially,

### 2020-05-14 paper/disc20clustering.tex
Comment

@@
 each node constitutes a single-node component which we refer to as a \emph{singleton} component, 
-and the contained node is an \emph{isolated} node.
+and the contained node is an \emph{isolated} node \maciek{unclear: isolated = contained?}.
 We define larger components in terms of smaller components.

### 2020-05-14 paper/disc20clustering.tex
Comment

@@
 we move the whole component that contains it.
-A~partition that maintains this invariant is a~\emph{component respecting} partition.
+A~partition that maintains this invariant is a~\emph{component respecting} partition \maciek{wrong use of "this".
+Better to define component respecting, and only then say that ALG only uses such partitions (but it is also implied in the learning model...)}.
 %Intuitively, a component preserves the history of requests between its nodes.

### 2020-05-14 paper/disc20clustering.tex
Comment

@@
 In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).
-A generic sketch of these two algorithms can be found in the appendix (Algorithm \ref{alg:ppl}).
+A generic sketch of these two algorithms can be found in the appendix (Algorithm \ref{alg:ppl}) \maciek{We don't have the generic. Also in the next paragraph you refer to the code}.
 
@@
 at least $\Delta(P)$ node migrations are required in order to reach the partition $P$ from $P_I$, and thus
-$\OPT \geq \Delta^*:= \Delta(P_F) $.
+$\OPT \geq \Delta^*:= \Delta(P_F) $ \maciek{Confusing. We have space. First define :=, then say that OPT bounds}.
 %We  use $ \Delta(P_F) $ wherever the initial partition is clear from the context

### 2020-05-14 paper/disc20clustering.tex
Comment

@@
 \begin{proof}
+	\maciek{Maybe: "On each inter-cluster request, the algorithm enumerates all component respecting $\ell$-way partitions in the order of increasing distance to $P_I$" Then we don't need to say i.e.}
 	On each inter-cluster request,

### 2020-05-14 paper/disc20clustering.tex
Comment

--- 2020-05-14 paper/disc20clustering.tex
@@
 \begin{proof}
+	\maciek{Maybe: "On each inter-cluster request, the algorithm enumerates all component respecting $\ell$-way partitions in the order of increasing distance to $P_I$" Then we don't need to say "i.e." and two sentences that explain the same thing}
 	On each inter-cluster request,

### 2020-05-14 paper/disc20clustering.tex
wip

@@
 	each repartitioning costs at most $2\cdot\OPT$.
-	The total cost is therefore at most $2\cdot\OPT\cdot (k-1) \cdot\ell$, which implies the competitive ratio.
+	The total cost is thus at most $2\cdot\OPT\cdot (k-1) \cdot\ell$, which implies the competitive ratio.
 \end{proof}
@@
 
-The algorithm analyzed in this section is a modified version of the algorithm DET proposed by Avin et al.~\cite{repartition-disc}, that for $k=3$ is $O(\ell^2)$-competitive.
+The algorithm analyzed in this section is a modified version of the algorithm DET proposed by Avin et al.~\cite{repartition-disc}, which for $k=3$ is $O(\ell^2)$-competitive.
 In our algorithm, we choose the closest partition after a component merge instead of an arbitrary one.

### 2020-05-14 paper/disc20clustering.tex
wip

@@
 	\label{lem:1req}
-	In a~single repartitioning of nodes (after a~merge of components), \TAlg exchanges at most two pairs of nodes.
+	In a~single repartition of nodes (after a~merge of components), \TAlg exchanges at most two pairs of nodes.
 \end{lemma}
@@
 \begin{proof}
-	If no component respecting partition exists after the merge of components, then \TAlg resets all components, ends the phase and performs no repartitioning.
+	If no component respecting partition exists after the merge of components, then \TAlg resets all components, ends the phase and performs no migrations.
 	It suffices to show that the merged component has size at least $4$ to conclude that \TAlg incurs no cost.
@@
 
-	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the repartitioning is local within $U$ and $V$.
-	In this case, a~local repartition is possible (within $U$ and $V$), for a~cost of at most $2$ swaps.
+	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the repartitioning is local within $U$ and $V$, for the~cost of at most $2$ swaps.
 	If either $U$ or $V$ is of type $C_3$, a~component of size $3$ participates in a~merge, and we have a~component of size at least $4$, and \TAlg ends the phase with no repartitioning.

### 2020-05-14 paper/disc20clustering.tex
wip

@@
 This means that the algorithm must eventually collocate all nodes of a~ground set to be competitive.
-We reveal the next ground set only after the collocation, hence we can repeat the analysis of the algorithm for the learing problem.
+We reveal the next ground set only after the collocation, hence we can repeat the analysis of the algorithm for the learning problem.
 Finally, we note that the construction is oblivious to the choice of the reconfiguration cost $\alpha$: we compare the 	number of node exchanges of \ALG and \OPT.
@@
 \textbf{Perfect Partition Learner algorithm.}
+\maciek{There should be a description of the algoritm. What steps it takes. Instead, we have PPL's properties here, and the algorithm description is not present at all.}
 Fix the initial partition
@@
 
-With each repartitioning,
+With each repartitioning \maciek{Unclear what repartitioning means. Maybe with each merge action?},
 \PPL moves to a~component respecting partition that minimizes the distance to the initial partition $P_I$.
@@
 \begin{proof}
-	If no component respecting partition exists after the merge of components, then \TAlg resets all components, ends the phase and performs no migrations.
+	If no component respecting partition exists after the merge of components, then \TAlg resets all components, ends the phase and performs no repartition.
 	It suffices to show that the merged component has size at least $4$ to conclude that \TAlg incurs no cost.
@@
 
-	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the repartitioning is local within $U$ and $V$, for the~cost of at most $2$ swaps.
-	If either $U$ or $V$ is of type $C_3$, a~component of size $3$ participates in a~merge, and we have a~component of size at least $4$, and \TAlg ends the phase with no repartitioning.
+	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the repartition is local within $U$ and $V$, for the~cost of at most $2$ swaps.
+	If either $U$ or $V$ is of type $C_3$, a~component of size $3$ participates in a~merge, and we have a~component of size at least $4$, and \TAlg ends the phase with no repartition.
 
@@
 	Otherwise, if one of $u,v$ belongs to a~component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
-
 	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a~cluster different from $U$ and $V$.
@@
 
-	In each case, we showed that a~partition is reachable in at most two swaps.
+	In each case, we showed that a~component respecting partition is reachable in at most two swaps.
 \end{proof}
@@
 
-	As \TAlg is component respecting, it never increases any counter above $\alpha$.
+	\TAlg is component respecting, hence it never increases any counter above $\alpha$.
 	We say that the pair $(u, v)$ is \emph{saturated} if the counter's value is $\alpha$, and \emph{unsaturated} otherwise (saturation of a~pair leads to a~merge action).
@@
 	exceeding this number would mean that a~component of size $4$ exists, and the phase should have ended already.
-	We bound the cost of each repartitioning after a~merge by Lemma~\ref{lem:1req}, obtaining $\TAlg(M) \leq 8\alpha\cdot\ell$.
+	We bound the cost of each repartition after a~merge by Lemma~\ref{lem:1req}, obtaining $\TAlg(M) \leq 8\alpha\cdot\ell$.
 	%Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.

### 2020-05-14 paper/disc20clustering.tex
wip

@@
 \textbf{Perfect Partition Learner algorithm.}
-\maciek{There should be a description of the algoritm. What steps it takes. Instead, we have PPL's properties here, and the algorithm description is not present at all.}
+\maciek{There should be a short description of the algorithm: what steps it takes, how it treats the input.
+Instead, we have PPL's properties here, and the algorithm description is not present at all.}
 Fix the initial partition
@@
 
-	Now we bound $\TAlg(\sigma_I)$.
-	A~cluster of type $C_3$ contributes at most $3 \alpha - 1$ to $\TAlg(\sigma_I)$, as two pairs of nodes from the component are saturated and incurred the cost $\alpha$ each, and the third, unsaturated pair incurs the cost at most $\alpha-1$.
-	Other cluster types contribute less: $C_1$ contributes~$0$ and $C_2$ contributes $\alpha$.
+	Now we bound $\TAlg(\sigma_I)$ by summing the intra-component counters of each cluster at the end of the phase.
+	The sum of intra-component counters in a cluster of type $C_3$ is at most $3 \alpha - 1$: two pairs of nodes from the component are saturated and its counter is $\alpha$ each, and the counter of the third, unsaturated pair is at most $\alpha-1$.
+	The sum of counters inside $C_1$ is $0$, and inside $C_2$ it is $\alpha$.
 	Summing over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq (3 \alpha-1)\cdot \ell \leq 3\alpha\cdot\ell$.

### 2020-05-14 paper/disc20clustering.tex
wip

@@
 
-	Now we bound $\TAlg(\sigma_I)$ by summing the intra-component counters of each cluster at the end of the phase.
+	We bound $\TAlg(\sigma_I)$ by summing the intra-component counters of each cluster at the end of the phase.
 	The sum of intra-component counters in a cluster of type $C_3$ is at most $3 \alpha - 1$: two pairs of nodes from the component are saturated and its counter is $\alpha$ each, and the counter of the third, unsaturated pair is at most $\alpha-1$.
-	The sum of counters inside $C_1$ is $0$, and inside $C_2$ it is $\alpha$.
+	The sum of counters inside $C_1$ is~$0$, and inside $C_2$ it is $\alpha$.
 	Summing over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq (3 \alpha-1)\cdot \ell \leq 3\alpha\cdot\ell$.
@@
 
-	Moreover, \TAlg paid for all requests from $\sigma_E$, and thus $\TAlg(\sigma_E) = |\sigma_E|$.
+	Furthermore, \TAlg paid for all requests from $\sigma_E$, and thus $\TAlg(\sigma_E) = |\sigma_E|$.
 	In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 11\alpha\cdot \ell + |\sigma_E|$ during this phase.
@@
 
-	%First, we estimate the cost related to $\sigma_I$.
-	We have $\OPT(M) + \OPT(\sigma_I) \geq \alpha$, as the phase ended when the components of \TAlg{} could not be partitioned without splitting them.
-	Hence, for every possible partition of $\OPT$, there exists a~non-collocated pair of nodes with at least $\alpha$ requests between them, and
-	$\OPT$ either served them remotely or performed a~migration.
-
-	\medskip
-	Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
+	First, we show $\OPT(M) + \OPT(\sigma_I) \geq \alpha$.
+	Assume that \OPT's partition is fixed throughout the phase (as otherwise \OPT pays $\alpha$ for a migration).
+	The phase ended when the components of \TAlg{} could not be partitioned without splitting them.
+	Hence, for every possible partition of $\OPT$, there exists a~non-collocated saturated pair, and
+	$\OPT$ paid for $\alpha$ requests that saturated the pair.
+
+	Next, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
 	Fix a~partition for \OPT prior serving the sequence $\sigma_E$.

### 2020-05-14 paper/disc20clustering.tex
refactoring the LB

@@
 	\label{th:lowerbound}
-	The competitive ratio of any deterministic online algorithm for the learning variant of Dynamic Balanced Graph Partitioning is at least
-	 $(k-2)(\ell-1)/2 - 3/2$ for any $k\geq 3$ and $\ell \geq 2$.
+	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least
+	 $(k-2)(\ell-1)/2 - 2$ for any $k\geq 3$ and $\ell \geq 2$.
 \end{theorem}
@@
 
-	We proceed by joining the current pivot node to a~larger ground set in a similar way, as long as a feasible partition of ground sets exists after the join.
-	In a single step, we merge the ground sets of the current pivot $x_i$ and the largest ground set $C$ s.t. $C \neq \{x_0,y_0\}$ and $I(C) = I(x_i)$.
-	This forces \ALG to collocate $\{x_i\} \cup C_i $ and to replace $x_i$ with some isolated node $x_{i+1}$.
-	(Note that these ground sets contain nodes that originate from the same cluster, whereas the nodes $x_0$ and $y_0$ originate from distinct clusters.)
-
-	 Once the process terminates,
-	 there are at most $\ell$ isolated nodes left. 
-	 Assume for contradiction that this is not the case,
-	 i.e., there are at least $\ell+1$ isolated nodes left.
-	 Then there must be two isolated nodes $u$ and $v$,
+	We proceed by joining the current pivot node to a~larger ground set in  similar steps.
+	Consider the step $i$,
+	when the isolated node $x_i$ is collocated with $B$.
+	 We issue a request between $x_i$ and some node $y_i \in C_i$,
+	 where $C_i$ is the largest ground set
+	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$.
+%	 only if  a feasible partition exists
+%	 that collocates all nodes in $\{x_i\} \cup C_i $.
+	Any feasible partition replaces $x_i$ with some isolated node $x_{i+1}$,
+	as the new ground set $\{x_i\} \cup C_i $  may not be ever split.
+%	Otherwise,
+%	no such feasible partition exists and
+%	we terminate the process.
+	We terminate the process once the number of remaining isolated nodes 
+	is less than $\ell+3$.
+	At each step $i$,
+	the number of isolated nodes decrease by one or two.
+	Then, once the process terminates,
+	there are at least $\ell+1$ isolated nodes left.
+
+	  Next, we argue that a partition exists when the process terminates.
+	 This implies that a feasible partition exists after any earlier step as well.
+%	 Assume for contradiction that the partition is not feasible.	 
+	 Since there are at least $\ell+1$ isolated nodes left,
+	 there must be two isolated nodes $x^*$ and $y^*$,
 	  with the same cluster of origin,
-	 i.e., $I(\{u\}) = I(\{v\})$.
-	 Consider the partition $P$
+	 i.e., $I(\{x^*\}) = I(\{y^*\})$.
+	 Consider the partition $P^*$
 	  obtained from the initial partition after swapping $x_0$ and $y_0$
@@
 	 In this partition,
-	 the ground set $\{x_0,y_0\}$ is collocated in the cluster $I(\{u,v\})$.
+	 the ground set $\{x_0,y_0\}$ is collocated in the cluster $I(\{x^*,y^*\})$.
 	 Note that after the first request $\{x_0,y_0\}$,
 	   we issue requests only between nodes that have the same cluster of origin and
-	   all these nodes are collocated in $P$.
-	 Therefore all ground sets constructed so far are collocated in  $P$,
-	 i.e., $P$ is a feasible partition,
-	 which contradicts our assumption.
-	 Hence, there are at most $\ell$ isolated nodes after the termination.
-
-	The number of isolated nodes decreases after each step.
-	Consider the last step, $j$, when exactly $\ell+1$ isolated nodes exist before  $x_{j}$ joins to another ground set.
-	That is, there are less than $\ell+1$ isolated nodes after the step $j$.
-	Then, at the end of step $j-1$,
-	there exist two isolated nodes $x^*$ and $y^*$
-	with the same cluster of origin, i.e., $I(\{x^*\})=I(\{y^*\})$.
+	   all these nodes are collocated in $P^*$.
+	 Therefore all ground sets constructed so far are collocated in  $P^*$,
+	 i.e., $P^*$ is a feasible partition.
+%	 Hence, there are at most $\ell$ isolated nodes after the termination.
+
+%	The number of isolated nodes decrease after each step.
+%	Consider the last step, $j$, when  $\ell+1$ isolated nodes exist before  $x_{j}$ joins to another ground set.
+%	That is, there are less than $\ell+1$ isolated nodes after the step $j$.
+%	Then, at the end of step $j-1$,
+%	there exist two isolated nodes $x^*$ and $y^*$
+%	with the same cluster of origin, i.e., $I(\{x^*\})=I(\{y^*\})$.
 %
-   Next,
-    we bound costs incurred by \OPT and \ALG over the  sequence of requests
-    $\sigma := \langle \{x_0,y_0\}, \dots, \{x_{j-1},y_{j-1}\}  \rangle$.
-
-	The optimal strategy is to perform two node swaps.
+%   Next,
+%    we bound costs incurred by \OPT and \ALG over the  sequence of requests
+%    $\sigma := \langle \{x_0,y_0\}, \dots, \{x_{j-1},y_{j-1}\}  \rangle$.
+
+	Consider  nodes $x^*$ and $y^*$ and the partition $P^*$ obtained previously.
+	\OPT moves to $P^*$ by performing only two node swaps.
 	Precisely, \OPT collocates $\set{x_0,y_0}$ 
 	by swapping them with $x^*$ and $y^*$.
-	In this partition, no ground set is split and
-%	there is no request to $x^*$ or $y^*$ in $\sigma$.
-	therefore \OPT pays only for the two swaps.
+	No ground set is split in $P^*$ and
+	 \OPT pays only for the two swaps.
 
 	\ALG performs at least one swap each time a~ground set grows,
-	i.e., each time it joins the isolated node $x_i$ (collocated with $B$) to  the ground set $C_i$ as specified before.
+	i.e., at each step $i$.
 %	In other words,
 %	some ground set grows by one node at each of these steps.
-	Consider any ground set $C^*$ after serving $\sigma$.
-	This ground set has grown exactly $|C^*|-1$ times until the end of $\sigma$. 
-	Let $\mathcal{S}$ be the set of all ground sets after serving the last request in $\sigma$.
+	Consider any ground set $C^* \neq B$ after the termination.
+	This ground set has grown exactly $|C^*|-1$ times until the the termination. 
+	Let $\mathcal{S}$ be the set of all ground sets after the process terminates.
 	Thus, $\mathcal{S}$ includes ground sets $B$,
 	$\{x_0,y_0\}$,
-	and $\ell+1$ singleton ground sets.
+	and (up to) $\ell+2$ singleton ground sets.
 	Among the remaining ground sets in $\mathcal{S}$,
@@
    the smaller ground set is either a singleton,
-   which contradicts the bound $\ell+1$ on the number of singletons,
+   which contradicts the bound $\ell+2$ on the number of singletons,
 	or we have joined nodes to it at some step,
-	contradicting our choice of the largest ground set of the same origin.
+	contradicting our choice of the largest $C_i$ at step $i$.
 	Hence, every  other ground set in
-	$\mathcal{S}$ has grown \emph{maximally} over $\sigma$.
+	$\mathcal{S}$ has grown \emph{maximally}.
 	There are at most $\ell-1$ maximally grown ground sets,
@@
 	 excluding the cluster containing $B$. 
-	Therefore,  $|\mathcal{S}| \leq 1 + 1 + (\ell+1)  + (\ell-1) = 2\ell+2$.
+	Therefore,  $|\mathcal{S}| \leq 1 + 1 + (\ell+2)  + (\ell-1) = 2\ell+3$.
 	Note that among all non-singleton ground sets in $\mathcal{S}$,
@@
 		\sum_{C^* \in \mathcal{S} } 1 - (k-1) \\
-	&\geq  k  \ell - (2\ell+2) - (k-1) =  (k-2)(\ell-1) - 3,
+	&\geq  k  \ell - (2\ell+3) - (k-1) =  (k-2)(\ell-1) - 4,
 	\end{align*}
@@
 	The competitive ratio is then
-	$\ALG / \OPT \geq ((k-2)(\ell-1) - 3) / 2$.
+	$\ALG / \OPT \geq ((k-2)(\ell-1) - 4) / 2$.
 	%%

### 2020-05-14 paper/disc20clustering.tex
wip

--- 2020-05-14 paper/disc20clustering.tex
@@
 
-	Next, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
-	Fix a~partition for \OPT prior serving the sequence $\sigma_E$.
-	Even without migrations, \OPT may mitigate paying for requests between at most $3\ell$ pairs of nodes that can be collocated in its clusters.
-	Recall that $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), and the counter of an unsaturated pair is at most $\alpha - 1$.
-	Hence, $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
-	Finally, during the phase, \OPT is faced with at least $\chi := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests from $\sigma_E$ that it cannot mitigate.
-
-	By migrating nodes, \OPT may decrease its cost, and pay less than $\chi$ for requests from $\sigma_E$.
+	Next, we bound $\OPT(\sigma_E) + \OPT(M)$.
+	The sequence $\sigma_E$ accounts only for unsaturated edges, thus there are at most $\alpha-1$ requests to each pair in $\sigma_E$.
+	\OPT may have at most $3\ell$ pairs of nodes collocated in its clusters, and thus avoid paying for $3\ell\cdot(\alpha-1)$ requests from $\sigma_E$.
+	Hence, at least $\chi := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests from $\sigma_E$ are external requests with respect to \OPT's configuration at the beginning of the phase.
+	Faced with these requests, \OPT may serve them remotely or perform migrations to decrease its cost.
 	By swapping a~pair of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
 	This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterward.
-	As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
-	By performing a~single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$
-	Hence, for serving at least $\chi$ unavoidable requests from $\sigma_E$, $\OPT$ pays at least
+	Hence, by performing a~single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
+	The total cost of \OPT is then at least
 	%
@@
 	\]
-	%	
-	To bound the competitive ratio, we transform the above inequality in the following way: $|\sigma_E| \leq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
+
+	Finally, to bound the competitive ratio, we transform the above inequality in the following way: $|\sigma_E| \leq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
 	For succintness, let $\xi := \OPT(\sigma_E) + \OPT(M)$.

### 2020-05-14 paper/disc20clustering.tex
cleanup

--- 2020-05-14 paper/disc20clustering.tex
@@
 	Since \ALG does not have $\{x_0,y_0\}$ collocated, 
-	the adversary issues external requests to this pair until \ALG collocates it.
+	the adversary issues an external request to this pair so that \ALG collocates it.
 	\ALG cannot collocate $\{x_0, y_0\}$ with $B$ (as $B$'s size is $k-1$), hence it collocates them in a~different cluster.
@@
 
-	We proceed by joining the current pivot node to a~larger ground set in  similar steps.
+	We proceed in similar steps by joining the current pivot node to a ground set of the same origin residing in a different cluster.
 	Consider the step $i$,
 	when the isolated node $x_i$ is collocated with $B$.
-	 We issue a request between $x_i$ and some node $y_i \in C_i$,
+	 We issue a request between $x_i$ and some node in $C_i$,
 	 where $C_i$ is the largest ground set
@@
 	the number of isolated nodes decrease by one or two.
-	Then, once the process terminates,
+	Therefore, once the process terminates,
 	there are at least $\ell+1$ isolated nodes left.
@@
 	  obtained from the initial partition after swapping $x_0$ and $y_0$
-	 with $u$ and $v$ (respectively).
+	 with $x^*$ and $y^*$ (respectively).
 	 In this partition,

### 2020-05-14 paper/disc20clustering.tex
one line LB formula
minor

@@
 
-	  Next, we argue that a partition exists when the process terminates.
+	  Next, we argue that a feasible partition exists when the process terminates.
 	 This implies that a feasible partition exists after any earlier step as well.
@@
 	a ground set in $\mathcal{S}$  has grown is
+	$	\sum_{C^* \in \mathcal{S} } (|C^*|-1) - (k-1) $
 	\begin{align*}
-	\sum_{C^* \in \mathcal{S} } (|C^*|-1) - (k-1) &= 
-	\sum_{C^* \in \mathcal{S} } |C^*| - 
-		\sum_{C^* \in \mathcal{S} } 1 - (k-1) \\
-	&\geq  k  \ell - (2\ell+3) - (k-1) =  (k-2)(\ell-1) - 4,
+	=\sum_{C^* \in \mathcal{S} } |C^*| - 
+		\sum_{C^* \in \mathcal{S} } 1 - (k-1) 
+	\geq  k  \ell - (2\ell+3) - (k-1) =  (k-2)(\ell-1) - 4,
 	\end{align*}

### 2020-05-14 paper/disc20clustering.tex
rewording
Comments

@@
 	let $I(C)$ denote the cluster,
-	i.e., the cluster that is the \emph{origin} of $C$.
+	i.e., the cluster that is the \emph{origin} of $C$ \maciek{This sentence requires a rewrite}.
 %	If $C$ contains nodes that originate from different clusters

### 2020-05-14 paper/disc20clustering.tex
removing maximally

--- 2020-05-14 paper/disc20clustering.tex
@@
 
-	\ALG performs at least one swap each time a~ground set grows,
-	i.e., at each step $i$.
+	\ALG performs at least one swap at each step $i$,
+	and some ground set grows.
 %	In other words,
@@
 	contradicting our choice of the largest $C_i$ at step $i$.
-	Hence, every  other ground set in
-	$\mathcal{S}$ has grown \emph{maximally}.
-	There are at most $\ell-1$ maximally grown ground sets,
+	Hence,  there are at most $\ell-1$ such grown ground sets,
 	one per cluster,

### 2020-05-14 paper/disc20clustering.tex
minor
removing maximally (1)
cluster of origin

--- 2020-05-14 paper/disc20clustering.tex
@@
 	For a ground set $C$ of nodes that are initially collocated in one cluster,
-	let $I(C)$ denote the cluster,
-	i.e., the cluster that is the \emph{origin} of $C$.
+	let $I(C)$ denote the cluster.
+	We refer to $I(C)$ as the cluster of \emph{origin},
+	when $C$ is clear from the context.
 %	If $C$ contains nodes that originate from different clusters
@@
 	contradicting our choice of the largest $C_i$ at step $i$.
-	Hence,  there are at most $\ell-1$ such grown ground sets,
-	one per cluster,
+	Hence,  there are at most $\ell-1$ such ground sets,
+	one per possible cluster of origin,
 	 excluding the cluster containing $B$. 

### 2020-05-14 paper/disc20clustering.tex
Unfinished phase

--- 2020-05-14 paper/disc20clustering.tex
@@
 	 where $C_i$ is the largest ground set
-	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$.
+	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$ \maciek{Ground sets merges instead of requests}.
 %	 only if  a feasible partition exists
@@
 	At each step $i$,
-	the number of isolated nodes decrease by one or two.
+	the number of isolated nodes decrease by one or two \maciek{Is this clear why? I am not, sure, take a look and decide. Possibly we need to say about two singletons merge}.
 	Therefore, once the process terminates,
@@
 	contradicting our choice of the largest $C_i$ at step $i$.
-	Hence,  there are at most $\ell-1$ such grown ground sets,
+	Hence, every  other ground set in
+	$\mathcal{S}$ has grown \emph{maximally}.
+	There are at most $\ell-1$ maximally grown ground sets,
 	one per cluster,
@@
 	Finally, to bound the competitive ratio, we transform the above inequality in the following way: $|\sigma_E| \leq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
-	For succintness, let $\xi := \OPT(\sigma_E) + \OPT(M)$.
+	For succinctness, let $\xi := \OPT(\sigma_E) + \OPT(M)$.
 	Combining the bounds on the cost of \TAlg and \OPT during each finished phase, the competitive ratio is
@@
 	If the input finished before the first $\alpha$ external requests, then \TAlg is $1$-competitive.
-	If at least $\alpha$ external requests were issued, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a~migration; for the cost of \TAlg we follow the analysis of the unsaturated pairs.
-	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
+	If at least $\alpha$ external requests were issued, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a~migration.
+	Charging this cost to \OPT serves the purpose of charging $\alpha$ at the end of a finished phase, and thus we can repeat the analysis of a finished phase.
+	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and we repeat the analysis of a finished phase.
 	This way, the competitive ratio increases at most twofold in comparison to a~finished phase, and the competitive ratio is $\TAlg(\sigma)/\OPT(\sigma) \leq 60\ell$.

### 2020-05-14 paper/disc20clustering.tex

@@
 	For a ground set $C$ of nodes that are initially collocated in one cluster,
-	let $I(C)$ denote the cluster,
-	i.e., the cluster that is the \emph{origin} of $C$ \maciek{This sentence requires a rewrite}.
+	let $I(C)$ denote the cluster.
+	We refer to $I(C)$ as the cluster of \emph{origin},
+	when $C$ is clear from the context.
 %	If $C$ contains nodes that originate from different clusters
@@
 	contradicting our choice of the largest $C_i$ at step $i$.
-	Hence, every  other ground set in
-	$\mathcal{S}$ has grown \emph{maximally}.
-	There are at most $\ell-1$ maximally grown ground sets,
-	one per cluster,
+	Hence,  there are at most $\ell-1$ such ground sets,
+	one per possible cluster of origin,
 	 excluding the cluster containing $B$. 

### 2020-05-14 paper/disc20clustering.tex
Distributed
We do not have OBRP definition

@@
 In the following, we show how to efficiently find a~fresh cluster in a~distributed manner.
-We organize the clusters into an arbitrary rooted balanced binary tree, and we broadcast the root to each cluster.
+To this end, we organize the clusters into an arbitrary rooted balanced binary tree, and we broadcast the root to each cluster.
 Each cluster maintains the counter of fresh clusters in its subtree.
@@
 we present \RM,
- an algorithm for the \OBRP restricted to clusters of capacity $k=2$.
+ an algorithm for clusters of capacity $k=2$.
 We interpret a~pair of nodes collocated in one cluster as a~``matched'' pair.

### 2020-05-14 paper/disc20clustering.tex
comments

@@
 Rematching is necessary for  maximizing intra-cluster communications,
-which is dual to minimizing inter-cluster communications.
+which is dual \maciek{dual has specific meaning. Is it correct here? I think you meant "equivalent" here} to minimizing inter-cluster communications.
 This is known as the  \emph{online rematching} 
 problem and a~non-strict 7-competitive algorithm is already given by~\cite{repartition-disc},
-in which the ratio comes with an additive factor in $O(\alpha\ell^2)$.
+in which the ratio comes with an additive factor $O(\alpha\ell^2)$.
 We do not only improve upon their competitive ratio,

### 2020-05-14 paper/disc20clustering.tex
comments

@@
 while our analysis is significantly  simpler and more concise,
-thanks to the charging scheme we devise here.
+thanks to the charging scheme we devise here \maciek{Way too strong. At best, analysis is slightly simpler. Just compare volumes and the notation needed.}.
 

### 2020-05-14 paper/disc20clustering.tex
wip

@@
 	Dashed lines represent requests.
-	An arrow, e.g.~from $b_1$ to $a_1$ means  that $b_1$ replaces $a_1$. }
+	An arrow, e.g.~from $b_1$ to $a_1$ means  that $b_1$ replaces $a_1$.\maciek{Maybe: "An arrow from $a$ to $b$ means that $b$ replaces $a$.". Or "$b$ migrates in place of $a$."} }
 	\label{fig:cyclic}
@@
 % before they inflict too much communication cost.
- Therefore it pays the price of only one migration per pair  (see Figure \ref{fig:cyclic}).
-Hence,
-whenever \OPT collocates a~pair,
-we charge it only the cost of moving a~single node to the cluster of the other,
-i.e., $\alpha$ (in contrast to $2\alpha$ on \RM).
+ Thus it pays the price of only one migration per pair  (see Figure \ref{fig:cyclic}).
+Whenever \OPT collocates a~pair,
+we charge it only the cost $\alpha$ of moving a~single node to the other cluster (in contrast to the cost $2\alpha$ incurred by \RM).
 

### 2020-05-14 paper/disc20clustering.tex
wip

@@
  Thus it pays the price of only one migration per pair  (see Figure \ref{fig:cyclic}).
-Whenever \OPT collocates a~pair,
+For this reason, whenever \OPT collocates a~pair,
 we charge it only the cost $\alpha$ of moving a~single node to the other cluster (in contrast to the cost $2\alpha$ incurred by \RM).

### 2020-05-14 paper/disc20clustering.tex
wip

@@
 However,
-we can charge the whole cost to \OPT only the first time \RM collocates a~pair,
+we can charge the whole cost \maciek{Whole cost unclear, I think} to \OPT only the first time \RM collocates a~pair,
 and not at any consequent time when \RM collocates it a~second time.

### 2020-05-14 paper/disc20clustering.tex
comments

@@
 $\sigma:=\{\sigma_1,\dots, \sigma_m\}$.
-	Assume on request $\sigma_{t}=\set{u,v}$, 
-	i.e., at (time) $t$,
-	\RM collocates the pair $\set{u,v}$.
-	This means,
-	the value of $C_{\set{u,v}}$ at $t$,
-	denoted by $C^t_{\set{u,v}}$, 
-	reaches $\lambda$ immediately before \RM resets the counter,
-	i.e.,
-	$ C^t_{\set{u,v}} = \lambda$.
-	We denote the set of all requests to a~pair $\set{x,y}$ that arrive
-	during $[t_1,t_2]$ by $\sigma_{\set{x,y}}[t_1,t_2]$,
-	and its overall cost to \OPT by
-	$\OPT (\sigma_{\{u,v\}}[t_1,t_2])$.
+	Assume that \RM collocates a pair $\set{u,v}$ at time $t$.
+	The value of $C_{\set{u,v}}$ at $t$,
+	denoted $C^t_{\set{u,v}}$, 
+	reaches $\lambda$ immediately before \RM resets the counter.
+	For any interval $[t_1,t_2]$, by $\sigma_{\set{x,y}}[t_1,t_2]$ we denote the set of all requests to a~pair $\set{x,y}$ that arrive
+	during $[t_1,t_2]$.
+	%and its overall cost to \OPT by
+	%$\OPT (\sigma_{\{u,v\}}[t_1,t_2])$.
 	We may use $\sigma_{\set{x,y}}$ whenever
-	 the interval $[t_1,t_2]$ is clear from the context.
+	 the interval $[t_1,t_2]$ is clear from the context \maciek{As a reader I would be horrified seeing this sentence. Better to fix it in the paragraph}.
 	%

### 2020-05-14 paper/disc20clustering.tex
comments

@@
 	We may use $\sigma_{\set{x,y}}$ whenever
-	 the interval $[t_1,t_2]$ is clear from the context \maciek{As a reader I would be horrified seeing this sentence. Better to fix it in the paragraph}.
+	 the interval $[t_1,t_2]$ is clear from the context \maciek{As a reader I would be horrified seeing this sentence, expecting shenanigans. Better to fix it in the paragraph}.
 	%

### 2020-05-14 paper/disc20clustering.tex
wip

@@
 	it is not the first collocation and
-	consider times $t'$ and $t''$ as define previously,
+	consider times $t'$ and $t''$ as defined previously,
 	 and let 

### 2020-05-14 paper/disc20clustering.tex
wip

@@
 	it must pay $\alpha$ for collocating the other pair, in this case $\{u,w\}$,
-	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$. 
+	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$ \maciek{Rewrite without these resp.s}. 
 	Therefore in any case,
@@
 	separating $\{u,v\}$.
-	We define times $t'' < t' < m$ similarly to the former case.
+	We define times $t'' < t' < m$ similarly to the former case \maciek{doing thing "similarly" is a trap zone in proofs, and we should avoid even suggesting it.
+I think that here you mean defining exactly the same.}.
 	Let $R^*_{\{u,v\}} := \sigma_{\{u,w\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.

### 2020-05-14 paper/disc20clustering.tex
wip

@@
 	it must pay $\alpha$ for collocating the other pair, in this case $\{u,w\}$,
-	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$ \maciek{Rewrite without these resp.s}. 
+	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$ \maciek{Rewrite without these resp.s. Also, see the comment on "or (and)" below}. 
 	Therefore in any case,
@@
 	Then,
-	\OPT must pay for collocating at least one pair or (and) serving requests 
+	\OPT must pay for collocating at least one pair or (and) \maciek{"or (and)" usually should be "and possibly". Does this make sense here?} serving requests 
 	to the other pair remotely.

### 2020-05-14 paper/disc20clustering.tex
wip

@@
 	and  
-	$\OPT (R_{t'}) + \OPT(R^*_{\{u,v\}})$.
+	$\OPT (R_{t'}) + \OPT(R^*_{\{u,v\}})$ \maciek{Propose how to formalize (but up to you).
+I think this is classic bound of maximum by an average argument.
+We actually use it in this paper already in \TAlg, line 407 at the time of writing.}.
 	Hence,

### 2020-05-14 paper/disc20clustering.tex
some comments done

--- 2020-05-14 paper/disc20clustering.tex
@@
 	 where $C_i$ is the largest ground set
-	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$ \maciek{Ground sets merges instead of requests}.
+	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$.
+	 Then \ALG must collocate the new ground set
+	  $\{x_i\} \cup C_i $ in one cluster.
 %	 only if  a feasible partition exists
@@
 	At each step $i$,
-	the number of isolated nodes decrease by one or two \maciek{Is this clear why? I am not, sure, take a look and decide. Possibly we need to say about two singletons merge}.
+	the number of isolated nodes decrease either by one,
+	 or by two if $C_i$ is a singleton.
 	Therefore, once the process terminates,
-	there are at least $\ell+1$ isolated nodes left.
+	in any case at least $\ell+1$ isolated nodes are left.
 
@@
 each node constitutes a single-node component which we refer to as a \emph{singleton} component, 
-and the contained node is an \emph{isolated} node \maciek{unclear: isolated = contained?}.
+and the node in such component is an \emph{isolated} node.
 We define larger components in terms of smaller components.
@@
 we move the whole component that contains it.
-A~partition that maintains this invariant is a~\emph{component respecting} partition \maciek{wrong use of "this".
-Better to define component respecting, and only then say that ALG only uses such partitions (but it is also implied in the learning model...)}.
+A~partition that has every component collocated 
+is a~\emph{component respecting} partition.
 %Intuitively, a component preserves the history of requests between its nodes.
@@
 In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).
-A generic sketch of these two algorithms can be found in the appendix (Algorithm \ref{alg:ppl}) \maciek{We don't have the generic. Also in the next paragraph you refer to the code}.
+%A generic sketch of these two algorithms can be found in the appendix (Algorithm \ref{alg:ppl}).
 
@@
 \textbf{Perfect Partition Learner algorithm.}
-\maciek{There should be a short description of the algorithm: what steps it takes, how it treats the input.
-Instead, we have PPL's properties here, and the algorithm description is not present at all.}
 Fix the initial partition
@@
 at least $\Delta(P)$ node migrations are required in order to reach the partition $P$ from $P_I$, and thus
-$\OPT \geq \Delta^*:= \Delta(P_F) $ \maciek{Confusing. We have space. First define :=, then say that OPT bounds}.
-%We  use $ \Delta(P_F) $ wherever the initial partition is clear from the context
+$\OPT \geq \Delta(P_F) $.
+
+Now we describe the algorithm \PPL.
+On each inter-cluster request $\{u,v\}$, 
+\PPL creates a new components by merging
+ the two components that contain nodes $u$ and $v$.
+In order to collocate nodes of the new component,
+\PPL moves to a~component respecting partition that minimizes the distance to the initial partition $P_I$.
 The scheme of the algorithm can be found in the  appendix (Algorithm \ref{alg:ppl}).
 
-With each repartitioning \maciek{Unclear what repartitioning means. Maybe with each merge action?},
-\PPL moves to a~component respecting partition that minimizes the distance to the initial partition $P_I$.
-As a~result,
-\PPL never moves to a~partition that is more than $\Delta^*$  node migrations away from $P_I$.
+\PPL never moves to a~partition that is more than $ \Delta^*:= \Delta(P_F) $   migrations away from $P_I$.
 This invariant latter ensures us that \PPL does not pay too much while recovering $P_F$.
-We emphasize that a~repartitioning by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
+We emphasize that a~\emph{repartitioning} by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
 This way \PPL never moves to a~partition beyond the distance $\Delta^*$.
@@
 \begin{proof}
-	\maciek{Maybe: "On each inter-cluster request, the algorithm enumerates all component respecting $\ell$-way partitions in the order of increasing distance to $P_I$" Then we don't need to say "i.e." and two sentences that explain the same thing}
 	On each inter-cluster request,
@@
 Rematching is necessary for  maximizing intra-cluster communications,
-which is dual \maciek{dual has specific meaning. Is it correct here? I think you meant "equivalent" here} to minimizing inter-cluster communications.
+which is equivalent to minimizing inter-cluster communications.
 This is known as the  \emph{online rematching} 
@@
 while our analysis is significantly simpler and more concise,
-thanks to the charging scheme we devise here \maciek{Way too strong. At best, analysis is slightly simpler. Just compare volumes and the notation needed.}.
+thanks to the charging scheme we devise here.
 
@@
 However,
-we can charge the whole cost \maciek{Whole cost unclear, I think} to \OPT only the first time \RM collocates a~pair,
+we can charge this cost to \OPT only the first time \RM collocates a~pair,
 and not at any consequent time when \RM collocates it a~second time.
@@
 	We may use $\sigma_{\set{x,y}}$ whenever
-	 the interval $[t_1,t_2]$ is clear from the context \maciek{As a reader I would be horrified seeing this sentence, expecting shenanigans. Better to fix it in the paragraph}.
+	 the interval $[t_1,t_2]$ is clear from the context.
 	%
@@
 	it must pay $\alpha$ for collocating the other pair, in this case $\{u,w\}$,
-	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$ \maciek{Rewrite without these resp.s}. 
+	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$. 
 	Therefore in any case,
@@
 	separating $\{u,v\}$.
-	We define times $t'' < t' < m$ similarly to the former case \maciek{doing thing "similarly" is a trap zone in proofs, and we should avoid even suggesting it.
-I think that here you mean defining exactly the same.}.
+	We define times $t'' < t' < m$ analogously to the former case.
 	Let $R^*_{\{u,v\}} := \sigma_{\{u,w\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.

### 2020-05-14 paper/disc20clustering.tex
correcting ground set def

@@
 We provide a~lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm for the learning problem.
-Later we elaborate on how to efficiently transform it to a~lower bound for the general partitioning problem.
+Later, we elaborate on how to efficiently transform it to a~lower bound for the general partitioning problem.
 The lower bound requires $k\geq 3$.
 In contrast, for $k=2$ the learning problem is trivial: immediate collocation of communicating pairs is $1$-competitive.
-(A partitioning problem for $k=2$ is non-trivial, a~lower bound of $3$ is known~\cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)
+In contrast,
+the general partitioning problem for $k=2$ is non-trivial (see Section~\ref{sec:k2}).
+%, a~lower bound of $3$ is known~\cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)
 
 Across this paper, we often refer to groups of communicating nodes.
-We use this concept slightly differently in the lower bound and the upper bounds.
-In our algorithms, we group nodes with a~communication history into \emph{components}, and in the lower bound, we group nodes that may communicate into \emph{ground sets}.
-
-A~\emph{ground set} is a group of nodes that repeatedly communicate if they are not collocated.
-If an algorithm splits any pair of nodes from a ground set, then the adversary continues to produce requests to this pair until the algorithm collocates it \maciek{learning model: does not need to continue issuing requests.}
-Any competitive algorithm must maintain a~perfect partition of ground sets into clusters.
-We emphasize that ground sets are revealed only upon their nodes are split.
+We use this concept slightly differently in the lower bound than the upper bounds.
+In our algorithms, we group nodes with a~communication history into \emph{components}.
+In this section,
+we group nodes that may ever communicate, into \emph{ground sets}.
+
+Given a perfect partition,
+every subset of nodes that belong to the same cluster in this partition
+is a~\emph{ground set}. 
+On each inter-cluster request a ground set is revealed.
+Any competitive algorithm under the learning model,
+ maintains a~perfect partition of ground sets into clusters.
+ An algorithm recovers the (hidden) perfect partition gradually over inter-cluster requests,
+  by merging pairs of ground sets involved in these requests.
+
 The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
 Once we construct a ground set, it lasts until the end of the input sequence.
-We say that a ground set is \emph{singleton} if it contains exactly one node.
-We say that the node that belongs to a singleton ground set is \emph{isolated}.
+We say that a ground set is \emph{singleton} if it contains exactly one node,
+that is an \emph{isolated} node.
 

### 2020-05-14 backup/disc20clustering.tex
stefan iteration

--- 2020-05-14 paper/disc20clustering.tex
@@
 
- \author{Mahmoud Parham}{mahmoud.parham@univie.ac.at}{Faculty of Computer Science, University of Vienna, Austria}{0000-0002-6211-077X}{}
-
- \author{Stefan Schmid}{stefan\_schmid@univie.ac.at}{Faculty of Computer Science, University of Vienna, Austria}{}{}
+% \author{Mahmoud Parham}{mahmoud.parham@univie.ac.at}{Faculty of Computer Science, University of Vienna, Austria}{0000-0002-6211-077X}{}
+
+% \author{Stefan Schmid}{stefan\_schmid@univie.ac.at}{Faculty of Computer Science, University of Vienna, Austria}{}{}
 
@@
 
-\subsection{Related work}
+\subsection{Related Work}
 
@@
 A straightforward analysis of the algorithm from Section~\ref{sec:k3} results in the bound $O(\ell^2)$, and the analysis must be improved in two places.
-One must bound the cost of each of $O(\ell)$ reconfigurations in~phase by a~constant, and show that the optimum offline algorithm must pay a~significant cost for inter-cluster requests.
+One must bound the cost of each of $O(\ell)$ reconfigurations per~phase by a~constant, and show that the optimum offline algorithm must pay a~significant cost for inter-cluster requests.
 We bound the cost of the latter by estimating the capabilities of the optimum offline algorithm to \emph{prepare} for an incoming sequence of requests.
 Furthermore, in the analysis of the algorithm from Section~\ref{sec:k2}, we propose a~novel charging scheme for edges that share a~vertex.
-
-We emphasize the implications of the tight result for the learning model.
-If one hopes to improve the lower bound for the general model, the ground sets must be dynamically added and \emph{removed}: if ground set are only added, the instance is solvable by the algorithm from Section~\ref{sec:ppl} within the cost $O(k\cdot \ell)$.
+We also emphasize the implications of the tight result for the learning model.
+If one hopes to improve the lower bound for the general model, the ground sets must be dynamically added and \emph{removed}: if ground sets are only added, the instance is solvable by the algorithm from Section~\ref{sec:ppl} within the cost $O(k\cdot \ell)$.
 
@@
 
-Across this paper, we often refer to groups of communicating nodes.
+Throughout this paper, we often refer to groups of communicating nodes.
 We use this concept slightly differently in the lower bound than the upper bounds.
@@
 is a~\emph{ground set}. 
-On each inter-cluster request a ground set is revealed.
 Any competitive algorithm under the learning model,
  maintains a~perfect partition of ground sets into clusters.
+ On each inter-cluster request a ground set is revealed.
  An algorithm recovers the (hidden) perfect partition gradually over inter-cluster requests,
@@
 \textbf{Resource augmentation.}
-The majority of work on Online Balanced Partitioning so far~\cite{repartition-disc,sigmetrics19_partitioning} focuses on the scenario with resource augmentation, where the clusters of an online algorithm are larger than the clusters of the offline optimal algorithm that we compare the performance to.
+The majority of work on the online balanced partitioning problem so far~\cite{repartition-disc,sigmetrics19_partitioning} focuses on the scenario with resource augmentation, where the clusters of an online algorithm are larger than the clusters of the offline optimal algorithm that we compare the performance to.
 We can adjust our construction to show a~lower bound of $\Omega(\ell)$ for a~setting with resource augmentation.
@@
 
+\stefan{I think we can delete this paragraph: now we discussed everything
+in the individual sections, no?}
 Although we have described our algorithms globally so far, we note that these allow for efficient distributed implementations. 

### 2020-05-14 paper/disc20clustering.tex
correcting ground set def (1)
tikz

@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{figs/cyclic}
+\begin{minipage}{.5\textwidth}
+	\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1.25,xscale=1.25]
+	\useasboundingbox  (-3mm,-5mm) rectangle (0mm,10mm);
+	%cluster 1
+	\def\x{30}
+	\def\w{15}
+	%Shape: Rectangle
+	\draw   (0,0) -- (15,0) -- (15,40) -- (0,40) -- cycle ;
+	%Straight Lines
+	\draw   (0,20) -- (15,20) ;
+	% Text Node
+	\draw (8,10) node (b1)  {$b_{1}$} ;
+	% Text Node
+	\draw (8,30) node (a4)   {$a_{4}$};
+
+	%cluster 2
+	\draw   (\x,0) -- (\w+\x,0) -- (\w+\x,40) -- (\x,40) -- cycle ;
+	\draw   (\x,20) -- (\w+\x,20) ;
+	% Text Node
+	\draw (\x+8,10) node (b2)   {$b_{2}$};
+	% Text Node
+	\draw (\x+8,30) node (a1)   {$a_{1}$};
+
+	\draw[dotted, shorten <= -6pt, shorten >= -6pt] (a4)  -- (a1);
+	\draw [->,blue, shorten <= -5pt, shorten >= -2pt] (a4) -- (b2);
+
+	%cluster 3
+	\draw   (2*\x,0) -- (\w+2*\x,0) -- (\w+2*\x,40) -- (2*\x,40) -- cycle ;
+	\draw    (2*\x,20) -- (\w+2*\x,20) ;
+	% Text Node
+	\draw (2*\x+8,10) node (b3)   {$b_{3}$};
+	% Text Node
+	\draw (2*\x+8,30) node (a2)   {$a_{2}$};
+	\draw[dotted, shorten <= -4pt, shorten >= -3pt] (b2)  -- (b3);
+	\draw [->,blue, shorten <= -5pt, shorten >= -2pt] (b2) -- (a2);
+
+	%cluster 4
+	\draw   (3*\x,0) -- (\w+3*\x,0) -- (\w+3*\x,40) -- (3*\x,40) -- cycle ;
+	\draw    (3*\x,20) -- (\w+3*\x,20) ;
+	% Text Node
+	\draw (3*\x+8,10) node (b4)   {$b_{4}$};
+	% Text Node
+	\draw (3*\x+8,30) node (a3)   {$a_{3}$};
+	\draw[dotted, shorten <= -4pt, shorten >= -3pt] (a2)  -- (a3);
+	\draw [->,blue, shorten <= -5pt, shorten >= -2pt] (a2) -- (b4);
+
+	\draw[dotted,  shorten <= -3pt, shorten >= -5pt]  (b4) to[out=-100,in=-80] (b1);
+	% arrow
+	\draw[->, blue,shorten <= -3pt, shorten >= -4pt]  (b4) 
+	to [out=-108,in=-130, distance=17mm]  (a4);
+	\end{tikzpicture}
+\end{minipage}\hfill
+%
+\begin{minipage}{.5\textwidth}
+	\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1.25,xscale=1.25]
+	\useasboundingbox  (0mm,-5mm) rectangle (0mm,10mm);
+%cluster 1
+\def\x{30}
+\def\w{15}
+%Shape: Rectangle
+\draw   (0,0) -- (15,0) -- (15,40) -- (0,40) -- cycle ;
+%Straight Lines
+\draw   (0,20) -- (15,20) ;
+% Text Node
+\draw (8,10) node (b1)  {$b_{1}$} ;
+% Text Node
+\draw (8,30) node (b4)   {$b_{4}$};
+
+%cluster 2
+\draw   (\x,0) -- (\w+\x,0) -- (\w+\x,40) -- (\x,40) -- cycle ;
+\draw   (\x,20) -- (\w+\x,20) ;
+% Text Node
+\draw (\x+8,10) node (a4)   {$a_{4}$};
+% Text Node
+\draw (\x+8,30) node (a1)   {$a_{1}$};
+
+%cluster 3
+\draw   (2*\x,0) -- (\w+2*\x,0) -- (\w+2*\x,40) -- (2*\x,40) -- cycle ;
+\draw    (2*\x,20) -- (\w+2*\x,20) ;
+% Text Node
+\draw (2*\x+8,10) node (b3)   {$b_{3}$};
+% Text Node
+\draw (2*\x+8,30) node (b2)   {$b_{2}$};
+
+%cluster 4
+\draw   (3*\x,0) -- (\w+3*\x,0) -- (\w+3*\x,40) -- (3*\x,40) -- cycle ;
+\draw    (3*\x,20) -- (\w+3*\x,20) ;
+% Text Node
+\draw (3*\x+8,10) node (a2)   {$a_{2}$};
+% Text Node
+\draw (3*\x+8,30) node (a3)   {$a_{3}$};
+\end{tikzpicture}
+\end{minipage}
 	\caption{
-		\OPT  collocates 5 pairs, paying one migration per pair.
 			Dashed lines represent requests.
-	An arrow, e.g.~from $b_1$ to $a_1$ means  that $b_1$ replaces $a_1$.\maciek{Maybe: "An arrow from $a$ to $b$ means that $b$ replaces $a$.". Or "$b$ migrates in place of $a$."} }
+		An arrow from node $x$ to node $y$ indicates that $x$  replaces $y$.
+		\OPT collocates 4 pairs from the left partition, with  4 migrations,
+		 resulting in the right partition.
+	}
 	\label{fig:cyclic}

### 2020-05-14 paper/disc20clustering.tex
hspace

@@
 	\centering
-\begin{minipage}{.5\textwidth}
-	\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1.25,xscale=1.25]
+\begin{minipage}[b]{.5\textwidth}
+	\hspace{15mm}
+	\begin{tikzpicture}[x=0.8pt,y=0.75pt,yscale=-1.25,xscale=1.25]
 	\useasboundingbox  (-3mm,-5mm) rectangle (0mm,10mm);
@@
 %
-\begin{minipage}{.5\textwidth}
-	\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1.25,xscale=1.25]
+\begin{minipage}[b]{.5\textwidth}
+	\begin{tikzpicture}[x=0.8pt,y=0.75pt,yscale=-1.25,xscale=1.25]
 	\useasboundingbox  (0mm,-5mm) rectangle (0mm,10mm);

### 2020-05-15 paper/disc20clustering.tex
hspace
new tikz, commented

@@
  \usepackage{amsmath}
- \usetikzlibrary{patterns} %
+ \usetikzlibrary{decorations.pathreplacing} %
  \usepackage{algorithm}
@@
 	\centering
-\begin{minipage}[b]{.5\textwidth}
+\begin{minipage}[b]{.55\textwidth}
 	\hspace{15mm}
@@
 %
-\begin{minipage}[b]{.5\textwidth}
+\begin{minipage}[b]{.45\textwidth}
 	\begin{tikzpicture}[x=0.8pt,y=0.75pt,yscale=-1.25,xscale=1.25]
@@
 
+%\section{Additional Figures}
+\begin{comment}
+\begin{figure}[H]	\label{fig:timeline}
+	\centering
+\begin{tikzpicture}
+\draw [->] (-10mm,0) -- (5,0);
+\draw [dotted] (-13mm,0) -- (-10mm,0);
+
+\draw [decorate,decoration={brace,amplitude=3pt}] (5mm,0.15) -- (45mm,0.15)
+node [black,midway,above=1pt,xshift=1pt] {$R_{t}$};
+
+\draw [decorate,decoration={brace,amplitude=3pt}, yshift=-18pt] (25mm,0.15) -- (-7mm,0.15)
+node [black,midway,below=1pt,xshift=1pt] {$R_{t'}$};
+
+\draw [] (45mm,2pt) -- (45mm,-2pt)
+node [black,midway,below=0pt,xshift=1pt] {$t$};
+
+% tick
+\draw [] (25mm,2pt) -- (25mm,-2pt)
+node [black,midway,below=0pt,xshift=1pt] {$t'$};
+% tick
+\draw [] (5mm,2pt) -- (5mm,-2pt)
+node [black,midway,below=0pt,xshift=1pt] {$t''$};
+% tick
+\draw [] (-7mm,2pt) -- (-7mm,-2pt)
+node [black,midway,below=0pt,xshift=1pt] {};
+
+\draw [line width=0.33mm, red] (5mm,0) -- (25mm,0) 
+node [black,midway,below=-2pt,xshift=1pt] {$\sigma_{\{u,w\}}$};
+
+\draw [] (25mm,0) -- (45mm,0) 
+node [black,midway,below=-2pt,xshift=1pt] {$\sigma_{\{u,v\}}$};
+
+\draw [] (5mm,2pt) -- (5mm,-2pt)
+node [black,midway,below=0pt,xshift=1pt] {$t''$};
+%% vertical arrow 
+%\draw [->] (3mm,14pt) -- (4.4mm,4pt)
+%node [black,midway,above=0pt,xshift=-4pt] {separation of  $\{u,w\}$};
+%% vertical arrow 
+%\draw [->] (30mm,-24pt) -- (26.5mm,-7pt)
+%node [black,midway,below=0pt,xshift=5pt, yshift=-5pt] {separation of  $\{u,v\}$};
+\end{tikzpicture}
+\caption{
+Illustration of the timeline used in the proof of Theorem \ref{thm:k=2}.
+Requests in $R_t$  are only to  $\{u,w\}$ and $\{u,v\}$,
+which  arrive  during the interval $(t'',t]$.
+Similarly,
+request in $R_{t'}$  are to $\{u,w\}$ and some other pair, irrelevant to the analysis.
+Hence, requests to $\{u,w\}$ are included in at most two such sets,
+which are $R_{t'}$ and $R_t$.
+This is because their intervals overlap on $(t'',t']$, shown with red thick line.
+}
+\end{figure}
+\end{comment}
+
 \end{document}

### 2020-05-15 paper/disc20clustering.tex
Finalizations

@@
 	 Therefore all ground sets constructed so far are collocated in  $P^*$,
-	 i.e., $P^*$ is a feasible partition.
+	 and it is a feasible partition.
 %	 Hence, there are at most $\ell$ isolated nodes after the termination.
@@
 	or we have joined nodes to it at some step,
-	contradicting our choice of the largest $C_i$ at step $i$.
+	contradicting our choice of the largest~$C_i$ at step $i$.
 	Hence,  there are at most $\ell-1$ such ground sets,
@@
 we choose a~component respecting partition closest to the initial partition.
-In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).
+In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a~repartition of minimum cost).
 %A generic sketch of these two algorithms can be found in the appendix (Algorithm \ref{alg:ppl}).
@@
 On each inter-cluster request $\{u,v\}$, 
-\PPL creates a new components by merging
+\PPL creates new components by merging
  the two components that contain nodes $u$ and $v$.
@@
 
-\PPL never moves to a~partition that is more than $ \Delta^*:= \Delta(P_F) $   migrations away from $P_I$.
+\PPL never moves to a~partition that is more than $ \Delta^*:= \Delta(P_F) $   migrations away from~$P_I$.
 This invariant latter ensures us that \PPL does not pay too much while recovering $P_F$.
@@
 
-\stefan{I think we can delete this paragraph: now we discussed everything
-in the individual sections, no?}
-Although we have described our algorithms globally so far, we note that these allow for efficient distributed implementations. 
+Our algorithms allow for efficient distributed implementations. 
 The algorithm PPL from Section~\ref{sec:ppl} can be distributed
@@
 %\section{Additional Figures}
-\begin{comment}
+%\begin{comment}
 \begin{figure}[H]	\label{fig:timeline}
@@
 \end{figure}
-\end{comment}
+%\end{comment}
 

### 2020-05-15 paper/disc20clustering.tex
Grammar & PPL

@@
 by dynamically migrating nodes (e.g., virtual machines) which communicate frequently topologically closer to each other, network traffic can be reduced significantly.  
-However, migrations entail overhead, and should be used moderately.
+However, migrations entail overhead and should be used moderately.
 
@@
 is a~\emph{ground set}. 
-Any competitive algorithm under the learning model,
+Any competitive algorithm under the learning model
  maintains a~perfect partition of ground sets into clusters.
- On each inter-cluster request a ground set is revealed.
+ On each inter-cluster request, a ground set is revealed.
  An algorithm recovers the (hidden) perfect partition gradually over inter-cluster requests,
@@
 Once we construct a ground set, it lasts until the end of the input sequence.
-We say that a ground set is \emph{singleton} if it contains exactly one node,
-that is an \emph{isolated} node.
+We say that a ground set is a~\emph{singleton} if it contains exactly one node,
+which is an \emph{isolated} node.
 
@@
 	At each step $i$,
-	the number of isolated nodes decrease either by one,
+	the number of isolated nodes decreases either by one
 	 or by two if $C_i$ is a singleton.
@@
 	Consider any ground set $C^* \neq B$ after the termination.
-	This ground set has grown exactly $|C^*|-1$ times until the the termination. 
+	This ground set has grown exactly $|C^*|-1$ times until the termination. 
 	Let $\mathcal{S}$ be the set of all ground sets after the process terminates.
@@
 
+\medskip
+
 \noindent
 \textbf{Perfect Partition Learner algorithm.}
+Now we describe the algorithm \PPL.
+On each inter-cluster request $\{u,v\}$, 
+\PPL creates new components by merging
+ the two components that contain nodes $u$ and $v$.
+In order to collocate nodes of the new component,
+\PPL moves to a~component respecting partition that minimizes the distance to the initial partition $P_I$.
+The scheme of the algorithm can be found in the  appendix (Algorithm \ref{alg:ppl}).
+
+\medskip
+
 Fix the initial partition
@@
 $\OPT \geq \Delta(P_F) $.
-
-Now we describe the algorithm \PPL.
-On each inter-cluster request $\{u,v\}$, 
-\PPL creates new components by merging
- the two components that contain nodes $u$ and $v$.
-In order to collocate nodes of the new component,
-\PPL moves to a~component respecting partition that minimizes the distance to the initial partition $P_I$.
-The scheme of the algorithm can be found in the  appendix (Algorithm \ref{alg:ppl}).
 

### 2020-05-15 paper/disc20clustering.tex
Grammar

@@
 In particular, pairs of nodes that did not reach the collocation threshold $\alpha$ (called external requests) incur the cost $O(\ell^2)$ for the algorithm in each phase.
-The novel part of the analysis lower-bounds the cost of \OPT on external requests, while considering its savings from migrations and possibly different configuration at the beginning of the phase.
+The novel part of the analysis lower-bounds the cost of \OPT on external requests while considering its savings from migrations and possibly different configuration at the beginning of the phase.
 This way, we show that \OPT paid a significant portion of the algorithm's cost on external requests.
@@
 \begin{proof}
-	If no component respecting partition exists after the merge of components, then \TAlg resets all components, ends the phase and performs no repartition.
+	If no component respecting partition exists after the merge of components, then \TAlg resets all components, ends the phase, and performs no repartition.
 	It suffices to show that the merged component has size at least $4$ to conclude that \TAlg incurs no cost.
@@
 	By $\sigma$ we denote the input sequence that arrived during the phase.
-	In our analysis, we focus on the requests that were external to \TAlg at the moment of their arrival; these are the only requests that incur cost for \TAlg.
+	In our analysis, we focus on the requests that were external to \TAlg at the moment of their arrival; these are the only requests that incur a~cost for \TAlg.
 	We denote these external requests by $\sigma_{cost}$.
@@
 	If the input finished before the first $\alpha$ external requests, then \TAlg is $1$-competitive.
-	If at least $\alpha$ external requests were issued, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a~migration.
+	If at least $\alpha$ external requests were issued, then $\OPT$ either paid $\alpha$ for serving them remotely or paid $\alpha$ for a~migration.
 	Charging this cost to \OPT serves the purpose of charging $\alpha$ at the end of a finished phase, and thus we can repeat the analysis of a finished phase.
-	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and we repeat the analysis of a finished phase.
+	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into the last two phases, and we repeat the analysis of a finished phase.
 	This way, the competitive ratio increases at most twofold in comparison to a~finished phase, and the competitive ratio is $\TAlg(\sigma)/\OPT(\sigma) \leq 60\ell$.
@@
 \textbf{Distributed implementation.}
-While we have described the algorithm globally so far, we note that it allows for an efficient distributed implementations. 
+While we have described the algorithm globally so far, we note that it allows for efficient distributed implementations. 
 The algorithm performs two types of operations that require communication with other clusters: a~component merge, and a~broadcast of the end of the phase.
@@
 We charge both \OPT and \RM whenever \RM collocates a~pair.%(i.e., on a~collocation event).
-\RM collocates a~pair always with a~swap,%(i.e., two simultaneous migrations),
-which  costs $2\alpha$,
+\RM collocates a~pair always with a~swap %(i.e., two simultaneous migrations),
+(that costs $2\alpha$),
 while \OPT may save some costs by collocating multiple pairs at once.
@@
 	for  requests in $\sigma_{\set{u,v}}(t',m]$.
-    Then the overall cost to \RM is
+    Then the overall cost of \RM is
 $	\RM (\sigma)

### 2020-05-15 paper/disc20clustering.tex
small tikz chnage

@@
 
-	\draw[dotted, shorten <= -6pt, shorten >= -6pt] (a4)  -- (a1);
+	\draw[thick,dotted, shorten <= -6pt, shorten >= -6pt] (a4)  -- (a1);
 	\draw [->,blue, shorten <= -5pt, shorten >= -2pt] (a4) -- (b2);
@@
 	\draw (2*\x+8,30) node (a2)   {$a_{2}$};
-	\draw[dotted, shorten <= -4pt, shorten >= -3pt] (b2)  -- (b3);
+	\draw[thick,dotted, shorten <= -4pt, shorten >= -3pt] (b2)  -- (b3);
 	\draw [->,blue, shorten <= -5pt, shorten >= -2pt] (b2) -- (a2);
@@
 	\draw (3*\x+8,30) node (a3)   {$a_{3}$};
-	\draw[dotted, shorten <= -4pt, shorten >= -3pt] (a2)  -- (a3);
+	\draw[thick,dotted, shorten <= -4pt, shorten >= -3pt] (a2)  -- (a3);
 	\draw [->,blue, shorten <= -5pt, shorten >= -2pt] (a2) -- (b4);
 
-	\draw[dotted,  shorten <= -3pt, shorten >= -5pt]  (b4) to[out=-100,in=-80] (b1);
+	\draw[thick, dotted,  shorten <= -3pt, shorten >= -5pt]  (b4) to[out=-100,in=-80] (b1);
 	% arrow
@@
 %\section{Additional Figures}
-%\begin{comment}
+\begin{comment}
 \begin{figure}[H]	\label{fig:timeline}
@@
 \end{figure}
-%\end{comment}
+\end{comment}
 

### 2020-05-15 disc_paper/disc20clustering.tex
Journal

--- 2020-05-15 paper/disc20clustering.tex
@@
  \usepackage{enumitem}
+ \usepackage{standalone}
 
@@
 Ideally, this algorithm should perform close to an optimal offline algorithm, without requiring any information about future traffic demands.
+
+\pagebreak
 
@@
 
-\medskip
-
 The algorithm in this section and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from~\cite{repartition-disc}.
@@
 %A generic sketch of these two algorithms can be found in the appendix (Algorithm \ref{alg:ppl}).
-
-\medskip
 
@@
 The scheme of the algorithm can be found in the  appendix (Algorithm \ref{alg:ppl}).
-
-\medskip
 
@@
 
+	\pagebreak
+
 	It remains to consider the case where both $U$ and $V$ are of type $C_2$.
@@
 	%Let $O^E$ be the cost serving requests between nodes that \TAlg did not put in the same component during that phase.
+
+	\pagebreak
 
@@
 
-\begin{figure}[H]
-	\centering
-\begin{minipage}[b]{.55\textwidth}
-	\hspace{15mm}
-	\begin{tikzpicture}[x=0.8pt,y=0.75pt,yscale=-1.25,xscale=1.25]
-	\useasboundingbox  (-3mm,-5mm) rectangle (0mm,10mm);
-	%cluster 1
-	\def\x{30}
-	\def\w{15}
-	%Shape: Rectangle
-	\draw   (0,0) -- (15,0) -- (15,40) -- (0,40) -- cycle ;
-	%Straight Lines
-	\draw   (0,20) -- (15,20) ;
-	% Text Node
-	\draw (8,10) node (b1)  {$b_{1}$} ;
-	% Text Node
-	\draw (8,30) node (a4)   {$a_{4}$};
-
-	%cluster 2
-	\draw   (\x,0) -- (\w+\x,0) -- (\w+\x,40) -- (\x,40) -- cycle ;
-	\draw   (\x,20) -- (\w+\x,20) ;
-	% Text Node
-	\draw (\x+8,10) node (b2)   {$b_{2}$};
-	% Text Node
-	\draw (\x+8,30) node (a1)   {$a_{1}$};
-
-	\draw[thick,dotted, shorten <= -6pt, shorten >= -6pt] (a4)  -- (a1);
-	\draw [->,blue, shorten <= -5pt, shorten >= -2pt] (a4) -- (b2);
-
-	%cluster 3
-	\draw   (2*\x,0) -- (\w+2*\x,0) -- (\w+2*\x,40) -- (2*\x,40) -- cycle ;
-	\draw    (2*\x,20) -- (\w+2*\x,20) ;
-	% Text Node
-	\draw (2*\x+8,10) node (b3)   {$b_{3}$};
-	% Text Node
-	\draw (2*\x+8,30) node (a2)   {$a_{2}$};
-	\draw[thick,dotted, shorten <= -4pt, shorten >= -3pt] (b2)  -- (b3);
-	\draw [->,blue, shorten <= -5pt, shorten >= -2pt] (b2) -- (a2);
-
-	%cluster 4
-	\draw   (3*\x,0) -- (\w+3*\x,0) -- (\w+3*\x,40) -- (3*\x,40) -- cycle ;
-	\draw    (3*\x,20) -- (\w+3*\x,20) ;
-	% Text Node
-	\draw (3*\x+8,10) node (b4)   {$b_{4}$};
-	% Text Node
-	\draw (3*\x+8,30) node (a3)   {$a_{3}$};
-	\draw[thick,dotted, shorten <= -4pt, shorten >= -3pt] (a2)  -- (a3);
-	\draw [->,blue, shorten <= -5pt, shorten >= -2pt] (a2) -- (b4);
-
-	\draw[thick, dotted,  shorten <= -3pt, shorten >= -5pt]  (b4) to[out=-100,in=-80] (b1);
-	% arrow
-	\draw[->, blue,shorten <= -3pt, shorten >= -4pt]  (b4) 
-	to [out=-108,in=-130, distance=17mm]  (a4);
-	\end{tikzpicture}
-\end{minipage}\hfill
-%
-\begin{minipage}[b]{.45\textwidth}
-	\begin{tikzpicture}[x=0.8pt,y=0.75pt,yscale=-1.25,xscale=1.25]
-	\useasboundingbox  (0mm,-5mm) rectangle (0mm,10mm);
-%cluster 1
-\def\x{30}
-\def\w{15}
-%Shape: Rectangle
-\draw   (0,0) -- (15,0) -- (15,40) -- (0,40) -- cycle ;
-%Straight Lines
-\draw   (0,20) -- (15,20) ;
-% Text Node
-\draw (8,10) node (b1)  {$b_{1}$} ;
-% Text Node
-\draw (8,30) node (b4)   {$b_{4}$};
-
-%cluster 2
-\draw   (\x,0) -- (\w+\x,0) -- (\w+\x,40) -- (\x,40) -- cycle ;
-\draw   (\x,20) -- (\w+\x,20) ;
-% Text Node
-\draw (\x+8,10) node (a4)   {$a_{4}$};
-% Text Node
-\draw (\x+8,30) node (a1)   {$a_{1}$};
-
-%cluster 3
-\draw   (2*\x,0) -- (\w+2*\x,0) -- (\w+2*\x,40) -- (2*\x,40) -- cycle ;
-\draw    (2*\x,20) -- (\w+2*\x,20) ;
-% Text Node
-\draw (2*\x+8,10) node (b3)   {$b_{3}$};
-% Text Node
-\draw (2*\x+8,30) node (b2)   {$b_{2}$};
-
-%cluster 4
-\draw   (3*\x,0) -- (\w+3*\x,0) -- (\w+3*\x,40) -- (3*\x,40) -- cycle ;
-\draw    (3*\x,20) -- (\w+3*\x,20) ;
-% Text Node
-\draw (3*\x+8,10) node (a2)   {$a_{2}$};
-% Text Node
-\draw (3*\x+8,30) node (a3)   {$a_{3}$};
-\end{tikzpicture}
-\end{minipage}
-	\caption{
-			Dashed lines represent requests.
-		An arrow from node $x$ to node $y$ indicates that $x$  replaces $y$.
-		\OPT collocates 4 pairs from the left partition, with  4 migrations,
-		 resulting in the right partition.
-	}
-	\label{fig:cyclic}
-\end{figure}
-
 %\medskip
@@
 %%
+
+\begin{figure}[H]
+	\centering
+	\begin{minipage}[b]{.5\textwidth}	
+				\centering	
+		\includestandalone{figs/k=2_a}
+	\end{minipage}%
+	%
+	\begin{minipage}[b]{.5\textwidth}
+		\centering
+				\includestandalone{figs/k=2_b}
+	\end{minipage}
+	\caption{
+		Dashed lines represent requests.
+		An arrow from node $x$ to node $y$ indicates that $x$  replaces $y$.
+		\OPT collocates 4 pairs from the left partition, with  4 migrations,
+		resulting in the right partition.
+	}
+	\label{fig:cyclic}
+\end{figure}
 
@@
 
+\begin{figure}[t]
+	\centering
+	\includestandalone{figs/timeline}
+	\caption{
+		Illustration of the timeline used in the proof of Theorem \ref{thm:k=2}.
+		Requests in $R_t$  are only to  $\{u,w\}$ and $\{u,v\}$,
+		which  arrive  during the interval $(t'',t]$.
+		Similarly,
+		request in $R_{t'}$  are to $\{u,w\}$ and some other pair, irrelevant to the analysis.
+		Hence, requests to $\{u,w\}$ are included in at most two such sets,
+		which are $R_{t'}$ and $R_t$.
+		This is because their intervals overlap on $(t'',t']$, shown with red thick line.
+	}
+	\label{fig:timeline}
+\end{figure}
+
 	Next,
@@
 	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
-	We define $R_{t'}$ for the collocation at $t'$  analogously.
+	We define $R_{t'}$ for the collocation at $t'$  analogously
+	(see Figure \ref{fig:timeline}).
 	Then,
@@
 	and for final requests at $t=m$,
-	\OPT pays at least 
+	\OPT pays at~least 
 	$\frac{1}{2}(
@@
 
+\pagebreak
+
 \section{Pseudocode of Perfect Partition Learner}
-
-\begin{algorithm}
+%
+\begin{algorithm}[h]
 	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
@@
 
-%\section{Additional Figures}
-\begin{comment}
-\begin{figure}[H]	\label{fig:timeline}
-	\centering
-\begin{tikzpicture}
-\draw [->] (-10mm,0) -- (5,0);
-\draw [dotted] (-13mm,0) -- (-10mm,0);
-
-\draw [decorate,decoration={brace,amplitude=3pt}] (5mm,0.15) -- (45mm,0.15)
-node [black,midway,above=1pt,xshift=1pt] {$R_{t}$};
-
-\draw [decorate,decoration={brace,amplitude=3pt}, yshift=-18pt] (25mm,0.15) -- (-7mm,0.15)
-node [black,midway,below=1pt,xshift=1pt] {$R_{t'}$};
-
-\draw [] (45mm,2pt) -- (45mm,-2pt)
-node [black,midway,below=0pt,xshift=1pt] {$t$};
-
-% tick
-\draw [] (25mm,2pt) -- (25mm,-2pt)
-node [black,midway,below=0pt,xshift=1pt] {$t'$};
-% tick
-\draw [] (5mm,2pt) -- (5mm,-2pt)
-node [black,midway,below=0pt,xshift=1pt] {$t''$};
-% tick
-\draw [] (-7mm,2pt) -- (-7mm,-2pt)
-node [black,midway,below=0pt,xshift=1pt] {};
-
-\draw [line width=0.33mm, red] (5mm,0) -- (25mm,0) 
-node [black,midway,below=-2pt,xshift=1pt] {$\sigma_{\{u,w\}}$};
-
-\draw [] (25mm,0) -- (45mm,0) 
-node [black,midway,below=-2pt,xshift=1pt] {$\sigma_{\{u,v\}}$};
-
-\draw [] (5mm,2pt) -- (5mm,-2pt)
-node [black,midway,below=0pt,xshift=1pt] {$t''$};
-%% vertical arrow 
-%\draw [->] (3mm,14pt) -- (4.4mm,4pt)
-%node [black,midway,above=0pt,xshift=-4pt] {separation of  $\{u,w\}$};
-%% vertical arrow 
-%\draw [->] (30mm,-24pt) -- (26.5mm,-7pt)
-%node [black,midway,below=0pt,xshift=5pt, yshift=-5pt] {separation of  $\{u,v\}$};
-\end{tikzpicture}
-\caption{
-Illustration of the timeline used in the proof of Theorem \ref{thm:k=2}.
-Requests in $R_t$  are only to  $\{u,w\}$ and $\{u,v\}$,
-which  arrive  during the interval $(t'',t]$.
-Similarly,
-request in $R_{t'}$  are to $\{u,w\}$ and some other pair, irrelevant to the analysis.
-Hence, requests to $\{u,w\}$ are included in at most two such sets,
-which are $R_{t'}$ and $R_t$.
-This is because their intervals overlap on $(t'',t']$, shown with red thick line.
-}
-\end{figure}
-\end{comment}
-
 \end{document}

### 2020-05-18 podc_NEW/figs/timeline.tex
New BA

--- 2020-05-15 paper/disc20clustering.tex
@@
- \documentclass[a4paper,anonymous,USenglish]{lipics-v2019}
-
- \usepackage[utf8]{inputenc}
- \usepackage{xspace}
- \usepackage{balance}
- \usepackage{amsmath,amsfonts,mathtools,amsthm}
- \usepackage{algorithmic}
- \usepackage{algorithm}
-
- \usepackage{balance}
- \usepackage{amsthm,amsmath,array,colortbl,graphicx,multirow}
- \usepackage{comment}
- \usepackage{balance}
+\documentclass{standalone}
 \usepackage{tikz}
- \usepackage{amsmath}
  \usetikzlibrary{decorations.pathreplacing} %
- \usepackage{algorithm}
- \usepackage[font={footnotesize}]{subcaption}
- \usepackage[font={footnotesize}]{caption}
- \usepackage{breakcites}
- \usepackage{booktabs}
- \usepackage{diagbox}
- \usepackage{xcolor}
- \usepackage{colortbl}
- \usepackage{cleveref}
- \usepackage{enumitem}
-
- \mathchardef\mhyphen="2D
-
- \title{Optimal Algorithms for Dynamic Balanced Graph Partitioning}
-
- \author{Maciej Pacut}{maciej.pacut@univie.ac.at}{Faculty of Computer Science, University of Vienna,Austria}{0000-0002-6379-1490}{}
-
-% \author{Mahmoud Parham}{mahmoud.parham@univie.ac.at}{Faculty of Computer Science, University of Vienna, Austria}{0000-0002-6211-077X}{}
-
-% \author{Stefan Schmid}{stefan\_schmid@univie.ac.at}{Faculty of Computer Science, University of Vienna, Austria}{}{}
-
- \authorrunning{M. Pacut, M. Parham, and S. Schmid}
- \Copyright{Maciej Pacut, Mahmoud Parham, and Stefan Schmid}
- \keywords{online algorithms, competitive analysis, distributed computing, graph partitioning, clustering, self-adjusting networks}
-
-% \EventEditors{}
-%\EventNoEds{3}
- \EventLongTitle{34th International Symposium on DIStributed Computing (DISC) 2020}
- \EventShortTitle{DISC 2020}
- \EventAcronym{DISC}
- \EventYear{2020}
- \EventDate{October 12--16, 2020}
- \EventLocation{Freiburg, Germany (virtual conference)}
- \EventLogo{}
- \SeriesVolume{}
- \ArticleNo{}
-
- \ccsdesc[500]{Theory of computation~Online algorithms}
- \ccsdesc[500]{Networks~Network algorithms}
- %\ccsdesc[300]{Computer systems organization~Cloud computing}
- \ccsdesc[300]{Computer systems organization~Distributed architectures}
-
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%&&
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%&&
- %  our macros start
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%&&
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%&&
-
- \newcommand{\OPT}{\textsc{OPT}\xspace}
- \newcommand{\ALG}{\textsc{ALG}\xspace}
- \newcommand{\PPL}{\textsc{PPL}\xspace}
- \newcommand{\OBRP}{BRP\xspace}
- \newcommand{\PPOBRP}{PP-BRP}
- \newcommand{\dist}{\textsc{dist}}
- \newcommand{\TAlg}{{\ensuremath{\textsf{ALG}_{3}}}\xspace}
-  \newcommand{\RM}{\textsc{RM}\xspace} % rematching alg
-
- \newcommand{\Rep}{\textsc{Rep}}
-
- %\newtheorem{claim}{Claim}
- \newtheorem{fact}{Fact}
- \newtheorem{rem}{Remark}
- \newtheorem{observation}{Observation}
- \newtheorem{property}{Property}
-
- \DeclarePairedDelimiter\pair{(}{)}
- \DeclarePairedDelimiter\set{\{}{\}}
-
- \DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
- \DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
-
- \newcommand\mahmoud[1]{\color{orange}\textbf{Mahmoud: #1~}\color{black}}
- \newcommand\stefan[1]{\color{blue}\textbf{Stefan: #1}\color{black}}
- \newcommand\maciek[1]{\color{brown}\textbf{(Maciek: #1)}\color{black}}
- %\newcommand\mahmoud[1]{}
- %\newcommand\stefan[1]{}
- %\newcommand\maciek[1]{}
-
- \newcommand{\todo}[1]{\noindent\color{brown}{todo: #1}\color{black}}
-
- \begin{CCSXML}
-	<ccs2012>
-	<concept>
-	<concept_id>10003033.10003068</concept_id>
-	<concept_desc>Networks~Network algorithms</concept_desc>
-	<concept_significance>500</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10010520.10010521.10010537.10003100</concept_id>
-	<concept_desc>Computer systems organization~Cloud computing</concept_desc>
-	<concept_significance>300</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10010520.10010521.10010537</concept_id>
-	<concept_desc>Computer systems organization~Distributed architectures</concept_desc>
-	<concept_significance>300</concept_significance>
-	</concept>
-	</ccs2012>
-\end{CCSXML}
-
-% for my small screen
-%\setlength{\textwidth}{8cm}
 
 \begin{document}
-\maketitle
-
-\begin{abstract}
-	Distributed   applications,  including  batch  processing, streaming, scale-out databases,
-	or machine learning, generate a~significant amount of network traffic.
-	By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), we can reduce the network load and  improve application performance. 
-	However, the communication pattern of different applications is often unknown a~priori and may change over time, hence it needs to be learned in an~online manner.
-	%
-	This paper revisits the online 
-	balanced partitioning problem 
-	(introduced by Avin et al.~at DISC 2016)
-	that asks for an algorithm that strikes
-	an optimal tradeoff between the benefits
-	of collocation (i.e., lower network load) 
-	and its costs (i.e., migrations). 
-	%
-	Our first contribution is a~significantly improved deterministic
-	lower bound of $\Omega(k\cdot \ell)$ on the
-	competitive ratio, where $\ell$ is the number
-	of clusters and $k$ is the cluster size,
-	even for a~scenario in which the communication
-	pattern is static and can be perfectly partitioned;
-	we also provide an asymptotically tight upper bound 
-	of $O(k\cdot \ell)$ for this scenario.
-	For $k=3$, we contribute an asymptotically tight upper bound
-	of $\Theta(\ell)$
-	for the general model in which the
-	communication pattern can change arbitrarily over time.
-	We improve the result for $k=2$ by providing a~strictly $6$-competitive upper bound for the general model.
-	In contrast to most prior work, our algorithms respect all capacity constraints and do not require resource augmentation.
-
-\end{abstract}
-
-%\renewcommand{\shortauthors}{M.~Pacut, M.~Parham, S.~Schmid}
-
-\section{Introduction}
-
-The popularity of data-centric, distributed applications has led to an explosive growth of network traffic, especially in data centers~\cite{roy2015inside,singh2015jupiter}.
-The performance of these distributed applications often critically depends on the underlying network~\cite{mogul2012we}, and efficient operation of these networks is important.
-At the same time, distributed systems are often highly virtualized today, and provide interesting new opportunities for resource optimization.
-In particular, it has become possible to operate data centers in a~more demand-aware manner: 
-by dynamically migrating nodes (e.g., virtual machines) which communicate frequently topologically closer to each other, network traffic can be reduced significantly.  
-However, migrations entail overhead and should be used moderately.
-
-This paper studies the algorithmic problem underlying such demand-aware
-optimizations, aiming to strike a~balance between the benefits of migrations (e.g., reduced network load) and their costs.
-In particular, we are interested in an online variant of the problem: since communication patterns can change over time, an online algorithm needs to react dynamically to new traffic patterns, and migrate nodes  accordingly.
-Ideally, this algorithm should perform close to an optimal offline algorithm, without requiring any information about future traffic demands.
-
-This problem is known as the dynamic balanced graph partitioning problem and was introduced by Avin et al.~\cite{repartition-disc, sidma-arxiv} at DISC 2016.
-a~special variant of the general problem has later been studied by Henzinger et al.~\cite{sigmetrics19_partitioning} at SIGMETRICS 2019.
-We refer to the latter as the learning model.
-
-\subsection{Model}
-
-We study two models in this paper: the \emph{general partitioning} model, and its subproblem, the \emph{learning} model.
-In both models, we assume that communication patterns are not known to our algorithms at the beginning.
-We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.
-In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a~given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the entire input sequence in advance.
-%We emphasize that all our algorithms have a~\emph{strict} competitive ratio (without an additive term).
-
-\noindent
-\textbf{General partitioning model.}
-In the \emph{dynamic balanced graph partitioning} problem, we are given a~set $V$ of $n$ nodes 
-(e.g., virtual machines or processes),
-initially arbitrarily partitioned into $\ell$~clusters
-(e.g., servers or entire racks),
-each of size~$k$.
-The nodes interact using
-a~sequence of pairwise communication requests
-$\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
-where a~pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a~certain amount of data.
-Nodes in $C \subset V$ are \emph{collocated}
-if they reside in the same cluster.
-
-An algorithm serves a~communication request between two nodes
-either \emph{locally} at cost~0
-if they are collocated,
-or \emph{remotely} at cost~1
-if they are located in different clusters.
-We refer to these two types of requests as \emph{internal}
-and \emph{external} requests, respectively.
-Before serving a~request,
-an online algorithm may perform a~\emph{repartition},
-%(i.e., \emph{reconfigure}).
-i.e.,
-it may move (``migrate'') some nodes into clusters different from their current clusters, while respecting the capacity of every cluster. 
-Afterward, 
-the algorithm serves the  request.
-The cost of migrating a~node from one cluster to another
-is~$\alpha \in \mathbb{Z}^+$.
-For any algorithm $\ALG$,
-its cost,
-denoted by $\ALG(\sigma)$,
-is the total cost of communications and
-the cost of migrations performed by $\ALG$ while serving the sequence $\sigma$.
-
-\noindent
-\textbf{Learning model.}
-We study a~\emph{learning} variant of dynamic balanced graph partitioning,
-where the communication pattern is \emph{static}:
-whether a~pair of  nodes ever communicate or not, 
-is determined a~priori and is unknown to algorithms,
- and such pairs communicate forever.
-Any algorithm must eventually collocate  pairs of communicating nodes,
-as otherwise it cannot be competitive.
-As in Henzinger et al.~\cite{sigmetrics19_partitioning}, we assume that the communication graph admits a~\emph{perfect partition},
-i.e., a~partition in which no inter-cluster request ever occurs.
-%any algorithm must learn this perfect partition.
-The algorithm's objective is to \emph{learn} the (static) communication graph
- while serving all requests,
-and without executing too many migrations.
-For the learning model, we  assume that the migration cost is $\alpha=1$.
-
-\subsection{Related Work}
-
-The two works closest to ours are by Avin et al.
-(on the general partitioning model)~\cite{repartition-disc, sidma-arxiv} and by Henzinger et al.
-(on the learning model)~\cite{sigmetrics19_partitioning}.
-However, the focus of these papers is primarily on models with resource augmentation: the online algorithm can use slightly larger clusters than the offline algorithm.  
-Avin et al.
-actually showed that their lower bound $\Omega(k)$ holds even for a~significant resource augmentation, and they provided an algorithm with the competitive ratio $O(k \log k)$ using the $(2+\epsilon)$-augmented cluster capacity.
-Their ratio is independent of $\ell$, which is impossible without significant resource augmentation.
-
-In contrast, we study the non-augmented setting, where the nodes need to be perfectly balanced  among the clusters.
-This assumption is not only more realistic but also significantly more challenging, as it is related to hard problems such as integer partitioning~\cite{integer-partitions-book}.
-In terms of results without augmentation, so far, it is only known that there exists an~$O(k^2 \cdot \ell^2)$-competitive algorithm~\cite{repartition-disc}; the best known lower bound is significantly lower, namely $\Omega(k)$.
-For $k=2$, Avin et al.~\cite{repartition-disc} presented a~$7$-competitive algorithm with a~substantial ($\Omega(\ell^2)$) additive constant.
-
-The problem has also been studied in a~weaker
-model where the adversary can only sample
-requests from a~fixed distribution~\cite{stochastic-ring}.
-
-The static offline version of~the~partitioning~problem, i.e., a~problem variant where
-migration is not allowed, where all requests are known in advance, and where
-the goal is to find an assignment of $n$ nodes to $\ell$~physical machines, each of~capacity $n/\ell$, is known as the
-\emph{$\ell$-balanced graph partitioning problem}. The problem is 
-NP-complete, and cannot even be approximated within any finite factor unless P
-= NP~\cite{AndRae06}.  The static
-variant where $\ell = 2$ corresponds to the minimum bisection problem, which
-is already NP-hard~\cite{GaJoSt76}, and 
-the currently best approximation ratio is $O(\log n)$~\cite{SarVaz95,ArKaKa99,FeKrNi00,FeiKra02,KraFei06,Raec08}.
-
-Our problem is further related to some classic online problems.
-In particular, it is related to online paging~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}, sometimes also referred to
-as online caching, where requests for data items (nodes) arrive over time and
-need to be served from a~cache of finite capacity, and where the number of
-cache misses must be minimized. Classic problem variants usually boil down to
-finding a~smart eviction strategy, such as Least Recently Used (LRU)~\cite{SleTar85}. In our
-setting, requests can be served remotely (i.e.,~without fetching the
-corresponding nodes to a~single physical machine). In this light, our model is more
-reminiscent of caching models \emph{with
-bypassing}~\cite{EpImLN11,EpImLN15,Irani02}.
-A major difference between  these problems is that in the caching problems, each request involves a~single element of the universe, while in our model \emph{both} endpoints of a~communication request are subject to~optimization.
-In this light, we can see our model as a~"symmetric" version of online paging.
-
-Dynamic graph partitioning problems are generally fundamental in computer science, and arise in many different contexts~\cite{streaming-soda,streaming1}.
-
-%More generally, the model is related to online
-%caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00},
-%see~\cite{repartition-disc} for a~discussion.
-%The static offline version of~the~problem, called the
-%\emph{$\ell$-balanced graph partitioning problem} is 
-%NP-complete, and cannot even be approximated within %any finite factor unless P
-%= NP~\cite{AndRae06}.
-
-\subsection{Our Contributions}
-
-This paper presents several new results on the dynamic graph partitioning problem  without augmentation.
-For the learning model, we present a~lower bound of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
-(that holds also in the general partitioning model).
-The best known lower bound so far was $\Omega(k)$~\cite{repartition-disc} that holds only in the general partitioning model.
-We complement this result with an asymptotically optimal, 
-$O(k\cdot \ell)$-competitive algorithm
-for the learning model.
-
-For the general partitioning model, we design
-an asymptotically optimal,
-$\Theta(\ell)$-competitive algorithm for $k=3$, improving the best known upper bound 
-so far $O(\ell^2)$~\cite{repartition-disc}.
-We further present a~strictly $6$-competitive algorithm for $k=2$ that improves upon the previous $7$-competitive algorithm with $O(\alpha\ell^2)$ additive constant.
-
-All algorithms in this paper have a strict competitive ratio (i.e., without an additive term).
-Table \ref{tab:overview} provides an overview of our contributions compared to prior work.
-
-\begin{table*}[h]
-	\centering
-	\renewcommand{\arraystretch}{1.5}
-	\begin{tabular}{>{\centering\arraybackslash}p{4.25cm}|>{\centering\arraybackslash}p{4.25cm}>{\centering\arraybackslash}p{4.25cm}}
-		\rowcolor{gray!50}
-		\textbf{Variant} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
-		\textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & 6\hspace{0.3cm}(\S \ref{sec:k2}) \\ 
-		\rowcolor{gray!25}
-		\textbf{$k=3$}&  $\Omega(\ell)$ \hspace{0.3cm}(\S \ref{sec:lowerbound})& $O(\ell) $\hspace{0.3cm}(\S \ref{sec:k3})\\
-		$k > 3$ & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k^2 \cdot \ell^2)$\hspace{0.1cm}~\cite{repartition-disc} \\
-		\rowcolor{gray!25}
-		Learning model & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell)$\hspace{0.1cm} (\S \ref{sec:ppl}) \\
-	\end{tabular}
-	\caption{Overview of known results and our contributions.
-The table summarizes the results for the general partitioning model, except for the last row that summarizes the results for the learning model for arbitrary $k$ and $\ell$.
-	}
-	\label{tab:overview}
-	\vspace{-7mm}
-\end{table*}
-
-\medskip
-\noindent
-\textbf{Algorithmic techniques.} The variant for general $k$ is still unresolved, hence it is vital to summarize algorithmic techniques used in this paper.
-A straightforward analysis of the algorithm from Section~\ref{sec:k3} results in the bound $O(\ell^2)$, and the analysis must be improved in two places.
-One must bound the cost of each of $O(\ell)$ reconfigurations per~phase by a~constant, and show that the optimum offline algorithm must pay a~significant cost for inter-cluster requests.
-We bound the cost of the latter by estimating the capabilities of the optimum offline algorithm to \emph{prepare} for an incoming sequence of requests.
-Furthermore, in the analysis of the algorithm from Section~\ref{sec:k2}, we propose a~novel charging scheme for edges that share a~vertex.
-We also emphasize the implications of the tight result for the learning model.
-If one hopes to improve the lower bound for the general model, the ground sets must be dynamically added and \emph{removed}: if ground sets are only added, the instance is solvable by the algorithm from Section~\ref{sec:ppl} within the cost $O(k\cdot \ell)$.
-
-\section{The Learning Model} %$\Omega(k\cdot \ell)$ for Competitive Ratio of Any Deterministic Algorithm}
-
-In this section, we consider the learning variant of dynamic balanced graph partitioning problem.
-For this setting, we show a~surprisingly high lower bound of $\Omega(k \cdot \ell)$ for $k\geq 3$.
-The lower bound holds also in the general partitioning model (studied in Section~\ref{sec:part}).
-At the end of this section, we discuss an asymptotically optimal upper bound for the learning variant.
-
-\subsection{Lower Bound}
-
-\label{sec:lowerbound}
-
-We provide a~lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm for the learning problem.
-Later, we elaborate on how to efficiently transform it to a~lower bound for the general partitioning problem.
-The lower bound requires $k\geq 3$.
-In contrast, for $k=2$ the learning problem is trivial: immediate collocation of communicating pairs is $1$-competitive.
-In contrast,
-the general partitioning problem for $k=2$ is non-trivial (see Section~\ref{sec:k2}).
-%, a~lower bound of $3$ is known~\cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)
-
-Throughout this paper, we often refer to groups of communicating nodes.
-We use this concept slightly differently in the lower bound than the upper bounds.
-In our algorithms, we group nodes with a~communication history into \emph{components}.
-In this section,
-we group nodes that may ever communicate, into \emph{ground sets}.
-
-Given a perfect partition,
-every subset of nodes that belong to the same cluster in this partition
-is a~\emph{ground set}. 
-Any competitive algorithm under the learning model
- maintains a~perfect partition of ground sets into clusters.
- On each inter-cluster request, a ground set is revealed.
- An algorithm recovers the (hidden) perfect partition gradually over inter-cluster requests,
-  by merging pairs of ground sets involved in these requests.
-
-The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
-Once we construct a ground set, it lasts until the end of the input sequence.
-We say that a ground set is a~\emph{singleton} if it contains exactly one node,
-which is an \emph{isolated} node.
-
-We start by constructing a~ground set of size $k-1$ on an arbitrarily chosen cluster.
-In any partition, there must exist an isolated node collocated with the ground set of size $k-1$.
-%, and we refer to it as the \emph{pivot} node.
-We issue requests between this node and some node that was initially collocated with it.
-By repeating such requests, almost every node is once collocated with the first ground set.
-In comparison, we show that there exists an optimal offline algorithm \OPT that performs only two node exchanges ("swaps").
-%Figure~\ref{fig:lb} illustrates the constructs used in the lower bound.
-
-\begin{theorem}
-	\label{th:lowerbound}
-	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least
-	 $(k-2)(\ell-1)/2 - 2$ for any $k\geq 3$ and $\ell \geq 2$.
-\end{theorem}
-
-\begin{proof}
-	Fix any online algorithm \ALG.
-	For a ground set $C$ of nodes that are initially collocated in one cluster,
-	let $I(C)$ denote the cluster.
-	We refer to $I(C)$ as the cluster of \emph{origin},
-	when $C$ is clear from the context.
-%	If $C$ contains nodes that originate from different clusters
-%	then $I(C)$ is undefined.
-%	Throughout the construction, we maintain ground sets for each node.
-	Initially, all nodes are isolated, i.e., each node is in a singleton ground set.
-	First, we choose a cluster arbitrarily and
-	 create a~ground set $B$ of $k-1$ nodes in this cluster.
-	%We never dismiss any constructed ground set.
-	Each cluster hosts exactly $k$ nodes, and in any feasible partition,
-	 a~single isolated node must be collocated with $B$.
-	At any time,
-	we refer to the isolated node currently collocated with $B$ as the \emph{pivot} node.
-	Let $x_0$ denote the first pivot node.
-
-	Then, we join the pivot node to a~larger ground set to force its eviction.
-	Precisely,
-	we create a~ground set $\{x_0, y_0\}$, 	where
-%	 $x_0$ is the current pivot node and
-	 $y_0$ is an arbitrary isolated node.
-	Since \ALG does not have $\{x_0,y_0\}$ collocated, 
-	the adversary issues an external request to this pair so that \ALG collocates it.
-	\ALG cannot collocate $\{x_0, y_0\}$ with $B$ (as $B$'s size is $k-1$), hence it collocates them in a~different cluster.
-	In order to preserve a~feasible partition of nodes after collocating $\{x_0, y_0\}$,
-	\ALG must replace $x_0$ with another isolated node that becomes the new pivot.
-%	denoted by $x_1$.
-
-	We proceed in similar steps by joining the current pivot node to a ground set of the same origin residing in a different cluster.
-	Consider the step $i$,
-	when the isolated node $x_i$ is collocated with $B$.
-	 We issue a request between $x_i$ and some node in $C_i$,
-	 where $C_i$ is the largest ground set
-	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$.
-	 Then \ALG must collocate the new ground set
-	  $\{x_i\} \cup C_i $ in one cluster.
-%	 only if  a feasible partition exists
-%	 that collocates all nodes in $\{x_i\} \cup C_i $.
-	Any feasible partition replaces $x_i$ with some isolated node $x_{i+1}$,
-	as the new ground set $\{x_i\} \cup C_i $  may not be ever split.
-%	Otherwise,
-%	no such feasible partition exists and
-%	we terminate the process.
-	We terminate the process once the number of remaining isolated nodes 
-	is less than $\ell+3$.
-	At each step $i$,
-	the number of isolated nodes decreases either by one
-	 or by two if $C_i$ is a singleton.
-	Therefore, once the process terminates,
-	in any case at least $\ell+1$ isolated nodes are left.
-
-	  Next, we argue that a feasible partition exists when the process terminates.
-	 This implies that a feasible partition exists after any earlier step as well.
-%	 Assume for contradiction that the partition is not feasible.	 
-	 Since there are at least $\ell+1$ isolated nodes left,
-	 there must be two isolated nodes $x^*$ and $y^*$,
-	  with the same cluster of origin,
-	 i.e., $I(\{x^*\}) = I(\{y^*\})$.
-	 Consider the partition $P^*$
-	  obtained from the initial partition after swapping $x_0$ and $y_0$
-	 with $x^*$ and $y^*$ (respectively).
-	 In this partition,
-	 the ground set $\{x_0,y_0\}$ is collocated in the cluster $I(\{x^*,y^*\})$.
-	 Note that after the first request $\{x_0,y_0\}$,
-	   we issue requests only between nodes that have the same cluster of origin and
-	   all these nodes are collocated in $P^*$.
-	 Therefore all ground sets constructed so far are collocated in  $P^*$,
-	 and it is a feasible partition.
-%	 Hence, there are at most $\ell$ isolated nodes after the termination.
-
-%	The number of isolated nodes decrease after each step.
-%	Consider the last step, $j$, when  $\ell+1$ isolated nodes exist before  $x_{j}$ joins to another ground set.
-%	That is, there are less than $\ell+1$ isolated nodes after the step $j$.
-%	Then, at the end of step $j-1$,
-%	there exist two isolated nodes $x^*$ and $y^*$
-%	with the same cluster of origin, i.e., $I(\{x^*\})=I(\{y^*\})$.
-%
-%   Next,
-%    we bound costs incurred by \OPT and \ALG over the  sequence of requests
-%    $\sigma := \langle \{x_0,y_0\}, \dots, \{x_{j-1},y_{j-1}\}  \rangle$.
-
-	Consider  nodes $x^*$ and $y^*$ and the partition $P^*$ obtained previously.
-	\OPT moves to $P^*$ by performing only two node swaps.
-	Precisely, \OPT collocates $\set{x_0,y_0}$ 
-	by swapping them with $x^*$ and $y^*$.
-	No ground set is split in $P^*$ and
-	 \OPT pays only for the two swaps.
-
-	\ALG performs at least one swap at each step $i$,
-	and some ground set grows.
-%	In other words,
-%	some ground set grows by one node at each of these steps.
-	Consider any ground set $C^* \neq B$ after the termination.
-	This ground set has grown exactly $|C^*|-1$ times until the termination. 
-	Let $\mathcal{S}$ be the set of all ground sets after the process terminates.
-	Thus, $\mathcal{S}$ includes ground sets $B$,
-	$\{x_0,y_0\}$,
-	and (up to) $\ell+2$ singleton ground sets.
-	Among the remaining ground sets in $\mathcal{S}$,
-	no two ground sets have the same origin.
-	Otherwise,
-   the smaller ground set is either a singleton,
-   which contradicts the bound $\ell+2$ on the number of singletons,
-	or we have joined nodes to it at some step,
-	contradicting our choice of the largest~$C_i$ at step $i$.
-	Hence,  there are at most $\ell-1$ such ground sets,
-	one per possible cluster of origin,
-	 excluding the cluster containing $B$. 
-	Therefore,  $|\mathcal{S}| \leq 1 + 1 + (\ell+2)  + (\ell-1) = 2\ell+3$.
-	Note that among all non-singleton ground sets in $\mathcal{S}$,
-	 only $B$ does not grow during the process.
-	 Thus,
-    the total number of times that 
-	a ground set in $\mathcal{S}$  has grown is
-	$	\sum_{C^* \in \mathcal{S} } (|C^*|-1) - (k-1) $
-	\begin{align*}
-	=\sum_{C^* \in \mathcal{S} } |C^*| - 
-		\sum_{C^* \in \mathcal{S} } 1 - (k-1) 
-	\geq  k  \ell - (2\ell+3) - (k-1) =  (k-2)(\ell-1) - 4,
-	\end{align*}
-	which bounds the number of swaps performed by \ALG.
-	The competitive ratio is then
-	$\ALG / \OPT \geq ((k-2)(\ell-1) - 4) / 2$.
-	%%
-	\begin{comment}
-	\ALG performs at least one swap  every time   a~ground set is constructed.
-	Each expansion reduces the number of isolated nodes by $1$ or $2$.
-	The decrease of two isolated nodes occurs when the largest component on  $I(x_i)$ is still singleton; this occurs at most $\ell$ times.
-	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
-	Hence, the total number of swaps of \ALG is at least $k\cdot \ell - k - 2 \ell$.
-	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - 2\ell) / 2$.
-	\end{comment}
-	%%
-\end{proof}
-
-\noindent
-\textbf{Lower bound for the general problem.}
-For a~lower bound $\Omega(k\cdot \ell)$ for the general partitioning problem, the adversary continues to issue requests to split nodes of a~ground set until the algorithm collocates them.
-Note that the ground sets constructed in our lower bound can be perfectly partitioned into the clusters.
-Hence, the optimal algorithm moves to a~perfect partition at the beginning (where requests incur the cost~$0$), and its cost is bounded.
-This means that the algorithm must eventually collocate all nodes of a~ground set to be competitive.
-We reveal the next ground set only after the collocation, hence we can repeat the analysis of the algorithm for the learning problem.
-Finally, we note that the construction is oblivious to the choice of the reconfiguration cost $\alpha$: we compare the 	number of node exchanges of \ALG and \OPT.
-
-\medskip
-
-\noindent
-\textbf{Resource augmentation.}
-The majority of work on the online balanced partitioning problem so far~\cite{repartition-disc,sigmetrics19_partitioning} focuses on the scenario with resource augmentation, where the clusters of an online algorithm are larger than the clusters of the offline optimal algorithm that we compare the performance to.
-We can adjust our construction to show a~lower bound of $\Omega(\ell)$ for a~setting with resource augmentation.
-
-Consider a~partitioning problem with resource augmentation $1+1/3-\epsilon$.
-Fix $k$ divisible by $3$, and construct $3$ ground sets of size $k/3$ in each cluster.
-Note that no more than $3$ such ground sets fit in one cluster.
-Then, apply the construction from the lower bound for $k=3$, using these ground sets in the way we used individual nodes.
-The cost of any algorithm (including \OPT) scales up by $k/3$, and the lower bound $\Omega(\ell)$ holds.
-
-Finally, we note the possibility of improvement. The algorithm CREP~\cite{repartition-disc} requires $(2+\epsilon)$-augmentation to guarantee the competitive ratio independent of $\ell$.
-In contrast, our construction shows that the linear term $\ell$ is inevitable if the augmentation is smaller than~$1+1/3$.
-
-\subsection{Upper Bound}
-\label{sec:ppl}
-
-We present an asymptotically optimal algorithm for the learning problem.
-The algorithm  collocates  a~pair as soon as they communicate and it never separates them.
-In order to preserve collocated pairs,
-we employ the concept of components,
-introduced by Avin~et~al.~\cite{repartition-disc}.
-
-We maintain subsets of frequently communicating nodes as \emph{components}.
-Initially,
-each node constitutes a single-node component which we refer to as a \emph{singleton} component, 
-and the node in such component is an \emph{isolated} node.
-We define larger components in terms of smaller components.
-Concretely,
-given a (sub)sequence of requests,
-two components $C_1$ and $C_2$
- \emph{merge} into one component as soon as
- for some pair of nodes $v_1 \in C_1$ and $v_2 \in C_2$,
-the frequency of requests $\{v_1,v_2\}$ reaches a certain threshold through the sequence.
-We keep all nodes of a~component always collocated in the same cluster,
-i.e., when we  move a~node,
-we move the whole component that contains it.
-A~partition that has every component collocated 
-is a~\emph{component respecting} partition.
-%Intuitively, a component preserves the history of requests between its nodes.
-
-In addition,
-we maintain a~balanced partition of our components as long as such partition exists,
-a reminiscent of partitioning given integers into sets of equal sum~\cite{integer-partitions-book}.
-In contrast, our partition is time-varying:  two components are merged into one component once they communicate, and we adjust the partition accordingly.
-
-\medskip
-
-The algorithm in this section and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from~\cite{repartition-disc}.
-The difference is in the choice of partition after a~component merge.
-In DET, the partition was arbitrary.
-In the algorithm for the learning model,
-we choose a~component respecting partition closest to the initial partition.
-In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a~repartition of minimum cost).
-%A generic sketch of these two algorithms can be found in the appendix (Algorithm \ref{alg:ppl}).
-
-\medskip
-
-\noindent
-\textbf{Perfect Partition Learner algorithm.}
-Now we describe the algorithm \PPL.
-On each inter-cluster request $\{u,v\}$, 
-\PPL creates new components by merging
- the two components that contain nodes $u$ and $v$.
-In order to collocate nodes of the new component,
-\PPL moves to a~component respecting partition that minimizes the distance to the initial partition $P_I$.
-The scheme of the algorithm can be found in the  appendix (Algorithm \ref{alg:ppl}).
-
-\medskip
-
-Fix the initial partition
-$P_I :=\{ I_1, \dots, I_{\ell}\}$ and \OPT's final partition
-$P_F := \{F_1, \dots, F_{\ell}\}$.
-The \emph{distance} of a~partition
- $P = \{C_1, \dots, C_{\ell}\}$ from the initial partition,
-defined as 
-$\Delta(P) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$,
- is the number of nodes in $P$ that do not reside in their initial cluster.
-In other words,
-at least $\Delta(P)$ node migrations are required in order to reach the partition $P$ from $P_I$, and thus
-$\OPT \geq \Delta(P_F) $.
-
-\PPL never moves to a~partition that is more than $ \Delta^*:= \Delta(P_F) $   migrations away from~$P_I$.
-This invariant latter ensures us that \PPL does not pay too much while recovering $P_F$.
-We emphasize that a~\emph{repartitioning} by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
-This way \PPL never moves to a~partition beyond the distance $\Delta^*$.
-
-\begin{property} \label{prop:dist<OPT}
-	Let $P$ be any partition chosen by \PPL at any time.
-	Then, $\Delta(P) \leq \Delta^*$.	
-\end{property}
-
-\begin{lemma}	\label{lemma:rebalancecost}
-	The cost of each repartitioning by \PPL is $2\cdot\OPT$.
-\end{lemma}
-\begin{proof}
-	Let $P_i$ denote the partition of \PPL immediately after serving $\sigma_{i}$.
-	Consider the repartitioning that transforms 
-	$P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
-	Let $M \subset V$ denote the set of nodes that migrate during this process.
-	Let $M^-$ and $M^+$ denote the subset of nodes that, respectively,
-	enter or leave their initial cluster during the repartitioning.    
-	Then,
-	$M = M^+ \cup M^-$.
-	Since at least $|M^-|$ nodes are not in their initial cluster before the repartitioning (i.e., in $P_{t-1}$),
-	the distance before the repartitioning is $\Delta(P_{t-1}) \geq | M^-|$.
-	Analogously,
-	the distance afterward is $\Delta(P_{t}) \geq | M^+|$.
-	Thus,
-	$|M| \leq \Delta(P_{t-1}) + \Delta(P_{t})$.
-	By Property \ref{prop:dist<OPT},
-	$\Delta(P_{t-1})  \leq \Delta^*$ and
-	$\Delta(P_{t}) \leq \Delta^*$.
-	Since $\Delta^* \leq \OPT$,
-	we obtain	 $|M| \leq 2\cdot\OPT$.
-\end{proof}
-
-\begin{theorem}	\label{thm:upperbound}
-	\PPL reaches the final partition $P_F$
-	 and it is $(2\cdot (k-1)\cdot\ell)$-competitive.
-\end{theorem}
-\begin{proof}
-	On each inter-cluster request,
-	the algorithm enumerates all component respecting $\ell$-way partitions of components
-	that are in the same (closest) distance to $P_I$.
-	That is, 
-	once it reaches a~partition $P$ at distance $\Delta^* = \Delta(P)$,
-	it does not move to a~partition
-	$P', \Delta(P') > \Delta^*$,
-	before it enumerates all partitions at distance $\Delta^*$.
-	Therefore,
-	\PPL eventually reaches  the partition $P_F$ at distance $\Delta^*=\OPT$.
-	With each distinct request, 
-	the size of some component increases by one.
-	For any cluster $F_i \in P_F$,
-	we have $\sum_{C \in F_i} |C| = k$.
-	A component $C \in F_i$,
-	 initially begins as an isolated node and it grows by gaining $|C|-1$ more nodes.
-	Hence, the total number of times a~component in $F_i$ grows is 
-	$\sum_{C \in F_i} (|C|-1) \leq k-1$.
-	Therefore, there are at most  $(k-1)\cdot\ell $ distinct requests
-	for which \PPL performs a~repartitioning and
-	 \PPL performs at most $(k-1)\cdot\ell $ repartitions.
-	By Lemma \ref{lemma:rebalancecost},
-	each repartitioning costs at most $2\cdot\OPT$.
-	The total cost is thus at most $2\cdot\OPT\cdot (k-1) \cdot\ell$, which implies the competitive ratio.
-\end{proof}
-
-% In Section~\ref{sec:lowerbound} we constructed a~$\Omega(k \cdot \ell)$ for \OBRP{}.
-% Note that the lower bound holds also in the perfect partition model, as the constructed input sequence allows \OPT to move to a~perfect partition.
-% The corollary is that PPL is optimal.
-
-\section{General Partitioning Model}
-\label{sec:part}
-
-Now  we discuss the general online
-model where the request sequence
-can be arbitrary.
-In Section~\ref{sec:k3}, we show an $O(k \cdot \ell)$-competitive algorithm for $k=3$ using the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental}.
-Prior to this section, we showed a~lower bound of $\Omega(k \cdot \ell)$  that holds for the general model (cf.
-Section~\ref{sec:lowerbound}), hence the result from this section is asymptotically optimal.
-Furthermore, in Section~\ref{sec:k2}, we show a~strictly $6$-competitive algorithm for $k=2$.
-
-\subsection{Optimal Algorithm for Clusters of Size 3}
-\label{sec:k3}
-
-The algorithm analyzed in this section is a modified version of the algorithm DET proposed by Avin et al.~\cite{repartition-disc}, which for $k=3$ is $O(\ell^2)$-competitive.
-In our algorithm, we choose the closest partition after a component merge instead of an arbitrary one.
-This allows to bound the cost of repartition by a constant (Lemma~\ref{lem:1req}).
-
-This modification alone is insufficient to obtain $O(\ell)$-competitive algorithm, and the analysis must be further improved.
-In particular, pairs of nodes that did not reach the collocation threshold $\alpha$ (called external requests) incur the cost $O(\ell^2)$ for the algorithm in each phase.
-The novel part of the analysis lower-bounds the cost of \OPT on external requests while considering its savings from migrations and possibly different configuration at the beginning of the phase.
-This way, we show that \OPT paid a significant portion of the algorithm's cost on external requests.
-
-\medskip
-
-\noindent
-\textbf{Component-based algorithm.}
-The algorithm \TAlg partitions nodes into components, and
-initially, each node is isolated (belongs to its own component).
-For each pair of nodes $\set{x,y}$, \TAlg maintains a~counter $C_{\set{x,y}}$ and increments it on every external request between $x$ and $y$.
-Once $C_{\set{x,y}} = \alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
-If no such partitioning exists, \TAlg resets all components to singleton components, resets all counters to $0$, and ends the phase.
-
-\begin{theorem}
-	\TAlg is $60\ell$-competitive for $k=3$.
-	\label{thm:k=3}
-\end{theorem}
-
-%\begin{figure}[H]
-%	\centering
-%	\includegraphics[width=0.3\textwidth]{figs/substitute}
-%	\caption{Cluster types used in the analysis of \TAlg.}
-%\end{figure}
-
-Before bounding the competitive ratio of \TAlg, we upper-bound the cost of a~single repartition of \TAlg.
-In our analysis, we distinguish among three types of clusters: $C_1, C_2$ and $C_3$. In a~cluster of type $C_i$, the size of the largest component contained in this cluster is~$i$.
-
-\begin{lemma}
-	\label{lem:1req}
-	In a~single repartition of nodes (after a~merge of components), \TAlg exchanges at most two pairs of nodes.
-\end{lemma}
-
-\begin{proof}
-	If no component respecting partition exists after the merge of components, then \TAlg resets all components, ends the phase, and performs no repartition.
-	It suffices to show that the merged component has size at least $4$ to conclude that \TAlg incurs no cost.
-
-	%We distinguish between three types of clusters: $C_1, C_2, C_3$,
-	%which we define as follows.
-	%A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial partition of any cluster).
-	%A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
-	%Finally, a~cluster of type $C_3$  contains one component of size $3$.
-
-	Consider a~request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
-	The request triggered the repartition, hence it was external and $U\neq V$.
-	We consider cases based on the types of clusters $U$ and $V$.
-
-	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the repartition is local within $U$ and $V$, for the~cost of at most $2$ swaps.
-	If either $U$ or $V$ is of type $C_3$, a~component of size $3$ participates in a~merge, and we have a~component of size at least $4$, and \TAlg ends the phase with no repartition.
-
-	It remains to consider the case where both $U$ and $V$ are of type $C_2$.
-	If $(u,v)$ both belong to components of size $2$, then the merged component has size $4$, and \TAlg incurs no cost. 
-	Otherwise, if one of $u,v$ belongs to a~component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
-	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a~cluster different from $U$ and $V$.
-	Note that if $C_1$-type cluster does not exist, then no component respecting partitioning exists.
-	Otherwise, \TAlg performs one swap --- it exchanges the nodes $u$ and $v$ with any two nodes~of~any cluster of type $C_1$.
-
-	In each case, we showed that a~component respecting partition is reachable in at most two swaps.
-\end{proof}
-
-%\begin{figure}[H]
-%	\centering
-%	\includegraphics[width=0.3\textwidth]{figs/substitute}
-%	\caption{an unsaturated edge that is inside a~component. If there is a~space next to it, maybe cluster types should fit next to it? Not sure.}
-%\end{figure}
-
-\begin{proof}[Proof of Theorem \ref{thm:k=3}]
-	Fix a~completed phase, and consider the state of \TAlg's counters at the end of it (before the reset).
-	We consider the incomplete phase later in this proof.
-
-	\TAlg is component respecting, hence it never increases any counter above $\alpha$.
-	We say that the pair $(u, v)$ is \emph{saturated} if the counter's value is $\alpha$, and \emph{unsaturated} otherwise (saturation of a~pair leads to a~merge action).
-	By $\sigma$ we denote the input sequence that arrived during the phase.
-	In our analysis, we focus on the requests that were external to \TAlg at the moment of their arrival; these are the only requests that incur a~cost for \TAlg.
-	We denote these external requests by $\sigma_{cost}$.
-	We partition the sequence $\sigma_{cost}$ into subsequences $\sigma_I$ and $\sigma_E$.
-	The sequence $\sigma_I$ (inter-component requests) denotes the requests from $\sigma_{cost}$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
-	The sequence $\sigma_E$ (extra-component requests) denotes the requests from $\sigma_{cost}$ that do not appear in $\sigma_I$.
-
-	%Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
-	%Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
-	Let $\TAlg(M)$ denote the cost of migrations performed by \TAlg in this phase.
-	During the phase, \TAlg performs at most $2 \ell$ component merge operations ---
-	exceeding this number would mean that a~component of size $4$ exists, and the phase should have ended already.
-	We bound the cost of each repartition after a~merge by Lemma~\ref{lem:1req}, obtaining $\TAlg(M) \leq 8\alpha\cdot\ell$.
-	%Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
-
-	We bound $\TAlg(\sigma_I)$ by summing the intra-component counters of each cluster at the end of the phase.
-	The sum of intra-component counters in a cluster of type $C_3$ is at most $3 \alpha - 1$: two pairs of nodes from the component are saturated and its counter is $\alpha$ each, and the counter of the third, unsaturated pair is at most $\alpha-1$.
-	The sum of counters inside $C_1$ is~$0$, and inside $C_2$ it is $\alpha$.
-	Summing over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq (3 \alpha-1)\cdot \ell \leq 3\alpha\cdot\ell$.
-
-	%We bound $A^E$ by $k^2 \cdot (\alpha - 1)$, as no more than $k^2$ pairs are unsaturated, and each of them contributes at most $\alpha -1$.
-	%\maciek{not needed most likely}
-
-	Furthermore, \TAlg paid for all requests from $\sigma_E$, and thus $\TAlg(\sigma_E) = |\sigma_E|$.
-	In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 11\alpha\cdot \ell + |\sigma_E|$ during this phase.
-
-	\medskip
-
-	Now we lower-bound the cost of the optimal offline solution.
-	To this end, we fix any optimal offline algorithm $\OPT$.
-	By $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$ we denote the cost of $\OPT$ on requests from sequences $\sigma_I$ and $\sigma_E$, respectively.
-	Note that these costs are defined with respect to components of \TAlg in this phase.
-	By $\OPT(M)$ we denote the cost of migrations performed by $\OPT$ in this phase.
-
-	The cost of \OPT is lower-bounded by the cost of serving $\sigma_I$ and the cost of serving $\sigma_E$.
-	While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts: we separately bound $\OPT$ by $\OPT(\sigma_I) + \OPT(M)$ and $\OPT(\sigma_E) + \OPT(M)$.
-	Combining those bounds and using the relation between the maximum and the average, we obtain the bound
-	\begin{align*}
-		\OPT&\geq \max\{\OPT(\sigma_I) + \OPT(M), \OPT(\sigma_E) + \OPT(M)\}\\
-		&\geq (\OPT(\sigma_I) + \OPT(M)) / 2 + (\OPT(\sigma_E) + \OPT(M)) / 2.
-	\end{align*}
-
-	%Let $O^M$ be the cost of migrations performed by $\OPT$ during the phase.
-	%Let $O^I$ be the cost of serving requests between nodes that were put in one component by \TAlg during this phase.
-	%Let $O^E$ be the cost serving requests between nodes that \TAlg did not put in the same component during that phase.
-
-	First, we show $\OPT(M) + \OPT(\sigma_I) \geq \alpha$.
-	Assume that \OPT's partition is fixed throughout the phase (as otherwise \OPT pays $\alpha$ for a migration).
-	The phase ended when the components of \TAlg{} could not be partitioned without splitting them.
-	Hence, for every possible partition of $\OPT$, there exists a~non-collocated saturated pair, and
-	$\OPT$ paid for $\alpha$ requests that saturated the pair.
-
-	Next, we bound $\OPT(\sigma_E) + \OPT(M)$.
-	The sequence $\sigma_E$ accounts only for unsaturated edges, thus there are at most $\alpha-1$ requests to each pair in $\sigma_E$.
-	\OPT may have at most $3\ell$ pairs of nodes collocated in its clusters, and thus avoid paying for $3\ell\cdot(\alpha-1)$ requests from $\sigma_E$.
-	Hence, at least $\chi := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests from $\sigma_E$ are external requests with respect to \OPT's configuration at the beginning of the phase.
-	Faced with these requests, \OPT may serve them remotely or perform migrations to decrease its cost.
-	By swapping a~pair of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
-	This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterward.
-	Hence, by performing a~single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
-	The total cost of \OPT is then at least
-	%
-	\[
-		\OPT(\sigma_E) + \OPT(M) \geq \chi \cdot \frac{2\alpha}{4 (\alpha-1)}\geq \frac{|\sigma_E|} {2} - 2 \alpha \cdot \ell.
-	\]
-
-	Finally, to bound the competitive ratio, we transform the above inequality in the following way: $|\sigma_E| \leq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
-	For succinctness, let $\xi := \OPT(\sigma_E) + \OPT(M)$.
-	Combining the bounds on the cost of \TAlg and \OPT during each finished phase, the competitive ratio is
-%
-	\[
-		\frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + \xi/2} \leq \frac{30\alpha\cdot\ell + 4\cdot \xi}{\alpha + \xi} \leq 30 \ell.
-	\]
-%	
-	\medskip
-
-	It remains to consider the last, unfinished phase.
-	First, consider the case, where the unfinished phase is also the first one.
-	Then, we cannot charge $\OPT$ due to the inability to partition the components.
-	Instead, we use the fact that \TAlg and $\OPT$ started with the same initial partition.
-	If the input finished before the first $\alpha$ external requests, then \TAlg is $1$-competitive.
-	If at least $\alpha$ external requests were issued, then $\OPT$ either paid $\alpha$ for serving them remotely or paid $\alpha$ for a~migration.
-	Charging this cost to \OPT serves the purpose of charging $\alpha$ at the end of a finished phase, and thus we can repeat the analysis of a finished phase.
-	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into the last two phases, and we repeat the analysis of a finished phase.
-	This way, the competitive ratio increases at most twofold in comparison to a~finished phase, and the competitive ratio is $\TAlg(\sigma)/\OPT(\sigma) \leq 60\ell$.
-\end{proof}
-
-\medskip
-\noindent
-\textbf{Distributed implementation.}
-While we have described the algorithm globally so far, we note that it allows for efficient distributed implementations. 
-The algorithm performs two types of operations that require communication with other clusters: a~component merge, and a~broadcast of the end of the phase.
-We say that a~cluster containing $3$ isolated nodes is \emph{fresh}.
-A merge of two components may require finding a~fresh cluster (for details see the proof of Lemma~\ref{lem:1req}).
-In the following, we show how to efficiently find a~fresh cluster in a~distributed manner.
-To this end, we organize the clusters into an arbitrary rooted balanced binary tree, and we broadcast the root to each cluster.
-Each cluster maintains the counter of fresh clusters in its subtree.
-To find a~fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
-Upon encountering a~fresh cluster, we end the traversal and decrease the counters on the followed path by $1$.
-Summarizing, ending the phase requires a~single broadcast, and merging components has $O(\log \ell)$ communication complexity.
-
-\subsection{Improved Algorithm for Online Rematching} \label{sec:k2}
-In this section,
-we present \RM,
- an algorithm for clusters of capacity $k=2$.
-We interpret a~pair of nodes collocated in one cluster as a~``matched'' pair.
-Hence,
-the problem is an online variant of the maximal matching problem where
-%a match is retractable.
-%That is, 
-a matched pair can separate in order to ``rematch'' with two other nodes.
-Rematching is necessary for  maximizing intra-cluster communications,
-which is equivalent to minimizing inter-cluster communications.
-This is known as the  \emph{online rematching} 
-problem and a~non-strict 7-competitive algorithm is already given by~\cite{repartition-disc},
-in which the ratio comes with an additive factor $O(\alpha\ell^2)$.
-We do not only improve upon their competitive ratio,
-but also show that our ratio holds \emph{strictly}
-(i.e., with no additive factor).
-Our algorithm is slightly simpler than the one in~\cite{repartition-disc}, 
-while our analysis is significantly simpler and more concise,
-thanks to the charging scheme we devise here.
-
-\begin{figure}[H]
-	\centering
-\begin{minipage}[b]{.55\textwidth}
-	\hspace{15mm}
-	\begin{tikzpicture}[x=0.8pt,y=0.75pt,yscale=-1.25,xscale=1.25]
-	\useasboundingbox  (-3mm,-5mm) rectangle (0mm,10mm);
-	%cluster 1
-	\def\x{30}
-	\def\w{15}
-	%Shape: Rectangle
-	\draw   (0,0) -- (15,0) -- (15,40) -- (0,40) -- cycle ;
-	%Straight Lines
-	\draw   (0,20) -- (15,20) ;
-	% Text Node
-	\draw (8,10) node (b1)  {$b_{1}$} ;
-	% Text Node
-	\draw (8,30) node (a4)   {$a_{4}$};
-
-	%cluster 2
-	\draw   (\x,0) -- (\w+\x,0) -- (\w+\x,40) -- (\x,40) -- cycle ;
-	\draw   (\x,20) -- (\w+\x,20) ;
-	% Text Node
-	\draw (\x+8,10) node (b2)   {$b_{2}$};
-	% Text Node
-	\draw (\x+8,30) node (a1)   {$a_{1}$};
-
-	\draw[thick,dotted, shorten <= -6pt, shorten >= -6pt] (a4)  -- (a1);
-	\draw [->,blue, shorten <= -5pt, shorten >= -2pt] (a4) -- (b2);
-
-	%cluster 3
-	\draw   (2*\x,0) -- (\w+2*\x,0) -- (\w+2*\x,40) -- (2*\x,40) -- cycle ;
-	\draw    (2*\x,20) -- (\w+2*\x,20) ;
-	% Text Node
-	\draw (2*\x+8,10) node (b3)   {$b_{3}$};
-	% Text Node
-	\draw (2*\x+8,30) node (a2)   {$a_{2}$};
-	\draw[thick,dotted, shorten <= -4pt, shorten >= -3pt] (b2)  -- (b3);
-	\draw [->,blue, shorten <= -5pt, shorten >= -2pt] (b2) -- (a2);
-
-	%cluster 4
-	\draw   (3*\x,0) -- (\w+3*\x,0) -- (\w+3*\x,40) -- (3*\x,40) -- cycle ;
-	\draw    (3*\x,20) -- (\w+3*\x,20) ;
-	% Text Node
-	\draw (3*\x+8,10) node (b4)   {$b_{4}$};
-	% Text Node
-	\draw (3*\x+8,30) node (a3)   {$a_{3}$};
-	\draw[thick,dotted, shorten <= -4pt, shorten >= -3pt] (a2)  -- (a3);
-	\draw [->,blue, shorten <= -5pt, shorten >= -2pt] (a2) -- (b4);
-
-	\draw[thick, dotted,  shorten <= -3pt, shorten >= -5pt]  (b4) to[out=-100,in=-80] (b1);
-	% arrow
-	\draw[->, blue,shorten <= -3pt, shorten >= -4pt]  (b4) 
-	to [out=-108,in=-130, distance=17mm]  (a4);
-	\end{tikzpicture}
-\end{minipage}\hfill
-%
-\begin{minipage}[b]{.45\textwidth}
-	\begin{tikzpicture}[x=0.8pt,y=0.75pt,yscale=-1.25,xscale=1.25]
-	\useasboundingbox  (0mm,-5mm) rectangle (0mm,10mm);
-%cluster 1
-\def\x{30}
-\def\w{15}
-%Shape: Rectangle
-\draw   (0,0) -- (15,0) -- (15,40) -- (0,40) -- cycle ;
-%Straight Lines
-\draw   (0,20) -- (15,20) ;
-% Text Node
-\draw (8,10) node (b1)  {$b_{1}$} ;
-% Text Node
-\draw (8,30) node (b4)   {$b_{4}$};
-
-%cluster 2
-\draw   (\x,0) -- (\w+\x,0) -- (\w+\x,40) -- (\x,40) -- cycle ;
-\draw   (\x,20) -- (\w+\x,20) ;
-% Text Node
-\draw (\x+8,10) node (a4)   {$a_{4}$};
-% Text Node
-\draw (\x+8,30) node (a1)   {$a_{1}$};
-
-%cluster 3
-\draw   (2*\x,0) -- (\w+2*\x,0) -- (\w+2*\x,40) -- (2*\x,40) -- cycle ;
-\draw    (2*\x,20) -- (\w+2*\x,20) ;
-% Text Node
-\draw (2*\x+8,10) node (b3)   {$b_{3}$};
-% Text Node
-\draw (2*\x+8,30) node (b2)   {$b_{2}$};
-
-%cluster 4
-\draw   (3*\x,0) -- (\w+3*\x,0) -- (\w+3*\x,40) -- (3*\x,40) -- cycle ;
-\draw    (3*\x,20) -- (\w+3*\x,20) ;
-% Text Node
-\draw (3*\x+8,10) node (a2)   {$a_{2}$};
-% Text Node
-\draw (3*\x+8,30) node (a3)   {$a_{3}$};
-\end{tikzpicture}
-\end{minipage}
-	\caption{
-			Dashed lines represent requests.
-		An arrow from node $x$ to node $y$ indicates that $x$  replaces $y$.
-		\OPT collocates 4 pairs from the left partition, with  4 migrations,
-		 resulting in the right partition.
-	}
-	\label{fig:cyclic}
-\end{figure}
-
-%\medskip
-\noindent
-\textbf{Algorithm ReMatch.} 
-The algorithm ReMatch (\RM) maintains a~counter $C_{\set{x,y}}$ for each pair of nodes $\set{x,y}$ and increments it on every remote request between $x$ and $y$.
-Once $C_{\set{x,y}} = \lambda$,
-it resets the counter $C_{\set{x,y}}:=0$ and collocates the pair by swapping $x$ with the node collocated with $y$.
-
-\begin{theorem} \label{thm:k=2}
-	For $\lambda=\alpha$, the algorithm \RM  is strictly 6-competitive.
-\end{theorem}
-
-\noindent
-\textbf{The charging scheme.}
-We charge both \OPT and \RM whenever \RM collocates a~pair.%(i.e., on a~collocation event).
-\RM collocates a~pair always with a~swap %(i.e., two simultaneous migrations),
-(that costs $2\alpha$),
-while \OPT may save some costs by collocating multiple pairs at once.
-% before they inflict too much communication cost.
- Thus it pays the price of only one migration per pair  (see Figure \ref{fig:cyclic}).
-For this reason, whenever \OPT collocates a~pair,
-we charge it only the cost $\alpha$ of moving a~single node to the other cluster (in contrast to the cost $2\alpha$ incurred by \RM).
-
-Consider two  pairs that share the same node, 
-i.e.~\emph{intersecting pairs},
-and the set of requests that causes the collocation of both pairs,
-at some  times during  $[1,m]$.
-%i.e.~the entire duration of the sequence.
-Observe that \OPT must pay a~non-zero cost
-for these requests over the entire $\sigma$,
-since it cannot have both pairs initially collocated.
-However,
-we can charge this cost to \OPT only the first time \RM collocates a~pair,
-and not at any consequent time when \RM collocates it a~second time.
-Otherwise,
- \OPT is possibly charged for the same cost repeatedly.
-For this reason,
-we charge \OPT a~cost inflicted by a~pair,
-if and only if  \OPT incurs that cost after the latest separation of the pair by \RM.
-%%
-
-\begin{proof}[Proof of Theorem \ref{thm:k=2}]
-	Fix an input sequence of requests
-$\sigma:=\{\sigma_1,\dots, \sigma_m\}$.
-	Assume that \RM collocates a pair $\set{u,v}$ at time $t$.
-	The value of $C_{\set{u,v}}$ at $t$,
-	denoted $C^t_{\set{u,v}}$, 
-	reaches $\lambda$ immediately before \RM resets the counter.
-	For any interval $[t_1,t_2]$, by $\sigma_{\set{x,y}}[t_1,t_2]$ we denote the set of all requests to a~pair $\set{x,y}$ that arrive
-	during $[t_1,t_2]$.
-	%and its overall cost to \OPT by
-	%$\OPT (\sigma_{\{u,v\}}[t_1,t_2])$.
-	We may use $\sigma_{\set{x,y}}$ whenever
-	 the interval $[t_1,t_2]$ is clear from the context.
-	%
-	%Else, this is not the first time and
-	%Alg collocates $\{u,v\}$ also some time prior to $t$.
-
-	If $t$ is not the first time that \RM collocates $\{u,v\}$ then
-	let $0 < t' < t$ be the latest time when \RM separates $\set{u,v}$
-	in order to collocate some intersecting pair
-	$\{x,y\} \neq \{u,v\}, \{x,y\} \cap \{u,v\} \neq \emptyset$, 
-	e.g.,
-	$\{x,y\}=\{u,w\}$.
-	Else,
-	$t$ is the first time that \RM collocates $\{u,v\}$ and let $t' := 0$.
-	Similarly,
-	if $t' > 0$ is not the first time that \RM  collocates $\{u,w\}$ 
-	then let $0 < t'' < t'$ be the latest time before $t'$ when \RM separates $\set{u,w}$.
-	Else,
-	$t'$ is the first time that \RM collocates $\{u,w\}$ and we let $t''=0$.
-	%
-
-	First,
-	we bound  costs incurred by \RM for requests that
-	lead to the collocation of $\{u,v\}$ at time $t \in T$, where
-	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$
-	is the set of times when \RM performs a~collocation.
-	By definitions of $t$ and $t'$,
-	the overall cost of requests in $\sigma_{\set{u,v}}$ incurred by \RM,
-	i.e., the total cost of remote serving 	and the moving cost is
-	$\lambda + 2\alpha$.	
-	Next,
-	we bound costs incurred by \RM
-	for requests that do not lead to collocations until the  end of the sequence at $t=m$.
-	Assume $\{u,v\}$ is not collocated at $t=m$
-	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
-	which means \RM pays $C^{m}_{ \{u,v\} }$
-	for  requests in $\sigma_{\set{u,v}}(t',m]$.
-    Then the overall cost of \RM is
-$	\RM (\sigma)
-=
-\sum_{ t \in T}(\lambda + 2\alpha) +
-\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
-$.
-
-	Next,
-	we bound  costs incurred by \OPT for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
-	If $t$ is the first time that \RM collocates $\{u,v\}$,
-	 then  \OPT pays
-	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
-	or $\alpha$ for collocating the pair and
-	serving (some of) the requests with  cost zero.
-	Therefore in this case,
-	$\OPT(\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
-	%
-	Otherwise,
-	it is not the first collocation and
-	consider times $t'$ and $t''$ as defined previously,
-	 and let 
-	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
-	We define $R_{t'}$ for the collocation at $t'$  analogously.
-	Then,
-	$\OPT(R_t) =\OPT (\sigma_{\{u,w\}}) 
-	+ \OPT(\sigma_{\{u,v\}}) $.
-	If \OPT has both pairs separated during their respective intervals,
-	then obviously it pays $2\lambda$ during those intervals.
-	Note that \OPT cannot have both pairs collocated at the same time.
-%	However, it can have one of the pairs already collocated prior to its respective interval.
-	Let us assume \OPT has one of the pairs,
-	e.g.~$\{u,v\}$,
-	 collocated already prior its respective interval, $(t',t]$,
-	 and keeps it so during the interval.
-	 Then it pays zero while serving $\sigma_{\{u,v\}}$.
-	Hence,
-	it must pay $\alpha$ for collocating the other pair, in this case $\{u,w\}$,
-	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$. 
-	Therefore in any case,
-	$\OPT(R_t) \geq  \min{\{ \lambda,\alpha \}} = \alpha$.
-
-	It remains to bound the cost  incurred by \OPT due to requests to $\{u,v\}$ that do not lead to its collocation until the end of the sequence at $t=m$.
-	We bound the cost analogously to the case where \RM collocates $\{u,v\}$.
-	If $\{u,v\}$ is not collocated in the initial matching
-	and \RM never collocates it,
-	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[1,m] |$.
-	\OPT pays
-	$\OPT(\sigma_{\{u,v\}}[1,m]) 
-	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
-	for collocating this pair or (and) paying for (resp.~some of) requests in $\sigma_{\{u,v\}}[1,m]$.
-	Else,
-	either $\{u,v\}$ is collocated in the initial matching
-	or \RM collocates it at some point.
-	Then there exists an intersecting pair $\set{u,w}$
-	that is collocated by \RM at $t' < m$,
-	separating $\{u,v\}$.
-	We define times $t'' < t' < m$ analogously to the former case.
-	Let $R^*_{\{u,v\}} := \sigma_{\{u,w\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
-	Then,
-	\OPT must pay for collocating at least one pair or (and) serving requests 
-	to the other pair remotely.
-	Thus,
-	$\OPT (R^*_{\{u,v\}}) 
-	\geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.
-
-	Next, we sum up all costs incurred by \OPT.
-	By definitions of $R_t$ and $R^*_{\{u,v\}}$, we have either
-	$R_{t'} \cap R_t = \sigma_{\{u,w\}}$ or
-	$R_{t'} \cap R^*_{\{u,v\}} = \sigma_{\{u,w\}}$.
-%	(\mahmoud{See figure \ref{fig:doubleCount}}). 
-	This means,
-	$\OPT ( \sigma_{\{u,w\}})$
-	is counted at most twice in each of  the expressions
-	$\OPT (R_{t'}) + \OPT (R_t)$
-	and  
-	$\OPT (R_{t'}) + \OPT(R^*_{\{u,v\}})$.
-	Hence,
-	for all collocations performed by \RM,
-	and for final requests at $t=m$,
-	\OPT pays at least 
-	$\frac{1}{2}(
-	\sum_{ t \in T } \OPT (R_t) +
-	\sum_{\{u,v\}}\OPT (R^*_{\{u,v\}})
-	) $.
-	Then,
-	the total cost to \OPT is
-	\begin{align*} 	%\label{eq:costOPT}
-		\OPT (\sigma)
-		&=
-		\frac{1}{2}
-		\Big(
-		\sum_{ t \in T} \OPT (R_t) 
-		+ \sum_{\{u,v\}}\OPT (R^*_{\{u,v\}})
-		\Big)	
-		\geq
-%		\Big(
-%		\sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  +
-%		\sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \} } 
-%		\Big)		
-%		=
-		\frac{1}{2}		
-		\Big(
-		\sum_{ t \in T} \alpha  
-		+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}
-		\Big),
-	\end{align*}
-and
-$
-	\RM(\sigma)	/
-	\OPT (\sigma)
-	\leq
-	2\Big(
-	\sum_{ t \in T} 3\alpha +
-	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
-	\Big)	 \big /
-	\Big(
-	\sum_{ t \in T} \alpha  
-	+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}  
-	\Big) 	\leq 6.
-$
-\end{proof}
-
-\noindent
-%\textbf{Approximation for offline rematching.}
-%We briefly sketch an approximation for the offline problem,
-%that makes use of \RM as a~subroutine.
-%In our approximation algorithm,
-%we simulate \RM on the given sequence.
-%Whenever,
-%\RM collocoates a~pair
-
-\section{Discussion and Future Work}
-
-This paper revisited the dynamic graph partitioning problem and presented several tight bounds for the important model where capacities cannot be exceeded, both for a~general partitioning model and for a~special learning model.
-
-While our bounds are tight, there are several interesting avenues for future research.
-In particular, we have so far focused on deterministic algorithms, and it would be interesting to study the power of randomization in this context.
-On the practical side, it would also be interesting to study our algorithms empirically, under realistic workloads.
-
-Our algorithms allow for efficient distributed implementations. 
-The algorithm PPL from Section~\ref{sec:ppl} can be distributed
-similarly to the approach in~\cite{sigmetrics19_partitioning}.
-The algorithm for $k=2$ from Section~\ref{sec:k2} performs only local communication for each request: counters are kept on the clusters and updated locally, and each migration is local within two clusters that reached the collocation threshold $\lambda$.
-Furthermore, we proposed an efficient distributed implementation of the algorithm for $k=3$ in Section~\ref{sec:k3}.
-
-\bibliographystyle{plainurl}
-\bibliography{references}
-
-\appendix
-
-\section{Pseudocode of Perfect Partition Learner}
-
-\begin{algorithm}
-	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-	\renewcommand{\algorithmicensure}{\textbf{Output:}}
-	\begin{algorithmic}
-		%        \Require 
-		%        $k, \ell$,
-		%        initial partition $P_I$,
-		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
-		%        \Ensure a~final partition $P_F$ 
-		\STATE {For each node $v$ create a~singleton component $C_v = \{ v \}$ and add it to $\mathcal{C}$.}
-		%\STATE{$P := P_I$}
-		\label{line:initcomponents}
-		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
-		\STATE{Let $C_1 \ni u$ and $C_2 \ni v$ be the  components containing $u$ and $v$, respectively.}
-		\IF{$C_1 \neq C_2$}
-		\STATE {Merge $C_1$ and $C_2$ into one component $C'$ and
-			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$.} \label{line:mergecomponents}
-		\IF{$C_1$ and $C_2$ are not collocated}
-		\STATE {Move to a~partition closest to $P_I$ and respecting all components in $\mathcal{C}$.} 
-		%		\COMMENT{move to a~partition closest to $P_I$}
-		\label{line:rebalance} 
-		\ENDIF
-		\ENDIF
-		\ENDFOR
-	\end{algorithmic}
-	\caption{Perfect Partition Learner (\PPL)}
-	\label{alg:ppl}
-\end{algorithm}
-
-%\section{Additional Figures}
-\begin{comment}
-\begin{figure}[H]	\label{fig:timeline}
-	\centering
 		\begin{tikzpicture}
@@
 	\end{tikzpicture}
-\caption{
-Illustration of the timeline used in the proof of Theorem \ref{thm:k=2}.
-Requests in $R_t$  are only to  $\{u,w\}$ and $\{u,v\}$,
-which  arrive  during the interval $(t'',t]$.
-Similarly,
-request in $R_{t'}$  are to $\{u,w\}$ and some other pair, irrelevant to the analysis.
-Hence, requests to $\{u,w\}$ are included in at most two such sets,
-which are $R_{t'}$ and $R_t$.
-This is because their intervals overlap on $(t'',t']$, shown with red thick line.
-}
-\end{figure}
-\end{comment}
 

### 2020-05-18 podc_NEW/podc_ba.tex
NEW brief announcement

--- 2020-05-15 disc_paper/disc20clustering.tex
@@
- \documentclass[a4paper,anonymous,USenglish]{lipics-v2019}
+\documentclass[manuscript,screen=true]{acmart}
 
@@
 
- \title{Optimal Algorithms for Dynamic Balanced Graph Partitioning}
+\title{Brief Announcement: Tight Bounds for Online Balanced Repartitioning}
 
- \author{Maciej Pacut}{maciej.pacut@univie.ac.at}{Faculty of Computer Science, University of Vienna,Austria}{0000-0002-6379-1490}{}
+\author{Maciej Pacut}
+\email{maciej.pacut@univie.ac.at}
+\orcid{0000-0002-6379-1490}
+\affiliation{%
+  \institution{Faculty of Computer Science, University of Vienna}
+  \country{Austria}
+}
 
-% \author{Mahmoud Parham}{mahmoud.parham@univie.ac.at}{Faculty of Computer Science, University of Vienna, Austria}{0000-0002-6211-077X}{}
+\author{Mahmoud Parham} 
+\email{mahmoud.parham@univie.ac.at}
+\orcid{0000-0002-6211-077X}
+\affiliation{%
+  \institution{Faculty of Computer Science, University of Vienna}
+  \country{Austria}
+}
 
-% \author{Stefan Schmid}{stefan\_schmid@univie.ac.at}{Faculty of Computer Science, University of Vienna, Austria}{}{}
+\author{Stefan Schmid} 
+\email{stefan_schmid@univie.ac.at}
+\affiliation{%
+  \institution{Faculty of Computer Science, University of Vienna}
+  \country{Austria}
+}
 
- \authorrunning{M. Pacut, M. Parham, and S. Schmid}
- \Copyright{Maciej Pacut, Mahmoud Parham, and Stefan Schmid}
+\copyrightyear{2020} 
+\acmYear{2020} 
+\setcopyright{acmlicensed}
+\acmConference{PODC '20}{July 29-August 2, 2020}{Toronto, Canada}
+
+\keywords{online algorithms, competitive analysis, graph partitioning, clustering}
+\acmISBN{}\acmPrice{}
+\acmDOI{}
  \keywords{online algorithms, competitive analysis, distributed computing, graph partitioning, clustering, self-adjusting networks}
-
-% \EventEditors{}
-%\EventNoEds{3}
- \EventLongTitle{34th International Symposium on DIStributed Computing (DISC) 2020}
- \EventShortTitle{DISC 2020}
- \EventAcronym{DISC}
- \EventYear{2020}
- \EventDate{October 12--16, 2020}
- \EventLocation{Freiburg, Germany (virtual conference)}
- \EventLogo{}
- \SeriesVolume{}
- \ArticleNo{}
 
@@
 \begin{document}
-\maketitle
 
@@
 \end{abstract}
+\maketitle
 
@@
 Ideally, this algorithm should perform close to an optimal offline algorithm, without requiring any information about future traffic demands.
-
-\pagebreak
 
@@
 All algorithms in this paper have a strict competitive ratio (i.e., without an additive term).
-Table \ref{tab:overview} provides an overview of our contributions compared to prior work.
-
-\begin{table*}[h]
-	\centering
-	\renewcommand{\arraystretch}{1.5}
-	\begin{tabular}{>{\centering\arraybackslash}p{4.25cm}|>{\centering\arraybackslash}p{4.25cm}>{\centering\arraybackslash}p{4.25cm}}
-		\rowcolor{gray!50}
-		\textbf{Variant} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
-		\textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & 6\hspace{0.3cm}(\S \ref{sec:k2}) \\ 
-		\rowcolor{gray!25}
-		\textbf{$k=3$}&  $\Omega(\ell)$ \hspace{0.3cm}(\S \ref{sec:lowerbound})& $O(\ell) $\hspace{0.3cm}(\S \ref{sec:k3})\\
-		$k > 3$ & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k^2 \cdot \ell^2)$\hspace{0.1cm}~\cite{repartition-disc} \\
-		\rowcolor{gray!25}
-		Learning model & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell)$\hspace{0.1cm} (\S \ref{sec:ppl}) \\
-	\end{tabular}
-	\caption{Overview of known results and our contributions.
-The table summarizes the results for the general partitioning model, except for the last row that summarizes the results for the learning model for arbitrary $k$ and $\ell$.
-	}
-	\label{tab:overview}
-	\vspace{-7mm}
-\end{table*}
 
@@
 
-Given a perfect partition,
-every subset of nodes that belong to the same cluster in this partition
-is a~\emph{ground set}. 
-Any competitive algorithm under the learning model
- maintains a~perfect partition of ground sets into clusters.
- On each inter-cluster request, a ground set is revealed.
- An algorithm recovers the (hidden) perfect partition gradually over inter-cluster requests,
-  by merging pairs of ground sets involved in these requests.
+  \begin{theorem}
+	\label{th:lowerbound}
+	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least
+	 $(k-2)(\ell-1)/2 - 2$ for any $k\geq 3$ and $\ell \geq 2$.
+\end{theorem}
 
@@
 Once we construct a ground set, it lasts until the end of the input sequence.
-We say that a ground set is a~\emph{singleton} if it contains exactly one node,
-which is an \emph{isolated} node.
-
 We start by constructing a~ground set of size $k-1$ on an arbitrarily chosen cluster.
@@
 %Figure~\ref{fig:lb} illustrates the constructs used in the lower bound.
-
-\begin{theorem}
-	\label{th:lowerbound}
-	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least
-	 $(k-2)(\ell-1)/2 - 2$ for any $k\geq 3$ and $\ell \geq 2$.
-\end{theorem}
-
-\begin{proof}
-	Fix any online algorithm \ALG.
-	For a ground set $C$ of nodes that are initially collocated in one cluster,
-	let $I(C)$ denote the cluster.
-	We refer to $I(C)$ as the cluster of \emph{origin},
-	when $C$ is clear from the context.
-%	If $C$ contains nodes that originate from different clusters
-%	then $I(C)$ is undefined.
-%	Throughout the construction, we maintain ground sets for each node.
-	Initially, all nodes are isolated, i.e., each node is in a singleton ground set.
-	First, we choose a cluster arbitrarily and
-	 create a~ground set $B$ of $k-1$ nodes in this cluster.
-	%We never dismiss any constructed ground set.
-	Each cluster hosts exactly $k$ nodes, and in any feasible partition,
-	 a~single isolated node must be collocated with $B$.
-	At any time,
-	we refer to the isolated node currently collocated with $B$ as the \emph{pivot} node.
-	Let $x_0$ denote the first pivot node.
-
-	Then, we join the pivot node to a~larger ground set to force its eviction.
-	Precisely,
-	we create a~ground set $\{x_0, y_0\}$, 	where
-%	 $x_0$ is the current pivot node and
-	 $y_0$ is an arbitrary isolated node.
-	Since \ALG does not have $\{x_0,y_0\}$ collocated, 
-	the adversary issues an external request to this pair so that \ALG collocates it.
-	\ALG cannot collocate $\{x_0, y_0\}$ with $B$ (as $B$'s size is $k-1$), hence it collocates them in a~different cluster.
-	In order to preserve a~feasible partition of nodes after collocating $\{x_0, y_0\}$,
-	\ALG must replace $x_0$ with another isolated node that becomes the new pivot.
-%	denoted by $x_1$.
-
-	We proceed in similar steps by joining the current pivot node to a ground set of the same origin residing in a different cluster.
-	Consider the step $i$,
-	when the isolated node $x_i$ is collocated with $B$.
-	 We issue a request between $x_i$ and some node in $C_i$,
-	 where $C_i$ is the largest ground set
-	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$.
-	 Then \ALG must collocate the new ground set
-	  $\{x_i\} \cup C_i $ in one cluster.
-%	 only if  a feasible partition exists
-%	 that collocates all nodes in $\{x_i\} \cup C_i $.
-	Any feasible partition replaces $x_i$ with some isolated node $x_{i+1}$,
-	as the new ground set $\{x_i\} \cup C_i $  may not be ever split.
-%	Otherwise,
-%	no such feasible partition exists and
-%	we terminate the process.
-	We terminate the process once the number of remaining isolated nodes 
-	is less than $\ell+3$.
-	At each step $i$,
-	the number of isolated nodes decreases either by one
-	 or by two if $C_i$ is a singleton.
-	Therefore, once the process terminates,
-	in any case at least $\ell+1$ isolated nodes are left.
-
-	  Next, we argue that a feasible partition exists when the process terminates.
-	 This implies that a feasible partition exists after any earlier step as well.
-%	 Assume for contradiction that the partition is not feasible.	 
-	 Since there are at least $\ell+1$ isolated nodes left,
-	 there must be two isolated nodes $x^*$ and $y^*$,
-	  with the same cluster of origin,
-	 i.e., $I(\{x^*\}) = I(\{y^*\})$.
-	 Consider the partition $P^*$
-	  obtained from the initial partition after swapping $x_0$ and $y_0$
-	 with $x^*$ and $y^*$ (respectively).
-	 In this partition,
-	 the ground set $\{x_0,y_0\}$ is collocated in the cluster $I(\{x^*,y^*\})$.
-	 Note that after the first request $\{x_0,y_0\}$,
-	   we issue requests only between nodes that have the same cluster of origin and
-	   all these nodes are collocated in $P^*$.
-	 Therefore all ground sets constructed so far are collocated in  $P^*$,
-	 and it is a feasible partition.
-%	 Hence, there are at most $\ell$ isolated nodes after the termination.
-
-%	The number of isolated nodes decrease after each step.
-%	Consider the last step, $j$, when  $\ell+1$ isolated nodes exist before  $x_{j}$ joins to another ground set.
-%	That is, there are less than $\ell+1$ isolated nodes after the step $j$.
-%	Then, at the end of step $j-1$,
-%	there exist two isolated nodes $x^*$ and $y^*$
-%	with the same cluster of origin, i.e., $I(\{x^*\})=I(\{y^*\})$.
-%
-%   Next,
-%    we bound costs incurred by \OPT and \ALG over the  sequence of requests
-%    $\sigma := \langle \{x_0,y_0\}, \dots, \{x_{j-1},y_{j-1}\}  \rangle$.
-
-	Consider  nodes $x^*$ and $y^*$ and the partition $P^*$ obtained previously.
-	\OPT moves to $P^*$ by performing only two node swaps.
-	Precisely, \OPT collocates $\set{x_0,y_0}$ 
-	by swapping them with $x^*$ and $y^*$.
-	No ground set is split in $P^*$ and
-	 \OPT pays only for the two swaps.
-
-	\ALG performs at least one swap at each step $i$,
-	and some ground set grows.
-%	In other words,
-%	some ground set grows by one node at each of these steps.
-	Consider any ground set $C^* \neq B$ after the termination.
-	This ground set has grown exactly $|C^*|-1$ times until the termination. 
-	Let $\mathcal{S}$ be the set of all ground sets after the process terminates.
-	Thus, $\mathcal{S}$ includes ground sets $B$,
-	$\{x_0,y_0\}$,
-	and (up to) $\ell+2$ singleton ground sets.
-	Among the remaining ground sets in $\mathcal{S}$,
-	no two ground sets have the same origin.
-	Otherwise,
-   the smaller ground set is either a singleton,
-   which contradicts the bound $\ell+2$ on the number of singletons,
-	or we have joined nodes to it at some step,
-	contradicting our choice of the largest~$C_i$ at step $i$.
-	Hence,  there are at most $\ell-1$ such ground sets,
-	one per possible cluster of origin,
-	 excluding the cluster containing $B$. 
-	Therefore,  $|\mathcal{S}| \leq 1 + 1 + (\ell+2)  + (\ell-1) = 2\ell+3$.
-	Note that among all non-singleton ground sets in $\mathcal{S}$,
-	 only $B$ does not grow during the process.
-	 Thus,
-    the total number of times that 
-	a ground set in $\mathcal{S}$  has grown is
-	$	\sum_{C^* \in \mathcal{S} } (|C^*|-1) - (k-1) $
-	\begin{align*}
-	=\sum_{C^* \in \mathcal{S} } |C^*| - 
-		\sum_{C^* \in \mathcal{S} } 1 - (k-1) 
-	\geq  k  \ell - (2\ell+3) - (k-1) =  (k-2)(\ell-1) - 4,
-	\end{align*}
-	which bounds the number of swaps performed by \ALG.
-	The competitive ratio is then
-	$\ALG / \OPT \geq ((k-2)(\ell-1) - 4) / 2$.
-	%%
-	\begin{comment}
-	\ALG performs at least one swap  every time   a~ground set is constructed.
-	Each expansion reduces the number of isolated nodes by $1$ or $2$.
-	The decrease of two isolated nodes occurs when the largest component on  $I(x_i)$ is still singleton; this occurs at most $\ell$ times.
-	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
-	Hence, the total number of swaps of \ALG is at least $k\cdot \ell - k - 2 \ell$.
-	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - 2\ell) / 2$.
-	\end{comment}
-	%%
-\end{proof}
-
-\noindent
-\textbf{Lower bound for the general problem.}
-For a~lower bound $\Omega(k\cdot \ell)$ for the general partitioning problem, the adversary continues to issue requests to split nodes of a~ground set until the algorithm collocates them.
-Note that the ground sets constructed in our lower bound can be perfectly partitioned into the clusters.
-Hence, the optimal algorithm moves to a~perfect partition at the beginning (where requests incur the cost~$0$), and its cost is bounded.
-This means that the algorithm must eventually collocate all nodes of a~ground set to be competitive.
-We reveal the next ground set only after the collocation, hence we can repeat the analysis of the algorithm for the learning problem.
-Finally, we note that the construction is oblivious to the choice of the reconfiguration cost $\alpha$: we compare the 	number of node exchanges of \ALG and \OPT.
-
-\medskip
-
-\noindent
-\textbf{Resource augmentation.}
-The majority of work on the online balanced partitioning problem so far~\cite{repartition-disc,sigmetrics19_partitioning} focuses on the scenario with resource augmentation, where the clusters of an online algorithm are larger than the clusters of the offline optimal algorithm that we compare the performance to.
-We can adjust our construction to show a~lower bound of $\Omega(\ell)$ for a~setting with resource augmentation.
-
-Consider a~partitioning problem with resource augmentation $1+1/3-\epsilon$.
-Fix $k$ divisible by $3$, and construct $3$ ground sets of size $k/3$ in each cluster.
-Note that no more than $3$ such ground sets fit in one cluster.
-Then, apply the construction from the lower bound for $k=3$, using these ground sets in the way we used individual nodes.
-The cost of any algorithm (including \OPT) scales up by $k/3$, and the lower bound $\Omega(\ell)$ holds.
-
-Finally, we note the possibility of improvement. The algorithm CREP~\cite{repartition-disc} requires $(2+\epsilon)$-augmentation to guarantee the competitive ratio independent of $\ell$.
-In contrast, our construction shows that the linear term $\ell$ is inevitable if the augmentation is smaller than~$1+1/3$.
 
@@
 \end{lemma}
-\begin{proof}
-	Let $P_i$ denote the partition of \PPL immediately after serving $\sigma_{i}$.
-	Consider the repartitioning that transforms 
-	$P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
-	Let $M \subset V$ denote the set of nodes that migrate during this process.
-	Let $M^-$ and $M^+$ denote the subset of nodes that, respectively,
-	enter or leave their initial cluster during the repartitioning.    
-	Then,
-	$M = M^+ \cup M^-$.
-	Since at least $|M^-|$ nodes are not in their initial cluster before the repartitioning (i.e., in $P_{t-1}$),
-	the distance before the repartitioning is $\Delta(P_{t-1}) \geq | M^-|$.
-	Analogously,
-	the distance afterward is $\Delta(P_{t}) \geq | M^+|$.
-	Thus,
-	$|M| \leq \Delta(P_{t-1}) + \Delta(P_{t})$.
-	By Property \ref{prop:dist<OPT},
-	$\Delta(P_{t-1})  \leq \Delta^*$ and
-	$\Delta(P_{t}) \leq \Delta^*$.
-	Since $\Delta^* \leq \OPT$,
-	we obtain	 $|M| \leq 2\cdot\OPT$.
-\end{proof}
 
@@
 \end{theorem}
-\begin{proof}
-	On each inter-cluster request,
-	the algorithm enumerates all component respecting $\ell$-way partitions of components
-	that are in the same (closest) distance to $P_I$.
-	That is, 
-	once it reaches a~partition $P$ at distance $\Delta^* = \Delta(P)$,
-	it does not move to a~partition
-	$P', \Delta(P') > \Delta^*$,
-	before it enumerates all partitions at distance $\Delta^*$.
-	Therefore,
-	\PPL eventually reaches  the partition $P_F$ at distance $\Delta^*=\OPT$.
-	With each distinct request, 
-	the size of some component increases by one.
-	For any cluster $F_i \in P_F$,
-	we have $\sum_{C \in F_i} |C| = k$.
-	A component $C \in F_i$,
-	 initially begins as an isolated node and it grows by gaining $|C|-1$ more nodes.
-	Hence, the total number of times a~component in $F_i$ grows is 
-	$\sum_{C \in F_i} (|C|-1) \leq k-1$.
-	Therefore, there are at most  $(k-1)\cdot\ell $ distinct requests
-	for which \PPL performs a~repartitioning and
-	 \PPL performs at most $(k-1)\cdot\ell $ repartitions.
-	By Lemma \ref{lemma:rebalancecost},
-	each repartitioning costs at most $2\cdot\OPT$.
-	The total cost is thus at most $2\cdot\OPT\cdot (k-1) \cdot\ell$, which implies the competitive ratio.
-\end{proof}
-
-% In Section~\ref{sec:lowerbound} we constructed a~$\Omega(k \cdot \ell)$ for \OBRP{}.
-% Note that the lower bound holds also in the perfect partition model, as the constructed input sequence allows \OPT to move to a~perfect partition.
-% The corollary is that PPL is optimal.
 
@@
 In Section~\ref{sec:k3}, we show an $O(k \cdot \ell)$-competitive algorithm for $k=3$ using the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental}.
-Prior to this section, we showed a~lower bound of $\Omega(k \cdot \ell)$  that holds for the general model (cf.
-Section~\ref{sec:lowerbound}), hence the result from this section is asymptotically optimal.
-Furthermore, in Section~\ref{sec:k2}, we show a~strictly $6$-competitive algorithm for $k=2$.
 
@@
 
+\begin{lemma}
+	\label{lem:1req}
+	In a~single repartition of nodes (after a~merge of components), \TAlg exchanges at most two pairs of nodes.
+\end{lemma}
+
 \begin{theorem}
@@
 
-Before bounding the competitive ratio of \TAlg, we upper-bound the cost of a~single repartition of \TAlg.
-In our analysis, we distinguish among three types of clusters: $C_1, C_2$ and $C_3$. In a~cluster of type $C_i$, the size of the largest component contained in this cluster is~$i$.
-
-\begin{lemma}
-	\label{lem:1req}
-	In a~single repartition of nodes (after a~merge of components), \TAlg exchanges at most two pairs of nodes.
-\end{lemma}
-
-\begin{proof}
-	If no component respecting partition exists after the merge of components, then \TAlg resets all components, ends the phase, and performs no repartition.
-	It suffices to show that the merged component has size at least $4$ to conclude that \TAlg incurs no cost.
-
-	%We distinguish between three types of clusters: $C_1, C_2, C_3$,
-	%which we define as follows.
-	%A cluster of type $C_1$ the cluster contains $3$ singleton components (this is also the initial partition of any cluster).
-	%A cluster of type $C_2$  contains one component of size $2$ and one component of size $1$.
-	%Finally, a~cluster of type $C_3$  contains one component of size $3$.
-
-	Consider a~request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
-	The request triggered the repartition, hence it was external and $U\neq V$.
-	We consider cases based on the types of clusters $U$ and $V$.
-
-	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the repartition is local within $U$ and $V$, for the~cost of at most $2$ swaps.
-	If either $U$ or $V$ is of type $C_3$, a~component of size $3$ participates in a~merge, and we have a~component of size at least $4$, and \TAlg ends the phase with no repartition.
-
-	\pagebreak
-
-	It remains to consider the case where both $U$ and $V$ are of type $C_2$.
-	If $(u,v)$ both belong to components of size $2$, then the merged component has size $4$, and \TAlg incurs no cost. 
-	Otherwise, if one of $u,v$ belongs to a~component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
-	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a~cluster different from $U$ and $V$.
-	Note that if $C_1$-type cluster does not exist, then no component respecting partitioning exists.
-	Otherwise, \TAlg performs one swap --- it exchanges the nodes $u$ and $v$ with any two nodes~of~any cluster of type $C_1$.
-
-	In each case, we showed that a~component respecting partition is reachable in at most two swaps.
-\end{proof}
-
-%\begin{figure}[H]
-%	\centering
-%	\includegraphics[width=0.3\textwidth]{figs/substitute}
-%	\caption{an unsaturated edge that is inside a~component. If there is a~space next to it, maybe cluster types should fit next to it? Not sure.}
-%\end{figure}
-
-\begin{proof}[Proof of Theorem \ref{thm:k=3}]
-	Fix a~completed phase, and consider the state of \TAlg's counters at the end of it (before the reset).
-	We consider the incomplete phase later in this proof.
-
-	\TAlg is component respecting, hence it never increases any counter above $\alpha$.
-	We say that the pair $(u, v)$ is \emph{saturated} if the counter's value is $\alpha$, and \emph{unsaturated} otherwise (saturation of a~pair leads to a~merge action).
-	By $\sigma$ we denote the input sequence that arrived during the phase.
-	In our analysis, we focus on the requests that were external to \TAlg at the moment of their arrival; these are the only requests that incur a~cost for \TAlg.
-	We denote these external requests by $\sigma_{cost}$.
-	We partition the sequence $\sigma_{cost}$ into subsequences $\sigma_I$ and $\sigma_E$.
-	The sequence $\sigma_I$ (inter-component requests) denotes the requests from $\sigma_{cost}$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
-	The sequence $\sigma_E$ (extra-component requests) denotes the requests from $\sigma_{cost}$ that do not appear in $\sigma_I$.
-
-	%Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
-	%Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
-	Let $\TAlg(M)$ denote the cost of migrations performed by \TAlg in this phase.
-	During the phase, \TAlg performs at most $2 \ell$ component merge operations ---
-	exceeding this number would mean that a~component of size $4$ exists, and the phase should have ended already.
-	We bound the cost of each repartition after a~merge by Lemma~\ref{lem:1req}, obtaining $\TAlg(M) \leq 8\alpha\cdot\ell$.
-	%Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.
-
-	We bound $\TAlg(\sigma_I)$ by summing the intra-component counters of each cluster at the end of the phase.
-	The sum of intra-component counters in a cluster of type $C_3$ is at most $3 \alpha - 1$: two pairs of nodes from the component are saturated and its counter is $\alpha$ each, and the counter of the third, unsaturated pair is at most $\alpha-1$.
-	The sum of counters inside $C_1$ is~$0$, and inside $C_2$ it is $\alpha$.
-	Summing over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq (3 \alpha-1)\cdot \ell \leq 3\alpha\cdot\ell$.
-
-	%We bound $A^E$ by $k^2 \cdot (\alpha - 1)$, as no more than $k^2$ pairs are unsaturated, and each of them contributes at most $\alpha -1$.
-	%\maciek{not needed most likely}
-
-	Furthermore, \TAlg paid for all requests from $\sigma_E$, and thus $\TAlg(\sigma_E) = |\sigma_E|$.
-	In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 11\alpha\cdot \ell + |\sigma_E|$ during this phase.
-
-	\medskip
-
-	Now we lower-bound the cost of the optimal offline solution.
-	To this end, we fix any optimal offline algorithm $\OPT$.
-	By $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$ we denote the cost of $\OPT$ on requests from sequences $\sigma_I$ and $\sigma_E$, respectively.
-	Note that these costs are defined with respect to components of \TAlg in this phase.
-	By $\OPT(M)$ we denote the cost of migrations performed by $\OPT$ in this phase.
-
-	The cost of \OPT is lower-bounded by the cost of serving $\sigma_I$ and the cost of serving $\sigma_E$.
-	While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts: we separately bound $\OPT$ by $\OPT(\sigma_I) + \OPT(M)$ and $\OPT(\sigma_E) + \OPT(M)$.
-	Combining those bounds and using the relation between the maximum and the average, we obtain the bound
-	\begin{align*}
-		\OPT&\geq \max\{\OPT(\sigma_I) + \OPT(M), \OPT(\sigma_E) + \OPT(M)\}\\
-		&\geq (\OPT(\sigma_I) + \OPT(M)) / 2 + (\OPT(\sigma_E) + \OPT(M)) / 2.
-	\end{align*}
-
-	%Let $O^M$ be the cost of migrations performed by $\OPT$ during the phase.
-	%Let $O^I$ be the cost of serving requests between nodes that were put in one component by \TAlg during this phase.
-	%Let $O^E$ be the cost serving requests between nodes that \TAlg did not put in the same component during that phase.
-
-	\pagebreak
-
-	First, we show $\OPT(M) + \OPT(\sigma_I) \geq \alpha$.
-	Assume that \OPT's partition is fixed throughout the phase (as otherwise \OPT pays $\alpha$ for a migration).
-	The phase ended when the components of \TAlg{} could not be partitioned without splitting them.
-	Hence, for every possible partition of $\OPT$, there exists a~non-collocated saturated pair, and
-	$\OPT$ paid for $\alpha$ requests that saturated the pair.
-
-	Next, we bound $\OPT(\sigma_E) + \OPT(M)$.
-	The sequence $\sigma_E$ accounts only for unsaturated edges, thus there are at most $\alpha-1$ requests to each pair in $\sigma_E$.
-	\OPT may have at most $3\ell$ pairs of nodes collocated in its clusters, and thus avoid paying for $3\ell\cdot(\alpha-1)$ requests from $\sigma_E$.
-	Hence, at least $\chi := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests from $\sigma_E$ are external requests with respect to \OPT's configuration at the beginning of the phase.
-	Faced with these requests, \OPT may serve them remotely or perform migrations to decrease its cost.
-	By swapping a~pair of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
-	This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterward.
-	Hence, by performing a~single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
-	The total cost of \OPT is then at least
-	%
-	\[
-		\OPT(\sigma_E) + \OPT(M) \geq \chi \cdot \frac{2\alpha}{4 (\alpha-1)}\geq \frac{|\sigma_E|} {2} - 2 \alpha \cdot \ell.
-	\]
-
-	Finally, to bound the competitive ratio, we transform the above inequality in the following way: $|\sigma_E| \leq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
-	For succinctness, let $\xi := \OPT(\sigma_E) + \OPT(M)$.
-	Combining the bounds on the cost of \TAlg and \OPT during each finished phase, the competitive ratio is
-%
-	\[
-		\frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + \xi/2} \leq \frac{30\alpha\cdot\ell + 4\cdot \xi}{\alpha + \xi} \leq 30 \ell.
-	\]
-%	
-	\medskip
-
-	It remains to consider the last, unfinished phase.
-	First, consider the case, where the unfinished phase is also the first one.
-	Then, we cannot charge $\OPT$ due to the inability to partition the components.
-	Instead, we use the fact that \TAlg and $\OPT$ started with the same initial partition.
-	If the input finished before the first $\alpha$ external requests, then \TAlg is $1$-competitive.
-	If at least $\alpha$ external requests were issued, then $\OPT$ either paid $\alpha$ for serving them remotely or paid $\alpha$ for a~migration.
-	Charging this cost to \OPT serves the purpose of charging $\alpha$ at the end of a finished phase, and thus we can repeat the analysis of a finished phase.
-	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into the last two phases, and we repeat the analysis of a finished phase.
-	This way, the competitive ratio increases at most twofold in comparison to a~finished phase, and the competitive ratio is $\TAlg(\sigma)/\OPT(\sigma) \leq 60\ell$.
-\end{proof}
-
-\medskip
-\noindent
-\textbf{Distributed implementation.}
-While we have described the algorithm globally so far, we note that it allows for efficient distributed implementations. 
-The algorithm performs two types of operations that require communication with other clusters: a~component merge, and a~broadcast of the end of the phase.
-We say that a~cluster containing $3$ isolated nodes is \emph{fresh}.
-A merge of two components may require finding a~fresh cluster (for details see the proof of Lemma~\ref{lem:1req}).
-In the following, we show how to efficiently find a~fresh cluster in a~distributed manner.
-To this end, we organize the clusters into an arbitrary rooted balanced binary tree, and we broadcast the root to each cluster.
-Each cluster maintains the counter of fresh clusters in its subtree.
-To find a~fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
-Upon encountering a~fresh cluster, we end the traversal and decrease the counters on the followed path by $1$.
-Summarizing, ending the phase requires a~single broadcast, and merging components has $O(\log \ell)$ communication complexity.
-
-\subsection{Improved Algorithm for Online Rematching} \label{sec:k2}
-In this section,
-we present \RM,
- an algorithm for clusters of capacity $k=2$.
-We interpret a~pair of nodes collocated in one cluster as a~``matched'' pair.
-Hence,
-the problem is an online variant of the maximal matching problem where
-%a match is retractable.
-%That is, 
-a matched pair can separate in order to ``rematch'' with two other nodes.
-Rematching is necessary for  maximizing intra-cluster communications,
-which is equivalent to minimizing inter-cluster communications.
-This is known as the  \emph{online rematching} 
-problem and a~non-strict 7-competitive algorithm is already given by~\cite{repartition-disc},
-in which the ratio comes with an additive factor $O(\alpha\ell^2)$.
-We do not only improve upon their competitive ratio,
-but also show that our ratio holds \emph{strictly}
-(i.e., with no additive factor).
-Our algorithm is slightly simpler than the one in~\cite{repartition-disc}, 
-while our analysis is significantly simpler and more concise,
-thanks to the charging scheme we devise here.
-
-%\medskip
-\noindent
-\textbf{Algorithm ReMatch.} 
-The algorithm ReMatch (\RM) maintains a~counter $C_{\set{x,y}}$ for each pair of nodes $\set{x,y}$ and increments it on every remote request between $x$ and $y$.
-Once $C_{\set{x,y}} = \lambda$,
-it resets the counter $C_{\set{x,y}}:=0$ and collocates the pair by swapping $x$ with the node collocated with $y$.
-
-\begin{theorem} \label{thm:k=2}
-	For $\lambda=\alpha$, the algorithm \RM  is strictly 6-competitive.
-\end{theorem}
-
-\noindent
-\textbf{The charging scheme.}
-We charge both \OPT and \RM whenever \RM collocates a~pair.%(i.e., on a~collocation event).
-\RM collocates a~pair always with a~swap %(i.e., two simultaneous migrations),
-(that costs $2\alpha$),
-while \OPT may save some costs by collocating multiple pairs at once.
-% before they inflict too much communication cost.
- Thus it pays the price of only one migration per pair  (see Figure \ref{fig:cyclic}).
-For this reason, whenever \OPT collocates a~pair,
-we charge it only the cost $\alpha$ of moving a~single node to the other cluster (in contrast to the cost $2\alpha$ incurred by \RM).
-
-Consider two  pairs that share the same node, 
-i.e.~\emph{intersecting pairs},
-and the set of requests that causes the collocation of both pairs,
-at some  times during  $[1,m]$.
-%i.e.~the entire duration of the sequence.
-Observe that \OPT must pay a~non-zero cost
-for these requests over the entire $\sigma$,
-since it cannot have both pairs initially collocated.
-However,
-we can charge this cost to \OPT only the first time \RM collocates a~pair,
-and not at any consequent time when \RM collocates it a~second time.
-Otherwise,
- \OPT is possibly charged for the same cost repeatedly.
-For this reason,
-we charge \OPT a~cost inflicted by a~pair,
-if and only if  \OPT incurs that cost after the latest separation of the pair by \RM.
-%%
-
-\begin{figure}[H]
-	\centering
-	\begin{minipage}[b]{.5\textwidth}	
-				\centering	
-		\includestandalone{figs/k=2_a}
-	\end{minipage}%
-	%
-	\begin{minipage}[b]{.5\textwidth}
-		\centering
-				\includestandalone{figs/k=2_b}
-	\end{minipage}
-	\caption{
-		Dashed lines represent requests.
-		An arrow from node $x$ to node $y$ indicates that $x$  replaces $y$.
-		\OPT collocates 4 pairs from the left partition, with  4 migrations,
-		resulting in the right partition.
-	}
-	\label{fig:cyclic}
-\end{figure}
-
-\begin{proof}[Proof of Theorem \ref{thm:k=2}]
-	Fix an input sequence of requests
-$\sigma:=\{\sigma_1,\dots, \sigma_m\}$.
-	Assume that \RM collocates a pair $\set{u,v}$ at time $t$.
-	The value of $C_{\set{u,v}}$ at $t$,
-	denoted $C^t_{\set{u,v}}$, 
-	reaches $\lambda$ immediately before \RM resets the counter.
-	For any interval $[t_1,t_2]$, by $\sigma_{\set{x,y}}[t_1,t_2]$ we denote the set of all requests to a~pair $\set{x,y}$ that arrive
-	during $[t_1,t_2]$.
-	%and its overall cost to \OPT by
-	%$\OPT (\sigma_{\{u,v\}}[t_1,t_2])$.
-	We may use $\sigma_{\set{x,y}}$ whenever
-	 the interval $[t_1,t_2]$ is clear from the context.
-	%
-	%Else, this is not the first time and
-	%Alg collocates $\{u,v\}$ also some time prior to $t$.
-
-	If $t$ is not the first time that \RM collocates $\{u,v\}$ then
-	let $0 < t' < t$ be the latest time when \RM separates $\set{u,v}$
-	in order to collocate some intersecting pair
-	$\{x,y\} \neq \{u,v\}, \{x,y\} \cap \{u,v\} \neq \emptyset$, 
-	e.g.,
-	$\{x,y\}=\{u,w\}$.
-	Else,
-	$t$ is the first time that \RM collocates $\{u,v\}$ and let $t' := 0$.
-	Similarly,
-	if $t' > 0$ is not the first time that \RM  collocates $\{u,w\}$ 
-	then let $0 < t'' < t'$ be the latest time before $t'$ when \RM separates $\set{u,w}$.
-	Else,
-	$t'$ is the first time that \RM collocates $\{u,w\}$ and we let $t''=0$.
-	%
-
-	First,
-	we bound  costs incurred by \RM for requests that
-	lead to the collocation of $\{u,v\}$ at time $t \in T$, where
-	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$
-	is the set of times when \RM performs a~collocation.
-	By definitions of $t$ and $t'$,
-	the overall cost of requests in $\sigma_{\set{u,v}}$ incurred by \RM,
-	i.e., the total cost of remote serving 	and the moving cost is
-	$\lambda + 2\alpha$.	
-	Next,
-	we bound costs incurred by \RM
-	for requests that do not lead to collocations until the  end of the sequence at $t=m$.
-	Assume $\{u,v\}$ is not collocated at $t=m$
-	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
-	which means \RM pays $C^{m}_{ \{u,v\} }$
-	for  requests in $\sigma_{\set{u,v}}(t',m]$.
-    Then the overall cost of \RM is
-$	\RM (\sigma)
-=
-\sum_{ t \in T}(\lambda + 2\alpha) +
-\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
-$.
-
-\begin{figure}[t]
-	\centering
-	\includestandalone{figs/timeline}
-	\caption{
-		Illustration of the timeline used in the proof of Theorem \ref{thm:k=2}.
-		Requests in $R_t$  are only to  $\{u,w\}$ and $\{u,v\}$,
-		which  arrive  during the interval $(t'',t]$.
-		Similarly,
-		request in $R_{t'}$  are to $\{u,w\}$ and some other pair, irrelevant to the analysis.
-		Hence, requests to $\{u,w\}$ are included in at most two such sets,
-		which are $R_{t'}$ and $R_t$.
-		This is because their intervals overlap on $(t'',t']$, shown with red thick line.
-	}
-	\label{fig:timeline}
-\end{figure}
-
-	Next,
-	we bound  costs incurred by \OPT for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
-	If $t$ is the first time that \RM collocates $\{u,v\}$,
-	 then  \OPT pays
-	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
-	or $\alpha$ for collocating the pair and
-	serving (some of) the requests with  cost zero.
-	Therefore in this case,
-	$\OPT(\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
-	%
-	Otherwise,
-	it is not the first collocation and
-	consider times $t'$ and $t''$ as defined previously,
-	 and let 
-	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
-	We define $R_{t'}$ for the collocation at $t'$  analogously
-	(see Figure \ref{fig:timeline}).
-	Then,
-	$\OPT(R_t) =\OPT (\sigma_{\{u,w\}}) 
-	+ \OPT(\sigma_{\{u,v\}}) $.
-	If \OPT has both pairs separated during their respective intervals,
-	then obviously it pays $2\lambda$ during those intervals.
-	Note that \OPT cannot have both pairs collocated at the same time.
-%	However, it can have one of the pairs already collocated prior to its respective interval.
-	Let us assume \OPT has one of the pairs,
-	e.g.~$\{u,v\}$,
-	 collocated already prior its respective interval, $(t',t]$,
-	 and keeps it so during the interval.
-	 Then it pays zero while serving $\sigma_{\{u,v\}}$.
-	Hence,
-	it must pay $\alpha$ for collocating the other pair, in this case $\{u,w\}$,
-	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$. 
-	Therefore in any case,
-	$\OPT(R_t) \geq  \min{\{ \lambda,\alpha \}} = \alpha$.
-
-	It remains to bound the cost  incurred by \OPT due to requests to $\{u,v\}$ that do not lead to its collocation until the end of the sequence at $t=m$.
-	We bound the cost analogously to the case where \RM collocates $\{u,v\}$.
-	If $\{u,v\}$ is not collocated in the initial matching
-	and \RM never collocates it,
-	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[1,m] |$.
-	\OPT pays
-	$\OPT(\sigma_{\{u,v\}}[1,m]) 
-	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
-	for collocating this pair or (and) paying for (resp.~some of) requests in $\sigma_{\{u,v\}}[1,m]$.
-	Else,
-	either $\{u,v\}$ is collocated in the initial matching
-	or \RM collocates it at some point.
-	Then there exists an intersecting pair $\set{u,w}$
-	that is collocated by \RM at $t' < m$,
-	separating $\{u,v\}$.
-	We define times $t'' < t' < m$ analogously to the former case.
-	Let $R^*_{\{u,v\}} := \sigma_{\{u,w\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
-	Then,
-	\OPT must pay for collocating at least one pair or (and) serving requests 
-	to the other pair remotely.
-	Thus,
-	$\OPT (R^*_{\{u,v\}}) 
-	\geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.
-
-	Next, we sum up all costs incurred by \OPT.
-	By definitions of $R_t$ and $R^*_{\{u,v\}}$, we have either
-	$R_{t'} \cap R_t = \sigma_{\{u,w\}}$ or
-	$R_{t'} \cap R^*_{\{u,v\}} = \sigma_{\{u,w\}}$.
-%	(\mahmoud{See figure \ref{fig:doubleCount}}). 
-	This means,
-	$\OPT ( \sigma_{\{u,w\}})$
-	is counted at most twice in each of  the expressions
-	$\OPT (R_{t'}) + \OPT (R_t)$
-	and  
-	$\OPT (R_{t'}) + \OPT(R^*_{\{u,v\}})$.
-	Hence,
-	for all collocations performed by \RM,
-	and for final requests at $t=m$,
-	\OPT pays at~least 
-	$\frac{1}{2}(
-	\sum_{ t \in T } \OPT (R_t) +
-	\sum_{\{u,v\}}\OPT (R^*_{\{u,v\}})
-	) $.
-	Then,
-	the total cost to \OPT is
-	\begin{align*} 	%\label{eq:costOPT}
-		\OPT (\sigma)
-		&=
-		\frac{1}{2}
-		\Big(
-		\sum_{ t \in T} \OPT (R_t) 
-		+ \sum_{\{u,v\}}\OPT (R^*_{\{u,v\}})
-		\Big)	
-		\geq
-%		\Big(
-%		\sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  +
-%		\sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \} } 
-%		\Big)		
-%		=
-		\frac{1}{2}		
-		\Big(
-		\sum_{ t \in T} \alpha  
-		+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}
-		\Big),
-	\end{align*}
-and
-$
-	\RM(\sigma)	/
-	\OPT (\sigma)
-	\leq
-	2\Big(
-	\sum_{ t \in T} 3\alpha +
-	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
-	\Big)	 \big /
-	\Big(
-	\sum_{ t \in T} \alpha  
-	+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}  
-	\Big) 	\leq 6.
-$
-\end{proof}
-
-\noindent
-%\textbf{Approximation for offline rematching.}
-%We briefly sketch an approximation for the offline problem,
-%that makes use of \RM as a~subroutine.
-%In our approximation algorithm,
-%we simulate \RM on the given sequence.
-%Whenever,
-%\RM collocoates a~pair
-
-\section{Discussion and Future Work}
-
-This paper revisited the dynamic graph partitioning problem and presented several tight bounds for the important model where capacities cannot be exceeded, both for a~general partitioning model and for a~special learning model.
-
-While our bounds are tight, there are several interesting avenues for future research.
-In particular, we have so far focused on deterministic algorithms, and it would be interesting to study the power of randomization in this context.
-On the practical side, it would also be interesting to study our algorithms empirically, under realistic workloads.
-
-Our algorithms allow for efficient distributed implementations. 
-The algorithm PPL from Section~\ref{sec:ppl} can be distributed
-similarly to the approach in~\cite{sigmetrics19_partitioning}.
-The algorithm for $k=2$ from Section~\ref{sec:k2} performs only local communication for each request: counters are kept on the clusters and updated locally, and each migration is local within two clusters that reached the collocation threshold $\lambda$.
-Furthermore, we proposed an efficient distributed implementation of the algorithm for $k=3$ in Section~\ref{sec:k3}.
-
 \bibliographystyle{plainurl}
@@
 
-\pagebreak
-
-\section{Pseudocode of Perfect Partition Learner}
-%
-\begin{algorithm}[h]
-	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-	\renewcommand{\algorithmicensure}{\textbf{Output:}}
-	\begin{algorithmic}
-		%        \Require 
-		%        $k, \ell$,
-		%        initial partition $P_I$,
-		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
-		%        \Ensure a~final partition $P_F$ 
-		\STATE {For each node $v$ create a~singleton component $C_v = \{ v \}$ and add it to $\mathcal{C}$.}
-		%\STATE{$P := P_I$}
-		\label{line:initcomponents}
-		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
-		\STATE{Let $C_1 \ni u$ and $C_2 \ni v$ be the  components containing $u$ and $v$, respectively.}
-		\IF{$C_1 \neq C_2$}
-		\STATE {Merge $C_1$ and $C_2$ into one component $C'$ and
-			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$.} \label{line:mergecomponents}
-		\IF{$C_1$ and $C_2$ are not collocated}
-		\STATE {Move to a~partition closest to $P_I$ and respecting all components in $\mathcal{C}$.} 
-		%		\COMMENT{move to a~partition closest to $P_I$}
-		\label{line:rebalance} 
-		\ENDIF
-		\ENDIF
-		\ENDFOR
-	\end{algorithmic}
-	\caption{Perfect Partition Learner (\PPL)}
-	\label{alg:ppl}
-\end{algorithm}
-
 \end{document}

### 2020-05-18 podc_NEW/podc_ba.tex
Cutting a lot

@@
 
-This problem is known as the dynamic balanced graph partitioning problem and was introduced by Avin et al.~\cite{repartition-disc, sidma-arxiv} at DISC 2016.
-a~special variant of the general problem has later been studied by Henzinger et al.~\cite{sigmetrics19_partitioning} at SIGMETRICS 2019.
-We refer to the latter as the learning model.
-
-\subsection{Model}
-
-We study two models in this paper: the \emph{general partitioning} model, and its subproblem, the \emph{learning} model.
-In both models, we assume that communication patterns are not known to our algorithms at the beginning.
-We measure the~quality of~presented algorithmic solutions by competitive analysis~\cite{borodin-book}, which is well-suited for problems that are online by their nature.
-In the competitive analysis, the goal is to~optimize \emph{the competitive ratio} of a~given online algorithm: the ratio of its cost to the cost of~an~optimal offline algorithm that knows the entire input sequence in advance.
-%We emphasize that all our algorithms have a~\emph{strict} competitive ratio (without an additive term).
-
-\noindent
-\textbf{General partitioning model.}
-In the \emph{dynamic balanced graph partitioning} problem, we are given a~set $V$ of $n$ nodes 
+\noindent \textbf{Model.}
+The \emph{dynamic balanced graph partitioning} problem (\OBRP{})
+is a fundamental learning problem
+that finds applications in the context of
+distributed systems optimization~\cite{repartition-disc}.
+We are given a~set $V$ of $n$ nodes 
 (e.g., virtual machines or processes),
@@
 
-\noindent
-\textbf{Learning model.}
-We study a~\emph{learning} variant of dynamic balanced graph partitioning,
-where the communication pattern is \emph{static}:
-whether a~pair of  nodes ever communicate or not, 
-is determined a~priori and is unknown to algorithms,
- and such pairs communicate forever.
-Any algorithm must eventually collocate  pairs of communicating nodes,
-as otherwise it cannot be competitive.
-As in Henzinger et al.~\cite{sigmetrics19_partitioning}, we assume that the communication graph admits a~\emph{perfect partition},
-i.e., a~partition in which no inter-cluster request ever occurs.
-%any algorithm must learn this perfect partition.
-The algorithm's objective is to \emph{learn} the (static) communication graph
- while serving all requests,
-and without executing too many migrations.
-For the learning model, we  assume that the migration cost is $\alpha=1$.
+We consider two flavors of the problem
+in this paper. In the first one, which
+was also studied by Henzinger et al.~\cite{sigmetrics19_partitioning}
+(SIGMETRICS 2019) and which
+we here mainly consider for the sake of a stronger
+lower bound, we assume that $\sigma$
+simply reveals the edges of a static graph
+that can be perfectly partitioned
+(i.e., in principle, no external requests
+are required). 
+In the second one, we consider a general
+model where $\sigma$ can be arbitrary;
+this was the original problem introduced
+by Avin et al.~\cite{repartition-disc} at DISC 2016.
+For simplicity, we will refer to the former
+model as the \emph{learning} model (as
+one has to learn the static communication graph) 
+and to the latter as the \emph{online} model.
 
-\subsection{Related Work}
-
+\noindent \textbf{Related work.}
+%
 The two works closest to ours are by Avin et al.
@@
 (on the learning model)~\cite{sigmetrics19_partitioning}.
-However, the focus of these papers is primarily on models with resource augmentation: the online algorithm can use slightly larger clusters than the offline algorithm.  
-Avin et al.
-actually showed that their lower bound $\Omega(k)$ holds even for a~significant resource augmentation, and they provided an algorithm with the competitive ratio $O(k \log k)$ using the $(2+\epsilon)$-augmented cluster capacity.
-Their ratio is independent of $\ell$, which is impossible without significant resource augmentation.
-
-In contrast, we study the non-augmented setting, where the nodes need to be perfectly balanced  among the clusters.
-This assumption is not only more realistic but also significantly more challenging, as it is related to hard problems such as integer partitioning~\cite{integer-partitions-book}.
-In terms of results without augmentation, so far, it is only known that there exists an~$O(k^2 \cdot \ell^2)$-competitive algorithm~\cite{repartition-disc}; the best known lower bound is significantly lower, namely $\Omega(k)$.
-For $k=2$, Avin et al.~\cite{repartition-disc} presented a~$7$-competitive algorithm with a~substantial ($\Omega(\ell^2)$) additive constant.
-
-The problem has also been studied in a~weaker
-model where the adversary can only sample
-requests from a~fixed distribution~\cite{stochastic-ring}.
-
 The static offline version of~the~partitioning~problem, i.e., a~problem variant where
@@
 the goal is to find an assignment of $n$ nodes to $\ell$~physical machines, each of~capacity $n/\ell$, is known as the
-\emph{$\ell$-balanced graph partitioning problem}. The problem is 
-NP-complete, and cannot even be approximated within any finite factor unless P
-= NP~\cite{AndRae06}.  The static
-variant where $\ell = 2$ corresponds to the minimum bisection problem, which
-is already NP-hard~\cite{GaJoSt76}, and 
-the currently best approximation ratio is $O(\log n)$~\cite{SarVaz95,ArKaKa99,FeKrNi00,FeiKra02,KraFei06,Raec08}.
-
-Our problem is further related to some classic online problems.
-In particular, it is related to online paging~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00}, sometimes also referred to
-as online caching, where requests for data items (nodes) arrive over time and
-need to be served from a~cache of finite capacity, and where the number of
-cache misses must be minimized. Classic problem variants usually boil down to
-finding a~smart eviction strategy, such as Least Recently Used (LRU)~\cite{SleTar85}. In our
-setting, requests can be served remotely (i.e.,~without fetching the
-corresponding nodes to a~single physical machine). In this light, our model is more
-reminiscent of caching models \emph{with
-bypassing}~\cite{EpImLN11,EpImLN15,Irani02}.
-A major difference between  these problems is that in the caching problems, each request involves a~single element of the universe, while in our model \emph{both} endpoints of a~communication request are subject to~optimization.
-In this light, we can see our model as a~"symmetric" version of online paging.
-
+\emph{$\ell$-balanced graph partitioning problem}~\cite{AndRae06}.
 Dynamic graph partitioning problems are generally fundamental in computer science, and arise in many different contexts~\cite{streaming-soda,streaming1}.
 
-%More generally, the model is related to online
-%caching~\cite{SleTar85,FKLMSY91,McGSle91,AcChNo00},
-%see~\cite{repartition-disc} for a~discussion.
-%The static offline version of~the~problem, called the
-%\emph{$\ell$-balanced graph partitioning problem} is 
-%NP-complete, and cannot even be approximated within %any finite factor unless P
-%= NP~\cite{AndRae06}.
-
-\subsection{Our Contributions}
-
+\noindent \textbf{Contributions.}
+%
 This paper presents several new results on the dynamic graph partitioning problem  without augmentation.
@@
 for the learning model.
-
 For the general partitioning model, we design
@@
 We further present a~strictly $6$-competitive algorithm for $k=2$ that improves upon the previous $7$-competitive algorithm with $O(\alpha\ell^2)$ additive constant.
-
-All algorithms in this paper have a strict competitive ratio (i.e., without an additive term).
-
-\medskip
-\noindent
-\textbf{Algorithmic techniques.} The variant for general $k$ is still unresolved, hence it is vital to summarize algorithmic techniques used in this paper.
-A straightforward analysis of the algorithm from Section~\ref{sec:k3} results in the bound $O(\ell^2)$, and the analysis must be improved in two places.
-One must bound the cost of each of $O(\ell)$ reconfigurations per~phase by a~constant, and show that the optimum offline algorithm must pay a~significant cost for inter-cluster requests.
-We bound the cost of the latter by estimating the capabilities of the optimum offline algorithm to \emph{prepare} for an incoming sequence of requests.
-Furthermore, in the analysis of the algorithm from Section~\ref{sec:k2}, we propose a~novel charging scheme for edges that share a~vertex.
-We also emphasize the implications of the tight result for the learning model.
-If one hopes to improve the lower bound for the general model, the ground sets must be dynamically added and \emph{removed}: if ground sets are only added, the instance is solvable by the algorithm from Section~\ref{sec:ppl} within the cost $O(k\cdot \ell)$.
 

### 2020-05-21 podc_NEW/podc_ba.tex
wip

@@
 
-We consider two flavors of the problem
-in this paper. In the first one, which
-was also studied by Henzinger et al.~\cite{sigmetrics19_partitioning}
-(SIGMETRICS 2019) and which
-we here mainly consider for the sake of a stronger
-lower bound, we assume that $\sigma$
-simply reveals the edges of a static graph
-that can be perfectly partitioned
-(i.e., in principle, no external requests
-are required). 
-In the second one, we consider a general
-model where $\sigma$ can be arbitrary;
-this was the original problem introduced
-by Avin et al.~\cite{repartition-disc} at DISC 2016.
-For simplicity, we will refer to the former
-model as the \emph{learning} model (as
-one has to learn the static communication graph) 
-and to the latter as the \emph{online} model.
-
 \noindent \textbf{Related work.}
@@
 
-In this section, we consider the learning variant of dynamic balanced graph partitioning problem.
-For this setting, we show a~surprisingly high lower bound of $\Omega(k \cdot \ell)$ for $k\geq 3$.
-The lower bound holds also in the general partitioning model (studied in Section~\ref{sec:part}).
-At the end of this section, we discuss an asymptotically optimal upper bound for the learning variant.
+In this section, we study a~\emph{learning} variant of dynamic balanced graph partitioning,
+where the communication pattern is \emph{static}:
+whether a~pair of  nodes ever communicate or not, 
+is determined a~priori and is unknown to algorithms,
+ and such pairs communicate forever.
+Any algorithm must eventually collocate  pairs of communicating nodes,
+as otherwise it cannot be competitive.
+As in Henzinger et al.~\cite{sigmetrics19_partitioning}, we assume that the communication graph admits a~\emph{perfect partition},
+i.e., a~partition in which no inter-cluster request ever occurs.
+%any algorithm must learn this perfect partition.
+The algorithm's objective is to \emph{learn} the (static) communication graph
+ while serving all requests,
+and without executing too many migrations.
 
@@
 We provide a~lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm for the learning problem.
-Later, we elaborate on how to efficiently transform it to a~lower bound for the general partitioning problem.
 The lower bound requires $k\geq 3$.
 In contrast, for $k=2$ the learning problem is trivial: immediate collocation of communicating pairs is $1$-competitive.
-In contrast,
-the general partitioning problem for $k=2$ is non-trivial (see Section~\ref{sec:k2}).
 %, a~lower bound of $3$ is known~\cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)
-
-Throughout this paper, we often refer to groups of communicating nodes.
-We use this concept slightly differently in the lower bound than the upper bounds.
-In our algorithms, we group nodes with a~communication history into \emph{components}.
-In this section,
-we group nodes that may ever communicate, into \emph{ground sets}.
 
@@
 
+A~\emph{ground set} is a group of nodes that communicate if they are not collocated.
 The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
-Once we construct a ground set, it lasts until the end of the input sequence.
-We start by constructing a~ground set of size $k-1$ on an arbitrarily chosen cluster.
+We start by constructing a~ground set $B$ of size $k-1$ on an arbitrarily chosen cluster.
 In any partition, there must exist an isolated node collocated with the ground set of size $k-1$.
@@
 We issue requests between this node and some node that was initially collocated with it.
-By repeating such requests, almost every node is once collocated with the first ground set.
-In comparison, we show that there exists an optimal offline algorithm \OPT that performs only two node exchanges ("swaps").
-%Figure~\ref{fig:lb} illustrates the constructs used in the lower bound.
+By repeating such requests, almost every node is once collocated with the ground set $B$.
+The cost of the algorithm is linear in terms of the number of nodes, and there exists an optimal offline algorithm \OPT that performs only two node exchanges.
 
@@
 
-We present an asymptotically optimal algorithm for the learning problem.
+There exists an asymptotically optimal algorithm for the learning problem.
 The algorithm  collocates  a~pair as soon as they communicate and it never separates them.
-In order to preserve collocated pairs,
+To preserve collocated pairs,
 we employ the concept of components,
@@
 
-We maintain subsets of frequently communicating nodes as \emph{components}.
-Initially,
-each node constitutes a single-node component which we refer to as a \emph{singleton} component, 
-and the node in such component is an \emph{isolated} node.
-We define larger components in terms of smaller components.
-Concretely,
-given a (sub)sequence of requests,
-two components $C_1$ and $C_2$
- \emph{merge} into one component as soon as
- for some pair of nodes $v_1 \in C_1$ and $v_2 \in C_2$,
-the frequency of requests $\{v_1,v_2\}$ reaches a certain threshold through the sequence.
-We keep all nodes of a~component always collocated in the same cluster,
-i.e., when we  move a~node,
-we move the whole component that contains it.
-A~partition that has every component collocated 
-is a~\emph{component respecting} partition.
-%Intuitively, a component preserves the history of requests between its nodes.
-
-In addition,
-we maintain a~balanced partition of our components as long as such partition exists,
-a reminiscent of partitioning given integers into sets of equal sum~\cite{integer-partitions-book}.
-In contrast, our partition is time-varying:  two components are merged into one component once they communicate, and we adjust the partition accordingly.
-
-The algorithm in this section and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from~\cite{repartition-disc}.
-The difference is in the choice of partition after a~component merge.
-In DET, the partition was arbitrary.
-In the algorithm for the learning model,
-we choose a~component respecting partition closest to the initial partition.
-In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a~repartition of minimum cost).
-%A generic sketch of these two algorithms can be found in the appendix (Algorithm \ref{alg:ppl}).
-
 \noindent
 \textbf{Perfect Partition Learner algorithm.}
-Now we describe the algorithm \PPL.
 On each inter-cluster request $\{u,v\}$, 
-\PPL creates new components by merging
+\PPL  merges
  the two components that contain nodes $u$ and $v$.
-In order to collocate nodes of the new component,
-\PPL moves to a~component respecting partition that minimizes the distance to the initial partition $P_I$.
-The scheme of the algorithm can be found in the  appendix (Algorithm \ref{alg:ppl}).
-
-Fix the initial partition
-$P_I :=\{ I_1, \dots, I_{\ell}\}$ and \OPT's final partition
-$P_F := \{F_1, \dots, F_{\ell}\}$.
-The \emph{distance} of a~partition
- $P = \{C_1, \dots, C_{\ell}\}$ from the initial partition,
-defined as 
-$\Delta(P) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$,
- is the number of nodes in $P$ that do not reside in their initial cluster.
-In other words,
-at least $\Delta(P)$ node migrations are required in order to reach the partition $P$ from $P_I$, and thus
-$\OPT \geq \Delta(P_F) $.
-
-\PPL never moves to a~partition that is more than $ \Delta^*:= \Delta(P_F) $   migrations away from~$P_I$.
-This invariant latter ensures us that \PPL does not pay too much while recovering $P_F$.
-We emphasize that a~\emph{repartitioning} by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
-This way \PPL never moves to a~partition beyond the distance $\Delta^*$.
-
-\begin{property} \label{prop:dist<OPT}
-	Let $P$ be any partition chosen by \PPL at any time.
-	Then, $\Delta(P) \leq \Delta^*$.	
-\end{property}
+To collocate nodes of the new component,
+\PPL moves to partition where no component is split, that minimizes the distance to the initial partition $P_I$.
 
 \begin{lemma}	\label{lemma:rebalancecost}
-	The cost of each repartitioning by \PPL is $2\cdot\OPT$.
+	The cost of serving each request is at most $2\cdot\OPT$.
 \end{lemma}
 
+Let $\Delta$ be the cost of \OPT.
+\PPL never moves to a~partition that is more than $ \Delta$   migrations away from~the initial configuration.
+
 \begin{theorem}	\label{thm:upperbound}
-	\PPL reaches the final partition $P_F$
-	 and it is $(2\cdot (k-1)\cdot\ell)$-competitive.
+	\PPL is $(2\cdot (k-1)\cdot\ell)$-competitive.
 \end{theorem}

### 2020-05-21 podc_NEW/podc_ba.tex
wip

@@
 	communication pattern can change arbitrarily over time.
-	We improve the result for $k=2$ by providing a~strictly $6$-competitive upper bound for the general model.
 	In contrast to most prior work, our algorithms respect all capacity constraints and do not require resource augmentation.
@@
 so far $O(\ell^2)$~\cite{repartition-disc}.
-We further present a~strictly $6$-competitive algorithm for $k=2$ that improves upon the previous $7$-competitive algorithm with $O(\alpha\ell^2)$ additive constant.
 
@@
 
-We provide a~lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm for the learning problem.
-The lower bound requires $k\geq 3$.
-In contrast, for $k=2$ the learning problem is trivial: immediate collocation of communicating pairs is $1$-competitive.
-%, a~lower bound of $3$ is known~\cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)
-
 \begin{theorem}
@@
 	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least
-	 $(k-2)(\ell-1)/2 - 2$ for any $k\geq 3$ and $\ell \geq 2$.
+	 $\Omega(k\cdot \ell)$ for any $k\geq 3$ and $\ell \geq 2$.
 \end{theorem}
 
-A~\emph{ground set} is a group of nodes that communicate if they are not collocated.
-The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
-We start by constructing a~ground set $B$ of size $k-1$ on an arbitrarily chosen cluster.
-In any partition, there must exist an isolated node collocated with the ground set of size $k-1$.
+We choose a set $B$ of $k-1$ collocated nodes and issue requests between them so they must stay collocated.
+In any partition, there must exist an isolated node collocated with $B$.
 %, and we refer to it as the \emph{pivot} node.
 We issue requests between this node and some node that was initially collocated with it.
-By repeating such requests, almost every node is once collocated with the ground set $B$.
+We repeat such requests to the current node collocated with $B$, and this node get replaced by another node that was never requested.
+We issue $(k-2)(\ell - 1)-4$ requests this way.
 The cost of the algorithm is linear in terms of the number of nodes, and there exists an optimal offline algorithm \OPT that performs only two node exchanges.
@@
 
-There exists an asymptotically optimal algorithm for the learning problem.
-The algorithm  collocates  a~pair as soon as they communicate and it never separates them.
-To preserve collocated pairs,
-we employ the concept of components,
+We present the algorithm \PPL that is  asymptotically optimal for the learning problem.
+It utilizes the concept of components,
 introduced by Avin~et~al.~\cite{repartition-disc}.
@@
 
-\begin{lemma}	\label{lemma:rebalancecost}
-	The cost of serving each request is at most $2\cdot\OPT$.
-\end{lemma}
+\medskip
 
@@
 
-\begin{theorem}	\label{thm:upperbound}
-	\PPL is $(2\cdot (k-1)\cdot\ell)$-competitive.
-\end{theorem}
+\begin{lemma}	\label{lemma:rebalancecost}
+	The cost of a single repartition is at most $2\cdot\OPT$.
+	The corollary is that \PPL is $(2\cdot (k-1)\cdot\ell)$-competitive.
+\end{lemma}
 
-\section{General Partitioning Model}
+\section{General Partitioning Model: Optimal Algorithm for Clusters of Size 3}
+
 \label{sec:part}
+\label{sec:k3}
 
@@
 can be arbitrary.
-In Section~\ref{sec:k3}, we show an $O(k \cdot \ell)$-competitive algorithm for $k=3$ using the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental}.
-
-\subsection{Optimal Algorithm for Clusters of Size 3}
-\label{sec:k3}
-
 The algorithm analyzed in this section is a modified version of the algorithm DET proposed by Avin et al.~\cite{repartition-disc}, which for $k=3$ is $O(\ell^2)$-competitive.
 In our algorithm, we choose the closest partition after a component merge instead of an arbitrary one.
-This allows to bound the cost of repartition by a constant (Lemma~\ref{lem:1req}).
+This allows to bound the cost of repartition by a constant:
+\begin{lemma}
+	\label{lem:1req}
+	In a~single repartition of nodes, the algorithm exchanges at most two pairs of nodes.
+\end{lemma}
 
@@
 
-\medskip
-
-\noindent
-\textbf{Component-based algorithm.}
-The algorithm \TAlg partitions nodes into components, and
-initially, each node is isolated (belongs to its own component).
-For each pair of nodes $\set{x,y}$, \TAlg maintains a~counter $C_{\set{x,y}}$ and increments it on every external request between $x$ and $y$.
-Once $C_{\set{x,y}} = \alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
-If no such partitioning exists, \TAlg resets all components to singleton components, resets all counters to $0$, and ends the phase.
-
-\begin{lemma}
-	\label{lem:1req}
-	In a~single repartition of nodes (after a~merge of components), \TAlg exchanges at most two pairs of nodes.
-\end{lemma}
-
 \begin{theorem}
-	\TAlg is $60\ell$-competitive for $k=3$.
+	The proposed algorithm for $k=3$ is $O(\ell)$-competitive.
 	\label{thm:k=3}

### 2020-05-21 podc_NEW/podc_ba.tex
BA 3 pages

@@
 introduced by Avin~et~al.~\cite{repartition-disc}.
-
-\noindent
-\textbf{Perfect Partition Learner algorithm.}
 On each inter-cluster request $\{u,v\}$, 
@@
 To collocate nodes of the new component,
-\PPL moves to partition where no component is split, that minimizes the distance to the initial partition $P_I$.
+\PPL moves to partition where no component is split, that minimizes the distance to the initial partition.
 
@@
 
+\noindent
 Let $\Delta$ be the cost of \OPT.
@@
 \begin{lemma}	\label{lemma:rebalancecost}
-	The cost of a single repartition is at most $2\cdot\OPT$.
-	The corollary is that \PPL is $(2\cdot (k-1)\cdot\ell)$-competitive.
+	The cost of a single repartition is at most $2\cdot\Delta$.
+	The corollary is that \PPL is $O(k\cdot\ell)$-competitive.
 \end{lemma}
@@
 
-This modification alone is insufficient to obtain $O(\ell)$-competitive algorithm, and the analysis must be further improved.
-In particular, pairs of nodes that did not reach the collocation threshold $\alpha$ (called external requests) incur the cost $O(\ell^2)$ for the algorithm in each phase.
-The novel part of the analysis lower-bounds the cost of \OPT on external requests while considering its savings from migrations and possibly different configuration at the beginning of the phase.
-This way, we show that \OPT paid a significant portion of the algorithm's cost on external requests.
+This modification alone is insufficient to obtain $O(\ell)$-competitive algorithm: pairs of nodes that did not reach the collocation threshold $\alpha$ incur the cost $O(\ell^2)$.
+We carefully analyse this cost and relate it to the cost of \OPT.
 
 \begin{theorem}
-	The proposed algorithm for $k=3$ is $O(\ell)$-competitive.
+	The  algorithm for $k=3$ is $O(\ell)$-competitive.
 	\label{thm:k=3}
